### 2020-04-01 motivation-talk/talk.tex
Added slides on motivation for this, wrt. past work

% !TeX program = lualatex

\documentclass[aspectratio=169,xcolor={dvipsnames}
%,notes=only
%,notes
%,show notes on second screen=right
,handout
]{beamer}
\usetheme[background=light, numbering=fraction]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{pgfpages}

%\usepackage[T1]{fontenc}

\usepackage{bm}

\usepackage[labelfont=bf,textfont={it}]{caption}
\usepackage{subcaption}
\captionsetup[figure]{justification=centering}
\captionsetup[subfigure]{justification=centering}

\usepackage{tikz}
\usetikzlibrary{arrows.meta, calc, fit, positioning}

%\usepackage{fontspec}
%\setsansfont{Fira Sans Mono}

\usepackage{etoolbox}
\usepackage[binary-units]{siunitx}
\robustify\bfseries
\sisetup{detect-all, range-phrase=--, range-units=single}

\usepackage[UKenglish]{babel}
\usepackage{csquotes}

\usepackage{amssymb}

\usepackage{lipsum}
\usepackage[basic]{complexity}
\usepackage[super,negative]{nth}

\usepackage{booktabs}

%bib
\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,sortcites
,backend=bibtex
,style=authortitle
]{biblatex}
%\addbibresource{papers.bib}
%\addbibresource{confs-journs.bib}

\newcommand{\acval}[3]{\ensuremath{\operatorname{\hat{q}}(#1, #2, #3)}}
\newcommand{\wvec}[1]{\ensuremath{\bm{w}_{#1}}}

\makeatletter
\DeclareRobustCommand{\rvdots}{%
	\vbox{
		\baselineskip4\p@\lineskiplimit\z@
		\kern-\p@
		\hbox{.}\hbox{.}\hbox{.}
}}
\makeatother

% official colours
\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}

\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.317647, 0.2}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}

\definecolor{inferno0}{rgb}{0.001462 0.000466 0.013866}
\definecolor{inferno64}{rgb}{0.341500 0.062325 0.429425}
\definecolor{inferno128}{rgb}{0.735683 0.215906 0.330245}
\definecolor{inferno192}{rgb}{0.978422 0.557937 0.034931}
\definecolor{inferno255}{rgb}{0.988362 0.998364 0.644924}

%picky abt et al.
\usepackage{xpatch}

\makeatletter\let\expandableinput\@@input\makeatother

\xpatchbibmacro{name:andothers}{%
	\bibstring{andothers}%
}{%
	\bibstring[\emph]{andothers}%
}{}{}

%opening!

\usepackage{cleveref}
\newcommand{\crefrangeconjunction}{--}

\usepackage{fontawesome}

\addtobeamertemplate{footnote}{\vspace{-6pt}\advance\hsize-0.5cm}{\vspace{6pt}}
\makeatletter
% Alternative A: footnote rule
\renewcommand*{\footnoterule}{\kern -3pt \hrule \@width 2in \kern 8.6pt}
% Alternative B: no footnote rule
% \renewcommand*{\footnoterule}{\kern 6pt}
\makeatother

\usepackage[export]{adjustbox}
\usetikzlibrary{arrows.meta, calc, fit, positioning, shapes.misc}

%-------------------------------------%
%-------------------------------------%

\title{Reinforcement Learning in Future Networks}
\author{\vspace{-1em}\textbf{Kyle A. Simpson}\\
	\faEnvelopeO{} \href{mailto:k.simpson.1@research.gla.ac.uk}{\nolinkurl{k.simpson.1@research.gla.ac.uk}}\\
	\vspace{1em}\small{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}}
\institute{University of Glasgow}
\date{\nth{27} March, 2020}

\begin{document}
\begin{frame}
\maketitle
\end{frame}

\begin{frame}{Data-driven networking: RL in networks}
	\alert{Data-driven networking}: Automate control, optimisation, configuration of the network.
	\begin{itemize}
		\item Flow performance optimisation.
		\item Resource allocation.
		\item Adaptive response to load, intrusions, etc.
		\item Feedback loop-like.
	\end{itemize}
\end{frame}

\begin{frame}{Data-driven networking: DDoS Prevention}
	\begin{columns}
		\begin{column}{0.6\linewidth}
			\begin{figure}
				\centering
				\resizebox{\linewidth}{!}{
					\begin{tikzpicture}[
					texts/.style = {text=black},
					labeltexts/.style = {text=gray},
					treeline/.style = {draw=uofgburgundy},
					treenode/.style = {texts, circle, centered, fill=white, treeline},
					load/.style = {fill=uofgcobalt},
					loadline/.style = {draw=uofgcobalt, line width=0.75mm},
					loadhide/.style = {fill=uofgcobalt!40!white},
					external/.style = {fill=uofgrust},
					externalhide/.style = {fill=uofgrust!40!white},
					hideline/.style = {draw=uofgsandstone!40!white},
					hidenode/.style = {treenode, hideline},
					grow'=right
					]
					\node [treenode, loadhide, label={[texts]above:Agent 1}] (mainagent) {};
					\node [treenode, loadhide, right = of mainagent] (inner0) {};
					\node [treenode, loadhide, above right = 0.2cm and 1cm of inner0] (inner1) {};
					\node [hidenode, below right = 0.2cm and 1cm of inner0] (inner2) {};
					\node [treenode, loadhide, below right = 0.2cm and 1cm of inner1] (inner3) {};
					\node [treenode, loadhide, right = of inner3] (inner4) {};
					\node [treenode, loadhide, above right = 0.2cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
					\node [hidenode, below right = 0.2cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};

					\node [hidenode, above left = 0.2cm and 0.5cm of inner1, label={[labeltexts]above:Agent 2}] (agent2) {};
					\node [hidenode, below left = 0.2cm and 0.5cm of inner2, label={[labeltexts]below:Agent 3}] (agent3) {};

					\draw[hideline, -] ($ (mainagent) + (-1,0) $) -- (mainagent);
					\draw[hideline, -] ($ (agent2) + (-1,0) $) -- (agent2);
					\draw[hideline, -] ($ (agent3) + (-1,0) $) -- (agent3);
					\draw[hideline, -] (inner1) -- (agent2);
					\draw[hideline, -] (inner2) -- (agent3);

					\draw[loadline, -] (mainagent) -- (inner0);
					\draw[loadline, -] (inner0) -- (inner1);
					\draw[hideline, -] (inner0) -- (inner2);
					\draw[-] (inner1) -- (inner3);
					\draw[hideline, -] (inner2) -- (inner3);
					\draw[loadline, -] (inner3) -- (inner4);
					\draw[loadline, -] (inner4) -- node [texts, above] {$U_{s_0}$} (s0) ;
					\draw[hideline, -] (inner4) -- node [labeltexts, below] {$U_{s_1}$} (s1);

					\end{tikzpicture}
				}
			\caption{`Global' state for any flow (\textbf{bold} means used by policy).}
			\end{figure}
		\end{column}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item Monitor flow statistics at ingress/egress points.
				\item Use load measurements from flow paths as global state.
				\item OUTPUT: throttle individual flows.
			\end{itemize}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Data-driven networking: DDoS Prevention}
	\begin{columns}
		\begin{column}{0.3\linewidth}
			How does this perform compared to state of the art, SPIFFY...?
			\begin{itemize}
				\item Worse for congestion-aware, but strong.
				\item \alert{Improvement on VoIP/UDP traffic!}
				\item Learning time on order of minutes.
				\begin{itemize}
					\item Classical RL.
				\end{itemize}
			\end{itemize}
		\end{column}
		\begin{column}{0.7\linewidth}
			\begin{figure}
				\centering
				\includegraphics[width=0.75\linewidth]{tnsm-tcp-16-single}
				\caption{
					Online performance of RL agents for TCP traffic, with and without policy sharing.
					\label{fig:tcp-tree-16}
				}
			\end{figure}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Programmable data-planes: Motivation}
\begin{columns}
	\begin{column}{0.6\linewidth}
		\begin{figure}
			\resizebox{\linewidth}{!}{
			\begin{tikzpicture}
			\node (iLabel) {Theory};

			\node (iDiag) at ($(0, -1.3) + (iLabel)$) {
			\resizebox{2cm}{!}{\begin{tikzpicture}
				\node[circle, draw] (state) {$S$};
				\node[circle, draw] (state') at ($(state) + (0, -2)$) {$S'$};

				\node (agent) at ($(2, 0) + (state)$) {Agent};
				\node[below of=agent] (action) {Action $A$};

				\node[right of=state'] {$+ R$};

				\draw[->] (state) -- (state') node[midway, right] {$A$};

				\draw[dotted, ->, bend left = 30] (state) -- (agent);
				\draw[->] (agent) -- (action);
				\draw[dotted, ->] (action) -- (state);
			\end{tikzpicture}}};

			\node (rLabel) at ($(3, 0) + (iLabel)$) {Reality};

			\node (rDiag) at ($(0, -1.5) + (rLabel)$) {
				\resizebox{2cm}{!}{\begin{tikzpicture}
					\node[circle, draw] (state) {$S$};
					\node[circle, draw] (state') at ($(state) + (0, -1.5)$) {$S'$};
					\node[circle, draw] (state'') at ($(state') + (0, -1.5)$) {$S''$};

					\node (agent) at ($(2, 0) + (state)$) {Agent};
					\node[below of=agent] (action) {Action $A$};

					\node[right of=state''] {$+ R$};

					\draw[->] (state) -- (state') node[midway, right] {$\varnothing$};
					\draw[->] (state') -- (state'') node[midway, right] {$A$};

					\draw[dotted, ->, bend left = 30] (state) -- (agent) node[midway, above] {$t_1$};
					\draw[->] (agent) -- (action) node[midway, right] {$t_2$};
					\draw[dotted, ->] (action) -- (state') node[midway, below] {$t_3$};
					\end{tikzpicture}}};
			\end{tikzpicture}
			}
			\caption{Asynchronous RL delays and state slippage (policy updates omitted).}
		\end{figure}
	\end{column}
	\begin{column}{0.4\linewidth}
		\begin{itemize}
			\item Programmable network hardware: arbitrary in-network compute.
			\item In data-driven, want to \alert{minimise time to act}.
			\item RL assumes that action \& policy update are zero-cost.
			\begin{itemize}
				\item Not so in real deployments!
				\item State drift, etc.
			\end{itemize}
		\end{itemize}
	\end{column}
\end{columns}
\end{frame}

\begin{frame}{DDoS Prevention: VNF + SDN  architecture}
	\begin{columns}
%		\begin{column}{0.3\linewidth}
%			\begin{itemize}
%				\item Monitor flow statistics at ingress/egress points.
%				\item Use load measurements from flow paths as global state.
%				\item OUTPUT: throttle individual flows.
%			\end{itemize}
%		\end{column}
		\begin{column}{1.0\linewidth}
			\begin{figure}
				\centering
				\resizebox{0.7\linewidth}{!}{\begin{tikzpicture}[scale=1, every node/.style={scale=0.7}]
					\node(remote){Remote};

					%%%

					\node[below=0.05 of remote](swpos){};
					\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
					\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
					\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
					\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
					\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Load\\Collector\end{varwidth}};

					%

					\node[right=2 of swpos](epos){};
					\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
					\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
					\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
					\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.65,0)$) {\begin{varwidth}{1.5 cm}\scriptsize \centering Load\\Collector\end{varwidth}};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\scriptsize \centering Estimator\\$g(\cdot)$\end{varwidth}};

					%

					\node[right=3.5 of epos](apos){};
					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
					\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent vNF};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};

					%

					\node[below=2.3 of remote.west](linestart){};
					\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
					\draw [dashed] (linestart) -- (lineend);

					%%%

					\node[below=2.4 of remote](local){Local};

					%%%

					\node[below=0.05 of local](aswpos){};
					\node[fill=white!80!uofgthistle, draw=black, minimum height=3.5cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
					\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering OpenFlow\\Tables\end{varwidth}};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\scriptsize \centering Stats\\Collector\end{varwidth}};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\scriptsize \centering Load\\Collector\end{varwidth}};

					%

					\node[right=3 of aswpos](avfpos){};
					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
					\node[below right, inner sep=2pt] at (avf.north west) {\small Agent vNF};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\scriptsize \centering Flowstate\\Database\end{varwidth}};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\scriptsize \centering TRS\\Scheduler\end{varwidth}};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};

					%%%

					\tikzset{>=stealth}

					\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
					\draw[thick, <->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};

					\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};

					\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
					\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};

					\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};

					\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
					\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};

					\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
					\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
					\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
					\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
					\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};

					\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
					\node[right=0.3] at (avfsa.east) {\tiny Experience};
					\end{tikzpicture}}
				\caption{
					System architecture for RL-driven DDoS defence.
					\label{fig:sys-arch}
				}
				\vspace{-1em}
			\end{figure}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Programmable data-planes: An Architecture for RL}
\begin{columns}
	\begin{column}{0.6\linewidth}
		\begin{figure}
			\resizebox{1.2\linewidth}{!}{\begin{tikzpicture}
				\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2.9cm, below right= 0.1 of sw2.north west](p4-master){};
				\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Master Island};
				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (p4-rules) at (p4-master.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};

				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2cm, below = 0.1 of p4-master](sw1){};
				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
				\node[below right, inner sep=2pt] at (switch.north west) {\small P4 Island};
				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};

				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2.9cm, below= 0.1 of switch](others){};
				\node[below right, inner sep=2pt] at (others.north west) {\small Other Islands};
				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-etc) at (others.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering \textellipsis \end{varwidth}};
				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-int) at (others-etc.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering In-band Network Telemetry\end{varwidth}};

				\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (3, -2)$){};
				\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};

				\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
				\node[below right, inner sep=2pt] at (wq-out.north west) {\small CLS};
				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Work Queue\end{varwidth}};

				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (loop) at ($(wq-out.east) + (1.5,-0.5)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Control ME\end{varwidth}};

				\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-cls-out) at ($(loop.south) + (0,-1.5)$){};
				\node[below right, inner sep=2pt] at (policy-cls-out.north west) {\small CLS};
				\node[above right, inner sep=2pt] at (policy-cls-out.north west) {\small Policy};
				\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-ctm-out) at ($(policy-cls-out.south) + (0,-0.25)$){};
				\node[below right, inner sep=2pt] at (policy-ctm-out.north west) {\small CTM};
				\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-emem-out) at ($(policy-ctm-out.south) + (0,-0.25)$){};
				\node[below right, inner sep=2pt] at (policy-emem-out.north west) {\small EMEM};

				\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t1) at (policy-cls-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T1\end{varwidth}};
				\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t2) at (policy-ctm-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T2\end{varwidth}};
				\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t3) at (policy-emem-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T3\end{varwidth}};

				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-1) at ($(wq-out.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-2) at ($(po-me-1.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-3) at ($(po-me-2.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

				\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2]  node[pos=0.5, rotate=90, anchor=center, above] {Rules}(sw1);
				\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (sw2);
				\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (switch);

				\draw[->] (oswlc.east)  to node[midway, above, sloped] {\tiny Control, Requests} (wq-in.west);
				\draw[->] (others-int.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
				\draw[<->, dotted] (wq-in.east)  to (loop.west);
				\draw[->, dotted] (loop.south) to (policy-cls-out);

				\draw[<->, dotted] (po-me-1.east) to (policy-cls-out.west);
				\draw[<->, dotted] (po-me-2.east) to (policy-ctm-out.west);
				\draw[<->, dotted] (po-me-3.east) to (policy-emem-out.west);

				\draw[->] (loop) to node[midway, above, sloped] {\tiny NN} (po-me-1);
				\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NN} (po-me-2);
				\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NN} (po-me-3);
				\draw[->] (po-me-3) to [out=-45,in=0,looseness=2] node[midway, above, sloped] {\tiny NN} (loop);
			\end{tikzpicture}}
			\caption{Architecture for generalised RL agent on Netronome hardware.}
		\end{figure}
	\end{column}
	\begin{column}{0.4\linewidth}
		\begin{itemize}
			\item Make decisions \alert{where they must be applied}.
			\item ``Bump-in-the-wire'' use of Netronome SmartNIC.
			\item P4 parsing/forwarding/filtering, offload to microcode for full programmability.
			\begin{itemize}
				\item P4 with offload common.
			\end{itemize}
				\item Solution: \alert{make use of on-device parallelism, externs to implement fast, adaptive data-driven applications}.
		\end{itemize}
	\end{column}
\end{columns}
\end{frame}

\end{document}

### 2020-04-24 paper/paper.tex
Starting a sense of a writeup...

% !TeX document-id = {5530719d-34df-4dd8-b4b5-e6ed092c2b36}
% !TeX program = pdflatex
% !BIB program = biber
\documentclass[sigconf,natbib=false]{acmart}

% Imports Ahoy

%\usepackage[dvipsnames]{xcolor}
%\usepackage{graphicx}
\usepackage{tikz}
\usepackage{varwidth}
\usetikzlibrary{arrows.meta, calc, fit, positioning}

\usepackage{etoolbox}
\usepackage[binary-units, per-mode=symbol]{siunitx}
\robustify\bfseries
\robustify\emph
%\robustify\uline
\sisetup{detect-all, range-phrase=--, range-units=single, detect-weight=true,detect-inline-weight=math, table-format=1.3}

\makeatletter
\let\MYcaption\@makecaption
\makeatother

\usepackage[font=footnotesize]{subcaption}

\makeatletter
\let\@makecaption\MYcaption
\makeatother

%\usepackage[basic]{complexity}
\usepackage[super,negative]{nth}

\usepackage[british]{babel}
\usepackage{csquotes}

\usepackage{booktabs}
%\usepackage[
%activate={true,nocompatibility},
%final,
%tracking=true,
%kerning=true,spacing=true
%]{microtype}
%\microtypecontext{spacing=nonfrench}

\usepackage[maxnames=2,maxbibnames=99,mincrossrefs=99,sortcites,style=numeric-comp
%,backend=biber
]{biblatex}
\addbibresource{papers-off.bib}
\addbibresource{confs-off.bib}
\addbibresource{books-off.bib}
\addbibresource{rfc.bib}
\addbibresource{misc.bib}

%\DeclareFieldFormat[inproceedings]{doi}{}
%\DeclareFieldFormat[article]{doi}{}
%\DeclareFieldFormat*{url}{}
%\DeclareFieldFormat[online]{url}{\mkbibacro{URL}\addcolon\space\url{#1}}
%\DeclareFieldFormat[report]{url}{\mkbibacro{URL}\addcolon\space\url{#1}}

%picky abt et al.
\usepackage{xpatch}

\xpatchbibmacro{name:andothers}{%
	\bibstring{andothers}%
}{%
	\bibstring[\emph]{andothers}%
}{}{}

\usepackage{url}
\usepackage{hyperref}
\usepackage[nameinlink]{cleveref}
\newcommand{\crefrangeconjunction}{--}
\crefname{table}{table}{tables}

% Official colours!

\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}

\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.317647, 0.2}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}

% End Imports

%\usepackage[english]{babel}
\usepackage{blindtext}

% Copyright
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

\settopmatter{printacmref=false, printccs=false, printfolios=true}

% DOI
\acmDOI{}

% ISBN
\acmISBN{}

%Conference
%\acmConference[Submitted for review to SIGCOMM]{}
%\acmYear{2018}
%\copyrightyear{}

%% {} with no args suppresses printing of the price
\acmPrice{}

\begin{document}
	\title{Revisiting the Classics: Online RL in the Programmable Data-Plane}

	%\titlenote{Produces the permission block, and copyright information}
	%\subtitle{Extended Abstract}

%	\author{Paper \# XXX, XXX pages}
	 \author{Kyle A. Simpson}
	 \orcid{0000-0001-8068-9909}
	 \affiliation{%
	   \institution{University of Glasgow}
	   \city{Glasgow} 
	   \state{Scotland}
	 }
	 \email{k.simpson.1@research.gla.ac.uk}
	 \author{Dimitrios P. Pezaros}
\orcid{0000-0003-0939-378X}
	 \affiliation{%
	\institution{University of Glasgow}
	\city{Glasgow} 
	\state{Scotland}
}
\email{Dimitrios.Pezaros@gla.ac.uk}

	% The default list of authors is too long for headers}
	\renewcommand{\shortauthors}{Simpson \emph{et al}.}

\begin{abstract}
Note: below text is just the UK systems workshop abstract I submitted.	
\end{abstract}

\maketitle

\section{Introduction}

Automatic optimisation, control, and defence of networks are at last becoming commonplace. Data-driven networking has led the charge in traffic optimisation, congestion control and packet classification via adaptive techniques such as Reinforcement Learning (RL), where every change and its measured effects further improve future decisions. Considering that the network evolves in its use and deployed protocols, this flexibility is essential. Yet data-driven methods suffer from a key weakness: they are dependent on both the recency and accuracy of input state. An out-of-date view of the world will lead to suboptimal choices, as will long processing times. These can result in worse performance and slower adaptation to the evolving network.

Programmable data-planes and in-switch compute, then, hold promise for integrating these new techniques in a feasible and efficient manner (beyond dedicated servers or virtualised network functions). For RL, key tasks include policy evaluation, online training, state collection, and action execution -- each of these introduces some degree of sensitivity to state accuracy. Ideally then, all logic would run on these programmable devices. Yet, there is often a finite budget in microcode/FPGA space, per-packet processing times, and available cores for execution. Moreover, necessary hardware, such as floating-point units, is unavailable in almost all cases.

The precise costs and trade-offs which operators and designers must make have yet to be identified. This follows from a design space explosion induced by many necessary workarounds. For instance, quantisation or fixed-point arithmetic will allow training and control on all devices, but introduces further questions: what degree of quantisation is most appropriate? What effect would this have on training accuracy, communications cost, or storage requirements? More concerns arise when we consider core allocation, local vs. distributed training, and reliable lightweight communication in multi-agent scenarios. In all cases, network operators will not consider tools which affect underlying traffic.

I aim to examine the effects of these choices on an existing RL-based DDoS attack mitigation system. To protect legitimate traffic, this controls packet drop and filtering for each flow using individual metrics observed by RL agents at ingress routers, and load measurements from several points along the path taken by the flow in question. The examined metrics include the throughput and latency of its individual components alongside system metrics: flow arrival-to-judgement time, and knowledge propagation time. Beyond this, it's crucial to identify what we can implement on pure P4-capable hardware. While extensions to P4 are fairly common in commodity hardware, the need for full design access (as in NetFPGA) or Micro-C (as in Netronome SmartNICs) represent a break from the clean, loop-free semantics of P4. As the market matures, these may represent different feature classes and price points.

I will discuss early development efforts (including challenges and results) on the Netronome Agilio SmartNIC, which supports the P4 language as well as some degree of arbitrary microcode. I intend to present how reinforcement learning execution and network telemetry will differ in the above metrics when compared to a vNF-based deployment (i.e., software on external commodity servers).

\section{Motivation}

\begin{figure}
	\begin{subfigure}{0.45\linewidth}
		\resizebox{\linewidth}{!}{
			\begin{tikzpicture}
					\node[circle, draw] (state) {$S$};
					\node[circle, draw] (state') at ($(state) + (0, -2)$) {$S'$};

					\node (agent) at ($(2, 0) + (state)$) {Agent};
					\node[below of=agent] (action) {Action $A$};

					\node[right of=state'] {$+ R$};

					\draw[->] (state) -- (state') node[midway, right] {$A$};

					\draw[dotted, ->, bend left = 30] (state) -- (agent);
					\draw[->] (agent) -- (action);
					\draw[dotted, ->] (action) -- (state);
		\end{tikzpicture}}
		\caption{Theory: state measurement, action computation, and learning are zero-cost.}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\resizebox{\linewidth}{!}{
			\begin{tikzpicture}
					\node[circle, draw] (state) {$S$};
					\node[circle, draw] (state') at ($(state) + (0, -1.5)$) {$S'$};
					\node[circle, draw] (state'') at ($(state') + (0, -1.5)$) {$S''$};

					\node (agent) at ($(2, 0) + (state)$) {Agent};
					\node[below of=agent] (action) {Action $A$};

					\node[right of=state''] {$+ R$};

					\draw[->] (state) -- (state') node[midway, right] {$\varnothing$};
					\draw[->] (state') -- (state'') node[midway, right] {$A$};

					\draw[dotted, ->, bend left = 30] (state) -- (agent) node[midway, above] {$t_1$};
					\draw[->] (agent) -- (action) node[midway, right] {$t_2$};
					\draw[dotted, ->] (action) -- (state') node[midway, below] {$t_3$};
		\end{tikzpicture}}
		\caption{Reality: costs of measurement and action lead to \emph{state drift}---over a time delay $t_1+t_2+t_3$, inaction transforms state $S$ into $S'$.}
	\end{subfigure}
\caption{Asynchronous RL delays and state slippage (policy updates omitted).\label{fig:state-slip}}
\end{figure}

What is our goal?

\begin{itemize}
	\item Research into asynchronous RL shows that clever reordering and mitigation can reduce the effects of state slippage on learning time and on accuracy of actions. However, it cannot completely eliminate the effect of state evolving due to the costs of action computation, execution, and policy updates~\cite{DBLP:journals/firai/TravnikMSP18}.

	\item What if we do stop the world (i.e., block operation) to ensure this doesn't happen? In case of congestion control, significant throughput drops (as DNN inference takes \SI{30}{\milli\second}), so asynchrony is necessary~\cite{DBLP:journals/corr/abs-1910-04054}.

	\item Training deep RL models has its own costs (need many parallel simulations). What if we only have one stream of experience to learn from? E.g., evolution of the system is too difficult to predict and model. In general, DNNs can't be trained effectively from a single, online, real-time stream of experience.

	\item PDP hardware is advertised as supporting P4, EBPF... But almost all implementations compile to a more complex/expressive language. Presence of externs is almost universal when it comes to in-switch compute---i.e., Tofino Native Architecture\footnote{\url{    https://www.barefootnetworks.com/products/brief-p4-studio/}}, Netronome Micro-C\footnote{\url{https://www.netronome.com/products/datapath-programming-tools/}}, P4$\rightarrow$NetFPGA toolchain\footnote{\url{https://github.com/NetFPGA/P4-NetFPGA-public}}.
\end{itemize}

What aspects of the state-of-the-art are relevant here?

\begin{itemize}
	\item Hmmm...
\end{itemize}

\section{Techniques}

What does the hardware give us, and what can we assume are present on many other classes of device?
\begin{itemize}
	\item 96-core (\emph{management engine}, ME) (? check), grouped into \emph{islands} based on local capabilities. Fast transfer between MEs on same island (\emph{nearest neighbour registers}, NNRs)

	\begin{itemize}
		\item Multicore model common.
		\item Arch-specific: one program per ME. Enables cooperative scheduling between \numrange{4}{8} threads, zero-cost context switch due to register file (below).
	\end{itemize}

	\item Explicit memory hierarchy:

	\begin{itemize}
		\item Large register file split disjointly between threads. (likely common)
		\item Transparent cache.
		\item In order of speed? Reg file, External Mem
	\end{itemize}

	\item One program to a core---enables cooperative scheduling with zero-cost context switches.

\end{itemize}
Now how do we make this work?
\begin{itemize}
	\item No floating point math. Solution? Quantisation (cite some other papers who've used this to decent effect).

	\item Bump-in-the-wire design for SmartNIC.

	\item P4 for easy packet parsing, control packet extraction, passthrough and manipulation of other traffic.

	\item RL Island keeps freelist, P4 PIF Islands copy packets over (could in theory hack P4 firmware micro-C libraries to stop freeing resources and just pass those?).

	\item Control packets marked by custom \emph{experimenter} DSCP (as in \textcite{DBLP:conf/isca/LiLYCSH19}). Infrastructure can ensure that only trusted internal nodes can produce such packets.

	\item Control allows adaptation to any policy design assuming standard tile-coding, for any target application. Policies may be installed via block copy, or trained online. (So long as policy meets compiled maximum dimensions, constraints.)

	\item (not impl'd yet) How does state work? Packet passed in over work queue to control ME. source can be either the network (same mechanism as control packets)

	\item (not impl'd yet) How do actions work? 2 options.
	\begin{enumerate}
		\item Produce packet on wire according to config (custom format).
		\item Pass packet to local island/ME (requires both config and custom build).
	\end{enumerate}

	\item Built with \Textcite{DBLP:journals/tnsm/SimpsonRP20} in mind, but applicable to any RL application.

	\item New PDPs have encryption primitives: scope to include authentication etc. around control packets.

	\item (Not impl'd yet) Event loop can easily run on Control ME: timing APIs exposed to generate periodic events in lieu of something like \textcite{DBLP:conf/hotnets/IbanezABM19}.
\end{itemize}

\begin{figure}
	\centering
	\resizebox{\linewidth}{!}{\begin{tikzpicture}
		\node(remote){Remote};

		%%%

		\node[below=0.05 of remote](swpos){};
		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
		\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};

		%

		\node[right=2 of swpos](epos){};
		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
		\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.85,0)$) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\small \centering Estimator\\$g(\cdot)$\end{varwidth}};

		%

		\node[right=3.5 of epos](apos){};
		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
		\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent vNF};
		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};

		%

		\node[below=2.3 of remote.west](linestart){};
		\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
		\draw [dashed] (linestart) -- (lineend);

		%%%

		\node[below=2.4 of remote](local){Local};

		%%%

		\node[below=0.25 of local](aswpos){};
		\node[fill=white!80!uofgthistle, draw=black, minimum height=3cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
		\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\small \centering OpenFlow\\Tables\end{varwidth}};
		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\small \centering Stats\\Collector\end{varwidth}};
		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};

		%

		\node (avfpos) at ($(aswpos) + (3,0.2)$){};
		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
		\node[below right, inner sep=2pt] at (avf.north west) {\small Agent vNF};
		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\small \centering Flowstate\\Database\end{varwidth}};
		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\small \centering TRS\\Scheduler\end{varwidth}};
		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};

		%%%

		\tikzset{>=stealth}

		\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
		\draw[thick, ->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};

		\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};

		\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
		\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};

		\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};

		\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
		\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};

		\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
		\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
		\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
		\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
		\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};

		\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
		\node[right=0.3] at (avfsa.east) {\tiny Experience};
		\end{tikzpicture}}
	\caption{
		System architecture  for our RL-driven DDoS defence system.
		\label{fig:sys-arch}
	}
\end{figure}

\begin{figure}
	\centering
	\resizebox{1.2\linewidth}{!}{\begin{tikzpicture}
		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2.9cm](p4-master){};
		\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Master Island};
		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (p4-rules) at (p4-master.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};

		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2cm, below = 0.1 of p4-master](sw1){};
		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
		\node[below right, inner sep=2pt] at (switch.north west) {\small P4 Island};
		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};

		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2.9cm, below= 0.1 of switch](others){};
		\node[below right, inner sep=2pt] at (others.north west) {\small Other Islands};
		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-etc) at (others.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering \textellipsis \end{varwidth}};
		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-int) at (others-etc.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering In-band Network Telemetry\end{varwidth}};

		\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (3, -2)$){};
		\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};

		\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
		\node[below right, inner sep=2pt] at (wq-out.north west) {\small CLS};
		\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Work Queue\end{varwidth}};

		\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (loop) at ($(wq-out.east) + (1.5,-0.5)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Control ME\end{varwidth}};

		\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-cls-out) at ($(loop.south) + (0,-1.5)$){};
		\node[below right, inner sep=2pt] at (policy-cls-out.north west) {\small CLS};
		\node[above right, inner sep=2pt] at (policy-cls-out.north west) {\small Policy};
		\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-ctm-out) at ($(policy-cls-out.south) + (0,-0.25)$){};
		\node[below right, inner sep=2pt] at (policy-ctm-out.north west) {\small CTM};
		\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-emem-out) at ($(policy-ctm-out.south) + (0,-0.25)$){};
		\node[below right, inner sep=2pt] at (policy-emem-out.north west) {\small EMEM};

		\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t1) at (policy-cls-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T1\end{varwidth}};
		\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t2) at (policy-ctm-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T2\end{varwidth}};
		\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t3) at (policy-emem-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T3\end{varwidth}};

		\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-1) at ($(wq-out.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

		\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-2) at ($(po-me-1.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

		\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-3) at ($(po-me-2.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

		\draw[->] (loop.north) [out=90,in=45,looseness=0.75] to node[pos=0.5, rotate=0, anchor=center, above] {Rules, Tables}(p4-rules.north east);

		\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2]  node[pos=0.5, rotate=90, anchor=center, above] {Rules}(sw1);
		\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (sw2);
		\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (switch);

		\draw[->] (oswlc.east)  to node[midway, above, sloped] {\begin{varwidth}{2.5cm}\tiny \centering Control, Requests,\\State\end{varwidth}} (wq-in.west);
		\draw[->] (others-int.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
		\draw[<->, dotted] (wq-in.east)  to (loop.west);
		\draw[->, dotted] (loop.south) to (policy-cls-out);

		\draw[<->, dotted] (po-me-1.east) to (policy-cls-out.west);
		\draw[<->, dotted] (po-me-2.east) to (policy-ctm-out.west);
		\draw[<->, dotted] (po-me-3.east) to (policy-emem-out.west);

		\draw[->] (loop) to node[midway, above, sloped] {\tiny NNR} (po-me-1);
		\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NNR} (po-me-2);
		\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NNR} (po-me-3);
		\draw[->] (po-me-3) to [out=-45,in=0,looseness=2.15] node[midway, above, sloped] {\tiny NNR} (loop);
		\end{tikzpicture}}
	\caption{Architecture for generalised RL agent on Netronome hardware.\label{fig:netro-arch}}
\end{figure}

\section{Evaluation}

How do we evaluate this without looking into classification performance?
\begin{itemize}
	\item Analytically?
	\begin{enumerate}
		\item Prove that tile splitting scheme over memory hierarchy can improve performance.
		\begin{itemize}
			\item Hit rate of different tiles according to tile dimensionality?
			\item Access time for each tier of memory hierarchy.
		\end{itemize}
	\end{enumerate}
	\item Experimentally?
	\begin{enumerate}
		\item Create ``VNF'' testbed: mirror traffic from switch to a host for telemetry processing/flow state. Send flow state over to another node who computes RL actions. RL actions installed on switch.
		\begin{itemize}
			\item Why not co-host RL computation with telemetry? Could be separate functions, could want to ensure that neither impacts the performance of the other.
		\end{itemize}
		\item Create ``PDP'' testbed: Load firmware, pass in traffic. Traffic processing/action compute/table mod occurs in PDP.
		\item What traffic? Healthy mix of attack (UDP), Discord legit Opus VoIP traffic, HTTP bulk-transfer traffic.
		\item Measure $t_{1 \cdots 3}$ according to \cref{fig:state-slip} in both VNF and PDP contexts.
		\begin{description}
			\item[$t_1$]
			\begin{description}
				\item[PDP] Packet ingress $\rightarrow$ INT Island $\rightarrow$ State arrival in Control ME.
				\item[VNF] Packet mirror time $\rightarrow$ telemetry VNF $\rightarrow$ state arrival in agent vnf.
			\end{description}

			\item[$t_2$]
			\begin{description}
				\item[PDP] Time to compute action over parallel Policy MEs (RL Island, SmartNIC).
				\item[VNF] Time to compute action in host program (on VNF) from received state.
			\end{description}

			\item[$t_3$]
			\begin{description}
				\item[PDP] Modify local tables across ME/Island boundaries (same device), verify rule present (or get ACK).
				\item[VNF] Generate OpenFlow control messages, send OF message to switch, verify rule present (or get ACK).
			\end{description}
		\end{description}
		\item HYPOTHESIS: PDP will present a reduction in these three times for per-flow state.
		\item QUESTION: Does same apply to learning?
		\item NOTE: since we're not actually assessing the \emph{quality} of output decisions, we can cut corners on what data is sent out and where. I.e., in-band network telemetry island can be skipped, if we move over representative volumes of state to/from the RL island with approximately correct delay as recorded by other studies.
	\end{enumerate}
\end{itemize}
I can think of one way to evaluate this if we employ classification performance:
\begin{enumerate}
	\item Run the existing work~\cite{DBLP:journals/tnsm/SimpsonRP20}, acquire a policy for one node after full info sharing.
	\item Load this policy onto the RL Island (\cref{fig:netro-arch}).
	\item Either feed in raw state packets, or feed in traffic if we can get INT working.
	\item For each level of quantisation needed, examine average performance.
	\item Don't compare against old numbers for final perf: use a ``VNF'' setup similar to above---this will probably suffer a bit compared to old results, where everything was on the same machine. Make it relative/normalised.
\end{enumerate}

\section{Related Work}

\begin{itemize}
	\item A switch/NIC architecture like \textcite{DBLP:conf/hotnets/StephensAS18} would remove the need for hard-coded island-island relationships.
\end{itemize}

\section{Conclusion}

%\bibliographystyle{ACM-Reference-Format}
%\bibliography{reference}
\printbibliography

\end{document}

### 2020-10-20 paper/paper.tex
Design notes, some fixes.
Currently, issue in tile coding to unravel...

@@
 
+\newpage
+
+\section{Bullets on (intended) design.}
+See also: \cref{fig:netro-arch}.
+\begin{itemize}
+	\item High-level overview.
+	\begin{itemize}
+		\item RL operations \emph{do not happen on the core path}.
+		\item Another island (or set of cores) is dedicated to handling inputs, configuration, computation.
+		\item Cross-island communication happens over \emph{rings} (Netronome-specific).
+		\begin{itemize}
+			\item E.g., this is how the P4 firmware propagates rules from master island to P4 workers.
+		\end{itemize}
+		\item Controller is \emph{always} co-hosted, interfaces over PCI-e. This also applies to Barefoot Tofino.
+	\end{itemize}
+	\item Policy configuration.
+	\begin{itemize}
+		\item (Re)-configurable at runtime.
+		\item Custom DSCP marker acts as indicator of RL packets (i.e., bother with parsing past UDP).
+		\item P4 parser extracts headers, parser state and payload passed over ring to another core dedicated to RL operations.
+		\item Policy dimensions, size, action count, tiling strategies, learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) can all be chosen and changed live.
+	\end{itemize}
+	\item State Handling.
+	\begin{itemize}
+		\item State input to model \emph{always} should produce an action.
+		\item State vectors should come from elsewhere on the device: this may either be a dedicated island/core cluster (i.e., also performing INT management), or it may be a packet from the network.
+	\end{itemize}
+	\item Action Handling.
+	\begin{itemize}
+		\item A state vector is tile coded, action probabilities are produced, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
+		\item \emph{Open question: how to configure action destination?
+This should either be another core (after an optional transformation), a packet sent out to a destination, maybe talk back to the controller.}
+	\end{itemize}
+	\item Learning.
+	\begin{itemize}
+		\item \emph{Open question: how to safely \& reasonably map state back to incoming rewards?}
+		\item \emph{Open question: Can/should this also support on-chip reward signals?}
+	\end{itemize}
+\end{itemize}
+
+\section{What's implemented 'til now?}
+System?
+\begin{itemize}
+	\item Core-to-core communication.
+	\item P4 parser for relevant packets.
+	\item Packet copy + handoff from P4 to RL MEs.
+	\item Dynamic base configuration.
+	\item Dynamic tiling configuration.
+	\item Policy installation via block-copy.
+	\item Packet generation for above config functions, as well as policy quantisation and splitting.
+	\item Tile-coding of quantised states on-chip, and state packet receipt.
+\end{itemize}
+
+Experiments?
+\begin{itemize}
+	\item Policy quantisation divergence \& accuracy (via mininet).
+\end{itemize}
+
+\section{Intermediate steps?}
+System?
+\begin{itemize}
+	\item Action computation on-chip.
+	\begin{itemize}
+		\item \emph{\num{1} week.}
+	\end{itemize}
+	\item Figure out PIF rule installation.
+	\begin{itemize}
+		\item \emph{\num{1} week.}
+		\item Rationale: understand and describe mechanism, do NOT implement. We know that we can save 1 controller RTT.
+	\end{itemize}
+	\item Setup other core as event source.
+	\begin{itemize}
+		\item \emph{\num{2} week.}
+		\item Idea: proxy for idea of INT core mentioned elsewhere, combined with the main periodic event sending that the last RL work relied upon.
+So, say "you can put that state management and extraction on another core, here's what the comms cost is" or similar.
+	\end{itemize}
+	\item Implement rudimentary training.
+	\begin{itemize}
+		\item \emph{\num{2} week.}
+	\end{itemize}
+\end{itemize}
+
+Experiments?
+\begin{itemize}
+	\item Measure DPDK SmartNIC RTT as a lower bound on controller access time.
+	\begin{itemize}
+		\item \emph{\num{0.5} weeks.}
+		\item Same mechanism applies to other manufacturers (Barefoot), so generalisable.
+	\end{itemize}
+	\item Measure on-switch times: ring-ring communication, tile-coding cost, action compute cost.
+	\begin{itemize}
+		\item \emph{\num{1} week.}
+	\end{itemize}
+	\item Measure times $t_{1...3}$ described elsewhere, or closest analogues in impl'd system.
+	\begin{itemize}
+		\item \emph{\num{1} week.}
+	\end{itemize}
+\end{itemize}
+
 \section{Related Work}

### 2020-10-20 code/nfp/rl-test/rl.c

@@ -28,6 +28,9 @@ volatile __declspec(export, emem, addr40, aligned(sizeof(unsigned int))) uint8_t
 
 volatile __declspec(export, emem) uint64_t really_really_bad = 0;
 
+__declspec(export, emem) uint32_t global_tc_indices[RL_MAX_TILE_HITS] = {0};
+__declspec(export, emem) uint16_t global_tc_count;
+
 __declspec(export, emem) struct rl_work_item test_work;
 
 /** Convert a state vector into a list of tile indices.
@@ -332,6 +335,12 @@ void state_packet(__addr40 _declspec(emem) struct rl_config *cfg, __declspec(xfe
 	}
 
 	tc_count = tile_code(state, cfg, tc_indices);
+	global_tc_count = tc_count;
+
+	i=0;
+	for (i=0; i<tc_count; i++) {
+		global_tc_indices[i] = tc_indices[i];
+	}
 }
 
 main() {

### 2020-10-20 code/nfp/rl-test/rl.h

@@ -27,7 +27,7 @@ typedef int32_t tile_t;
 #define T2_MAX_SETS 8
 #define T3_MAX_SETS 1
 
-#define RL_MAX_SETS T1_MAX_SETS+T2_MAX_SETS+T3_MAX_SETS
+#define RL_MAX_SETS (T1_MAX_SETS+T2_MAX_SETS+T3_MAX_SETS)
 
 enum tile_location {
 	TILE_LOCATION_T1,

### 2020-10-21 code/nfp/rl-test/rl-pkt-store.h
Tile coding working marvelously (on single-tiling configs).
On that note, also special cased them to maximise coverage of the state space.

@@ -5,7 +5,7 @@
 #include <nfp.h>
 
 #define RL_PKT_MAX_SZ 1504
-#define RL_PKT_STORE_COUNT 10
+#define RL_PKT_STORE_COUNT 100
 
 struct rl_pkt_store {
 	// Need a mutex

### 2020-10-21 code/nfp/rl-test/rl.c

@@ -41,20 +41,31 @@ uint16_t tile_code(tile_t *state, __addr40 _declspec(emem) struct rl_config *cfg
 	uint16_t tiling_set_idx;
 	uint16_t out_idx = 0;
 
+	// Tiling sets: each relies upon its own dimension list.
+	// Each of these is one entry in a tiling packet.
 	for (tiling_set_idx = 0; tiling_set_idx < cfg->num_tilings; ++tiling_set_idx) {
 		uint16_t tiling_idx;
 		uint32_t local_tile_base = cfg->tiling_sets[tiling_set_idx].start_tile;
 
+		// Repeat the selected tiling set, changing shift.
+		// Each of these is a *copy* of the above loop, shifted along by the relevant indices.
 		for (tiling_idx = 0; tiling_idx < cfg->tilings_per_set; ++tiling_idx) {
 			uint16_t dim_idx;
-			uint32_t width_product = 1;
+			// We can either divide the tile base, and mult all the tile indices later.
+			// Or... we could do it here, and save the trouble!
+			// (ordinarily, this would start at 1 for "true tile indices" (action-free))
+			uint32_t width_product = cfg->num_actions;
 			uint32_t local_tile = local_tile_base;
 
-			for (dim_idx = 0; dim_idx < cfg->tiling_sets[tiling_idx].num_dims; ++dim_idx) {
-				uint16_t active_dim = cfg->tiling_sets[tiling_idx].dims[dim_idx];
+			// These goes over each dimension in the top-level loop
+			for (dim_idx = 0; dim_idx < cfg->tiling_sets[tiling_set_idx].num_dims; ++dim_idx) {
+				uint16_t active_dim = cfg->tiling_sets[tiling_set_idx].dims[dim_idx];
 				uint8_t reduce_tile;
 				tile_t val = state[active_dim];
+
+				// This is what varies per tiling in a set.
 				tile_t shift = tiling_idx * cfg->shift_amt[active_dim];
+
 				tile_t max = cfg->adjusted_maxes[active_dim] - shift;
 				tile_t min = cfg->mins[active_dim] - shift;
 
@@ -165,10 +176,22 @@ void setup_packet(__addr40 _declspec(emem) struct rl_config *cfg, __declspec(xfe
 	);
 
[…]

### 2020-10-21 code/nfp/rl-test/rl.h

@@ -97,7 +97,7 @@ struct rl_config {
 
 	// 1-D width of a tile in each dimension.
 	// must be computed!
-	// this is w' = (max - min) / (n_tiles_per_dim + 1)
+	// this is w' = (max - min) / (n_tiles_per_dim - 1)
 	tile_t width[RL_DIMENSION_MAX]; // 592 -> 672
 	// every tiling in the same set is then shifted by n * this.
 	// This is w'/tilings_per_set

### 2020-10-21 code/control/state.sh
State packets 'r' here.

@@ -1,0 +1,1 @@
+cd target/release && sudo ./control -i vf0_0 state --src-ip 192.168.0.4 ../../config/example-state.json

### 2020-10-21 code/nfp/rl-test/rl.c
Action preference computation.
Seems to be correct, going back to manual examination.

@@ -31,6 +31,8 @@ volatile __declspec(export, emem) uint64_t really_really_bad = 0;
 __declspec(export, emem) uint32_t global_tc_indices[RL_MAX_TILE_HITS] = {0};
 __declspec(export, emem) uint16_t global_tc_count;
 
+__declspec(export, emem) tile_t global_prefs[MAX_ACTIONS] = {0};
+
 __declspec(export, emem) struct rl_work_item test_work;
 
 /** Convert a state vector into a list of tile indices.
@@ -86,6 +88,51 @@ uint16_t tile_code(tile_t *state, __addr40 _declspec(emem) struct rl_config *cfg
 		}
 	}
 	return out_idx;
+}
+
+/** Convert a tile list into a list of action preferences.
+*
+* Length written into act_list is guaranteed to be equal to cfg->num_actions.
+* Note, that act_list must be at least of size MAX_ACTIONS.
+*/
+void action_preferences(uint32_t *tile_indices, uint16_t tile_hit_count, __addr40 _declspec(emem) struct rl_config *cfg, tile_t *act_list) {
+	enum tile_location loc = TILE_LOCATION_T1;
+	uint16_t i = 0;
+	uint16_t j = 0;
+	uint16_t act_count = cfg->num_actions;
+
+	for (i = 0; i < tile_hit_count; i++) {
+		uint32_t target = tile_indices[i];
+		uint32_t loc_base = cfg->last_tier_tile[loc];
+		uint32_t base;
+
+		while (target >= loc_base) {
+			loc++;
+			loc_base = cfg->last_tier_tile[loc];
+		}
+
+		// find location of tier in memory.
+		switch (loc) {
+			case TILE_LOCATION_T1:
+				base = target;
+				for (j = 0; j < act_count; j++) {
+					act_list[j] += t1_tiles[base + j];
+				}
+				break;
+			case TILE_LOCATION_T2:
+				base = target - cfg->last_tier_tile[loc-1];
+				for (j = 0; j < act_count; j++) {
+					act_list[j] += t2_tiles[base + j];
+				}
+				break;
+			case TILE_LOCATION_T3:
+				base = target - cfg->last_tier_tile[loc-1];
+				for (j = 0; j < act_count; j++) {
+					act_list[j] += t3_tiles[base + j];
+				}
+				break;
[…]

### 2020-10-30 paper/paper.tex
Action preference computation.
Seems to be correct, going back to manual examination.
Fallback (non-dpdk) mirror timing.

@@
 %Conference
-%\acmConference[Submitted for review to SIGCOMM]{}
-%\acmYear{2018}
+\acmConference[SOSR '21]{}
+%\acmYear{2021}
 %\copyrightyear{}
@@
 However, it cannot completely eliminate the effect of state evolving due to the costs of action computation, execution, and policy updates~\cite{DBLP:journals/firai/TravnikMSP18}.
+	\begin{itemize}
+		\item What can I do with these observations?
+		\begin{itemize}
+			\item Reordering steps in an actively learning system is a necessary part in reducing state-slippage on learning.
+			\item Minimising these effects needs us to build our system explicitly to reduce this further.
+		\end{itemize}
+		\item What can I do to address this?
+		\begin{itemize}
+			\item Reduce the time required to compute an action, to react to state, and to install that action in the network.
+			\item Ensure that state is measured at agent (on-chip, or close by).
+			\item Ensure actions are computed as close to where they occur as possible.
+		\end{itemize}
+		\item How are these being addressed?
+		\begin{itemize}
+			\item Place agents/decision-making as close to the switch as possible.
+		\end{itemize}
+		\item What applications is this affecting?
+		\begin{itemize}
+			\item Time-sensitive RL: DDoS mitigation, Congestion-control algorithms, ...
+		\end{itemize}
+		\item What would its solution be useful for?
+		\begin{itemize}
+			\item Make actions taken more ``correct'' (as underlying state change has been reduced).
+			\item In online settings, reduce the noise of the state-action-reward function being learned.
+		\end{itemize}
+	\end{itemize}
 
@@
 In case of congestion control, significant throughput drops (as DNN inference takes \SI{30}{\milli\second}), so asynchrony is necessary~\cite{DBLP:journals/corr/abs-1910-04054}.
+	\begin{itemize}
+		\item What can I do with these observations?
+		\begin{itemize}
+			\item We can't pause operation of a system to collect more state, while updating the model, or while choosing an action.
+		\end{itemize}
+		\item What can I do to address this?
+		\begin{itemize}
+			\item Use state coalescing measures like in \textcite{DBLP:journals/tnsm/SimpsonRP20} and \textcite{DBLP:journals/corr/abs-1910-04054}.
+		\end{itemize}
+		\item What applications is this affecting?
+		\begin{itemize}
+			\item Congestion control algorithms primarily, others where some system depends on the output of an RL agent.
+		\end{itemize}
+	\end{itemize}
 
@@
 In general, DNNs can't be trained effectively from a single, online, real-time stream of experience.
+	\begin{itemize}
+		\item What can I do with these observations?
+		\begin{itemize}
+			\item This suggests that we should also look into non-DNN based RL approaches, if we find that RL is the best tool for the job.
+		\end{itemize}
+		\item What can I do to address this?
+		\begin{itemize}
+			\item Employ classical RL methods, which don't need many streams of parallel execution traces to learn effectively.
+		\end{itemize}
+		\item If addressed, what do I use them for?
+		\begin{itemize}
+			\item There's ongoing research which attempts to execute NN/DNN-based RL/ML models in switches, but these work under heavy concessions.
+			\item Training and format conversion must happen off of the device, so can't be done online.
+		\end{itemize}
+		\item What applications is this affecting?
+		\begin{itemize}
+			\item In-network RL applications needing real-time evolution, where DNN online training is not doable.
+			\item Likely invites some accuracy impact---linear (e.g., tile-coded) function approximation not as expressive or powerful as DNNs.
+		\end{itemize}
+		\item What would its solution be useful for?
+		\begin{itemize}
+			\item Ensures that online learning is possible in the contexts we're interested in.
+		\end{itemize}
+	\end{itemize}
 
@@
 Presence of externs is almost universal when it comes to in-switch compute---i.e., Tofino Native Architecture\footnote{\url{    https://www.barefootnetworks.com/products/brief-p4-studio/}}, Netronome Micro-C\footnote{\url{https://www.netronome.com/products/datapath-programming-tools/}}, P4$\rightarrow$NetFPGA toolchain\footnote{\url{https://github.com/NetFPGA/P4-NetFPGA-public}}.
+	\begin{itemize}
+		\item What can I do with these observations?
+		\begin{itemize}
+			\item We should use P4 where possible to make implementation of in-network compute possible.
+			\item \emph{But} we would be remiss not to make use of the multicore, general-purpose compute that the core on many of these devices are fundamentally built to do.
+			\item In-network compute and/or PDPs give us the tools we need to put action computation, the mechanism of action, and state measurement at the same location in the network!
+		\end{itemize}
+		\item If addressed, what do I use them for?
+		\begin{itemize}
+			\item Identify the common aspects of non-NetFPGA P4-capable devices.
+			\item Determine whether these allow a common architecture (but not common reimplementation since different languages, low-level specifics).
+			\item Might need to describe the same mechanism as implemented in both Netronome and Barefoot?
+		\end{itemize}
+		\item What applications is this affecting?
+		\begin{itemize}
+			\item PDP applications in general.
+		\end{itemize}
+	\end{itemize}
 \end{itemize}
@@
 			\item E.g., this is how the P4 firmware propagates rules from master island to P4 workers.
+			\item My impl.\ specific---shared freelist for memory pointers in EMEM.
 		\end{itemize}

### 2020-10-30 code/dpdk-mirror-timer/main.c

@@ -1,47 +1,318 @@
 /* SPDX-License-Identifier: BSD-3-Clause
- * Copyright(c) 2010-2014 Intel Corporation
+ * Copyright(c) 2010-2015 Intel Corporation
  */
 
-#include <stdio.h>
-#include <string.h>
 #include <stdint.h>
-#include <errno.h>
-#include <sys/queue.h>
-
-#include <rte_memory.h>
-#include <rte_launch.h>
+#include <inttypes.h>
+#include <getopt.h>
 #include <rte_eal.h>
-#include <rte_per_lcore.h>
+#include <rte_ethdev.h>
+#include <rte_cycles.h>
 #include <rte_lcore.h>
-#include <rte_debug.h>
+#include <rte_mbuf.h>
 
-static int
-lcore_hello(__rte_unused void *arg)
+#define RX_RING_SIZE 1024
+#define TX_RING_SIZE 1024
+
+#define NUM_MBUFS 8191
+#define MBUF_CACHE_SIZE 250
+#define BURST_SIZE 32
+
+static const char usage[] =
+	"%s EAL_ARGS -- [-t]\n";
+
+static const struct rte_eth_conf port_conf_default = {
+	.rxmode = {
+		.max_rx_pkt_len = RTE_ETHER_MAX_LEN,
+	},
+};
+
+static struct {
+	uint64_t total_cycles;
+	uint64_t total_queue_cycles;
+	uint64_t total_pkts;
+} latency_numbers;
+
+int hw_timestamping;
+
+#define NUM_MEASUREMENTS 100000
+
+struct packet_times {
+	uint64_t out;
+	uint64_t in;
+};
+
+static packet_times time_store[NUM_MEASUREMENTS] = {0};
+
+#define TICKS_PER_CYCLE_SHIFT 16
+static uint64_t ticks_per_cycle_mult;
+
+static uint16_t
+add_timestamps(uint16_t port __rte_unused, uint16_t qidx __rte_unused,
+		struct rte_mbuf **pkts, uint16_t nb_pkts,
+		uint16_t max_pkts __rte_unused, void *_ __rte_unused)
 {
-	unsigned lcore_id;
-	lcore_id = rte_lcore_id();
-	printf("hello from core %u\n", lcore_id);
+	unsigned i;
+	uint64_t now = rte_rdtsc();
+
+	for (i = 0; i < nb_pkts; i++)
+		pkts[i]->udata64 = now;
+	return nb_pkts;
+}
+
[…]

### 2020-10-30 code/mirror-timer/src/config.rs

@@ -1,0 +1,60 @@
+use super::*;
+
+pub struct GlobalConfig {
+	pub iface: NetworkInterface,
+	pub channel: Channel,
+}
+
+impl GlobalConfig {
+	pub fn new(iface_name: Option<&str>) -> Self {
+		let iface = datalink::interfaces()
+			.into_iter()
+			.find(move |iface: &NetworkInterface| {
+				if let Some(name) = iface_name {
+					iface.name == name
+				} else {
+					true
+				}
+			})
+			.unwrap_or_else(|| panic!("Couldn't bind interface \"{}\"", iface_name.unwrap_or("")));
+
+		let channel = datalink::channel(&iface, Default::default())
+			.map_err(|e| eprintln!("{:?}", e))
+			.unwrap_or_else(|_| {
+				panic!("Failed to open channel on \"{}\"", iface_name.unwrap_or(""))
+			});
+
+		Self { iface, channel }
+	}
+}
+
+pub struct TransportConfig {
+	pub src_addr: SocketAddrV4,
+	pub dst_addr: SocketAddrV4,
+}
+
+impl Default for TransportConfig {
+	fn default() -> Self {
+		let a = SocketAddrV4::new(Ipv4Addr::new(0, 0, 0, 0), 0);
+		Self {
+			src_addr: a,
+			dst_addr: a,
+		}
+	}
+}
+
+pub struct Config<'a> {
+	pub global: &'a mut GlobalConfig,
+	pub transport: TransportConfig,
+	pub iterations: u64,
+}
+
+impl<'a> Config<'a> {
+	pub fn new(global: &'a mut GlobalConfig, iterations: u64) -> Self {
+		Self {
+			global,
+			transport: Default::default(),
+			iterations,
+		}
+	}
+}

### 2020-10-30 code/mirror-timer/src/constants.rs

@@ -1,0 +1,1 @@
+pub const MAX_PKT_SIZE: usize = 1500;

### 2020-10-30 code/mirror-timer/src/lib.rs

@@ -1,0 +1,65 @@
+mod config;
+mod constants;
+mod packet;
+
+pub use self::{config::*, constants::*, packet::*};
+
+use byteorder::{BigEndian, WriteBytesExt};
+use enum_primitive::*;
+use pnet::{
+	datalink::{self, Channel, NetworkInterface},
+	packet::{
+		ethernet::*,
+		ip::IpNextHeaderProtocols,
+		ipv4::{self, *},
+		udp::{self, *},
+		PacketSize,
+	},
+	util::MacAddr,
+};
+use std::{
+	net::{Ipv4Addr, SocketAddrV4},
+	time::Instant,
+};
+
+pub fn list() {
+	println!("{:#?}", datalink::interfaces());
+}
+
+pub fn run(cfg: &mut Config) {
+	let mut measurements = vec![];
+	let mut buffer = [0u8; MAX_PKT_SIZE];
+
+	let (cursor, offsets) = build_transport(cfg.global, &cfg.transport, &mut buffer);
+
+	for i in 0..cfg.iterations {
+		let l_cursor = cursor + {
+			let mut body = &mut buffer[cursor..];
+			let space_start = body.len();
+
+			body.write_u64::<BigEndian>(i)
+				.expect("Eh?");
+			let space_end = body.len();
+
+			space_start - space_end
+		};
+
+		finalize(&mut buffer[..l_cursor], &offsets, &cfg.transport);
+
+		// Now send.
+
+		if let Channel::Ethernet(ref mut tx, ref mut rx) = &mut cfg.global.channel {
+			let t = Instant::now();
+			tx.send_to(&buffer[..l_cursor], None);
+
+			let _bytes = rx.next().expect("Argh?!");
+			let elapsed = t.elapsed();
+
+			measurements.push(elapsed.as_nanos());
+		} else {
+			eprintln!("Failed to send packet: no channel found");
+		}
+	}
+
+	println!("{:?}", measurements);
+}

### 2020-10-30 code/mirror-timer/src/main.rs

@@ -1,0 +1,110 @@
+use clap::{App, Arg, SubCommand};
+use mirror_timer::*;
+use std::net::IpAddr;
+
+fn main() {
+	let matches = App::new("(Netronome) Mirror timer")
+		.version("0.1")
+		.author("Kyle A. Simpson")
+		.about("Time RTT for in-switch RL firmware.")
+		.arg(Arg::with_name("interface")
+			.short("i")
+			.long("interface")
+			.value_name("IFACE")
+			.help(
+				"Interface to send packets over.\
+				If none is named, the first available interface will be used.")
+			.takes_value(true))
+		.subcommand(SubCommand::with_name("list")
+			.about("List all bindable interfaces."))
+		.subcommand(SubCommand::with_name("time")
+			.about("Sends packets to a netronome interface, timing the RTT to mirror each.")
+			.arg(Arg::with_name("src-ip")
+				.short("s")
+				.long("src-ip")
+				.value_name("IPV4")
+				.help("Source IP for control packets. Defaults to selected interface.")
+				.takes_value(true)
+				.default_value("192.168.1.2"))
+			.arg(Arg::with_name("dst-ip")
+				.short("d")
+				.long("dst-ip")
+				.value_name("IPV4")
+				.help("Destination IP for control packets. Defaults to selected interface.")
+				.takes_value(true)
+				.default_value("192.168.1.141"))
+			.arg(Arg::with_name("src-port")
+				.short("p")
+				.long("src-port")
+				.value_name("U16")
+				.help("Source UDP port for control packets.")
+				.takes_value(true)
+				.default_value("16767"))
+			.arg(Arg::with_name("dst-port")
+				.short("P")
+				.long("dst-port")
+				.value_name("U16")
+				.help("Destination port for control packets.")
+				.takes_value(true)
+				.default_value("16768"))
+			.arg(Arg::with_name("num-points")
+				.help("Number of RTT measurements to take.")
+				.takes_value(true)
+				.required(true)
+				.default_value("100000")))
+		.get_matches();
+
+	match matches.subcommand() {
+		("list", Some(_sub_m)) => {
+			mirror_timer::list();
+		},
+		("time", Some(sub_m)) => {
+			let mut g_cfg = GlobalConfig::new(matches.value_of("interface"));
+
+			let iterations = sub_m.value_of("num-points")
[…]

### 2020-10-30 code/mirror-timer/src/packet.rs

@@ -1,0 +1,87 @@
+use super::*;
+
+pub struct TransportOffsets {
+	ip: usize,
+	udp: usize,
+	data: usize,
+}
+
+pub fn build_transport(
+	cfg: &GlobalConfig,
+	t_cfg: &TransportConfig,
+	buf: &mut [u8],
+) -> (usize, TransportOffsets) {
+	let mut out = TransportOffsets {
+		ip: 0,
+		udp: 0,
+		data: 0,
+	};
+
+	let mut cursor = 0;
+	{
+		let mut eth_pkt =
+			MutableEthernetPacket::new(&mut buf[cursor..]).expect("Plenty of room...");
+		eth_pkt.set_destination(MacAddr::broadcast());
+		eth_pkt.set_ethertype(EtherTypes::Ipv4);
+		eth_pkt.set_source(cfg.iface.mac_address());
+
+		cursor += eth_pkt.packet_size();
+	}
+
+	{
+		out.ip = cursor;
+
+		let mut ipv4_pkt = MutableIpv4Packet::new(&mut buf[cursor..]).expect("Plenty of room...");
+		ipv4_pkt.set_version(4);
+		ipv4_pkt.set_header_length(5);
+		ipv4_pkt.set_ttl(64);
+		ipv4_pkt.set_next_level_protocol(IpNextHeaderProtocols::Udp);
+		ipv4_pkt.set_destination(*t_cfg.dst_addr.ip());
+		ipv4_pkt.set_source(*t_cfg.src_addr.ip());
+
+		// ipv4_pkt.set_dscp(RL_DSCP_TRAP);
+
+		cursor += ipv4_pkt.packet_size();
+	}
+
+	{
+		out.udp = cursor;
+
+		let mut udp_pkt = MutableUdpPacket::new(&mut buf[cursor..]).expect("Plenty of room...");
+		udp_pkt.set_source(t_cfg.src_addr.port());
+		udp_pkt.set_destination(t_cfg.dst_addr.port());
+		// checksum is optional in ipv4
+		udp_pkt.set_checksum(0);
+
+		cursor += udp_pkt.packet_size();
+	}
+
+	out.data = cursor;
+
+	(cursor, out)
+}
+
+pub fn finalize(buf: &mut [u8], off: &TransportOffsets, t_cfg: &TransportConfig) {
+	{
+		let region = &mut buf[off.ip..];
+		let len = region.len();
[…]

### 2020-10-30 code/mirror-timer/src/lib.rs
Non-DPDK numbers

@@ -37,8 +37,7 @@ pub fn run(cfg: &mut Config) {
 			let mut body = &mut buffer[cursor..];
 			let space_start = body.len();
 
-			body.write_u64::<BigEndian>(i)
-				.expect("Eh?");
+			body.write_u64::<BigEndian>(i).expect("Eh?");
 			let space_end = body.len();
 
 			space_start - space_end
@@ -61,5 +60,7 @@ pub fn run(cfg: &mut Config) {
 		}
 	}
 
-	println!("{:?}", measurements);
+	for el in &measurements[..] {
+		println!("{}", el);
+	}
 }

### 2020-10-30 code/mirror-timer/src/main.rs

@@ -7,51 +7,65 @@ fn main() {
 		.version("0.1")
 		.author("Kyle A. Simpson")
 		.about("Time RTT for in-switch RL firmware.")
-		.arg(Arg::with_name("interface")
-			.short("i")
-			.long("interface")
-			.value_name("IFACE")
-			.help(
-				"Interface to send packets over.\
-				If none is named, the first available interface will be used.")
-			.takes_value(true))
-		.subcommand(SubCommand::with_name("list")
-			.about("List all bindable interfaces."))
-		.subcommand(SubCommand::with_name("time")
-			.about("Sends packets to a netronome interface, timing the RTT to mirror each.")
-			.arg(Arg::with_name("src-ip")
-				.short("s")
-				.long("src-ip")
-				.value_name("IPV4")
-				.help("Source IP for control packets. Defaults to selected interface.")
-				.takes_value(true)
-				.default_value("192.168.1.2"))
-			.arg(Arg::with_name("dst-ip")
-				.short("d")
-				.long("dst-ip")
-				.value_name("IPV4")
-				.help("Destination IP for control packets. Defaults to selected interface.")
-				.takes_value(true)
-				.default_value("192.168.1.141"))
-			.arg(Arg::with_name("src-port")
-				.short("p")
-				.long("src-port")
-				.value_name("U16")
-				.help("Source UDP port for control packets.")
-				.takes_value(true)
-				.default_value("16767"))
-			.arg(Arg::with_name("dst-port")
-				.short("P")
-				.long("dst-port")
-				.value_name("U16")
-				.help("Destination port for control packets.")
-				.takes_value(true)
-				.default_value("16768"))
-			.arg(Arg::with_name("num-points")
-				.help("Number of RTT measurements to take.")
-				.takes_value(true)
-				.required(true)
-				.default_value("100000")))
+		.arg(
+			Arg::with_name("interface")
+				.short("i")
+				.long("interface")
+				.value_name("IFACE")
+				.help(
+					"Interface to send packets over.\
+				If none is named, the first available interface will be used.",
+				)
+				.takes_value(true),
+		)
+		.subcommand(SubCommand::with_name("list").about("List all bindable interfaces."))
+		.subcommand(
+			SubCommand::with_name("time")
+				.about("Sends packets to a netronome interface, timing the RTT to mirror each.")
+				.arg(
+					Arg::with_name("src-ip")
+						.short("s")
+						.long("src-ip")
[…]

### 2020-10-30 code/nfp/bump/altmirror.sh

@@ -1,0 +1,1 @@
+rtecli design-load -f out/nfp_6x_40.nffw -p out/pif_design.json -c truemirror.p4cfg

### 2020-10-30 paper/paper.tex
Appended notes to motivation.

@@
 			\item There's ongoing research which attempts to execute NN/DNN-based RL/ML models in switches, but these work under heavy concessions.
+Need binarisation, reduced accuracy, can't run backpropagation on SmartNIC hardware...
 			\item Training and format conversion must happen off of the device, so can't be done online.
@@
 			\item Ensures that online learning is possible in the contexts we're interested in.
+			\item I.e., DDoS prevention (evolving problem).
 		\end{itemize}
@@
 			\item We should use P4 where possible to make implementation of in-network compute possible.
-			\item \emph{But} we would be remiss not to make use of the multicore, general-purpose compute that the core on many of these devices are fundamentally built to do.
-			\item In-network compute and/or PDPs give us the tools we need to put action computation, the mechanism of action, and state measurement at the same location in the network!
+			\item \emph{But} we would be remiss not to make use of the multicore, general-purpose compute that many of these devices are fundamentally built to do.
+			\item In-network compute and/or PDPs give us the tools we need to put RL action computation, the mechanism of action, and state measurement at the same location in the network!
 		\end{itemize}
@@
 
-\newpage
-
 \section{Bullets on (intended) design.}

### 2020-12-02 interim-talk/talk.tex
Talk!

--- 2020-04-01 motivation-talk/talk.tex
@@
 %,show notes on second screen=right
-,handout
+%,handout
 ]{beamer}
@@
 \usepackage{lipsum}
-\usepackage[basic]{complexity}
+%\usepackage[basic]{complexity}
 \usepackage[super,negative]{nth}
@@
 \usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,sortcites
-,backend=bibtex
+%,backend=bibtex
 ,style=authortitle
 ]{biblatex}
-%\addbibresource{papers.bib}
-%\addbibresource{confs-journs.bib}
+\addbibresource{../paper/papers-off.bib}
+\addbibresource{../paper/confs-off.bib}
 
@@
 
-\title{Reinforcement Learning in Future Networks}
+\title{Towards In-Switch Reinforcement Learning}
 \author{\vspace{-1em}\textbf{Kyle A. Simpson}\\
@@
 \institute{University of Glasgow}
-\date{\nth{27} March, 2020}
+\date{\nth{2} December, 2020}
 
@@
 
-\begin{frame}{Data-driven networking: DDoS Prevention}
-	\begin{columns}
-		\begin{column}{0.6\linewidth}
-			\begin{figure}
-				\centering
-				\resizebox{\linewidth}{!}{
-					\begin{tikzpicture}[
-					texts/.style = {text=black},
-					labeltexts/.style = {text=gray},
-					treeline/.style = {draw=uofgburgundy},
-					treenode/.style = {texts, circle, centered, fill=white, treeline},
-					load/.style = {fill=uofgcobalt},
-					loadline/.style = {draw=uofgcobalt, line width=0.75mm},
-					loadhide/.style = {fill=uofgcobalt!40!white},
-					external/.style = {fill=uofgrust},
-					externalhide/.style = {fill=uofgrust!40!white},
-					hideline/.style = {draw=uofgsandstone!40!white},
-					hidenode/.style = {treenode, hideline},
-					grow'=right
-					]
-					\node [treenode, loadhide, label={[texts]above:Agent 1}] (mainagent) {};
-					\node [treenode, loadhide, right = of mainagent] (inner0) {};
-					\node [treenode, loadhide, above right = 0.2cm and 1cm of inner0] (inner1) {};
-					\node [hidenode, below right = 0.2cm and 1cm of inner0] (inner2) {};
-					\node [treenode, loadhide, below right = 0.2cm and 1cm of inner1] (inner3) {};
-					\node [treenode, loadhide, right = of inner3] (inner4) {};
-					\node [treenode, loadhide, above right = 0.2cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
-					\node [hidenode, below right = 0.2cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};
-
-					\node [hidenode, above left = 0.2cm and 0.5cm of inner1, label={[labeltexts]above:Agent 2}] (agent2) {};
-					\node [hidenode, below left = 0.2cm and 0.5cm of inner2, label={[labeltexts]below:Agent 3}] (agent3) {};
-
-					\draw[hideline, -] ($ (mainagent) + (-1,0) $) -- (mainagent);
-					\draw[hideline, -] ($ (agent2) + (-1,0) $) -- (agent2);
-					\draw[hideline, -] ($ (agent3) + (-1,0) $) -- (agent3);
-					\draw[hideline, -] (inner1) -- (agent2);
-					\draw[hideline, -] (inner2) -- (agent3);
-
-					\draw[loadline, -] (mainagent) -- (inner0);
-					\draw[loadline, -] (inner0) -- (inner1);
-					\draw[hideline, -] (inner0) -- (inner2);
-					\draw[-] (inner1) -- (inner3);
-					\draw[hideline, -] (inner2) -- (inner3);
-					\draw[loadline, -] (inner3) -- (inner4);
-					\draw[loadline, -] (inner4) -- node [texts, above] {$U_{s_0}$} (s0) ;
-					\draw[hideline, -] (inner4) -- node [labeltexts, below] {$U_{s_1}$} (s1);
-
-					\end{tikzpicture}
-				}
-			\caption{`Global' state for any flow (\textbf{bold} means used by policy).}
-			\end{figure}
-		\end{column}
-		\begin{column}{0.4\linewidth}
-			\begin{itemize}
-				\item Monitor flow statistics at ingress/egress points.
-				\item Use load measurements from flow paths as global state.
-				\item OUTPUT: throttle individual flows.
-			\end{itemize}
-		\end{column}
-	\end{columns}
-\end{frame}
-
-\begin{frame}{Data-driven networking: DDoS Prevention}
-	\begin{columns}
-		\begin{column}{0.3\linewidth}
-			How does this perform compared to state of the art, SPIFFY...?
-			\begin{itemize}
-				\item Worse for congestion-aware, but strong.
-				\item \alert{Improvement on VoIP/UDP traffic!}
-				\item Learning time on order of minutes.
-				\begin{itemize}
-					\item Classical RL.
-				\end{itemize}
-			\end{itemize}
-		\end{column}
-		\begin{column}{0.7\linewidth}
-			\begin{figure}
-				\centering
-				\includegraphics[width=0.75\linewidth]{tnsm-tcp-16-single}
-				\caption{
-					Online performance of RL agents for TCP traffic, with and without policy sharing.
-					\label{fig:tcp-tree-16}
-				}
-			\end{figure}
-		\end{column}
-	\end{columns}
-\end{frame}
-
-\begin{frame}{Programmable data-planes: Motivation}
+\begin{frame}{Why programmable data-planes?}
 	\begin{columns}
@@
 			\begin{itemize}
-			\item Programmable network hardware: arbitrary in-network compute.
 				\item In data-driven, want to \alert{minimise time to act}.
 				\item RL assumes that action \& policy update are zero-cost.
+				\vspace{-1em}
 				\begin{itemize}
@@
 				\end{itemize}
+				\item Controller contact time, serialisation, ...
+				\item In other ML, often need line rate inference.
+				\item Programmable network hardware fills this niche.
 			\end{itemize}
@@
 
-\begin{frame}{DDoS Prevention: VNF + SDN  architecture}
+\begin{frame}{Recent Programmable Trends in Data-Driven Networks}
 	\begin{columns}
-%		\begin{column}{0.3\linewidth}
-%			\begin{itemize}
-%				\item Monitor flow statistics at ingress/egress points.
-%				\item Use load measurements from flow paths as global state.
-%				\item OUTPUT: throttle individual flows.
-%			\end{itemize}
-%		\end{column}
-		\begin{column}{1.0\linewidth}
+		\begin{column}{0.6\linewidth}
+			\begin{itemize}
+				\item ML acceleration, line-rate packet classification.
+				\item How? Train model off-NIC, convert to \alert{binary neural network}\footnotemark, or \alert{decision tree}\footnotemark.
+				\item Limits? No online training, cost of backprop algo (expensive!), vast data needs.
+				\item \alert{What if we need online learning?}
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.4\linewidth}
+			\centering
+			\resizebox{0.6\linewidth}{!}{
+				\includegraphics{images/smartnic.png}
+			}
+
+			\resizebox{0.5\linewidth}{!}{
+				\includegraphics{images/nn}
+			}
+		\end{column}
+	\end{columns}
+\setcounter{footnote}{1}
+\footcitetext{DBLP:journals/corr/abs-2009-02353}
+\setcounter{footnote}{2}
+\footcitetext{DBLP:conf/hotnets/XiongZ19}
+\end{frame}
+
+\begin{frame}{Case study: DDoS Prevention}
+	\begin{columns}
+		\begin{column}{0.4\linewidth}
+			\begin{figure}
+				\centering
+				\resizebox{\linewidth}{!}{
+					\begin{tikzpicture}[
+					texts/.style = {text=black},
+					labeltexts/.style = {text=gray},
+					treeline/.style = {draw=uofgburgundy},
+					treenode/.style = {texts, circle, centered, fill=white, treeline},
+					load/.style = {fill=uofgcobalt},
+					loadline/.style = {draw=uofgcobalt, line width=0.75mm},
+					loadhide/.style = {fill=uofgcobalt!40!white},
+					external/.style = {fill=uofgrust},
+					externalhide/.style = {fill=uofgrust!40!white},
+					hideline/.style = {draw=uofgsandstone!40!white},
+					hidenode/.style = {treenode, hideline},
+					grow'=right
+					]
+					\node [treenode, loadhide, label={[texts]above:Agent 1}] (mainagent) {};
+					\node [treenode, loadhide, right = of mainagent] (inner0) {};
+					\node [treenode, loadhide, above right = 0.2cm and 1cm of inner0] (inner1) {};
+					\node [hidenode, below right = 0.2cm and 1cm of inner0] (inner2) {};
+					\node [treenode, loadhide, below right = 0.2cm and 1cm of inner1] (inner3) {};
+					\node [treenode, loadhide, right = of inner3] (inner4) {};
+					\node [treenode, loadhide, above right = 0.2cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
+					\node [hidenode, below right = 0.2cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};
+
+					\node [hidenode, above left = 0.2cm and 0.5cm of inner1, label={[labeltexts]above:Agent 2}] (agent2) {};
+					\node [hidenode, below left = 0.2cm and 0.5cm of inner2, label={[labeltexts]below:Agent 3}] (agent3) {};
+
+					\draw[hideline, -] ($ (mainagent) + (-1,0) $) -- (mainagent);
+					\draw[hideline, -] ($ (agent2) + (-1,0) $) -- (agent2);
+					\draw[hideline, -] ($ (agent3) + (-1,0) $) -- (agent3);
+					\draw[hideline, -] (inner1) -- (agent2);
+					\draw[hideline, -] (inner2) -- (agent3);
+
+					\draw[loadline, -] (mainagent) -- (inner0);
+					\draw[loadline, -] (inner0) -- (inner1);
+					\draw[hideline, -] (inner0) -- (inner2);
+					\draw[-] (inner1) -- (inner3);
+					\draw[hideline, -] (inner2) -- (inner3);
+					\draw[loadline, -] (inner3) -- (inner4);
+					\draw[loadline, -] (inner4) -- node [texts, above] {$U_{s_0}$} (s0) ;
+					\draw[hideline, -] (inner4) -- node [labeltexts, below] {$U_{s_1}$} (s1);
+
+					\end{tikzpicture}
+				}
+			\caption{`Global' state for any flow (\textbf{bold} means used by policy).}
+			\end{figure}
+		\end{column}
+		\begin{column}{0.6\linewidth}
+			\begin{itemize}
+				\item Prior work\footnotemark.
+				\item Monitor flow statistics at ingress/egress points.
+				\item Use load measurements from flow paths as global state.
+				\item OUTPUT: throttle individual flows.
+				\item Classical RL.
+				\begin{itemize}
+					\item Learning time on order of minutes.
+				\end{itemize}
+			\end{itemize}
+		\end{column}
+	\end{columns}
+\footcitetext{DBLP:journals/tnsm/SimpsonRP20}
+\end{frame}
+
+\begin{frame}{Case Study: DDoS Prevention (Architecture)}
+	\begin{columns}
+		\begin{column}{0.3\linewidth}
+			\begin{itemize}
+				\item Reward measurements come from network.
+				\item Input state mixes local flow data with global load data.
+				\item Flow measurements combined, decisions scheduled to prevent overload.
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.7\linewidth}
 			\begin{figure}
@@
 				\caption{
-					System architecture for RL-driven DDoS defence.
+					(Non-PDP) system architecture for RL-driven DDoS defence.
 					\label{fig:sys-arch}
@@
 
-\begin{frame}{Programmable data-planes: An Architecture for RL}
+\begin{frame}{Goals... and Reality}
+	\centering
+	\begin{tabular}{c c c}
+		\toprule
+		Goal & Solution & Drawback \\ \midrule
+		$\bullet$ On-NIC action compute & \onslide+<2->{\alert<2>{Quantisation}} & \onslide+<2->{\alert<2>{Platform Model}} \\
+		$\bullet$ On-NIC learning & \onslide+<3->{\alert<3>{Classical RL}} & \onslide+<3->{\alert<3>{Info capacity}} \\
+		$\bullet$ Direct action installation & \onslide+<4->{\alert<4>{Custom actions}} & \onslide+<4->{\alert<4>{No intrinsic support}} \\
+		$\bullet$ (Partly) External State, Rewards, & \onslide+<5->{\alert<5>{Independent Module}} & \onslide+<5->{\alert<5>{---}} \\
+		and Task Independence & & \\
+		$\bullet$ Runtime reconfiguration & \onslide+<6->{\alert<6>{P4 Parser + RTE}} & \onslide+<6->{\alert<6>{---}} \\ \bottomrule
+	\end{tabular}
+\end{frame}
+
+\begin{frame}{Device classes}
+	\begin{columns}
+		\begin{column}{0.5\linewidth}
+			\alert{SmartNICs (e.g., Netronome)}
+
+			\begin{itemize}[<+->]
+				\item Low port density.
+				\item Easy to have asynchrony:
+				\item SOC-based $\rightarrow$ extra cores off the datapath.
+				\item NetFPGA $\rightarrow$ can define extra functional units and interconnect.
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.5\linewidth}
+			\alert{Programmable Switches (e.g., Tofino)}
+
+			\begin{itemize}[<+->]
+				\item High port density.
+				\item No additional compute units. Very close to P4 PSA.
+				\item But still more powerful--e.g., Tofino supports MUL operations.
+				\item Might be doable... deadline-aware.
+			\end{itemize}
+		\end{column}
+	\end{columns}
+
+	\pause
+	\vspace{1em}
+	We'll mainly focus on SmartNICs.
+\end{frame}
+
+\begin{frame}{Timing: Why not offload to the controller?}
+	\begin{columns}
+		\begin{column}{0.5\linewidth}
+			For SmartNICs, the attached host \emph{is the (closest) controller}.
+
+			\begin{itemize}
+				\item PCIe access times $\mathcal{O}(\si{\micro\second})$.\footnotemark
+				\item MATs recompiled $\mathcal{O}(\si{\second})$. Many rules $\implies$ batching.
+				\item Thrift serde time $\mathcal{O}(\si{\milli\second})$.
+			\end{itemize}
+
+			%	?? https://github.com/rpclib/benchmarks/blob/master/plots/get_blob_real_log.svg
+
+			\alert{Meanwhile core-to-core on chip around \SI{100}{\nano\second} @ \SI{1.2}{\giga\hertz}.}
+		\end{column}
+		\begin{column}{0.5\linewidth}
+			\begin{figure}
+				\includegraphics[width=0.9\linewidth]{../plots/build/rte-timer/rte-times-logx}
+				\caption{Netronome rule installation cost \\(1 table, \numrange{1}{65536} rules).}
+			\end{figure}
+		\end{column}
+	\end{columns}
+
+	\footcitetext{DBLP:conf/sigcomm/NeugebauerAZAL018}
+\end{frame}
+
+\begin{frame}{Quantisation}
+	\begin{columns}
+		\begin{column}{0.5\linewidth}
+			\begin{figure}
+				\includegraphics[width=0.9\linewidth]{../plots/build/marl-quant/accuracy-binary}
+				\caption{DDoS throttle accuracy wrt. fixed-point Q$n$.}
+			\end{figure}
+		\end{column}
+		\begin{column}{0.5\linewidth}
+			No FPU... so use \alert{Q$n$ fixed-point}.
+			\begin{itemize}
+				\item For RL updates: need only shift, multiply, add.
+				\item \alert{Policy exec only needs add}.
+				\item \SI{16}{\bit} fraction has negligible loss.
+			\end{itemize}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Dynamic Action Installation}
+	\begin{itemize}
+		\item Netronome w/ P4?
+		\begin{itemize}
+			\item MAT structure (DCFL) computed on host...
+			\item So can't directly interop with P4.
+			\item ...But we can have other actions/externs maintain their own lookup tables.
+		\end{itemize}
+		\item Tofino has features to make this more feasible.
+		\begin{itemize}
+			\item Action Profiles/Selects
+		\end{itemize}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Architecting On-NIC RL}
 	\begin{columns}
@@
 			\begin{figure}
-			\resizebox{1.2\linewidth}{!}{\begin{tikzpicture}
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2.9cm, below right= 0.1 of sw2.north west](p4-master){};
-				\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Master Island};
-				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (p4-rules) at (p4-master.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};
+				\resizebox{1\linewidth}{!}{\begin{tikzpicture}
+						\node[fill=white!80!uofgcobalt, draw=black, minimum height=5.5cm, minimum width=2.9cm, below right= 0.1 of sw2.north west](p4-master){};
+						\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Datapath};
+						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.8] (p4-rules) at (p4-master.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};
+						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n1) at (p4-rules.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Extern-based Action\end{varwidth}};
+						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n2) at (n1.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
+						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (oswlc) at (n2.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};
+						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n3) at (oswlc.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
+						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n4) at (n3.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Network Telemetry\end{varwidth}};
 
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2cm, below = 0.1 of p4-master](sw1){};
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
-				\node[below right, inner sep=2pt] at (switch.north west) {\small P4 Island};
-				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};
+						\node (pkt-text) at ($(p4-master.north) + (0.0,0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet In\end{varwidth}};
 
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2.9cm, below= 0.1 of switch](others){};
-				\node[below right, inner sep=2pt] at (others.north west) {\small Other Islands};
-				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-etc) at (others.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering \textellipsis \end{varwidth}};
-				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-int) at (others-etc.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering In-band Network Telemetry\end{varwidth}};
+						\node (pkt-text-out) at ($(p4-master.south) + (0.0,-0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet Out\end{varwidth}};
 
-				\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (3, -2)$){};
+						\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (5, -1)$){};
 						\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};
@@
 						\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
-				\node[below right, inner sep=2pt] at (wq-out.north west) {\small CLS};
-				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Work Queue\end{varwidth}};
+						\node[below right, inner sep=2pt] at (wq-out.north west) {\small EMEM};
+						\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Input Queue\end{varwidth}};
 
 						\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (loop) at ($(wq-out.east) + (1.5,-0.5)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Control ME\end{varwidth}};
+
+						\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-act) at ($(loop.north) + (0.0,1.5)$){};
+						\node[below right, inner sep=2pt] at (wq-act.north west) {\small EMEM};
+						\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-actin) at (wq-act.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Action Queue\end{varwidth}};
 
@@
 
-				\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2]  node[pos=0.5, rotate=90, anchor=center, above] {Rules}(sw1);
-				\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (sw2);
-				\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (switch);
-
 						\draw[->] (oswlc.east)  to node[midway, above, sloped] {\tiny Control, Requests} (wq-in.west);
-				\draw[->] (others-int.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
+						\draw[->] (n4.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
 						\draw[<->, dotted] (wq-in.east)  to (loop.west);
 						\draw[->, dotted] (loop.south) to (policy-cls-out);
+
+						\draw[->] (loop.north)  to node[midway, above, sloped] {\tiny Actions} (wq-actin.south);
+
+						\draw[->] (wq-actin.west)  to node[midway, above, sloped] {\tiny Actions} (n1.east);
 
@@
 						\draw[<->, dotted] (po-me-3.east) to (policy-emem-out.west);
+
+						\draw[->] (pkt-text)  to  (p4-master.north);
+						\draw[->] (p4-master.south)  to  (pkt-text-out);
 
@@
 			\begin{itemize}
-			\item Make decisions \alert{where they must be applied}.
-			\item ``Bump-in-the-wire'' use of Netronome SmartNIC.
-			\item P4 parsing/forwarding/filtering, offload to microcode for full programmability.
-			\begin{itemize}
-				\item P4 with offload common.
-			\end{itemize}
-				\item Solution: \alert{make use of on-device parallelism, externs to implement fast, adaptive data-driven applications}.
+				\item Classical RL built on tile-coding---\alert{online}.
+				\item Async wrt. datapath.
+				\item Dynamic selection of last reward, trace info.
+				\item \alert{Runtime configurable} (policy, size, application) from data/control-plane.
+				\item \alert{Task independent.}
 			\end{itemize}
@@
 
+\begin{frame}{How could this (hypothetically) fit the case study?}
+	\begin{itemize}
+		\item P4 action to gather state on collected flows.
+		\begin{itemize}
+			\item Flow telemetry in P4 well-documented.
+			\item Pass in/schedule state vectors every $\delta{}t$.
+		\end{itemize}
+		\item Directly pass reward measurements to RL core.
+		\item Base policy installed by controller, updated live.
+		\item Matched packets poll actions from RL core.
+		\begin{itemize}
+			\item Map RL actions to state machine, maintain throttling hash table.
+			\item Every $\delta{}t'$, batch actions to controller.
+		\end{itemize}
+	\end{itemize}
+\end{frame}
+
+%\begin{frame}{What about higher port-density devices?}
+%	content...
+%\end{frame}
+
+\begin{frame}{Execution costs (Q15.16)}
+	\begin{columns}
+		\begin{column}{0.4\linewidth}
+			\begin{itemize}
+				\item Single-threaded---still to be accelerated.
+				\item Small? One tiling per memory tier.
+				\item Versus use-case: \SI{330}{\micro\second} on commodity i7.
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.6\linewidth}
+			\centering
+			\begin{tabular}{ l c c }
+				\toprule
+				Policy Size & Action (\si{\micro\second}) & Plus Update (\si{\micro\second}) \\ \midrule
+				Small & \num{10.66} & \num{14.60} \\
+				Max (DDoS) & \num{512.67} & \num{612.55} \\ \bottomrule
+			\end{tabular}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}[standout]
+	\begin{itemize}
+		\item Online in-NIC RL is possible!
+		\item In-switch? Less so.
+		\item Arch-specific, but similar design for similar hardware class.
+		\item Work-in-progress: end-to-end timing, training accuracy, other use-cases, optimisation.
+	\end{itemize}
+	\alert{Questions?}
+\end{frame}
+
 \end{document}

### 2020-12-02 interim-talk/talk.tex
Finalise talk.

@@
 %,show notes on second screen=right
-%,handout
+,handout
 ]{beamer}
@@
 
-\begin{frame}{Timing: Why not offload to the controller?}
-	\begin{columns}
-		\begin{column}{0.5\linewidth}
-			For SmartNICs, the attached host \emph{is the (closest) controller}.
-
-			\begin{itemize}
-				\item PCIe access times $\mathcal{O}(\si{\micro\second})$.\footnotemark
-				\item MATs recompiled $\mathcal{O}(\si{\second})$. Many rules $\implies$ batching.
-				\item Thrift serde time $\mathcal{O}(\si{\milli\second})$.
-			\end{itemize}
-
-			%	?? https://github.com/rpclib/benchmarks/blob/master/plots/get_blob_real_log.svg
-
-			\alert{Meanwhile core-to-core on chip around \SI{100}{\nano\second} @ \SI{1.2}{\giga\hertz}.}
-		\end{column}
-		\begin{column}{0.5\linewidth}
-			\begin{figure}
-				\includegraphics[width=0.9\linewidth]{../plots/build/rte-timer/rte-times-logx}
-				\caption{Netronome rule installation cost \\(1 table, \numrange{1}{65536} rules).}
-			\end{figure}
-		\end{column}
-	\end{columns}
-
-	\footcitetext{DBLP:conf/sigcomm/NeugebauerAZAL018}
-\end{frame}
-
-\begin{frame}{Quantisation}
-	\begin{columns}
-		\begin{column}{0.5\linewidth}
-			\begin{figure}
-				\includegraphics[width=0.9\linewidth]{../plots/build/marl-quant/accuracy-binary}
-				\caption{DDoS throttle accuracy wrt. fixed-point Q$n$.}
-			\end{figure}
-		\end{column}
-		\begin{column}{0.5\linewidth}
-			No FPU... so use \alert{Q$n$ fixed-point}.
-			\begin{itemize}
-				\item For RL updates: need only shift, multiply, add.
-				\item \alert{Policy exec only needs add}.
-				\item \SI{16}{\bit} fraction has negligible loss.
-			\end{itemize}
-		\end{column}
-	\end{columns}
-\end{frame}
-
-\begin{frame}{Dynamic Action Installation}
-	\begin{itemize}
-		\item Netronome w/ P4?
-		\begin{itemize}
-			\item MAT structure (DCFL) computed on host...
-			\item So can't directly interop with P4.
-			\item ...But we can have other actions/externs maintain their own lookup tables.
-		\end{itemize}
-		\item Tofino has features to make this more feasible.
-		\begin{itemize}
-			\item Action Profiles/Selects
-		\end{itemize}
-	\end{itemize}
-\end{frame}
-
 \begin{frame}{Architecting On-NIC RL}
@@
 	\begin{itemize}
-		\item P4 action to gather state on collected flows.
+		\item P4 table-action to gather state on collected flows.
 		\begin{itemize}
@@
 
+\begin{frame}{Timing: Why not offload to the controller?}
+	\begin{columns}
+		\begin{column}{0.5\linewidth}
+			For SmartNICs, the attached host \emph{is the (closest) controller}.
+
+			\begin{itemize}
+				\item PCIe access times $\mathcal{O}(\si{\micro\second})$.\footnotemark
+				\item MATs recompiled $\mathcal{O}(\si{\second})$. Many rules $\implies$ batching.
+				\item Thrift serde time $\mathcal{O}(\si{\milli\second})$.
+			\end{itemize}
+
+			%	?? https://github.com/rpclib/benchmarks/blob/master/plots/get_blob_real_log.svg
+
+			\alert{Meanwhile core-to-core on chip around \SI{100}{\nano\second} @ \SI{1.2}{\giga\hertz}.}
+		\end{column}
+		\begin{column}{0.5\linewidth}
+			\begin{figure}
+				\includegraphics[width=0.9\linewidth]{../plots/build/rte-timer/rte-times-logx}
+				\caption{Netronome rule installation cost \\(1 table, \numrange{1}{65536} rules).}
+			\end{figure}
+		\end{column}
+	\end{columns}
+
+	\footcitetext{DBLP:conf/sigcomm/NeugebauerAZAL018}
+\end{frame}
+
+\begin{frame}{Quantisation}
+	\begin{columns}
+		\begin{column}{0.5\linewidth}
+			\begin{figure}
+				\includegraphics[width=0.9\linewidth]{../plots/build/marl-quant/accuracy-binary}
+				\caption{DDoS throttle accuracy wrt. fixed-point Q$n$.}
+			\end{figure}
+		\end{column}
+		\begin{column}{0.5\linewidth}
+			No FPU... so use \alert{Q$n$ fixed-point}.
+			\begin{itemize}
+				\item For RL updates: need only shift, multiply, add.
+				\item \alert{Policy exec only needs add}.
+				\item \SI{16}{\bit} fraction has negligible loss.
+			\end{itemize}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Dynamic Action Installation}
+	\begin{itemize}
+		\item Netronome w/ P4?
+		\begin{itemize}
+			\item MAT structure (DCFL) computed on host...
+			\item So can't directly interop with P4.
+			\item ...But we can have other actions/externs maintain their own lookup tables.
+		\end{itemize}
+		\item Tofino has features to make this more feasible.
+		\begin{itemize}
+			\item Action Profiles/Selects
+		\end{itemize}
+	\end{itemize}
+\end{frame}
+
 %\begin{frame}{What about higher port-density devices?}
@@
 				\item Small? One tiling per memory tier.
-				\item Versus use-case: \SI{330}{\micro\second} on commodity i7.
+				\item Max? 20-element state, 17 mixed-dim tilings, 8 per set.
+				\item Versus use-case: \SI{330}{\micro\second} on commodity i7 (\SI{4.2}{\giga\hertz}).
 			\end{itemize}
@@
 \begin{frame}[standout]
+	Takeaways:
 	\begin{itemize}
-		\item Online in-NIC RL is possible!
+		\item \alert{Online in-NIC RL is possible!}
 		\item In-switch? Less so.
-		\item Arch-specific, but similar design for similar hardware class.
-		\item Work-in-progress: end-to-end timing, training accuracy, other use-cases, optimisation.
+		\item Platform-specific, but similar design for SmartNIC hardware class.
+		\item \alert{Work-in-progress}: end-to-end timing, training accuracy, other use-cases (AQM?), optimisation.
 	\end{itemize}

### 2020-12-11 code/control/reward.sh
Reward packet gen.

@@ -1,0 +1,1 @@
+cd target/release && sudo ./control -i vf0_0 reward --src-ip 192.168.0.4 1.0 8  --dst-ip 192.168.1.5 --dst-port 10602 --src-port 9001 --value-location 0

### 2020-12-11 code/control/src/config.rs

@@ -73,6 +73,20 @@ impl<'a> SendStateConfig<'a> {
 	}
 }
 
+pub struct SendRewardConfig<'a> {
+	pub global: &'a mut GlobalConfig,
+	pub transport: TransportConfig,
+}
+
+impl<'a> SendRewardConfig<'a> {
+	pub fn new(global: &'a mut GlobalConfig) -> Self {
+		Self {
+			global,
+			transport: Default::default(),
+		}
+	}
+}
+
 pub struct TilingsConfig<'a> {
 	pub global: &'a mut GlobalConfig,
 	pub tiling: TilingSet,

### 2020-12-11 code/control/src/lib.rs

@@ -162,3 +162,18 @@ pub fn generate_state(cfg: &FakeStateGeneratorConfig) {
 
 	println!("{}", serde_json::to_string_pretty(&out).unwrap());
 }
+
+pub fn send_reward(cfg: &mut SendRewardConfig, value_loc: i32, reward: f32, shift: usize) {
+	let mut buffer = [0u8; MAX_PKT_SIZE];
+
+	let quant_reward = (reward * ((1 << shift) as f32)) as i32;
+
+	let sz = build_reward_packet(cfg, &mut buffer[..], quant_reward, value_loc)
+		.expect("Packet building failed...");
+
+	if let Channel::Ethernet(ref mut tx, _rx) = &mut cfg.global.channel {
+		tx.send_to(&buffer[..sz], None);
+	} else {
+		eprintln!("Failed to send packet: no channel found");
+	}
+}

### 2020-12-11 code/control/src/main.rs

@@ -173,6 +173,50 @@ fn main() {
 				.help("Input JSON file used to build setup packet.")
 				.takes_value(true)
 				.required(true)))
+		.subcommand(SubCommand::with_name("reward")
+			.about("Sends reward packet.")
+			.arg(Arg::with_name("src-ip")
+				.short("s")
+				.long("src-ip")
+				.value_name("IPV4")
+				.help("Source IP for control packets. Defaults to selected interface.")
+				.takes_value(true))
+			.arg(Arg::with_name("dst-ip")
+				.short("d")
+				.long("dst-ip")
+				.value_name("IPV4")
+				.help("Destination IP for control packets. Defaults to selected interface.")
+				.takes_value(true)
+				.default_value("192.168.1.141"))
+			.arg(Arg::with_name("src-port")
+				.short("p")
+				.long("src-port")
+				.value_name("U16")
+				.help("Source UDP port for control packets.")
+				.takes_value(true)
+				.default_value("16767"))
+			.arg(Arg::with_name("dst-port")
+				.short("P")
+				.long("dst-port")
+				.value_name("U16")
+				.help("Destination port for control packets.")
+				.takes_value(true)
+				.default_value("16768"))
+			.arg(Arg::with_name("value-location")
+				.short("l")
+				.long("value-location")
+				.value_name("I32")
+				.help("Destination port for control packets.")
+				.takes_value(true)
+				.default_value("0"))
+			.arg(Arg::with_name("REWARD")
+				.help("Floating point reward value.")
+				.takes_value(true)
+				.required(true))
+			.arg(Arg::with_name("SHIFT")
+				.help("Quantiser value (i.e., bitshift).")
+				.takes_value(true)
+				.required(true)))
 		.get_matches();
 
 	match matches.subcommand() {
@@ -414,6 +458,63 @@ fn main() {
 				.expect("Invalid tiling packet config file!");
 
 			control::generate_state(&cfg);
+		},
+		("reward", Some(sub_m)) => {
+			let mut g_cfg = GlobalConfig::new(matches.value_of("interface"));
+			let mut cfg = SendRewardConfig::new(&mut g_cfg);
+
+			if let Some(ip) = sub_m.value_of("src-ip") {
+				cfg.transport
+					.src_addr
+					.set_ip(ip.parse().expect("Invalid source IpV4Addr."));
+			} else {
[…]

### 2020-12-11 code/control/src/packet.rs

@@ -5,6 +5,7 @@ pub enum RlType {
 	Config = 0,
 	Insert,
 	State,
+	Reward,
 }
 }
 
@@ -317,6 +318,31 @@ pub fn build_state_packet(
 		for el in state.iter() {
 			body.write_i32::<BigEndian>(*el)?;
 		}
+
+		let space_end = body.len();
+		cursor += space_start - space_end;
+	}
+
+	finalize(&mut buf[..cursor], &offsets, &tile_cfg.transport);
+
+	Ok(cursor)
+}
+
+pub fn build_reward_packet(
+	tile_cfg: &SendRewardConfig,
+	buf: &mut [u8],
+	quant_reward: i32,
+	value_loc: i32,
+) -> IoResult<usize> {
+	let (mut cursor, offsets) = build_transport(tile_cfg.global, &tile_cfg.transport, buf);
+	cursor += build_type(&mut buf[cursor..], RlType::Reward);
+
+	{
+		let mut body = &mut buf[cursor..];
+		let space_start = body.len();
+
+		body.write_i32::<BigEndian>(quant_reward)?;
+		body.write_i32::<BigEndian>(value_loc)?;
 
 		let space_end = body.len();
 		cursor += space_start - space_end;

### 2021-02-05 paper/paper.tex
Reward packet gen.
Preliminary fixes to signalling tests, paper expansion

--- 2020-10-20 paper/paper.tex
@@
 
-\xpatchbibmacro{name:andothers}{%
-	\bibstring{andothers}%
-}{%
-	\bibstring[\emph]{andothers}%
-}{}{}
-
 \usepackage{url}
@@
 \crefname{table}{table}{tables}
+
+\DefineBibliographyStrings{english}{%
+	andothers = {\emph{et al}\adddot}
+}
+\DeclareFieldFormat[inproceedings]{url}{}
+\DeclareFieldFormat[article]{url}{}
+%\DeclareFieldFormat[inproceedings]{doi}{}
+%\DeclareFieldFormat[article]{doi}{}
+%\DeclareFieldFormat[inproceedings]{editor}{}
+%\DeclareFieldFormat[proceedings]{editor}{}
+%\DeclareFieldFormat[article]{editor}{}
+
+\newcommand{\fakepara}[1]{\noindent\textbf{#1:}}
 
@@
 % DOI
-\acmDOI{}
+\acmDOI{12345}
 
 % ISBN
-\acmISBN{}
+\acmISBN{67890}
 
 %Conference
-%\acmConference[Submitted for review to SIGCOMM]{}
-%\acmYear{2018}
+\acmConference[SOSR '21]{}
+%\acmYear{2021}
 %\copyrightyear{}
@@
 	   \city{Glasgow} 
-	   \state{Scotland}
+	   \country{Scotland}
 	 }
@@
 	\city{Glasgow} 
-	\state{Scotland}
+	\country{Scotland}
 }
@@
 
-\section{Motivation}
+This paper contributes:
+\begin{itemize}
+	\item A thing (?? REFERENCE SECTION). Here's how that thing is utterly fantastic and will change the world, or not.
+\end{itemize}
+
+\section{Motivation}\label{sec:motivation}
+\fakepara{Reinforcement learning in asynchronous environments}
+RL techniques allow an \emph{agent} to learn how to iteratively interact with a system in an optimal fashion.
+However, there is some degree of divergence between their theory and implementation.
+Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
+The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
+Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
+
+These delays (and variance thereof) add noise to the state-action mapping being learned, which has a potent reduction to learning rate and final accuracy, even for simple grid world tasks according to \textcite{DBLP:journals/firai/TravnikMSP18}.
+They in turn show that reordering algorithmic steps can reduce these costs for online 1-step algorithms, but that reducing this further requires detailed agent-environment co-design.
+This principle has influenced the design of real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, showing that asynchrony is necessary for high-speed applications.
+Achieving this often requires that state measurements are combined or coalesced~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
+`Stopping the world' in the algorithmic sense causes significant performance degradation, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control, or any time-sensitive control problems.
 
@@
 		\resizebox{\linewidth}{!}{
+			\centering
 			\begin{tikzpicture}
@@
 	\begin{subfigure}{0.45\linewidth}
-		\resizebox{\linewidth}{!}{
+		\resizebox{0.75\linewidth}{!}{
+			\centering
 			\begin{tikzpicture}
@@
 
-What is our goal?
+?? Find some cites citing the relevance of this problem wrt. self-driving cars, robotics, etc.
 
-\begin{itemize}
-	\item Research into asynchronous RL shows that clever reordering and mitigation can reduce the effects of state slippage on learning time and on accuracy of actions.
-However, it cannot completely eliminate the effect of state evolving due to the costs of action computation, execution, and policy updates~\cite{DBLP:journals/firai/TravnikMSP18}.
+The solution we propose is to make use of the recent wave of programmable network devices to \emph{bring reinforcement learning to the data plane}---placing state measurement ($t_1$), low-cost decision-making processes ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
+In networks, actions are most likely to be installed on backbone switches, bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts.
+Ideally, these functions which comprise an RL agent would all be collocated on the same chip or device, but this is easier said than done. 
+Both programmable devices and the network environment make this more difficult, as we'll examine in the sequel.
 
-	\item What if we do stop the world (i.e., block operation) to ensure this doesn't happen?
-In case of congestion control, significant throughput drops (as DNN inference takes \SI{30}{\milli\second}), so asynchrony is necessary~\cite{DBLP:journals/corr/abs-1910-04054}.
+\fakepara{Programmable hardware capabilities}
+The introduction of the P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} language has led to explosive growth in the research community surrounding in-network computation and offloading.
+Providing a single language which is compatible with network devices of many form factors has been instrumental in the development of novel fine-grained traffic measurement approaches~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
+However, this requires that there be a reasonably consistent interface and behaviour between device classes.
+This need for underpins the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}, which defines a conceptual model of match-action tables divided into ingress and egress pipelines.
+The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.
 
-	\item Training deep RL models has its own costs (need many parallel simulations).
-What if we only have one stream of experience to learn from?
-E.g., evolution of the system is too difficult to predict and model.
-In general, DNNs can't be trained effectively from a single, online, real-time stream of experience.
+We would be remiss to believe that the P4 and the PSA define all capabilities that programmable hardware can support.
+Many compatible devices have legacies predating these developments, for instance many-core SOC-based Netronome SmartNIC~\parencite{netronome-smartnic}, and NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14} NICs.
+The first of these compiles from P4$\rightarrow$Micro-C$\rightarrow$bytecode (with Micro-C externs), while the latter can combine the P4$\rightarrow$NetFPGA toolchain~\parencite{DBLP:conf/fpga/IbanezBMZ19} with arbitrary circuit externs---both exposing further device capabilities beyond the specification.
+Even Intel Tofino~\parencite{barefoot-intel} ASICs, which architecturally mirror the PSA, expose additional matching capabilities and ALU functions via the Tofino Native Architecture.
+Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of DL accelerators (which, in turn, lack packet switching functionality).
+Moreover, the limited memory/block RAM and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
 
-	\item PDP hardware is advertised as supporting P4, EBPF...
-But almost all implementations compile to a more complex/expressive language.
-Presence of externs is almost universal when it comes to in-switch compute---i.e., Tofino Native Architecture\footnote{\url{    https://www.barefootnetworks.com/products/brief-p4-studio/}}, Netronome Micro-C\footnote{\url{https://www.netronome.com/products/datapath-programming-tools/}}, P4$\rightarrow$NetFPGA toolchain\footnote{\url{https://github.com/NetFPGA/P4-NetFPGA-public}}.
-\end{itemize}
+Luckily, the task of bringing machine learning models to resource-constrained environments is well-studied.
+Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
+Lower bit-depths reduce memory footprints, and improve throughput in designs such as \emph{bfloat16}~\parencite{bfloat16-blog} in Google TPUs~\parencite{DBLP:journals/sigops/XieDMKVZT18}, \emph{hfp8}~\parencite{DBLP:conf/nips/SunCCWVSCZG19}, and other floating point formats~\parencite{DBLP:journals/corr/abs-2007-01530}.
+In many cases, accuracy losses for doing so are negligible.
+Much of this work goes further still towards integer~\parencite{tensorrt-8bit} or binarised~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16} representations, sacrificing dynamic range for simpler arithmetic operations.
+This has the important side effect of making ML policy inference tractable on hardware without floating-point capabilities, which other works have leveraged to bring neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
 
-What aspects of the state-of-the-art are relevant here?
+To make RL workloads feasible under such constraints, we propose the use of quantised, fixed-point representations such as $Qm.n$, which allow us to evaluate and update policies using only integer arithmetic.
+This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation ($t_2$, \cref{fig:state-slip}).
+Crucially, we train our focus on devices with a low port density, such as SmartNICs and NetFPGA devices.
+The designs of these devices make it reasonable to move policy processing \emph{outside of the packet pipeline}; SmartNICs expose many additional programmable cores, and NetFPGAs allow for the synthesis of independent functional units.
+The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
+When combined with the above P4-driven techniques for in-network flow/device measurement, we at last have the mechanisms to collocate the key processes of an RL agent.
+Moreover, the base P4 dataplane can be used to simplify parsing logic and offer runtime control over which flows/packets are monitored, alongside other packet actions.
+This again fits our goals of integrating RL techniques directly within bump-in-the-wire installations and at end-hosts.
 
-\begin{itemize}
-	\item Hmmm...
-\end{itemize}
+?? relate update/execute operations here? i.e., policy execute needs only ALU add, train needs add+mul+shr. or in RL section? Should that come first?
+
+\fakepara{Online training}
+Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
+This can then make \emph{online} learning of the task an attractive prospect, using a single available stream of experience.
+DNNs, particularly when used as the basis for an RL agent's policy, require vast amounts of experience to converge on an accurate parameter set.
+In many problem domains, this equates to training from compute-years worth of distributed offline simulations, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
+Achieving stable learning requires sizeable batches of gradients to be computed, potentially using the entire experience replay from each simulation.
+Still, the computational cost of gradient computation via backpropagation is significant on embedded hardware.
+
+To then achieve \emph{online} learning, we return our focus to older \emph{classical} reinforcement learning methods.
+In particular, we suggest tile-coding~\cite[pp. \numrange{217}{221}]{RL2E}, with one-step temporal-difference learning algorithms such as Sarsa or Q-learning~\cite[pp. \numrange{129}{132}]{RL2E}.
+These choices have important benefits for in-NIC execution.
+Simpler function approximators do not require batches of inputs to learn in a more stable way, negating the additional memory required to store experience replays.
+Tile-coding in particular admits many optimisations, being an embarrassingly parallel problem.
+Furthermore, gradient calculation is trivial for linear methods like tile coding, particularly when compared to the computational cost of backpropagation required by neural networks.
+Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair.
+
+?? Probably want some cites on the need for batching in NN methods, even though this is understood. DL book?
+
+\fakepara{To-be-organised}
+Interesting thing I realised from \textcite{DBLP:conf/sigcomm/NeugebauerAZAL018}: GPU and NIC might cause IOMMU contention?
+Making line-rate offload of DDN to GPU impractical.
+Unsure if I want to twin this with later measurements/plots.
+
+\section{Background}
+
+\subsection{Classical Reinforcement Learning}
+?? Move into background parent section?
+
+?? RL~\parencite{RL2E}
+
+?? Use this to explain differences between the modern state of the art and older works.
+
+?? Probably give a rough introduction to the algorithm, show how computation needs only addition---particularly when using Qm.n fixed point.
+
+\subsection{Hmm..}
+Should motivation be in here? Unsure what other preliminaries need to be explained.
+
+\section{Design and Implementation}\label{sec:design}
+Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system.
+\Cref{fig:netro-arch} outlines our design and implementation on Netronome SmartNIC hardware in pursuit of this goal.
+At a high level, ...
+
+\subsection{Policy Storage}
+?? Blocks of dense quantised tiles: each an action preference list.
+
+?? Tiling data is split into regions of memory based on access time: higher-dimension tilings are sent to further NUMA regions (EMEM/IMEM).
+
+\subsection{Action Computation}
+?? Split across MEs.
+
+?? each ME can then subdivide by factor of 4 among contexts.
+
+\subsection{Communication}
+?? Free-list of packet handles shared between P4 pipeline threads for state/config/action handover.
+
+?? multi-producer-multi-consumer -- rings used to signal cores, place work, etc., combined with above free-list
+
+?? used for both directions of communication.
+
+\subsection{Notes from PGR}
+
+The main gist of this design is that RL operations \emph{do not happen on the core data path}, which ensures that action computation, installation, and updates don't stall packet transmission or affect throughput.
+In the default deployment of a P4 packet processing pipeline, several of these cores go unused, making this paradigm possible.
+These dedicated cores (\emph{micro-engines} grouped by \emph{island} in NFP terminology) are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
+The main interaction model is that platform-specific IPC mechanisms (message rings) are used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
+These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
+Both input and output can occur on any other core of the device, i.e., as part of P4 \texttt{extern} plugins, while the P4 control plane itself provides granular control over which flows are monitored or affected.
+An input state vector \emph{always} induces an action, and if desired updates the policy using either an included reward or one retrieved from memory according to a (configurable) key in the state.
+Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
+
+?? Expand: this is based on PGR Year 3 atm.
+
+\subsection{Targetting other device classes}
+While this design caters primarily to Netronome's SmartNIC devices, many aspects of this design have clear analogues in NICs of similar form-factor, such as their close NetFPGA SUME cousins.
+For instance, the availability of other cores can be handled by designing additional off-path functional units, much like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+Rather than general-purpose cross-core communication, an RL functional unit would be hard-wired to and from the custom actions that interface with it, using the P4$\rightarrow$NetFPGA toolchain to offer the base framework and granular control over packet and flow selection.
+
+?? Fill out above with more specific instances.
+
+Bringing this to high-port density devices such as Intel Tofino-powered switches is more difficult.
+As discussed in \cref{sec:motivation}, Tofino ASICs map closely to the P4 PSA, meaning that there are no spare asynchronous general-purpose compute units to place this work on.
+A potential solution is to divide RL actions across several received packets (i.e., iteratively computing a portion of the action preference list each time) until any further work would delay outbound transmission.
+This, however, would introduce new issues surrounding concurrent accesses, work splitting, and altered timescales for learning: we leave their treatment and examination to future work.
 
@@
 	\resizebox{1.2\linewidth}{!}{\begin{tikzpicture}
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2.9cm](p4-master){};
-		\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Master Island};
-		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (p4-rules) at (p4-master.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};
+			\node (sw2) {};
 
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2cm, below = 0.1 of p4-master](sw1){};
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
-		\node[below right, inner sep=2pt] at (switch.north west) {\small P4 Island};
-		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};
+			\node[fill=white!80!uofgcobalt, draw=black, minimum height=5.5cm, minimum width=2.9cm, below right= 0.1 of sw2.north west](p4-master){};
+			\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Datapath};
+			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.8] (p4-rules) at (p4-master.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};
+			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n1) at (p4-rules.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Extern-based Action\end{varwidth}};
+			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n2) at (n1.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
+			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (oswlc) at (n2.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};
+			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n3) at (oswlc.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
+			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n4) at (n3.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Network Telemetry\end{varwidth}};
 
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2.9cm, below= 0.1 of switch](others){};
-		\node[below right, inner sep=2pt] at (others.north west) {\small Other Islands};
-		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-etc) at (others.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering \textellipsis \end{varwidth}};
-		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-int) at (others-etc.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering In-band Network Telemetry\end{varwidth}};
+			\node (pkt-text) at ($(p4-master.north) + (0.0,0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet In\end{varwidth}};
 
-		\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (3, -2)$){};
+			\node (pkt-text-out) at ($(p4-master.south) + (0.0,-0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet Out\end{varwidth}};
+
+			\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (5, -1)$){};
 			\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};
@@
 			\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
-		\node[below right, inner sep=2pt] at (wq-out.north west) {\small CLS};
-		\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Work Queue\end{varwidth}};
+			\node[below right, inner sep=2pt] at (wq-out.north west) {\small EMEM};
+			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Input Queue\end{varwidth}};
 
 			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (loop) at ($(wq-out.east) + (1.5,-0.5)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Control ME\end{varwidth}};
+
+			\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-act) at ($(loop.north) + (0.0,1.5)$){};
+			\node[below right, inner sep=2pt] at (wq-act.north west) {\small EMEM};
+			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-actin) at (wq-act.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Action Queue\end{varwidth}};
 
@@
 
-		\draw[->] (loop.north) [out=90,in=45,looseness=0.75] to node[pos=0.5, rotate=0, anchor=center, above] {Rules, Tables}(p4-rules.north east);
-
-		\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2]  node[pos=0.5, rotate=90, anchor=center, above] {Rules}(sw1);
-		\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (sw2);
-		\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (switch);
-
-		\draw[->] (oswlc.east)  to node[midway, above, sloped] {\begin{varwidth}{2.5cm}\tiny \centering Control, Requests,\\State\end{varwidth}} (wq-in.west);
-		\draw[->] (others-int.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
+			\draw[->] (oswlc.east)  to node[midway, above, sloped] {\tiny Control, Requests} (wq-in.west);
+			\draw[->] (n4.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
 			\draw[<->, dotted] (wq-in.east)  to (loop.west);
 			\draw[->, dotted] (loop.south) to (policy-cls-out);
+
+			\draw[->] (loop.north)  to node[midway, above, sloped] {\tiny Actions} (wq-actin.south);
+
+			\draw[->] (wq-actin.west)  to node[midway, above, sloped] {\tiny Actions} (n1.east);
 
@@
 
-		\draw[->] (loop) to node[midway, above, sloped] {\tiny NNR} (po-me-1);
-		\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NNR} (po-me-2);
-		\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NNR} (po-me-3);
-		\draw[->] (po-me-3) to [out=-45,in=0,looseness=2.15] node[midway, above, sloped] {\tiny NNR} (loop);
+			\draw[->] (pkt-text)  to  (p4-master.north);
+			\draw[->] (p4-master.south)  to  (pkt-text-out);
+
+			\draw[->] (loop) to node[midway, above, sloped] {\tiny NN} (po-me-1);
+			\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NN} (po-me-2);
+			\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NN} (po-me-3);
+			\draw[->] (po-me-3) to [out=-45,in=0,looseness=2] node[midway, above, sloped] {\tiny NN} (loop);
 	\end{tikzpicture}}
@@
 \begin{enumerate}
-	\item Run the existing work~\cite{DBLP:journals/tnsm/SimpsonRP20}, acquire a policy for one node after full info sharing.
+	\item Run the existing work~\parencite{DBLP:journals/tnsm/SimpsonRP20}, acquire a policy for one node after full info sharing.
 	\item Load this policy onto the RL Island (\cref{fig:netro-arch}).
@@
 \end{enumerate}
-
-\newpage
 
@@
 	\item Tile-coding of quantised states on-chip, and state packet receipt.
+	\item Action computation on-chip.
+	\item Rudimentary, single-core training.
 \end{itemize}
@@
 	\item Policy quantisation divergence \& accuracy (via mininet).
+	\item Measure DPDK SmartNIC RTT as a lower bound on controller access time.
+	\begin{itemize}
+		\item Same mechanism applies to other manufacturers (Barefoot), so generalisable.
+		\item My results are non-DPDK, but there is a good study~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018} on exactly this.
+	\end{itemize}
+	\item Measure basic on-switch times: ring-ring communication, tile-coding cost, action compute cost, learning cost.
+\end{itemize}
+
+Known/understood (but not implemented)?
+\begin{itemize}
+	\item Figure out PIF rule installation.
+	\begin{itemize}
+		\item Not doable through P4 na\"{i}vely. Handing back rules to controller is expensive: need another action to deliberately use its own hash table etc., then batch back as faster P4 rules.
+	\end{itemize}
 \end{itemize}
@@
 \begin{itemize}
-	\item Action computation on-chip.
-	\begin{itemize}
-		\item \emph{\num{1} week.}
-	\end{itemize}
-	\item Figure out PIF rule installation.
-	\begin{itemize}
-		\item \emph{\num{1} week.}
-		\item Rationale: understand and describe mechanism, do NOT implement. We know that we can save 1 controller RTT.
-	\end{itemize}
 	\item Setup other core as event source.
@@
 	\end{itemize}
-	\item Implement rudimentary training.
+	\item Fixes to hash table use, signalling.
+	\begin{itemize}
+		\item \emph{\num{1} week.}
+		\item Blocks progress on optimisation, correct multi-stream RL.
+	\end{itemize}
+	\item Optimise/parallelise action compute/learning.
 	\begin{itemize}
 		\item \emph{\num{2} week.}
+		\item Value? Proves this approach can scale according to resource availability/demand on the deployment device, at compile-time anyhow.
 	\end{itemize}
@@
 \begin{itemize}
-	\item Measure DPDK SmartNIC RTT as a lower bound on controller access time.
-	\begin{itemize}
-		\item \emph{\num{0.5} weeks.}
-		\item Same mechanism applies to other manufacturers (Barefoot), so generalisable.
-	\end{itemize}
-	\item Measure on-switch times: ring-ring communication, tile-coding cost, action compute cost.
+	\item Measure sum of times $t_{1...3}$ described elsewhere.
 	\begin{itemize}
 		\item \emph{\num{1} week.}
+		\item Currently have individual, need ``end-to-end''.
 	\end{itemize}
-	\item Measure times $t_{1...3}$ described elsewhere, or closest analogues in impl'd system.
+	\item Throughput impact testing.
 	\begin{itemize}
 		\item \emph{\num{1} week.}
+		\item Requires current impl, creation of basic testbed and packet forwarding setup.
+		\item Use-case independent.
+	\end{itemize}
+	\item Measure execution, training costs for varying policy sizes.
+	\begin{itemize}
+		\item \emph{\num{1.5} weeks.}
+		\item Policies need not be meaningful, and can be garbage data. Use-case independent.
+		\item Partially blocked on multicore accelerated/optimised execution.
+	\end{itemize}
+	\item Measure rule installation cost on different platforms.
+	\begin{itemize}
+		\item \emph{\num{1.5} weeks.}
+		\item Blocked on creation of basic testbed and packet forwarding setup.
+	\end{itemize}
+	\item Investigate stationarity.
+	\begin{itemize}
+		\item \emph{\numrange{2}{3} weeks.}
+		\item Requires full use-case implementation.
+		\item Requires careful thought about sensible changes to normality.
+		\item Might be bolstered by comparison against an offline-trained system?
 	\end{itemize}
@@
 
+\section{Everything we could possibly evaluate}
+
+Evaluation will need to be divided into 2 main categories:
+\begin{enumerate}
+	\item Simulation/emulation based on mininet.
+	\begin{itemize}
+		\item These can only assess algorithmic properties, as performance measurements in mininet (i.e., a full mult-agent system) are not representative.
+	\end{itemize}
+	\item Testbed setup (2/3 configurations):
+	\begin{itemize}
+		\item These will assess key system performance properties of a \emph{single agent}, in control of a single switch/NIC as a bump in the wire.
+Bump-in-the-wire device will carry traffic between two hosts.
+		\item \emph{Time permitting, may need to send state vectors in packets rather than do feature/state extraction on-chip}.
+		\item \emph{Necessary.} RL runs on bump-in-the-wire SmartNIC.
+		\item \emph{Either/both.} P4-based off-path RL.
+Custom packet actions (e.g., random drop) will be implemented in P4 with externs.
+Packets/digests mirrored to state extraction machine.
+Agent may be co-hosted with state-extraction.
+Rules inserted via P4-device specific RTE.
+		\item \emph{Either/both.} OvS-based off-path RL.
+Custom packet actions are already implemented in OvS as OpenFlow extensions.
+Packets/digests mirrored to state extraction machine.
+Agent may be co-hosted with state-extraction.
+	\end{itemize}
+\end{enumerate}
+
+\subsection{What is this different from, or better than?}
+List below mainly for the benefit of helping enumerate everything I can compare against, and what their drawbacks are.
+
+\begin{itemize}
+	\item (Deep) neural network-backed RL.
+	\begin{itemize}
+		\item NNs infeasible to update online.
+Ordinarily this requires larger volumes of data than can be gathered in real time, but becomes substantially more difficult after change of representation (quantisation, binarisation).
+An online system (such as this) should be able to adapt to changes in underlying traffic, optimisation criteria, etc., dynamically.
+		\item Complexity of function approximation means that policy updates are more computationally complex to compute.
+For instance, NNs require use of the backpropagation algorithm, rather than the simpler classical schemes which need only the tile set already identified in action selection.
+		\item Only heavily quantised (i.e., binary neural networks) suitable for in-switch execution.
+	\end{itemize}
+
+	\item MARL and other ``agents as network functions'' designs.
+	\begin{itemize}
+		\item Action, state collection, updates etc.
+happen on nodes outside of the main packet path in this paradigm.
+Thus, state measurements, action computation, and so on, should induce the state drift described above.
+Also makes these approaches ``slower to react'' in theory.
+	\end{itemize}
+\end{itemize}
+
+At present, there are no other RL works in PDPs, let alone with online policy updates.
+Other existing RL use cases suitable for adaptation to the ``bump-in-the-wire'' model are proving tricky to find.
+
+\subsection{Emulation metrics}
+Key (MVP) metrics will be \textbf{bolded}, and expanded with a subsubsection.
+
+\textbf{Quantisation accuracy}, quantised training accuracy, \textbf{stationarity}.
+
+\subsubsection{Quantisation accuracy}
+Examination of whether lower resolution quantisation impacts RL agent decision accuracy substantially.
+
+This has been done using MARL in mininet, see \cref{fig:quant-acc}.
+
+\begin{figure}
+	\includegraphics{../plots/build/marl-quant/accuracy-binary}
+	\caption{Normalised accuracy of a converted, pre-trained floating-point tile-coded policy after conversion to $32Qn$ fixed-point.}\label{fig:quant-acc}
+\end{figure}
+
+\subsubsection{Interaction with stationarity}
+Learning time of the system as compared to known changes in normality.
+
+I.e., in the DDoS case: how long does an agent take to respond to the onset of an attack?
+A change in attack strategy?
+The conclusion of an attack?
+There are no citations on \emph{how} DDoS attacks evolve, just that they do, and that this can occur on various timescales.
+
+\subsection{Testbed metrics}
+Key (MVP) metrics will be \textbf{bolded}, and expanded with a subsubsection.
+
+\textbf{Traffic throughput (bytes, packets)}, \textbf{packet processing delay}, \textbf{action execution time}, \textbf{learning time}, from-scratch learning accuracy, \textbf{rule installation delay/cost}, host and NIC memory costs, policy installation time, ``accuracy'' in the DDoS case.
+
+\subsubsection{Traffic throughput}
+Main research question: does moving RL logic onto the NIC reduce maximum throughput of the NIC, even if calculation is on a different functional unit?
+This is relevant in both \si{\byte\per\second} and packets per second.
+
+This is application independent, and would assess whether the RL units (acting independently on the netronome) affect throughput compared with a basic P4 program at different work intensities.
+Ideally, this would show no effect on forwarding performance.
+
+\subsubsection{Packet processing delay}
+Main research question: does the presence of off-path (but on-chip) RL execution meaningfully increase packet processing delay?
+How does this compare to a P4 program with an equivalent number of tables?
+What about to OvS?
+
+\subsubsection{Action execution \& learning time}
+Main research question: does moving RL logic from the host to the NIC substantially reduce delays from state measurement, to state receipt, to action computation and installation?
+These constitute $t_{1...3}$ discussed above.
+Primarily, these need to be looked at using varying policy sizes and distribution across memory regions.
+
+We ideally want these both with and without any acceleration from extra cores on the netronome: the point of this being that we can demonstrate that the system (or a similar NetFPGA system) can scale upwards according to available system resources or demands.
+
+\subsubsection{Rule installation delay/cost.}
+Main research question: how long does it take to install rules in the above testbed setups? What is the downtime of so doing?
+
+This is application-dependent: see earlier discussions on rule batching and local extern/hash-table use.
+
 \section{Related Work}
 
-\begin{itemize}
-	\item A switch/NIC architecture like \textcite{DBLP:conf/hotnets/StephensAS18} would remove the need for hard-coded island-island relationships.
-\end{itemize}
+?? A switch/NIC architecture like \textcite{DBLP:conf/hotnets/StephensAS18} would remove the need for hard-coded island-island relationships.
+
+?? Taurus. They can't train online though!~\parencite{DBLP:journals/corr/abs-2002-08987}
+
+?? Do switches~\parencite{DBLP:conf/hotnets/XiongZ19}.
+?? Train offline, classical methods targetting all match-action P4 systems (no externs etc).
+
+?? N3IC run NNs on NIC~\parencite{DBLP:journals/corr/abs-2009-02353}, allow packet tagging to occur based on pre-trained BNN~\parencite{DBLP:conf/nips/HubaraCSEB16}.
+?? both NFP and NetFPGA.
+?? Or as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
+?? Binary NNs ``more suitable for network hardware''~\parencite{DBLP:journals/corr/MiyashitaLM16}?
+
+?? Can we collect some papers on per-packet ML/RL? Per-packet classification using NNs to build decision trees~\parencite{DBLP:conf/sigcomm/LiangZJS19}
+
+?? RL for congestion control. ~\parencite{DBLP:journals/corr/abs-1910-04054}
+
+?? Arbitrary training enhanced by in-network? General NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18}, RL-specfic~\parencite{DBLP:conf/isca/LiLYCSH19}.
 

### 2021-02-05 code/nfp/ring-comm-timing/ref-forward.c

@@ -9,7 +9,7 @@
 
 #include "shared_defs.h"
 
-__nnr uint32_t in_val;
+__declspec(export, emem) uint32_t ref_times_ct = 0;
 
 // See https://groups.google.com/g/open-nfp/c/gAIBsqolNdY/m/gjbe4VIzBwAJ
 // for reflector usage.
@@ -27,7 +27,7 @@ __declspec(visible) SIGNAL foreign_sig;
 #define READ_REG_TYPE 0
 
 // details of thread who sends to us
-#define PRODUCER_ME 5 // 1 + 4;
+#define PRODUCER_ME 4 // 0 + 4;
 #define PRODUCER_ISLAND 12
 #define PRODUCER_LOC (PRODUCER_ISLAND << 4) + PRODUCER_ME
 
@@ -45,9 +45,11 @@ main() {
 	unsigned int client_sig = __signal_number(&receiver_sig, CONSUMER_LOC);
 	unsigned int register_no = __xfer_reg_number(&receiver_x, CONSUMER_LOC) & 4095;
 
+	__declspec(xfer_write_reg) uint32_t cter;
+
 	client_sig |= CONSUMER_CTX << 4;
 
-	remote_sig = (1 << 13) | (client_sig << 9);
+	remote_sig = (client_sig << 9);
 
 	base_address = CONSUMER_ISLAND << 24;
 	base_address |= READ_REG_TYPE << 16;
@@ -75,5 +77,7 @@ main() {
 			alu[remote_sig, --, B, 3, <<13]
 			ct[reflect_write_sig_remote, me0_x, base_address, 0, 1], indirect_ref
 		}
+
+		mem_write32(&cter, &ref_times_ct, new_i);
 	}
 }

### 2021-02-05 code/nfp/ring-comm-timing/ref-src.c

@@ -18,10 +18,11 @@ __declspec(remote read_reg) uint32_t foreign_x;
 __declspec(remote) SIGNAL foreign_sig;
 
 // details of thread who sends to us
-#define READ_REG_TYPE 0
-#define PRODUCER_ME 5 // 1 + 4;
-#define PRODUCER_ISLAND 12
+#define READ_REG_TYPE 0 // 0 = transfer registers, 1 = csr registers. 	Bit 16
+#define PRODUCER_ME 5 // 1 + 4; FPC + 4 == "Master number"				Bits 13:10
+#define PRODUCER_ISLAND 12 // Island number. 							Bits 29:24
 #define PRODUCER_LOC (PRODUCER_ISLAND << 4) + PRODUCER_ME
+// Then register number is bits 9:2
 
 // Details for thread we send to
 #define CONSUMER_CTX 0
@@ -54,7 +55,7 @@ main() {
 	base_address |= CONSUMER_ME << 10;
 	base_address |= register_no << 2;
 
-	remote_sig = (1 << 13) | (client_sig << 9);
+	remote_sig = (client_sig << 9);
 
 	__implicit_write(&receiver_x);
 	__implicit_write((void*)&receiver_sig);

### 2021-02-05 code/nfp/rl-test/rl.c

@@ -54,7 +54,8 @@ __export __emem __align(SA_TABLE_SZ) struct mem_lkup_cam32_16B_table_bucket_entr
 
 // Need to use this due to size limits on struct in CAMHT def, even
 // if key is different.
-#define SA_ENTRIES 0x20000
+// #define SA_ENTRIES 0x20000
+#define SA_ENTRIES 0x10
 CAMHT_DECLARE(state_action_map, SA_ENTRIES, uint64_t)
 
 __declspec(emem) struct state_action_pair state_action_pairs[SA_ENTRIES];
@@ -627,6 +628,9 @@ void state_packet(__addr40 _declspec(emem) struct rl_config *cfg, __declspec(xfe
 			changed_key = 1;
 		}
 
+
+		nani = (((uint64_t) reward_found) << 32) | state_added;
+		mem_write64(&nani, &really_really_bad, sizeof(uint64_t));
 		if ((!(state_added || changed_key)) && state_found >= 0 && reward_found >= 0) {
 			tile_t matched_reward = reward_map_key_tbl[reward_found];
 			struct state_action_pair lsap = state_action_pairs[state_found];

### 2021-04-16 paper/paper.tex
Paper work.
Expanded most notes on design into full prose.

@@
 %Conference
-\acmConference[SOSR '21]{}
+\acmConference[SOSR '21]{Symposium of SDx Research 2021}{\numrange{13}{14} September}{Virtual Conference}
 %\acmYear{2021}
@@
 \acmPrice{}
+
+\newcommand{\approach}{On Path Learning}
+\newcommand{\approachshort}{OPaL}
 
@@
 
-?? Probably give a rough introduction to the algorithm, show how computation needs only addition---particularly when using Qm.n fixed point.
+?? Probably give a rough introduction to the algorithm, show what arithmetic operations are needed---particularly when using Qm.n fixed point.
+
+?? Explain how it is an embarrassingly parallel problem? Or at least very amenable to partition-aggregate.
+
+?? Critical section around shared writeback loc, but individual work items are disjoint in their input data (particularly useful for updates).
 
@@
 \section{Design and Implementation}\label{sec:design}
-Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system.
+Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{}}.
+At a high level, \approachshort{} is designed to use the auxiliary compute exposed by general SmartNIC devices to offer low-latency online learning, scaling according to available on-chip resources at build time.
+
 \Cref{fig:netro-arch} outlines our design and implementation on Netronome SmartNIC hardware in pursuit of this goal.
-At a high level, ...
-
-\subsection{Policy Storage}
-?? Blocks of dense quantised tiles: each an action preference list.
-
-?? Tiling data is split into regions of memory based on access time: higher-dimension tilings are sent to further NUMA regions (EMEM/IMEM).
-
-\subsection{Action Computation}
-?? Split across MEs.
-
-?? each ME can then subdivide by factor of 4 among contexts.
-
-\subsection{Communication}
-?? Free-list of packet handles shared between P4 pipeline threads for state/config/action handover.
-
-?? multi-producer-multi-consumer -- rings used to signal cores, place work, etc., combined with above free-list
-
-?? used for both directions of communication.
-
-\subsection{Notes from PGR}
-
-The main gist of this design is that RL operations \emph{do not happen on the core data path}, which ensures that action computation, installation, and updates don't stall packet transmission or affect throughput.
-In the default deployment of a P4 packet processing pipeline, several of these cores go unused, making this paradigm possible.
-These dedicated cores (\emph{micro-engines} grouped by \emph{island} in NFP terminology) are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
-The main interaction model is that platform-specific IPC mechanisms (message rings) are used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
-These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
-Both input and output can occur on any other core of the device, i.e., as part of P4 \texttt{extern} plugins, while the P4 control plane itself provides granular control over which flows are monitored or affected.
-An input state vector \emph{always} induces an action, and if desired updates the policy using either an included reward or one retrieved from memory according to a (configurable) key in the state.
-Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
-
-?? Expand: this is based on PGR Year 3 atm.
-
-\subsection{Targetting other device classes}
-While this design caters primarily to Netronome's SmartNIC devices, many aspects of this design have clear analogues in NICs of similar form-factor, such as their close NetFPGA SUME cousins.
-For instance, the availability of other cores can be handled by designing additional off-path functional units, much like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-Rather than general-purpose cross-core communication, an RL functional unit would be hard-wired to and from the custom actions that interface with it, using the P4$\rightarrow$NetFPGA toolchain to offer the base framework and granular control over packet and flow selection.
-
-?? Fill out above with more specific instances.
-
-Bringing this to high-port density devices such as Intel Tofino-powered switches is more difficult.
-As discussed in \cref{sec:motivation}, Tofino ASICs map closely to the P4 PSA, meaning that there are no spare asynchronous general-purpose compute units to place this work on.
-A potential solution is to divide RL actions across several received packets (i.e., iteratively computing a portion of the action preference list each time) until any further work would delay outbound transmission.
-This, however, would introduce new issues surrounding concurrent accesses, work splitting, and altered timescales for learning: we leave their treatment and examination to future work.
-
-\section{Techniques}
-
-What does the hardware give us, and what can we assume are present on many other classes of device?
-\begin{itemize}
-	\item 96-core (\emph{management engine}, ME) (? check), grouped into \emph{islands} based on local capabilities. Fast transfer between MEs on same island (\emph{nearest neighbour registers}, NNRs)
-
-	\begin{itemize}
-		\item Multicore model common.
-		\item Arch-specific: one program per ME. Enables cooperative scheduling between \numrange{4}{8} threads, zero-cost context switch due to register file (below).
-	\end{itemize}
-
-	\item Explicit memory hierarchy:
-
-	\begin{itemize}
-		\item Large register file split disjointly between threads. (likely common)
-		\item Transparent cache.
-		\item In order of speed? Reg file, External Mem
-	\end{itemize}
-
-	\item One program to a core---enables cooperative scheduling with zero-cost context switches.
-
-\end{itemize}
-Now how do we make this work?
-\begin{itemize}
-	\item No floating point math. Solution? Quantisation (cite some other papers who've used this to decent effect).
-
-	\item Bump-in-the-wire design for SmartNIC.
-
-	\item P4 for easy packet parsing, control packet extraction, passthrough and manipulation of other traffic.
-
-	\item RL Island keeps freelist, P4 PIF Islands copy packets over (could in theory hack P4 firmware micro-C libraries to stop freeing resources and just pass those?).
-
-	\item Control packets marked by custom \emph{experimenter} DSCP (as in \textcite{DBLP:conf/isca/LiLYCSH19}). Infrastructure can ensure that only trusted internal nodes can produce such packets.
-
-	\item Control allows adaptation to any policy design assuming standard tile-coding, for any target application.
-Policies may be installed via block copy, or trained online.
-(So long as policy meets compiled maximum dimensions, constraints.)
-
-	\item (not impl'd yet) How does state work? Packet passed in over work queue to control ME. source can be either the network (same mechanism as control packets)
-
-	\item (not impl'd yet) How do actions work? 2 options.
-	\begin{enumerate}
-		\item Produce packet on wire according to config (custom format).
-		\item Pass packet to local island/ME (requires both config and custom build).
-	\end{enumerate}
-
-	\item Built with \Textcite{DBLP:journals/tnsm/SimpsonRP20} in mind, but applicable to any RL application.
-
-	\item New PDPs have encryption primitives: scope to include authentication etc. around control packets.
-
-	\item (Not impl'd yet) Event loop can easily run on Control ME: timing APIs exposed to generate periodic events in lieu of something like \textcite{DBLP:conf/hotnets/IbanezABM19}.
-\end{itemize}
-
-\begin{figure}
-	\centering
-	\resizebox{\linewidth}{!}{\begin{tikzpicture}
-		\node(remote){Remote};
-
-		%%%
-
-		\node[below=0.05 of remote](swpos){};
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
-		\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
-		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
-
-		%
-
-		\node[right=2 of swpos](epos){};
-		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
-		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
-		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
-		\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.85,0)$) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\small \centering Estimator\\$g(\cdot)$\end{varwidth}};
-
-		%
-
-		\node[right=3.5 of epos](apos){};
-		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
-		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
-		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
-		\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent vNF};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
-
-		%
-
-		\node[below=2.3 of remote.west](linestart){};
-		\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
-		\draw [dashed] (linestart) -- (lineend);
-
-		%%%
-
-		\node[below=2.4 of remote](local){Local};
-
-		%%%
-
-		\node[below=0.25 of local](aswpos){};
-		\node[fill=white!80!uofgthistle, draw=black, minimum height=3cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
-		\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\small \centering OpenFlow\\Tables\end{varwidth}};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\small \centering Stats\\Collector\end{varwidth}};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
-
-		%
-
-		\node (avfpos) at ($(aswpos) + (3,0.2)$){};
-		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
-		\node[below right, inner sep=2pt] at (avf.north west) {\small Agent vNF};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\small \centering Flowstate\\Database\end{varwidth}};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\small \centering TRS\\Scheduler\end{varwidth}};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};
-
-		%%%
-
-		\tikzset{>=stealth}
-
-		\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
-		\draw[thick, ->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};
-
-		\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
-
-		\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
-		\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};
-
-		\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};
-
-		\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
-		\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};
-
-		\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
-		\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
-		\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
-		\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
-		\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};
-
-		\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
-		\node[right=0.3] at (avfsa.east) {\tiny Experience};
-		\end{tikzpicture}}
-	\caption{
-		System architecture  for our RL-driven DDoS defence system.
-		\label{fig:sys-arch}
-	}
-\end{figure}
+We describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.
 
@@
 	\end{tikzpicture}}
-	\caption{Architecture for generalised RL agent on Netronome hardware.\label{fig:netro-arch}}
+	\caption{Architecture for an \approachshort{} agent on Netronome hardware, using 4 cores for parallel policy processing. (FIXME: redo in illustrator? Show CTXes, PISA)\label{fig:netro-arch}}
 \end{figure}
+
+\subsection{System Model}
+\approachshort{} is a general, task-independent framework for in-network, online training and execution of \emph{any reinforcement learning agent design} using classical methods.
+The meaning 
+However, in-NIC/in-network execution specifically benefits packet-, flow-, and network-level learning, control, and optimisation tasks.
+\approachshort{} runs on one more cores of a SmartNIC to convert state measurements from the environment into a stream of actions using a stored policy.
+These dedicated cores (\emph{micro-engines} grouped by \emph{island} in NFP terminology) are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
+Combined with reward measurements, this policy can then be updated or trained from scratch entirely on the NIC, acting as a fully online RL agent.
+An input state vector \emph{always} induces an action, and if desired updates the policy using either an included reward or one retrieved from memory according to a (configurable) key placed alongside the state.
+
+%However, high-speed data networks impose inviolable per-packet deadlines.
+
+To protect traffic throughput and allow for effective deployment in as many environments as possible, \approachshort{} places RL execution off the main packet path, communicating and running parallel to the main P4 dataplane.
+This asynchrony allows coexistence with P4 programs, and imposes minimal/no impact on carried traffic for both bump-in-the-wire deployments and at end-points.
+For instance, in the default deployment of a P4 packet processing pipeline on Netronome NFP chips several of these cores go unused (as does spare area on an FPGA design), making this paradigm possible.
+
+The main interaction model is that platform-specific IPC mechanisms (message rings) are used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
+These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
+Both input and output can occur on any other core of the device, i.e., as part of P4 \texttt{extern} plugins or a dedicated flow state measurement subsystem, while the P4 control plane itself provides granular control over which flows are monitored or affected.
+
+Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
+
+?? Dynamic, runtime configuration via control and data planes.
+
+?? Compile-time config of bit-depth, 
+?? Allows learning from several live traces simultaneously.
+?? Allows execution of indepdnnt traces simultaneously.
+
+\subsection{Action Computation}
+Note: talk about single core AND parallel.
+
+\begin{itemize}
+	\item A state vector is tile coded, action probabilities are produced, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
+	\item \emph{Open question: how to configure action destination?
+This should either be another core (after an optional transformation), a packet sent out to a destination, maybe talk back to the controller.}
+\end{itemize}
+\begin{itemize}
+	\item \emph{Open question: how to safely \& reasonably map state back to incoming rewards?}
+	\item \emph{Open question: Can/should this also support on-chip reward signals?}
+\end{itemize}
+
+?? Split across MEs.
+
+?? each ME can then subdivide by factor of 8 among contexts.
+
+?? Scales from 1--(n-1) cores.
+
+\subsection{Agent-Environment Communication}\label{sec:agent-environment-communication}
+\approachshort{} uses \emph{multiple-producer/multiple-consumer} (MPMC) messaging channels to communicate with other computational elements; be they P4 programs on the packet path, or additional on-chip analysis and control modules.
+Through these channels the system \emph{pushes} state vectors, reward measures, and control/setup packets as its inputs, and \emph{pulls} a stream of state-action pairs as its outputs.
+This allows agent decisions to be made asynchronously---preventing packet stalling---and allowing for many parallel decision-making agents to be used if desired.
+The key insight of this mechanism is that on-chip reward/state signals enjoy first-class support in much the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-NIC/chip asynchronous processes or the controller.
+
+Our implementation uses platform-specific IPC mechanisms (\emph{EMEM ring buffers}) with hardware signalling mechanisms on work arrival to achieve this.
+Due to the lack of memory allocation in these environments, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state/configuration/policy data handover.
+Packet headers are extracted and parsed using the tooling and data types autogenerated by the P4 pipeline.
+We found that this costs a median \SIrange{126}{140}{\nano\second} communication time, comparable to message channels in the Rust and Go languages on commodity hardware.
+
+\subsection{Intra-Agent Communication}
+Even with embarrassingly parallel problems, optimising for latency requires meticulous care in how work is passed out and aggregated.
+This is truer still when moving from the moderately fine-grained, efficient control of classical RL methods ($\sim$\SI{2}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
+Ordinarily, the serialisation/marshalling of requests, responses, and safe shared data access can incur significant extra overheads.
+On-chip execution and the nature of action preference computation allow us to use lockless atomic aggregation, removing the overheads of explicit messaging/packetisation.
+Moreover, physically adjacent functional units may have special-purpose shared registers or share a small fast cache to accelerate communication further.
+
+Our implementation exploits the locality of threads, cores, and their parent islands in the NFP architecture.
+Groups of \numrange{4}{12} cores (MEs) in this platform are termed \emph{islands}, with shared access to low-latency scratch memory and the adjacent registers of neighbouring MEs.
+Each ME then has a number of threads fixed at compile time (\numrange{4}{8}), sharing one large register file to enable zero-cost context switches.
+Policy compute/update tasks and configuration updates are passed between cores using these direct \emph{next neighbour} registers, signalling all child threads in response.
+Each task performs atomic adds to a shared preference list, and atomically increments an acknowledgement counter to be periodically checked by the master thread.
+
+In earlier designs we had experimented with bounded buffers as in \cref{sec:agent-environment-communication} for this, modified to be located solely in local scratch memory, dedicating the master thread to result aggregation.
+We found that this created a performance bottleneck at this final stage, causing significant head-of-line blocking for each of the workers.
+Similarly, our next neighbour work notification scheme ($\sim$\SI{20}{\nano\second} latency per relayed message) was examined against \emph{reflector} and \emph{work queue} IPC mechanisms (\SIlist{58;126}{\nano\second} per messaged core).
+These slower messages can be used in theory to skip ahead into longer ME chains.
+
+\subsection{Reconfiguration}
+?? Compile-time config of bit-depth, what else?
+
+\begin{itemize}
+	\item (Re)-configurable at runtime.
+	\item Custom DSCP marker acts as indicator of RL packets (i.e., bother with parsing past UDP). This trick is copied from \textcite{DBLP:conf/isca/LiLYCSH19}.
+	\item P4 parser extracts headers, parser state and payload passed over ring to another core dedicated to RL operations.
+	\item Policy dimensions, size, action count, tiling strategies, learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) can all be chosen and changed live.
+\end{itemize}
+
+\subsection{Work Allocation}
+?? WHy
+
+?? How
+
+\subsection{Policy Storage}
+?? Why
+
+?? Trained online or installed by block copy from control/data planes.
+
+?? Blocks of dense quantised tiles: each an action preference list.
+
+?? Tiling data is split into regions of memory based on access time: higher-dimension tilings are sent to further NUMA regions (EMEM/IMEM).
+
+\subsection{Variable Quantisation Bitdepth}
+Compile time eh
+
+?? Mention that BB SIMD failed here.
+
+??
+
+?? Lower bitdepth can mean more actions, more tilings, more dimensions in tilings, higher perf.
+
+\subsection{Notes from chat with others}
+This approach can be used to generate transferable policies/samples for other models.
+
+Potential: low-latency train for some specific points, and high throughput for others? Think on the exact note I took...
+
+``can have accel'd offline, train online subset w/ my approach''
+
+THis might mean "do 32-bit online", then downsample to 8-bit policy for high-throughput mode.
+
+Can changes in trained policy be used to transfer to a more complex function approximator elsewhere?
+
+\subsection{Targeting other device classes}
+While this design caters primarily to Netronome's SmartNIC devices, many aspects of this design have clear analogues in NICs of similar form-factor, such as their close NetFPGA SUME cousins.
+For instance, the availability of other cores can be handled by designing additional off-path functional units, much like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+Rather than general-purpose cross-core communication, an RL functional unit would be hard-wired to and from the custom actions that interface with it, using the P4$\rightarrow$NetFPGA toolchain to offer the base framework and granular control over packet and flow selection.
+Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
+Consider a single, shared, read-only register between RL functional ...
+?? think about really really carefully lest Reviewer 2 murder you
+?? Parallel, atomic writes at the required resolution.
+?? No bit stuffing required.
+
+Bringing this to high-port density devices such as Intel Tofino-powered switches is more difficult.
+As discussed in \cref{sec:motivation}, Tofino ASICs map closely to the P4 PSA, meaning that there are no spare asynchronous general-purpose compute units to place this work on.
+A potential solution is to divide RL actions across several received packets (i.e., iteratively computing a portion of the action preference list each time) until any further work would delay outbound transmission.
+This, however, would introduce new issues surrounding concurrent accesses, work splitting, and altered timescales for learning: we leave their treatment and examination to future work.
+
+%\begin{figure}
+%	\centering
+%	\resizebox{\linewidth}{!}{\begin{tikzpicture}
+%		\node(remote){Remote};
+%		
+%		%%%
+%		
+%		\node[below=0.05 of remote](swpos){};
+%		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
+%		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
+%		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
+%		\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
+%		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
+%		
+%		%
+%		
+%		\node[right=2 of swpos](epos){};
+%		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
+%		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
+%		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
+%		\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
+%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.85,0)$) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
+%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\small \centering Estimator\\$g(\cdot)$\end{varwidth}};
+%		
+%		%
+%		
+%		\node[right=3.5 of epos](apos){};
+%		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
+%		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
+%		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
+%		\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent vNF};
+%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
+%		
+%		%
+%		
+%		\node[below=2.3 of remote.west](linestart){};
+%		\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
+%		\draw [dashed] (linestart) -- (lineend);
+%		
+%		%%%
+%		
+%		\node[below=2.4 of remote](local){Local};
+%		
+%		%%%
+%		
+%		\node[below=0.25 of local](aswpos){};
+%		\node[fill=white!80!uofgthistle, draw=black, minimum height=3cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
+%		\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
+%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\small \centering OpenFlow\\Tables\end{varwidth}};
+%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\small \centering Stats\\Collector\end{varwidth}};
+%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
+%		
+%		%
+%		
+%		\node (avfpos) at ($(aswpos) + (3,0.2)$){};
+%		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
+%		\node[below right, inner sep=2pt] at (avf.north west) {\small Agent vNF};
+%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
+%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\small \centering Flowstate\\Database\end{varwidth}};
+%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\small \centering TRS\\Scheduler\end{varwidth}};
+%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
+%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};
+%		
+%		%%%
+%		
+%		\tikzset{>=stealth}
+%		
+%		\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
+%		\draw[thick, ->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};
+%		
+%		\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
+%		
+%		\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
+%		\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};
+%		
+%		\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};
+%		
+%		\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
+%		\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};
+%		
+%		\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
+%		\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
+%		\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
+%		\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
+%		\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};
+%		
+%		\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
+%		\node[right=0.3] at (avfsa.east) {\tiny Experience};
+%		\end{tikzpicture}}
+%	\caption{
+%		System architecture  for our RL-driven DDoS defence system.
+%		\label{fig:sys-arch}
+%	}
+%\end{figure}
 
@@
 \end{enumerate}
-
-\section{Bullets on (intended) design.}
-See also: \cref{fig:netro-arch}.
-\begin{itemize}
-	\item High-level overview.
-	\begin{itemize}
-		\item RL operations \emph{do not happen on the core path}.
-		\item Another island (or set of cores) is dedicated to handling inputs, configuration, computation.
-		\item Cross-island communication happens over \emph{rings} (Netronome-specific).
-		\begin{itemize}
-			\item E.g., this is how the P4 firmware propagates rules from master island to P4 workers.
-			\item My impl.\ specific---shared freelist for memory pointers in EMEM.
-		\end{itemize}
-		\item Controller is \emph{always} co-hosted, interfaces over PCI-e. This also applies to Barefoot Tofino.
-	\end{itemize}
-	\item Policy configuration.
-	\begin{itemize}
-		\item (Re)-configurable at runtime.
-		\item Custom DSCP marker acts as indicator of RL packets (i.e., bother with parsing past UDP).
-		\item P4 parser extracts headers, parser state and payload passed over ring to another core dedicated to RL operations.
-		\item Policy dimensions, size, action count, tiling strategies, learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) can all be chosen and changed live.
-	\end{itemize}
-	\item State Handling.
-	\begin{itemize}
-		\item State input to model \emph{always} should produce an action.
-		\item State vectors should come from elsewhere on the device: this may either be a dedicated island/core cluster (i.e., also performing INT management), or it may be a packet from the network.
-	\end{itemize}
-	\item Action Handling.
-	\begin{itemize}
-		\item A state vector is tile coded, action probabilities are produced, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
-		\item \emph{Open question: how to configure action destination?
-This should either be another core (after an optional transformation), a packet sent out to a destination, maybe talk back to the controller.}
-	\end{itemize}
-	\item Learning.
-	\begin{itemize}
-		\item \emph{Open question: how to safely \& reasonably map state back to incoming rewards?}
-		\item \emph{Open question: Can/should this also support on-chip reward signals?}
-	\end{itemize}
-\end{itemize}
 
@@
 
+?? Event-driven model could better support many timer-driven RL use-cases \textcite{DBLP:conf/hotnets/IbanezABM19}.
+
 ?? Taurus. They can't train online though!~\parencite{DBLP:journals/corr/abs-2002-08987}
@@
 
+?? Mechanisms to support full in-switch DDoS detection~\cite{tnms-ddos-victim-ident}.
+
 \section{Conclusion}

### 2021-04-19 tables/src/latency-tput.py
Time/throughput table.

@@ -1,0 +1,79 @@
+import numpy as np
+import sys
+
+sys.stdout.reconfigure(encoding='utf-8')
+
+# sub in: bit depth, fw-type, measure_name
+results_fmt_str = "../results/rl-perf-tester/vary-work-ct/{}/{}.28d.31c.{}.dat"
+
+cycle_in_ns = 1e9 / 1.2e9
+
+core_ct_for_tput = 32
+types = [("single", "Single", True), ("balanced", "Multi", False)]
+measures = [("UpdateAll", R"\cmark", True), ("ComputeAndWriteout", R"\xmark", False)]
+bitdepths = [32, 16, 8]
+
+print(R"\begin{tabular}{@{}ccSSSSSS@{}}")
+print(R"\toprule\multicolumn{1}{c}{Bitdepth} & \multicolumn{1}{c}{Online} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{4}{c}{Completion (\si{\micro\second})}\\")
+
+# Just generate the next header line...
+t2_header = R"\cmidrule(lr){3-4}\cmidrule(lr){5-8} &"
+
+for i in range(2):
+	for (fname_part, pres_part, div_by_core_ct) in types:
+		parts = [pres_part]
+		if i != 0:
+			parts.append(pres_part + R" (\nth{99})")
+
+		for part in parts:
+			t2_header += R" & \multicolumn{1}{c}{"+part+R"}"
+
+t2_header += R"\\ \midrule"
+
+print(t2_header)
+
+for bits in bitdepths:
+	for online_i, (online_fname_part, online_latex_part, needs_lock) in enumerate(measures):
+		out = "{} & {}".format(bits if online_i == 0 else "", online_latex_part)
+
+		latencies = []
+		tputs = []
+
+		# Throughput
+		for type_fname_part, _p, do_div in types:
+			fname = results_fmt_str.format(bits, type_fname_part, online_fname_part)
+			indiv_data = []
+			with open(fname) as of:
+				lines = of.readlines()
+				indiv_data = np.array([float(x.strip()) for x in lines])
+			indiv_data *= cycle_in_ns;
+			indiv_data /= 1.0e3;
+
+			stats = [
+				np.median(indiv_data),
+				np.percentile(indiv_data, 99),
+			]
+
+			latencies.append("{:.3f}".format(stats[0]))
[…]

### 2021-04-19 tables/src/latency-tput.py
Plots cleaned up, ready for insertion after build...

@@ -9,12 +9,12 @@ results_fmt_str = "../results/rl-perf-tester/vary-work-ct/{}/{}.28d.31c.{}.dat"
 cycle_in_ns = 1e9 / 1.2e9
 
 core_ct_for_tput = 32
-types = [("single", "Single", True), ("balanced", "Multi", False)]
+types = [("single", "Single", True), ("balanced", "Parallel", False)]
 measures = [("UpdateAll", R"\cmark", True), ("ComputeAndWriteout", R"\xmark", False)]
 bitdepths = [32, 16, 8]
 
 print(R"\begin{tabular}{@{}ccSSSSSS@{}}")
-print(R"\toprule\multicolumn{1}{c}{Bitdepth} & \multicolumn{1}{c}{Online} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{4}{c}{Completion (\si{\micro\second})}\\")
+print(R"\toprule\multicolumn{1}{c}{Bits} & \multicolumn{1}{c}{Online} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{4}{c}{Completion (\si{\micro\second})}\\")
 
 # Just generate the next header line...
 t2_header = R"\cmidrule(lr){3-4}\cmidrule(lr){5-8} &"

### 2021-04-20 paper/paper.tex
Time/throughput table.
Latency-focussed plots?

@@
 \usepackage{csquotes}
+\usepackage{pifont}
 
@@
 \newcommand{\approachshort}{OPaL}
+
+\makeatletter\let\expandableinput\@@input\makeatother
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
 
@@
 
+?? Mention that you compute action, send action, AND THEN compute update exactly as in that one paper to reduce latency
+
 \begin{itemize}
@@
 
+\section{Potential Integrations}
+
+\subsection{DDoS Defence}
+?? Link to my own work
+?? Explain how every part may be substituted
+?? Bump-in-the-wire at edge nodes.
+
+\subsection{Something host-specific?}
+Accelerate something at an end-host...
+
 \section{Evaluation}
+
+\subsection{Experimental Setup}
+
+\section{Results}
+
+\begin{table}
+	\caption{Throughput and per-item compute time for an RL agent with/without policy learning.
+State$\rightarrow$action latency equals the offline completion cost.
+TODO: compare to host machine?
+Put completion times first?
+Put multi first?}
+	\resizebox{\linewidth}{!}{
+		\expandableinput ../tables/build/latency-tput
+	}
+\end{table}
+
+\begin{figure}
+	\resizebox{1.0\linewidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
+	}
+	\caption{?? Choice of work strategy is pretty important I guess}
+\end{figure}
+
+\begin{figure}
+	\resizebox{1.0\linewidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/vary-work}
+	}
+	\caption{?? Max Parallelisation. ?? Bitdepth affects independent cores more.}
+\end{figure}
+
+\begin{figure}
+	\resizebox{1.0\linewidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/vary-core}
+	}
+	\caption{?? Max-size work (129 work-items). ?? Need to dedicate 3 worker to beat independent cores. ?? Sharper perf increase when new physical core added. ?? up to 31 due to master thread.}
+\end{figure}
 

### 2021-04-20 paper/paper.tex
More work on design section, starting evaluation.

@@
 % Copyright
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
-\setcopyright{none}
-%\setcopyright{acmcopyright}
+%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
+%\setcopyright{none}
+\setcopyright{acmcopyright}
 %\setcopyright{acmlicensed}
@@
 
-\settopmatter{printacmref=false, printccs=false, printfolios=true}
+%\settopmatter{printacmref=false, printccs=false, printfolios=true}
+\settopmatter{printfolios=true}
 
@@
 %Conference
-\acmConference[SOSR '21]{Symposium of SDx Research 2021}{\numrange{13}{14} September}{Virtual Conference}
-%\acmYear{2021}
+\acmConference[SOSR '21]{Symposium of SDx Research 2021}{September \numrange{13}{14}, 2021}{Online}
+\acmYear{2021}
 %\copyrightyear{}
@@
 %% {} with no args suppresses printing of the price
-\acmPrice{}
+\acmPrice{15.00}
 
@@
 \begin{document}
-	\title{Revisiting the Classics: Online RL in the Programmable Data-Plane}
-
-	%\titlenote{Produces the permission block, and copyright information}
+	\title{Revisiting~the~Classics: Online~RL~in~the~Programmable~Data-Plane}
+
+%	\titlenote{Produces the permission block, and copyright information}
 	%\subtitle{Extended Abstract}
@@
 \begin{abstract}
-Note: below text is just the UK systems workshop abstract I submitted.	
+??
+In-NIC execution allows clever use of resources to bring action compute latency  from \SI{1}{\milli\second} to \SI{50}{\micro\second} (2 orders of magnitude!) in real-world policies for DDoS detection/prevention control.
 \end{abstract}
+
+\begin{CCSXML}
+	<ccs2012>
+	<concept>
+	<concept_id>10003033.10003068.10003069</concept_id>
+	<concept_desc>Networks~Data path algorithms</concept_desc>
+	<concept_significance>500</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10003033.10003099.10003102</concept_id>
+	<concept_desc>Networks~Programmable networks</concept_desc>
+	<concept_significance>500</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10003752.10010070.10010071.10010261</concept_id>
+	<concept_desc>Theory of computation~Reinforcement learning</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10010147.10010257.10010258.10010261</concept_id>
+	<concept_desc>Computing methodologies~Reinforcement learning</concept_desc>
+	<concept_significance>500</concept_significance>
+	</concept>
+	</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Networks~Data path algorithms}
+\ccsdesc[500]{Networks~Programmable networks}
+\ccsdesc[500]{Computing methodologies~Reinforcement learning}
+
+%\keywords{ACM proceedings, \LaTeX, text tagging}
 
@@
 
-\section{Motivation}\label{sec:motivation}
+\section{Background and Motivation}\label{sec:motivation}
+?? RL~\parencite{RL2E}
+
+?? Use this to explain differences between the modern state of the art and older works.
+
+?? Probably give a rough introduction to the algorithm, show what arithmetic operations are needed---particularly when using Qm.n fixed point.
+
+?? Tile coding requires div, though can be simplified using binary tile widths (relates down to bitshifts). What other ops?
+
+?? Explain how it is an embarrassingly parallel problem? Or at least very amenable to partition-aggregate.
+
+?? Critical section around shared writeback loc, but individual work items are disjoint in their input data (particularly useful for updates).
+
 \fakepara{Reinforcement learning in asynchronous environments}
@@
 
-\section{Background}
-
-\subsection{Classical Reinforcement Learning}
-?? Move into background parent section?
-
-?? RL~\parencite{RL2E}
-
-?? Use this to explain differences between the modern state of the art and older works.
-
-?? Probably give a rough introduction to the algorithm, show what arithmetic operations are needed---particularly when using Qm.n fixed point.
-
-?? Explain how it is an embarrassingly parallel problem? Or at least very amenable to partition-aggregate.
-
-?? Critical section around shared writeback loc, but individual work items are disjoint in their input data (particularly useful for updates).
-
-\subsection{Hmm..}
-Should motivation be in here? Unsure what other preliminaries need to be explained.
-
 \section{Design and Implementation}\label{sec:design}
-Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{}}.
+Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{} (\approach)}.
 At a high level, \approachshort{} is designed to use the auxiliary compute exposed by general SmartNIC devices to offer low-latency online learning, scaling according to available on-chip resources at build time.
@@
 \Cref{fig:netro-arch} outlines our design and implementation on Netronome SmartNIC hardware in pursuit of this goal.
+We open-source our firmware design and control programs for the benefit of the community\footnote{\url{https://github.com/FelixMcFelix/pdp-rl-paper}---currently private...}.
 We describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.
@@
 	\end{tikzpicture}}
-	\caption{Architecture for an \approachshort{} agent on Netronome hardware, using 4 cores for parallel policy processing. (FIXME: redo in illustrator? Show CTXes, PISA)\label{fig:netro-arch}}
+	\caption{Architecture for an \approachshort{} agent on Netronome hardware, using 4 cores for parallel policy processing.
+(FIXME: redo in illustrator?
+Show CTXes, PISA, single vs parallel)\label{fig:netro-arch}}
 \end{figure}
@@
 \approachshort{} is a general, task-independent framework for in-network, online training and execution of \emph{any reinforcement learning agent design} using classical methods.
-The meaning 
+\approachshort{} is agnostic to the meaning of state vectors it receives as inputs and the actions it produces, which are employed by other functional units or the dataplane.
 However, in-NIC/in-network execution specifically benefits packet-, flow-, and network-level learning, control, and optimisation tasks.
-\approachshort{} runs on one more cores of a SmartNIC to convert state measurements from the environment into a stream of actions using a stored policy.
-These dedicated cores (\emph{micro-engines} grouped by \emph{island} in NFP terminology) are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
+
+\approachshort{} runs on one more cores of a SmartNIC to convert fixed-point state measurements from the environment into a stream of actions using a stored policy.
+These dedicated cores (\emph{microengines} grouped by \emph{island} in NFP terminology) are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
 Combined with reward measurements, this policy can then be updated or trained from scratch entirely on the NIC, acting as a fully online RL agent.
 An input state vector \emph{always} induces an action, and if desired updates the policy using either an included reward or one retrieved from memory according to a (configurable) key placed alongside the state.
+This allows for simultaneous control and learning over independent systems by the same agent (i.e., optimising several flows with their own reward measures).
 
@@
 Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
-
-?? Dynamic, runtime configuration via control and data planes.
-
-?? Compile-time config of bit-depth, 
-?? Allows learning from several live traces simultaneously.
-?? Allows execution of indepdnnt traces simultaneously.
-
-\subsection{Action Computation}
-Note: talk about single core AND parallel.
-
-?? Mention that you compute action, send action, AND THEN compute update exactly as in that one paper to reduce latency
-
-\begin{itemize}
-	\item A state vector is tile coded, action probabilities are produced, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
-	\item \emph{Open question: how to configure action destination?
-This should either be another core (after an optional transformation), a packet sent out to a destination, maybe talk back to the controller.}
-\end{itemize}
-\begin{itemize}
-	\item \emph{Open question: how to safely \& reasonably map state back to incoming rewards?}
-	\item \emph{Open question: Can/should this also support on-chip reward signals?}
-\end{itemize}
-
-?? Split across MEs.
-
-?? each ME can then subdivide by factor of 8 among contexts.
-
-?? Scales from 1--(n-1) cores.
+Bitdepth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
+
+\subsection{Action and Update Computation}
+\approachshort{} applies the insights of \textcite{DBLP:journals/firai/TravnikMSP18} to minimise action latency; an action is computed, sent out into the environment, and only then is the underlying policy updated.
+Using one of the below strategies chosen at compile time, a state vector is tile coded, converted into action probabilities, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
+This is then written out to the environment as in \cref{sec:agent-environment-communication}.
+If updates are enabled, \approachshort{} then checks an internal hashmap for a previous state-action pair matching the current instruction source, and if found then a policy update is performed.
+Updates are computed using \emph{single-step semi-gradient Sarsa}, though modification to support other single-step methods would be trivial.
+The new state- or tiles-action pair is then written into storage
+
+\begin{figure}
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[keepaspectratio, width=\linewidth]{figures/single-core-arch-sketch}
+		\caption{Single (offline throughput-optimal).\label{fig:single-and-parallel:single}}
+	\end{subfigure}
+	\begin{subfigure}{0.45\linewidth}
+		\centering
+		\includegraphics[keepaspectratio, width=\linewidth]{figures/parallel-arch-sketch}
+		\caption{Parallel (online-optimal).\label{fig:single-and-parallel:parallel}}
+	\end{subfigure}
+	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to latency/throughput needs. ?? MAKE ME FOR REAL \label{fig:single-and-parallel}}
+\end{figure}
+
+?? Do we need better/snappier names than just Single/Parallel? It's awfully dry.
+
+In the NFP architecture we use, physical cores (MEs) each have 8 hardware threads (contexts): all contexts on an ME share the same program and employ zero-cost context switches on I/O.
+Each core then provides 8 worker threads, though in the parallel case one must be dedicated to monitoring work acknowledgements.
+Two main firmware models govern how the computation-intensive parts of these tasks (tile-coding, action preference list construction, policy updates) are carried out:
+\begin{description}
+	\item[Single (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each performs its work sequentially.
+Computing an action list requires a \emph{read lock} on the policy.
+If an update occurs, the core requests a \emph{write lock} before updating, greatly limiting online throughput.
+Tile lists are stored for update computation.
+	\item[Parallel (\cref{fig:single-and-parallel:parallel})] Threads cooperate on processing state vectors, minimising latency.
+Workers have a fixed list of work items, while a master thread sends compute/update commands before awaiting worker completion.
+Work items are disjoint, requiring no policy locks.
+State vectors are stored for update computation.
+\end{description}
+Each offers a different point of optimisation; if updates are disabled, then the \emph{single} model can maximise throughput, while the \emph{parallel} model is designed to minimise decision latency and needs no locks for updating the policy (increasing online learning throughput).
+
+We found that, paradoxically, it was more efficient in the parallel case \emph{to do more work} by having each worker recompute its tile subset from a stored state.
+It transpired that cacheing this data placed a larger \texttt{memcpy} in the serial section, whose size did not scale at all with bit depth.
 
@@
 
+?? Reconfig triggers cacheing of per-work item data: tiling locations, ....
+
+Written in rust.
+
 \subsection{Work Allocation}
@@
 
+?? Work index correlates with complexity (and can be estimated a-priori).
+
 \subsection{Policy Storage}
@@
 
+?? Batching of reads. ?? action value batching  (relate to bitdepth).
+
 ?? Lower bitdepth can mean more actions, more tilings, more dimensions in tilings, higher perf.
@@
 Can changes in trained policy be used to transfer to a more complex function approximator elsewhere?
+
+\subsection{Limitations}
+Known/understood (but not implemented)?
+\begin{itemize}
+	\item Figure out PIF rule installation.
+	\begin{itemize}
+		\item Not doable through P4 na\"{i}vely. Handing back rules to controller is expensive: need another action to deliberately use its own hash table etc., then batch back as faster P4 rules.
+	\end{itemize}
+\end{itemize}
 
@@
 
+?? Place Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} on the datapath as the main source of state vectors.
+
+?? Use the TNSM work you reviewed~\parencite{tnms-ddos-victim-ident} as a reward function source.
+
 \subsection{Something host-specific?}
@@
 \subsection{Experimental Setup}
+Evaluation of \approachshort{} was performed on server blades hosting a single Netronome Agilio LX 1x100GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
+These ran Ubuntu 18.04.5 LTS (4.15.0-140-generic x86\_64) on Intel Xeon Bronze 3204 CPUs (\SI{1.9}{\giga\hertz}) and \SI{32}{\gibi\byte} RAM.
+Control programs were built using rustc 1.51.0.
+All timing and throughput measurements were repeated over 5000 trials.
+
+Studies based on \textcite{DBLP:journals/tnsm/SimpsonRP20} mirror their experimental setup (Ubuntu 18.04.2 LTS (4.4.3-040403-generic x86\_64), Intel Core i7-6700K (\SI{4.2}{\giga\hertz}), \SI{32}{\gibi\byte} RAM).
 
@@
 	}
-	\caption{?? Max Parallelisation. ?? Bitdepth affects independent cores more.}
+	\caption{?? More work means More time. ?? Max Parallelisation. ?? Bitdepth affects independent cores more.}
 \end{figure}
@@
 	}
-	\caption{?? Max-size work (129 work-items). ?? Need to dedicate 3 worker to beat independent cores. ?? Sharper perf increase when new physical core added. ?? up to 31 due to master thread.}
+	\caption{??
+More cores means less time ??
+Max-size work (129 work-items).
+??
+Need to dedicate 3 worker to beat independent cores.
+??
+Sharper perf increase when new physical core added.
+??
+up to 31 due to master thread.}
 \end{figure}
@@
 \end{enumerate}
-
-\section{What's implemented 'til now?}
-System?
-\begin{itemize}
-	\item Core-to-core communication.
-	\item P4 parser for relevant packets.
-	\item Packet copy + handoff from P4 to RL MEs.
-	\item Dynamic base configuration.
-	\item Dynamic tiling configuration.
-	\item Policy installation via block-copy.
-	\item Packet generation for above config functions, as well as policy quantisation and splitting.
-	\item Tile-coding of quantised states on-chip, and state packet receipt.
-	\item Action computation on-chip.
-	\item Rudimentary, single-core training.
-\end{itemize}
-
-Experiments?
-\begin{itemize}
-	\item Policy quantisation divergence \& accuracy (via mininet).
-	\item Measure DPDK SmartNIC RTT as a lower bound on controller access time.
-	\begin{itemize}
-		\item Same mechanism applies to other manufacturers (Barefoot), so generalisable.
-		\item My results are non-DPDK, but there is a good study~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018} on exactly this.
-	\end{itemize}
-	\item Measure basic on-switch times: ring-ring communication, tile-coding cost, action compute cost, learning cost.
-\end{itemize}
-
-Known/understood (but not implemented)?
-\begin{itemize}
-	\item Figure out PIF rule installation.
-	\begin{itemize}
-		\item Not doable through P4 na\"{i}vely. Handing back rules to controller is expensive: need another action to deliberately use its own hash table etc., then batch back as faster P4 rules.
-	\end{itemize}
-\end{itemize}
 
@@
 
+\begin{acks}
+	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grant~\grantnum{gs-epsrc}{EP/N509668/1}].
+\end{acks}
+
 %\bibliographystyle{ACM-Reference-Format}

### 2021-04-22 paper/paper.tex
Partial work on evaluation and such.

@@
 \usetikzlibrary{arrows.meta, calc, fit, positioning}
+
+\usepackage[title]{appendix}
 
@@
 \usepackage[font=footnotesize]{subcaption}
+\usepackage{awesomebox}
 
@@
 
+\newcounter{insightc}
+\newenvironment{insight}
+	{
+		\begin{tipblock}\refstepcounter{insightc}\textbf{Insight \theinsightc:}\em
+	}
+	{
+		\end{tipblock}
+	}
+
 \begin{document}
-	\title{Revisiting~the~Classics: Online~RL~in~the~Programmable~Data-Plane}
+	\title{Revisiting~the~Classics: Online~RL~in~the~Programmable~Dataplane}
 
@@
 
+\setlength{\aweboxleftmargin}{0.12\linewidth}
+\setlength{\aweboxcontentwidth}{1.97\linewidth}
+
 \section{Introduction}
+?? REDO ME
 
@@
 \begin{itemize}
-	\item A thing (?? REFERENCE SECTION). Here's how that thing is utterly fantastic and will change the world, or not.
+	\item An analysis of why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility (\cref{sec:motivation}),
+	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands (\cref{sec:design}),
+	\item Potential integrations of \approachshort{} with state-of-the-art PDP applications to perform fully in-NIC, automated DDoS mitigation and ?? OTHER THING (\cref{sec:potential-integrations}),
+	\item In-depth evaluation of how our approach affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{0}{$\times$} latency reduction (\crefrange{sec:evaluation}{sec:results}).
+??actual val
 \end{itemize}
@@
 ?? Critical section around shared writeback loc, but individual work items are disjoint in their input data (particularly useful for updates).
+
+\begin{insight}
+	Classical RL methods are computationally simple to update online, and inference and learning are embarrassingly parallel.
+\end{insight}
 
@@
 
+\begin{insight}
+	Online control problems benefit from low-latency, local execution and training (i.e., in-NIC/switch).
+\end{insight}
+
 \fakepara{Programmable hardware capabilities}
@@
 Moreover, the limited memory/block RAM and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
+
+\begin{insight}
+	Low-port density programmable network devices often have spare resources and extra capabilities beyond the PISA specification which can aid asynchronous, local compute.
+\end{insight}
 
@@
 
-?? relate update/execute operations here? i.e., policy execute needs only ALU add, train needs add+mul+shr. or in RL section? Should that come first?
+?? relate update/execute operations here? i.e., policy execute needs only ALU add (and whatever is needed for tile code), train needs add+mul+shr. or in RL section? Should that come first?
 
@@
 Updates are computed using \emph{single-step semi-gradient Sarsa}, though modification to support other single-step methods would be trivial.
-The new state- or tiles-action pair is then written into storage
+The new state- or tiles-action pair is then written into storage.
 
@@
 \end{description}
-Each offers a different point of optimisation; if updates are disabled, then the \emph{single} model can maximise throughput, while the \emph{parallel} model is designed to minimise decision latency and needs no locks for updating the policy (increasing online learning throughput).
+Each offers a different point of optimisation; if updates are disabled, then the \emph{single} model can maximise throughput, while the \emph{parallel} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
 
@@
 
-\subsection{Reconfiguration}
-?? Compile-time config of bit-depth, what else?
-
-\begin{itemize}
-	\item (Re)-configurable at runtime.
-	\item Custom DSCP marker acts as indicator of RL packets (i.e., bother with parsing past UDP). This trick is copied from \textcite{DBLP:conf/isca/LiLYCSH19}.
-	\item P4 parser extracts headers, parser state and payload passed over ring to another core dedicated to RL operations.
-	\item Policy dimensions, size, action count, tiling strategies, learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) can all be chosen and changed live.
-\end{itemize}
-
-?? Reconfig triggers cacheing of per-work item data: tiling locations, ....
-
-Written in rust.
+\subsection{Reconfigurability}
+\approachshort{} allows for policy dimensions, size, action count, tiling strategies, and algorithmic learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) to be chosen and changed at runtime.
+This extends to policy data, which may be imported from an offline pre-trained model via such packets, and exported via the NFP binary support package on the host machine.
+Some aspects must be chosen at compile time; bit-depth, work allocation strategy, and maximum policy/tiling/state sizes---these govern core operation or allocated memory regions.
+
+In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved DSCP value as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+While we primarily use this to automate parser generation, this also allows for configuration packets to be received from only trusted hosts over the dataplane if needed via P4 rules.
+Our control packet generation library (and evaluation frameworks which build upon it) are written in Rust.
 
 \subsection{Work Allocation}
-?? WHy
-
-?? How
-
-?? Work index correlates with complexity (and can be estimated a-priori).
-
-\subsection{Policy Storage}
-?? Why
-
-?? Trained online or installed by block copy from control/data planes.
-
-?? Blocks of dense quantised tiles: each an action preference list.
-
-?? Tiling data is split into regions of memory based on access time: higher-dimension tilings are sent to further NUMA regions (EMEM/IMEM).
-
-\subsection{Variable Quantisation Bitdepth}
-Compile time eh
-
-?? Mention that BB SIMD failed here.
-
-??
-
-?? Batching of reads. ?? action value batching  (relate to bitdepth).
-
-?? Lower bitdepth can mean more actions, more tilings, more dimensions in tilings, higher perf.
-
-\subsection{Notes from chat with others}
-This approach can be used to generate transferable policies/samples for other models.
-
-Potential: low-latency train for some specific points, and high throughput for others? Think on the exact note I took...
-
-``can have accel'd offline, train online subset w/ my approach''
-
-THis might mean "do 32-bit online", then downsample to 8-bit policy for high-throughput mode.
-
-Can changes in trained policy be used to transfer to a more complex function approximator elsewhere?
+Due to the lack of dynamic memory allocation and to simplify action-value lookups, policies cannot be effectively stored sparsely.
+As a consequence, tiling space requirements scale exponentially with dimension count, and so higher-dimension tilings must be placed in memory regions large enough to allocate space for them.
+As a result, tilings are split across memory regions, giving different access and compute costs to different tasks---ranging over CLS, CTM and EMEM in increasing latency.
+
+We use a simple first-fit work placement algorithm run on the NFP, placing the largest work item into the least loaded thread of the least loaded core.
+The cost of any individual work item (according to its dimension count and memory location) was estimated offlin.
+This work allocation and any cached per-work-item values are recomputed when any configuration is installed or changed.
+
+\subsection{Variable Quantisation Bit Depth}
+At compile-time, \approachshort{} can be configured to use \SIlist[list-final-separator = { \translate{or} }]{32;16;8}{\bit} values in input states and its internal policies.
+Naturally, smaller bit depths reduce the storage required for policies, tiling data, and stored state-action pairs---allowing more complex problems to be modelled using more dimensions or fine-grained policies.
+Through the lens of computational efficiency, reduced bit depth allows action preference lists to be read using fewer I/O operations (as more values may fit into a single machine word).
+We had investigated bit-stuffing several such values into a single word for our atomic writeback mechanism (as the platform offers both \SIlist{32;64
+}{\bit} atomic addition).
+This is analogous to SIMD---through clever use of padding bits---but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
+
+\subsection{Notes from chat with Rhys}
+?? Where/how to fit this in? I think it's a useful point.
+
+The two compute models discussed here need not be homogeneously deployed.
+For instance, in a networked deployment a subset of \approachshort{} nodes could be \SI{32}{\bit} \emph{parallel} agents, training online, while most other nodes ran \SI{8}{\bit} \emph{single} designs to meet throughput guarantees.
+The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
+This can be taken further still, using policy deltas or execution traces to enable out-of-path transfer learning for more complex function approximators such as neural networks.
+
+%Potential: low-latency train for some specific points, and high throughput for others? Think on the exact note I took...
+%
+%``can have accel'd offline, train online subset w/ my approach''
+%
+%THis might mean "do 32-bit online", then downsample to 8-bit policy for high-throughput mode.
+%
+%Can changes in trained policy be used to transfer to a more complex function approximator elsewhere?
 
 \subsection{Limitations}
-Known/understood (but not implemented)?
-\begin{itemize}
-	\item Figure out PIF rule installation.
-	\begin{itemize}
-		\item Not doable through P4 na\"{i}vely. Handing back rules to controller is expensive: need another action to deliberately use its own hash table etc., then batch back as faster P4 rules.
-	\end{itemize}
-\end{itemize}
+Unfortunately, direct rule installation into P4 tables from the SmartNIC itself is not, in general, possible.
+To achieve line-rate match-action table lookup, platforms like NFP use accelerated datastructures (e.g., DCFL~\parencite{DBLP:conf/infocom/TaylorT05}) which must be computed by the controller over the \emph{entire rule set}.
+Even through externs,
+
+We instead suggest on NFP chips that externs or datapath stages which apply actions to packet processing should maintain a small local store of state-action pairs, and periodically send these back to the controller for batch rule installation.
+This ensures that the vast majority of installed rules enjoy accelerated performance, while preventing rule installation delay on the newest decisions.
+Platforms such as Intel Tofino greatly simplify this, with Tofino Native Architecture intrinsics such as \emph{Action Profiles/Selects} allowing a P4 action to be chosen based on a register value (i.e., an RL action).
 
@@
 ?? Parallel, atomic writes at the required resolution.
-?? No bit stuffing required.
+?? No bit stuffing required for SIMD-alike?.
+?? More chances for vectorisation? Is that where we're falling down atm with 32-bit being weirdly optimal?
 
@@
 
-\section{Potential Integrations}
-
-\subsection{DDoS Defence}
+\section{Potential Integrations}\label{sec:potential-integrations}
+To show the general applicability and use of \approachshort{}, we propose two ideal integrations which would benefit most from in-NIC reinforcement learning; online DDoS attack mitigation (\cref{sec:integ-1}), and ??
+OTHER THING (\cref{sec:integ-2}).
+We support these using other state-of-the-art P4/PDP developments.
+
+\subsection{DDoS Defence}\label{sec:integ-1}
 ?? Link to my own work
 ?? Explain how every part may be substituted
+
 ?? Bump-in-the-wire at edge nodes.
 
+?? How does increased attack flow ratio affect the RL core? How does changing its params to increase its own throughput?
+?? Remember the timed-random-selection; use the state coalescing trick indep discovered by you and those other guys.
+
+?? Can the P4 dataplane be used to reduce/increase the number of flows used as inputs? (YES)
+
 ?? Place Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} on the datapath as the main source of state vectors.
@@
 
-\subsection{Something host-specific?}
-Accelerate something at an end-host...
-
-\section{Evaluation}
+\subsection{Something host-specific?}\label{sec:integ-2}
+Accelerate something at an end-host... Flow control? Something datacentre-y? A novel CCA? What?
+
+\section{Evaluation}\label{sec:evaluation}
+We outline our performance evaluation of...
+
+?? find something meaningful to put here.
 
@@
 Control programs were built using rustc 1.51.0.
-All timing and throughput measurements were repeated over 5000 trials.
+We run \approachshort{} on an island with \num{4} MEs of the NFP-6480 for a total of \num{32} hardware threads.
+This is the maximum number of MEs on a single island which is not in use by an existing P4 pipeline.
+Where feasible, we test \SIlist{32;16;8}{\bit} quantised arithmetic firmware versions.
+All timing and throughput measurements were repeated over \num{5000} trials.
+
+Policy sizes are set to those matching a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dimension state vectors, a bias tile and 16 full tiling sets (7$\times$1-dim, 8$\times$2-dim, 1$\times$4-dim), 8 tilings per dimension set, 6 tiles per dimension, choosing between 10 actions.
 
@@
 
-\section{Results}
+\subsection{Experiments}
+
+\fakepara{Raw inference and learning performance}
+We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs.
+We compare these against floating point and quantised implementations running on a commodity host machine.
+This allows us to demonstrate the performance differences between the \emph{Single} and \emph{Parallel} configurations, particularly in how \emph{Single's} required policy locks impact throughput.
+We also report on the derived throughput measures accounting for the locking behaviour of each.
+
+Building on this, we vary the amount of allocated worker threads to show how \approachshort{} can scale to fit available compute resources on a device.
+This also demonstrates the number of cores needed to achieve a given latency bound on a policy of representative complexity.
+Moreover, to demonstrate how these costs vary as policy complexity increases, we vary the number of total dimensions included in the tiling.
+
+?? NOT GOT OTHERS TO COMPARE AGAINST YET, JUST SELF.
+?? Mention versus DNN latency in passing here?
+
+\fakepara{Work allocation}
+%?? Also our simple, ahead-of-time, work scheduler.
+We verify that our heuristic, runtime work scheduler (\emph{Balanced}) makes meaningful use of the explicit memory hierarchy and cost of each work item.
+We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} baseline allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ items given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
+
+\fakepara{End-to-end RL latency}
+We compare the key RL decision-making latencies we discuss in \cref{fig:state-slip} ($t_1,t_2,t_3$) across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to the controller of an OpenFlow-compatible switch (OvS).
+
+?? NOT DONE YET
+?? MOSTLY A PAPER EXERCISE FROM SOME OTHER MEASURES
+
+\fakepara{Co-existence with the dataplane}
+We measure maximum packet rate and bandwidth throughput of a P4 pipeline on our SmartNICs, both with and without \approachshort{}.
+This allows us to quantify whether on-chip (out-of-path) execution impacts ordinary dataplane behaviour through indirect means: e.g., EMEM cache evictions or hidden resource contention.
+In particular, we stress test both decisions and policy updates of our \emph{Parallel} design at full throughput.
+
+?? NOT DONE YET.
+
+\fakepara{Resource requirements}
+Using the maximum policy size defined above, we investigate how the memory requirements imposed by \approachshort{} vary with the number of dedicated MEs, over and above a base P4 forwarding plane.
+
+?? Not done yet: read these out from the IDE.
+
+\fakepara{The impact of bit depth}
+We further investigate how bit depth affects the accuracy of policy by varying the number of bits used to represent the mantissa of action values and input state.
+
+\subsection{Results and Discussion}\label{sec:results}
 
@@
 	\resizebox{1.0\linewidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
-	}
-	\caption{?? Choice of work strategy is pretty important I guess}
-\end{figure}
-
-\begin{figure}
-	\resizebox{1.0\linewidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-work}
 	}
-	\caption{?? More work means More time. ?? Max Parallelisation. ?? Bitdepth affects independent cores more.}
+	\caption{?? More work means More time. ?? Max Parallelisation. ?? Bitdepth affects independent cores more. ?? latency version in appendix: \Cref{fig:vary-work-latency}}
 \end{figure}
@@
 ??
-up to 31 due to master thread.}
+up to 31 due to master thread.
+??
+latency version in appendix: \Cref{fig:vary-core-latency}}
 \end{figure}
 
+\fakepara{Raw inference and learning performance}
+TEST
+
+\fakepara{Work allocation}
+?? \Cref{fig:work-alloc-32}
+
+\begin{figure}
+	\resizebox{1.0\linewidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
+	}
+	\caption{?? Choice of work strategy is pretty important I guess\label{fig:work-alloc-32}}
+\end{figure}
+
+\fakepara{End-to-end RL latency}
+TEST
+
+\fakepara{Co-existence with the dataplane}
+TEST
+
+\fakepara{Resource requirements}
+?? \Cref{tab:resources}
+
+\begin{table}
+\caption{NFP memory use due to \approachshort{}.
+CLS and CTM are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM are shared between all NFP programs on a NIC.\label{tab:resources}}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{@{}cSSSSSSSSSS@{}}
+	\toprule Firmware & \multicolumn{2}{c}{EMEM} & \multicolumn{2}{c}{EMEM Cache} & \multicolumn{2}{c}{IMEM} & \multicolumn{2}{c}{i5.CLS} & \multicolumn{2}{c}{i5.CTM}\\
+	& \multicolumn{1}{c}{\si{\mebi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} \\
+	\midrule Base P4 & 6794.73 & 88.47 & 0 & 0 & 858.28 & 10.48 & 0 & 0 & 0 & 0 \\
+	Single (1-core) & 6778.83 & 88.27 & 1737.91 & 18.86 & 1263.28 & 15.42 & 28.69 & 44.82 & 90 & 35.16 \\
+	Single (4-core) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+	Parallel (1-core) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+	Parallel (4-core) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+	\bottomrule
+\end{tabular}
+}
+\end{table}
+
+\fakepara{The impact of bit depth}
+?? \Cref{fig:quant-acc}
+
+\begin{figure}
+	\includegraphics{../plots/build/marl-quant/accuracy-binary}
+	\caption{Normalised accuracy of a converted, pre-trained floating-point tile-coded policy after conversion to $32Qn$ fixed-point.}\label{fig:quant-acc}
+\end{figure}
+
+\section{Related Work}
+
+?? A switch/NIC architecture like \textcite{DBLP:conf/hotnets/StephensAS18} would remove the need for hard-coded island-island relationships.
+
+?? Event-driven model could better support many timer-driven RL use-cases \textcite{DBLP:conf/hotnets/IbanezABM19}.
+
+?? Taurus. They can't train online though!~\parencite{DBLP:journals/corr/abs-2002-08987}
+
+?? Do switches~\parencite{DBLP:conf/hotnets/XiongZ19}.
+?? Train offline, classical methods targetting all match-action P4 systems (no externs etc).
+
+?? N3IC run NNs on NIC~\parencite{DBLP:journals/corr/abs-2009-02353}, allow packet tagging to occur based on pre-trained BNN~\parencite{DBLP:conf/nips/HubaraCSEB16}.
+?? both NFP and NetFPGA.
+?? Or as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
+?? Binary NNs ``more suitable for network hardware''~\parencite{DBLP:journals/corr/MiyashitaLM16}?
+
+?? Can we collect some papers on per-packet ML/RL? Per-packet classification using NNs to build decision trees~\parencite{DBLP:conf/sigcomm/LiangZJS19}
+
+?? RL for congestion control. ~\parencite{DBLP:journals/corr/abs-1910-04054}
+
+?? Arbitrary training enhanced by in-network? General NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18}, RL-specfic~\parencite{DBLP:conf/isca/LiLYCSH19}.
+
+?? Mechanisms to support full in-switch DDoS detection~\cite{tnms-ddos-victim-ident}.
+
+\section{Conclusion}
+
+\begin{acks}
+	The authors would like to thank Rhys Simpson for his comments and discussions on SIMD-like optimisations.
+	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grant~\grantnum{gs-epsrc}{EP/N509668/1}].
+\end{acks}
+
+%\bibliographystyle{ACM-Reference-Format}
+%\bibliography{reference}
+\printbibliography
+
+\begin{appendices}
+	\section{Old ``What could we do``?}
 	How do we evaluate this without looking into classification performance?
@@
 
-\begin{figure}
-	\includegraphics{../plots/build/marl-quant/accuracy-binary}
-	\caption{Normalised accuracy of a converted, pre-trained floating-point tile-coded policy after conversion to $32Qn$ fixed-point.}\label{fig:quant-acc}
-\end{figure}
-
 	\subsubsection{Interaction with stationarity}
@@
 
-\section{Related Work}
-
-?? A switch/NIC architecture like \textcite{DBLP:conf/hotnets/StephensAS18} would remove the need for hard-coded island-island relationships.
-
-?? Event-driven model could better support many timer-driven RL use-cases \textcite{DBLP:conf/hotnets/IbanezABM19}.
-
-?? Taurus. They can't train online though!~\parencite{DBLP:journals/corr/abs-2002-08987}
-
-?? Do switches~\parencite{DBLP:conf/hotnets/XiongZ19}.
-?? Train offline, classical methods targetting all match-action P4 systems (no externs etc).
-
-?? N3IC run NNs on NIC~\parencite{DBLP:journals/corr/abs-2009-02353}, allow packet tagging to occur based on pre-trained BNN~\parencite{DBLP:conf/nips/HubaraCSEB16}.
-?? both NFP and NetFPGA.
-?? Or as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
-?? Binary NNs ``more suitable for network hardware''~\parencite{DBLP:journals/corr/MiyashitaLM16}?
-
-?? Can we collect some papers on per-packet ML/RL? Per-packet classification using NNs to build decision trees~\parencite{DBLP:conf/sigcomm/LiangZJS19}
-
-?? RL for congestion control. ~\parencite{DBLP:journals/corr/abs-1910-04054}
-
-?? Arbitrary training enhanced by in-network? General NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18}, RL-specfic~\parencite{DBLP:conf/isca/LiLYCSH19}.
-
-?? Mechanisms to support full in-switch DDoS detection~\cite{tnms-ddos-victim-ident}.
-
-\section{Conclusion}
-
-\begin{acks}
-	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grant~\grantnum{gs-epsrc}{EP/N509668/1}].
-\end{acks}
-
-%\bibliographystyle{ACM-Reference-Format}
-%\bibliography{reference}
-\printbibliography
+	\begin{figure}
+		\resizebox{1.0\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
+		}
+		\caption{Latency! varying work.\label{fig:vary-work-latency}}
+	\end{figure}
+
+	\begin{figure}
+		\resizebox{1.0\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
+		}
+		\caption{Latency! varying core count\label{fig:vary-core-latency}}
+	\end{figure}
+\end{appendices}
 

### 2021-04-22 code/nfp/rl-test/policy/policy_mem.h
Modifications needed to get memory numbers for 4xsingle...
This WILL need the Single expts to be re-run.

@@ -3,9 +3,9 @@
 
 #include "../rl.h"
 
-__declspec(i5.cls, import, aligned(8))tile_t t1_tiles[MAX_CLS_TILES] = {0};
-__declspec(i5.ctm, import, aligned(8))tile_t t2_tiles[MAX_CTM_TILES] = {0};
-__declspec(import, imem, aligned(8))tile_t t3_tiles[MAX_IMEM_TILES] = {0};
+__declspec(cls, import, aligned(8))tile_t t1_tiles[MAX_CLS_TILES];
+__declspec(ctm, import, aligned(8))tile_t t2_tiles[MAX_CTM_TILES];
+__declspec(import, imem, aligned(8))tile_t t3_tiles[MAX_IMEM_TILES];
 
 __declspec(import, emem) struct rl_config cfg = {0};
 

### 2021-04-22 code/nfp/rl-test/policy/policy_mem_define.h

@@ -3,8 +3,8 @@
 
 #include "../rl.h"
 
-__declspec(i5.cls, export, aligned(8))tile_t t1_tiles[MAX_CLS_TILES] = {0};
-__declspec(i5.ctm, export, aligned(8))tile_t t2_tiles[MAX_CTM_TILES] = {0};
+__declspec(cls, export, aligned(8))tile_t t1_tiles[MAX_CLS_TILES] = {0};
+__declspec(ctm, export, aligned(8))tile_t t2_tiles[MAX_CTM_TILES] = {0};
 __declspec(export, imem, aligned(8))tile_t t3_tiles[MAX_IMEM_TILES] = {0};
 
 // Maybe keep one of these locally, too?

### 2021-04-22 code/nfp/rl-test/rl-import.c

@@ -1,0 +1,882 @@
+#include <memory.h>
+#include <nfp.h>
+#include <nfp/me.h>
+//#include <nfp_mem_ring.h>
+//#include <nfp_mem_workq.h>
+#include <nfp/mem_bulk.h>
+//#include <nfp/mem_lkup.h>
+#include <nfp/mem_ring.h>
+#include <rtl.h>
+#include <lu/cam_hash.h>
+// #include <lu/cam_hash_ctrl.h>
+#include "rl.h"
+#include "rl-pkt-store.h"
+#include "pif_parrep.h"
+
+#include "subtask/ack.h"
+#include "subtask/work.h"
+#include "quant_math.h"
+#include "worker_config.h"
+
+#include "policy/policy_mem.h"
+
+#include "util.h"
+
+// ring head and tail on i25 or emem1
+volatile __emem_n(0) __declspec(import, addr40, aligned(512*1024*sizeof(unsigned int))) uint32_t rl_mem_workq[512*1024] = {0};
+// _NFP_CHIPRES_ASM(.alloc_resource rl_mem_workq_rnum emem0_queues+RL_RING_NUMBER global 1)
+//_NFP_CHIPRES_ASM(.init_mu_ring rl_mem_workq_rnum rl_mem_workq)
+
+__declspec(import, emem) struct rl_pkt_store rl_pkts;
+volatile __declspec(import, emem, addr40, aligned(sizeof(unsigned int))) uint8_t inpkt_buffer[RL_PKT_MAX_SZ * RL_PKT_STORE_COUNT] = {0};
+
+// ring head and tail on i25 or emem1
+volatile __emem_n(0) __declspec(import, addr40, aligned(512*1024*sizeof(unsigned int))) uint32_t rl_out_workq[512*1024] = {0};
+// _NFP_CHIPRES_ASM(.alloc_resource rl_out_workq_rnum emem0_queues+RL_OUT_RING_NUMBER global 1)
+//_NFP_CHIPRES_ASM(.init_mu_ring rl_mem_workq_rnum rl_mem_workq)
+
+__declspec(import, emem) struct rl_pkt_store rl_actions;
+volatile __declspec(import, emem, addr40, aligned(sizeof(unsigned int))) uint8_t rl_out_state_buffer[RL_DIMENSION_MAX * sizeof(tile_t) * RL_PKT_STORE_COUNT] = {0};
+
+volatile __declspec(import, emem) uint64_t really_really_bad;
+
[…]

### 2021-04-22 code/nfp/rl-test/rl.c

@@ -18,6 +18,7 @@
 #include "worker_config.h"
 
 #include "policy/policy_mem_define.h"
+
 #include "util.h"
 
 // ring head and tail on i25 or emem1
@@ -55,7 +56,7 @@ __export __emem __align(SA_TABLE_SZ) struct mem_lkup_cam32_16B_table_bucket_entr
 #define SA_ENTRIES 0x10
 CAMHT_DECLARE(state_action_map, SA_ENTRIES, uint64_t)
 
-__declspec(emem) struct state_action_pair state_action_pairs[SA_ENTRIES];
+__declspec(export, emem) struct state_action_pair state_action_pairs[SA_ENTRIES];
 
 #define REWARD_ENTRIES 0x10
 CAMHT_DECLARE(reward_map, REWARD_ENTRIES, union pad_tile)
@@ -74,7 +75,7 @@ volatile __declspec(shared) struct work local_ctx_work = {0};
 *
 * Returns length of tile coded list.
 */
-uint16_t tile_code(tile_t *state, __addr40 _declspec(emem) struct rl_config *cfg, uint32_t *output) {
+uint16_t tile_code(__addr40 _declspec(emem) tile_t *state, __addr40 _declspec(emem) struct rl_config *cfg, __addr40 __declspec(emem) uint32_t *output) {
 	uint16_t tiling_set_idx;
 	uint16_t out_idx = 0;
 
@@ -145,7 +146,7 @@ uint16_t tile_code(tile_t *state, __addr40 _declspec(emem) struct rl_config *cfg
 * Length written into act_list is guaranteed to be equal to cfg->num_actions.
 * Note, that act_list must be at least of size MAX_ACTIONS.
 */
-void action_preferences(uint32_t *tile_indices, uint16_t tile_hit_count, __addr40 _declspec(emem) struct rl_config *cfg, tile_t *act_list) {
+void action_preferences(__addr40 __declspec(emem) uint32_t *tile_indices, uint16_t tile_hit_count, __addr40 _declspec(emem) struct rl_config *cfg, __addr40 __declspec(emem) tile_t *act_list) {
 	enum tile_location loc = TILE_LOCATION_T1;
 	uint16_t i = 0;
 	uint16_t j = 0;
@@ -260,7 +261,7 @@ tile_t fat_select_key(struct key_source key_src, __addr40 __declspec(emem) tile_
[…]

### 2021-04-22 code/nfp/rl-test/worker_config_marker.h

@@ -1,8 +1,7 @@
 #ifndef _CONFIG_MARKER_H
 #define _CONFIG_MARKER_H
 #define WORK_ALLOC_STRAT ALLOC_FILL_HEAVY
-// #define WORK_ALLOC_RANDOMISE
-// #define WORK_ALLOC_STRAT ALLOC_CHUNK
 #define _RL_CONFIG_SOLO_REMOTE_WORK
-// #define WORKER_BARGAIN_BUCKET_SIMD
+// #define _RL_CONFIG_SOLO_LOCAL_WORK
+// #define _RL_CONFIG_SOLO_INDEP_RL_CORES
 #endif /* !_CONFIG_MARKER_H_ */

### 2021-04-22 paper/paper.tex
Draft 4 Dimitris.

@@
 \newcommand{\approachshort}{OPaL}
+\newcommand{\coopfw}{parallel}
+\newcommand{\Coopfw}{Parallel}
+\newcommand{\indfw}{single}
+\newcommand{\Indfw}{Single}
 
@@
 \Cref{fig:netro-arch} outlines our design and implementation on Netronome SmartNIC hardware in pursuit of this goal.
-We open-source our firmware design and control programs for the benefit of the community\footnote{\url{https://github.com/FelixMcFelix/pdp-rl-paper}---currently private...}.
+We open-source our firmware design and control programs for the benefit of the community.\footnote{\url{https://github.com/FelixMcFelix/pdp-rl-paper}---currently private...}
 We describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.
@@
 (FIXME: redo in illustrator?
-Show CTXes, PISA, single vs parallel)\label{fig:netro-arch}}
+Show CTXes, PISA, \Coopfw{} vs \Indfw{})\label{fig:netro-arch}}
 \end{figure}
@@
 Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
-Bitdepth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
+Bit depth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
 
@@
 		\includegraphics[keepaspectratio, width=\linewidth]{figures/single-core-arch-sketch}
-		\caption{Single (offline throughput-optimal).\label{fig:single-and-parallel:single}}
+		\caption{\Indfw{} (offline throughput-optimal).\label{fig:single-and-parallel:single}}
 	\end{subfigure}
@@
 		\includegraphics[keepaspectratio, width=\linewidth]{figures/parallel-arch-sketch}
-		\caption{Parallel (online-optimal).\label{fig:single-and-parallel:parallel}}
+		\caption{\Coopfw{} (online-optimal).\label{fig:single-and-parallel:parallel}}
 	\end{subfigure}
@@
 
-?? Do we need better/snappier names than just Single/Parallel? It's awfully dry.
-
-In the NFP architecture we use, physical cores (MEs) each have 8 hardware threads (contexts): all contexts on an ME share the same program and employ zero-cost context switches on I/O.
-Each core then provides 8 worker threads, though in the parallel case one must be dedicated to monitoring work acknowledgements.
+?? Do we need better/snappier names than just \Indfw/\Coopfw? It's awfully dry.
+
+In the NFP architecture we use, physical cores (MEs) each have 8 hardware threads (contexts): all contexts on an ME share the same program and employ zero-cost context switches triggered by I/O.
+Each core then provides 8 worker threads, though in the \indfw{} case one must be dedicated to monitoring work acknowledgements.
 Two main firmware models govern how the computation-intensive parts of these tasks (tile-coding, action preference list construction, policy updates) are carried out:
 \begin{description}
-	\item[Single (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each performs its work sequentially.
+	\item[\Indfw{} (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each performs its work sequentially.
 Computing an action list requires a \emph{read lock} on the policy.
@@
 Tile lists are stored for update computation.
-	\item[Parallel (\cref{fig:single-and-parallel:parallel})] Threads cooperate on processing state vectors, minimising latency.
+	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel})] Threads cooperate on processing state vectors, minimising latency.
 Workers have a fixed list of work items, while a master thread sends compute/update commands before awaiting worker completion.
@@
 \end{description}
-Each offers a different point of optimisation; if updates are disabled, then the \emph{single} model can maximise throughput, while the \emph{parallel} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
+Each offers a different point of optimisation; if updates are disabled, then the \emph{\indfw{}} model can maximise throughput, while the \emph{\coopfw{}} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
 
@@
 This extends to policy data, which may be imported from an offline pre-trained model via such packets, and exported via the NFP binary support package on the host machine.
-Some aspects must be chosen at compile time; bit-depth, work allocation strategy, and maximum policy/tiling/state sizes---these govern core operation or allocated memory regions.
-
-In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved DSCP value as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+Some aspects must be chosen at compile time; bit depth, work allocation strategy, and maximum policy/tiling/state sizes---these govern core operation or allocated memory regions.
+
+In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved (pool 2~\parencite{rfc2474}) DSCP value as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
 While we primarily use this to automate parser generation, this also allows for configuration packets to be received from only trusted hosts over the dataplane if needed via P4 rules.
@@
 The two compute models discussed here need not be homogeneously deployed.
-For instance, in a networked deployment a subset of \approachshort{} nodes could be \SI{32}{\bit} \emph{parallel} agents, training online, while most other nodes ran \SI{8}{\bit} \emph{single} designs to meet throughput guarantees.
+For instance, in a networked deployment a subset of \approachshort{} nodes could be \SI{32}{\bit} \emph{\coopfw{}} agents, training online, while most other nodes ran \SI{8}{\bit} \emph{\indfw{}} designs to meet throughput guarantees.
 The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
@@
 To achieve line-rate match-action table lookup, platforms like NFP use accelerated datastructures (e.g., DCFL~\parencite{DBLP:conf/infocom/TaylorT05}) which must be computed by the controller over the \emph{entire rule set}.
-Even through externs,
+Even through externs, directly adding new rules is neither feasible nor safe.
 
@@
 Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
+
 Consider a single, shared, read-only register between RL functional ...
@@
 
-\subsection{DDoS Defence}\label{sec:integ-1}
-?? Link to my own work
-?? Explain how every part may be substituted
-
-?? Bump-in-the-wire at edge nodes.
+\subsection{In-Network DDoS Defence}\label{sec:integ-1}
+Classical reinforcement learning has seen some recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
+
+?? Bump-in-the-wire deployment of SmartNICs at edge nodes.
+
+?? Place Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} on the datapath as the main source of state vectors.
+?? This resides in the P4 dataplane, and exists on-chip alongside the \approachshort{} agent.
+?? Use INDDoS~\parencite{tnms-ddos-victim-ident} as a reward function source---i.e., using the number of detected victims as a loss value.
+?? Still receive load measures from elsewhere in the network; this is now the element at highest risk of being outdated, but the least likely to vary significantly in response to individual actions.
 
@@
 ?? Remember the timed-random-selection; use the state coalescing trick indep discovered by you and those other guys.
-
-?? Can the P4 dataplane be used to reduce/increase the number of flows used as inputs? (YES)
-
-?? Place Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} on the datapath as the main source of state vectors.
-
-?? Use the TNSM work you reviewed~\parencite{tnms-ddos-victim-ident} as a reward function source.
+?? As such, the state collection mechanism prevents flows from being acted upon too often and ensures that the RL decision-making
+?? As we shall see, this far exceeds the number of decisions per second that a host based deployment could handle, though using these techniques is still key in preventing overloads.
+
+?? The P4 dataplane is then used to reduce/increase the number of flows used as inputs? (YES)
 
@@
 We compare these against floating point and quantised implementations running on a commodity host machine.
-This allows us to demonstrate the performance differences between the \emph{Single} and \emph{Parallel} configurations, particularly in how \emph{Single's} required policy locks impact throughput.
+This allows us to demonstrate the performance differences between the \emph{\indfw} and \emph{\coopfw} configurations, particularly in how \emph{\indfw}'s required policy locks impact throughput.
 We also report on the derived throughput measures accounting for the locking behaviour of each.
@@
 ?? NOT GOT OTHERS TO COMPARE AGAINST YET, JUST SELF.
-?? Mention versus DNN latency in passing here?
 
@@
 
-?? Not done yet: read these out from the IDE.
-
 \fakepara{The impact of bit depth}
@@
 \subsection{Results and Discussion}\label{sec:results}
+\fakepara{Raw inference and learning performance}
+\Cref{tab:lat-tput} shows that \Coopfw{} \approachshort{} achieves sub-\SI{50}{\micro\second} median latency, with \nth{99} percentile latencies less than \SI{1}{\micro\second} worse.
+As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, while the simpler \Indfw{} succeeds when policies are fixed/known.
+?? Crucially, our using many extra functional units to tackle this task
+
+?? Bitdepth is important for single item computation in the simpler \Indfw{} strategy.
+
+??
+We observed that \SIlist{8;16}{\bit} have worse action compute/latency, but better update processing time in \SI{16}{\bit}.
+Why?
+Throughput and batching of value reads only goes so far, native register size for ALU is \SI{32}{\bit}, so these wins are ``undone'' to some extent by the compiler extracting/shifting parts  for correctness.
+Wins are mainly due to serial part scaling with bitsize.
+
+?? Mention versus DNN latency in passing here?
+
+?? NOT GOT OTHERS TO COMPARE AGAINST YET, JUST SELF. NEED TO COMPARE AGAINST SOMEONE TO SHOW SUPREMACY
 
@@
 Put completion times first?
-Put multi first?}
+Put multi first?\label{tab:lat-tput}}
 	\resizebox{\linewidth}{!}{
@@
 
-\begin{figure}
-	\resizebox{1.0\linewidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-work}
-	}
-	\caption{?? More work means More time. ?? Max Parallelisation. ?? Bitdepth affects independent cores more. ?? latency version in appendix: \Cref{fig:vary-work-latency}}
-\end{figure}
+?? \Cref{fig:vary-core} shows how action+update compute time scales with available cores.
+?? Limited by software.
+?? Extra \emph{physical core} additions (rather than logical workers) very visible.
+?? 3 workers needed to process items faster than \Indfw{}, but 2 workers needed to bring \emph{latency} below \Indfw{}
+?? TAKEAWAY: resource demand tunable to meet latency demands for workload.
 
@@
 ??
-latency version in appendix: \Cref{fig:vary-core-latency}}
+latency version in appendix: \Cref{fig:vary-core-latency}.\label{fig:vary-core}}
 \end{figure}
 
-\fakepara{Raw inference and learning performance}
-TEST
+?? \Cref{fig:vary-work} shows how action+update compute time of any individual work item scales with the complexity of the work.
+?? Note: input vectors still have full size of 20 elements.
+?? Lower bit depth leads to a \SIrange{5}{10}{\micro\second} improvement in less complex policies, but these converge with policy complexity.
+?? Relate this to latency? \SI{32}{\bit} marginally ``wins out'' there..
+?? In latency if \Indfw{}, 32 == 8 < 16 (hmm).
+
+\begin{figure}
+	\resizebox{1.0\linewidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/vary-work}
+	}
+	\caption{?? More work means More time. ?? Max Parallelisation. ?? Bitdepth affects independent cores more. ?? latency version in appendix: \Cref{fig:vary-work-latency}.\label{fig:vary-work}}
+\end{figure}
 
 \fakepara{Work allocation}
-?? \Cref{fig:work-alloc-32}
+\Cref{fig:work-alloc-32} shows that our heuristic first-fit allocator (\emph{Balanced}) is key in achieving consistent sub-\SIlist{50;100}{\micro\second} latencies/processing times respectively.
+The trend is repeated for all bit depths, noting the latency differences described above.
+The constant difference between \emph{Action} and \emph{Update Prep} was observed to scale linearly with bit depth, matching storage and lookup work in the serial portion between parallel jobs (plots omitted).
+The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its ID, as batching work in contiguous chunks leads to some workers receiving exclusively high-dimensionality tile sets.
+The minor gap in performance lower bounds between the \emph{Random} and \emph{Balanced} allocators suggests that further optimisations can be made in this way.
+We expect that closing or exceeding this gap may require more complex modelling of NFP's hardware thread interactions, which lies far beyond the scope of in-NIC scheduling.
+Some additional complexity may be tolerated though---subject to code store limits---scheduling runs exactly once per configuration change, so does not impact per-action code.
 
@@
 	}
-	\caption{?? Choice of work strategy is pretty important I guess\label{fig:work-alloc-32}}
+	\caption{Action/update compute times in a \SI{32}{\bit} \emph{\Coopfw{}} agent under variable work schedulers.\label{fig:work-alloc-32}}
 \end{figure}
@@
 \fakepara{End-to-end RL latency}
-TEST
+?? TODO
 
 \fakepara{Co-existence with the dataplane}
-TEST
+?? TODO
 
 \fakepara{Resource requirements}
-?? \Cref{tab:resources}
+\Cref{tab:resources} demonstrates how \approachshort{} consumes shared and local memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
+As one program is installed per ME, these represents the minimum and maximum resource use we can command on a single island (i.e., without replacing P4 workers).
+We observe negligible costs on shared bulk storage ($\sim$\SI{2}{\mebi\byte}).
+The most significant costs arise due to policy data (\SI{405}{\kibi\byte} capacity shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
+Thread local storage in CLS for compute/register spilling scales with the number of MEs required, and with \emph{\Coopfw} this is far below the ``fair share'' (allowing co-existence with other asynchronous dataplane programs).
+Notably, \emph{\Indfw{}} imposes far heavier requirements on the limited per-island CLS, due to the extra resources needed to maintain, operate on, and cache tile lists for optimal processing.
 
 \begin{table}
-\caption{NFP memory use due to \approachshort{}.
-CLS and CTM are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM are shared between all NFP programs on a NIC.\label{tab:resources}}
+\caption{NFP memory use due to \approachshort{} using \numlist{1;4} MEs (\SI{32}{\bit}).
+CLS and CTM regions are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM regions are shared between all NFP programs on a NIC.\label{tab:resources}}
 \resizebox{\linewidth}{!}{
@@
 	& \multicolumn{1}{c}{\si{\mebi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} \\
-	\midrule Base P4 & 6794.73 & 88.47 & 0 & 0 & 858.28 & 10.48 & 0 & 0 & 0 & 0 \\
-	Single (1-core) & 6778.83 & 88.27 & 1737.91 & 18.86 & 1263.28 & 15.42 & 28.69 & 44.82 & 90 & 35.16 \\
-	Single (4-core) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-	Parallel (1-core) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-	Parallel (4-core) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+	\midrule Base P4 & 6776.67 & 88.24 & 268.52 & 2.91 & 858.28 & 10.48 & 0 & 0 & 0 & 0 \\
+	\Indfw(1) & 6778.77 & 88.27 & 1738.01 & 18.86 & 1263.28 & 15.42 & 23.50 & 36.72 & 90.00 & 35.16 \\
+	\Indfw(4) & 6778.79 & 88.27 & 1746.88 & 18.95 & 1263.28 & 15.42 & 49.00 & 76.56 & 90.00 & 35.16 \\
+	\Coopfw(1) & 6778.76 & 88.27 & 1742.51 & 18.91 & 1263.28 & 15.42 & 19.66 & 30.72 & 90.00 & 35.16 \\
+	\Coopfw(4) & 6778.76 & 88.27 & 1738.04 & 18.86 & 1263.28 & 15.42 & 31.82 & 49.71 & 90.00 & 35.16 \\
 	\bottomrule
@@
 ?? \Cref{fig:quant-acc}
+?? \SI{5}{\bit} mantissa suffices for $\ge$ \SI{90}{\percent} relative accuracy, so \SI{8}{\bit} is fine (1S + 2E + 5M), \SI{16}{\bit} is better.
 
@@
 
+?? NOTES: try to strengthen emphasis on ``other works can't do online learning in the data plane''. This is absolutely, \SI{100}{\percent} the killer feature.
+
 ?? A switch/NIC architecture like \textcite{DBLP:conf/hotnets/StephensAS18} would remove the need for hard-coded island-island relationships.
@@
 
+?? Future work: allow scaling over multiple islands.
+?? Future work: NetFPGA it.
+
 \begin{acks}
-	The authors would like to thank Rhys Simpson for his comments and discussions on SIMD-like optimisations.
+	The authors would like to thank Rhys Simpson for his comments and discussions on the soundness of SIMD-like optimisations.
 	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grant~\grantnum{gs-epsrc}{EP/N509668/1}].
@@
 \begin{appendices}
-	\section{Old ``What could we do``?}
+	\begin{figure}
+		\resizebox{1.0\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
+		}
+		\caption{Latency! varying work.\label{fig:vary-work-latency}}
+	\end{figure}
+
+	\begin{figure}
+		\resizebox{1.0\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
+		}
+		\caption{Latency! varying core count\label{fig:vary-core-latency}}
+	\end{figure}
+	\section{Old ``What could we do''?}
 	How do we evaluate this without looking into classification performance?
@@
 	This is application-dependent: see earlier discussions on rule batching and local extern/hash-table use.
-
-	\begin{figure}
-		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
-		}
-		\caption{Latency! varying work.\label{fig:vary-work-latency}}
-	\end{figure}
-
-	\begin{figure}
-		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
-		}
-		\caption{Latency! varying core count\label{fig:vary-core-latency}}
-	\end{figure}
 \end{appendices}

### 2021-04-23 code/marl-quant/marl/marl.py
A method to time stuff on hosts at last?

@@ -170,6 +170,8 @@ def marlExperiment(
 		quant_iter_start = None,
 
 		train_using_quant = None,
+
+		sarsa_pairs_to_export = None,
 	):
 
 	agent_classes = {

### 2021-04-23 code/marl-quant/marl/quant-perf-runner.py

@@ -1,0 +1,335 @@
+import numpy as np
+import os
+import platform
+from quantiser import *
+from sarsa import SarsaLearner, QLearner
+import time
+
+machine_name = platform.node()
+results_dir = "../../../results/host-rl-perf/"
+quantisers = [
+	None,
+	("i8", np.dtype("i1"), Quantiser.binary(5)),
+	("i16", np.dtype("i2"), Quantiser.binary(11)),
+	("i32", np.dtype("i4"), Quantiser.binary(24)),
+],
+
+iters_to_run = 5000
+
+# Hacked config code taken from marl.py...
+
+def marlParams(
+		linkopts = {
+			#"bw": 10,
+			"delay": 10,
+		},
+		n_teams = 1,
+
+		# per-team options
+		n_inters = 2,
+		n_learners = 3,
+		host_range = [2, 2],
+
+		calc_max_capacity = None,
+
+		P_good = 0.6,
+		good_range = [0, 1],
+		evil_range = [2.5, 6],
+		good_file = "../../data/pcaps/bigFlows.pcap",
+		bad_file = None,
+
+		topol = "tree",
+		ecmp_servers = 8,
+		ecmp_k = 4,
+
+		explore_episodes = 80000,
+		episodes = 1000,#100000
+		episode_length = 5000,#1000
+		separate_episodes = False,
+
+		max_bw = None,
+		pdrop_magnitudes = [0.1*n for n in xrange(10)],
+
+		alpha = 0.05,
+		epsilon = 0.3,
+		discount = 0,
+		break_equal = None,
+
+		algo = "sarsa",
+		trace_decay = 0.0,
+		trace_threshold = 0.0001,
+		use_path_measurements = True,
+
+		model = "tcpreplay",
+		submodel = None,
+		rescale_opus = False,
+		mix_model = None,
+
+		use_controller = False,
+		moralise_ips = True,
+
+		dt = 0.001,
+
+		# These should govern tc on the links at various points in the network.
+		#
+		# Final hop is just what you'd think, old_style governs
+		# whether sender rates are mediated by tcpreplay or by tc on the link
+		# over from said host.
+		old_style = False,
+		force_host_tc = False,
+		protect_final_hop = True,
+
+		with_ratio = False,
+		override_action = None,
+		manual_early_limit = None,
+		estimate_const_limit = False,
+
+		rf = "ctl",
+
+		rand_seed = 0xcafed00d,
[…]

### 2021-04-23 code/marl-quant/marl/sarsa.py

@@ -5,6 +5,7 @@ import random
 from spf import *
 import sys
 import tilecoding.representation as r
+import time
 
 # FIXME: change update et al to use quantiser.
 
@@ -28,8 +29,10 @@ class SarsaLearner:
 				increase_fixed_math_alpha_if_narrowed=True,
 				always_include_bias=True,
 				AcTrans=MarlMachine,
-				quantiser=None):
+				quantiser=None,
+				quantiser_dt=None):
 		self.quantiser = quantiser
+		self.quantiser_dt = quantiser_dt
 
 		state_range = [
 			[0 for i in xrange(vec_size)] + extended_mins,
@@ -97,7 +100,7 @@ class SarsaLearner:
 				if self.quantiser is None:
 					self.values[tile] = np.full(len(self.actions), float(self.default_q))
 				else:
-					self.values[tile] = np.full(len(self.actions), float(self.default_q), dtype=np.dtype(int))
+					self.values[tile] = np.full(len(self.actions), float(self.default_q), dtype=self.quantiser_dt)
 
 	def _get_state_values(self, state):
 		self._ensure_state_vals_exist(state)
@@ -201,6 +204,7 @@ class SarsaLearner:
 		# FIXME: reconcile effect of narrowings upon learning rate when using 'intended' maths.
 		action_narrowing=None,
 		update_narrowing=None,
+		time_write_space=None
 	):
 		(last_state, last_action, last_z) = (self.last_act if subs_last_act is None else subs_last_act)
 
@@ -223,6 +227,9 @@ class SarsaLearner:
 		# but we need ALL values to be made available when we are handling "transitions"
 		# between narrowed/unnarrowed regions of the markov chain.
 		(new_action, new_values, argmax_values, ac_values) = self.select_action(state, action_narrowing)
+
+		if time_write_space is not None:
+			time_write_space["time"] = time.time()
 
 		next_vals = argmax_values if self._argmax_in_dt else new_values
 		argmax_chosen = np.all(new_values == argmax_values)
@@ -302,7 +309,7 @@ class SarsaLearner:
 
 		return tuple(out)
 
-	def as_quantised(self, quantiser):
+	def as_quantised(self, quantiser, dt=np.dtype(int)):
 		out = copy.deepcopy(self)
 
 		# remake tile coding
@@ -321,7 +328,7 @@ class SarsaLearner:
 		# update all entries of values?
 		nv = {}
[…]

### 2021-04-28 paper/paper.tex
A method to time stuff on hosts at last?
More principled t'put for hosts.

@@
 \section{Design and Implementation}\label{sec:design}
+?? Can we say ``in-situ'' anywhere? I like that term.
+
 Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{} (\approach)}.
@@
 Policy sizes are set to those matching a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dimension state vectors, a bias tile and 16 full tiling sets (7$\times$1-dim, 8$\times$2-dim, 1$\times$4-dim), 8 tilings per dimension set, 6 tiles per dimension, choosing between 10 actions.
+Although their more successful \emph{SPF} agent design uses 4 actions, we choose this to investigate the performance of larger agent designs.
 
@@
 \caption{NFP memory use due to \approachshort{} using \numlist{1;4} MEs (\SI{32}{\bit}).
-CLS and CTM regions are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM regions are shared between all NFP programs on a NIC.\label{tab:resources}}
+CLS and CTM regions are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM regions are shared between all NFP programs on a NIC.\label{tab:resources} ??NEEDS REWORKED}
 \resizebox{\linewidth}{!}{

### 2021-04-28 code/marl-quant/marl/qt-tput.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+SCRIPT="./quant-perf-tput-task.py"
+
+for concurrent_task_limit in `seq 1 32`;
+do
+	for quantiser_choice in `seq 0 3`;
+	do
+		for do_update in `seq 0 1`;
+		do
+			for attempt in `seq 0 0`;
+			do
+				for task_i in $(seq 0 $(expr $(concurrent_task_limit) - 1));
+				do
+					python2 $SCRIPT $concurrent_task_limit $task_i $quantiser_choice $do_update $attempt &
+				done
+
+				wait
+			done
+		done
+	done
+done
+

### 2021-04-28 code/marl-quant/marl/quant-perf-tput-task.py

@@ -3,11 +3,12 @@ import os
 import platform
 from quantiser import *
 from sarsa import SarsaLearner, QLearner
+import sys
 from spf import *
 import time
 
 machine_name = platform.node()
-results_dir = "../../../results/host-rl-perf/"
+results_dir = "../../../results/host-rl-perf-tput/raw"
 pquantisers = [
 	None,
 	("i8", np.dtype("i1"), Quantiser.binary(5)),
@@ -15,8 +16,9 @@ pquantisers = [
 	("i32", np.dtype("i4"), Quantiser.binary(24)),
 ]
 
-warmup_len = 1000
-iters_to_run = 5000
+warmup_len = 2.0
+iters_to_run = 10.0
+cooldown_len = 2.0
 
 # Hacked config code taken from marl.py...
 
@@ -227,124 +229,151 @@ def marlParams(
 
 	return sarsaParams
 
-for possible_quantiser in pquantisers:
-	params = {
-		#"P_good": 1.0,
-		"n_teams": 2,
-		"n_inters": 2,
-		"n_learners": 3,
+args = sys.argv()
+task_ct = int(args[1])
+task_i = int(args[2])
+quantiser_i = int(args[3])
+act_or_update = int(args[4])
+repetition = int(args[5])
 
-		"alpha": 0.05,
-		"epsilon": 0.2,
-		"discount": 0.0,
-		"dt": 0.05,
+possible_quantiser = pquantisers[quantiser_i]
 
-		"ecmp_servers": 2,
+params = {
+	#"P_good": 1.0,
+	"n_teams": 2,
+	"n_inters": 2,
+	"n_learners": 3,
 
-		"explore_episodes": 0.8,
-		"episodes": 10,
-		"episode_length": 10000,
-		"separate_episodes": True,
+	"alpha": 0.05,
+	"epsilon": 0.2,
+	"discount": 0.0,
+	"dt": 0.05,
 
-		"rf": "marl",
-		"use_controller": True,
-		"actions_target_flows": True,
-		"trs_maxtime": 0.001,
+	"ecmp_servers": 2,
 
-		"split_codings": True,
-		"feature_max": 20,
+	"explore_episodes": 0.8,
+	"episodes": 10,
+	"episode_length": 10000,
+	"separate_episodes": True,
 
-		# new shiny quantisation stuff.
-		"do_quant_testing": True,
-		"quantisers": [possible_quantiser],
[…]

### 2021-04-28 code/marl-quant/marl/sarsa.py

@@ -204,7 +204,8 @@ class SarsaLearner:
 		# FIXME: reconcile effect of narrowings upon learning rate when using 'intended' maths.
 		action_narrowing=None,
 		update_narrowing=None,
-		time_write_space=None
+		time_write_space=None,
+		do_not_update=False
 	):
 		(last_state, last_action, last_z) = (self.last_act if subs_last_act is None else subs_last_act)
 
@@ -227,6 +228,9 @@ class SarsaLearner:
 		# but we need ALL values to be made available when we are handling "transitions"
 		# between narrowed/unnarrowed regions of the markov chain.
 		(new_action, new_values, argmax_values, ac_values) = self.select_action(state, action_narrowing)
+
+		if do_not_update:
+			return (new_action, ac_values, None)
 
 		if time_write_space is not None:
 			time_write_space["time"] = time.time()

### 2021-04-28 code/marl-quant/marl/qt-tput.sh
Fixes to run on beast2

@@ -9,7 +9,7 @@ do
 		do
 			for attempt in `seq 0 0`;
 			do
-				for task_i in $(seq 0 $(expr $(concurrent_task_limit) - 1));
+				for task_i in $(seq 0 $(expr $concurrent_task_limit - 1));
 				do
 					python2 $SCRIPT $concurrent_task_limit $task_i $quantiser_choice $do_update $attempt &
 				done

### 2021-04-28 code/marl-quant/marl/quant-perf-tput-task.py

@@ -8,7 +8,7 @@ from spf import *
 import time
 
 machine_name = platform.node()
-results_dir = "../../../results/host-rl-perf-tput/raw"
+results_dir = "../../../results/host-rl-perf-tput/raw/"
 pquantisers = [
 	None,
 	("i8", np.dtype("i1"), Quantiser.binary(5)),
@@ -229,7 +229,7 @@ def marlParams(
 
 	return sarsaParams
 
-args = sys.argv()
+args = sys.argv
 task_ct = int(args[1])
 task_i = int(args[2])
 quantiser_i = int(args[3])
@@ -314,22 +314,14 @@ do_not_update = act_or_update == 0
 while t < end:
 	t_before = time.time()
 
-	s1 = q_states[i % len(q_states)]
+	s1 = learner.to_state_quanted(q_states[i % len(q_states)])
 
-	# this can be assumed to be done.
-	s2 = learner.to_state(s2)
-	last_pair_storage[i] = (s2, a, None)
-
-	last_act = None
+	last_act = (0, 0, None)
 
 	if not do_not_update:
 		last_act = coded_q_states[(i + 1) % len(q_states)]
 
 	# ------------
-
-	# do the state conversion work in here
-	# this includes tile coding
-	s1 = learner.to_state_quanted(qs1)
 
 	(new_ac, vals, new_z) = learner.update(
 		s1,
@@ -371,9 +363,9 @@ if act_or_update == 0:
 else:
 	name_part = "UpdateAll"
 
-for (name_part, data_to_write) in files:
-	# throughput can just be measured by counting rows (lmao)
-	outpath = "{}{}.{}.{}.{}({}).{}.dat".format(results_dir, outname, machine_name, name_part, task_ct, task_i, repetition)
-	with open(outpath, "w") as of:
-		for val in timing_measures:
-			of.write("{:.2f}\n".format(val))
+# throughput can just be measured by counting rows (lmao)
+outpath = "{}{}.{}.{}.{}({}).{}.dat".format(results_dir, outname, machine_name, name_part, task_ct, task_i, repetition)
+with open(outpath, "w") as of:
+	for val in timing_measures:
+		of.write("{:.2f}\n".format(val))
+

### 2021-04-28 code/marl-quant/marl/sarsa.py

@@ -212,8 +212,9 @@ class SarsaLearner:
 		# because of updates in parallel, we need to grab the current values.
 		# The update depends not on the value they once had, but only on the *current value* and the target (R).
 		# Otherwise, we're moving from an old start to the new target...
-		all_tile_action_vals = self._get_state_values(last_state)
-		last_values = np.array([av[last_action] for av in all_tile_action_vals])
+		if not do_not_update:
+			all_tile_action_vals = self._get_state_values(last_state)
+			last_values = np.array([av[last_action] for av in all_tile_action_vals])
 
 		## slight complication -- interaction between last & next?
 		# i.e. full -> reduced -> reduced -> full

### 2021-04-30 results/host-rl-perf-tput/process.py
Latency/Throughputs for host execution---processed.

@@ -1,0 +1,73 @@
+import numpy as np
+import scipy.stats as spss
+
+files_dir = "raw/"
+const_secs = 10.0
+
+bitdepths = ["i8", "i16", "i32", "f32"]
+machines = [
+	"perlman",
+	"beast2",
+	"beast1",
+]
+act_modes = ["ComputeAndWriteout", "UpdateAll"]
+trials = 10
+thread_max = 32
+
+for bd in bitdepths:
+	for machine in machines:
+		for act_mode in act_modes:
+			print("Working on {}-{}-{}".format(bd, machine, act_mode))
+			# out format: mean_tput, std_tput, med_time, 99_time, 9999_time, mean_time, std_time
+			expt_rows = []
+
+			last_all_tputs = None
+
+			for thread_ct in range(1, thread_max+1):
+				print("\t{} threads".format(thread_ct))
+				all_tputs = []
+				all_items = []
+
+				for trial_i in range(trials):
+					items = []
+					for thread_i in range(thread_ct):
+						fname = "{}{}.{}.{}.{}({}).{}.dat".format(
+							files_dir,
+							bd,
+							machine,
+							act_mode,
+							thread_ct,
+							thread_i,
+							trial_i,
+						)
+						with open(fname, "r") as of:
+							for line in of:
+								items.append(float(line.strip()))
+
+					all_tputs.append(float(len(items))/const_secs)
+					all_items += items
+
+				if last_all_tputs is None:
+					p_value = 0.0
+				else:
+					(_t_val, p_value) = spss.stats.ttest_ind(all_tputs, last_all_tputs, equal_var=False)
+
+				expt_rows.append([
+					np.mean(all_tputs),
+					np.std(all_tputs),
+					np.median(all_items),
+					np.percentile(all_items, 99.0),
+					np.percentile(all_items, 99.99),
+					np.mean(all_items),
+					np.std(all_items),
+					p_value,
+				])
+
+				last_all_tputs = all_tputs
+
+			# writeout
+			outfile = "SUMMARY.{}.{}.{}.dat".format(bd, machine, act_mode)
+			with open(outfile, "w") as of:
+				for row in expt_rows:
+					format_row = ["{:.3f}".format(x) for x in row]
+					of.write(" ".join(format_row) + "\n")

### 2021-04-30 tables/src/latency-tput-temp.py

@@ -51,7 +51,7 @@ for bits in bitdepths:
 
 			stats = [
 				np.median(indiv_data),
-				np.percentile(indiv_data, 99),
+				np.percentile(indiv_data, 99.99),
 			]
 
 			latencies.append("{:.3f}".format(stats[0]))
@@ -66,7 +66,7 @@ for bits in bitdepths:
 
 			indiv_tputs /= 1.0e3
 
-			tputs.append("{:.3f} \\pm {:.3f}".format(np.mean(indiv_tputs), np.var(indiv_tputs)))
+			tputs.append("{:.3f} \\pm {:.3f}".format(np.mean(indiv_tputs), np.std(indiv_tputs)))
 
 		for os in tputs + latencies:
 			out += " & {}".format(os)

### 2021-04-30 tables/src/latency-tput.py

@@ -66,7 +66,7 @@ for bits in bitdepths:
 
 			indiv_tputs /= 1.0e3
 
-			tputs.append("{:.3f} \\pm {:.3f}".format(np.mean(indiv_tputs), np.var(indiv_tputs)))
+			tputs.append("{:.3f} \\pm {:.3f}".format(np.mean(indiv_tputs), np.std(indiv_tputs)))
 
 		for os in tputs + latencies:
 			out += " & {}".format(os)

### 2021-05-03 paper/paper.tex
Modifying runner now that the hash table apparently works?

@@
 
+?? Probably mention the idea of including "state keys" and "reward keys" in the input state vector.
+
 In the NFP architecture we use, physical cores (MEs) each have 8 hardware threads (contexts): all contexts on an ME share the same program and employ zero-cost context switches triggered by I/O.
@@
 Policy sizes are set to those matching a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dimension state vectors, a bias tile and 16 full tiling sets (7$\times$1-dim, 8$\times$2-dim, 1$\times$4-dim), 8 tilings per dimension set, 6 tiles per dimension, choosing between 10 actions.
-Although their more successful \emph{SPF} agent design uses 4 actions, we choose this to investigate the performance of larger agent designs.
+Although their more successful \emph{SPF} agent design uses 3 actions, we choose this to investigate the performance of larger agent designs.
 
@@
 
-\fakepara{The impact of bit depth}
-We further investigate how bit depth affects the accuracy of policy by varying the number of bits used to represent the mantissa of action values and input state.
+%\fakepara{The impact of bit depth}
+%We further investigate how bit depth affects the accuracy of policy by varying the number of bits used to represent the mantissa of action values and input state.
 
@@
 	\resizebox{\linewidth}{!}{
-		\expandableinput ../tables/build/latency-tput
+		\expandableinput ../tables/build/latency-tput-temp
 	}
@@
 
-\fakepara{The impact of bit depth}
-?? \Cref{fig:quant-acc}
-?? \SI{5}{\bit} mantissa suffices for $\ge$ \SI{90}{\percent} relative accuracy, so \SI{8}{\bit} is fine (1S + 2E + 5M), \SI{16}{\bit} is better.
-
-\begin{figure}
-	\includegraphics{../plots/build/marl-quant/accuracy-binary}
-	\caption{Normalised accuracy of a converted, pre-trained floating-point tile-coded policy after conversion to $32Qn$ fixed-point.}\label{fig:quant-acc}
-\end{figure}
+%\fakepara{The impact of bit depth}
+%?? \Cref{fig:quant-acc}
+%?? \SI{5}{\bit} mantissa suffices for $\ge$ \SI{90}{\percent} relative accuracy, so \SI{8}{\bit} is fine (1S + 2E + 5M), \SI{16}{\bit} is better.
+%
+%\begin{figure}
+%	\includegraphics{../plots/build/marl-quant/accuracy-binary}
+%	\caption{Normalised accuracy of a converted, pre-trained floating-point tile-coded policy after conversion to $32Qn$ fixed-point.}\label{fig:quant-acc}
+%\end{figure}
 

### 2021-05-03 code/control/src/setup.rs

@@ -135,4 +135,11 @@ impl<T: Tile> KeySource<T> {
 			KeySource::Value(v) => v.wideint(),
 		}
 	}
+
+	pub fn field(&self) -> Option<i32> {
+		match self {
+			KeySource::Field(f) => Some(f),
+			_ => None,
+		}
+	}
 }

### 2021-05-03 code/rl-perf-tester/src/experiment.rs

@@ -421,6 +421,18 @@ pub fn generate_state<T: Tile + Clone>(setup: &Setup<T>, rng: &mut impl RngCore)
 		out.push(T::from_int(sample))
 	}
 
+	if let Some(f) = setup.state_key.field() {
+		if let Some(valid) = out.get_mut(f as usize) {
+			*valid = T::from_int(0);
+		}
+	}
+
+	if let Some(f) = setup.reward_key.field() {
+		if let Some(valid) = out.get_mut(f as usize) {
+			*valid = T::from_int(1);
+		}
+	}
+
 	out
 	//vec![T::from_int(0); setup.n_dims as usize]
 }

### 2021-05-03 code/nfp/rl-test/packet/packet.c
Bigger hashtables, working memcpys, Papa John's!

@@ -434,9 +434,9 @@ void state_packet(
 	// );
 
 	tc_count = tile_code((__declspec(emem) __addr40 tile_t *)pkt->packet_payload, cfg, tc_indices);
-	global_tc_count = tc_count;
+	/*global_tc_count = tc_count;
 
-	/*for (i=0; i<tc_count; i++) {
+	for (i=0; i<tc_count; i++) {
 		global_tc_indices[i] = tc_indices[i];
 	}*/
 
@@ -561,9 +561,9 @@ void state_packet(
 		// 	(void*)tc_indices, 0,
 		// 	tc_count * sizeof(tile_t)
 		// );
-		memcpy_mem40_mem40_al8(
-			(__declspec(emem) __addr40 uint64_t *) &(state_action_pairs[state_found].tiles[0]),
-			(__declspec(emem) __addr40 uint64_t *) tc_indices,
+		memcpy_emem_ctm_al8(
+			(__declspec(emem) uint64_t *) &(state_action_pairs[state_found].tiles[0]),
+			(__declspec(emem) uint64_t *) tc_indices,
 			tc_count * sizeof(uint32_t)
 		);
 	}

### 2021-05-03 code/nfp/rl-test/rl-prims.c

@@ -7,8 +7,65 @@ __intrinsic void memcpy_mem40_mem40_al8(
     uint64_t  p= (uint64_t)dest;
     uint64_t  q= (uint64_t)src;
 
-    while (n >= 8) {
-        *(__mem uint64_t *) p = *(__mem uint64_t *) q;
+    while (n > 0) {
+        *(__emem __addr40 uint64_t *) p = *(__emem __addr40 uint64_t *) q;
+
+	    p += 8;
+	    q += 8;
+
+        n -= 8;
+    }
+}
+
+__intrinsic void memcpy_emem_mem40_al8(
+	__declspec(emem) uint64_t *dest,
+	__declspec(emem) __addr40 uint64_t *src,
+	int32_t n
+)
+{
+    uint64_t  p= (uint64_t)dest;
+    uint64_t  q= (uint64_t)src;
+
+    while (n > 0) {
+        *(__emem uint64_t *) p = *(__mem __addr40 uint64_t *) q;
+
+	    p += 8;
+	    q += 8;
+
+        n -= 8;
+    }
+}
+
+__intrinsic void memcpy_mem40_ctm40_al8(
+	__declspec(emem) __addr40 uint64_t *dest,
+	__declspec(ctm) __addr40 uint64_t *src,
+	int32_t n
+)
+{
+    uint64_t  p= (uint64_t)dest;
+    uint64_t  q= (uint64_t)src;
+
+    while (n > 0) {
+        *(__emem __addr40 uint64_t *) p = *(__ctm __addr40 uint64_t *) q;
+
+	    p += 8;
+	    q += 8;
+
+        n -= 8;
+    }
+}
+
+__intrinsic void memcpy_emem_ctm_al8(
+	__declspec(emem) uint64_t *dest,
+	__declspec(ctm) uint64_t *src,
+	int32_t n
+)
+{
+    uint64_t  p= (uint64_t)dest;
+    uint64_t  q= (uint64_t)src;
+
+    while (n > 0) {
+        *(__emem uint64_t *) p = *(__ctm uint64_t *) q;
 
 	    p += 8;
 	    q += 8;
@@ -166,6 +223,8 @@ __intrinsic void update_action_preferences(__addr40 _declspec(emem) uint32_t *ti
 	uint16_t j = 0;
 	uint16_t act_count = cfg->num_actions;
 
+	// really_really_bad = 0;
+
 	for (i = 0; i < tile_hit_count; i++) {
[…]

### 2021-05-03 code/nfp/rl-test/rl.c

@@ -1,5 +1,6 @@
 #include <memory.h>
 #include <nfp.h>
+#include <nfp/cls.h>
 #include <nfp/me.h>
 //#include <nfp_mem_ring.h>
 //#include <nfp_mem_workq.h>
@@ -26,7 +27,7 @@ volatile __emem_n(0) __declspec(export, addr40, aligned(512*1024*sizeof(unsigned
 _NFP_CHIPRES_ASM(.alloc_resource rl_mem_workq_rnum emem0_queues+RL_RING_NUMBER global 1)
 //_NFP_CHIPRES_ASM(.init_mu_ring rl_mem_workq_rnum rl_mem_workq)
 
-__declspec(export, emem) struct rl_pkt_store rl_pkts;
+__declspec(export, emem, aligned(sizeof(uint64_t))) struct rl_pkt_store rl_pkts;
 volatile __declspec(export, emem, addr40, aligned(sizeof(uint64_t))) uint8_t inpkt_buffer[RL_PKT_MAX_SZ * RL_PKT_STORE_COUNT] = {0};
 
 // ring head and tail on i25 or emem1
@@ -34,7 +35,7 @@ volatile __emem_n(0) __declspec(export, addr40, aligned(512*1024*sizeof(unsigned
 _NFP_CHIPRES_ASM(.alloc_resource rl_out_workq_rnum emem0_queues+RL_OUT_RING_NUMBER global 1)
 //_NFP_CHIPRES_ASM(.init_mu_ring rl_mem_workq_rnum rl_mem_workq)
 
-__declspec(export, emem) struct rl_pkt_store rl_actions;
+__declspec(export, emem, aligned(sizeof(uint64_t))) struct rl_pkt_store rl_actions;
 volatile __declspec(export, emem, addr40, aligned(sizeof(uint64_t))) uint8_t rl_out_state_buffer[ALIGNED_OUT_STATE_SZ * RL_PKT_STORE_COUNT] = {0};
 
 volatile __declspec(export, emem) uint64_t really_really_bad = 0;
@@ -53,7 +54,7 @@ __export __emem __align(SA_TABLE_SZ) struct mem_lkup_cam32_16B_table_bucket_entr
 // Need to use this due to size limits on struct in CAMHT def, even
 // if key is different.
 // #define SA_ENTRIES 0x20000
-#define SA_ENTRIES 0x10
+#define SA_ENTRIES 0x1000
 CAMHT_DECLARE(state_action_map, SA_ENTRIES, uint64_t)
 
[…]

### 2021-05-03 code/nfp/rl-test/worker_config_marker.h

@@ -1,5 +1,6 @@
 #ifndef _CONFIG_MARKER_H
 #define _CONFIG_MARKER_H
-#define _RL_CONFIG_SOLO_REMOTE_WORK
+// #define _RL_CONFIG_SOLO_REMOTE_WORK
+#define _RL_CONFIG_SOLO_INDEP_RL_CORES
 #define WORK_ALLOC_STRAT ALLOC_FILL_HEAVY
 #endif /* !_CONFIG_MARKER_H_ */

### 2021-05-07 paper/paper.tex
Architecture figures.

@@
 \newcommand{\approachshort}{OPaL}
-\newcommand{\coopfw}{parallel}
-\newcommand{\Coopfw}{Parallel}
-\newcommand{\indfw}{single}
-\newcommand{\Indfw}{Single}
+\newcommand{\Coopfw}{CoOp}
+\newcommand{\coopfw}{\Coopfw}
+\newcommand{\Indfw}{Ind}
+\newcommand{\indfw}{\Indfw}
 
@@
 	\centering
-	\resizebox{1.2\linewidth}{!}{\begin{tikzpicture}
-			\node (sw2) {};
-
-			\node[fill=white!80!uofgcobalt, draw=black, minimum height=5.5cm, minimum width=2.9cm, below right= 0.1 of sw2.north west](p4-master){};
-			\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Datapath};
-			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.8] (p4-rules) at (p4-master.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};
-			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n1) at (p4-rules.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Extern-based Action\end{varwidth}};
-			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n2) at (n1.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
-			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (oswlc) at (n2.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};
-			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n3) at (oswlc.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
-			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n4) at (n3.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Network Telemetry\end{varwidth}};
-
-			\node (pkt-text) at ($(p4-master.north) + (0.0,0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet In\end{varwidth}};
-
-			\node (pkt-text-out) at ($(p4-master.south) + (0.0,-0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet Out\end{varwidth}};
-
-			\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (5, -1)$){};
-			\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};
-
-			\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
-			\node[below right, inner sep=2pt] at (wq-out.north west) {\small EMEM};
-			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Input Queue\end{varwidth}};
-
-			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (loop) at ($(wq-out.east) + (1.5,-0.5)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Control ME\end{varwidth}};
-
-			\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-act) at ($(loop.north) + (0.0,1.5)$){};
-			\node[below right, inner sep=2pt] at (wq-act.north west) {\small EMEM};
-			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-actin) at (wq-act.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Action Queue\end{varwidth}};
-
-			\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-cls-out) at ($(loop.south) + (0,-1.5)$){};
-			\node[below right, inner sep=2pt] at (policy-cls-out.north west) {\small CLS};
-			\node[above right, inner sep=2pt] at (policy-cls-out.north west) {\small Policy};
-			\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-ctm-out) at ($(policy-cls-out.south) + (0,-0.25)$){};
-			\node[below right, inner sep=2pt] at (policy-ctm-out.north west) {\small CTM};
-			\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-emem-out) at ($(policy-ctm-out.south) + (0,-0.25)$){};
-			\node[below right, inner sep=2pt] at (policy-emem-out.north west) {\small EMEM};
-
-			\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t1) at (policy-cls-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T1\end{varwidth}};
-			\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t2) at (policy-ctm-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T2\end{varwidth}};
-			\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t3) at (policy-emem-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T3\end{varwidth}};
-
-			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-1) at ($(wq-out.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};
-
-			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-2) at ($(po-me-1.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};
-
-			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-3) at ($(po-me-2.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};
-
-			\draw[->] (oswlc.east)  to node[midway, above, sloped] {\tiny Control, Requests} (wq-in.west);
-			\draw[->] (n4.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
-			\draw[<->, dotted] (wq-in.east)  to (loop.west);
-			\draw[->, dotted] (loop.south) to (policy-cls-out);
-
-			\draw[->] (loop.north)  to node[midway, above, sloped] {\tiny Actions} (wq-actin.south);
-
-			\draw[->] (wq-actin.west)  to node[midway, above, sloped] {\tiny Actions} (n1.east);
-
-			\draw[<->, dotted] (po-me-1.east) to (policy-cls-out.west);
-			\draw[<->, dotted] (po-me-2.east) to (policy-ctm-out.west);
-			\draw[<->, dotted] (po-me-3.east) to (policy-emem-out.west);
-
-			\draw[->] (pkt-text)  to  (p4-master.north);
-			\draw[->] (p4-master.south)  to  (pkt-text-out);
-
-			\draw[->] (loop) to node[midway, above, sloped] {\tiny NN} (po-me-1);
-			\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NN} (po-me-2);
-			\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NN} (po-me-3);
-			\draw[->] (po-me-3) to [out=-45,in=0,looseness=2] node[midway, above, sloped] {\tiny NN} (loop);
-	\end{tikzpicture}}
+	\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/arch-with-p4}
 	\caption{Architecture for an \approachshort{} agent on Netronome hardware, using 4 cores for parallel policy processing.
@@
 
-\begin{figure}
-	\begin{subfigure}{0.3\linewidth}
+\begin{figure*}
 	\centering
-		\includegraphics[keepaspectratio, width=\linewidth]{figures/single-core-arch-sketch}
+	\begin{subfigure}{0.45\linewidth}
+		\centering
+		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/ind}
 		\caption{\Indfw{} (offline throughput-optimal).\label{fig:single-and-parallel:single}}
@@
 		\centering
-		\includegraphics[keepaspectratio, width=\linewidth]{figures/parallel-arch-sketch}
+		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/coop}
 		\caption{\Coopfw{} (online-optimal).\label{fig:single-and-parallel:parallel}}
@@
 	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to latency/throughput needs. ?? MAKE ME FOR REAL \label{fig:single-and-parallel}}
-\end{figure}
+\end{figure*}
 
@@
 
+\begin{figure}
+	\resizebox{1.0\linewidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
+	}
+	\caption{Latency! varying work.\label{fig:vary-work-latency}}
+\end{figure}
+
+\begin{figure}
+	\resizebox{1.0\linewidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
+	}
+	\caption{Latency! varying core count\label{fig:vary-core-latency}}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\begin{subfigure}{0.45\linewidth}
+		\resizebox{1.0\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/latency-cumul}
+		}
+		\caption{Just Pars.}
+	\end{subfigure}
+	\begin{subfigure}{0.45\linewidth}
+		\resizebox{1.0\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/latency-cumul-broad}
+		}
+		\caption{Wider View.}
+	\end{subfigure}
+	\caption{BigCap}
+\end{figure}
+
 \fakepara{Work allocation}
@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}{@{}cSSSSSSSSSS@{}}
+\begin{tabular}{
+		@{}c
+		S[table-format=4.2] S[table-format=2.2]
+		S[table-format=4.2] S[table-format=2.2]
+		S[table-format=4.2] S[table-format=2.2]
+		S[table-format=2.2] S[table-format=2.2]
+		S[table-format=3.2] S[table-format=2.2]
+		@{}
+	}
 	\toprule Firmware & \multicolumn{2}{c}{EMEM} & \multicolumn{2}{c}{EMEM Cache} & \multicolumn{2}{c}{IMEM} & \multicolumn{2}{c}{i5.CLS} & \multicolumn{2}{c}{i5.CTM}\\
 	& \multicolumn{1}{c}{\si{\mebi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} \\
-	\midrule Base P4 & 6776.67 & 88.24 & 268.52 & 2.91 & 858.28 & 10.48 & 0 & 0 & 0 & 0 \\
-	\Indfw(1) & 6778.77 & 88.27 & 1738.01 & 18.86 & 1263.28 & 15.42 & 23.50 & 36.72 & 90.00 & 35.16 \\
-	\Indfw(4) & 6778.79 & 88.27 & 1746.88 & 18.95 & 1263.28 & 15.42 & 49.00 & 76.56 & 90.00 & 35.16 \\
-	\Coopfw(1) & 6778.76 & 88.27 & 1742.51 & 18.91 & 1263.28 & 15.42 & 19.66 & 30.72 & 90.00 & 35.16 \\
-	\Coopfw(4) & 6778.76 & 88.27 & 1738.04 & 18.86 & 1263.28 & 15.42 & 31.82 & 49.71 & 90.00 & 35.16 \\
+	\midrule Base P4 & 6776.67 & 88.24 & 268.52 & 2.91 & 858.28 & 10.48 & 0.00 & 0.00 & 0.00 & 0.00 \\
+	\Indfw(1) & 6780.21 & 88.28 & 2541.07 & 27.57 & 1263.28 & 15.42 & 24.75 & 38.67 & 94.25 & 36.82 \\
+	\Indfw(4) & 6780.21 & 88.28 & 2545.32 & 27.62 & 1263.28 & 15.42 & 51.18 & 79.97 & 107.00 & 41.80 \\
+	\Coopfw(1) & 6779.12 & 88.27 & 1769.83 & 19.20 & 1263.28 & 15.42 & 20.25 & 31.64 & 90.00 & 35.16 \\
+	\Coopfw(4) & 6779.12 & 88.27 & 1769.84 & 19.20 & 1263.28 & 15.42 & 31.91 & 49.85 & 90.00 & 35.16 \\
 	\bottomrule
@@
 
+??
+Simialr HW has been used to get NN \emph{inference} \textcite{langlet-ml-netronome}---although \emph{actual NN latency} in his Fig 4.4 (IPG is inter-packet gap: his is not truly async but achieves perf by lying on the packet path.
+Note that 0 IPG murders his latencies).
+
 \section{Conclusion}
@@
 \begin{appendices}
-	\begin{figure}
-		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
-		}
-		\caption{Latency! varying work.\label{fig:vary-work-latency}}
-	\end{figure}
-
-	\begin{figure}
-		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
-		}
-		\caption{Latency! varying core count\label{fig:vary-core-latency}}
-	\end{figure}
 	\section{Old ``What could we do''?}

### 2021-05-07 paper/paper.tex
Some diagram descriptions

@@
 
+\begin{figure}
+	\centering
+	\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/arch-with-p4}
+	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
+SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible.\label{fig:netro-arch}}
+\end{figure}
+
 Automatic optimisation, control, and defence of networks are at last becoming commonplace.
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/arch-with-p4}
-	\caption{Architecture for an \approachshort{} agent on Netronome hardware, using 4 cores for parallel policy processing.
-(FIXME: redo in illustrator?
-Show CTXes, PISA, \Coopfw{} vs \Indfw{})\label{fig:netro-arch}}
-\end{figure}
-
 \subsection{System Model}
@@
 		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/ind}
-		\caption{\Indfw{} (offline throughput-optimal).\label{fig:single-and-parallel:single}}
+		\caption{\Indfw{} (offline throughput-optimal).
+\emph{Workers} independently pull commands from and push state-action pairs to the environment (locking access to when updates must occur).\label{fig:single-and-parallel:single}}
 	\end{subfigure}
+\hspace{0.04\linewidth}
 	\begin{subfigure}{0.45\linewidth}
@@
 		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/coop}
-		\caption{\Coopfw{} (online-optimal).\label{fig:single-and-parallel:parallel}}
+		\caption{\Coopfw{} (online-optimal). A single \emph{controller} ... to many \emph{minion} threads, who operate on independent subtasks.\label{fig:single-and-parallel:parallel}}
 	\end{subfigure}
-	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to latency/throughput needs. ?? MAKE ME FOR REAL \label{fig:single-and-parallel}}
+	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to latency/throughput needs.\label{fig:single-and-parallel}}
 \end{figure*}

### 2021-05-10 paper/paper.tex
New tables, expanded writing!

@@
 	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
-SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible.\label{fig:netro-arch}}
+SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible alongside P4 dataplanes.
+Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
 \end{figure}
@@
 \section{Background and Motivation}\label{sec:motivation}
+\subsection{Reinforcement Learning}
 ?? RL~\parencite{RL2E}
@@
 
-\fakepara{Programmable hardware capabilities}
+\subsection{Programmable hardware capabilities}
 The introduction of the P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} language has led to explosive growth in the research community surrounding in-network computation and offloading.
@@
 
+\subsection{Netronome Platform Fundamentals}
+Explain MEs, Threads, CTXs, so on so on.
+
 \section{Design and Implementation}\label{sec:design}
-?? Can we say ``in-situ'' anywhere? I like that term.
-
 Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{} (\approach)}.
@@
 
-\Cref{fig:netro-arch} outlines our design and implementation on Netronome SmartNIC hardware in pursuit of this goal.
+\begin{figure*}
+	\centering
+	\begin{subfigure}{0.45\linewidth}
+		\centering
+		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/ind}
+		\caption{\Indfw{} (offline throughput-optimal).
+\emph{Workers} independently pull commands from (and push state-action pairs to) the environment, locking access to the policy when updates must occur.\label{fig:single-and-parallel:single}}
+	\end{subfigure}
+	\hspace{0.04\linewidth}
+	\begin{subfigure}{0.45\linewidth}
+		\centering
+		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/coop}
+		\caption{\Coopfw{} (online-optimal).
+A single \emph{controller} delegates RL computation and updates to many \emph{minion} threads, who operate on independent subtasks.\label{fig:single-and-parallel:parallel}}
+	\end{subfigure}
+	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to either latency or throughput needs.\label{fig:single-and-parallel}}
+\end{figure*}
+
+\Cref{fig:netro-arch,fig:single-and-parallel} outline our design and implementation on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
 We open-source our firmware design and control programs for the benefit of the community.\footnote{\url{https://github.com/FelixMcFelix/pdp-rl-paper}---currently private...}
-We describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.
+Moving beyond the Netronome platform, we describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.
 
@@
 \approachshort{} runs on one more cores of a SmartNIC to convert fixed-point state measurements from the environment into a stream of actions using a stored policy.
-These dedicated cores (\emph{microengines} grouped by \emph{island} in NFP terminology) are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
+As an example, this might be to map flow state and performance measurements into a queue priority for future packets from that flow, or to compute and apply a rate limit to preserve quality of service.
+These dedicated cores are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
 Combined with reward measurements, this policy can then be updated or trained from scratch entirely on the NIC, acting as a fully online RL agent.
@@
 
-To protect traffic throughput and allow for effective deployment in as many environments as possible, \approachshort{} places RL execution off the main packet path, communicating and running parallel to the main P4 dataplane.
-This asynchrony allows coexistence with P4 programs, and imposes minimal/no impact on carried traffic for both bump-in-the-wire deployments and at end-points.
+To protect traffic throughput and allow for effective deployment in as many environments as possible, \approachshort{} places RL execution on-chip, \emph{but off the main packet path}, communicating and running parallel to the main P4 dataplane.
+As shown in \cref{fig:netro-arch}, this asynchrony allows coexistence with P4 programs, and imposes minimal/no impact on carried traffic for both bump-in-the-wire deployments and at end-points.
 For instance, in the default deployment of a P4 packet processing pipeline on Netronome NFP chips several of these cores go unused (as does spare area on an FPGA design), making this paradigm possible.
@@
 The new state- or tiles-action pair is then written into storage.
-
-\begin{figure*}
-	\centering
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/ind}
-		\caption{\Indfw{} (offline throughput-optimal).
-\emph{Workers} independently pull commands from and push state-action pairs to the environment (locking access to when updates must occur).\label{fig:single-and-parallel:single}}
-	\end{subfigure}
-\hspace{0.04\linewidth}
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/coop}
-		\caption{\Coopfw{} (online-optimal). A single \emph{controller} ... to many \emph{minion} threads, who operate on independent subtasks.\label{fig:single-and-parallel:parallel}}
-	\end{subfigure}
-	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to latency/throughput needs.\label{fig:single-and-parallel}}
-\end{figure*}
-
-?? Do we need better/snappier names than just \Indfw/\Coopfw? It's awfully dry.
 
@@
 
-We found that, paradoxically, it was more efficient in the parallel case \emph{to do more work} by having each worker recompute its tile subset from a stored state.
+Latency and throughput, as in many networked systems, have different effects upon RL agents according to their design and target problem.
+Higher throughput for an RL agent is a necessity for per-flow or per-packet applications, which can require high decision-per-second rates even after combining state measurements received from the environment, such as flow control in DDoS prevention.
+Equally, lower latency affords an RL agent finer-grained control and learning of a problem, being able to react sooner to new information (e.g., device state in a routing optimisation problem, or queue depth when trying to enforce packet pacing).
+
+We found that, paradoxically, it was more efficient in the \Coopfw{} case \emph{to do more work} by having each worker recompute its tile subset from a stored state.
 It transpired that cacheing this data placed a larger \texttt{memcpy} in the serial section, whose size did not scale at all with bit depth.
@@
 
-\subsection{Notes from chat with Rhys}
-?? Where/how to fit this in? I think it's a useful point.
-
-The two compute models discussed here need not be homogeneously deployed.
-For instance, in a networked deployment a subset of \approachshort{} nodes could be \SI{32}{\bit} \emph{\coopfw{}} agents, training online, while most other nodes ran \SI{8}{\bit} \emph{\indfw{}} designs to meet throughput guarantees.
-The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
-This can be taken further still, using policy deltas or execution traces to enable out-of-path transfer learning for more complex function approximators such as neural networks.
-
 %Potential: low-latency train for some specific points, and high throughput for others? Think on the exact note I took...
@@
 
-\subsection{Limitations}
+\subsection{Limitations}\label{sec:limitations}
 Unfortunately, direct rule installation into P4 tables from the SmartNIC itself is not, in general, possible.
@@
 OTHER THING (\cref{sec:integ-2}).
-We support these using other state-of-the-art P4/PDP developments.
+We support these using other state-of-the-art P4/PDP developments, and discuss how best to balance the concerns of online training with throughput in widespread deployment.
 
 \subsection{In-Network DDoS Defence}\label{sec:integ-1}
-Classical reinforcement learning has seen some recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
-
-?? Bump-in-the-wire deployment of SmartNICs at edge nodes.
-
-?? Place Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} on the datapath as the main source of state vectors.
-?? This resides in the P4 dataplane, and exists on-chip alongside the \approachshort{} agent.
-?? Use INDDoS~\parencite{tnms-ddos-victim-ident} as a reward function source---i.e., using the number of detected victims as a loss value.
-?? Still receive load measures from elsewhere in the network; this is now the element at highest risk of being outdated, but the least likely to vary significantly in response to individual actions.
-
-?? How does increased attack flow ratio affect the RL core? How does changing its params to increase its own throughput?
-?? Remember the timed-random-selection; use the state coalescing trick indep discovered by you and those other guys.
-?? As such, the state collection mechanism prevents flows from being acted upon too often and ensures that the RL decision-making
-?? As we shall see, this far exceeds the number of decisions per second that a host based deployment could handle, though using these techniques is still key in preventing overloads.
-
-?? The P4 dataplane is then used to reduce/increase the number of flows used as inputs? (YES)
+Classical reinforcement learning has seen recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
+\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and local, per-flow state to monitor how flows respond to bandwidth changes and packet drop---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} with more allowance for congestion-unaware protocols.
+Actions then move flows up or down in punishment levels according to a finite state machine.
+
+\fakepara{Why in-NIC?}
+The above approach relies upon co-hosting traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
+However, packet mirroring, offloading RL computation to a host (potentially over layers of virtualisation), and CPU performance are all sources of additional, consistent state-action latency.
+Both traffic statistics collection \emph{and} data-driven learning must be executed on such hosts/network functions.
+Unless running these stages in a dedicated pipeline (adding further processing latency), resource contention between these processes will further impact tail latencies.
+Naturally, this requires high-performance hardware to be stacked at network egress points, potentially beyond reasonable space, power, or ventilation constraints.
+The solution to implement and improve upon this work using \approachshort{} is to place its RL agents on SmartNICs at AS edge nodes---a bump-in-the-wire deployment.
+
+\fakepara{Inputs}
+To collate the required inputs and state, we then examine the recent innovations of the community.
+Low-latency, pure-P4 solutions to extract and record per-flow TCP state directly in the dataplane such as Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} are well-studied.
+In fact, the statistics offered by Dapper are a super-set of the local input state values employed by \emph{Guarded} agents, offering an opportunity to further improve their efficacy. 
+We propose placing such monitors in the P4 dataplane, existing on-chip alongside the \approachshort{} agent.
+The required global state (load measures from network paths) must still come from elsewhere in the network; this is now the element at highest risk of becoming stale, but the least likely to vary significantly in response to individual actions.
+
+The original work suggests theoretical ``ground truth'' reward functions, whose correct implementation and designs were left as an open challenge. 
+We posit that INDDoS~\parencite{tnms-ddos-victim-ident}, which uses Count-min Sketches to estimate DDoS victim cardinality, could be used as an effective reward function source---i.e., using the number of detected victims as a loss value.
+
+\fakepara{Integrating \approachshort}
+Before each Dapper monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{OUT Ring} for any generated actions.
+As we note in \cref{sec:limitations}, these actions would be placed into a small local hash table and simultaneously exported to the attached controller to be inserted as P4 rules in batches.
+After the main monitoring step, packet ingress timestamps would be used to emulate the \emph{Timed Random Sequential} (TRS) scheduler used by the anti-DDoS agents for rate-controlled work, where state vectors would be selected and passed into \approachshort.
+By design, active flows which are not \emph{judged} after a configurable time are discarded to prune the work set and allow new flows to be seen by the agent.
+The tight bounds on execution time known \emph{a priori} make it easy to calculate the maximum number of decisions which can be made per deadline.
+Reward values would then be separately inserted by a modified INDDoS table.
+
+?? Can I have a P4 code snippet here?
+
+A reduction in state-action latency (i.e., with \Coopfw) is, as always, useful for minimising the noise inherent in learning.
+However, an agent is limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
+Accordingly, this use-case benefits most from an increase in \emph{throughput} using \Indfw{}.
+In this context, higher throughput means that network flows are more likely to be judged in every timestep, even when flow cardinality is high---making it more likely that changes in flow behaviour will be observed, acted on, and learned from.
+
+We note that the TRS scheduler is designed to handle cases where the number of flows created by an attacker, combining state vectors and information over time while the asynchronous agent is itself busy.
+An increased number of (attacking) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
+As we shall examine in \Cref{sec:evaluation}, \approachshort{} far exceeds the action throughput of a host based deployment could handle.
+The control plane is then free to use wildcards or specific matches to narrow down (or expand) the set of flows to be considered and controlled, though explicit scheduling techniques are still key in such adversarial environments.
 
@@
 
+\subsection{Network Deployment Considerations}
+The two compute models discussed above need not be homogeneously deployed.
+For instance, in a networked deployment a subset of \approachshort{} nodes could be \SI{32}{\bit} \emph{\coopfw{}} agents, training online, while most other nodes ran \SI{8}{\bit} \emph{\indfw{}} designs to meet throughput guarantees.
+The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
+This can be taken further still, using policy deltas or execution traces to enable out-of-path transfer learning for more complex function approximators such as neural networks.
+
 \section{Evaluation}\label{sec:evaluation}
-We outline our performance evaluation of...
-
-?? find something meaningful to put here.
+We outline our performance evaluation of \approachshort.
 
@@
 Evaluation of \approachshort{} was performed on server blades hosting a single Netronome Agilio LX 1x100GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
-These ran Ubuntu 18.04.5 LTS (4.15.0-140-generic x86\_64) on Intel Xeon Bronze 3204 CPUs (\SI{1.9}{\giga\hertz}) and \SI{32}{\gibi\byte} RAM.
+These mid-tier servers (\emph{MidServer}) ran Ubuntu 18.04.5 LTS (4.15.0-140-generic x86\_64) on Intel Xeon Bronze 3204 CPUs ($6\times$\SI{1.9}{\giga\hertz}) and \SI{32}{\gibi\byte} RAM.
+We additionally use a more powerful consumer-grade machine (\emph{Collector}) for estimating host performance when offloaded to a network function, running Ubuntu 18.04.4 LTS (4.15.0-96-generic) on Intel Core i7-6700K ($4\times$\SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} RAM.
 Control programs were built using rustc 1.51.0.
@@
 
+\fakepara{Deployability}
+By timing policy installation, compile times and other metrics, we describe the runtime costs needed for an administrator to repurpose an installed agent in a live network.
+
 \subsection{Results and Discussion}\label{sec:results}
@@
 \begin{table}
-	\caption{Throughput and per-item compute time for an RL agent with/without policy learning.
-State$\rightarrow$action latency equals the offline completion cost.
-TODO: compare to host machine?
-Put completion times first?
-Put multi first?\label{tab:lat-tput}}
+	\caption{Latencies. On-device execution is crucial in not only lowering latencies, but in reducing tail latencies. Lower is better.\label{tab:lats}}
 	\resizebox{\linewidth}{!}{
-		\expandableinput ../tables/build/latency-tput-temp
+		\expandableinput ../tables/build/latency
+	}
+\end{table}
+
+\begin{table}
+	\caption{Puts! Designs which do not take advantage..... Higher is better.\label{tab:tputs}}
+	\resizebox{\linewidth}{!}{
+		\expandableinput ../tables/build/tput
 	}
@@
 \caption{NFP memory use due to \approachshort{} using \numlist{1;4} MEs (\SI{32}{\bit}).
-CLS and CTM regions are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM regions are shared between all NFP programs on a NIC.\label{tab:resources} ??NEEDS REWORKED}
+CLS and CTM regions are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM regions are shared between all NFP programs on a NIC.\label{tab:resources}}
 \resizebox{\linewidth}{!}{
@@
 %\end{figure}
+
+\fakepara{Deployability}
+?? Config times scale with work complexity in \Indfw, confirming the $\mathcal{O}{\left(n\right)}$ complexity of the allocator (??).
+?? Changing from online to offline has this cost.
+
+?? Anything else worth FW install time fairly constant (i.e., time to swap between bit-depths).
 

### 2021-05-10 tables/build/latency.tex
New tables, expanded writing!

\begin{tabular}{@{}ccS[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]@{}}
\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{3}{c}{State-Action Latency (\si{\micro\second})} & \multicolumn{3}{c}{State-Update Time (\si{\micro\second})}\\
\cmidrule(lr){3-5}\cmidrule(lr){6-8} & & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{\nth{99}} & \multicolumn{1}{c}{\num{99.99}\nthscript{th}} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{\nth{99}} & \multicolumn{1}{c}{\num{99.99}\nthscript{th}}\\ \midrule
Float & Collector & 515.94 & 606.06 & 725.03 & 606.06 & 636.82 & 833.99\\
 & MidServer & 1069.07 & 1125.1 & 1508.0 & 1260.04 & 1605.99 & 1719.864\\
\cmidrule(lr){1-8}
Int32 & \approachshort{}-\Indfw{} & 185.040 & 185.373 & 185.707 & 230.707 & 231.200 & 231.633\\
 & \approachshort{}-\Coopfw{} & \bfseries 44.133 & \bfseries 44.280 & \bfseries 44.320 & \bfseries 81.287 & \bfseries 81.760 & \bfseries 82.447\\
Int16 & \approachshort{}-\Indfw{} & 195.320 & 195.680 & 195.953 & 242.040 & 242.587 & 243.547\\
 & \approachshort{}-\Coopfw{} & 46.293 & 46.453 & 46.493 & 84.347 & 84.800 & 85.567\\
Int8 & \approachshort{}-\Indfw{} & 193.413 & 193.733 & 194.247 & 240.173 & 240.707 & 241.547\\
 & \approachshort{}-\Coopfw{} & 45.920 & 46.013 & 46.040 & 83.733 & 84.213 & 85.160\\
\bottomrule
\end{tabular}

### 2021-05-10 tables/build/tput.tex
New tables, expanded writing!

\begin{tabular}{@{}ccS[table-format=2]S[table-format=3.3]S[table-format=2.3]S[table-format=1.3]S[table-format=1.3]@{}}
\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{1}{c}{Workers} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{2}{c}{Throughput/core (k actions/s)}\\
\cmidrule(lr){4-5}\cmidrule(lr){6-7} & & & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online} & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online}\\ \midrule
Float & Collector & 4 & 7.673 \pm 0.049 & 1.627 \pm 0.031 & 1.918 \pm 0.012 & \multicolumn{1}{c}{---}\\
 & MidServer & 6 & 5.584 \pm 0.030 & 0.791 \pm 0.012 & 0.931 \pm 0.005 & \multicolumn{1}{c}{---}\\
\cmidrule(lr){1-7}
Int32 & \approachshort{}-\Indfw{} & 32 & \bfseries 172.956 \pm 0.171 & 4.335 \pm 0.005 & \bfseries 5.405 \pm 0.005 & \multicolumn{1}{c}{---}\\
 & \approachshort{}-\Coopfw{} & 32 & 22.670 \pm 0.073 & \bfseries 12.308 \pm 0.043 & 0.708 \pm 0.002 & \bfseries 0.385 \pm 0.001\\
Int16 & \approachshort{}-\Indfw{} & 32 & 163.822 \pm 0.115 & 4.132 \pm 0.004 & 5.119 \pm 0.004 & \multicolumn{1}{c}{---}\\
 & \approachshort{}-\Coopfw{} & 32 & 21.598 \pm 0.031 & 11.858 \pm 0.027 & 0.675 \pm 0.001 & 0.371 \pm 0.001\\
Int8 & \approachshort{}-\Indfw{} & 32 & 165.472 \pm 0.148 & 4.165 \pm 0.005 & 5.171 \pm 0.005 & \multicolumn{1}{c}{---}\\
 & \approachshort{}-\Coopfw{} & 32 & 21.792 \pm 0.063 & 11.953 \pm 0.041 & 0.681 \pm 0.002 & 0.374 \pm 0.001\\
\bottomrule
\end{tabular}

### 2021-05-10 tables/src/latency.py

@@ -1,0 +1,115 @@
+import numpy as np
+import sys
+
+sys.stdout.reconfigure(encoding='utf-8')
+
+# sub in: bit depth, fw-type, measure_name
+float_results_fmt_str = "../results/host-rl-perf-tput/SUMMARY.f32.{}.{}.dat"
+results_fmt_str = "../results/rl-perf-tester/vary-work-ct/{}/{}.28d.31c.{}.dat"
+
+cycle_in_ns = 1e9 / 1.2e9
+
+core_ct_for_tput = 32
+types = [("single", R"\Indfw{}", True), ("balanced", R"\Coopfw{}", False)]
+measures = [("ComputeAndWriteout", R"\xmark", False),("UpdateAll", R"\cmark", True)]
+bitdepths = [32, 16, 8]
+
+print(R"\begin{tabular}{@{}ccS[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]@{}}")
+print(R"\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{3}{c}{State-Action Latency (\si{\micro\second})} & \multicolumn{3}{c}{State-Update Time (\si{\micro\second})}\\")
+
+# Just generate the next header line...
+t2_header = R"\cmidrule(lr){3-5}\cmidrule(lr){6-8} &"
+
+measure_names = ["Median", R"\nth{99}", R"\num{99.99}\nthscript{th}"]
+
+for i in range(2):
+	for name in measure_names:
+		t2_header += R" & \multicolumn{1}{c}{"+name+R"}"	
+
+t2_header += R"\\ \midrule"
+
+best_row_for_each_column = [3, 3, 3, 3, 3, 3]
+curr_row = 0
+
+print(t2_header)
+
+# Deal with Float results here.
+# File names like: results/host-rl-perf-tput/SUMMARY.f32.beast1.ComputeAndWriteout.dat
+# These results are already in micros.
+
+machines = [("Collector", "beast2", 4), ("MidServer", "perlman", 6)]
+
+for i, (machine_pres, machine_name, core_limit) in enumerate(machines):
+	out = "{} & {}".format("Float" if i == 0 else "", machine_pres)
+
+	latencies = []
+
+	for online_i, (online_fname_part, online_latex_part, needs_lock) in enumerate(measures):
[…]

### 2021-05-10 tables/src/tput.py

@@ -1,0 +1,167 @@
+import numpy as np
+import sys
+
+sys.stdout.reconfigure(encoding='utf-8')
+
+# sub in: bit depth, fw-type, measure_name
+float_results_fmt_str = "../results/host-rl-perf-tput/SUMMARY.f32.{}.{}.dat"
+results_fmt_str = "../results/rl-perf-tester/vary-work-ct/{}/{}.28d.31c.{}.dat"
+
+cycle_in_ns = 1e9 / 1.2e9
+
+core_ct_for_tput = 32
+types = [("single", R"\Indfw{}", True), ("balanced", R"\Coopfw{}", False)]
+measures = [("ComputeAndWriteout", "Offline", False),("UpdateAll", "Online", True)]
+bitdepths = [32, 16, 8]
+
+print(R"\begin{tabular}{@{}ccS[table-format=2]S[table-format=3.3]S[table-format=2.3]S[table-format=1.3]S[table-format=1.3]@{}}")
+print(R"\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{1}{c}{Workers} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{2}{c}{Throughput/core (k actions/s)}\\")
+
+# Just generate the next header line...
+t2_header = R"\cmidrule(lr){4-5}\cmidrule(lr){6-7} & &"
+
+for i in range(2):
+	for (_f, name, _n) in measures:
+		t2_header += R" & \multicolumn{1}{c}{"+name+R"}"	
+
+t2_header += R"\\ \midrule"
+
+print(t2_header)
+
+best_row_for_each_column = [2, 3, 2, 3]
+curr_row = 0
+
+# Deal with Float results here.
+# File names like: results/host-rl-perf-tput/SUMMARY.f32.beast1.ComputeAndWriteout.dat
+# These results are already in micros.
+
+machines = [("Collector", "beast2", 4), ("MidServer", "perlman", 6)]
+
+for i, (machine_pres, machine_name, core_limit) in enumerate(machines):
+	out = "{} & {}".format("Float" if i == 0 else "", machine_pres)
+
+	best_core = None
+	tputs = []
+	stds = []
+	core_cts = []
+
+	for online_i, (online_fname_part, online_latex_part, needs_lock) in enumerate(measures):
+		fname = float_results_fmt_str.format(machine_name, online_fname_part)
+
+		best_core_ct = 0
+		best_dats = None
[…]

### 2021-05-11 paper/paper.tex
Intro + BG made a reality

@@
 \documentclass[sigconf,natbib=false]{acmart}
+
+\usepackage{amsmath}
+\let\openbox\relax
+\usepackage{amsthm}
+\newcommand\bmmax{2}
+\usepackage{bm}
 
@@
 \newcommand{\approachshort}{OPaL}
-\newcommand{\Coopfw}{CoOp}
+\newcommand{\Coopfw}{\emph{CoOp}}
 \newcommand{\coopfw}{\Coopfw}
-\newcommand{\Indfw}{Ind}
+\newcommand{\Indfw}{\emph{Ind}}
 \newcommand{\indfw}{\Indfw}
@@
 \newcommand{\xmark}{\ding{55}}%
+
+% make math easy
+\newcommand{\acval}[3]{\ensuremath{\operatorname{\hat{q}}(#1, #2, #3)}}
+\newcommand{\acvalblank}{\ensuremath{\operatorname{\hat{q}}(\cdot)}}
+\newcommand{\wvec}[1]{\ensuremath{\bm{w}_{#1}}}
 
@@
 ??
-In-NIC execution allows clever use of resources to bring action compute latency  from \SI{1}{\milli\second} to \SI{50}{\micro\second} (2 orders of magnitude!) in real-world policies for DDoS detection/prevention control.
+In-NIC execution allows clever use of resources to bring 99.99th percentile action compute latency  from \SI{725}{\micro\second} to \SI{44}{\micro\second} (1.5 orders of magnitude!) in real-world policies for DDoS detection/prevention control.
+
+?? We are the first to bring fully online RL to the dataplane!
 \end{abstract}
@@
 \section{Introduction}
-?? REDO ME
+?? REDO ME: this is an old abstract
 
@@
 	\item Potential integrations of \approachshort{} with state-of-the-art PDP applications to perform fully in-NIC, automated DDoS mitigation and ?? OTHER THING (\cref{sec:potential-integrations}),
-	\item In-depth evaluation of how our approach affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{0}{$\times$} latency reduction (\crefrange{sec:evaluation}{sec:results}).
-??actual val
+	\item In-depth evaluation of how our approach affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{11.7}{$\times$} latency reduction compared to commodity hardware (\SI{16.4}{$\times$} at 99.99\nthscript{th} tail) (\cref{sec:evaluation}).
 \end{itemize}
@@
 \section{Background and Motivation}\label{sec:motivation}
+We discuss recent trends in programmable switch hardware
+By combining this with insights from the ML/RL communities (past and present), we discuss why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility.
+
+Crucially, we train our focus on devices with a low port density, such as SmartNICs and NetFPGA devices.
+The designs of these devices make it reasonable to move policy processing \emph{outside of the packet pipeline}; SmartNICs expose many additional programmable cores, and NetFPGAs allow for the synthesis of independent functional units.
+The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
+
+\subsection{Programmable hardware capabilities}
+The introduction of the P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} language has led to explosive growth in the research community surrounding in-network computation and offloading.
+Providing a single language which is compatible with network devices of many form factors has been instrumental in the development of novel fine-grained traffic measurement approaches~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
+However, this requires that there be a reasonably consistent interface and behaviour between device classes.
+This need for consistency underpins the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}, which defines a conceptual model of match-action tables divided into ingress and egress pipelines.
+The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.
+
+We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
+Many compatible devices have legacies long predating these developments. 
+For instance, many-core SOC-based Netronome SmartNIC~\parencite{netronome-smartnic}, and NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14} NICs allow virtually arbitrary programs to be specified and executed.
+The first of these compiles from P4$\rightarrow$Micro-C$\rightarrow$bytecode (with Micro-C externs), while the latter can combine the P4$\rightarrow$NetFPGA toolchain~\parencite{DBLP:conf/fpga/IbanezBMZ19} with arbitrary circuit externs---both exposing further device capabilities beyond the specification.
+This pattern extends to other SmartNICs, such as NVIDIA BlueField~\parencite{nvidia-bluefield} and Xilinx Alveo~\parencite{xilinx-alveo} devices.
+Even Intel Tofino~\parencite{barefoot-intel} ASICs, which architecturally mirror the PSA, expose additional matching capabilities and ALU functions via the Tofino Native Architecture.
+Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of deep learning accelerators (which, in turn, lack packet switching functionality).
+Moreover, the limited memory/block RAM and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
+
+\begin{insight}
+	Low-port density programmable network devices often have spare resources and extra capabilities beyond the PSA specification which can aid asynchronous, local compute.
+\end{insight}
+
+\fakepara{To-be-organised}
+Interesting thing I realised from \textcite{DBLP:conf/sigcomm/NeugebauerAZAL018}: GPU and NIC might cause IOMMU contention?
+Making line-rate offload of DDN to GPU impractical.
+Unsure if I want to twin this with later measurements/plots.
+
 \subsection{Reinforcement Learning}
-?? RL~\parencite{RL2E}
-
-?? Use this to explain differences between the modern state of the art and older works.
-
-?? Probably give a rough introduction to the algorithm, show what arithmetic operations are needed---particularly when using Qm.n fixed point.
-
-?? Tile coding requires div, though can be simplified using binary tile widths (relates down to bitshifts). What other ops?
-
-?? Explain how it is an embarrassingly parallel problem? Or at least very amenable to partition-aggregate.
-
-?? Critical section around shared writeback loc, but individual work items are disjoint in their input data (particularly useful for updates).
+\emph{Reinforcement learning} (RL) algorithms are methods of training an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task \cite{RL2E}.
+Like most machine learning methods, an RL algorithm uses gradient information to update the parameters used to approximate a function.
+In RL contexts, this is a state$\rightarrow$action function known as a \emph{policy}.
+When training, agents are given \emph{reward measurements} and a learned policy acts to maximise the \emph{expected discounted reward} received.
+When a pre-trained policy is deployed, this signal is not required.
+
+It's useful to consider how these algorithms differ from other ML use cases, such as classifiers.
+The main differences lie in how this gradient information is used, and combined with \emph{reward measurements} received from the environment.
+Rather than adapting the learned parameters along the gradient using an error value from a target value, because the optimal actions aren't known they adjust values using a \emph{Markov Decision Process} (MDP)---capturing state trajectories to adjust value based on past and future decisions.
+
+Consider the single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E}:
+\begin{subequations}
+	\begin{gather}
+		\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
+		\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
+	\end{gather}%
+	\label{eqn:sg-sarsa}%
+	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, $\acval{S}{A}{\wvec{}}$ denotes the \emph{value} of an action $A$ taken in state $S$ according to the policy $\wvec{}$, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
+$\gamma,\alpha\in[0,1]$ are the discount factor and learning rate (governing the degree of forward planning and policy stability, respectively).
+\end{subequations}
+In essence, at each timestep the policy parameters ($\wvec{}$) are increased along the gradient ($\nabla{\acvalblank}$) using a fixed learning rate ($\alpha$) and a computed adjustment ($\delta_t$).
+This adjustment is equal to the difference between the chosen action $A$'s value in state $S$ and the reward received ($R_{t+1} - \acval{S_t}{A_t}{\wvec{}}$), \emph{plus} some part of the \emph{next} action's value ($\gamma\acval{S_{t+1}}{A_{t+1}}{\wvec{}}$).
+To give some context on the design space here, other algorithms may employ separate state value approximations, use the entirety of an agent's trace, or be tailored to characteristics of the policy approximator (e.g., how neural networks benefit from batching).
 
 \begin{insight}
-	Classical RL methods are computationally simple to update online, and inference and learning are embarrassingly parallel.
+	Classical, single-step RL algorithms are computationally simple and are agnostic to the policy approximation used.
+Furthermore, they require only additions and multiplication to update and learn a policy online.
 \end{insight}
 
-\fakepara{Reinforcement learning in asynchronous environments}
-RL techniques allow an \emph{agent} to learn how to iteratively interact with a system in an optimal fashion.
-However, there is some degree of divergence between their theory and implementation.
+\fakepara{RL in asynchronous environments}
+There remains some degree of divergence between the theory and implementation of RL agents.
 Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
@@
 
-?? Find some cites citing the relevance of this problem wrt. self-driving cars, robotics, etc.
+%?? Find some cites citing the relevance of this problem wrt. self-driving cars, robotics, etc.
 
@@
 
-\subsection{Programmable hardware capabilities}
-The introduction of the P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} language has led to explosive growth in the research community surrounding in-network computation and offloading.
-Providing a single language which is compatible with network devices of many form factors has been instrumental in the development of novel fine-grained traffic measurement approaches~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
-However, this requires that there be a reasonably consistent interface and behaviour between device classes.
-This need for underpins the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}, which defines a conceptual model of match-action tables divided into ingress and egress pipelines.
-The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.
-
-We would be remiss to believe that the P4 and the PSA define all capabilities that programmable hardware can support.
-Many compatible devices have legacies predating these developments, for instance many-core SOC-based Netronome SmartNIC~\parencite{netronome-smartnic}, and NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14} NICs.
-The first of these compiles from P4$\rightarrow$Micro-C$\rightarrow$bytecode (with Micro-C externs), while the latter can combine the P4$\rightarrow$NetFPGA toolchain~\parencite{DBLP:conf/fpga/IbanezBMZ19} with arbitrary circuit externs---both exposing further device capabilities beyond the specification.
-Even Intel Tofino~\parencite{barefoot-intel} ASICs, which architecturally mirror the PSA, expose additional matching capabilities and ALU functions via the Tofino Native Architecture.
-Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of DL accelerators (which, in turn, lack packet switching functionality).
-Moreover, the limited memory/block RAM and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
+\subsection{Efficient policy approximation}
+Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
+This can then make \emph{online} learning of the task an attractive prospect, using a single available stream of experience.
+\emph{Deep neural networks} (DNNs), particularly when used as the basis for an RL agent's policy, require vast amounts of experience to converge on an accurate parameter set.
+In many problem domains, this equates to training from compute-years worth of distributed offline simulations, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
+Achieving stable learning requires sizeable batches of gradients to be computed, potentially using the entire experience replay from each simulation.
+Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the computational cost of gradient computation via backpropagation remains too significant on embedded hardware.
+
+%?? Probably want some cites on the need for batching in NN methods, even though this is understood. DL book?
+
+We then return again to what classical RL methods can offer us; in particular, the simple linear function approximation given by \emph{tile coding} \cite[pp.\ \numrange{217}{221}]{RL2E}.
+The idea is simple: a policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
+A state vector produces a single `hit' in each tiling, which provides both the set of action preference lists to sum during inference \emph{and} the set of tiles to update at the next timestep.
+Crucially, there are no data dependencies between calculating any tile hits for an input state vector.
+This has important benefits for parallelisation.
+To select an action, we can produce a task for each tiling---retrieving the list of action values contributed by the tile activated by the input state.
+An aggregate step sums up all action preference lists, and the action with the highest value is then selected.
+Updating the policy (with a new $\delta_t$ value) has no aggregate step, and as tile indices map to disjoint memory regions no locks are required.
+
+?? Algo block/diagram?
+
+For instance, a policy with $m$ sets of tiles (each having its own set of input dimensions), each containing $n$ overlapping tilings then creates $m \times n$ separate tasks.\footnote{Retrieving each individual action's value can be considered a work item (giving $m \times n \times a$ tasks when there are $a$ actions).
+However, this requires the tile coding operation to repeated $a$ times per tile, which is likely extremely wasteful outside of extreme degrees of parallelism.}
+As an example, a real-world policy size which we examine later (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets containing \num{8} tilings creates \num{129} work items.
+
+Computing a tile hit requires one division per dimension, which is an expensive addition to the ALU operations we require from a platform.
+However, if the width of tiles are fixed at powers-of-two, then these may be replaced with right bitshifts.
+As we shall see shortly, this aligns with our choice of numerical representation.
 
 \begin{insight}
-	Low-port density programmable network devices often have spare resources and extra capabilities beyond the PISA specification which can aid asynchronous, local compute.
+	Tile coded policies are embarrassingly parallel in both inference \emph{and} learning. Moreover, in some instances the process can be altered to require only bitshifts, additions, and multiplication.
 \end{insight}
 
-Luckily, the task of bringing machine learning models to resource-constrained environments is well-studied.
+\subsection{Numerical representation for embedded ML}
+A key feature universally lacking from the target environments is floating-point arithmetic support.
+Luckily, the task of bringing machine learning models to resource-constrained environments without these capabilities is well-studied.
 Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
@@
 Much of this work goes further still towards integer~\parencite{tensorrt-8bit} or binarised~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16} representations, sacrificing dynamic range for simpler arithmetic operations.
-This has the important side effect of making ML policy inference tractable on hardware without floating-point capabilities, which other works have leveraged to bring neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
-
-To make RL workloads feasible under such constraints, we propose the use of quantised, fixed-point representations such as $Qm.n$, which allow us to evaluate and update policies using only integer arithmetic.
+The works mentions above use similar techniques to make inference tractable on network hardware, i.e., to bring neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
+
+To make RL workloads feasible under such constraints, we propose the use of quantised, fixed-point representations such as $Qm.n$ (i.e., $m$\si{bit} signed integers with an $n$\si{bit} fractional part), which allow us to evaluate and update policies using only integer arithmetic.
 This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation ($t_2$, \cref{fig:state-slip}).
-Crucially, we train our focus on devices with a low port density, such as SmartNICs and NetFPGA devices.
-The designs of these devices make it reasonable to move policy processing \emph{outside of the packet pipeline}; SmartNICs expose many additional programmable cores, and NetFPGAs allow for the synthesis of independent functional units.
-The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
+
 When combined with the above P4-driven techniques for in-network flow/device measurement, we at last have the mechanisms to collocate the key processes of an RL agent.
@@
 
-?? relate update/execute operations here? i.e., policy execute needs only ALU add (and whatever is needed for tile code), train needs add+mul+shr. or in RL section? Should that come first?
-
-\fakepara{Online training}
-Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
-This can then make \emph{online} learning of the task an attractive prospect, using a single available stream of experience.
-DNNs, particularly when used as the basis for an RL agent's policy, require vast amounts of experience to converge on an accurate parameter set.
-In many problem domains, this equates to training from compute-years worth of distributed offline simulations, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
-Achieving stable learning requires sizeable batches of gradients to be computed, potentially using the entire experience replay from each simulation.
-Still, the computational cost of gradient computation via backpropagation is significant on embedded hardware.
-
-To then achieve \emph{online} learning, we return our focus to older \emph{classical} reinforcement learning methods.
-In particular, we suggest tile-coding~\cite[pp. \numrange{217}{221}]{RL2E}, with one-step temporal-difference learning algorithms such as Sarsa or Q-learning~\cite[pp. \numrange{129}{132}]{RL2E}.
+\begin{insight}
+	Fixed point quantisation allows us to perform numerical calculations using only integer arithmetic.
+	Existing literature to-date supports the effectiveness and use of these techniques on embedded ML deployments.
+\end{insight}
+
+\subsection{Netronome Platform Fundamentals}
+?? Explain MEs, Threads, CTXs, so on so on.
+
+\subsection{Discussion}
+%?? relate update/execute operations here? i.e., policy execute needs only ALU add (and whatever is needed for tile code), train needs add+mul+shr. or in RL section? Should that come first?
+
+To achieve \emph{online in-NIC learning}, we return our focus to older \emph{classical} reinforcement learning methods.
+In particular, we suggest tile-coding, with one-step temporal-difference learning algorithms such as Sarsa or Q-learning.
 These choices have important benefits for in-NIC execution.
@@
 Tile-coding in particular admits many optimisations, being an embarrassingly parallel problem.
-Furthermore, gradient calculation is trivial for linear methods like tile coding, particularly when compared to the computational cost of backpropagation required by neural networks.
-Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair.
-
-?? Probably want some cites on the need for batching in NN methods, even though this is understood. DL book?
-
-\fakepara{To-be-organised}
-Interesting thing I realised from \textcite{DBLP:conf/sigcomm/NeugebauerAZAL018}: GPU and NIC might cause IOMMU contention?
-Making line-rate offload of DDN to GPU impractical.
-Unsure if I want to twin this with later measurements/plots.
-
-\subsection{Netronome Platform Fundamentals}
-Explain MEs, Threads, CTXs, so on so on.
+Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair, safeguarding the limited memory of the target devices.
 
@@
 If updates are enabled, \approachshort{} then checks an internal hashmap for a previous state-action pair matching the current instruction source, and if found then a policy update is performed.
-Updates are computed using \emph{single-step semi-gradient Sarsa}, though modification to support other single-step methods would be trivial.
+Updates are computed using \emph{single-step semi-gradient Sarsa} (\cref{eqn:sg-sarsa}), though modification to support other single-step methods would be trivial.
 The new state- or tiles-action pair is then written into storage.
@@
 \section{Evaluation}\label{sec:evaluation}
-We outline our performance evaluation of \approachshort.
-
-\subsection{Experimental Setup}
+We outline our performance evaluation of \approachshort...
+
+\subsection{Experimental Setup}\label{sec:experimental-setup}
 Evaluation of \approachshort{} was performed on server blades hosting a single Netronome Agilio LX 1x100GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
@@
 We compare these against floating point and quantised implementations running on a commodity host machine.
-This allows us to demonstrate the performance differences between the \emph{\indfw} and \emph{\coopfw} configurations, particularly in how \emph{\indfw}'s required policy locks impact throughput.
+This allows us to demonstrate the performance differences between the \indfw and \coopfw configurations, particularly in how \indfw's required policy locks impact throughput.
 We also report on the derived throughput measures accounting for the locking behaviour of each.
@@
 Moreover, to demonstrate how these costs vary as policy complexity increases, we vary the number of total dimensions included in the tiling.
-
-?? NOT GOT OTHERS TO COMPARE AGAINST YET, JUST SELF.
 
@@
 \fakepara{Raw inference and learning performance}
-\Cref{tab:lat-tput} shows that \Coopfw{} \approachshort{} achieves sub-\SI{50}{\micro\second} median latency, with \nth{99} percentile latencies less than \SI{1}{\micro\second} worse.
+\Cref{tab:lats} shows that \Coopfw{} \approachshort{} achieves sub-\SI{50}{\micro\second} median latency, with \nth{99} percentile latencies less than \SI{1}{\micro\second} worse.
 As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, while the simpler \Indfw{} succeeds when policies are fixed/known.
@@
 
-?? NOT GOT OTHERS TO COMPARE AGAINST YET, JUST SELF. NEED TO COMPARE AGAINST SOMEONE TO SHOW SUPREMACY
-
 \begin{table}
-	\caption{Latencies. On-device execution is crucial in not only lowering latencies, but in reducing tail latencies. Lower is better.\label{tab:lats}}
+	\caption{Latencies and computation times for \approachshort{} versus commodity hardware hosts.
+On-device execution is crucial in not only lowering latencies, but in reducing tail latencies.
+Lower is better, with the best marked \emph{in bold}.\label{tab:lats}}
 	\resizebox{\linewidth}{!}{
@@
 \begin{table}
-	\caption{Puts! Designs which do not take advantage..... Higher is better.\label{tab:tputs}}
+	\caption{Action and update throughputs for \approachshort{} versus commodity hardware hosts.
+Most designs cannot scale online performance with additional cores.
+Higher is better, with the best marked \emph{in bold}.\label{tab:tputs}}
 	\resizebox{\linewidth}{!}{
@@
 
+\Cref{tab:tputs} shows how \approachshort{} compares in terms of throughput with a numpy-based implementation of an RL policy.
+We choose the worker count on host machines with the largest number of parallel agents which produced a statistically significant throughput increase ($p < 0.05$).
+We observed that this equalled the number of physical cores on each device, while moving beyond this (even below the number of hyper-threads) would hamper tail latencies by an order of magnitude.
+To make the comparison fairer in the context of many-core CPU environments, we include per-core throughput.
+We observed that ...
+
 ?? \Cref{fig:vary-core} shows how action+update compute time scales with available cores.
-?? Limited by software.
+?? (Core count Limited by software).
 ?? Extra \emph{physical core} additions (rather than logical workers) very visible.
-?? 3 workers needed to process items faster than \Indfw{}, but 2 workers needed to bring \emph{latency} below \Indfw{}
+?? 11 workers needed to process items faster than \Indfw{}, but 5 workers needed to bring \emph{latency} below \Indfw{} ()
 ?? TAKEAWAY: resource demand tunable to meet latency demands for workload.
@@
 		}
-		\caption{Just Pars.}
+		\caption{\approachshort's \Coopfw{} design achieves consistent, tight latency bounds.}
 	\end{subfigure}
@@
 		}
-		\caption{Wider View.}
+		\caption{Tail latencies suffer in host execution---particularly when oversubscribed.}
 	\end{subfigure}
-	\caption{BigCap}
+	\caption{Cumulative state-action latency plots for \approachshort{} and host-based execution.}
 \end{figure}
@@
 	}
-	\caption{Action/update compute times in a \SI{32}{\bit} \emph{\Coopfw{}} agent under variable work schedulers.\label{fig:work-alloc-32}}
+	\caption{Action/update compute times in a \SI{32}{\bit} \Coopfw{} agent under variable work schedulers.\label{fig:work-alloc-32}}
 \end{figure}
+
+\begin{figure}
+	\resizebox{1.0\linewidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/vary-core-tput-on-per}
+	}
+	\caption{Throughput per added worker in a \Coopfw{} agent. \label{fig:tput-per-core}}
+\end{figure}
+
+\Cref{fig:tput-per-core}
+?? downticks in tput/core correspond to a `plateau' in throughput: the heuristic scheduler cannot find a better solution until extra cores are allocated
+?? Why? granularity of work items. Can't really subdivide high-dim policy items. Pushing past is tricky as described...
 
@@
 As one program is installed per ME, these represents the minimum and maximum resource use we can command on a single island (i.e., without replacing P4 workers).
-We observe negligible costs on shared bulk storage ($\sim$\SI{2}{\mebi\byte}).
+We observe negligible costs on shared bulk storage ($\sim$\SI{4}{\mebi\byte}), including storage space for a hash-table of \num{4096} state-action pairs and \num{16} reward values.
 The most significant costs arise due to policy data (\SI{405}{\kibi\byte} capacity shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
-Thread local storage in CLS for compute/register spilling scales with the number of MEs required, and with \emph{\Coopfw} this is far below the ``fair share'' (allowing co-existence with other asynchronous dataplane programs).
-Notably, \emph{\Indfw{}} imposes far heavier requirements on the limited per-island CLS, due to the extra resources needed to maintain, operate on, and cache tile lists for optimal processing.
+Thread local storage in CLS for compute/register spilling scales with the number of MEs required, and with \Coopfw{} this is far below the ``fair share'' (allowing co-existence with other asynchronous dataplane programs).
+Notably, \Indfw{} imposes far heavier requirements on the limited per-island CLS, due to the extra resources needed to maintain, operate on, and cache tile lists for optimal processing.
 
@@
 \fakepara{Deployability}
-?? Config times scale with work complexity in \Indfw, confirming the $\mathcal{O}{\left(n\right)}$ complexity of the allocator (??).
+??
+Setup and tiling packets take a mean \SI{26.65}{\micro\second} to be processed on \Indfw{}, while this processing and heuristic work scheduling takes \SI{417.25}{\micro\second} for the full-size policy described above.
+
+?? Config times scale with work complexity in \Coopfw{}, confirming the $\mathcal{O}{\left(n\right)}$ complexity of the allocator (??). Is it O mlogn?
 ?? Changing from online to offline has this cost.
 
-?? Anything else worth FW install time fairly constant (i.e., time to swap between bit-depths).
+We observed that firmware installation (i.e., changing from \Indfw{} to \Coopfw{}, bit-depth, or increasing maximum policy sizes) took a mean time of \SI{38.83}{\second}.
+?? What about compile time?
 
@@
 ?? Train offline, classical methods targetting all match-action P4 systems (no externs etc).
+
+\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}
 

### 2021-05-21 code/nfp/rl-test/worker/worker_body.c
Precalc work_idxes, tile idexes (~25% improvement)

@@ -29,12 +29,12 @@ __intrinsic uint32_t tile_code_with_cfg_single(
 	__addr40 __declspec(emem) tile_t *state,
 	__declspec(cls) struct rl_config *cfg,
 	uint8_t bias_tile_exists,
-	uint8_t work_idx
+	uint8_t work_idx,
+	uint16_t tiling_set_idx,
+	uint16_t tiling_idx
 ) {
 	uint16_t dim_idx;
-	uint16_t tiling_set_idx;
 	uint32_t local_tile;
-	uint16_t tiling_idx;
 	uint32_t width_product = cfg->num_actions;
 
 	tile_t emergency_vals[4] = {0};
@@ -46,16 +46,8 @@ __intrinsic uint32_t tile_code_with_cfg_single(
 		return 0;
 	}
 
-	tiling_set_idx = bias_tile_exists
-		? (((work_idx - 1) / cfg->tilings_per_set) + 1)
-		: (work_idx / cfg->tilings_per_set);
-
 	// can also get by remul'ing
 	local_tile = cfg->tiling_sets[tiling_set_idx].start_tile;
-	tiling_idx = work_idx - (bias_tile_exists
-		? 1 + ((tiling_set_idx-1) * cfg->tilings_per_set)
-		: tiling_set_idx * cfg->tilings_per_set);
-		// work_idx - local_tile;
 
 	local_tile += tiling_idx * cfg->tiling_sets[tiling_set_idx].tiling_size;
 
@@ -494,6 +486,9 @@ void work(uint8_t is_master, unsigned int parent_sig) {
 	uint16_t work_idxes[RL_MAX_TILE_HITS/2] = {0};
 	uint32_t tile_hits[RL_MAX_TILE_HITS/2] = {0};
 
+	uint16_t tiling_set_idxes[RL_MAX_TILE_HITS/2] = {0};
+	uint16_t tiling_idxes[RL_MAX_TILE_HITS/2] = {0};
+
 	uint32_t last_work_t;
 	uint32_t b_t0;
 	uint32_t b_t1;
@@ -637,7 +632,14 @@ void work(uint8_t is_master, unsigned int parent_sig) {
 						: TILE_LOCATION_T3 + 1;
 					//uint32_t t0 = local_csr_read(local_csr_timestamp_low);
 					//uint32_t t1;
[…]

### 2021-05-21 code/nfp/rl-test/rl.c
Make new optimisations safer/bounded to precache
There's now always a fallback option; though I do lose out on 0.6us due to branching... Better safe than sorry.

@@ -651,7 +651,7 @@ main() {
 					rl_pkt_return_slot(&rl_actions, workq_dump.state);
 				}
 				#endif /* !ENABLE_STRESS_TEST */
-				
+
 				break;
 			case PIF_PARREP_TYPE_in_reward:
 				pt._pad = workq_read_register.parsed_fields.reward.measured_value;
@@ -736,6 +736,12 @@ main() {
 			config_cycle_estimate = t1 - t0;
 			full_config_cycle_estimate = t2 - t0;
 		}
+
+		#else
+
+		t2 = t1;
+		config_cycle_estimate = t1 - t0;
+		full_config_cycle_estimate = t2 - t0;
 
 		#endif /* !_RL_WORKER_DISABLED */
 	}

### 2021-05-21 code/nfp/rl-test/rl.h

@@ -40,7 +40,7 @@ enum tile_location {
 // maximum number of tiles that can be returned as active
 // calls to `tile_code` must have at least this size.
 #define RL_MAX_TILE_HITS MAX_TILINGS_PER_SET*RL_MAX_SETS
-#define RL_MAX_PRECACHE 8
+#define RL_MAX_PRECACHE (RL_MAX_TILE_HITS/4)
 
 // I don't think there's a conveneient way to macro this
 #define T1_MAX_TILES_PER_TILING MAX_TILES_PER_DIM

### 2021-05-21 code/nfp/rl-test/worker/worker_body.c

@@ -225,6 +225,8 @@ __intrinsic void action_preferences_with_cfg_single(
 			loc++;
 			loc_base = cfg->last_tier_tile[loc];
 		}
+	} else {
+		__critical_path(90);
 	}
 
 	// WORKER_BARGAIN_BUCKET_SIMD
@@ -252,6 +254,8 @@ __intrinsic void action_preferences_with_cfg_single(
 		case TILE_LOCATION_T3:
 			base = tile_index - cfg->last_tier_tile[loc-1];
 			break;
+		default:
+			__impossible_path();
 	}
 
 	while (j < act_count) {
@@ -384,10 +388,13 @@ __intrinsic void update_action_preference_with_cfg_single(
 
 	if (loc > TILE_LOCATION_T3) {
 		loc = TILE_LOCATION_T1;
-		while (target >= loc_base) {
+		loc_base = cfg->last_tier_tile[loc];
+		while (tile_index >= loc_base) {
 			loc++;
 			loc_base = cfg->last_tier_tile[loc];
 		}
+	} else {
+		__critical_path(90);
 	}
 
 	// find location of tier in memory.
@@ -404,6 +411,8 @@ __intrinsic void update_action_preference_with_cfg_single(
 			base = target - cfg->last_tier_tile[loc-1];
 			t3_tiles[base + action] += delta;
 			break;
+		default:
+			__impossible_path();
 	}
 }
 
@@ -483,11 +492,10 @@ void work(uint8_t is_master, unsigned int parent_sig) {
 	uint8_t allocs_with_spill = 0;
 	uint8_t has_bias = 0;
 	uint8_t iter = 0;
-	uint16_t work_idxes[RL_MAX_TILE_HITS/2] = {0};
-	uint32_t tile_hits[RL_MAX_TILE_HITS/2] = {0};
+	uint16_t work_idxes[RL_MAX_TILE_HITS] = {0};
 
-	uint16_t tiling_set_idxes[RL_MAX_TILE_HITS/2] = {0};
-	uint16_t tiling_idxes[RL_MAX_TILE_HITS/2] = {0};
+	uint16_t tiling_set_idxes[RL_MAX_PRECACHE] = {0};
+	uint16_t tiling_idxes[RL_MAX_PRECACHE] = {0};
 
 	uint32_t last_work_t;
 	uint32_t b_t0;
@@ -627,18 +635,34 @@ void work(uint8_t is_master, unsigned int parent_sig) {
[…]

### 2021-05-21 paper/paper.tex
Shorter tables (and stop making Windows cry)

@@
 ?? We are the first to bring fully online RL to the dataplane!
+
+?? Emphasise here and elsewhere: scales with core count, but with some designs (\Indfw) achieve better latency and per-core throughput with just one core/worker!
 \end{abstract}
@@
 
+?? TODO: trim down. Too ``thesis-y'', rather than ``conference-y''.
+?? Save this elsewhere 'til then?
+
 \subsection{Programmable hardware capabilities}
@@
 This pattern extends to other SmartNICs, such as NVIDIA BlueField~\parencite{nvidia-bluefield} and Xilinx Alveo~\parencite{xilinx-alveo} devices.
-Even Intel Tofino~\parencite{barefoot-intel} ASICs, which architecturally mirror the PSA, expose additional matching capabilities and ALU functions via the Tofino Native Architecture.
+
+Currently, low port density devices like the above are most likely to have specific general-purpose compute, as they are designed to suit high-performance offloading and middlebox development.
+Even Intel Tofino~\parencite{barefoot-intel} ASICs, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
 Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of deep learning accelerators (which, in turn, lack packet switching functionality).
@@
 \begin{insight}
-	Low-port density programmable network devices often have spare resources and extra capabilities beyond the PSA specification which can aid asynchronous, local compute.
+	Current low-port density programmable network devices often have spare resources and extra capabilities beyond the PSA specification which can aid asynchronous, local compute.
 \end{insight}
@@
 	\end{subfigure}
+	\hspace{0.05\linewidth}
 	\begin{subfigure}{0.45\linewidth}
@@
 
-The solution we propose is to make use of the recent wave of programmable network devices to \emph{bring reinforcement learning to the data plane}---placing state measurement ($t_1$), low-cost decision-making processes ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
+The solution we propose is to make use of the recent wave of programmable network devices to \emph{bring reinforcement learning to the data plane}---referring again to \cref{fig:state-slip}, we would place place state measurement ($t_1$), low-cost decision-making processes ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
 In networks, actions are most likely to be installed on backbone switches, bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts.
@@
 A state vector produces a single `hit' in each tiling, which provides both the set of action preference lists to sum during inference \emph{and} the set of tiles to update at the next timestep.
-Crucially, there are no data dependencies between calculating any tile hits for an input state vector.
+Crucially, this internal representation ensures that there are no data dependencies between any tile hit calculations for an input state vector.
 This has important benefits for parallelisation.
@@
 At a high level, \approachshort{} is designed to use the auxiliary compute exposed by general SmartNIC devices to offer low-latency online learning, scaling according to available on-chip resources at build time.
+As the allocation of cores/chip area is set ahead of time by a framework or system administrator, \approachshort{} (\Coopfw) agents enumerate themselves at runtime initialisation.
 
@@
 
-\approachshort{} runs on one more cores of a SmartNIC to convert fixed-point state measurements from the environment into a stream of actions using a stored policy.
+\approachshort{} runs on one or more cores of a SmartNIC to convert fixed-point state measurements from the environment into a stream of actions using a stored policy.
 As an example, this might be to map flow state and performance measurements into a queue priority for future packets from that flow, or to compute and apply a rate limit to preserve quality of service.
@@
 ?? Probably mention the idea of including "state keys" and "reward keys" in the input state vector.
+
+?? Move a bunch of this to the NFP basics chapter.
 
@@
 
+?? While we \emph{could} use bitshifts in tile coding, we don't due to the strict limits it places upon policy design.
+
 \subsection{Agent-Environment Communication}\label{sec:agent-environment-communication}
@@
 Even with embarrassingly parallel problems, optimising for latency requires meticulous care in how work is passed out and aggregated.
-This is truer still when moving from the moderately fine-grained, efficient control of classical RL methods ($\sim$\SI{2}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
+This is truer still when moving from the moderately fine-grained, efficient control of classical RL methods ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
 Ordinarily, the serialisation/marshalling of requests, responses, and safe shared data access can incur significant extra overheads.
@@
 Moreover, physically adjacent functional units may have special-purpose shared registers or share a small fast cache to accelerate communication further.
+
+?? Move much of the below to the NFP basics section?
 
@@
 We use a simple first-fit work placement algorithm run on the NFP, placing the largest work item into the least loaded thread of the least loaded core.
-The cost of any individual work item (according to its dimension count and memory location) was estimated offlin.
+The cost of any individual work item (according to its dimension count and memory location) was estimated offline.
 This work allocation and any cached per-work-item values are recomputed when any configuration is installed or changed.
+Naturally, for $n$ tasks and $m$ threads this is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per task, storing $m/8$ and $\le8$ costs respectively.
 
@@
 
+?? Need to mention that this is a theoretical improvement; the results are basically opposite to this.
+?? Main benefit is then, I guess, the redux in storage requirements?
+
 %Potential: low-latency train for some specific points, and high throughput for others? Think on the exact note I took...
@@
 Platforms such as Intel Tofino greatly simplify this, with Tofino Native Architecture intrinsics such as \emph{Action Profiles/Selects} allowing a P4 action to be chosen based on a register value (i.e., an RL action).
+
+?? Achieving very tight latency bounds on \Coopfw{} requires quite a decent core allocation. (i.e., need at least 11 workers to make a gain in online t'put, 5 cores needed to win in throughput).
 
@@
 
+?? Will these look better \emph{after} the results?
+
 \subsection{In-Network DDoS Defence}\label{sec:integ-1}
@@
 The solution to implement and improve upon this work using \approachshort{} is to place its RL agents on SmartNICs at AS edge nodes---a bump-in-the-wire deployment.
+
+?? Any way to move some of this sooner? I.e., this is ``why move DDoS prevention here'', Stefanos suggested backporting some of this to justify ``why \emph{RL} in-NIC''?
 
@@
 	\resizebox{\linewidth}{!}{
-		\expandableinput ../tables/build/latency
+		\expandableinput ../tables/build/latency-only32
 	}
@@
 	\resizebox{\linewidth}{!}{
-		\expandableinput ../tables/build/tput
+		\expandableinput ../tables/build/tput-only32
 	}
@@
 
+?? Urgently describe and explain why lower bitdepth gives *worse* metrics.
+?? Mainly, register width.
+?? We also do stuff to batch read as many action vals as possible!! And yet this happens!!
+?? Also explain that Integer math does worse due to it falling back to BLAS in numpy (i.e., a float optimiser).
+?? We might have good int results if I do it in Rust?
+
 \fakepara{Work allocation}
@@
 ?? Future work: NetFPGA it.
+
+?? Future directions for community: want to see more general-purpose async compute on-chip! Larger form-factors!
+
+?? Are there other ``old ways of doing things'' worth looking at?
 

### 2021-05-21 tables/build/latency-only32.tex
Shorter tables (and stop making Windows cry)

\begin{tabular}{@{}ccS[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]@{}}
\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{3}{c}{State-Action Latency (\si{\micro\second})} & \multicolumn{3}{c}{State-Update Time (\si{\micro\second})}\\
\cmidrule(lr){3-5}\cmidrule(lr){6-8} & & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{\nth{99}} & \multicolumn{1}{c}{\num{99.99}\nthscript{th}} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{\nth{99}} & \multicolumn{1}{c}{\num{99.99}\nthscript{th}}\\ \midrule
Float & Collector & 515.94 & 606.06 & 725.03 & 606.06 & 636.82 & 833.99\\
 & MidServer & 1069.07 & 1125.1 & 1508.0 & 1260.04 & 1605.99 & 1719.864\\
Int32 & \approachshort{}-\Indfw{} & 185.040 & 185.373 & 185.707 & 230.707 & 231.200 & 231.633\\
 & \approachshort{}-\Coopfw{} & \bfseries 44.133 & \bfseries 44.280 & \bfseries 44.320 & \bfseries 81.287 & \bfseries 81.760 & \bfseries 82.447\\
\bottomrule
\end{tabular}

### 2021-05-21 tables/build/tput-only32.tex
Shorter tables (and stop making Windows cry)

\begin{tabular}{@{}ccS[table-format=2]S[table-format=3.3]S[table-format=2.3]S[table-format=1.3]S[table-format=1.3]@{}}
\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{1}{c}{Workers} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{2}{c}{Throughput/core (k actions/s)}\\
\cmidrule(lr){4-5}\cmidrule(lr){6-7} & & & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online} & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online}\\ \midrule
Float & Collector & 4 & 7.673 \pm 0.049 & 1.627 \pm 0.031 & 1.918 \pm 0.012 & \multicolumn{1}{c}{---}\\
 & MidServer & 6 & 5.584 \pm 0.030 & 0.791 \pm 0.012 & 0.931 \pm 0.005 & \multicolumn{1}{c}{---}\\
Int32 & \approachshort{}-\Indfw{} & 32 & \bfseries 172.956 \pm 0.171 & 4.335 \pm 0.005 & \bfseries 5.405 \pm 0.005 & \multicolumn{1}{c}{---}\\
 & \approachshort{}-\Coopfw{} & 32 & 22.670 \pm 0.073 & \bfseries 12.308 \pm 0.043 & 0.708 \pm 0.002 & \bfseries 0.385 \pm 0.001\\
\bottomrule
\end{tabular}

### 2021-05-21 code/rl-perf-tester/src/main.rs

@@ -224,7 +224,8 @@ fn run_core<'a>(
 		// TODO: make folder for install times.
 		// TODO: make files with current date in 'em.
 		let out_dir = format!("{}/setup-times/", OUTPUT_DIR);
-		let time_str = chrono::offset::Utc::now().to_rfc3339();
+		let time = chrono::offset::Utc::now();
+		let time_str = time.format("%FT %H %M %S");
 
 		let _ = std::fs::create_dir_all(&out_dir);
 

### 2021-05-21 tables/src/latency-only32.py

@@ -1,0 +1,115 @@
+import numpy as np
+import sys
+
+sys.stdout.reconfigure(encoding='utf-8')
+
+# sub in: bit depth, fw-type, measure_name
+float_results_fmt_str = "../results/host-rl-perf-tput/SUMMARY.f32.{}.{}.dat"
+results_fmt_str = "../results/rl-perf-tester/vary-work-ct/{}/{}.28d.31c.{}.dat"
+
+cycle_in_ns = 1e9 / 1.2e9
+
+core_ct_for_tput = 32
+types = [("single", R"\Indfw{}", True), ("balanced", R"\Coopfw{}", False)]
+measures = [("ComputeAndWriteout", R"\xmark", False),("UpdateAll", R"\cmark", True)]
+bitdepths = [32]#, 16, 8]
+
+print(R"\begin{tabular}{@{}ccS[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]@{}}")
+print(R"\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{3}{c}{State-Action Latency (\si{\micro\second})} & \multicolumn{3}{c}{State-Update Time (\si{\micro\second})}\\")
+
+# Just generate the next header line...
+t2_header = R"\cmidrule(lr){3-5}\cmidrule(lr){6-8} &"
+
+measure_names = ["Median", R"\nth{99}", R"\num{99.99}\nthscript{th}"]
+
+for i in range(2):
+	for name in measure_names:
+		t2_header += R" & \multicolumn{1}{c}{"+name+R"}"	
+
+t2_header += R"\\ \midrule"
+
+best_row_for_each_column = [3, 3, 3, 3, 3, 3]
+curr_row = 0
+
+print(t2_header)
+
+# Deal with Float results here.
+# File names like: results/host-rl-perf-tput/SUMMARY.f32.beast1.ComputeAndWriteout.dat
+# These results are already in micros.
+
+machines = [("Collector", "beast2", 4), ("MidServer", "perlman", 6)]
+
+for i, (machine_pres, machine_name, core_limit) in enumerate(machines):
+	out = "{} & {}".format("Float" if i == 0 else "", machine_pres)
+
+	latencies = []
+
+	for online_i, (online_fname_part, online_latex_part, needs_lock) in enumerate(measures):
[…]

### 2021-05-21 tables/src/tput-only32.py

@@ -1,0 +1,167 @@
+import numpy as np
+import sys
+
+sys.stdout.reconfigure(encoding='utf-8')
+
+# sub in: bit depth, fw-type, measure_name
+float_results_fmt_str = "../results/host-rl-perf-tput/SUMMARY.f32.{}.{}.dat"
+results_fmt_str = "../results/rl-perf-tester/vary-work-ct/{}/{}.28d.31c.{}.dat"
+
+cycle_in_ns = 1e9 / 1.2e9
+
+core_ct_for_tput = 32
+types = [("single", R"\Indfw{}", True), ("balanced", R"\Coopfw{}", False)]
+measures = [("ComputeAndWriteout", "Offline", False),("UpdateAll", "Online", True)]
+bitdepths = [32]#, 16, 8]
+
+print(R"\begin{tabular}{@{}ccS[table-format=2]S[table-format=3.3]S[table-format=2.3]S[table-format=1.3]S[table-format=1.3]@{}}")
+print(R"\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{1}{c}{Workers} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{2}{c}{Throughput/core (k actions/s)}\\")
+
+# Just generate the next header line...
+t2_header = R"\cmidrule(lr){4-5}\cmidrule(lr){6-7} & &"
+
+for i in range(2):
+	for (_f, name, _n) in measures:
+		t2_header += R" & \multicolumn{1}{c}{"+name+R"}"	
+
+t2_header += R"\\ \midrule"
+
+print(t2_header)
+
+best_row_for_each_column = [2, 3, 2, 3]
+curr_row = 0
+
+# Deal with Float results here.
+# File names like: results/host-rl-perf-tput/SUMMARY.f32.beast1.ComputeAndWriteout.dat
+# These results are already in micros.
+
+machines = [("Collector", "beast2", 4), ("MidServer", "perlman", 6)]
+
+for i, (machine_pres, machine_name, core_limit) in enumerate(machines):
+	out = "{} & {}".format("Float" if i == 0 else "", machine_pres)
+
+	best_core = None
+	tputs = []
+	stds = []
+	core_cts = []
+
+	for online_i, (online_fname_part, online_latex_part, needs_lock) in enumerate(measures):
+		fname = float_results_fmt_str.format(machine_name, online_fname_part)
+
+		best_core_ct = 0
+		best_dats = None
[…]

### 2021-05-24 paper/paper.tex
Light perf-tester refactoring, fw-install server support.

@@
 
+?? NOTE: this aggregastion step is specifically WAIT-FREE, not just lockless etc. because progress always made.
+
 ?? Move much of the below to the NFP basics section?
@@
 We additionally use a more powerful consumer-grade machine (\emph{Collector}) for estimating host performance when offloaded to a network function, running Ubuntu 18.04.4 LTS (4.15.0-96-generic) on Intel Core i7-6700K ($4\times$\SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} RAM.
-Control programs were built using rustc 1.51.0.
+Control programs were built using rustc 1.52.1.
 We run \approachshort{} on an island with \num{4} MEs of the NFP-6480 for a total of \num{32} hardware threads.
@@
 Where feasible, we test \SIlist{32;16;8}{\bit} quantised arithmetic firmware versions.
-All timing and throughput measurements were repeated over \num{5000} trials.
+All \approachshort{} timing measurements were repeated over \num{10000} trials (preceded by \num{1000} warmup packets), from which throughput was derived.
+Host throughput measures were observed over \num{10} trials of \SI{10}{\second}.
+?? verify these numbers against code?
 

### 2021-05-24 code/rl-perf-tester/src/config.rs

@@ -13,3 +13,7 @@ pub struct Config<'a> {
 	pub skip_setup: bool,
 	pub skip_retiling: bool,
 }
+
+pub struct HostConfig {
+	pub rtecli_path: String,
+}

### 2021-05-24 code/rl-perf-tester/src/lib.rs

@@ -1,8 +1,9 @@
 mod config;
 mod constants;
 mod experiment;
+mod stress;
 
-pub use self::{config::*, constants::*, experiment::*};
+pub use self::{config::*, constants::*, experiment::*, stress::*};
 
 use control::{GlobalConfig, SendStateConfig, Setup, SetupConfig, Tile, TilingsConfig};
 use core::fmt::Debug;
@@ -12,7 +13,7 @@ use rand_chacha::ChaChaRng;
 use serde::{de::DeserializeOwned, Serialize};
 use std::{
 	fs::File,
-	io::Write,
+	io::{Error as IoError, ErrorKind as IoErrorKind, Result as IoResult, Write},
 	process::Command,
 	thread,
 	time::{Duration, Instant},
@@ -124,38 +125,13 @@ fn run_experiment_with_datatype<T>(
 		}
 
 		if !config.skip_firmware_install {
-			// NOTE: I do this here to try and mitigate any randomness between sets foe now.
-			// And/or strange bugs that lead to permanent early exit.
-			eprint!("\tInstalling firmware... ");
-			std::io::stderr().flush().unwrap();
+			let fw_install_time = install_fw(
+				config.rtecli_path,
+				&in_file,
+				&format!("{}/{}bit/{}.json", FIRMWARE_DIR, bit_depth, fw.fw_name())[..],
+			)
+			.expect("Failed to install firmware.");
 
-			let fw_t0 = Instant::now();
-			let cmd_output = Command::new(config.rtecli_path)
-				.args(&[
-					"design-load",
-					"-f",
-					&in_file[..],
-					"-p",
-					&format!("{}/{}bit/{}.json", FIRMWARE_DIR, bit_depth, fw.fw_name())[..],
-					"-c",
-					P4CFG_PATH,
-				])
-				.output()
-				.expect("Failed to install firmware.");
-			let fw_t1 = Instant::now();
-
-			if !cmd_output.status.success() {
-				panic!(
-					"Failed to install firmware (code {:?}): {:?} {:?}",
-					cmd_output.status.code(),
-					std::str::from_utf8(&cmd_output.stdout[..]),
-					std::str::from_utf8(&cmd_output.stderr[..]),
-				);
-			}
-
-			eprintln!("Installed!");
-
-			let fw_install_time = fw_t1 - fw_t0;
[…]

### 2021-05-24 code/rl-perf-tester/src/main.rs

@@ -1,10 +1,25 @@
 use clap::{App, Arg, ArgMatches, SubCommand};
 use control::TransportConfig;
 use core::iter::Iterator;
-use rl_perf_tester::{Config, ExperimentFile, InstallTimes, OUTPUT_DIR, RTE_PATH, RTSYM_PATH};
-use std::{fs::File, io::{BufReader, Write}};
+use rl_perf_tester::{
+	Config,
+	ExperimentFile,
+	HostConfig,
+	InstallTimes,
+	OUTPUT_DIR,
+	RTE_PATH,
+	RTSYM_PATH,
+};
+use std::{
+	error::Error,
+	fs::File,
+	io::{BufReader, Write},
+};
 
-fn main() {
+fn main() -> Result<(), Box<dyn Error>> {
+	#[cfg(unix)]
+	sudo::escalate_if_needed()?;
+
 	let run_base = SubCommand::with_name("run")
 		.about("Runs the given experiments.")
 		.arg(
@@ -122,6 +137,33 @@ fn main() {
 			),
 		)
 		.subcommand(run_base.clone().name("run-all"))
+		.subcommand(
+			SubCommand::with_name("stress-host")
+				.arg(
+					Arg::with_name("port")
+						.short("p")
+						.long("port")
+						.value_name("PORT")
+						.help("Port number to host the WebSocket server on.")
+						.takes_value(true)
+						.default_value("3188"),
+				)
+				.arg(
+					Arg::with_name("rtecli-path")
+						.short("c")
+						.long("rtecli-path")
+						.value_name("PATH")
+						.help("Path to the executable `rtecli`.")
+						.takes_value(true)
+						.default_value(RTE_PATH),
+				)
+				.about("Host a server for firmware installation and management on the device under test.")
+		)
+		.subcommand(
+			SubCommand::with_name("stress-client")
+				.arg(Arg::with_name("blah"))
+				.about("Host a server for firmware installation and management on the device under test.")
+		)
 		.get_matches();
 
 	match matches.subcommand() {
@@ -145,10 +187,30 @@ fn main() {
 				println!("No experiment names given.");
 				println!("{}", matches.usage());
 			},
+		("stress-host", Some(sub_m)) => {
+			let port = sub_m
+				.value_of("port")
+				.and_then(|p_str| p_str.parse().ok())
+				.expect("Invalid source port (u16).");
+
+			let config = HostConfig {
+				rtecli_path: matches
[…]

### 2021-05-24 code/rl-perf-tester/src/stress/host.rs

@@ -1,0 +1,106 @@
+use super::protocol::*;
+
+use crate::{HostConfig, FIRMWARE_DIR};
+use std::{
+	io::ErrorKind as IoErrorKind,
+	net::TcpListener,
+	sync::{
+		atomic::{AtomicBool, Ordering},
+		Arc,
+	},
+};
+use tungstenite::error::Error as WsError;
+
+pub fn host_stress_server(port: u16, config: HostConfig) {
+	let listener = TcpListener::bind(("127.0.0.1", port)).unwrap();
+	let busy = Arc::new(AtomicBool::new(false));
+
+	let cfg = Arc::new(config);
+
+	for stream in listener.incoming() {
+		if let Ok(stream) = stream {
+			let l_busy = busy.clone();
+			stream.set_nonblocking(true).unwrap();
+
+			let cfg = Arc::clone(&cfg);
+
+			std::thread::spawn(move || {
+				let mut websocket = tungstenite::server::accept(stream).unwrap();
+
+				let was_busy = l_busy.fetch_or(true, Ordering::Relaxed);
+
+				if was_busy {
+					// Send Message.
+					let _ = websocket.write_message(for_ws(&HostToClient::Fail(FailReason::Busy)));
+					// Close Conn.
+					let _ = websocket.close(None);
+					return;
+				}
+
+				let (tx, rx) = flume::unbounded();
+
+				loop {
+					match rx.try_recv() {
+						Ok(None) => {
+							let _ = websocket.write_message(for_ws(&HostToClient::Success));
+						},
+						Ok(Some(mistake)) => {
+							let _ = websocket.write_message(for_ws(&HostToClient::Fail(mistake)));
+						},
+						_ => {},
+					}
+
+					match websocket.read_message() {
+						Ok(msg) =>
+							if let Ok(text) = msg.to_text() {
+								match serde_json::from_str::<ClientToHost>(text) {
+									Ok(ClientToHost::Fw(f)) => {
+										let my_tx = tx.clone();
+										let my_cfg = Arc::clone(&cfg);
+										std::thread::spawn(move || {
+											install_fw(my_tx, f, &my_cfg);
+										});
+									},
+									_ => {
+										let _ = websocket
+											.write_message(for_ws(&HostToClient::IllegalRequest));
+										let _ = websocket.close(None);
+
+										break;
+									},
+								}
+							},
[…]

### 2021-05-24 code/rl-perf-tester/src/stress/mod.rs

@@ -1,0 +1,4 @@
+mod host;
+pub mod protocol;
+
+pub use self::host::*;

### 2021-05-24 code/rl-perf-tester/src/stress/protocol.rs

@@ -1,0 +1,30 @@
+use serde::{Deserialize, Serialize};
+use tungstenite::protocol::Message;
+
+#[derive(Clone, Deserialize, Serialize)]
+pub enum ClientToHost {
+	Fw(FwInstall),
+}
+
+#[derive(Clone, Deserialize, Serialize)]
+pub enum HostToClient {
+	Success,
+	Fail(FailReason),
+	IllegalRequest,
+}
+
+#[derive(Clone, Deserialize, Serialize)]
+pub enum FailReason {
+	Busy,
+	Generic(String),
+}
+
+#[derive(Clone, Deserialize, Serialize)]
+pub struct FwInstall {
+	pub name: String,
+	pub p4cfg_override: Option<String>,
+}
+
+pub fn for_ws(msg: &impl Serialize) -> Message {
+	Message::Text(serde_json::to_string(msg).unwrap())
+}

### 2021-05-24 code/rl-perf-tester/src/config.rs
All the machinery *around* pktgen automation...

@@ -17,3 +17,14 @@ pub struct Config<'a> {
 pub struct HostConfig {
 	pub rtecli_path: String,
 }
+
+pub struct StressConfig<'a> {
+	pub transport_cfg: TransportConfig,
+	pub host_server: &'a str,
+	pub port: u16,
+	pub pci_addr: &'a str,
+	pub bind_driver: &'a str,
+	pub min_rate: u32,
+	pub max_rate: u32,
+	pub num_trials: u32,
+}

### 2021-05-24 code/rl-perf-tester/src/main.rs

@@ -6,6 +6,7 @@ use rl_perf_tester::{
 	ExperimentFile,
 	HostConfig,
 	InstallTimes,
+	StressConfig,
 	OUTPUT_DIR,
 	RTE_PATH,
 	RTSYM_PATH,
@@ -148,21 +149,116 @@ fn main() -> Result<(), Box<dyn Error>> {
 						.takes_value(true)
 						.default_value("3188"),
 				)
-				.arg(
-					Arg::with_name("rtecli-path")
-						.short("c")
-						.long("rtecli-path")
-						.value_name("PATH")
-						.help("Path to the executable `rtecli`.")
-						.takes_value(true)
-						.default_value(RTE_PATH),
-				)
 				.about("Host a server for firmware installation and management on the device under test.")
 		)
 		.subcommand(
 			SubCommand::with_name("stress-client")
 				.arg(Arg::with_name("blah"))
-				.about("Host a server for firmware installation and management on the device under test.")
+				.arg(
+					Arg::with_name("HOST")
+						.help("Address of the host machine for the target firmware. This must be addressable over the local control plane.")
+						.takes_value(true)
+						.min_values(1)
+						.required(true),
+				)
+				.arg(
+					Arg::with_name("ws-port")
+						.short("w")
+						.long("ws-port")
+						.value_name("PORT")
+						.help("Port number to conenct to the WebSocket server.")
+						.takes_value(true)
+						.default_value("3188"),
+				)
+				.arg(
+					Arg::with_name("src-ip")
+						.short("s")
+						.long("src-ip")
+						.value_name("IPV4")
+						.help("Source IP for control packets. Defaults to selected interface.")
+						.takes_value(true)
+						.default_value("192.168.0.4"),
+				)
+				.arg(
+					Arg::with_name("dst-ip")
+						.short("d")
+						.long("dst-ip")
+						.value_name("IPV4")
+						.help("Destination IP for control packets. Defaults to selected interface.")
+						.takes_value(true)
+						.default_value("192.168.1.141"),
+				)
+				.arg(
+					Arg::with_name("src-port")
+						.short("p")
+						.long("src-port")
+						.value_name("U16")
+						.help("Source UDP port for control packets.")
+						.takes_value(true)
+						.default_value("16767"),
+				)
+				.arg(
+					Arg::with_name("dst-port")
+						.short("P")
+						.long("dst-port")
+						.value_name("U16")
[…]

### 2021-05-24 code/rl-perf-tester/src/stress/client.rs

@@ -1,0 +1,159 @@
+use super::protocol::*;
+
+use control::{GlobalConfig, Setup, SetupConfig, TilingsConfig};
+use crate::{ProtoSetup, StressConfig};
+use std::{error::Error, io::Result as IoResult, process::Command, thread, time::Duration};
+use tungstenite::error::Error as WsError;
+
+type AnyRes<A> = Result<A, Box<dyn Error>>;
+
+pub fn run_stress_test(config: &StressConfig, if_name: &str) -> AnyRes<()> {
+	const PKT_SIZES: [usize; 10] = [64, 128, 256, 512, 1024, 1500, 2048, 4096, 8192, 9000];
+
+	for rate in config.min_rate..=config.max_rate {
+		eprintln!("Rate set to {}k pkts/sec!", rate);
+
+		unbind_and_reset_nfp(config)?;
+
+		eprintln!("Device {} unbound from DPDK!", config.pci_addr);
+
+		while tell_dut_to_install_fw(config, rate)? {
+			// Accidentally in use... so sleep it off.
+			eprintln!("DUT busy; retrying in 10s...");
+			thread::sleep(Duration::from_secs(10));
+		}
+
+		eprintln!("Firmware {}.0k installed!", rate);
+
+		thread::sleep(Duration::from_secs(1));
+
+		setup_dut(config, if_name)?;
+
+		eprintln!("Control packets sent over {}!", if_name);
+
+		thread::sleep(Duration::from_secs(3));
+
+		bind_nfp(config)?;
+
+		for i in 0..config.num_trials {
+			for pkt_sz in PKT_SIZES.iter() {
+				// TODO: Pktgen stuff.
+				eprintln!("Trial {} for {}B pkts would be here!", i, pkt_sz);
+			}
+		}
+	}
+
+	let _ = unbind_and_reset_nfp(config);
+
+	Ok(())
+}
+
+fn bind_nfp(config: &StressConfig) -> IoResult<()> {
+	Command::new("dpdk-devbind.py")
+		.args(&[&format!("--bind={}", config.bind_driver), config.pci_addr])
+		.output()
+		.map(|_| ())
+}
+
+fn unbind_and_reset_nfp(config: &StressConfig) -> IoResult<()> {
+	Command::new("dpdk-devbind.py")
+		.args(&["-u", config.pci_addr])
+		.output()?;
+
+	Command::new("rmmod").args(&["nfp"]).output()?;
[…]

### 2021-05-24 code/rl-perf-tester/src/stress/mod.rs

@@ -1,4 +1,5 @@
+mod client;
 mod host;
 pub mod protocol;
 
-pub use self::host::*;
+pub use self::{client::*, host::*};

### 2021-05-24 code/rl-perf-tester/src/main.rs
Fix to client mode...

@@ -153,7 +153,6 @@ fn main() -> Result<(), Box<dyn Error>> {
 		)
 		.subcommand(
 			SubCommand::with_name("stress-client")
-				.arg(Arg::with_name("blah"))
 				.arg(
 					Arg::with_name("HOST")
 						.help("Address of the host machine for the target firmware. This must be addressable over the local control plane.")

### 2021-05-25 code/rl-perf-tester/src/config.rs
Hook Pktgen running into the main control program.

@@ -27,4 +27,5 @@ pub struct StressConfig<'a> {
 	pub min_rate: u32,
 	pub max_rate: u32,
 	pub num_trials: u32,
+	pub dpdk_pktgen_home: &'a str,
 }

### 2021-05-25 code/rl-perf-tester/src/constants.rs

@@ -4,3 +4,4 @@ pub const FIRMWARE_DIR: &str = "../nfp/rl-test/fws";
 pub const RTE_PATH: &str = "/opt/netronome/p4/bin/rtecli";
 pub const P4CFG_PATH: &str = "../nfp/rl-test/testcfg.p4cfg";
 pub const RTSYM_PATH: &str = "/opt/netronome/bin/nfp-rtsym";
+pub const PKTGEN_PATH: &str = "/home/clouduser/dpdk/pktgen-dpdk";

### 2021-05-25 code/rl-perf-tester/src/main.rs

@@ -8,6 +8,7 @@ use rl_perf_tester::{
 	InstallTimes,
 	StressConfig,
 	OUTPUT_DIR,
+	PKTGEN_PATH,
 	RTE_PATH,
 	RTSYM_PATH,
 };
@@ -125,6 +126,15 @@ fn main() -> Result<(), Box<dyn Error>> {
 				.help("Path to the executable `nfp-rtsym`.")
 				.takes_value(true)
 				.default_value(RTSYM_PATH),
+		)
+		.arg(
+			Arg::with_name("pktgen-home-path")
+				.short("H")
+				.long("pktgen-home-path")
+				.value_name("PATH")
+				.help("Path to the base directory for `pktgen-dpdk`.")
+				.takes_value(true)
+				.default_value(PKTGEN_PATH),
 		)
 		.subcommand(SubCommand::with_name("list").about("List all bindable interfaces."))
 		.subcommand(SubCommand::with_name("expts").about("List all experiment names for `run`."))
@@ -246,7 +256,7 @@ fn main() -> Result<(), Box<dyn Error>> {
 						.value_name("U32")
 						.help("Maximum rate, inclusive, for FW selection (in k decisions/s).")
 						.takes_value(true)
-						.default_value("11"),
+						.default_value("16"),
 				)
 				.arg(
 					Arg::with_name("num-trials")
@@ -359,6 +369,10 @@ fn main() -> Result<(), Box<dyn Error>> {
 				.and_then(|p_str| p_str.parse().ok())
 				.expect("Number of trials required!");
 
+			let dpdk_pktgen_home = matches
+				.value_of("pktgen-home-path")
+				.expect("I really, really need to know where to find Pktgen-DPDK!");
+
 			let config = StressConfig {
 				transport_cfg: t_cfg,
 
@@ -371,6 +385,8 @@ fn main() -> Result<(), Box<dyn Error>> {
 				min_rate,
 				max_rate,
 				num_trials,
+
+				dpdk_pktgen_home,
 			};
 
 			rl_perf_tester::run_stress_test(&config, if_name)

### 2021-05-25 code/rl-perf-tester/src/stress/client.rs

@@ -14,8 +14,6 @@ use tungstenite::error::Error as WsError;
 type AnyRes<A> = Result<A, Box<dyn Error>>;
 
 pub fn run_stress_test(config: &StressConfig, if_name: &str) -> AnyRes<()> {
-	const PKT_SIZES: [usize; 10] = [64, 128, 256, 512, 1024, 1500, 2048, 4096, 8192, 9000];
-
 	for rate in config.min_rate..=config.max_rate {
 		eprintln!("Rate set to {}k pkts/sec!", rate);
 
@@ -41,17 +39,36 @@ pub fn run_stress_test(config: &StressConfig, if_name: &str) -> AnyRes<()> {
 
 		bind_nfp(config)?;
 
-		for i in 0..config.num_trials {
-			for pkt_sz in PKT_SIZES.iter() {
-				// TODO: Pktgen stuff.
-				eprintln!("Trial {} for {}B pkts would be here!", i, pkt_sz);
-			}
-		}
+		run_dpdk_expt_set(config, rate)?;
 	}
 
 	let _ = unbind_and_reset_nfp(config, if_name);
 
 	Ok(())
+}
+
+fn run_dpdk_expt_set(config: &StressConfig, rate: u32) -> IoResult<()> {
+	// sudo -E $PKTGEN_HOME/build/app/pktgen -n 6 -l 0-15 -- -m "{1}.0,{2:3}.1" -f opal-stress.lua
+
+	rpo(
+		Command::new(&format!("{}/build/app/pktgen", config.dpdk_pktgen_home))
+			.args(&[
+				"-n",
+				"6",
+				"-l",
+				"0-15",
+				"--",
+				"-m",
+				"\"{1}.0,{2:3}.1\"",
+				"-f",
+				"opal-stress.lua",
+			])
+			.env("PKTGEN_HOME", config.dpdk_pktgen_home)
+			.env("RL_TEST_STRESS_K", rate.to_string())
+			.env("RL_TEST_STRESS_ITERS", config.num_trials.to_string())
+			.output()?,
+		"run full DPDK experiment set",
+	)
 }
 
 fn bind_nfp(config: &StressConfig) -> IoResult<()> {

### 2021-05-26 paper/paper.tex
Hook Pktgen running into the main control program.
Experimental changes needing a re-run...
No spoolup period now; this should let me just sum ipackets and imissed to establish how many packet Rxs (thus losses) there really were. Also tries to get double the latency measures in the same time period.

@@
 \subsection{Variable Quantisation Bit Depth}
-At compile-time, \approachshort{} can be configured to use \SIlist[list-final-separator = { \translate{or} }]{32;16;8}{\bit} values in input states and its internal policies.
+At compile-time, \approachshort{} can be configured to use \SIlist[list-final-separator = { or }]{32;16;8}{\bit} values in input states and its internal policies.
 Naturally, smaller bit depths reduce the storage required for policies, tiling data, and stored state-action pairs---allowing more complex problems to be modelled using more dimensions or fine-grained policies.
@@
 Host throughput measures were observed over \num{10} trials of \SI{10}{\second}.
+
 ?? verify these numbers against code?
+
+?? NOTE: Ritchie is actually Intel(R) Xeon(R) Silver 4208 CPU (8 times 2.1GHz)
 

### 2021-05-26 code/rl-perf-tester/src/lib.rs

@@ -271,7 +271,7 @@ fn run_experiment_with_datatype<T>(
 }
 
 fn install_fw(rtecli_path: &str, in_file: &str, json_file: &str) -> IoResult<Duration> {
-	eprint!("\tInstalling firmware... ");
+	eprint!("\tInstalling firmware {}... ", in_file);
 	std::io::stderr().flush()?;
 
 	let fw_t0 = Instant::now();

### 2021-05-26 code/rl-perf-tester/src/stress/client.rs

@@ -105,7 +105,7 @@ fn run_dpdk_expt_set(config: &StressConfig, rate: u32) -> IoResult<()> {
 				"0-15",
 				"--",
 				"-m",
-				"{1}.0,{2:3}.1",
+				"{1}.0,{1:1-7}.1",
 				"-f",
 				"opal-stress.lua",
 			])

### 2021-05-26 results/process-stress.py

@@ -1,0 +1,93 @@
+import numpy as np
+import scipy.stats as spss
+
+stress_result_dir = "stress/"
+
+# pkt sz, stress level, iter
+stress_file_lat = "l-{}B-{}k-{}.dat"
+stress_file_tpu = "t-{}B-{}k-{}.dat"
+
+pkt_sizes = [64, 128, 256, 512, 1024, 1280, 1518];
+iter_ct = 10
+rate_max = 16 + 1
+
+def write_matrix(name, matrix):
+	pass
+
+print("Latencies!")
+
+# process latencies first.
+for pkt_size in pkt_sizes:
+	lats_baseline = None
+	for rate in range(rate_max):
+		lats_for_this_cell = []
+		for i in range(iter_ct):
+			with open(stress_result_dir + stress_file_lat.format(pkt_size, rate, i)) as of:
+				next(of)
+				lats_for_this_cell += [float(x.strip()) for x in of]
+
+		p_val = None
+		if lats_baseline is None:
+			lats_baseline = lats_for_this_cell
+		else:
+			(_t_val, p_val) = spss.stats.ttest_ind(lats_baseline, lats_for_this_cell, equal_var=False)
+
+		metrics = [
+			np.median(lats_for_this_cell),
+			np.percentile(lats_for_this_cell, 99.0),
+			np.percentile(lats_for_this_cell, 99.99),
+			np.mean(lats_for_this_cell),
+			np.std(lats_for_this_cell),
+			p_val,
+		]
+		print("{}B {}k: {}".format(pkt_size, rate, metrics))
+
+# { {
+#     ibytes = 28044446236,
+#     ierrors = 0,
+#     imissed = 36943107,
+#     ipackets = 226164889,
+#     obytes = 32625360380,
+#     oerrors = 0,
+#     opackets = 263107745,
+#     rx_nombuf = 0
+#   },
+#   n = 1
+# }
+
+def extract_from_line(line):
+	return float(line.split(",")[0].split("=")[1].strip())
+
+print("Loss rates!")
+
+# process mean losses?
+for pkt_size in pkt_sizes:
+	losses_baseline = None
+	for rate in range(rate_max):
+		losses_for_this_cell = []
+		for i in range(iter_ct):
+			with open(stress_result_dir + stress_file_tpu.format(pkt_size, rate, i)) as of:
+				lines = of.readlines()
+				imissed = extract_from_line(lines[3])
+				ipackets = extract_from_line(lines[4])
[…]

### 2021-05-29 paper/paper.tex
Today's paper work.

@@
 \usepackage{bm}
+
+\usepackage[ruled,vlined]{algorithm2e}
+
+% Get Booktab style ruling in algorithm blocks.
+% Values shamelessly taken from egreg's answer in 
+% https://tex.stackexchange.com/a/345745/82917
+\makeatletter
+\renewcommand*{\@algocf@pre@ruled}{\hrule height\heavyrulewidth depth0pt \kern\belowrulesep}
+\renewcommand*{\algocf@caption@ruled}{\box\algocf@capbox\kern\aboverulesep\hrule height\lightrulewidth\kern\belowrulesep}
+\renewcommand*{\@algocf@post@ruled}{\kern\aboverulesep\hrule height\heavyrulewidth\relax}
+\makeatother
 
@@
 %Conference
-\acmConference[SOSR '21]{Symposium of SDx Research 2021}{September \numrange{13}{14}, 2021}{Online}
+\acmConference[SOSR '21]{Symposium of SDx Research 2021}{September \numrange{20}{21}, 2021}{Online}
 \acmYear{2021}
@@
 
+?? Make a point of explaining why in-NIC... in general!
+
 \begin{figure}
@@
 \subsection{Netronome Platform Fundamentals}
-?? Explain MEs, Threads, CTXs, so on so on.
+As we implement this work on Netronome SmartNICs, it is necessary to explain the basics of the NFP architecture.
+NFP chips are \emph{system-on-a-chip} (SoC) devices, and achieve scalable packet processing through sheer parallelism.
+Aside from an ARM management chip and application specific functional units, most of the chip is composed of \emph{microengines} (MEs), grouped into \emph{islands} of 4 or 12 MEs.
+All 12-ME islands are used by a default P4 pipeline.
+Each ME has \numrange{4}{8} \emph{contexts} (hardware threads) which share a code store and large register file, offering zero-cost context switches triggered by I/O.
+Contexts and MEs may send one another numbered signals, and MEs have a small \emph{next-neighbour} register file for passing values in one direction to the next ME on the same island.
+MEs run a proprietary instruction set, compiled to via a \emph{(Micro-)C} compiler.
+Beyond registers, the platform implements an explicit memory hierarchy scaling in size, location, and access cost as below:
+$$\text{LMEM (ME)} < \text{CLS (Island)} < \text{CTM} < \text{IMEM (Chip)} < \text{EMEM}$$
 
@@
 \approachshort{} runs on one or more cores of a SmartNIC to convert fixed-point state measurements from the environment into a stream of actions using a stored policy.
-As an example, this might be to map flow state and performance measurements into a queue priority for future packets from that flow, or to compute and apply a rate limit to preserve quality of service.
+As an example, this might be to map flow state and performance measurements into a queueing priority for future packets from that flow, or to compute and apply a rate limit to preserve quality of service.
 These dedicated cores are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
@@
 
-Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
+Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match-action tables and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
 Bit depth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
@@
 The new state- or tiles-action pair is then written into storage.
-
-?? Probably mention the idea of including "state keys" and "reward keys" in the input state vector.
-
-?? Move a bunch of this to the NFP basics chapter.
-
-In the NFP architecture we use, physical cores (MEs) each have 8 hardware threads (contexts): all contexts on an ME share the same program and employ zero-cost context switches triggered by I/O.
-Each core then provides 8 worker threads, though in the \indfw{} case one must be dedicated to monitoring work acknowledgements.
+\approachshort{} can be configured to automatically select values of the input state vector as keys for state and reward storage.
+
 Two main firmware models govern how the computation-intensive parts of these tasks (tile-coding, action preference list construction, policy updates) are carried out:
@@
 If an update occurs, the core requests a \emph{write lock} before updating, greatly limiting online throughput.
-Tile lists are stored for update computation.
-	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel})] Threads cooperate on processing state vectors, minimising latency.
-Workers have a fixed list of work items, while a master thread sends compute/update commands before awaiting worker completion.
+\emph{Tile lists} are stored for update computation.
+	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel,alg:parsa})] Threads cooperate on processing state vectors, minimising latency.
+Minions have a fixed list of work items, while a controller thread sends compute/update commands before awaiting worker completion.
 Work items are disjoint, requiring no policy locks.
-State vectors are stored for update computation.
+\emph{State vectors} are stored for update computation.
 \end{description}
 Each offers a different point of optimisation; if updates are disabled, then the \emph{\indfw{}} model can maximise throughput, while the \emph{\coopfw{}} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
+We note that our parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is made possible only by the use of quantisation to integers, as atomic operations on floating point numbers typically do not exist outside of graphics- or tensor-processing hardware.
 
@@
 It transpired that cacheing this data placed a larger \texttt{memcpy} in the serial section, whose size did not scale at all with bit depth.
-
-?? While we \emph{could} use bitshifts in tile coding, we don't due to the strict limits it places upon policy design.
+Additionally, we do not use bitshifts in place of division operations in our implementation, due to the strict limits that power-of-two tile widths place upon policy design.
 
@@
 
-?? NOTE: this aggregastion step is specifically WAIT-FREE, not just lockless etc. because progress always made.
-
-?? Move much of the below to the NFP basics section?
-
 Our implementation exploits the locality of threads, cores, and their parent islands in the NFP architecture.
-Groups of \numrange{4}{12} cores (MEs) in this platform are termed \emph{islands}, with shared access to low-latency scratch memory and the adjacent registers of neighbouring MEs.
-Each ME then has a number of threads fixed at compile time (\numrange{4}{8}), sharing one large register file to enable zero-cost context switches.
 Policy compute/update tasks and configuration updates are passed between cores using these direct \emph{next neighbour} registers, signalling all child threads in response.
-Each task performs atomic adds to a shared preference list, and atomically increments an acknowledgement counter to be periodically checked by the master thread.
-
-In earlier designs we had experimented with bounded buffers as in \cref{sec:agent-environment-communication} for this, modified to be located solely in local scratch memory, dedicating the master thread to result aggregation.
+Each task performs atomic adds to a shared preference list, and atomically increments an acknowledgement counter to be periodically checked by the master thread, implementing the wait-free \emph{ParSa} algorithm we introduce (\cref{alg:parsa}).
+
+In earlier designs we had experimented with bounded buffers as in \cref{sec:agent-environment-communication} for this, modified to be located solely in CLS memory, dedicating the master thread to result aggregation.
 We found that this created a performance bottleneck at this final stage, causing significant head-of-line blocking for each of the workers.
@@
 
+\begin{algorithm}
+	\caption{ParSa---\emph{Par}allel \emph{Sa}rsa\label{alg:parsa}}
+	\SetKwProg{generate}{Function \emph{generate}}{}{end}
+
+	Map store=new Map(obj, queue)\;
+	\generate{Object pivot}{
+		\ForAll{child $c$ in pivot}{
+			\If{ $c$'s FieldContext is not set and $c$ is fusible}{
+				generate($c$)\;
+			}
+		}
+		build pivot's fieldContext $fc$\;
+		EmitClassName\;
+		EmitFields($fc$)\;
+		EmitMethods($fc$)\;
+	}
+\end{algorithm}
+
 \subsection{Reconfigurability}
-\approachshort{} allows for policy dimensions, size, action count, tiling strategies, and algorithmic learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) to be chosen and changed at runtime.
-This extends to policy data, which may be imported from an offline pre-trained model via such packets, and exported via the NFP binary support package on the host machine.
+\approachshort{} allows for policy dimensions, size, action count, tiling strategies, and algorithmic learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) to be chosen and changed at runtime using at most two control packets.
+This extends to policy data, which may be imported from an offline pre-trained model via many such packets, and exported via the NFP binary support package on the host machine.
 Some aspects must be chosen at compile time; bit depth, work allocation strategy, and maximum policy/tiling/state sizes---these govern core operation or allocated memory regions.
@@
 In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved (pool 2~\parencite{rfc2474}) DSCP value as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-While we primarily use this to automate parser generation, this also allows for configuration packets to be received from only trusted hosts over the dataplane if needed via P4 rules.
+While we primarily use this to automate parser generation, this also allows for configuration packets to be received from only trusted hosts (over the dataplane if needed) via P4 rules.
 Our control packet generation library (and evaluation frameworks which build upon it) are written in Rust.
@@
 As a consequence, tiling space requirements scale exponentially with dimension count, and so higher-dimension tilings must be placed in memory regions large enough to allocate space for them.
-As a result, tilings are split across memory regions, giving different access and compute costs to different tasks---ranging over CLS, CTM and EMEM in increasing latency.
-
-We use a simple first-fit work placement algorithm run on the NFP, placing the largest work item into the least loaded thread of the least loaded core.
-The cost of any individual work item (according to its dimension count and memory location) was estimated offline.
+As a result, tilings are split across memory regions, giving different access and compute costs to different tasks---ranging over CLS, CTM and IMEM in increasing latency.
+
+We use a simple first-fit work placement algorithm run on the NFP, placing the largest work item into the least loaded context of the least loaded ME.
+The cost of any individual work item (according to its dimension count and memory location) was empirically measured offline, and we weight the total cost per ME to account for the number of minion threads available.
+This weighting specifically accounts for the presence of the controller thread on the first ME.
 This work allocation and any cached per-work-item values are recomputed when any configuration is installed or changed.
-Naturally, for $n$ tasks and $m$ threads this is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per task, storing $m/8$ and $\le8$ costs respectively.
+Naturally, for $n$ tilings/tasks and $m$ threads this procedure is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per task, storing $m/8$ and $\le8$ costs respectively.
 
@@
 Naturally, smaller bit depths reduce the storage required for policies, tiling data, and stored state-action pairs---allowing more complex problems to be modelled using more dimensions or fine-grained policies.
-Through the lens of computational efficiency, reduced bit depth allows action preference lists to be read using fewer I/O operations (as more values may fit into a single machine word).
+Through the lens of computational efficiency, reduced bit depth should allows action preference lists to be read using fewer I/O operations (as more values may fit into a single machine word).
+
 We had investigated bit-stuffing several such values into a single word for our atomic writeback mechanism (as the platform offers both \SIlist{32;64
@@
 This is analogous to SIMD---through clever use of padding bits---but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
-
-?? Need to mention that this is a theoretical improvement; the results are basically opposite to this.
-?? Main benefit is then, I guess, the redux in storage requirements?
+We investigate further performance implications of changing bit-depth in the sequel.
 
@@
 
-We instead suggest on NFP chips that externs or datapath stages which apply actions to packet processing should maintain a small local store of state-action pairs, and periodically send these back to the controller for batch rule installation.
+We instead suggest on NFP chips that \texttt{extern}s or datapath stages which apply actions to packet processing should maintain a small local store of state-action pairs, and periodically send these back to the controller for batch rule installation.
 This ensures that the vast majority of installed rules enjoy accelerated performance, while preventing rule installation delay on the newest decisions.
@@
 
-?? Achieving very tight latency bounds on \Coopfw{} requires quite a decent core allocation. (i.e., need at least 11 workers to make a gain in online t'put, 5 cores needed to win in throughput).
+As is expected of parallel algorithms, efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
+Even in wait-free algorithms, this requires a minimum number of workers to improve upon the latency bounds of a serial approach.
+We investigate the exact worker count requirements imposed to break even or improve upon single-threaded execution in the sequel.
 
@@
 Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
-
-Consider a single, shared, read-only register between RL functional ...
-?? think about really really carefully lest Reviewer 2 murder you
-?? Parallel, atomic writes at the required resolution.
-?? No bit stuffing required for SIMD-alike?.
-?? More chances for vectorisation? Is that where we're falling down atm with 32-bit being weirdly optimal?
-
-Bringing this to high-port density devices such as Intel Tofino-powered switches is more difficult.
+As our later results show, working on a bitdepth below the native word size forces the compiler or programmer to emit excess code to extract and re-align preference values.
+This reduces overall efficiency even though considerably fewer reads are made.
+We expect that more bespoke hardware/FPGA co-design would allow this boiler-plate to be removed and enable the SIMD-like optimisations we discuss above without creative and costly bit-stuffing.
+
+Bringing \approachshort{} to high-port density devices such as Intel Tofino-powered switches is more difficult.
 As discussed in \cref{sec:motivation}, Tofino ASICs map closely to the P4 PSA, meaning that there are no spare asynchronous general-purpose compute units to place this work on.
-A potential solution is to divide RL actions across several received packets (i.e., iteratively computing a portion of the action preference list each time) until any further work would delay outbound transmission.
+Taking inspiration from real-time programming techniques, a potential solution is to divide RL actions across several received packets (i.e., iteratively computing a portion of the action preference list each time) until any further work would delay outbound transmission.
 This, however, would introduce new issues surrounding concurrent accesses, work splitting, and altered timescales for learning: we leave their treatment and examination to future work.
 
-%\begin{figure}
-%	\centering
-%	\resizebox{\linewidth}{!}{\begin{tikzpicture}
-%		\node(remote){Remote};
-%		
-%		%%%
-%		
-%		\node[below=0.05 of remote](swpos){};
-%		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
-%		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
-%		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
-%		\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
-%		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
-%		
-%		%
-%		
-%		\node[right=2 of swpos](epos){};
-%		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
-%		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
-%		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
-%		\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
-%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.85,0)$) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
-%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\small \centering Estimator\\$g(\cdot)$\end{varwidth}};
-%		
-%		%
-%		
-%		\node[right=3.5 of epos](apos){};
-%		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
-%		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
-%		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
-%		\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent vNF};
-%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
-%		
-%		%
-%		
-%		\node[below=2.3 of remote.west](linestart){};
-%		\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
-%		\draw [dashed] (linestart) -- (lineend);
-%		
-%		%%%
-%		
-%		\node[below=2.4 of remote](local){Local};
-%		
-%		%%%
-%		
-%		\node[below=0.25 of local](aswpos){};
-%		\node[fill=white!80!uofgthistle, draw=black, minimum height=3cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
-%		\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
-%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\small \centering OpenFlow\\Tables\end{varwidth}};
-%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\small \centering Stats\\Collector\end{varwidth}};
-%		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
-%		
-%		%
-%		
-%		\node (avfpos) at ($(aswpos) + (3,0.2)$){};
-%		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
-%		\node[below right, inner sep=2pt] at (avf.north west) {\small Agent vNF};
-%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
-%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\small \centering Flowstate\\Database\end{varwidth}};
-%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\small \centering TRS\\Scheduler\end{varwidth}};
-%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
-%		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};
-%		
-%		%%%
-%		
-%		\tikzset{>=stealth}
-%		
-%		\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
-%		\draw[thick, ->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};
-%		
-%		\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
-%		
-%		\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
-%		\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};
-%		
-%		\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};
-%		
-%		\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
-%		\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};
-%		
-%		\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
-%		\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
-%		\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
-%		\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
-%		\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};
-%		
-%		\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
-%		\node[right=0.3] at (avfsa.east) {\tiny Experience};
-%		\end{tikzpicture}}
-%	\caption{
-%		System architecture  for our RL-driven DDoS defence system.
-%		\label{fig:sys-arch}
-%	}
-%\end{figure}
-
-\section{Potential Integrations}\label{sec:potential-integrations}
-To show the general applicability and use of \approachshort{}, we propose two ideal integrations which would benefit most from in-NIC reinforcement learning; online DDoS attack mitigation (\cref{sec:integ-1}), and ??
-OTHER THING (\cref{sec:integ-2}).
-We support these using other state-of-the-art P4/PDP developments, and discuss how best to balance the concerns of online training with throughput in widespread deployment.
-
-?? Will these look better \emph{after} the results?
-
-\subsection{In-Network DDoS Defence}\label{sec:integ-1}
-Classical reinforcement learning has seen recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
-\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and local, per-flow state to monitor how flows respond to bandwidth changes and packet drop---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} with more allowance for congestion-unaware protocols.
-Actions then move flows up or down in punishment levels according to a finite state machine.
-
-\fakepara{Why in-NIC?}
-The above approach relies upon co-hosting traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
-However, packet mirroring, offloading RL computation to a host (potentially over layers of virtualisation), and CPU performance are all sources of additional, consistent state-action latency.
-Both traffic statistics collection \emph{and} data-driven learning must be executed on such hosts/network functions.
-Unless running these stages in a dedicated pipeline (adding further processing latency), resource contention between these processes will further impact tail latencies.
-Naturally, this requires high-performance hardware to be stacked at network egress points, potentially beyond reasonable space, power, or ventilation constraints.
-The solution to implement and improve upon this work using \approachshort{} is to place its RL agents on SmartNICs at AS edge nodes---a bump-in-the-wire deployment.
-
-?? Any way to move some of this sooner? I.e., this is ``why move DDoS prevention here'', Stefanos suggested backporting some of this to justify ``why \emph{RL} in-NIC''?
-
-\fakepara{Inputs}
-To collate the required inputs and state, we then examine the recent innovations of the community.
-Low-latency, pure-P4 solutions to extract and record per-flow TCP state directly in the dataplane such as Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} are well-studied.
-In fact, the statistics offered by Dapper are a super-set of the local input state values employed by \emph{Guarded} agents, offering an opportunity to further improve their efficacy. 
-We propose placing such monitors in the P4 dataplane, existing on-chip alongside the \approachshort{} agent.
-The required global state (load measures from network paths) must still come from elsewhere in the network; this is now the element at highest risk of becoming stale, but the least likely to vary significantly in response to individual actions.
-
-The original work suggests theoretical ``ground truth'' reward functions, whose correct implementation and designs were left as an open challenge. 
-We posit that INDDoS~\parencite{tnms-ddos-victim-ident}, which uses Count-min Sketches to estimate DDoS victim cardinality, could be used as an effective reward function source---i.e., using the number of detected victims as a loss value.
-
-\fakepara{Integrating \approachshort}
-Before each Dapper monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{OUT Ring} for any generated actions.
-As we note in \cref{sec:limitations}, these actions would be placed into a small local hash table and simultaneously exported to the attached controller to be inserted as P4 rules in batches.
-After the main monitoring step, packet ingress timestamps would be used to emulate the \emph{Timed Random Sequential} (TRS) scheduler used by the anti-DDoS agents for rate-controlled work, where state vectors would be selected and passed into \approachshort.
-By design, active flows which are not \emph{judged} after a configurable time are discarded to prune the work set and allow new flows to be seen by the agent.
-The tight bounds on execution time known \emph{a priori} make it easy to calculate the maximum number of decisions which can be made per deadline.
-Reward values would then be separately inserted by a modified INDDoS table.
-
-?? Can I have a P4 code snippet here?
-
-A reduction in state-action latency (i.e., with \Coopfw) is, as always, useful for minimising the noise inherent in learning.
-However, an agent is limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
-Accordingly, this use-case benefits most from an increase in \emph{throughput} using \Indfw{}.
-In this context, higher throughput means that network flows are more likely to be judged in every timestep, even when flow cardinality is high---making it more likely that changes in flow behaviour will be observed, acted on, and learned from.
-
-We note that the TRS scheduler is designed to handle cases where the number of flows created by an attacker, combining state vectors and information over time while the asynchronous agent is itself busy.
-An increased number of (attacking) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
-As we shall examine in \Cref{sec:evaluation}, \approachshort{} far exceeds the action throughput of a host based deployment could handle.
-The control plane is then free to use wildcards or specific matches to narrow down (or expand) the set of flows to be considered and controlled, though explicit scheduling techniques are still key in such adversarial environments.
-
-\subsection{Something host-specific?}\label{sec:integ-2}
-Accelerate something at an end-host... Flow control? Something datacentre-y? A novel CCA? What?
-
-\subsection{Network Deployment Considerations}
-The two compute models discussed above need not be homogeneously deployed.
-For instance, in a networked deployment a subset of \approachshort{} nodes could be \SI{32}{\bit} \emph{\coopfw{}} agents, training online, while most other nodes ran \SI{8}{\bit} \emph{\indfw{}} designs to meet throughput guarantees.
-The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
-This can be taken further still, using policy deltas or execution traces to enable out-of-path transfer learning for more complex function approximators such as neural networks.
-
 \section{Evaluation}\label{sec:evaluation}
-We outline our performance evaluation of \approachshort...
+We investigate the performance of \approachshort{} as compared with classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{$\times$} speedup in median--\nth{99} state-action latency and \SI{9.9}{$\times$} greater online learning throughput.
+Crucially, we find that in-NIC execution offers tight tail latency bounds compared with host-based approaches.
+We report on how \approachshort{} scales to fit available device resources as cores/threads are added, noting that both \Indfw{} and \Coopfw{} designs outperform commodity hosts using just one core in latency and online throughput.
+Furthermore, \Indfw{} provides higher per-core offline throughput than host-based approaches, even though our measured hosts exhibit higher clock speeds.
+Finally, we show that \approachshort{} has minimal impact on dataplane cross-traffic carried by its parent device.
 
 \subsection{Experimental Setup}\label{sec:experimental-setup}
-Evaluation of \approachshort{} was performed on server blades hosting a single Netronome Agilio LX 1x100GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
-These mid-tier servers (\emph{MidServer}) ran Ubuntu 18.04.5 LTS (4.15.0-140-generic x86\_64) on Intel Xeon Bronze 3204 CPUs ($6\times$\SI{1.9}{\giga\hertz}) and \SI{32}{\gibi\byte} RAM.
-We additionally use a more powerful consumer-grade machine (\emph{Collector}) for estimating host performance when offloaded to a network function, running Ubuntu 18.04.4 LTS (4.15.0-96-generic) on Intel Core i7-6700K ($4\times$\SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} RAM.
+Testing machines had the following specifications, all having \SI{32}{\gibi\byte} RAM:
+\begin{description}
+	\item[\emph{MidServer}] Intel Xeon Bronze 3204 CPUs ($6\times$\SI{1.9}{\giga\hertz}),
+	\item[\emph{HighServer}] Intel Xeon Silver 4208 CPU ($8\times$\SI{2.1}{\giga\hertz}),
+	\item[\emph{Collector}] Intel Core i7-6700K ($4\times$\SI{4.2}{\giga\hertz}).
+\end{description}
+Evaluation of \approachshort{} was performed on server blades (\emph{Mid/HighServer}), each hosting a single Netronome Agilio LX 1x40GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
+These servers ran Ubuntu 18.04.5 LTS (4.15.0-140-generic).
+We additionally use a more powerful consumer-grade machine (\emph{Collector}) for estimating host performance when offloaded to a network function, running Ubuntu 18.04.4 LTS (4.15.0-96-generic).
 Control programs were built using rustc 1.52.1.
+
 We run \approachshort{} on an island with \num{4} MEs of the NFP-6480 for a total of \num{32} hardware threads.
@@
 Host throughput measures were observed over \num{10} trials of \SI{10}{\second}.
-
-?? verify these numbers against code?
-
-?? NOTE: Ritchie is actually Intel(R) Xeon(R) Silver 4208 CPU (8 times 2.1GHz)
-
 Policy sizes are set to those matching a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dimension state vectors, a bias tile and 16 full tiling sets (7$\times$1-dim, 8$\times$2-dim, 1$\times$4-dim), 8 tilings per dimension set, 6 tiles per dimension, choosing between 10 actions.
-Although their more successful \emph{SPF} agent design uses 3 actions, we choose this to investigate the performance of larger agent designs.
-
-Studies based on \textcite{DBLP:journals/tnsm/SimpsonRP20} mirror their experimental setup (Ubuntu 18.04.2 LTS (4.4.3-040403-generic x86\_64), Intel Core i7-6700K (\SI{4.2}{\giga\hertz}), \SI{32}{\gibi\byte} RAM).
+For \Coopfw{} for instance, this creates \num{129} work items across \num{31} workers.
+Although their more successful \emph{SPF} agent design uses 3 actions, we choose a larger action set to investigate the performance of more complex agents.
 
@@
 We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs.
-We compare these against floating point and quantised implementations running on a commodity host machine.
-This allows us to demonstrate the performance differences between the \indfw and \coopfw configurations, particularly in how \indfw's required policy locks impact throughput.
+We compare these against floating point (numpy-based) implementations running on a commodity host machine.
+This allows us to demonstrate the performance differences between the \indfw{} and \coopfw{} configurations, particularly in how \indfw's required policy locks impact throughput.
 We also report on the derived throughput measures accounting for the locking behaviour of each.
@@
 We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} baseline allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ items given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
+We use maximum-size, \SI{32}{\bit} policies as described above.
 
@@
 
-?? NOT DONE YET
-?? MOSTLY A PAPER EXERCISE FROM SOME OTHER MEASURES
-
 \fakepara{Co-existence with the dataplane}
-We measure maximum packet rate and bandwidth throughput of a P4 pipeline on our SmartNICs, both with and without \approachshort{}.
+While varying the rate of full RL updates performed by \approachshort{}-\Coopfw{} (\SI{32}{\bit}) from \numrange{0}{16000} actions/s, we measure packet losses and sample latencies of cross traffic forwarded over a P4 pipeline hosted on our SmartNICs.
 This allows us to quantify whether on-chip (out-of-path) execution impacts ordinary dataplane behaviour through indirect means: e.g., EMEM cache evictions or hidden resource contention.
-In particular, we stress test both decisions and policy updates of our \emph{Parallel} design at full throughput.
-
-?? NOT DONE YET.
+%In particular, we stress test both decisions and policy updates of our \emph{Parallel} design at full throughput.
+
+We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, using an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source via the default firmware.
+We perform throughput/loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements.
+Sent traffic was comprised of fixed-size \SIrange{64}{1518}{\byte} packets~\cite{rfc2544}.
+CPU clock scaling was disabled on \emph{HighServer}.
 
@@
 Using the maximum policy size defined above, we investigate how the memory requirements imposed by \approachshort{} vary with the number of dedicated MEs, over and above a base P4 forwarding plane.
+We report resource use for \SI{32}{\bit} \Indfw{} and \Coopfw{} agents with maximum policy sizes as above, with hash-tables sized to \num{4096} state-action pairs and \num{16} separate reward values.
 
@@
 \fakepara{Deployability}
-By timing policy installation, compile times and other metrics, we describe the runtime costs needed for an administrator to repurpose an installed agent in a live network.
+By timing agent setup and compile times, we describe the runtime costs needed for an administrator to repurpose an installed agent in a live network.
 
@@
 \fakepara{Raw inference and learning performance}
-\Cref{tab:lats} shows that \Coopfw{} \approachshort{} achieves sub-\SI{50}{\micro\second} median latency, with \nth{99} percentile latencies less than \SI{1}{\micro\second} worse.
+\Cref{tab:lats} shows that \Coopfw{} \approachshort{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse.
 As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, while the simpler \Indfw{} succeeds when policies are fixed/known.
@@
 ?? Mention versus DNN latency in passing here?
+
+?? Need to mention some of the omitted results here for brevity; namely host integer implementations.
 
@@
 
-\Cref{tab:tputs} shows how \approachshort{} compares in terms of throughput with a numpy-based implementation of an RL policy.
+\Cref{tab:tputs} shows how \approachshort{} compares in terms of throughput with a numpy-based (i.e., BLAS-accelerated) implementation of an RL policy.
 We choose the worker count on host machines with the largest number of parallel agents which produced a statistically significant throughput increase ($p < 0.05$).
@@
 
-?? Urgently describe and explain why lower bitdepth gives *worse* metrics.
-?? Mainly, register width.
-?? We also do stuff to batch read as many action vals as possible!! And yet this happens!!
-?? Also explain that Integer math does worse due to it falling back to BLAS in numpy (i.e., a float optimiser).
-?? We might have good int results if I do it in Rust?
+A somewhat remarkable trend is that the \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit} designs, with the exception of smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
+We found this to be rather surprising, given that our implementation is optimised to read and write policy tile data in batches (achieving fewer I/O operations as required).
+Our understanding is that this is due to the native register width on the NFP being \SI{32}{\bit}, and so the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
+This matches \SI{32}{\bit} becoming dominant in more complex workloads: higher dimension tilings require more arithmetic operations to compute the hit tile index.
+Most of the I/O comes after this step, causing ALU use to dominate.
+Regardless, the reduction in memory needed to store a policy remain.
 
 \fakepara{Work allocation}
-\Cref{fig:work-alloc-32} shows that our heuristic first-fit allocator (\emph{Balanced}) is key in achieving consistent sub-\SIlist{50;100}{\micro\second} latencies/processing times respectively.
+\Cref{fig:work-alloc-32} shows that our heuristic first-fit allocator (\emph{Balanced}) is key in achieving consistent sub-\SIlist{35;65}{\micro\second} latencies/processing times respectively.
 The trend is repeated for all bit depths, noting the latency differences described above.
@@
 We expect that closing or exceeding this gap may require more complex modelling of NFP's hardware thread interactions, which lies far beyond the scope of in-NIC scheduling.
-Some additional complexity may be tolerated though---subject to code store limits---scheduling runs exactly once per configuration change, so does not impact per-action code.
+Some additional complexity may be tolerated, subject to code store limits---scheduling runs exactly once per configuration change, so does not impact per-action code.
 
@@
 
-\Cref{fig:tput-per-core}
-?? downticks in tput/core correspond to a `plateau' in throughput: the heuristic scheduler cannot find a better solution until extra cores are allocated
-?? Why? granularity of work items. Can't really subdivide high-dim policy items. Pushing past is tricky as described...
+An interesting aspect of \Coopfw{} and the \emph{ParSa} algorithm in general is that adding additional cores has both diminishing returns and key thresholds to pass.
+Consider \cref{fig:tput-per-core}, where the throughput per worker decreases over time but occasionally increases sharply.
+Later downward ticks (\numrange{25}{29} workers) correspond to plateaus in throughput.
+This is a common problem stemming from the granularity of work items, where the heuristic scheduler cannot find a better solution to a bottleneck until extra cores are allocated.
+We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
+While there is still have a \SI{4.2}{$\times$} factor of task oversubscription at maximum worker count, it is clear that latencies are bound below by the length of the longest task.
 
@@
 \fakepara{Co-existence with the dataplane}
-?? TODO
+No packet loss incurred
+?? mention actually achieved rates?
+
+?? Now talk about statistically significant deltas etc.
+
+\begin{figure}
+	\centering
+	\begin{subfigure}{0.45\linewidth}
+		\resizebox{1.0\linewidth}{!}{
+%			\includegraphics{../plots/build/stress/heat-latency-median}
+			\includegraphics{../plots/build/stress/heat-latency-two-9}
+		}
+		\caption{Ayy?}
+	\end{subfigure}
+	\hspace{0.05\linewidth}
+		\begin{subfigure}{0.45\linewidth}
+		\resizebox{1.0\linewidth}{!}{
+			\includegraphics{../plots/build/stress/histo-1518B-0-10}
+		}
+		\caption{Ayy?}
+	\end{subfigure}
+	\caption{Ayy.}
+\end{figure}
 
@@
 	\midrule Base P4 & 6776.67 & 88.24 & 268.52 & 2.91 & 858.28 & 10.48 & 0.00 & 0.00 & 0.00 & 0.00 \\
-	\Indfw(1) & 6780.21 & 88.28 & 2541.07 & 27.57 & 1263.28 & 15.42 & 24.75 & 38.67 & 94.25 & 36.82 \\
-	\Indfw(4) & 6780.21 & 88.28 & 2545.32 & 27.62 & 1263.28 & 15.42 & 51.18 & 79.97 & 107.00 & 41.80 \\
-	\Coopfw(1) & 6779.12 & 88.27 & 1769.83 & 19.20 & 1263.28 & 15.42 & 20.25 & 31.64 & 90.00 & 35.16 \\
-	\Coopfw(4) & 6779.12 & 88.27 & 1769.84 & 19.20 & 1263.28 & 15.42 & 31.91 & 49.85 & 90.00 & 35.16 \\
+	\Indfw(1) & 6780.21 & 88.28 & 2541.08 & 27.57 & 1263.28 & 15.42 & 24.75 & 38.67 & 94.25 & 36.82 \\
+	\Indfw(4) & 6780.22 & 88.28 & 2545.33 & 27.62 & 1263.28 & 15.42 & 51.18 & 79.97 & 107.00 & 41.80 \\
+	\Coopfw(1) & 6779.12 & 88.27 & 1773.59 & 19.24 & 1263.28 & 15.42 & 22.41 & 35.01 & 90.00 & 35.16 \\
+	\Coopfw(4) & 6779.12 & 88.27 & 1769.84 & 19.20 & 1263.28 & 15.42 & 52.16 & 81.49 & 90.00 & 35.16 \\
 	\bottomrule
@@
 
+\section{Potential Integrations}\label{sec:potential-integrations}
+To show the general applicability and use of \approachshort{}, we propose an ideal integrations which would benefit from in-NIC reinforcement learning; online DDoS attack mitigation.
+We support this using other state-of-the-art P4/PDP developments, and discuss how best to balance the concerns of online training with throughput in widespread deployment.
+Additionally, we discuss how network administrators may combine offline (\Indfw) and online (\Coopfw) agents within a network to achieve online learning while maximising action throughput in the majority of the network.
+
+\subsection{In-Network DDoS Defence}\label{sec:integ-1}
+Classical reinforcement learning has seen recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
+\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and local, per-flow state to monitor how flows respond to bandwidth changes and packet drop---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} with more allowance for congestion-unaware protocols.
+Actions then move flows up or down in punishment levels according to a finite state machine.
+
+\fakepara{Why in-NIC?}
+The above approach relies upon co-hosting traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
+However, packet mirroring, offloading RL computation to a host (potentially over layers of virtualisation), and CPU performance are all sources of additional, consistent state-action latency.
+Both traffic statistics collection \emph{and} data-driven learning must be executed on such hosts/network functions.
+Unless running these stages in a dedicated pipeline (adding further processing latency), resource contention between these processes will further impact tail latencies.
+Naturally, this requires high-performance hardware to be stacked at network egress points, potentially beyond reasonable space, power, or ventilation constraints.
+The solution to implement and improve upon this work using \approachshort{} is to place its RL agents on SmartNICs at AS edge nodes---a bump-in-the-wire deployment.
+
+?? Any way to move some of this sooner? I.e., this is ``why move DDoS prevention here'', Stefanos suggested backporting some of this to justify ``why \emph{RL} in-NIC''?
+
+\fakepara{Inputs}
+To collate the required inputs and state, we then examine the recent innovations of the community.
+Low-latency, pure-P4 solutions to extract and record per-flow TCP state directly in the dataplane such as Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} are well-studied.
+In fact, the statistics offered by Dapper are a super-set of the local input state values employed by \emph{Guarded} agents, offering an opportunity to further improve their efficacy. 
+We propose placing such monitors in the P4 dataplane, existing on-chip alongside the \approachshort{} agent.
+The required global state (load measures from network paths) must still come from elsewhere in the network; this is now the element at highest risk of becoming stale, but the least likely to vary significantly in response to individual actions.
+
+The original work suggests theoretical ``ground truth'' reward functions, whose correct implementation and designs were left as an open challenge. 
+We posit that INDDoS~\parencite{tnms-ddos-victim-ident}, which uses Count-min Sketches to estimate DDoS victim cardinality, could be used as an effective reward function source---i.e., using the number of detected victims as a loss value.
+
+\fakepara{Integrating \approachshort}
+Before each Dapper monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{OUT Ring} for any generated actions.
+As we note in \cref{sec:limitations}, these actions would be placed into a small local hash table and simultaneously exported to the attached controller to be inserted as P4 rules in batches.
+After the main monitoring step, packet ingress timestamps would be used to emulate the \emph{Timed Random Sequential} (TRS) scheduler used by the anti-DDoS agents for rate-controlled work, where state vectors would be selected and passed into \approachshort.
+By design, active flows which are not \emph{judged} after a configurable time are discarded to prune the work set and allow new flows to be seen by the agent.
+The tight bounds on execution time known \emph{a priori} make it easy to calculate the maximum number of decisions which can be made per deadline.
+Reward values would then be separately inserted by a modified INDDoS table.
+
+?? Can I have a P4 code snippet here?
+
+A reduction in state-action latency (i.e., with \Coopfw) is, as always, useful for minimising the noise inherent in learning.
+However, an agent is limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
+Accordingly, this use-case benefits most from an increase in \emph{throughput} using \Indfw{}.
+In this context, higher throughput means that network flows are more likely to be judged in every timestep, even when flow cardinality is high---making it more likely that changes in flow behaviour will be observed, acted on, and learned from.
+
+We note that the TRS scheduler is designed to handle cases where the number of flows created by an attacker, combining state vectors and information over time while the asynchronous agent is itself busy.
+An increased number of (attacking) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
+As we shall examine in \Cref{sec:evaluation}, \approachshort{} far exceeds the action throughput of a host based deployment could handle.
+The control plane is then free to use wildcards or specific matches to narrow down (or expand) the set of flows to be considered and controlled, though explicit scheduling techniques are still key in such adversarial environments.
+
+%\subsection{Something host-specific?}\label{sec:integ-2}
+%Accelerate something at an end-host... Flow control? Something datacentre-y? A novel CCA? What?
+
+\subsection{Network Deployment Considerations}
+The two compute models discussed above need not be homogeneously deployed.
+For instance, in a networked deployment a subset of \approachshort{} nodes could be \SI{32}{\bit} \coopfw{} agents, training online, while most other nodes ran \SI{8}{\bit} \indfw{} designs to meet throughput guarantees.
+The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
+This can be taken further still, using policy deltas or execution traces to enable out-of-path transfer learning for more complex function approximators such as neural networks.
+
 \section{Related Work}
@@
 \begin{acks}
-	The authors would like to thank Rhys Simpson for his comments and discussions on the soundness of SIMD-like optimisations.
+	The authors would like to thank Rhys Simpson for his comments, and for discussions on the soundness of SIMD-like optimisations.
+	They would additionally like to thank Stefanos Sagkriotis, Mircea Iordache-\c{S}ic\u{a}, and Haruna Adoga for their comments and feedback.
 	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grant~\grantnum{gs-epsrc}{EP/N509668/1}].
@@
 
-\begin{appendices}
-	\section{Old ``What could we do''?}
-	How do we evaluate this without looking into classification performance?
-	\begin{itemize}
-		\item Analytically?
-		\begin{enumerate}
-			\item Prove that tile splitting scheme over memory hierarchy can improve performance.
-			\begin{itemize}
-				\item Hit rate of different tiles according to tile dimensionality?
-				\item Access time for each tier of memory hierarchy.
-			\end{itemize}
-		\end{enumerate}
-		\item Experimentally?
-		\begin{enumerate}
-			\item Create ``VNF'' testbed: mirror traffic from switch to a host for telemetry processing/flow state. Send flow state over to another node who computes RL actions. RL actions installed on switch.
-			\begin{itemize}
-				\item Why not co-host RL computation with telemetry? Could be separate functions, could want to ensure that neither impacts the performance of the other.
-			\end{itemize}
-			\item Create ``PDP'' testbed: Load firmware, pass in traffic. Traffic processing/action compute/table mod occurs in PDP.
-			\item What traffic? Healthy mix of attack (UDP), Discord legit Opus VoIP traffic, HTTP bulk-transfer traffic.
-			\item Measure $t_{1 \cdots 3}$ according to \cref{fig:state-slip} in both VNF and PDP contexts.
-			\begin{description}
-				\item[$t_1$]
-				\begin{description}
-					\item[PDP] Packet ingress $\rightarrow$ INT Island $\rightarrow$ State arrival in Control ME.
-					\item[VNF] Packet mirror time $\rightarrow$ telemetry VNF $\rightarrow$ state arrival in agent vnf.
-				\end{description}
-
-				\item[$t_2$]
-				\begin{description}
-					\item[PDP] Time to compute action over parallel Policy MEs (RL Island, SmartNIC).
-					\item[VNF] Time to compute action in host program (on VNF) from received state.
-				\end{description}
-
-				\item[$t_3$]
-				\begin{description}
-					\item[PDP] Modify local tables across ME/Island boundaries (same device), verify rule present (or get ACK).
-					\item[VNF] Generate OpenFlow control messages, send OF message to switch, verify rule present (or get ACK).
-				\end{description}
-			\end{description}
-			\item HYPOTHESIS: PDP will present a reduction in these three times for per-flow state.
-			\item QUESTION: Does same apply to learning?
-			\item NOTE: since we're not actually assessing the \emph{quality} of output decisions, we can cut corners on what data is sent out and where.
+%\begin{appendices}
+%	\section{Old ``What could we do''?}
+%	How do we evaluate this without looking into classification performance?
+%	\begin{itemize}
+%		\item Analytically?
+%		\begin{enumerate}
+%			\item Prove that tile splitting scheme over memory hierarchy can improve performance.
+%			\begin{itemize}
+%				\item Hit rate of different tiles according to tile dimensionality?
+%				\item Access time for each tier of memory hierarchy.
+%			\end{itemize}
+%		\end{enumerate}
+%		\item Experimentally?
+%		\begin{enumerate}
+%			\item Create ``VNF'' testbed: mirror traffic from switch to a host for telemetry processing/flow state.
+Send flow state over to another node who computes RL actions.
+RL actions installed on switch.
+%			\begin{itemize}
+%				\item Why not co-host RL computation with telemetry? Could be separate functions, could want to ensure that neither impacts the performance of the other.
+%			\end{itemize}
+%			\item Create ``PDP'' testbed: Load firmware, pass in traffic. Traffic processing/action compute/table mod occurs in PDP.
+%			\item What traffic? Healthy mix of attack (UDP), Discord legit Opus VoIP traffic, HTTP bulk-transfer traffic.
+%			\item Measure $t_{1 \cdots 3}$ according to \cref{fig:state-slip} in both VNF and PDP contexts.
+%			\begin{description}
+%				\item[$t_1$]
+%				\begin{description}
+%					\item[PDP] Packet ingress $\rightarrow$ INT Island $\rightarrow$ State arrival in Control ME.
+%					\item[VNF] Packet mirror time $\rightarrow$ telemetry VNF $\rightarrow$ state arrival in agent vnf.
+%				\end{description}
+%				
+%				\item[$t_2$]
+%				\begin{description}
+%					\item[PDP] Time to compute action over parallel Policy MEs (RL Island, SmartNIC).
+%					\item[VNF] Time to compute action in host program (on VNF) from received state.
+%				\end{description}
+%				
+%				\item[$t_3$]
+%				\begin{description}
+%					\item[PDP] Modify local tables across ME/Island boundaries (same device), verify rule present (or get ACK).
+%					\item[VNF] Generate OpenFlow control messages, send OF message to switch, verify rule present (or get ACK).
+%				\end{description}
+%			\end{description}
+%			\item HYPOTHESIS: PDP will present a reduction in these three times for per-flow state.
+%			\item QUESTION: Does same apply to learning?
+%			\item NOTE: since we're not actually assessing the \emph{quality} of output decisions, we can cut corners on what data is sent out and where.
 I.e., in-band network telemetry island can be skipped, if we move over representative volumes of state to/from the RL island with approximately correct delay as recorded by other studies.
-		\end{enumerate}
-	\end{itemize}
-	I can think of one way to evaluate this if we employ classification performance:
-	\begin{enumerate}
-		\item Run the existing work~\parencite{DBLP:journals/tnsm/SimpsonRP20}, acquire a policy for one node after full info sharing.
-		\item Load this policy onto the RL Island (\cref{fig:netro-arch}).
-		\item Either feed in raw state packets, or feed in traffic if we can get INT working.
-		\item For each level of quantisation needed, examine average performance.
-		\item Don't compare against old numbers for final perf: use a ``VNF'' setup similar to above---this will probably suffer a bit compared to old results, where everything was on the same machine.
+%		\end{enumerate}
+%	\end{itemize}
+%	I can think of one way to evaluate this if we employ classification performance:
+%	\begin{enumerate}
+%		\item Run the existing work~\parencite{DBLP:journals/tnsm/SimpsonRP20}, acquire a policy for one node after full info sharing.
+%		\item Load this policy onto the RL Island (\cref{fig:netro-arch}).
+%		\item Either feed in raw state packets, or feed in traffic if we can get INT working.
+%		\item For each level of quantisation needed, examine average performance.
+%		\item Don't compare against old numbers for final perf: use a ``VNF'' setup similar to above---this will probably suffer a bit compared to old results, where everything was on the same machine.
 Make it relative/normalised.
-	\end{enumerate}
-
-	\section{Intermediate steps?}
-	System?
-	\begin{itemize}
-		\item Setup other core as event source.
-		\begin{itemize}
-			\item \emph{\num{2} week.}
-			\item Idea: proxy for idea of INT core mentioned elsewhere, combined with the main periodic event sending that the last RL work relied upon.
+%	\end{enumerate}
+%
+%	\section{Intermediate steps?}
+%	System?
+%	\begin{itemize}
+%		\item Setup other core as event source.
+%		\begin{itemize}
+%			\item \emph{\num{2} week.}
+%			\item Idea: proxy for idea of INT core mentioned elsewhere, combined with the main periodic event sending that the last RL work relied upon.
 So, say "you can put that state management and extraction on another core, here's what the comms cost is" or similar.
-		\end{itemize}
-		\item Fixes to hash table use, signalling.
-		\begin{itemize}
-			\item \emph{\num{1} week.}
-			\item Blocks progress on optimisation, correct multi-stream RL.
-		\end{itemize}
-		\item Optimise/parallelise action compute/learning.
-		\begin{itemize}
-			\item \emph{\num{2} week.}
-			\item Value? Proves this approach can scale according to resource availability/demand on the deployment device, at compile-time anyhow.
-		\end{itemize}
-	\end{itemize}
-
-	Experiments?
-	\begin{itemize}
-		\item Measure sum of times $t_{1...3}$ described elsewhere.
-		\begin{itemize}
-			\item \emph{\num{1} week.}
-			\item Currently have individual, need ``end-to-end''.
-		\end{itemize}
-		\item Throughput impact testing.
-		\begin{itemize}
-			\item \emph{\num{1} week.}
-			\item Requires current impl, creation of basic testbed and packet forwarding setup.
-			\item Use-case independent.
-		\end{itemize}
-		\item Measure execution, training costs for varying policy sizes.
-		\begin{itemize}
-			\item \emph{\num{1.5} weeks.}
-			\item Policies need not be meaningful, and can be garbage data. Use-case independent.
-			\item Partially blocked on multicore accelerated/optimised execution.
-		\end{itemize}
-		\item Measure rule installation cost on different platforms.
-		\begin{itemize}
-			\item \emph{\num{1.5} weeks.}
-			\item Blocked on creation of basic testbed and packet forwarding setup.
-		\end{itemize}
-		\item Investigate stationarity.
-		\begin{itemize}
-			\item \emph{\numrange{2}{3} weeks.}
-			\item Requires full use-case implementation.
-			\item Requires careful thought about sensible changes to normality.
-			\item Might be bolstered by comparison against an offline-trained system?
-		\end{itemize}
-	\end{itemize}
-
-	\section{Everything we could possibly evaluate}
-
-	Evaluation will need to be divided into 2 main categories:
-	\begin{enumerate}
-		\item Simulation/emulation based on mininet.
-		\begin{itemize}
-			\item These can only assess algorithmic properties, as performance measurements in mininet (i.e., a full mult-agent system) are not representative.
-		\end{itemize}
-		\item Testbed setup (2/3 configurations):
-		\begin{itemize}
-			\item These will assess key system performance properties of a \emph{single agent}, in control of a single switch/NIC as a bump in the wire.
+%		\end{itemize}
+%		\item Fixes to hash table use, signalling.
+%		\begin{itemize}
+%			\item \emph{\num{1} week.}
+%			\item Blocks progress on optimisation, correct multi-stream RL.
+%		\end{itemize}
+%		\item Optimise/parallelise action compute/learning.
+%		\begin{itemize}
+%			\item \emph{\num{2} week.}
+%			\item Value? Proves this approach can scale according to resource availability/demand on the deployment device, at compile-time anyhow.
+%		\end{itemize}
+%	\end{itemize}
+%	
+%	Experiments?
+%	\begin{itemize}
+%		\item Measure sum of times $t_{1...3}$ described elsewhere.
+%		\begin{itemize}
+%			\item \emph{\num{1} week.}
+%			\item Currently have individual, need ``end-to-end''.
+%		\end{itemize}
+%		\item Throughput impact testing.
+%		\begin{itemize}
+%			\item \emph{\num{1} week.}
+%			\item Requires current impl, creation of basic testbed and packet forwarding setup.
+%			\item Use-case independent.
+%		\end{itemize}
+%		\item Measure execution, training costs for varying policy sizes.
+%		\begin{itemize}
+%			\item \emph{\num{1.5} weeks.}
+%			\item Policies need not be meaningful, and can be garbage data. Use-case independent.
+%			\item Partially blocked on multicore accelerated/optimised execution.
+%		\end{itemize}
+%		\item Measure rule installation cost on different platforms.
+%		\begin{itemize}
+%			\item \emph{\num{1.5} weeks.}
+%			\item Blocked on creation of basic testbed and packet forwarding setup.
+%		\end{itemize}
+%		\item Investigate stationarity.
+%		\begin{itemize}
+%			\item \emph{\numrange{2}{3} weeks.}
+%			\item Requires full use-case implementation.
+%			\item Requires careful thought about sensible changes to normality.
+%			\item Might be bolstered by comparison against an offline-trained system?
+%		\end{itemize}
+%	\end{itemize}
+%	
+%	\section{Everything we could possibly evaluate}
+%	
+%	Evaluation will need to be divided into 2 main categories:
+%	\begin{enumerate}
+%		\item Simulation/emulation based on mininet.
+%		\begin{itemize}
+%			\item These can only assess algorithmic properties, as performance measurements in mininet (i.e., a full mult-agent system) are not representative.
+%		\end{itemize}
+%		\item Testbed setup (2/3 configurations):
+%		\begin{itemize}
+%			\item These will assess key system performance properties of a \emph{single agent}, in control of a single switch/NIC as a bump in the wire.
 Bump-in-the-wire device will carry traffic between two hosts.
-			\item \emph{Time permitting, may need to send state vectors in packets rather than do feature/state extraction on-chip}.
-			\item \emph{Necessary.} RL runs on bump-in-the-wire SmartNIC.
-			\item \emph{Either/both.} P4-based off-path RL.
+%			\item \emph{Time permitting, may need to send state vectors in packets rather than do feature/state extraction on-chip}.
+%			\item \emph{Necessary.} RL runs on bump-in-the-wire SmartNIC.
+%			\item \emph{Either/both.} P4-based off-path RL.
 Custom packet actions (e.g., random drop) will be implemented in P4 with externs.
@@
 Rules inserted via P4-device specific RTE.
-			\item \emph{Either/both.} OvS-based off-path RL.
+%			\item \emph{Either/both.} OvS-based off-path RL.
 Custom packet actions are already implemented in OvS as OpenFlow extensions.
@@
 Agent may be co-hosted with state-extraction.
-		\end{itemize}
-	\end{enumerate}
-
-	\subsection{What is this different from, or better than?}
-	List below mainly for the benefit of helping enumerate everything I can compare against, and what their drawbacks are.
-
-	\begin{itemize}
-		\item (Deep) neural network-backed RL.
-		\begin{itemize}
-			\item NNs infeasible to update online.
+%		\end{itemize}
+%	\end{enumerate}
+%	
+%	\subsection{What is this different from, or better than?}
+%	List below mainly for the benefit of helping enumerate everything I can compare against, and what their drawbacks are.
+%	
+%	\begin{itemize}
+%		\item (Deep) neural network-backed RL.
+%		\begin{itemize}
+%			\item NNs infeasible to update online.
 Ordinarily this requires larger volumes of data than can be gathered in real time, but becomes substantially more difficult after change of representation (quantisation, binarisation).
 An online system (such as this) should be able to adapt to changes in underlying traffic, optimisation criteria, etc., dynamically.
-			\item Complexity of function approximation means that policy updates are more computationally complex to compute.
+%			\item Complexity of function approximation means that policy updates are more computationally complex to compute.
 For instance, NNs require use of the backpropagation algorithm, rather than the simpler classical schemes which need only the tile set already identified in action selection.
-			\item Only heavily quantised (i.e., binary neural networks) suitable for in-switch execution.
-		\end{itemize}
-
-		\item MARL and other ``agents as network functions'' designs.
-		\begin{itemize}
-			\item Action, state collection, updates etc.
+%			\item Only heavily quantised (i.e., binary neural networks) suitable for in-switch execution.
+%		\end{itemize}
+%		
+%		\item MARL and other ``agents as network functions'' designs.
+%		\begin{itemize}
+%			\item Action, state collection, updates etc.
 happen on nodes outside of the main packet path in this paradigm.
@@
 Also makes these approaches ``slower to react'' in theory.
-		\end{itemize}
-	\end{itemize}
-
-	At present, there are no other RL works in PDPs, let alone with online policy updates.
-	Other existing RL use cases suitable for adaptation to the ``bump-in-the-wire'' model are proving tricky to find.
-
-	\subsection{Emulation metrics}
-	Key (MVP) metrics will be \textbf{bolded}, and expanded with a subsubsection.
-
-	\textbf{Quantisation accuracy}, quantised training accuracy, \textbf{stationarity}.
-
-	\subsubsection{Quantisation accuracy}
-	Examination of whether lower resolution quantisation impacts RL agent decision accuracy substantially.
-
-	This has been done using MARL in mininet, see \cref{fig:quant-acc}.
-
-	\subsubsection{Interaction with stationarity}
-	Learning time of the system as compared to known changes in normality.
-
-	I.e., in the DDoS case: how long does an agent take to respond to the onset of an attack?
-	A change in attack strategy?
-	The conclusion of an attack?
-	There are no citations on \emph{how} DDoS attacks evolve, just that they do, and that this can occur on various timescales.
-
-	\subsection{Testbed metrics}
-	Key (MVP) metrics will be \textbf{bolded}, and expanded with a subsubsection.
-
-	\textbf{Traffic throughput (bytes, packets)}, \textbf{packet processing delay}, \textbf{action execution time}, \textbf{learning time}, from-scratch learning accuracy, \textbf{rule installation delay/cost}, host and NIC memory costs, policy installation time, ``accuracy'' in the DDoS case.
-
-	\subsubsection{Traffic throughput}
-	Main research question: does moving RL logic onto the NIC reduce maximum throughput of the NIC, even if calculation is on a different functional unit?
-	This is relevant in both \si{\byte\per\second} and packets per second.
-
-	This is application independent, and would assess whether the RL units (acting independently on the netronome) affect throughput compared with a basic P4 program at different work intensities.
-	Ideally, this would show no effect on forwarding performance.
-
-	\subsubsection{Packet processing delay}
-	Main research question: does the presence of off-path (but on-chip) RL execution meaningfully increase packet processing delay?
+%		\end{itemize}
+%	\end{itemize}
+%	
+%	At present, there are no other RL works in PDPs, let alone with online policy updates.
+%	Other existing RL use cases suitable for adaptation to the ``bump-in-the-wire'' model are proving tricky to find.
+%	
+%	\subsection{Emulation metrics}
+%	Key (MVP) metrics will be \textbf{bolded}, and expanded with a subsubsection.
+%	
+%	\textbf{Quantisation accuracy}, quantised training accuracy, \textbf{stationarity}.
+%	
+%	\subsubsection{Quantisation accuracy}
+%	Examination of whether lower resolution quantisation impacts RL agent decision accuracy substantially.
+%	
+%	This has been done using MARL in mininet, see \cref{fig:quant-acc}.
+%	
+%	\subsubsection{Interaction with stationarity}
+%	Learning time of the system as compared to known changes in normality.
+%	
+%	I.e., in the DDoS case: how long does an agent take to respond to the onset of an attack?
+%	A change in attack strategy?
+%	The conclusion of an attack?
+%	There are no citations on \emph{how} DDoS attacks evolve, just that they do, and that this can occur on various timescales.
+%	
+%	\subsection{Testbed metrics}
+%	Key (MVP) metrics will be \textbf{bolded}, and expanded with a subsubsection.
+%	
+%	\textbf{Traffic throughput (bytes, packets)}, \textbf{packet processing delay}, \textbf{action execution time}, \textbf{learning time}, from-scratch learning accuracy, \textbf{rule installation delay/cost}, host and NIC memory costs, policy installation time, ``accuracy'' in the DDoS case.
+%	
+%	\subsubsection{Traffic throughput}
+%	Main research question: does moving RL logic onto the NIC reduce maximum throughput of the NIC, even if calculation is on a different functional unit?
+%	This is relevant in both \si{\byte\per\second} and packets per second.
+%	
+%	This is application independent, and would assess whether the RL units (acting independently on the netronome) affect throughput compared with a basic P4 program at different work intensities.
+%	Ideally, this would show no effect on forwarding performance.
+%	
+%	\subsubsection{Packet processing delay}
+%	Main research question: does the presence of off-path (but on-chip) RL execution meaningfully increase packet processing delay?
 How does this compare to a P4 program with an equivalent number of tables?
 What about to OvS?
-
-	\subsubsection{Action execution \& learning time}
-	Main research question: does moving RL logic from the host to the NIC substantially reduce delays from state measurement, to state receipt, to action computation and installation?
-	These constitute $t_{1...3}$ discussed above.
-	Primarily, these need to be looked at using varying policy sizes and distribution across memory regions.
-
-	We ideally want these both with and without any acceleration from extra cores on the netronome: the point of this being that we can demonstrate that the system (or a similar NetFPGA system) can scale upwards according to available system resources or demands.
-
-	\subsubsection{Rule installation delay/cost.}
-	Main research question: how long does it take to install rules in the above testbed setups? What is the downtime of so doing?
-
-	This is application-dependent: see earlier discussions on rule batching and local extern/hash-table use.
-\end{appendices}
+%	
+%	\subsubsection{Action execution \& learning time}
+%	Main research question: does moving RL logic from the host to the NIC substantially reduce delays from state measurement, to state receipt, to action computation and installation?
+%	These constitute $t_{1...3}$ discussed above.
+%	Primarily, these need to be looked at using varying policy sizes and distribution across memory regions.
+%	
+%	We ideally want these both with and without any acceleration from extra cores on the netronome: the point of this being that we can demonstrate that the system (or a similar NetFPGA system) can scale upwards according to available system resources or demands.
+%	
+%	\subsubsection{Rule installation delay/cost.}
+%	Main research question: how long does it take to install rules in the above testbed setups? What is the downtime of so doing?
+%	
+%	This is application-dependent: see earlier discussions on rule batching and local extern/hash-table use.
+%\end{appendices}
 

### 2021-05-29 code/nfp/rl-test/stress.c
Accidentally had all the stress FWs going at full pelt...
now THIS needs a rerun...

@@ -28,6 +28,8 @@ volatile __declspec(import, emem, addr40, aligned(sizeof(unsigned int))) uint8_t
 
 volatile __declspec(import, emem) uint32_t begin_stress;
 
+__declspec(export, emem) uint64_t pkt_interval;
+
 main() {
 	uint32_t t0, t1;
 	__declspec(write_reg) struct rl_work_item workq_write_register = {0};
@@ -45,6 +47,8 @@ main() {
 
 	work_to_send.ctldata.t4_type = PIF_PARREP_TYPE_in_state;
 	work_to_send.parsed_fields.state.dim_count = RL_DIMENSION_MAX;
+
+	pkt_interval = STRESS_TEST_INDIV_GAP_RCYCLES;
 
 	// wait for valid config
 	while (1) {
@@ -100,9 +104,11 @@ main() {
 		// record t1
 		t1 = local_csr_read(local_csr_timestamp_low);
 
+		// elapsed = t1 - t0
 		if (t1 - t0 < STRESS_TEST_INDIV_GAP_RCYCLES) {
 			// Big note: this takes actual cycles, but the asm blocks take 16*cycles.
-			sleep((t1 - t0) << 4);
+			// sleep for (total - elapsed)
+			sleep((STRESS_TEST_INDIV_GAP_RCYCLES - (t1 - t0)) << 4);
 		}
 	}
 }

### 2021-05-29 paper/paper.tex
Some results work.

@@
 
-Programmable data-planes and in-switch compute, then, hold promise for integrating these new techniques in a feasible and efficient manner (beyond dedicated servers or virtualised network functions).
+Programmable dataplanes and in-switch compute, then, hold promise for integrating these new techniques in a feasible and efficient manner (beyond dedicated servers or virtualised network functions).
 For RL, key tasks include policy evaluation, online training, state collection, and action execution -- each of these introduces some degree of sensitivity to state accuracy.
@@
 
-The solution we propose is to make use of the recent wave of programmable network devices to \emph{bring reinforcement learning to the data plane}---referring again to \cref{fig:state-slip}, we would place place state measurement ($t_1$), low-cost decision-making processes ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
+The solution we propose is to make use of the recent wave of programmable network devices to \emph{bring reinforcement learning to the dataplane}---referring again to \cref{fig:state-slip}, we would place place state measurement ($t_1$), low-cost decision-making processes ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
 In networks, actions are most likely to be installed on backbone switches, bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts.
@@
 Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
-Lower bit-depths reduce memory footprints, and improve throughput in designs such as \emph{bfloat16}~\parencite{bfloat16-blog} in Google TPUs~\parencite{DBLP:journals/sigops/XieDMKVZT18}, \emph{hfp8}~\parencite{DBLP:conf/nips/SunCCWVSCZG19}, and other floating point formats~\parencite{DBLP:journals/corr/abs-2007-01530}.
+Lower bit depths reduce memory footprints, and improve throughput in designs such as \emph{bfloat16}~\parencite{bfloat16-blog} in Google TPUs~\parencite{DBLP:journals/sigops/XieDMKVZT18}, \emph{hfp8}~\parencite{DBLP:conf/nips/SunCCWVSCZG19}, and other floating point formats~\parencite{DBLP:journals/corr/abs-2007-01530}.
 In many cases, accuracy losses for doing so are negligible.
@@
 This is analogous to SIMD---through clever use of padding bits---but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
-We investigate further performance implications of changing bit-depth in the sequel.
+We investigate further performance implications of changing bit depth in the sequel.
 
@@
 Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
-As our later results show, working on a bitdepth below the native word size forces the compiler or programmer to emit excess code to extract and re-align preference values.
+As our later results show, working on a bit depth below the native word size forces the compiler or programmer to emit excess code to extract and re-align preference values.
 This reduces overall efficiency even though considerably fewer reads are made.
@@
 
-?? Bitdepth is important for single item computation in the simpler \Indfw{} strategy.
+?? Bit depth is important for single item computation in the simpler \Indfw{} strategy.
 
@@
 	}
-	\caption{??
-More cores means less time ??
-Max-size work (129 work-items).
-??
-Need to dedicate 3 worker to beat independent cores.
-??
-Sharper perf increase when new physical core added.
-??
-up to 31 due to master thread.
-??
-latency version in appendix: \Cref{fig:vary-core-latency}.\label{fig:vary-core}}
+	\caption{
+		\Coopfw{}'s online learning performance improves as additional cores are made available, on max size tasks (\num{129} work items).
+This requires \num{8} workers to offer greater online throughput than single-threaded in-NIC RL.
+Sharper performance increases occur when new physical core is added (\numrange{7}{8}) or the scheduler works around a bottleneck (\numrange{13}{14}).\label{fig:vary-core}}
 \end{figure}
@@
 	}
-	\caption{?? More work means More time. ?? Max Parallelisation. ?? Bitdepth affects independent cores more. ?? latency version in appendix: \Cref{fig:vary-work-latency}.\label{fig:vary-work}}
+	\caption{?? More work means More time. ?? Max Parallelisation. ?? Bit depth affects independent cores more. ?? latency version in appendix: \Cref{fig:vary-work-latency}.\label{fig:vary-work}}
 \end{figure}
+
+We found that a key aspect of in-NIC execution is that it allows far tighter bounds on tail latency to be offered compared to host offloading.
+Examining the state-action latencies in \cref{tab:lats}, we see that \num{99.99}\nthscript{th} percentile latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{} respectively.
+Similar results were observed for other bit depths.
+By contrast, host-based tail latencies are at least \SI{40.53}{\percent} greater in our test setup when parallel workers are at or below the physical core count.
+We show the cumulative distributions of these in detail in \cref{fig:lat-cumul}, noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF or traffic processing task---can impact tail latencies further (\emph{Float(Over)}).
 
@@
 	}
-	\caption{Latency! varying core count\label{fig:vary-core-latency}}
+	\caption{\Coopfw{}'s action latencies improve as additional cores are made available.
+This requires \num{3} workers (4 total contexts) to improve upon single-threaded in-NIC RL.\label{fig:vary-core-latency}}
 \end{figure}
@@
 	\end{subfigure}
-	\caption{Cumulative state-action latency plots for \approachshort{} and host-based execution.}
+	\caption{Cumulative state-action latency plots for \approachshort{} and host-based execution.\label{fig:lat-cumul}}
 \end{figure}
@@
 \Cref{tab:resources} demonstrates how \approachshort{} consumes shared and local memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
-As one program is installed per ME, these represents the minimum and maximum resource use we can command on a single island (i.e., without replacing P4 workers).
+As one program is installed per ME, these represent the minimum and maximum resource use we can command on a single island (i.e., without replacing P4 workers).
 We observe negligible costs on shared bulk storage ($\sim$\SI{4}{\mebi\byte}), including storage space for a hash-table of \num{4096} state-action pairs and \num{16} reward values.
 The most significant costs arise due to policy data (\SI{405}{\kibi\byte} capacity shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
-Thread local storage in CLS for compute/register spilling scales with the number of MEs required, and with \Coopfw{} this is far below the ``fair share'' (allowing co-existence with other asynchronous dataplane programs).
-Notably, \Indfw{} imposes far heavier requirements on the limited per-island CLS, due to the extra resources needed to maintain, operate on, and cache tile lists for optimal processing.
+Thread local storage in CLS for compute/register spilling scales with the number of MEs required, particularly when caching  and with \Coopfw{} this is far below the ``fair share'' (allowing co-existence with other asynchronous dataplane programs).
 
@@
 
-We observed that firmware installation (i.e., changing from \Indfw{} to \Coopfw{}, bit-depth, or increasing maximum policy sizes) took a mean time of \SI{38.83}{\second}.
+We observed that firmware installation (i.e., changing from \Indfw{} to \Coopfw{}, bit depth, or increasing maximum policy sizes) took a mean time of \SI{38.83}{\second}.
 ?? What about compile time?
@@
 
-?? Can I have a P4 code snippet here?
+%?? Can I have a P4 code snippet here?
 
@@
 
-?? NOTES: try to strengthen emphasis on ``other works can't do online learning in the data plane''. This is absolutely, \SI{100}{\percent} the killer feature.
+?? NOTES: try to strengthen emphasis on ``other works can't do online learning in the dataplane''. This is absolutely, \SI{100}{\percent} the killer feature.
 
@@
 ?? N3IC run NNs on NIC~\parencite{DBLP:journals/corr/abs-2009-02353}, allow packet tagging to occur based on pre-trained BNN~\parencite{DBLP:conf/nips/HubaraCSEB16}.
-?? both NFP and NetFPGA.
+?? both NFP (\SI{45}{\micro\second}) and NetFPGA (\SI{0.3}{\micro\second}).
+?? We handle larger inputs (versus their \SI{256}{\bit}), offer online training.
+?? Use our 5-dim as a point of comparison?
+
 ?? Or as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.

### 2021-05-31 results/process-stress.py
Trying out relative plots...

@@ -33,6 +33,26 @@ def write_matrix(name, matrix, w, h):
 			#	of.write("{} {} {}\n".format(x, y, val))
 			of.write(" ".join([str(x) for x in row]) + "\n")
 
+	rel = shaped - shaped[0]
+	with open(out_dir + name + "rel.dat", "w") as of:
+		for x, row in enumerate(rel):
+			#if not x == 0:
+			#	of.write("\n")
+
+			#for y, val in enumerate(row):
+			#	of.write("{} {} {}\n".format(x, y, val))
+			of.write(" ".join([str(x) for x in row]) + "\n")
+
+	perc = (shaped / shaped[0]) - 1.0
+	with open(out_dir + name + "perc.dat", "w") as of:
+		for x, row in enumerate(perc):
+			#if not x == 0:
+			#	of.write("\n")
+
+			#for y, val in enumerate(row):
+			#	of.write("{} {} {}\n".format(x, y, val))
+			of.write(" ".join([str(x) for x in row]) + "\n")
+
 print("Latencies!")
 
 matrices = {}

### 2021-05-31 results/process-stress.py
Tried to replot with Big Fonts

@@ -43,7 +43,7 @@ def write_matrix(name, matrix, w, h):
 			#	of.write("{} {} {}\n".format(x, y, val))
 			of.write(" ".join([str(x) for x in row]) + "\n")
 
-	perc = (shaped / shaped[0]) - 1.0
+	perc = ((shaped / shaped[0]) - 1.0) * 100.0
 	with open(out_dir + name + "perc.dat", "w") as of:
 		for x, row in enumerate(perc):
 			#if not x == 0:

### 2021-05-31 results/process-stress.py
Change to microseconds in histos (for my sanity)

@@ -130,7 +130,7 @@ for pkt_size in pkt_sizes:
 			histo = np.histogram(lats_for_this_cell, bins)
 
 			for (bin_x, freq) in zip(bins, histo[0]):
-				of.write("{},{}\n".format(int(bin_x), int(freq)))
+				of.write("{},{}\n".format(bin_x / 1000.0, int(freq)))
 
 for key, value in matrices.items():
 	write_matrix(key, value, len(pkt_sizes), rate_max)

### 2021-05-31 paper/paper.tex
Some minor changes to stats extraction.

@@
 \sisetup{detect-all, range-phrase=--, range-units=single, detect-weight=true,detect-inline-weight=math, table-format=1.3}
+\DeclareSIUnit{\packet}{p}
 
@@
 \fakepara{Co-existence with the dataplane}
-No packet loss incurred
-?? mention actually achieved rates?
+We were able to saturate the \SI{40}{\giga\bit\per\second} link for packet sizes above \SI{256}{\byte}.
+For frame sizes of \SIlist{64;128;256}{\byte} input traffic rates were \SIlist{17.4;32.9;37.0}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2;18.1}{\mega\packet\per\second}).
+Passing traffic over the loaded NFP device running \approachshort, no packet losses were incurred at any rate of RL actions.
+
+We show the effect of RL workloads on the round-trip latencies of cross traffic through \cref{fig:dataplane-coop}.
+As observed latency measures do not obey a normal distribution, we employ a one-tailed \emph{Mann-Whitney U test} to mark statistically significant population increases in latency ($p < 0.05$) with a "+".
+In general, we observe that statistically significant latency increases concentrate around smaller packet sizes.
+All (bar one) of these affected \nth{99} percentile latencies by under \SI{0.38}{\percent} (at most \SI{78}{\nano\second}).
+
+Is this anomaly just burstiness? It's \SI{222}{\nano\second} (1.175 percent)
+
+?? Cross-island/chip traffic must cross command push pull bus.
+
+\cref{fig:dataplane-example} shows one such example...
 
 ?? Now talk about statistically significant deltas etc.
+
+\cref{fig:dataplane-coop}
+
+?? Odd two-tailed behaviour on 1518B
+
+?? Why spike in significance at 256? Likely due to packet segmentation in NFP (i.e., split between memory regions with body in EMEM).
 
@@
 %			\includegraphics{../plots/build/stress/heat-latency-median}
-			\includegraphics{../plots/build/stress/heat-latency-two-9}
-		}
-		\caption{Ayy?}
+			\includegraphics{../plots/build/stress/heat-latency-two-9-rel}
+%			\includegraphics{../plots/build/stress/heat-latency-two-9-perc}
+		}
+		\caption{Deviations in \nth{99} percentile cross-traffic latencies.\label{fig:dataplane-heat}}
 	\end{subfigure}
@@
 		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/stress/histo-1518B-0-10}
-		}
-		\caption{Ayy?}
+			\includegraphics{../plots/build/stress/histo-128B-0-3}
+		}
+		\caption{Distribution of round-trip latencies for \SI{128}{\byte} packets.\label{fig:dataplane-example}}
 	\end{subfigure}
-	\caption{Ayy.}
+	\caption{At most \SI{1.5}{\percent}... "+" symbols denote statistically significant ($p < 0.05$) increases in population latency, concentrated on smaller packet sizes.\label{fig:dataplane-coop}}
 \end{figure}

### 2021-05-31 results/process-stress.py

@@ -156,18 +156,25 @@ print("Loss rates!")
 # process mean losses?
 for pkt_size in pkt_sizes:
 	losses_baseline = None
+	rates = []
+	pkts = []
 	for rate in range(rate_max):
 		losses_for_this_cell = []
+
 		for i in range(iter_ct):
 			with open(stress_result_dir + stress_file_tpu.format(pkt_size, rate, i)) as of:
 				lines = of.readlines()
 				imissed = extract_from_line(lines[3])
 				ipackets = extract_from_line(lines[4])
+				obytes = extract_from_line(lines[5])
 				opackets = extract_from_line(lines[7])
 
 				made_it_to_nic = imissed + ipackets
 
 				losses_for_this_cell += [1.0 - (made_it_to_nic / opackets)]
+
+				pkts.append(opackets)
+				rates.append(((float(obytes) / 30.0) * 8.0) / 1e9)
 
 		p_val = None
 		if losses_baseline is None:
@@ -184,3 +191,4 @@ for pkt_size in pkt_sizes:
 			p_val,
 		]
 		print("{}B {}k: {}".format(pkt_size, rate, metrics))
+	print("SPEEDS: {}B: {} {}".format(pkt_size, np.mean(rates), np.mean(pkts) / 30e6))

### 2021-05-31 paper/paper.tex
A hard day's work
Most of the results spiel in.

--- 2021-05-29 paper/paper.tex
@@
 We also report on the derived throughput measures accounting for the locking behaviour of each.
+We compare online learning performance (input states produce an output, and then update the policy with appropriate locking) with offline (input states \emph{only} produce an output) in these cases.
+State-action latency is a shared property of both cases, with the main difference being the impact on throughput of the update step.
 
@@
 
-We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, using an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source via the default firmware.
+We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, placing an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source via the default firmware.
 We perform throughput/loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements.
+DPDK was setup using four \SI{1}{\gibi\byte} hugepages.
 Sent traffic was comprised of fixed-size \SIrange{64}{1518}{\byte} packets~\cite{rfc2544}.
@@
 \fakepara{Raw inference and learning performance}
-\Cref{tab:lats} shows that \Coopfw{} \approachshort{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse.
-As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, while the simpler \Indfw{} succeeds when policies are fixed/known.
-?? Crucially, our using many extra functional units to tackle this task
-
-?? Bit depth is important for single item computation in the simpler \Indfw{} strategy.
-
-??
-We observed that \SIlist{8;16}{\bit} have worse action compute/latency, but better update processing time in \SI{16}{\bit}.
-Why?
-Throughput and batching of value reads only goes so far, native register size for ALU is \SI{32}{\bit}, so these wins are ``undone'' to some extent by the compiler extracting/shifting parts  for correctness.
-Wins are mainly due to serial part scaling with bitsize.
-
-?? Mention versus DNN latency in passing here?
-
-?? Need to mention some of the omitted results here for brevity; namely host integer implementations.
+\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based (BLAS-accelerated) implementation of an RL policy\footnote{For brevity, we omit numpy-based integer results---summarised against a floating point implementation, median state-action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}.
+We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using one island (4 MEs) of the NFP-6480.
+This corresponds to \SIlist{15.0;21.0}{$\times$} speedups over a \emph{Collector} host.
+Importantly, \Indfw{} achieves lower median state-action latencies (\SI{2.79}{$\times$}) \emph{and} update times (\SI{2.63}{$\times$}) than a dedicated \emph{Collector} host \emph{while requiring only a single core or dedicated functional unit}.
 
@@
 
-\Cref{tab:tputs} shows how \approachshort{} compares in terms of throughput with a numpy-based (i.e., BLAS-accelerated) implementation of an RL policy.
-We choose the worker count on host machines with the largest number of parallel agents which produced a statistically significant throughput increase ($p < 0.05$).
-We observed that this equalled the number of physical cores on each device, while moving beyond this (even below the number of hyper-threads) would hamper tail latencies by an order of magnitude.
+\Cref{tab:tputs} compares \approachshort{}'s throughput against host-based execution.
+We choose the worker count on host machines as the number of parallel agents that maximises throughput.
+This equalled the amount of physical cores on each device, while moving beyond this (even below the number of hyper-threads) would hamper tail latencies by an order of magnitude.
 To make the comparison fairer in the context of many-core CPU environments, we include per-core throughput.
-We observed that ...
-
-?? \Cref{fig:vary-core} shows how action+update compute time scales with available cores.
-?? (Core count Limited by software).
-?? Extra \emph{physical core} additions (rather than logical workers) very visible.
-?? 11 workers needed to process items faster than \Indfw{}, but 5 workers needed to bring \emph{latency} below \Indfw{} ()
-?? TAKEAWAY: resource demand tunable to meet latency demands for workload.
+\Indfw{} achieves \SI{2.82}{$\times$} higher offline throughput than commodity \emph{Collector} hardware in spite of the NFP-6480 having a considerably slower clock speed (\SI{0.29}{$\times$}).
+When compounded with the abundance of such weaker chips, in-NIC RL is able to deliver much higher throughput.
+As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} a dedicated collector machine, due to the locking requirements mandating coordinated division of work.
+
+By limiting the available workers in software, we show how \Coopfw{}'s policy update time (thus  online throughput---\cref{fig:vary-core}) and state-action latency (\cref{fig:vary-core-latency}) scale with available cores.
+While the compute times and latency of \Coopfw{} always outperform the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency.
+Some artefacts of our environment and design choices are visible, such as the addition of new physical cores (MEs) being more significant than contexts, and the presence of some schedule bottlenecks.
+Most importantly however, \approachshort{}-\Coopfw{}'s resource demand is tunable at compile time to meet the online training rate and/or action latency required by a task/environment.
 
@@
 
-?? \Cref{fig:vary-work} shows how action+update compute time of any individual work item scales with the complexity of the work.
-?? Note: input vectors still have full size of 20 elements.
-?? Lower bit depth leads to a \SIrange{5}{10}{\micro\second} improvement in less complex policies, but these converge with policy complexity.
-?? Relate this to latency? \SI{32}{\bit} marginally ``wins out'' there..
-?? In latency if \Indfw{}, 32 == 8 < 16 (hmm).
+\begin{figure}
+	\resizebox{1.0\linewidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
+	}
+	\caption{\Coopfw{}'s action latencies improve as additional cores are made available.
+This requires \num{3} workers (4 total contexts) to improve upon the state-action latency of single-threaded in-NIC RL.\label{fig:vary-core-latency}}
+\end{figure}
+
+\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects complete update costs and state-action latency respectively, scaling from a bias tile up to the full policy size used throughout\footnote{Input vectors here all have 20 elements regardless of the policy design.}.
+\Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning performance.
+We note that this is a trivial case, in that the use of \emph{only} a bias tile negates the need for any input state (similar to a multi-arm bandit problem).
 
@@
 	}
-	\caption{?? More work means More time. ?? Max Parallelisation. ?? Bit depth affects independent cores more. ?? latency version in appendix: \Cref{fig:vary-work-latency}.\label{fig:vary-work}}
+	\caption{\Coopfw{} can fully process individual items faster than \Indfw{}---thus has higher online performance---on almost all policy sizes.
+Lower bit depths are more effective on simpler policies.\label{fig:vary-work}}
 \end{figure}
@@
 	}
-	\caption{Latency! varying work.\label{fig:vary-work-latency}}
-\end{figure}
-
-\begin{figure}
-	\resizebox{1.0\linewidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
-	}
-	\caption{\Coopfw{}'s action latencies improve as additional cores are made available.
-This requires \num{3} workers (4 total contexts) to improve upon single-threaded in-NIC RL.\label{fig:vary-core-latency}}
+	\caption{State-action latency scales with additional work in a similar manner to overall processing time; though \SI{32}{\bit} firmwares become more effective sooner (at \num{3} work items).\label{fig:vary-work-latency}}
 \end{figure}
@@
 Regardless, the reduction in memory needed to store a policy remain.
+This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) \Coopfw{} agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the algorithm.
 
@@
 ?? TODO
+?? BIGGEST TO DO.
+
+?? PCIe RTT is 10us (neugebauer, BNN NFP paper), NFP RTT is 18us (own measurements). EMEM Ring one-way is 120ns.
+
+?? In the name of fairness, assume that rule installation uses same mechanism we recommend, but show how bad it can get? I.e. with rule installation costs etc.
 
 \fakepara{Co-existence with the dataplane}
-No packet loss incurred
-?? mention actually achieved rates?
-
-?? Now talk about statistically significant deltas etc.
+We were able to saturate the \SI{40}{\giga\bit\per\second} link for packet sizes at or above \SI{256}{\byte}.
+%For frame sizes of \SIlist{64;128;256}{\byte} input traffic rates were \SIlist{17.4;32.9;37.0}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2;18.1}{\mega\packet\per\second}).
+For frame sizes of \SIlist{64;128}{\byte} input traffic rates were \SIlist{17.4;32.9}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2}{\mega\packet\per\second}).
+Passing this traffic over the loaded NFP device running \approachshort, no packet losses were incurred at any rate of RL actions.
+
+We show the effect of RL workloads on the round-trip latencies of cross traffic through \cref{fig:dataplane-coop}.
+As observed latency measures do not obey a normal distribution (particularly \SIlist{256;1518}{\byte}, which are bimodal), we employ a one-tailed \emph{Mann-Whitney U test} to mark statistically significant population increases in latency ($p < 0.05$) with a ``+''.
+In general, we observe that statistically significant latency increases concentrate around smaller packet sizes.
+All (bar one) of these affected \nth{99} percentile latencies by under \SI{0.38}{\percent} (at most \SI{78}{\nano\second}).
+This slight general degradation can be explained by increased pressure on the NFP's \emph{command push-pull} (CPP) bus, which is responsible for handling cross-island accesses to memory (particularly IMEM/EMEM) and other resources.
+\approachshort{} places load on the CPP bus through use of its IN/OUT rings for external communication and through last-tier policy accesses.
+This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---the Netronome P4 toolchain segments packets, storing metadata (e.g., MAC prepend data) and the first bytes of a packet in a $\SI{256}{\byte}$ CTM block and parking the bodies of processed packets in EMEM.
+\SI{256}{\byte} packets overshoot this due to metadata, causing small I/O accesses at a high rate for packets sized around this cutoff point.
+Regardless, this effect is small when observed.
+
+The anomalous result is \SI{128}{\byte} packets at \num{3000} RL action/update computations per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase, shown in \cref{fig:dataplane-example}.
+In the above context, we believe that the inbound request rate is weakly synchronised with inbound packets, causing a higher level of burstiness around accesses to the CPP bus.
+We expect that dedicated hardware/FPGA designs can avoid this problem by having dedicated In/Out access mechanisms for an \approachshort{} agent.
 
@@
 %			\includegraphics{../plots/build/stress/heat-latency-median}
-			\includegraphics{../plots/build/stress/heat-latency-two-9}
-		}
-		\caption{Ayy?}
+			\includegraphics{../plots/build/stress/heat-latency-two-9-rel}
+%			\includegraphics{../plots/build/stress/heat-latency-two-9-perc}
+		}
+		\caption{Deviations in \nth{99} percentile cross-traffic RTTs.\label{fig:dataplane-heat}}
 	\end{subfigure}
@@
 		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/stress/histo-1518B-0-10}
-		}
-		\caption{Ayy?}
+			\includegraphics{../plots/build/stress/histo-128B-0-3-trim}
+		}
+		\caption{Distribution of RTTs for \SI{128}{\byte} packets for baseline and \num{3000} RL actions/s.\label{fig:dataplane-example}}
 	\end{subfigure}
-	\caption{Ayy.}
+	\caption{Effects on tail latency of cross-traffic caused by different loads of off-path RL compute.
+Statistically significant increases in population latency are concentrated on smaller packet sizes, and are typically sub-\SI{77}{\nano\second}.\label{fig:dataplane-coop}}
 \end{figure}
@@
 \fakepara{Deployability}
-??
-Setup and tiling packets take a mean \SI{26.65}{\micro\second} to be processed on \Indfw{}, while this processing and heuristic work scheduling takes \SI{417.25}{\micro\second} for the full-size policy described above.
-
-?? Config times scale with work complexity in \Coopfw{}, confirming the $\mathcal{O}{\left(n\right)}$ complexity of the allocator (??). Is it O mlogn?
-?? Changing from online to offline has this cost.
+We found that setup and tiling packets take a mean \SIlist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing for an agent to be swapped between online and offline operation (or repurposed for another task/policy) painlessly.
+Tiling packet processing was found to scale linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
+Online-offline swaps for \Coopfw{} exhibit similar cost, however the need for explicit ahead-of-time heuristic scheduling means that policy/tiling \emph{structure} changes (including the first complete setup) take \SI{422.63}{\micro\second} for the full-size policy described above.
+The time taken for \Coopfw{} to schedule its tasks was found to scale with the number of workers ($m$) and work items ($n$) as described above ($\mathcal{O}{\left(n\log{m}\right)}$).
+Ignoring the trivial solutions, reducing the worker count to \num{1} costs a mean \SI{238.22}{\micro\second}, while placing a single task incurs \SI{53.79}{\micro\second}.
+Policy \emph{data} changes require no additional work in any case, resolving purely to \texttt{memcpy}s.
 
 We observed that firmware installation (i.e., changing from \Indfw{} to \Coopfw{}, bit depth, or increasing maximum policy sizes) took a mean time of \SI{38.83}{\second}.
-?? What about compile time?
+In the event that appropriate firmwares are not pre-compiled, we found that compiling and linking both \approachshort{} and the P4 toolchain took around \SI{35}{\second}, while changing only \approachshort{} parameters required around \SI{25}{\second}.
+These results show that \approachshort{} can be easily adapted and altered by network administrators once in place, and illustrates .
 

### 2021-06-01 paper/paper.tex
Presentation pass of results section

@@
 \newcommand{\indfw}{\Indfw}
+\newcommand{\inring}{\textsc{In}}
+\newcommand{\outring}{\textsc{Out}}
 
@@
 For \Coopfw{} for instance, this creates \num{129} work items across \num{31} workers.
-Although their more successful \emph{SPF} agent design uses 3 actions, we choose a larger action set to investigate the performance of more complex agents.
+Although the model application's more successful \emph{SPF} agent design uses 3 actions, we choose a larger action set to investigate the performance of more complex agents.
 
@@
 This equalled the amount of physical cores on each device, while moving beyond this (even below the number of hyper-threads) would hamper tail latencies by an order of magnitude.
-To make the comparison fairer in the context of many-core CPU environments, we include per-core throughput.
+To make the comparison fair in the context of many-core CPU environments, we include per-core throughput.
 \Indfw{} achieves \SI{2.82}{$\times$} higher offline throughput than commodity \emph{Collector} hardware in spite of the NFP-6480 having a considerably slower clock speed (\SI{0.29}{$\times$}).
 When compounded with the abundance of such weaker chips, in-NIC RL is able to deliver much higher throughput.
-As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} a dedicated collector machine, due to the locking requirements mandating coordinated division of work.
+As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} that of a dedicated collector machine, due to the locking requirements mandating coordinated division of work.
 
@@
 This requires \num{8} workers to offer greater online throughput than single-threaded in-NIC RL.
-Sharper performance increases occur when new physical core is added (\numrange{7}{8}) or the scheduler works around a bottleneck (\numrange{13}{14}).\label{fig:vary-core}}
+Sharper performance increases occur when a new physical core is added (\numrange{7}{8}) or the scheduler works around a bottleneck (\numrange{13}{14}).\label{fig:vary-core}}
 \end{figure}
@@
 Most of the I/O comes after this step, causing ALU use to dominate.
-Regardless, the reduction in memory needed to store a policy remain.
-This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) \Coopfw{} agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the algorithm.
+This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) \Coopfw{} agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the algorithm in the online case.
+Regardless, the \SIrange{2}{4}{$\times$} reduction in memory needed to store a policy holds.
 
@@
 	}
-	\caption{Action/update compute times in a \SI{32}{\bit} \Coopfw{} agent under variable work schedulers.\label{fig:work-alloc-32}}
+	\caption{Action/update compute times in a \SI{32}{\bit} \Coopfw{} agent under different work schedulers.\label{fig:work-alloc-32}}
 \end{figure}
@@
 	}
-	\caption{Throughput per added worker in a \Coopfw{} agent. \label{fig:tput-per-core}}
+	\caption{Throughput per added worker in a \Coopfw{} agent.\label{fig:tput-per-core}}
 \end{figure}
@@
 We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
-While there is still have a \SI{4.2}{$\times$} factor of task oversubscription at maximum worker count, it is clear that latencies are bound below by the length of the longest task.
+While there is still a \SI{4.2}{$\times$} factor of task oversubscription at maximum worker count, it is clear that latencies are bound below by the length of the longest task.
 
@@
 ?? In the name of fairness, assume that rule installation uses same mechanism we recommend, but show how bad it can get? I.e. with rule installation costs etc.
+
+?? Cziva~\parencite{DBLP:phd/ethos/Cziva18} discusses vNF times.
 
@@
 This slight general degradation can be explained by increased pressure on the NFP's \emph{command push-pull} (CPP) bus, which is responsible for handling cross-island accesses to memory (particularly IMEM/EMEM) and other resources.
-\approachshort{} places load on the CPP bus through use of its IN/OUT rings for external communication and through last-tier policy accesses.
+\approachshort{} places load on the CPP bus through use of its \inring{}/\outring{} rings for external communication and through last-tier policy accesses.
 This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---the Netronome P4 toolchain segments packets, storing metadata (e.g., MAC prepend data) and the first bytes of a packet in a $\SI{256}{\byte}$ CTM block and parking the bodies of processed packets in EMEM.
@@
 The anomalous result is \SI{128}{\byte} packets at \num{3000} RL action/update computations per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase, shown in \cref{fig:dataplane-example}.
+This is observed through a shift of some packet latencies from the mode towards the tail, but no other changes in the distribution.
 In the above context, we believe that the inbound request rate is weakly synchronised with inbound packets, causing a higher level of burstiness around accesses to the CPP bus.
-We expect that dedicated hardware/FPGA designs can avoid this problem by having dedicated In/Out access mechanisms for an \approachshort{} agent.
+We expect that dedicated hardware/FPGA designs can avoid this problem by having dedicated \inring{}/\outring{} access mechanisms for an \approachshort{} agent.
 
@@
 	\caption{Effects on tail latency of cross-traffic caused by different loads of off-path RL compute.
-Statistically significant increases in population latency are concentrated on smaller packet sizes, and are typically sub-\SI{77}{\nano\second}.\label{fig:dataplane-coop}}
+Statistically significant increases in population latency are concentrated on smaller packet sizes, and are typically sub-\SI{78}{\nano\second}.\label{fig:dataplane-coop}}
 \end{figure}
@@
 \Cref{tab:resources} demonstrates how \approachshort{} consumes shared and local memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
-As one program is installed per ME, these represent the minimum and maximum resource use we can command on a single island (i.e., without replacing P4 workers).
-We observe negligible costs on shared bulk storage ($\sim$\SI{4}{\mebi\byte}), including storage space for a hash-table of \num{4096} state-action pairs and \num{16} reward values.
+As one program is installed per ME, these results represent the minimum and maximum resource use we can command on a single island (i.e., without replacing P4 workers).
+We observe negligible costs on shared bulk EMEM storage ($\sim$\SI{4}{\mebi\byte}), including storage space for a hash-table of \num{4096} state-action pairs and \num{16} reward values.
 The most significant costs arise due to policy data (\SI{405}{\kibi\byte} capacity shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
-Thread local storage in CLS for compute/register spilling scales with the number of MEs required, particularly when caching  and with \Coopfw{} this is far below the ``fair share'' (allowing co-existence with other asynchronous dataplane programs).
+Thread local storage in CLS for compute/register spilling scales with the number of MEs required (requiring \SIlist{13.77;15.49}{\percent} per-ME for \Indfw{}/\Coopfw) due to precaching and the space needed to hold tile lists.
+Due to the initial policy cost this falls below the fair share at 3 MEs, but always allows for co-existence with other asynchronous dataplane programs.
 
@@
 In the event that appropriate firmwares are not pre-compiled, we found that compiling and linking both \approachshort{} and the P4 toolchain took around \SI{35}{\second}, while changing only \approachshort{} parameters required around \SI{25}{\second}.
-These results show that \approachshort{} can be easily adapted and altered by network administrators once in place, and illustrates .
+These results show that \approachshort{} can be easily adapted and altered by network administrators once in place, and illustrates an advantage of SoC-based SmartNICs.
 

### 2021-06-01 paper/paper.tex
An algorithm
Mother of god.

@@
 	\caption{ParSa---\emph{Par}allel \emph{Sa}rsa\label{alg:parsa}}
-	\SetKwProg{generate}{Function \emph{generate}}{}{end}
-
-	Map store=new Map(obj, queue)\;
-	\generate{Object pivot}{
-		\ForAll{child $c$ in pivot}{
-			\If{ $c$'s FieldContext is not set and $c$ is fusible}{
-				generate($c$)\;
-			}
-		}
-		build pivot's fieldContext $fc$\;
-		EmitClassName\;
-		EmitFields($fc$)\;
-		EmitMethods($fc$)\;
-	}
+	\SetKw{Let}{let}
+	\SetKw{Enum}{enum}
+	\SetKw{In}{in}
+	\SetKw{Await}{await}
+	\SetKw{Const}{const}
+	\SetKwProg{parsa}{Function \emph{ParSa}}{}{end}
+	\SetKwProg{control}{Function \emph{Ctl}}{}{end}
+	\SetKwProg{minion}{Function \emph{Minion}}{}{end}
+	\SetKwProg{tilecode}{Function \emph{TileCode}}{}{end}
+
+	\tcc{Given message passing mechanisms \emph{scatter} and \emph{recv}, quantised arithmetic functions $Q_\mathit{mul}$ and TileCode, and omitting schedule/config/precache updates.}
+
+	\Enum Par \{ Act(\emph{state}), Upd(\emph{delta, action, state}) \}\;
+	\Const \emph{cfg, policy} = /* ... */\;
+	\Let \emph{values}: [AtomicI32; cfg.n\_actions] = \{0\}\;
+	\Let \emph{acks}: AtomicI32 = 0\;
+	\parsa{id, schedule}{
+		\uIf{id==0}{
+			\ForAll{state\_pkt \In \inring{}}{
+				Ctl(state\_pkt)\;
+			}
+		}
+		\Else{
+			\While{true}{Minion(schedule[id - 1], recv())\;}
+		}
+	}
+
+	\control{state}{
+		\emph{values, acks} = \{0\}, scatter(Par::Act(\emph{state}))\;
+		acquire slot for \outring, copy \emph{state} into slot\;
+		\Await acks == \emph{cfg.n\_minions}\;
+		\Let \emph{action} = argmax(\emph{values})\;
+		write \emph{action} into \outring{} slot, enqueue\;
+		\If{cfg.online}{
+			\Let \emph{((l\_state, l\_act, l\_val), found\_s)} = \emph{cfg.lookup\_state\_from\_key(state)}\;
+			\Let \emph{(reward, found\_r)} = \emph{cfg.lookup\_reward\_from\_key(state)}\;
+			\If{found\_s \&\& found\_r}{
+				\Let $\delta_t$ = $\mathit{reward} + Q_\mathit{mul}$(\emph{cfg}.$\gamma$, \emph{values}[\emph{action}]) $-$ \emph{l\_val}\;
+				$\delta_t$ = $Q_\mathit{mul}$(\emph{cfg}.$\alpha$, $\delta_t$)\;
+				\emph{acks} = 0, scatter(Par::Upd($\delta_t$, \emph{action}, \emph{state}))\;
+				\Await acks == \emph{cfg.n\_minions}\;
+			}
+			\emph{cfg}.store\_state(\emph{state}, \emph{action}, \emph{values}[\emph{action}])\;
+		}
+	}
+
+	\minion{tasks, msg}{
+		\Switch{msg}{
+		\uCase{Par::Act(\emph{s})}{
+			\ForAll{task \In tasks}{
+				\Let \emph{hit} = TileCode(\emph{s}, \emph{task})\;
+				\For{i \In [0..cfg.n\_actions)}{
+					\emph{values}[\emph{i}].atomic\_add(\emph{policy}[\emph{hit}][\emph{a}])\;
+				}
+			}
+		}
+		\uCase{Par::Upd($\delta$, \emph{a}, \emph{s})}{
+			\ForAll{task \In tasks}{
+				\Let \emph{hit} = TileCode(\emph{s}, \emph{task})\;
+				\emph{policy}[\emph{hit}][\emph{a}] += $\delta$\;
+			}
+		}
+		}
+		acks.atomic\_add(1)\;
+	}
+
+%	\tilecode{state, task}{
+%		\Let \emph{hit} = \emph{cfg}.get\_task(task)\;
+%	}
 \end{algorithm}
+
+?? Where to mention that algo 1 omits epsilon greedy etc?
 
@@
 ?? Future directions for community: want to see more general-purpose async compute on-chip! Larger form-factors!
+
+?? Future directions; transfer learning between classical methods into faster offline-only?
 

### 2021-06-02 paper/paper.tex
Related Work.,..

@@
 In-NIC execution allows clever use of resources to bring 99.99th percentile action compute latency  from \SI{725}{\micro\second} to \SI{44}{\micro\second} (1.5 orders of magnitude!) in real-world policies for DDoS detection/prevention control.
+
+We present \emph{\approachshort{}}---\approach{}---to the best of our knowledge the first online RL in the dataplane...
 
@@
 We note that our parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is made possible only by the use of quantisation to integers, as atomic operations on floating point numbers typically do not exist outside of graphics- or tensor-processing hardware.
+ParSa is described here at a high level, as an algorithm suited for \emph{any multiprocessor environment}.
+We detail our communication primitives in \cref{sec:intra-agent-communication}, and our work scheduling strategy in \cref{sec:work-allocation}, eliding the details of tile-coding as they are well-understood.
 
@@
 
-\subsection{Intra-Agent Communication}
+\subsection{Intra-Agent Communication}\label{sec:intra-agent-communication}
 Even with embarrassingly parallel problems, optimising for latency requires meticulous care in how work is passed out and aggregated.
@@
 
-\subsection{Work Allocation}
+\subsection{Work Allocation}\label{sec:work-allocation}
 Due to the lack of dynamic memory allocation and to simplify action-value lookups, policies cannot be effectively stored sparsely.
@@
 We use a simple first-fit work placement algorithm run on the NFP, placing the largest work item into the least loaded context of the least loaded ME.
+Each work item is a separate \emph{tiling} over a number of dimensions, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
 The cost of any individual work item (according to its dimension count and memory location) was empirically measured offline, and we weight the total cost per ME to account for the number of minion threads available.
@@
 \section{Related Work}
-
-?? NOTES: try to strengthen emphasis on ``other works can't do online learning in the dataplane''. This is absolutely, \SI{100}{\percent} the killer feature.
-
-?? A switch/NIC architecture like \textcite{DBLP:conf/hotnets/StephensAS18} would remove the need for hard-coded island-island relationships.
-
-?? Event-driven model could better support many timer-driven RL use-cases \textcite{DBLP:conf/hotnets/IbanezABM19}.
-
-?? Taurus. They can't train online though!~\parencite{DBLP:journals/corr/abs-2002-08987}
-
-?? Do switches~\parencite{DBLP:conf/hotnets/XiongZ19}.
-?? Train offline, classical methods targetting all match-action P4 systems (no externs etc).
-
-\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}
-
-?? N3IC run NNs on NIC~\parencite{DBLP:journals/corr/abs-2009-02353}, allow packet tagging to occur based on pre-trained BNN~\parencite{DBLP:conf/nips/HubaraCSEB16}.
-?? both NFP (\SI{45}{\micro\second}) and NetFPGA (\SI{0.3}{\micro\second}).
-?? We handle larger inputs (versus their \SI{256}{\bit}), offer online training.
-?? Use our 5-dim as a point of comparison?
-
-?? Or as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
-?? Binary NNs ``more suitable for network hardware''~\parencite{DBLP:journals/corr/MiyashitaLM16}?
-
-?? Can we collect some papers on per-packet ML/RL? Per-packet classification using NNs to build decision trees~\parencite{DBLP:conf/sigcomm/LiangZJS19}
-
-?? RL for congestion control. ~\parencite{DBLP:journals/corr/abs-1910-04054}
-
-?? Arbitrary training enhanced by in-network? General NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18}, RL-specfic~\parencite{DBLP:conf/isca/LiLYCSH19}.
-
-?? Mechanisms to support full in-switch DDoS detection~\cite{tnms-ddos-victim-ident}.
-
-??
-Simialr HW has been used to get NN \emph{inference} \textcite{langlet-ml-netronome}---although \emph{actual NN latency} in his Fig 4.4 (IPG is inter-packet gap: his is not truly async but achieves perf by lying on the packet path.
-Note that 0 IPG murders his latencies).
+\fakepara{In-network ML}
+Taurus~\parencite{DBLP:journals/corr/abs-2002-08987} proposes that efficient line-rate, low-latency per-packet inference can be achieved by placing a run-time configurable grid topology of map-reduce processing units into the packet processing pipeline (implementing CNNs, LSTMs, and some classical approaches such as SVMs).
+Their CGRA-based implementation is able to be repurposed at runtime and achieves sub-\si{\micro\second} extra latency, though online model training is not considered.
+
+\emph{IIsy}~\parencite{DBLP:conf/hotnets/XiongZ19} shows how \emph{classical ML inference} techniques (SVMs, Na\"{i}ve Bayes, Decision Trees, K-Means) can be converted into match-action table structures which are compatible with \emph{any} P4 deployment.
+They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA devices (at line rate in most cases), but can only investigate the use of pre-trained models.
+
+A recent line of research in the community has been to investigate the use of \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate, low latency packet classification.
+\emph{BaNaNa SPLIT} shows the initial steps of this trend investigating their use as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} in an effort to reduce latency imposed by batching and bulk transfer to the GPU.
+However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}, with the hypothesis put forward that networking hardware would be a better fit for some parts of the inference pipeline.
+This line of work has moved onto full, in-network packet tagging and classification by pre-trained BNNs with \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
+N3IC achieves packet inference in \SI{45}{\micro\second} on the NFP, and \SI{0.3}{\micro\second} on NetFPGA for \SI{256}{\bit} inputs.
+Comparatively, \approachshort{}-\Coopfw{} can process an identically-sized input (8-dim) in a median \SI{13.83}{\micro\second}.
+Our work handles larger inputs (\SI{640}{\bit}) at lower latencies of \SI{34}{\micro\second}, and offers online learning.
+We expect that a NetFPGA implementation of \approachshort{} would enjoy a similar factor of speedup.
+No works investigate the runtime training of BNN function approximators in-NIC.
+
+\textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantised weights and inputs on NFP hardware, using \emph{in-path} compute rather than our asynchronous model.
+Inference latency on small networks (3 layers, \num{30} layers) can be as high as \SI{500}{\micro\second} on line rate traffic, stressing the value of the asynchronous compute and quantisation we adopt.
+
+We note that none of the other approaches listed here (or that we have seen) tackle the issue of \emph{online learning and control} in-network---we believe \approachshort{} has broken new ground in this regard.
+
+\fakepara{In-network ML \emph{acceleration}}
+Optimisation of distributed neural network training is an area where in-network compute has been key in the case of general NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18} and RL-specific procedures~\parencite{DBLP:conf/isca/LiLYCSH19} using NetFPGAs to introduce floating point adders to process dedicated packet classes.
+Distributed training is a partition-aggregate workload, where gradient vectors are sent back to a controller to update the `true' model---in-NIC processing allows these to be aggregated \emph{in-network}, overcoming incast behaviour and host bottlenecks.
+
+\fakepara{RL for network control}
+\Textcite{DBLP:conf/sigcomm/LiangZJS19} use deep RL techniques on a host machine to train an agent which can build efficient decision tree packet classifiers.
+Their approach, \emph{NeuroCuts}, is able to produce time/space efficient classifiers for use in constrained environments (e.g., network hardware) beyond existing heuristic solutions.
+
+Deep RL techniques have been used for QUIC congestion control optimisation~\parencite{DBLP:journals/corr/abs-1910-04054}.
+A key facet of this work is the need for asynchronous RL in networks, where pauses for DNN-based inference can significantly harm throughput.
+
+\fakepara{PDP design for asynchronous compute}
+\Textcite{DBLP:conf/hotnets/StephensAS18} have proposed \emph{PANIC}, which places a routing fabric between distinct packet/data processing elements \emph{in a SmartNIC}.
+Such designs would more generally enable asynchronous and novel compute in SmartNICs and programmable switches, for instance offering consistent and easy to use communication between workers versus hard-coded ME relationships.
+
+Event-driven versions of the P4 pipeline have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}, 
+Timer events and device state changes would empower in-network RL use-cases, signalling timesteps for RL agents or a new, effective, fine-grained sources of input state.
 
@@
 
-?? Future directions for community: want to see more general-purpose async compute on-chip! Larger form-factors!
+?? Future directions for community: want to see more general-purpose async compute on-chip! Larger form-factors! More asynchrony!
 

### 2021-06-03 ngni-talk/talk.tex
NGNI talk draft.

--- 2020-04-01 motivation-talk/talk.tex
@@
 \usepackage{bm}
+\usepackage{mathtools}
 
@@
 
+\newcommand{\approach}{On Path Learning}
+\newcommand{\approachshort}{OPaL}
+\newcommand{\Coopfw}{\emph{CoOp}}
+\newcommand{\coopfw}{\Coopfw}
+\newcommand{\Indfw}{\emph{Ind}}
+\newcommand{\indfw}{\Indfw}
+\newcommand{\inring}{\textsc{In}}
+\newcommand{\outring}{\textsc{Out}}
+
+\makeatletter\let\expandableinput\@@input\makeatother
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+
 %-------------------------------------%
@@
 
-\title{Reinforcement Learning in Future Networks}
-\author{\vspace{-1em}\textbf{Kyle A. Simpson}\\
+\title{Revisiting the Classics: Online RL in the Programmable Dataplane}
+\author{\vspace{-1em}\textbf{Kyle A. Simpson}, Dimitrios P. Pezaros\\
 	\faEnvelopeO{} \href{mailto:k.simpson.1@research.gla.ac.uk}{\nolinkurl{k.simpson.1@research.gla.ac.uk}}\\
@@
 \institute{University of Glasgow}
-\date{\nth{27} March, 2020}
+\date{\nth{4} June, 2021}
 
 \begin{document}
+% title fun, including Org logos....
 \begin{frame}
 	\maketitle
+	\begin{tikzpicture}[overlay, remember picture]
+%		\node[above right=0.8cm and 0.9cm of current page.south west] (esnet-logo) {\includegraphics[width=2.75cm]{netlab-trim}};
+%		\node[right=1cm of esnet-logo] {\adjincludegraphics[height=2cm,trim={0 {.4\height} 0 {.05\height}},clip]{uofg}};
+		\node[above right=-0.1cm and 0.8cm of current page.south west] (uofg-logo) {\adjincludegraphics[height=2cm,trim={0 {.4\height} 0 {.05\height}},clip]{uofg}};
+		\node[right=0.5cm of uofg-logo] {\includegraphics[width=2.75cm]{netlab-trim}};
+	\end{tikzpicture}
 \end{frame}
@@
 
-\begin{frame}{DDoS Prevention: VNF + SDN  architecture}
+\begin{frame}{Recent Programmable Trends in Data-Driven Networks}
 	\begin{columns}
-%		\begin{column}{0.3\linewidth}
-%			\begin{itemize}
-%				\item Monitor flow statistics at ingress/egress points.
-%				\item Use load measurements from flow paths as global state.
-%				\item OUTPUT: throttle individual flows.
-%			\end{itemize}
-%		\end{column}
-		\begin{column}{1.0\linewidth}
+		\begin{column}{0.6\linewidth}
+			\begin{itemize}
+				\item ML acceleration, line-rate packet classification.
+				\item How? Train model off-NIC, convert to \alert{binary neural network}\footnotemark, or \alert{decision tree}\footnotemark.
+				\item Limits? No online training, cost of backprop algo (expensive!), vast data needs.
+				\item \alert{What if we need online learning?}
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.4\linewidth}
+			\centering
+			\resizebox{0.6\linewidth}{!}{
+				\includegraphics{images/smartnic.png}
+			}
+
+			\resizebox{0.5\linewidth}{!}{
+				\includegraphics{images/nn}
+			}
+		\end{column}
+	\end{columns}
+\setcounter{footnote}{1}
+\footcitetext{DBLP:journals/corr/abs-2009-02353}
+\setcounter{footnote}{2}
+\footcitetext{DBLP:conf/hotnets/XiongZ19}
+\end{frame}
+
+\begin{frame}{Timing: Why not offload to the controller?}
+	\begin{columns}
+		\begin{column}{0.5\linewidth}
+			For SmartNICs, the attached host \emph{is the (closest) controller}.
+
+			\begin{itemize}
+				\item PCIe access times $\mathcal{O}(\si{\micro\second})$\footnotemark
+				\item Crossing VMs/vNFS has $\sim$\SI{10}{$\times$} higher cost\footnotemark
+				\item MATs recompiled $\mathcal{O}(\si{\second})$. Many rules $\implies$ batching.
+%				\item Thrift serde time $\mathcal{O}(\si{\milli\second})$.
+			\end{itemize}
+
+			%	?? https://github.com/rpclib/benchmarks/blob/master/plots/get_blob_real_log.svg
+
+			\alert{Meanwhile core-to-core on NFP around \SI{140}{\nano\second} @ \SI{1.2}{\giga\hertz}.}
+		\end{column}
+		\begin{column}{0.5\linewidth}
 			\begin{figure}
-				\centering
-				\resizebox{0.7\linewidth}{!}{\begin{tikzpicture}[scale=1, every node/.style={scale=0.7}]
-					\node(remote){Remote};
-
-					%%%
-
-					\node[below=0.05 of remote](swpos){};
-					\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
-					\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
-					\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
-					\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
-					\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Load\\Collector\end{varwidth}};
-
-					%
-
-					\node[right=2 of swpos](epos){};
-					\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
-					\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
-					\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
-					\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
-					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.65,0)$) {\begin{varwidth}{1.5 cm}\scriptsize \centering Load\\Collector\end{varwidth}};
-					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\scriptsize \centering Estimator\\$g(\cdot)$\end{varwidth}};
-
-					%
-
-					\node[right=3.5 of epos](apos){};
-					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
-					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
-					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
-					\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent vNF};
-					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
-
-					%
-
-					\node[below=2.3 of remote.west](linestart){};
-					\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
-					\draw [dashed] (linestart) -- (lineend);
-
-					%%%
-
-					\node[below=2.4 of remote](local){Local};
-
-					%%%
-
-					\node[below=0.05 of local](aswpos){};
-					\node[fill=white!80!uofgthistle, draw=black, minimum height=3.5cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
-					\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
-					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering OpenFlow\\Tables\end{varwidth}};
-					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\scriptsize \centering Stats\\Collector\end{varwidth}};
-					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\scriptsize \centering Load\\Collector\end{varwidth}};
-
-					%
-
-					\node[right=3 of aswpos](avfpos){};
-					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
-					\node[below right, inner sep=2pt] at (avf.north west) {\small Agent vNF};
-					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
-					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\scriptsize \centering Flowstate\\Database\end{varwidth}};
-					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\scriptsize \centering TRS\\Scheduler\end{varwidth}};
-					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
-					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};
-
-					%%%
-
-					\tikzset{>=stealth}
-
-					\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
-					\draw[thick, <->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};
-
-					\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
-
-					\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
-					\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};
-
-					\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};
-
-					\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
-					\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};
-
-					\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
-					\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
-					\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
-					\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
-					\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};
-
-					\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
-					\node[right=0.3] at (avfsa.east) {\tiny Experience};
-					\end{tikzpicture}}
-				\caption{
-					System architecture for RL-driven DDoS defence.
-					\label{fig:sys-arch}
-				}
-				\vspace{-1em}
+				\includegraphics[width=0.9\linewidth]{../plots/build/rte-timer/rte-times-logx}
+				\caption{Netronome rule installation cost \\(1 table, \numrange{1}{65536} rules).}
 			\end{figure}
@@
 	\end{columns}
+	\setcounter{footnote}{3}
+	\footcitetext{DBLP:conf/sigcomm/NeugebauerAZAL018}
+	\setcounter{footnote}{4}
+	\footcitetext{DBLP:journals/cm/CzivaP17}
 \end{frame}
 
-\begin{frame}{Programmable data-planes: An Architecture for RL}
+\begin{frame}{How do we bring online, in-NIC RL?}
 	\begin{columns}
@@
 			\begin{figure}
-			\resizebox{1.2\linewidth}{!}{\begin{tikzpicture}
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=1cm, minimum width=2.9cm, below right= 0.1 of sw2.north west](p4-master){};
-				\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Master Island};
-				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (p4-rules) at (p4-master.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};
-
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2cm, below = 0.1 of p4-master](sw1){};
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
-				\node[below right, inner sep=2pt] at (switch.north west) {\small P4 Island};
-				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};
-
-				\node[fill=white!80!uofgcobalt, draw=black, minimum height=2cm, minimum width=2.9cm, below= 0.1 of switch](others){};
-				\node[below right, inner sep=2pt] at (others.north west) {\small Other Islands};
-				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-etc) at (others.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering \textellipsis \end{varwidth}};
-				\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (others-int) at (others-etc.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering In-band Network Telemetry\end{varwidth}};
-
-				\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (3, -2)$){};
-				\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};
-
-				\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
-				\node[below right, inner sep=2pt] at (wq-out.north west) {\small CLS};
-				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Work Queue\end{varwidth}};
-
-				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (loop) at ($(wq-out.east) + (1.5,-0.5)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Control ME\end{varwidth}};
-
-				\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-cls-out) at ($(loop.south) + (0,-1.5)$){};
-				\node[below right, inner sep=2pt] at (policy-cls-out.north west) {\small CLS};
-				\node[above right, inner sep=2pt] at (policy-cls-out.north west) {\small Policy};
-				\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-ctm-out) at ($(policy-cls-out.south) + (0,-0.25)$){};
-				\node[below right, inner sep=2pt] at (policy-ctm-out.north west) {\small CTM};
-				\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-emem-out) at ($(policy-ctm-out.south) + (0,-0.25)$){};
-				\node[below right, inner sep=2pt] at (policy-emem-out.north west) {\small EMEM};
-
-				\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t1) at (policy-cls-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T1\end{varwidth}};
-				\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t2) at (policy-ctm-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T2\end{varwidth}};
-				\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t3) at (policy-emem-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T3\end{varwidth}};
-
-				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-1) at ($(wq-out.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};
-
-				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-2) at ($(po-me-1.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};
-
-				\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-3) at ($(po-me-2.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};
-
-				\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2]  node[pos=0.5, rotate=90, anchor=center, above] {Rules}(sw1);
-				\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (sw2);
-				\draw[->] (p4-rules.west)  to [out=180,in=180,looseness=2] (switch);
-
-				\draw[->] (oswlc.east)  to node[midway, above, sloped] {\tiny Control, Requests} (wq-in.west);
-				\draw[->] (others-int.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
-				\draw[<->, dotted] (wq-in.east)  to (loop.west);
-				\draw[->, dotted] (loop.south) to (policy-cls-out);
-
-				\draw[<->, dotted] (po-me-1.east) to (policy-cls-out.west);
-				\draw[<->, dotted] (po-me-2.east) to (policy-ctm-out.west);
-				\draw[<->, dotted] (po-me-3.east) to (policy-emem-out.west);
-
-				\draw[->] (loop) to node[midway, above, sloped] {\tiny NN} (po-me-1);
-				\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NN} (po-me-2);
-				\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NN} (po-me-3);
-				\draw[->] (po-me-3) to [out=-45,in=0,looseness=2] node[midway, above, sloped] {\tiny NN} (loop);
-			\end{tikzpicture}}
-			\caption{Architecture for generalised RL agent on Netronome hardware.}
+				\includegraphics[width=\linewidth]{../paper/figures/arch-with-p4}
 			\end{figure}
@@
 			\begin{itemize}
-			\item Make decisions \alert{where they must be applied}.
-			\item ``Bump-in-the-wire'' use of Netronome SmartNIC.
-			\item P4 parsing/forwarding/filtering, offload to microcode for full programmability.
-			\begin{itemize}
-				\item P4 with offload common.
-			\end{itemize}
-				\item Solution: \alert{make use of on-device parallelism, externs to implement fast, adaptive data-driven applications}.
+				\item Classical RL built on tile-coding---\alert{online}.
+				\item \alert{Fixed-point arithmetic}.
+				\item Async wrt. datapath.
+				\item Dynamic selection of last reward, trace info.
+				\item \alert{Runtime configurable} (policy, size, application) from data/control-plane. \alert{Task independent.}
 			\end{itemize}
@@
 
+\section{Background and Design}
+
+\begin{frame}{Single-step (classical) RL---Sarsa}
+	A simple explanation:\pause
+
+	\begin{subequations}
+		\begin{gather}
+			\delta_t = \overbrace{R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}}}^{\mathclap{\substack{\text{New target value:}\\ \text{Reward} + \text{some of the next action's value}}}} - \underbrace{\acval{S_t}{A_t}{\wvec{t}}}_{\mathclap{\text{Current value estimate}}},\\\pause
+			\bm{w}_{t+1} = \bm{w}_{t} + \overbrace{\alpha \delta_t}^{\mathclap{\text{Move a little bit of } \delta_t \text{ along...}}} \underbrace{\nabla{\acval{S_t}{A_t}{\wvec{t}}}}_{\mathclap{\text{...the policy's gradient}}},\pause
+		\end{gather}%
+	\end{subequations}
+
+	Design implications?
+\end{frame}
+
+\begin{frame}{Tile Coding}
+\centering
+\resizebox{0.7\linewidth}{!}{
+	\begin{tikzpicture}
+		\node at (0.35,0.4){$s = \begin{pmatrix}
+				0.7 \\
+				0.3
+			\end{pmatrix}$};
+		\node at (0,-1.1) {$\bm{x}(s,\cdot) = \begin{Bmatrix}
+				T_{1,9}, \\
+				T_{2,5}, \\
+				T_{\mathit{bias}}
+			\end{Bmatrix}$};
+
+		\node at (2.5,-0.5) {
+			\begin{tikzpicture}
+				\draw[step=0.5cm,color=uofgcobalt,opacity=0.7,shift={(0,0)},label=above:{Tiling 0}] (-0.5,-0.5) grid (1,1);
+				\fill[uofgcobalt,opacity=0.5] (0.5,-0.5) rectangle (1,0);
+				\node[color=uofgcobalt] at (0,1.2) {\footnotesize Tiling 1};
+
+				\draw[step=0.5cm,color=uofgpumpkin,opacity=0.9,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
+				\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
+				\node[color=uofgpumpkin!50!uofgrust] at (0.25,-0.95) {\footnotesize Tiling 2};
+
+				\node[circle, black, draw,
+				fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
+				%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];
+
+				\draw[->] (-0.25,-0.5)--(-0.25,0.85);
+				\draw[->] (-0.25,-0.5)--(1.1,-0.5);
+
+				\node at (1,-0.7) {\footnotesize 1};
+				\node at (-0.4,0.75) {\footnotesize 1};
+				\node at (-0.35,-0.6) {\footnotesize 0};
+			\end{tikzpicture}
+		};
+
+	\end{tikzpicture}
+}
+\end{frame}
+
+\begin{frame}{Tile Coding (ii)}
+	\begin{columns}
+		\begin{column}{0.6\linewidth}
+			\centering
+			\resizebox{0.8\linewidth}{!}{
+				\begin{tikzpicture}
+					\node at (0.35,0.4){$s = \begin{pmatrix}
+							0.7 \\
+							0.3
+						\end{pmatrix}$};
+					\node at (0,-1.1) {$\bm{x}(s,\cdot) = \begin{Bmatrix}
+							T_{1,9}, \\
+							T_{2,5}, \\
+							T_{\mathit{bias}}
+						\end{Bmatrix}$};
+
+					\node at (2.5,-0.5) {
+						\begin{tikzpicture}
+							\draw[step=0.5cm,color=uofgcobalt,opacity=0.7,shift={(0,0)},label=above:{Tiling 0}] (-0.5,-0.5) grid (1,1);
+							\fill[uofgcobalt,opacity=0.5] (0.5,-0.5) rectangle (1,0);
+							\node[color=uofgcobalt] at (0,1.2) {\footnotesize Tiling 1};
+
+							\draw[step=0.5cm,color=uofgpumpkin,opacity=0.9,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
+							\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
+							\node[color=uofgpumpkin!50!uofgrust] at (0.25,-0.95) {\footnotesize Tiling 2};
+
+							\node[circle, black, draw,
+							fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
+							%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];
+
+							\draw[->] (-0.25,-0.5)--(-0.25,0.85);
+							\draw[->] (-0.25,-0.5)--(1.1,-0.5);
+
+							\node at (1,-0.7) {\footnotesize 1};
+							\node at (-0.4,0.75) {\footnotesize 1};
+							\node at (-0.35,-0.6) {\footnotesize 0};
+						\end{tikzpicture}
+					};
+
+				\end{tikzpicture}
+			}
+		\end{column}
+		\begin{column}{0.4\linewidth}
+			\begin{itemize}[<+->]
+				\item Operations? $+, -, \times, \div$
+				\begin{itemize}[<+->]
+					\item PoT tile width? $\div$ replaced w/ shift.
+				\end{itemize}
+				\item Tiling set: identical dimensions, different shifts
+				\item Gradient for RL?
+				\begin{itemize}[<+->]
+					\item List of hit tiles.
+				\end{itemize}
+				\item Can be more complex...
+			\end{itemize}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Tile Coding: Parallelism}
+	\centering
+	\resizebox{0.7\linewidth}{!}{
+		\begin{tikzpicture}
+			\node at (0,0) {
+				\begin{tikzpicture}
+					\draw[step=0.5cm,color=uofgcobalt,opacity=0.7,shift={(0,0)},label=above:{Tiling 0}] (-0.5,-0.5) grid (1,1);
+					\fill[uofgcobalt,opacity=0.5] (0.5,-0.5) rectangle (1,0);
+					\node[color=uofgcobalt] (t1g) at (0,1.2) {\footnotesize Tiling 1};
+
+					\draw[step=0.5cm,color=uofgpumpkin,opacity=0.9,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
+					\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
+					\node[color=uofgpumpkin!50!uofgrust] (t2g) at (0.25,-0.95) {\footnotesize Tiling 2};
+
+					\node[circle, black, draw,
+					fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
+					%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];
+
+					\draw[->] (-0.25,-0.5)--(-0.25,0.85);
+					\draw[->] (-0.25,-0.5)--(1.1,-0.5);
+
+					\node at (1,-0.7) {\footnotesize 1};
+					\node at (-0.4,0.75) {\footnotesize 1};
+					\node at (-0.35,-0.6) {\footnotesize 0};
+				\end{tikzpicture}
+			};
+
+		\node (policy-head) at (2.5,1.2) {Policy};
+		\draw[color=uofgcobalt,opacity=0.7] (2,0) rectangle ++(2,1) node[pos=.5] (t1p) {Tiling 1};
+		\fill[uofgcobalt,opacity=0.25] (3.33,0) rectangle ++(0.67,0.33);
+
+		\draw[color=uofgpumpkin,opacity=0.9] (2,-1) rectangle ++(2,1) node[pos=.5] (t2p) {Tiling 2};
+		\fill[uofgpumpkin,opacity=0.25] (2.67,-0.67) rectangle ++(0.67,0.33);
+
+		\draw (2,-2) rectangle ++(2,1) node[pos=.5] {$\cdots$};
+
+		\draw [->,color=uofgcobalt, bend left] (t1g) to (t1p.west);
+		\draw [->,color=uofgpumpkin, bend right] (t2g) to (t2p.west);
+
+		\end{tikzpicture}
+	}
+\end{frame}
+
+\begin{frame}{Tile Coding: Parallelism (ii)}
+	\centering
+	\resizebox{0.6\linewidth}{!}{
+		\begin{tikzpicture}
+			\node at (0,0) {
+				\begin{tikzpicture}
+					\draw[step=0.5cm,color=uofgcobalt,opacity=0.7,shift={(0,0)},label=above:{Tiling 0}] (-0.5,-0.5) grid (1,1);
+					\fill[uofgcobalt,opacity=0.5] (0.5,-0.5) rectangle (1,0);
+					\node[color=uofgcobalt] (t1g) at (0,1.2) {\footnotesize Tiling 1};
+
+					\draw[step=0.5cm,color=uofgpumpkin,opacity=0.9,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
+					\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
+					\node[color=uofgpumpkin!50!uofgrust] (t2g) at (0.25,-0.95) {\footnotesize Tiling 2};
+
+					\node[circle, black, draw,
+					fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
+					%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];
+
+					\draw[->] (-0.25,-0.5)--(-0.25,0.85);
+					\draw[->] (-0.25,-0.5)--(1.1,-0.5);
+
+					\node at (1,-0.7) {\footnotesize 1};
+					\node at (-0.4,0.75) {\footnotesize 1};
+					\node at (-0.35,-0.6) {\footnotesize 0};
+				\end{tikzpicture}
+			};
+
+			\node (policy-head) at (2.5,1.2) {Policy};
+			\draw[color=uofgcobalt,opacity=0.7] (2,0) rectangle ++(2,1) node[pos=.5] (t1p) {Tiling 1};
+			\fill[uofgcobalt,opacity=0.25] (3.33,0) rectangle ++(0.67,0.33);
+
+			\draw[color=uofgpumpkin,opacity=0.9] (2,-1) rectangle ++(2,1) node[pos=.5] (t2p) {Tiling 2};
+			\fill[uofgpumpkin,opacity=0.25] (2.67,-0.67) rectangle ++(0.67,0.33);
+
+			\draw (2,-2) rectangle ++(2,1) node[pos=.5] (tdot) {$\cdots$};
+
+			\draw [->,color=uofgcobalt, bend left] (t1g) to (t1p.west);
+			\draw [->,color=uofgpumpkin, bend right] (t2g) to (t2p.west);
+
+			\node (act-list) at (1,-2.5) {$\bm{a}=\left[ \cdots \right]$};
+
+			\draw [->, bend right] (t1p.west) to (act-list);
+			\draw [->, bend right] (t2p.west) to (act-list);
+			\draw [->, bend right] (tdot.west) to (act-list);
+
+		\end{tikzpicture}
+	}
+\end{frame}
+
+\begin{frame}{Designs: How to exploit on-NIC parallelism?}
+	\begin{itemize}[<+->]
+		\item Netronome SmartNICs \emph{very} multi-core.
+		\begin{itemize}[<+->]
+			\item NetFPGAs also allow creating separate, effectively async functional units.
+		\end{itemize}
+		\item Two ways to take advantage:
+		\begin{itemize}[<+->]
+			\item Available threads process \Indfw{}ependently.
+			\item Available threads \Coopfw{}erate on each inference or learning task.
+		\end{itemize}
+		\item Basic algorithm:
+		\begin{itemize}[<+->]
+			\item (Parallel) action compute.
+			\item Output action.
+			\item Check for trace in progress for this input.
+			\item If found: compute $\delta$, do (parallel) policy update.
+		\end{itemize}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Designs: \emph{Ind}}
+	\centering
+	\includegraphics[width=0.52\linewidth]{../paper/figures/ind}
+\end{frame}
+
+\begin{frame}{Designs: \emph{CoOp}}
+	\centering
+	\includegraphics[width=0.535\linewidth]{../paper/figures/coop}
+\end{frame}
+
+\section{Evaluation}
+
+\begin{frame}{Metrics of interest}
+	Versus commodity hosts...
+	\begin{itemize}[<+->]
+		\item State-Action/Update latency
+		\item Online/Offline throughput
+		\item Impact on cross-traffic
+		\item Device resource use
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Latency}
+	\centering
+	\resizebox{\linewidth}{!}{\expandableinput ../tables/build/latency-only32}\pause
+
+	\textbf{Lower latency with \alert{just one (slower) core}.}\pause
+
+	\textbf{One NFP island $\implies$ \alert{\SI{15}{$\times$} median S$\rightarrow$A speedup}.}\pause
+
+	\textbf{Far tighter tail than host offload!}
+\end{frame}
+
+\begin{frame}{Latency---Bit Depth}
+	\begin{figure}
+		\centering
+		\begin{subfigure}{0.49\linewidth}
+			\resizebox{\linewidth}{!}{
+				\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
+			}
+		\end{subfigure}
+		\begin{subfigure}{0.49\linewidth}
+			\resizebox{\linewidth}{!}{
+				\includegraphics{../plots/build/rl-perf-tester/vary-work}
+			}
+		\end{subfigure}
+	\caption{\approachshort{} action and update latencies based on work size (crossover at 3-dim, 10-dim).}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Latency---Core Count}
+	\begin{figure}
+		\centering
+		\begin{subfigure}{0.49\linewidth}
+			\resizebox{\linewidth}{!}{
+				\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
+			}
+		\end{subfigure}
+		\begin{subfigure}{0.49\linewidth}
+			\resizebox{\linewidth}{!}{
+				\includegraphics{../plots/build/rl-perf-tester/vary-core}
+			}
+		\end{subfigure}
+		\caption{\approachshort{} action and update latencies based on worker count (crossover at 3-core, 8-core).}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Examining Tail Latencies}
+	\begin{figure}
+		\centering
+		\begin{subfigure}{0.45\linewidth}
+			\resizebox{1.0\linewidth}{!}{
+				\includegraphics{../plots/build/rl-perf-tester/latency-cumul}
+			}
+			\caption{\Coopfw{} latency CDFs}
+		\end{subfigure}
+		\begin{subfigure}{0.45\linewidth}
+			\resizebox{1.0\linewidth}{!}{
+				\includegraphics{../plots/build/rl-perf-tester/latency-cumul-broad}
+			}
+			\caption{\approachshort{} and host latency CDFs}
+		\end{subfigure}
+		\caption{Cumulative state-action latency plots for \approachshort{} and host-based execution.\label{fig:lat-cumul}}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Throughput}
+	\centering
+	\resizebox{\linewidth}{!}{\expandableinput ../tables/build/tput-only32}\pause
+
+	\textbf{In-NIC and quantised offers \alert{higher throughput per core}.}\pause
+
+	\textbf{\emph{Par}allel \emph{Sa}rsa key to maximising \alert{online} throughput.}
+\end{frame}
+
+\begin{frame}{Impact on cross-traffic}
+	\centering
+	\begin{figure}
+		\centering
+		\begin{subfigure}{0.45\linewidth}
+			\resizebox{1.0\linewidth}{!}{
+				%			\includegraphics{../plots/build/stress/heat-latency-median}
+				\includegraphics{../plots/build/stress/heat-latency-two-9-rel}
+				%			\includegraphics{../plots/build/stress/heat-latency-two-9-perc}
+			}
+			\caption{Deviations in \nth{99} percentile RTTs.\label{fig:dataplane-heat}}
+		\end{subfigure}
+		\hspace{0.05\linewidth}
+		\begin{subfigure}{0.45\linewidth}
+			\resizebox{1.0\linewidth}{!}{
+				\includegraphics{../plots/build/stress/histo-128B-0-3-trim}
+			}
+			\caption{\SI{128}{\byte} outlier (\nth{99}$+$\SI{222}{\nano\second})}
+		\end{subfigure}
+		\caption{Effects on tail latency of cross-traffic---typically sub-\SI{78}{\nano\second}.\label{fig:dataplane-coop}}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Resource Use}
+	\resizebox{\linewidth}{!}{
+		\begin{tabular}{
+				@{}c
+				S[table-format=4.2] S[table-format=2.2]
+				S[table-format=4.2] S[table-format=2.2]
+				S[table-format=4.2] S[table-format=2.2]
+				S[table-format=2.2] S[table-format=2.2]
+				S[table-format=3.2] S[table-format=2.2]
+				@{}
+			}
+			\toprule Firmware & \multicolumn{2}{c}{EMEM} & \multicolumn{2}{c}{EMEM Cache} & \multicolumn{2}{c}{IMEM} & \multicolumn{2}{c}{i5.CLS} & \multicolumn{2}{c}{i5.CTM}\\
+			& \multicolumn{1}{c}{\si{\mebi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} \\
+			\midrule Base P4 & 6776.67 & 88.24 & 268.52 & 2.91 & 858.28 & 10.48 & 0.00 & 0.00 & 0.00 & 0.00 \\
+			\Indfw(1) & 6780.21 & 88.28 & 2541.08 & 27.57 & 1263.28 & 15.42 & 24.75 & 38.67 & 94.25 & 36.82 \\
+			\Indfw(4) & 6780.22 & 88.28 & 2545.33 & 27.62 & 1263.28 & 15.42 & 51.18 & 79.97 & 107.00 & 41.80 \\
+			\Coopfw(1) & 6779.12 & 88.27 & 1773.59 & 19.24 & 1263.28 & 15.42 & 22.41 & 35.01 & 90.00 & 35.16 \\
+			\Coopfw(4) & 6779.12 & 88.27 & 1769.84 & 19.20 & 1263.28 & 15.42 & 52.16 & 81.49 & 90.00 & 35.16 \\
+			\bottomrule
+		\end{tabular}
+	}
+\end{frame}
+
+\begin{frame}{Network deployment/configurability}
+	\begin{itemize}[<+->]
+		\item \textbf{\Indfw{}}: \SI{27}{\micro\second}
+		\item \textbf{\Coopfw{}}: \SIrange{54}{238}{\micro\second}
+		\item \textbf{New policy data?} \alert{Just memcopies}.
+		\item Only design, bit depth, max policy sizes need recompile.
+		\item Can \alert{mix and match agent types in the network}, export learned policy over control plane.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}[standout]
+	Takeaways:
+	\begin{itemize}
+		\item \alert{Online in-NIC RL is possible!}
+		\item \alert{Order-of-magnitude latency improvement} over offloading, higher online throughput.
+		\item Platform-specific, but similar design for SmartNIC hardware class.
+		\item \alert{Future work}: accuracy, use-cases (AQM, DDoS prevention), NetFPGA, transfer learning.
+	\end{itemize}
+	\alert{Questions?}\\
+	{
+		\scriptsize
+		\vspace{2em}\faEnvelopeO{} \href{mailto:k.simpson.1@research.gla.ac.uk}{\nolinkurl{k.simpson.1@research.gla.ac.uk}}\\
+		\vspace{-0.8em}	\small{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}
+	}
+\end{frame}
+
+\appendix
+
+\begin{frame}{Scheduler performance}
+	\begin{figure}
+		\resizebox{0.8\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
+		}
+		\caption{Action/update compute times in a \SI{32}{\bit} \Coopfw{} agent under different work schedulers.\label{fig:work-alloc-32}}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Per-worker throughput}
+	\begin{figure}
+		\resizebox{0.7\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/vary-core-tput-on-per}
+		}
+		\caption{Throughput per added worker in a \Coopfw{} agent.\label{fig:tput-per-core}}
+	\end{figure}
+\end{frame}
+
+\begin{frame}[allowframebreaks]{References}
+	\printbibliography[heading=none]
+\end{frame}
+
 \end{document}

### 2021-06-03 ngni-talk/talk.tex
Pres mode...

@@
 
-	\begin{subequations}
-		\begin{gather}
-			\delta_t = \overbrace{R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}}}^{\mathclap{\substack{\text{New target value:}\\ \text{Reward} + \text{some of the next action's value}}}} - \underbrace{\acval{S_t}{A_t}{\wvec{t}}}_{\mathclap{\text{Current value estimate}}},\\\pause
-			\bm{w}_{t+1} = \bm{w}_{t} + \overbrace{\alpha \delta_t}^{\mathclap{\text{Move a little bit of } \delta_t \text{ along...}}} \underbrace{\nabla{\acval{S_t}{A_t}{\wvec{t}}}}_{\mathclap{\text{...the policy's gradient}}},\pause
-		\end{gather}%
-	\end{subequations}
+	How should we adjust the value of selected items?
+	$$
+			\delta_t = \overbrace{R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}}}^{\mathclap{\substack{\text{New target value:}\\ \text{Reward} + \text{some of the next action's value}}}} - \underbrace{\acval{S_t}{A_t}{\wvec{t}}}_{\mathclap{\text{Current value estimate}}},
+	$$\pause
+
+	Policy parameter update:
+	$$
+			\bm{w}_{t+1} = \bm{w}_{t} + \overbrace{\alpha \delta_t}^{\mathclap{\text{Move a little bit of } \delta_t \text{ along...}}} \underbrace{\nabla{\acval{S_t}{A_t}{\wvec{t}}}}_{\mathclap{\text{...the policy's gradient}}}.
+	$$\pause
 

### 2021-06-04 ngni-talk/talk.tex
Branding (tm)

@@
 
+	How to select an action? Pick largest from list: $\acval{S_t}{\cdot}{\wvec{t}}$\pause
+
 	How should we adjust the value of selected items?
@@
 	$$\pause
-
 	Policy parameter update:
@@
 	$$\pause
-
 	Design implications?
@@
 	\begin{itemize}[<+->]
-		\item Netronome SmartNICs \emph{very} multi-core.
+		\item Netronome SmartNICs \emph{very} multi-core (NFP-6480).
 		\begin{itemize}[<+->]
@@
 
-\begin{frame}{Designs: \emph{Ind}}
+\begin{frame}{Designs: \emph{Ind} (on NFP)}
 	\centering
@@
 
-\begin{frame}{Designs: \emph{CoOp}}
+\begin{frame}{Designs: \emph{CoOp} (on NFP)}
 	\centering
@@
 		\item Device resource use
-	\end{itemize}
+	\end{itemize}\pause
+	\textbf{On large-ish policies} (20D state, 10 actions, bias+(7$\times$1D, 8$\times$2D, 1$\times$4D)).
 \end{frame}
@@
 		\item Platform-specific, but similar design for SmartNIC hardware class.
-		\item \alert{Future work}: accuracy, use-cases (AQM, DDoS prevention), NetFPGA, transfer learning.
+		\item \alert{Future work}: use-cases (AQM, DDoS prevention \& accuracy), NetFPGA, transfer learning.
 	\end{itemize}
@@
 	}
+
+	\begin{tikzpicture}[overlay, remember picture]
+		%		\node[above right=0.8cm and 0.9cm of current page.south west] (esnet-logo) {\includegraphics[width=2.75cm]{netlab-trim}};
+		%		\node[right=1cm of esnet-logo] {\adjincludegraphics[height=2cm,trim={0 {.4\height} 0 {.05\height}},clip]{uofg}};
+		\node[above left=0.2cm and 0.8cm of current page.south east] {\includegraphics[width=2.75cm]{netlab-fulllogo}};
+		\node[above right=0.35cm and 0.8cm of current page.south west] {\includegraphics[width=2.75cm]{uofg-white}};
+	\end{tikzpicture}
 \end{frame}
@@
 
+\begin{frame}{Latency---All \approachshort}
+	\centering
+	\resizebox{\linewidth}{!}{\expandableinput ../tables/build/latency}
+\end{frame}
+
+\begin{frame}{Throughput---All \approachshort}
+	\centering
+	\resizebox{\linewidth}{!}{\expandableinput ../tables/build/tput}
+\end{frame}
+
 \begin{frame}[allowframebreaks]{References}

### 2021-06-06 paper/paper-old.tex
Paper-prog

--- 2021-06-02 paper/paper.tex
@@
 % The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{Simpson \emph{et al}.}
+%\renewcommand{\shortauthors}{Simpson \emph{et al}.}
 
 \begin{abstract}
-??
-In-NIC execution allows clever use of resources to bring 99.99th percentile action compute latency  from \SI{725}{\micro\second} to \SI{44}{\micro\second} (1.5 orders of magnitude!) in real-world policies for DDoS detection/prevention control.
-
-We present \emph{\approachshort{}}---\approach{}---to the best of our knowledge the first online RL in the dataplane...
-
-?? We are the first to bring fully online RL to the dataplane!
-
-?? Emphasise here and elsewhere: scales with core count, but with some designs (\Indfw) achieve better latency and per-core throughput with just one core/worker!
+Automatic, data-driven networking is becoming more capable and widely researched, driven by the efficacy of \emph{deep reinforcement learning} (DRL) algorithms.
+Yet the complexity of both DRL inference and learning force these tasks to be offloaded \emph{away from the dataplane} to hosts, harming latency-sensitive applications.
+Online learning and control have been lost, despite being useful techniques when problems evolve or are hard to model.
+
+We present \emph{\approachshort{}}---\approach{}---to our knowledge the first work to bring \emph{online reinforcement learning} to the dataplane.
+\approachshort{} makes online learning possible in constrained SmartNIC hardware by returning to classical RL techniques, which simplify update logic and enable parallel processing.
+Versus offloading, we achieve a \SI{21}{$\times$} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{$\times$} improvement in online throughput for real-world policy designs.
+In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
+\approachshort{}'s designs scale with additional resources at compile-time to improve upon both decision latency and throughput, improving upon offloading using \emph{just one worker}, are quickly reconfigurable at runtime, and are task-agnostic.
+
 \end{abstract}
@@
 \section{Introduction}
-?? REDO ME: this is an old abstract
-
-?? Make a point of explaining why in-NIC... in general!
-
 \begin{figure}
@@
 \end{figure}
-
 Automatic optimisation, control, and defence of networks are at last becoming commonplace.
-Data-driven networking has led the charge in traffic optimisation, congestion control and packet classification via adaptive techniques such as Reinforcement Learning (RL), where every change and its measured effects further improve future decisions.
-Considering that the network evolves in its use and deployed protocols, this flexibility is essential.
-Yet data-driven methods suffer from a key weakness: they are dependent on both the recency and accuracy of input state.
-An out-of-date view of the world will lead to suboptimal choices, as will long processing times.
-These can result in worse performance and slower adaptation to the evolving network.
-
-Programmable dataplanes and in-switch compute, then, hold promise for integrating these new techniques in a feasible and efficient manner (beyond dedicated servers or virtualised network functions).
+Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing the automation of traffic/multimedia optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
+In RL methods, every change and its measured effects improve future decisions made by an agent.
+
+Parallel to this, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and NetFPGA devices~\parencite{DBLP:journals/micro/ZilbermanACM14} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
+This has manifested in novel uses of programmable dataplane hardware to accelerate distributed tasks, fine-grained traffic measurement~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}, or per-packet processing.
+The latter is particularly attractive due to the costs of offloading to commodity host machines; input data must cross the PCIe bus several times, throughput requirements often impose the use of batching, and dedicated accelerators must again be contacted over PCIe~\parencite{DBLP:journals/corr/abs-2009-02353}.
+Additionally in multi-tenant environments, IOMMU contention for both NICs and inference accelerators may impact tail latencies~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
+In response, we have seen classical ML~\parencite{DBLP:conf/hotnets/XiongZ19} techniques, neural networks~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome}, and hardware design proposals for CGRA-based map-reduce blocks~\parencite{DBLP:journals/corr/abs-2002-08987} \emph{directly in the dataplane}.
+These works have shown the value in in-network ML: high-throughput, low latency response to network changes, particularly when co-hosted with state extraction.
+
+However, the task of \emph{training} these models online \emph{and} in-network presents an exciting (and challenging) lacuna in the field.
+For instance, \emph{deep neural network} (DNN) training relies upon the backpropagation algorithm to compute gradients, and for sizeable amounts of data to be held in mini-batches for stable training.
+Moreover, DNN-backed RL must often be trained by vast amounts of offline simulation---in some cases compute-years worth~\parencite{DBLP:journals/corr/abs-1912-06680}---which can be counter-productive if the target problem evolves over time or is hard to model.
+These space and compute requirements, at present, mandate that training must occur in commodity hosts or dedicated hardware.
+Moreover, high state-action latencies \emph{caused} by offloading and expensive policy approximations add significant noise to the learning process~\parencite{DBLP:journals/firai/TravnikMSP18} and impact application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
+
+%?? Make a point of explaining why in-NIC... in general!
+%?? Offloading costs
+%?? High-throughput, low response latency to network changes.
+%?? Mention PCIe offloading costs, DNN costs, potential IOMMU contention in virtualisation-based envs over NIC accesses and GPU accesses~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
+
+%?? Using classical RL? cost now is offloading latency
+%?? Using DRL? Other costs.
+%?? Deep RL training infamous for taking  
+%
+%Offloading to a host machine has several drawbacks; the additional latency
+%
+%?? In-network ML has focussed on inference, NOT online training...
+%?? But they have innovations in data formats etc.
+%Moreover, necessary hardware, such as floating-point units, is unavailable in almost all cases.
+%
+%?? P4 caused an explosion of suitable measurement techniques in plane...
+%?? Classes of device exist today
+%?? So... ripe for putting online learning agents here!
+%
+%?? So we do this...
+%?? Classical RL techniques using the advances and observations of other datapath ML.
+%?? path-adjacent, on-chip, in the dataplane (\cref{fig:netro-arch})
+%?? Exploits legacy of many p4 devices to provide general purpose compute.
+%
+%?? Can I fit in the state-slippage shit here?
+
+The question we investigate is: can online RL be brought to the dataplane by returning to computationally simpler methods, to act on locally extracted state?
+Can it be made more efficient \emph{by dataplane hardware}?
+Through this work, we can comfortably answer ``yes'' on both counts.
+In particular, we exploit the (pre-P4) legacy of general purpose compute in SmartNIC devices to offer path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
+By executing on spare compute units, we prevent packet stalling and offer quick runtime reconfigurability.
+We look backwards to classical RL techniques (tile-coded policies, single-step RL algorithms), alongside the adaptations introduced by existing in-network ML (such as quantisation).
+Such adaptations unlock further improvements in parallel inference and update processing.
+In summary, this paper contributes:
+\begin{itemize}
+	\item An analysis of why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility (\cref{sec:motivation}),
+	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands (\cref{sec:design}),
+	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm for tile-coded policies (\cref{alg:parsa}),
+	\item In-depth evaluation of how our approach affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
+	\item A potential integration of \approachshort{} with state-of-the-art PDP applications to perform fully in-NIC, automated DDoS mitigation (\cref{sec:potential-integrations}).
+\end{itemize}
+
+%Automatic optimisation, control, and defence of networks are at last becoming commonplace.
+%Data-driven networking has led the charge in traffic optimisation, congestion control and packet classification via adaptive techniques such as \emph{reinforcement learning} (RL), where every change and its measured effects further improve future decisions.
+%Considering that the network evolves in its use and deployed protocols, this flexibility is essential.
+%Yet data-driven methods suffer from a key weakness: they are dependent on both the recency and accuracy of input state.
+%An out-of-date view of the world will lead to suboptimal choices, as will long processing times.
+%These can result in worse performance and slower adaptation to the evolving network.
+%
+%Programmable dataplanes and in-switch compute, then, hold promise for integrating these new techniques in a feasible and efficient manner (beyond dedicated servers or virtualised network functions).
 For RL, key tasks include policy evaluation, online training, state collection, and action execution -- each of these introduces some degree of sensitivity to state accuracy.
@@
 Moreover, necessary hardware, such as floating-point units, is unavailable in almost all cases.
-
-The precise costs and trade-offs which operators and designers must make have yet to be identified.
+%
+%The precise costs and trade-offs which operators and designers must make have yet to be identified.
 This follows from a design space explosion induced by many necessary workarounds.
@@
 In all cases, network operators will not consider tools which affect underlying traffic.
-
-I aim to examine the effects of these choices on an existing RL-based DDoS attack mitigation system.
+%
+%I aim to examine the effects of these choices on an existing RL-based DDoS attack mitigation system.
 To protect legitimate traffic, this controls packet drop and filtering for each flow using individual metrics observed by RL agents at ingress routers, and load measurements from several points along the path taken by the flow in question.
@@
 As the market matures, these may represent different feature classes and price points.
-
-I will discuss early development efforts (including challenges and results) on the Netronome Agilio SmartNIC, which supports the P4 language as well as some degree of arbitrary microcode.
+%
+%I will discuss early development efforts (including challenges and results) on the Netronome Agilio SmartNIC, which supports the P4 language as well as some degree of arbitrary microcode.
 I intend to present how reinforcement learning execution and network telemetry will differ in the above metrics when compared to a vNF-based deployment (i.e., software on external commodity servers).
 
-This paper contributes:
-\begin{itemize}
-	\item An analysis of why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility (\cref{sec:motivation}),
-	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands (\cref{sec:design}),
-	\item Potential integrations of \approachshort{} with state-of-the-art PDP applications to perform fully in-NIC, automated DDoS mitigation and ?? OTHER THING (\cref{sec:potential-integrations}),
-	\item In-depth evaluation of how our approach affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{11.7}{$\times$} latency reduction compared to commodity hardware (\SI{16.4}{$\times$} at 99.99\nthscript{th} tail) (\cref{sec:evaluation}).
-\end{itemize}
-
 \section{Background and Motivation}\label{sec:motivation}
-We discuss recent trends in programmable switch hardware
+We discuss recent trends in programmable switch hardware.
 By combining this with insights from the ML/RL communities (past and present), we discuss why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility.
@@
 
-?? TODO: trim down. Too ``thesis-y'', rather than ``conference-y''.
-?? Save this elsewhere 'til then?
+%?? TODO: trim down. Too ``thesis-y'', rather than ``conference-y''.
+%?? Save this elsewhere 'til then?
 
 \subsection{Programmable hardware capabilities}
-The introduction of the P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} language has led to explosive growth in the research community surrounding in-network computation and offloading.
+The introduction of the P4 language~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to explosive growth in the research community surrounding in-network computation and offloading.
 Providing a single language which is compatible with network devices of many form factors has been instrumental in the development of novel fine-grained traffic measurement approaches~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
@@
 \end{insight}
-
-\fakepara{To-be-organised}
-Interesting thing I realised from \textcite{DBLP:conf/sigcomm/NeugebauerAZAL018}: GPU and NIC might cause IOMMU contention?
-Making line-rate offload of DDN to GPU impractical.
-Unsure if I want to twin this with later measurements/plots.
 
@@
 Updating the policy (with a new $\delta_t$ value) has no aggregate step, and as tile indices map to disjoint memory regions no locks are required.
-
-?? Algo block/diagram?
 
@@
 ParSa is described here at a high level, as an algorithm suited for \emph{any multiprocessor environment}.
+We omit the simple logic for $\epsilon$-greedy action selection (which we implement), and note that modification to other single-step algorithms such as \emph{Q-learning} would be trivial.
 We detail our communication primitives in \cref{sec:intra-agent-communication}, and our work scheduling strategy in \cref{sec:work-allocation}, eliding the details of tile-coding as they are well-understood.
+In general it is also possible for the \emph{Ctl} task to act as an additional \emph{Minion} in its parallel sections; we were limited here by code store requirements on the NFP.
 
@@
 Packet headers are extracted and parsed using the tooling and data types autogenerated by the P4 pipeline.
-We found that this costs a median \SIrange{126}{140}{\nano\second} communication time, comparable to message channels in the Rust and Go languages on commodity hardware.
+We found that this costs a median \SIrange{126}{140}{\nano\second} communication time (local--remote), comparable to message channels in the Rust and Go languages on commodity hardware.
 
@@
 
-?? Where to mention that algo 1 omits epsilon greedy etc?
-
 \subsection{Reconfigurability}
@@
 \fakepara{End-to-end RL latency}
-We compare the key RL decision-making latencies we discuss in \cref{fig:state-slip} ($t_1,t_2,t_3$) across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to the controller of an OpenFlow-compatible switch (OvS).
+We compare the key RL decision-making latencies we discuss in \cref{fig:state-slip} across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to a \emph{virtual network function} (vNF).
 
@@
 \fakepara{End-to-end RL latency}
-?? TODO
-?? BIGGEST TO DO.
-
-?? PCIe RTT is 10us (neugebauer, BNN NFP paper), NFP RTT is 18us (own measurements). EMEM Ring one-way is 120ns.
-
-?? In the name of fairness, assume that rule installation uses same mechanism we recommend, but show how bad it can get? I.e. with rule installation costs etc.
-
-?? Cziva~\parencite{DBLP:phd/ethos/Cziva18} discusses vNF times.
+%?? PCIe RTT is 10us (neugebauer, BNN NFP paper), NFP RTT is 18us (own measurements). EMEM Ring one-way is 120ns.
+%?? Cziva~\parencite[pp.113]{DBLP:phd/ethos/Cziva18} discusses vNF times.
+%?? reference the inference times brought up in Taurus?
+Referring to \cref{fig:state-slip}, we take $t_2$ from \cref{tab:lats} for host and in-NIC processing times, and substitute $t_1+t_3$ for the state packet round-trip time to the decision site:
+\begin{description}
+	\item[In-NIC.] As described in \cref{sec:agent-environment-communication}, EMEM rings have a median one way delay cross-island of \SI{140}{\nano\second}, giving a \emph{median \SI{34.63}{\micro\second} end-to-end inference latency}.
+	\item[Dedicated Collector.] Offloads hosted in this manner will employ DPDK to maximise performance, giving one-way PCIe delays of \SIrange{0.9}{2.3}{\micro\second} for network packets.
+	A UDP packet carrying \num{20} elements of state in \approachshort{} is \SI{128}{\byte}, so costs \SI{1}{\micro\second} to forward, and the reply state-action pair invokes a slightly higher cost.
+	\emph{This gives an end-to-end inference latency of \SI{517.9}{\micro\second}}.
+	\item[vNF Offload.] \Textcite{DBLP:journals/cm/CzivaP17} show that more lightweight vNF frameworks like GNF~\parencite{DBLP:journals/cm/CzivaP17} and ClickOS~\parencite{DBLP:conf/nsdi/MartinsAROHBH14} add \SIrange{45}{55}{\micro\second} \emph{additional} latency above PCIe costs.
+	\emph{This gives an end-to-end inference latency of \SI{572.9}{\micro\second}}.
+\end{description}
+Using these estimates, in-NIC classical RL inference offers \SIlist{14.96;16.54}{$\times$} speedups in latency over collector and vNF deployments respectively (assuming no steering cost in either case).
+We also contrast these against deep RL policies on network tasks, which can take \SI{3}{\milli\second} to compute~\parencite{DBLP:journals/corr/abs-1910-04054}---2 orders of magnitude above \approachshort{}.
+%In the name of fairness, we assume that rule installation uses same mechanism we recommend, but show how bad it can get? I.e. with rule installation costs etc.
 
@@
 \section{Potential Integrations}\label{sec:potential-integrations}
-To show the general applicability and use of \approachshort{}, we propose an ideal integrations which would benefit from in-NIC reinforcement learning; online DDoS attack mitigation.
+To show the general applicability and use of \approachshort{}, we propose an ideal integration which would benefit from in-NIC reinforcement learning; online DDoS attack mitigation.
 We support this using other state-of-the-art P4/PDP developments, and discuss how best to balance the concerns of online training with throughput in widespread deployment.
@@
 
-?? Any way to move some of this sooner? I.e., this is ``why move DDoS prevention here'', Stefanos suggested backporting some of this to justify ``why \emph{RL} in-NIC''?
+%?? Any way to move some of this sooner? I.e., this is ``why move DDoS prevention here'', Stefanos suggested backporting some of this to justify ``why \emph{RL} in-NIC''?
 
@@
 To collate the required inputs and state, we then examine the recent innovations of the community.
-Low-latency, pure-P4 solutions to extract and record per-flow TCP state directly in the dataplane such as Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} are well-studied.
+Low-latency, pure-P4 solutions to extract and record per-flow TCP state directly in the dataplane such as Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} and Sonata~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18} are well-studied.
 In fact, the statistics offered by Dapper are a super-set of the local input state values employed by \emph{Guarded} agents, offering an opportunity to further improve their efficacy. 
@@
 
-A recent line of research in the community has been to investigate the use of \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate, low latency packet classification.
+A recent line of research in the community has been to investigate \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate, low latency packet classification.
 \emph{BaNaNa SPLIT} shows the initial steps of this trend investigating their use as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} in an effort to reduce latency imposed by batching and bulk transfer to the GPU.
@@
 
-Event-driven versions of the P4 pipeline have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}, 
+Event-driven versions of the P4 pipeline have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}.
 Timer events and device state changes would empower in-network RL use-cases, signalling timesteps for RL agents or a new, effective, fine-grained sources of input state.
@@
 \section{Conclusion}
-
-?? Future work: allow scaling over multiple islands.
-?? Future work: NetFPGA it.
-
-?? Future directions for community: want to see more general-purpose async compute on-chip! Larger form-factors! More asynchrony!
-
-?? Future directions; transfer learning between classical methods into faster offline-only?
-
-?? Are there other ``old ways of doing things'' worth looking at?
+We have presented \emph{\approachshort{}}, bringing \emph{online reinforcement learning} to the dataplane.
+\approachshort{} has shown how classical RL techniques make online learning possible by simplifying update logic and enable parallel processing.
+These algorithms (and in-NIC deployment) enabled a \SIrange{15}{21}{$\times$} reduction in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput.
+The deployment environment and asynchronous design were shown to eliminate PCIe and impose minimal impact on carried dataplane traffic.
+\approachshort{}'s designs scale with additional compute resources at deployment to improve upon both decision latency and throughput.
+Our throughput-optimal design, \Indfw{}, improves upon these metrics with \emph{just one worker}.
+
+In future, we aim to examine the relative performance of individual applications driven by \approachshort---both classical and deep RL-based---and how a NetFPGA implementation can offer further latency and throughput improvements.
+A promising avenue here would be to investigate constant transfer learning between a handful of online \approachshort{} agents and other high-throughput offline function approximators such as BNNs.
+It is our hope that high port-density programmable devices will, in time, expose similar asynchronous, on-chip offloading capability to current SmartNIC devices.
+We also are interested in `classical' approaches for other tasks which may have been overlooked in favour of more performant (yet complex) solutions, yet enable in-network deployment.
 

### 2021-06-06 code/nfp/rl-test/packet/packet.c
Slight whoopsie...

@@ -540,7 +540,7 @@ void state_packet(
 			tile_t dt;
 			// don't copy the tile list, that's probably TOO heavy.
 		
-			dt = matched_reward + quant_mul(cfg->alpha, value_of_chosen_action, cfg->quantiser_shift) - state_action_pairs[state_found].val;
+			dt = matched_reward + quant_mul(cfg->gamma, value_of_chosen_action, cfg->quantiser_shift) - state_action_pairs[state_found].val;
 
 			adjustment = quant_mul(adjustment, dt, cfg->quantiser_shift);
 

### 2021-06-06 code/nfp/rl-test/rl.c

@@ -303,7 +303,7 @@ void state_packet_delegate(
 			tile_t adjustment = cfg->alpha;
 			tile_t dt;
 		
-			dt = matched_reward + quant_mul(cfg->alpha, value_of_chosen_action, cfg->quantiser_shift) - state_action_pairs[state_found].val;
+			dt = matched_reward + quant_mul(cfg->gamma, value_of_chosen_action, cfg->quantiser_shift) - state_action_pairs[state_found].val;
 
 			adjustment = quant_mul(adjustment, dt, cfg->quantiser_shift);
 

### 2021-06-06 paper/paper.tex
Most Dimitris feedback applied...

@@
 % The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{Simpson \emph{et al}.}
+%\renewcommand{\shortauthors}{Simpson \emph{et al}.}
 
 \begin{abstract}
-??
-In-NIC execution allows clever use of resources to bring 99.99th percentile action compute latency  from \SI{725}{\micro\second} to \SI{44}{\micro\second} (1.5 orders of magnitude!) in real-world policies for DDoS detection/prevention control.
-
-We present \emph{\approachshort{}}---\approach{}---to the best of our knowledge the first online RL in the dataplane...
-
-?? We are the first to bring fully online RL to the dataplane!
-
-?? Emphasise here and elsewhere: scales with core count, but with some designs (\Indfw) achieve better latency and per-core throughput with just one core/worker!
+Automatic, data-driven networking is becoming more capable and widely researched, mainly driven by the efficacy of \emph{Deep Reinforcement Learning} (DRL) algorithms.
+Yet the complexity of both DRL inference and learning force these tasks to be offloaded \emph{away from the dataplane} to hosts, harming latency-sensitive applications.
+Online learning of such policies cannot occur in the dataplane, despite being useful techniques when problems evolve or are hard to model.
+
+We present \emph{\approachshort{}}---\approach{}---the first work to bring \emph{online reinforcement learning} to the dataplane.
+\approachshort{} makes online learning possible in constrained SmartNIC hardware by returning to classical RL techniques which simplify update logic and enable parallel processing.
+Compared to offloading, we achieve a \SI{21}{$\times$} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{$\times$} improvement in online throughput for real-world policy designs.
+In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
+\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, improving on offloading, and are quickly reconfigurable at runtime compared to reinstalling device firmware/designs.
+
 \end{abstract}
@@
 \end{figure}
-
 Automatic optimisation, control, and defence of networks are at last becoming commonplace.
-Data-driven networking has led the charge in traffic optimisation, congestion control and packet classification via adaptive techniques such as Reinforcement Learning (RL), where every change and its measured effects further improve future decisions.
-Considering that the network evolves in its use and deployed protocols, this flexibility is essential.
-Yet data-driven methods suffer from a key weakness: they are dependent on both the recency and accuracy of input state.
-An out-of-date view of the world will lead to suboptimal choices, as will long processing times.
-These can result in worse performance and slower adaptation to the evolving network.
-
-Programmable dataplanes and in-switch compute, then, hold promise for integrating these new techniques in a feasible and efficient manner (beyond dedicated servers or virtualised network functions).
+Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing the automation of traffic/multimedia optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
+In RL methods, every change and its measured effects improve future decisions made by an agent.
+
+Parallel to this, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and programmable network hardware~\parencite{DBLP:journals/micro/ZilbermanACM14, netronome-smartnic, xilinx-alveo, barefoot-intel} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
+This has manifested in novel uses of programmable dataplane hardware to accelerate distributed tasks, fine-grained traffic measurement~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}, or per-packet processing.
+The latter is particularly attractive due to the costs of offloading to commodity host machines; input data must cross the PCIe bus several times, throughput requirements often impose the use of batching, and dedicated accelerators must again be contacted over PCIe~\parencite{DBLP:journals/corr/abs-2009-02353}.
+Additionally, in multi-tenant environments, IOMMU contention for both NICs and inference accelerators may impact tail latencies~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
+In response, we have seen non-neural ML techniques~\parencite{DBLP:conf/hotnets/XiongZ19}, neural networks~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome}, and hardware design proposals for CGRA-based map-reduce blocks~\parencite{DBLP:journals/corr/abs-2002-08987} \emph{directly in the dataplane}.
+These works have shown the value in in-network ML: high-throughput, low latency response to network changes, particularly when co-hosted with state extraction.
+
+However, the task of \emph{training} these models online \emph{and} in-network presents an exciting (and challenging) lacuna in the field.
+Online training is particularly important if an agent needs to react to unforeseen (failure) states, such as in DDoS/intrusion prevention.
+For context, \emph{Deep Neural Network} (DNN) training relies upon the backpropagation algorithm to compute gradients, and for sizeable amounts of data to be held in mini-batches for stable training.
+Moreover, DNN-backed RL must often be trained by vast amounts of offline simulation---in some cases compute-years worth~\parencite{DBLP:journals/corr/abs-1912-06680}---which can be counter-productive if the target problem evolves over time or is hard to model.
+These space and compute requirements, at present, mandate that training must occur in commodity hosts or dedicated hardware.
+Moreover, high state-action latencies \emph{caused} by offloading and expensive policy approximations add significant noise to the learning process~\parencite{DBLP:journals/firai/TravnikMSP18} and impact application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
+
+%?? Make a point of explaining why in-NIC... in general!
+%?? Offloading costs
+%?? High-throughput, low response latency to network changes.
+%?? Mention PCIe offloading costs, DNN costs, potential IOMMU contention in virtualisation-based envs over NIC accesses and GPU accesses~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
+
+%?? Using classical RL? cost now is offloading latency
+%?? Using DRL? Other costs.
+%?? Deep RL training infamous for taking  
+%
+%Offloading to a host machine has several drawbacks; the additional latency
+%
+%?? In-network ML has focussed on inference, NOT online training...
+%?? But they have innovations in data formats etc.
+%Moreover, necessary hardware, such as floating-point units, is unavailable in almost all cases.
+%
+%?? P4 caused an explosion of suitable measurement techniques in plane...
+%?? Classes of device exist today
+%?? So... ripe for putting online learning agents here!
+%
+%?? So we do this...
+%?? Classical RL techniques using the advances and observations of other datapath ML.
+%?? path-adjacent, on-chip, in the dataplane (\cref{fig:netro-arch})
+%?? Exploits legacy of many p4 devices to provide general purpose compute.
+%
+%?? Can I fit in the state-slippage shit here?
+
+The question we investigate is: can online RL be brought to the dataplane by returning to computationally simpler methods, to act on locally extracted state?
+Can it be made more efficient \emph{by dataplane hardware}?
+Through this work, we can comfortably answer ``yes'' on both counts.
+In particular, we exploit how SmartNIC devices often expose general purpose compute to provide path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
+As many of these devices have engineering and development histories which predate P4, general compute beyond P4's limits~\parencite{p4-psa} is surprisingly common.
+By executing on spare compute units, we prevent packet stalling and offer quick runtime reconfigurability.
+We look backwards to classical RL techniques (tile-coded policies, single-step RL algorithms), alongside the adaptations introduced by existing in-network ML (such as quantisation).
+Such adaptations unlock further improvements in parallel inference and update processing.
+In summary, this paper contributes:
+\begin{itemize}
+	\item An analysis of why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility (\cref{sec:motivation}),
+	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands (\cref{sec:design}),
+	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm for tile-coded policies (\cref{alg:parsa}),
+	\item In-depth evaluation of how our approach affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
+	\item A potential integration of \approachshort{} with state-of-the-art PDP applications to perform fully in-NIC, automated DDoS mitigation (\cref{sec:potential-integrations}).
+\end{itemize}
+
+%Automatic optimisation, control, and defence of networks are at last becoming commonplace.
+%Data-driven networking has led the charge in traffic optimisation, congestion control and packet classification via adaptive techniques such as \emph{reinforcement learning} (RL), where every change and its measured effects further improve future decisions.
+%Considering that the network evolves in its use and deployed protocols, this flexibility is essential.
+%Yet data-driven methods suffer from a key weakness: they are dependent on both the recency and accuracy of input state.
+%An out-of-date view of the world will lead to suboptimal choices, as will long processing times.
+%These can result in worse performance and slower adaptation to the evolving network.
+%
+%Programmable dataplanes and in-switch compute, then, hold promise for integrating these new techniques in a feasible and efficient manner (beyond dedicated servers or virtualised network functions).
 For RL, key tasks include policy evaluation, online training, state collection, and action execution -- each of these introduces some degree of sensitivity to state accuracy.
@@
 As the market matures, these may represent different feature classes and price points.
-
-I will discuss early development efforts (including challenges and results) on the Netronome Agilio SmartNIC, which supports the P4 language as well as some degree of arbitrary microcode.
+%
+%I will discuss early development efforts (including challenges and results) on the Netronome Agilio SmartNIC, which supports the P4 language as well as some degree of arbitrary microcode.
 I intend to present how reinforcement learning execution and network telemetry will differ in the above metrics when compared to a vNF-based deployment (i.e., software on external commodity servers).
 
-This paper contributes:
-\begin{itemize}
-	\item An analysis of why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility (\cref{sec:motivation}),
-	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands (\cref{sec:design}),
-	\item Potential integrations of \approachshort{} with state-of-the-art PDP applications to perform fully in-NIC, automated DDoS mitigation and ?? OTHER THING (\cref{sec:potential-integrations}),
-	\item In-depth evaluation of how our approach affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{11.7}{$\times$} latency reduction compared to commodity hardware (\SI{16.4}{$\times$} at 99.99\nthscript{th} tail) (\cref{sec:evaluation}).
-\end{itemize}
-
-\section{Background and Motivation}\label{sec:motivation}
-We discuss recent trends in programmable switch hardware
+\section{Preliminaries: RL for in-network computation}\label{sec:motivation}
+We discuss recent trends in programmable switch hardware.
 By combining this with insights from the ML/RL communities (past and present), we discuss why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility.
 
-Crucially, we train our focus on devices with a low port density, such as SmartNICs and NetFPGA devices.
+Crucially, we train our focus on devices with a low port density, such as SmartNICs and the NetFPGA.
 The designs of these devices make it reasonable to move policy processing \emph{outside of the packet pipeline}; SmartNICs expose many additional programmable cores, and NetFPGAs allow for the synthesis of independent functional units.
 The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
-
-?? TODO: trim down. Too ``thesis-y'', rather than ``conference-y''.
-?? Save this elsewhere 'til then?
+%?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
+As ethernet line rate moves beyond \SIlist{40;100}{\giga\bit\per\second}, packet processing deadlines grow tighter in tandem; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second}, giving a \SI{3.99}{\micro\second} deadline on Netronome SmartNICs (having \num{312} P4 pipelines).
+In-pipeline execution thus increases the risk of drops or stalled packet transmissions.
+
+%However, this requires that there be a reasonably consistent interface and behaviour between device classes.
+%This need for consistency underpins the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}, which defines a conceptual model of match-action tables divided into ingress and egress pipelines.
+%The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.
+
+%?? TODO: trim down. Too ``thesis-y'', rather than ``conference-y''.
+%?? Save this elsewhere 'til then?
 
 \subsection{Programmable hardware capabilities}
-The introduction of the P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} language has led to explosive growth in the research community surrounding in-network computation and offloading.
-Providing a single language which is compatible with network devices of many form factors has been instrumental in the development of novel fine-grained traffic measurement approaches~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
-However, this requires that there be a reasonably consistent interface and behaviour between device classes.
-This need for consistency underpins the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}, which defines a conceptual model of match-action tables divided into ingress and egress pipelines.
-The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.
-
+%The introduction of the P4 language~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to explosive growth in the research community surrounding in-network computation and offloading.
+%Providing a single language which is compatible with network devices of many form factors has been instrumental in the development of novel fine-grained traffic measurement approaches~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
+%However, this requires that there be a reasonably consistent interface and behaviour between device classes.
+%This need for consistency underpins the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}, which defines a conceptual model of match-action tables divided into ingress and egress pipelines.
+%The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.
+
+%We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
+%Many compatible devices have legacies long predating these developments. 
+%For instance, many-core SOC-based Netronome SmartNIC~\parencite{netronome-smartnic}, and NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14} NICs allow virtually arbitrary programs to be specified and executed.
+%The first of these compiles from P4$\rightarrow$Micro-C$\rightarrow$bytecode (with Micro-C externs), while the latter can combine the P4$\rightarrow$NetFPGA toolchain~\parencite{DBLP:conf/fpga/IbanezBMZ19} with arbitrary circuit externs---both exposing further device capabilities beyond the specification.
+%This pattern extends to other SmartNICs, such as NVIDIA BlueField~\parencite{nvidia-bluefield} and Xilinx Alveo~\parencite{xilinx-alveo} devices.
+
+While the P4 language~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to great interest in network programmability, it requires similar behaviour between device classes.
+This is encoded as the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}---a sensible lower bound on the device capabilities required to implement a P4 dataplane.
 We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
 Many compatible devices have legacies long predating these developments. 
-For instance, many-core SOC-based Netronome SmartNIC~\parencite{netronome-smartnic}, and NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14} NICs allow virtually arbitrary programs to be specified and executed.
-The first of these compiles from P4$\rightarrow$Micro-C$\rightarrow$bytecode (with Micro-C externs), while the latter can combine the P4$\rightarrow$NetFPGA toolchain~\parencite{DBLP:conf/fpga/IbanezBMZ19} with arbitrary circuit externs---both exposing further device capabilities beyond the specification.
-This pattern extends to other SmartNICs, such as NVIDIA BlueField~\parencite{nvidia-bluefield} and Xilinx Alveo~\parencite{xilinx-alveo} devices.
+For instance, many-core SOC-based Netronome SmartNIC~\parencite{netronome-smartnic}, and NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19} NICs allow virtually arbitrary programs to be specified and executed.
+This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
 
@@
 
-\begin{insight}
-	Current low-port density programmable network devices often have spare resources and extra capabilities beyond the PSA specification which can aid asynchronous, local compute.
-\end{insight}
-
-\fakepara{To-be-organised}
-Interesting thing I realised from \textcite{DBLP:conf/sigcomm/NeugebauerAZAL018}: GPU and NIC might cause IOMMU contention?
-Making line-rate offload of DDN to GPU impractical.
-Unsure if I want to twin this with later measurements/plots.
+%\begin{insight}
+%	Current low-port density programmable network devices often have spare resources and extra capabilities beyond the PSA specification which can aid asynchronous, local compute.
+%\end{insight}
 
 \subsection{Reinforcement Learning}
-\emph{Reinforcement learning} (RL) algorithms are methods of training an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task \cite{RL2E}.
-Like most machine learning methods, an RL algorithm uses gradient information to update the parameters used to approximate a function.
-In RL contexts, this is a state$\rightarrow$action function known as a \emph{policy}.
-When training, agents are given \emph{reward measurements} and a learned policy acts to maximise the \emph{expected discounted reward} received.
-When a pre-trained policy is deployed, this signal is not required.
-
-It's useful to consider how these algorithms differ from other ML use cases, such as classifiers.
-The main differences lie in how this gradient information is used, and combined with \emph{reward measurements} received from the environment.
-Rather than adapting the learned parameters along the gradient using an error value from a target value, because the optimal actions aren't known they adjust values using a \emph{Markov Decision Process} (MDP)---capturing state trajectories to adjust value based on past and future decisions.
-
-Consider the single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E}:
-\begin{subequations}
-	\begin{gather}
-		\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
-		\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
-	\end{gather}%
-	\label{eqn:sg-sarsa}%
-	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, $\acval{S}{A}{\wvec{}}$ denotes the \emph{value} of an action $A$ taken in state $S$ according to the policy $\wvec{}$, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
+%\emph{Reinforcement learning} (RL) algorithms are methods of training an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task \cite{RL2E}.
+%Like most machine learning methods, an RL algorithm uses gradient information to update the parameters used to approximate a function.
+%In RL contexts, this is a state$\rightarrow$action function known as a \emph{policy}.
+%When training, agents are given \emph{reward measurements} and a learned policy acts to maximise the \emph{expected discounted reward} received.
+%When a pre-trained policy is deployed, this signal is not required.
+%
+%It's useful to consider how these algorithms differ from other ML use cases, such as classifiers.
+%The main differences lie in how this gradient information is used, and combined with \emph{reward measurements} received from the environment.
+%Rather than adapting the learned parameters along the gradient using an error value from a target value, because the optimal actions aren't known they adjust values using a \emph{Markov Decision Process} (MDP)---capturing state trajectories to adjust value based on past and future decisions.
+
+\emph{Reinforcement learning} (RL) algorithms train an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task~\parencite{RL2E}.
+Like most ML methods, RL algorithms use gradient information to update the parameters used to approximate a function with \emph{reward measurements} from the environment.
+RL's \emph{Markov Decision Process} (MDP) structure allows online learning of a state-action map in a model-free way, and can step through local minima when needed compared to ML.
+In networking, this allows for learning to respond to on-device state or to handle rapidly evolving problems.
+Generally these are agnostic to the policy approximation used, and classical single-step RL algorithms are simple---most of the compute work is occupied by the policy itself.
+Consider the single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E} which requires only additions and multiplication to update and learn a policy online.
+
+%Consider the single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E}:
+%\begin{subequations}
+%	\begin{gather}
+%		\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
+%		\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
+%	\end{gather}%
+%	\label{eqn:sg-sarsa}%
+%	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, $\acval{S}{A}{\wvec{}}$ denotes the \emph{value} of an action $A$ taken in state $S$ according to the policy $\wvec{}$, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
 $\gamma,\alpha\in[0,1]$ are the discount factor and learning rate (governing the degree of forward planning and policy stability, respectively).
-\end{subequations}
-In essence, at each timestep the policy parameters ($\wvec{}$) are increased along the gradient ($\nabla{\acvalblank}$) using a fixed learning rate ($\alpha$) and a computed adjustment ($\delta_t$).
-This adjustment is equal to the difference between the chosen action $A$'s value in state $S$ and the reward received ($R_{t+1} - \acval{S_t}{A_t}{\wvec{}}$), \emph{plus} some part of the \emph{next} action's value ($\gamma\acval{S_{t+1}}{A_{t+1}}{\wvec{}}$).
-To give some context on the design space here, other algorithms may employ separate state value approximations, use the entirety of an agent's trace, or be tailored to characteristics of the policy approximator (e.g., how neural networks benefit from batching).
-
-\begin{insight}
-	Classical, single-step RL algorithms are computationally simple and are agnostic to the policy approximation used.
+%\end{subequations}
+%In essence, at each timestep the policy parameters ($\wvec{}$) are increased along the gradient ($\nabla{\acvalblank}$) using a fixed learning rate ($\alpha$) and a computed adjustment ($\delta_t$).
+%This adjustment is equal to the difference between the chosen action $A$'s value in state $S$ and the reward received ($R_{t+1} - \acval{S_t}{A_t}{\wvec{}}$), \emph{plus} some part of the \emph{next} action's value ($\gamma\acval{S_{t+1}}{A_{t+1}}{\wvec{}}$).
+%To give some context on the design space here, other algorithms may employ separate state value approximations, use the entirety of an agent's trace, or be tailored to characteristics of the policy approximator (e.g., how neural networks benefit from batching).
+
+%\begin{insight}
+%	Classical, single-step RL algorithms are computationally simple and are agnostic to the policy approximation used.
 Furthermore, they require only additions and multiplication to update and learn a policy online.
-\end{insight}
+%\end{insight}
 
 \fakepara{RL in asynchronous environments}
-There remains some degree of divergence between the theory and implementation of RL agents.
-Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
-The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
-Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
-
+%There remains some degree of divergence between the theory and implementation of RL agents.
+%Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
+%The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
+%Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
+Consider \cref{fig:state-slip}: the traditional RL formulation assumes that an agent receives a new state at fixed time intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
 These delays (and variance thereof) add noise to the state-action mapping being learned, which has a potent reduction to learning rate and final accuracy, even for simple grid world tasks according to \textcite{DBLP:journals/firai/TravnikMSP18}.
-They in turn show that reordering algorithmic steps can reduce these costs for online 1-step algorithms, but that reducing this further requires detailed agent-environment co-design.
+Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
 This principle has influenced the design of real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, showing that asynchrony is necessary for high-speed applications.
-Achieving this often requires that state measurements are combined or coalesced~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
+This often requires that state measurements are combined or coalesced~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
 `Stopping the world' in the algorithmic sense causes significant performance degradation, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control, or any time-sensitive control problems.
@@
 
-\begin{insight}
-	Online control problems benefit from low-latency, local execution and training (i.e., in-NIC/switch).
-\end{insight}
+?? Return to dimitris note: how is this important to final solution?
+
+%\begin{insight}
+%	Online control problems benefit from low-latency, local execution and training (i.e., in-NIC/switch).
+%\end{insight}
 
 \subsection{Efficient policy approximation}
-Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
-This can then make \emph{online} learning of the task an attractive prospect, using a single available stream of experience.
-\emph{Deep neural networks} (DNNs), particularly when used as the basis for an RL agent's policy, require vast amounts of experience to converge on an accurate parameter set.
-In many problem domains, this equates to training from compute-years worth of distributed offline simulations, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
-Achieving stable learning requires sizeable batches of gradients to be computed, potentially using the entire experience replay from each simulation.
-Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the computational cost of gradient computation via backpropagation remains too significant on embedded hardware.
+%Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
+%This can then make \emph{online} learning of the task an attractive prospect, using a single available stream of experience.
+%\emph{Deep neural networks} (DNNs), particularly when used as the basis for an RL agent's policy, require vast amounts of experience to converge on an accurate parameter set.
+%In many problem domains, this equates to training from compute-years worth of distributed offline simulations, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
+%Achieving stable learning requires sizeable batches of gradients to be computed, potentially using the entire experience replay from each simulation.
+%Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the computational cost of gradient computation via backpropagation remains too significant on embedded hardware.
+Complex \emph{Deep Neural Networks} (DNNs), particularly when used as the basis for an RL agent's policy, require vast amounts (e.g., in the worst case compute-years~\parencite{DBLP:journals/corr/abs-1912-06680}) of experience to converge on an accurate parameter set, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
+Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the storage and computational costs of gradient computation via backpropagation remain too significant on embedded hardware.
 
@@
 The idea is simple: a policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
-A state vector produces a single `hit' in each tiling, which provides both the set of action preference lists to sum during inference \emph{and} the set of tiles to update at the next timestep.
+This covers any input-output mapping, such as converting packet header data into a preference list for queue priorities.
+A state vector produces a single `hit' in each tiling, all of which then correspond to a list of action preference lists.
+Inference (choosing an action) is then simply summing over all such lists to obtian a final action preference list, selecting the maximum ($a$).
+Updating at the next timestep uses this same list of lists, adjusting the value of the last selected action using an adjustment value $\delta_t$ computed through Sarsa.
 Crucially, this internal representation ensures that there are no data dependencies between any tile hit calculations for an input state vector.
-This has important benefits for parallelisation.
+This has important benefits for parallelising the key parts of an RL algorithm; both action selection and policy updates are map-reduce problems.
 To select an action, we can produce a task for each tiling---retrieving the list of action values contributed by the tile activated by the input state.
@@
 Updating the policy (with a new $\delta_t$ value) has no aggregate step, and as tile indices map to disjoint memory regions no locks are required.
-
-?? Algo block/diagram?
+As we show in \cref{sec:design}, it is possible to make the inference aggregate step wait-free.
 
@@
 
-\begin{insight}
-	Tile coded policies are embarrassingly parallel in both inference \emph{and} learning. Moreover, in some instances the process can be altered to require only bitshifts, additions, and multiplication.
-\end{insight}
+%\begin{insight}
+%	Tile coded policies are embarrassingly parallel in both inference \emph{and} learning.
+Moreover, in some instances the process can be altered to require only bitshifts, additions, and multiplication.
+%\end{insight}
 
 \subsection{Numerical representation for embedded ML}
-A key feature universally lacking from the target environments is floating-point arithmetic support.
+A key feature universally lacking from programmable network hardware is floating-point arithmetic support.
 Luckily, the task of bringing machine learning models to resource-constrained environments without these capabilities is well-studied.
@@
 Much of this work goes further still towards integer~\parencite{tensorrt-8bit} or binarised~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16} representations, sacrificing dynamic range for simpler arithmetic operations.
-The works mentions above use similar techniques to make inference tractable on network hardware, i.e., to bring neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
+The works mentioned above use similar techniques to make inference tractable on network hardware, i.e., to bring neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
 
@@
 This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation ($t_2$, \cref{fig:state-slip}).
+We note that our later parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is made possible only by the use of quantisation to integers, as atomic operations on floating point numbers typically do not exist outside of graphics- or tensor-processing hardware.
 
@@
 
-\begin{insight}
-	Fixed point quantisation allows us to perform numerical calculations using only integer arithmetic.
-	Existing literature to-date supports the effectiveness and use of these techniques on embedded ML deployments.
-\end{insight}
+%\begin{insight}
+%	Fixed point quantisation allows us to perform numerical calculations using only integer arithmetic.
+%	Existing literature to-date supports the effectiveness and use of these techniques on embedded ML deployments.
+%\end{insight}
 
 \subsection{Netronome Platform Fundamentals}
-As we implement this work on Netronome SmartNICs, it is necessary to explain the basics of the NFP architecture.
+As we implement this work on Netronome SmartNICs (thus the NFP chip architecture), it is necessary to explain its basics.
 NFP chips are \emph{system-on-a-chip} (SoC) devices, and achieve scalable packet processing through sheer parallelism.
@@
 Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair, safeguarding the limited memory of the target devices.
+
+?? How to fit this through?
 
@@
 \Cref{fig:netro-arch,fig:single-and-parallel} outline our design and implementation on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
-We open-source our firmware design and control programs for the benefit of the community.\footnote{\url{https://github.com/FelixMcFelix/pdp-rl-paper}---currently private...}
+?? Explain \cref{fig:netro-arch},
+?? Wish-based justification for P4.
+?? \Cref{fig:single-and-parallel} is explained in \cref{sec:action-and-update-computation}
+
+%We open-source our firmware design and control programs for the benefit of the community.\footnote{\url{https://github.com/FelixMcFelix/pdp-rl-paper}---currently private...}
+We open-source our firmware design and control programs for the benefit of the community.\footnote{Link redacted for double-blind submission}.
 Moving beyond the Netronome platform, we describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.
@@
 Combined with reward measurements, this policy can then be updated or trained from scratch entirely on the NIC, acting as a fully online RL agent.
-An input state vector \emph{always} induces an action, and if desired updates the policy using either an included reward or one retrieved from memory according to a (configurable) key placed alongside the state.
-This allows for simultaneous control and learning over independent systems by the same agent (i.e., optimising several flows with their own reward measures).
+An input state vector \emph{always} induces an action and, if desired, updates the policy using either an included reward or one retrieved from memory according to a (configurable) key placed alongside the state.
+This allows for simultaneous control and learning over independent systems by the same agent (i.e., optimising several flows with their own reward measures, such as DDoS mitigation in an AS where each next-hop AS might have their own `health' metric).
 
@@
 
-The main interaction model is that platform-specific IPC mechanisms (message rings) are used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
+The main interaction model is that platform-specific IPC (message rings) is used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
 These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
@@
 Bit depth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
-
-\subsection{Action and Update Computation}
+?? Move to later?
+
+\subsection{Action and Update Computation}\label{sec:action-and-update-computation}
 \approachshort{} applies the insights of \textcite{DBLP:journals/firai/TravnikMSP18} to minimise action latency; an action is computed, sent out into the environment, and only then is the underlying policy updated.
@@
 Each offers a different point of optimisation; if updates are disabled, then the \emph{\indfw{}} model can maximise throughput, while the \emph{\coopfw{}} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
-We note that our parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is made possible only by the use of quantisation to integers, as atomic operations on floating point numbers typically do not exist outside of graphics- or tensor-processing hardware.
-ParSa is described here at a high level, as an algorithm suited for \emph{any multiprocessor environment}.
-We detail our communication primitives in \cref{sec:intra-agent-communication}, and our work scheduling strategy in \cref{sec:work-allocation}, eliding the details of tile-coding as they are well-understood.
+%ParSa is described here at a high level, as an algorithm suited for \emph{any multiprocessor environment}.
+%We omit the simple logic for $\epsilon$-greedy action selection (which we implement), and note that modification to other single-step algorithms such as \emph{Q-learning} would be trivial.
+%We detail our communication primitives in \cref{sec:intra-agent-communication}, and our work scheduling strategy in \cref{sec:work-allocation}, eliding the details of tile-coding as they are well-understood.
+%In general it is also possible for the \emph{Ctl} task to act as an additional \emph{Minion} in its parallel sections; we were limited here by code store requirements on the NFP.
 
@@
 Additionally, we do not use bitshifts in place of division operations in our implementation, due to the strict limits that power-of-two tile widths place upon policy design.
+
+?? Explain that above is a general property, probably still holds particularly at high cpu speed
 
@@
 
-Our implementation uses platform-specific IPC mechanisms (\emph{EMEM ring buffers}) with hardware signalling mechanisms on work arrival to achieve this.
-Due to the lack of memory allocation in these environments, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state/configuration/policy data handover.
+Our implementation uses platform-specific IPC (\emph{EMEM ring buffers}) with hardware signalling mechanisms on work arrival to achieve this.
+Due to lack of memory allocation in these environments, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state/configuration/policy data handover.
 Packet headers are extracted and parsed using the tooling and data types autogenerated by the P4 pipeline.
-We found that this costs a median \SIrange{126}{140}{\nano\second} communication time, comparable to message channels in the Rust and Go languages on commodity hardware.
+We found that this costs a median \SIrange{126}{140}{\nano\second} communication time (local--remote), comparable to message channels in the Rust and Go languages on commodity hardware.
 
@@
 \section{Evaluation}\label{sec:evaluation}
-We investigate the performance of \approachshort{} as compared with classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{$\times$} speedup in median--\nth{99} state-action latency and \SI{9.9}{$\times$} greater online learning throughput.
-Crucially, we find that in-NIC execution offers tight tail latency bounds compared with host-based approaches.
+We investigate the performance of \approachshort{} compared to classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{$\times$} speedup in median--\nth{99} state-action latency and \SI{9.9}{$\times$} greater online learning throughput.
+Crucially, we find that in-NIC execution offers tight tail latency bounds compared to host-based approaches.
 We report on how \approachshort{} scales to fit available device resources as cores/threads are added, noting that both \Indfw{} and \Coopfw{} designs outperform commodity hosts using just one core in latency and online throughput.
@@
 Where feasible, we test \SIlist{32;16;8}{\bit} quantised arithmetic firmware versions.
-All \approachshort{} timing measurements were repeated over \num{10000} trials (preceded by \num{1000} warmup packets), from which throughput was derived.
-Host throughput measures were observed over \num{10} trials of \SI{10}{\second}.
+All \approachshort{} timing measurements were repeated over \num{10000} input state packets (preceded by \num{1000} warmup packets), storing an item processing time which accessed via the control machine, from which throughput was derived.
+Host throughput and latency measures were observed over \num{10} trials of \SI{10}{\second} (with \SI{5}{\second} warmup/cooldown times).
+We differ this from the NFP as hosts need to run numpy-based agents in parallel via separate processes; this also allows us to investigate the effects of oversubscription on host latencies.
+
 Policy sizes are set to those matching a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dimension state vectors, a bias tile and 16 full tiling sets (7$\times$1-dim, 8$\times$2-dim, 1$\times$4-dim), 8 tilings per dimension set, 6 tiles per dimension, choosing between 10 actions.
+For context, such input state would be various aspects of per-flow state (e.g., IATs, up/down rate) which are combined with other state such as the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim)
 For \Coopfw{} for instance, this creates \num{129} work items across \num{31} workers.
@@
 We compare online learning performance (input states produce an output, and then update the policy with appropriate locking) with offline (input states \emph{only} produce an output) in these cases.
+Online performance marks the number of decisions that can be made per second (and associated latency) when still training a policy.
+Offline performance is crucial for pushing a trained, known-good policy to agents in the network with an expected higher raw decision throughput.
 State-action latency is a shared property of both cases, with the main difference being the impact on throughput of the update step.
@@
 Building on this, we vary the amount of allocated worker threads to show how \approachshort{} can scale to fit available compute resources on a device.
-This also demonstrates the number of cores needed to achieve a given latency bound on a policy of representative complexity.
+This is an important aspect for (automated) administration and allocation of compute resources in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of time work scheduling which we examine later.
+This also demonstrates the number of cores needed to achieve a given latency/throughput bound on a policy of representative complexity.
 Moreover, to demonstrate how these costs vary as policy complexity increases, we vary the number of total dimensions included in the tiling.
@@
 %?? Also our simple, ahead-of-time, work scheduler.
-We verify that our heuristic, runtime work scheduler (\emph{Balanced}) makes meaningful use of the explicit memory hierarchy and cost of each work item.
-We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} baseline allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ items given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
+We verify that our heuristic, runtime work scheduler (which we term \emph{Balanced}) makes meaningful use of the explicit memory hierarchy and cost of each work item.
+We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} baseline allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ tilings given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
 We use maximum-size, \SI{32}{\bit} policies as described above.
@@
 \fakepara{End-to-end RL latency}
-We compare the key RL decision-making latencies we discuss in \cref{fig:state-slip} ($t_1,t_2,t_3$) across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to the controller of an OpenFlow-compatible switch (OvS).
+We compare the key RL decision-making latencies we discuss in \cref{fig:state-slip} across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to a \emph{virtual Network Function} (vNF).
 
@@
 We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, placing an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source via the default firmware.
-We perform throughput/loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements.
+We perform throughput/loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements (sampling at \SI{2000}{\hertz} for \num{10}$\times$\SI{10}{\second}).
+These choices allow maximum throughput in the former case (relying solely on NIC counters for loss counting).
+In the latter case we aim to minimise host resource contention so as to observe exact latency measurements, have a high enough sample count to detect subtle (aggregate) latency effects, and eliminate \emph{host} receive drops.
 DPDK was setup using four \SI{1}{\gibi\byte} hugepages.
@@
 
+?? How do I safely say: these are arbitrary?
+
 %\fakepara{The impact of bit depth}
@@
 \fakepara{Raw inference and learning performance}
-\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based (BLAS-accelerated) implementation of an RL policy\footnote{For brevity, we omit numpy-based integer results---summarised against a floating point implementation, median state-action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}.
+\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based implementation of an RL policy\footnote{For brevity, we omit numpy-based integer results---summarised against a floating point implementation, median state-action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}.
 We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using one island (4 MEs) of the NFP-6480.
@@
 When compounded with the abundance of such weaker chips, in-NIC RL is able to deliver much higher throughput.
-As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} that of a dedicated collector machine, due to the locking requirements mandating coordinated division of work.
+%As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} that of a dedicated collector machine, due to the locking requirements mandating coordinated division of work.
+As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} that of a dedicated collector machine, as the write lock around policy updates creates a bottleneck.
 
@@
 
-\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects complete update costs and state-action latency respectively, scaling from a bias tile up to the full policy size used throughout\footnote{Input vectors here all have 20 elements regardless of the policy design.}.
+\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects complete update costs and state-action latency respectively, scaling from a bias tile up to the full DDoS policy size\footnote{Input vectors here all have 20 elements regardless of the policy design.}.
 \Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning performance.
-We note that this is a trivial case, in that the use of \emph{only} a bias tile negates the need for any input state (similar to a multi-arm bandit problem).
+%We note that this is a trivial case, in that the use of \emph{only} a bias tile negates the need for any input state (similar to a multi-arm bandit problem).
+We note that this is a trivial case, as using \emph{only} a bias tile provides/alters a single preference list regardless of input state.
 
@@
 We found that a key aspect of in-NIC execution is that it allows far tighter bounds on tail latency to be offered compared to host offloading.
-Examining the state-action latencies in \cref{tab:lats}, we see that \num{99.99}\nthscript{th} percentile latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{} respectively.
+Examining the state-action latencies in \cref{tab:lats}, we see that \num{99.99}\nthscript{th} percentile latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, respectively.
 Similar results were observed for other bit depths.
@@
 
-A somewhat remarkable trend is that the \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit} designs, with the exception of smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
-We found this to be rather surprising, given that our implementation is optimised to read and write policy tile data in batches (achieving fewer I/O operations as required).
-Our understanding is that this is due to the native register width on the NFP being \SI{32}{\bit}, and so the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
+A noteworthy trend is that the \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit} designs, with the exception of smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
+This held even though our implementation is optimised to read and write policy tile data in batches (achieving fewer I/O operations as required).
+This is due to the native register width on the NFP being \SI{32}{\bit}, and so the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
 This matches \SI{32}{\bit} becoming dominant in more complex workloads: higher dimension tilings require more arithmetic operations to compute the hit tile index.
@@
 This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) \Coopfw{} agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the algorithm in the online case.
-Regardless, the \SIrange{2}{4}{$\times$} reduction in memory needed to store a policy holds.
+As a result, these bit depths give a \SIrange{2}{4}{$\times$} reduction in memory needed to store a policy (enabling more fine-grained or more complex policies), in exchange for slightly worse latency and throughput.
 
 \fakepara{Work allocation}
-\Cref{fig:work-alloc-32} shows that our heuristic first-fit allocator (\emph{Balanced}) is key in achieving consistent sub-\SIlist{35;65}{\micro\second} latencies/processing times respectively.
+\Cref{fig:work-alloc-32} shows that our heuristic first-fit allocator (\emph{Balanced}) is key in achieving consistent sub-\SIlist{35;65}{\micro\second} latencies/processing times, respectively.
 The trend is repeated for all bit depths, noting the latency differences described above.
@@
 
-An interesting aspect of \Coopfw{} and the \emph{ParSa} algorithm in general is that adding additional cores has both diminishing returns and key thresholds to pass.
+An interesting aspect of \Coopfw{} and the \emph{ParSa} algorithm in general is that adding cores has both diminishing returns and key thresholds to pass.
 Consider \cref{fig:tput-per-core}, where the throughput per worker decreases over time but occasionally increases sharply.
 Later downward ticks (\numrange{25}{29} workers) correspond to plateaus in throughput.
-This is a common problem stemming from the granularity of work items, where the heuristic scheduler cannot find a better solution to a bottleneck until extra cores are allocated.
+This is a problem stemming from the granularity of work items (i.e., tilings in \emph{ParSa}), where the heuristic scheduler cannot find a better solution to a bottleneck until extra cores are allocated.
 We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
@@
 \fakepara{End-to-end RL latency}
-?? TODO
-?? BIGGEST TO DO.
-
-?? PCIe RTT is 10us (neugebauer, BNN NFP paper), NFP RTT is 18us (own measurements). EMEM Ring one-way is 120ns.
-
-?? In the name of fairness, assume that rule installation uses same mechanism we recommend, but show how bad it can get? I.e. with rule installation costs etc.
-
-?? Cziva~\parencite{DBLP:phd/ethos/Cziva18} discusses vNF times.
+%?? PCIe RTT is 10us (neugebauer, BNN NFP paper), NFP RTT is 18us (own measurements). EMEM Ring one-way is 120ns.
+%?? Cziva~\parencite[pp.113]{DBLP:phd/ethos/Cziva18} discusses vNF times.
+%?? reference the inference times brought up in Taurus?
+Referring to \cref{fig:state-slip}, we take $t_2$ from \cref{tab:lats} for host and in-NIC processing times, and substitute $t_1+t_3$ for the state packet round-trip time to the decision site:
+\begin{description}
+	\item[In-NIC.] As described in \cref{sec:agent-environment-communication}, EMEM rings have a median one way delay cross-island of \SI{140}{\nano\second}, giving a \emph{median \SI{34.63}{\micro\second} end-to-end inference latency}.
+	\item[Dedicated Collector.] Offloads hosted in this manner will employ DPDK to maximise performance, giving one-way PCIe delays of \SIrange{0.9}{2.3}{\micro\second} for network packets.
+	A UDP packet carrying \num{20} elements of state in \approachshort{} is \SI{128}{\byte}, so costs \SI{1}{\micro\second} to forward, and the reply state-action pair invokes a slightly higher cost.
+	\emph{This gives an end-to-end inference latency of \SI{517.9}{\micro\second}}.
+	\item[vNF Offload.] \Textcite{DBLP:journals/cm/CzivaP17} show that more lightweight vNF frameworks like GNF~\parencite{DBLP:journals/cm/CzivaP17} and ClickOS~\parencite{DBLP:conf/nsdi/MartinsAROHBH14} add \SIrange{45}{55}{\micro\second} \emph{additional} latency above PCIe costs.
+	\emph{This gives an end-to-end inference latency of \SI{572.9}{\micro\second}}.
+\end{description}
+Using these estimates, in-NIC classical RL inference offers \SIlist{14.96;16.54}{$\times$} speedups in latency over collector and vNF deployments respectively (assuming no steering cost in either case).
+We also contrast these against deep RL policies on network tasks, which can take \SI{3}{\milli\second} to compute~\parencite{DBLP:journals/corr/abs-1910-04054}---2 orders of magnitude above \approachshort{} with identically sized (20-dim) input vectors.
+%In the name of fairness, we assume that rule installation uses same mechanism we recommend, but show how bad it can get? I.e. with rule installation costs etc.
 
@@
 All (bar one) of these affected \nth{99} percentile latencies by under \SI{0.38}{\percent} (at most \SI{78}{\nano\second}).
-This slight general degradation can be explained by increased pressure on the NFP's \emph{command push-pull} (CPP) bus, which is responsible for handling cross-island accesses to memory (particularly IMEM/EMEM) and other resources.
+This slight general degradation can be explained by increased pressure on the NFP's \emph{Command Push-Pull} (CPP) bus, which is responsible for handling cross-island accesses to memory (particularly IMEM/EMEM) and other resources.
 \approachshort{} places load on the CPP bus through use of its \inring{}/\outring{} rings for external communication and through last-tier policy accesses.
@@
 The most significant costs arise due to policy data (\SI{405}{\kibi\byte} capacity shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
-Thread local storage in CLS for compute/register spilling scales with the number of MEs required (requiring \SIlist{13.77;15.49}{\percent} per-ME for \Indfw{}/\Coopfw) due to precaching and the space needed to hold tile lists.
-Due to the initial policy cost this falls below the fair share at 3 MEs, but always allows for co-existence with other asynchronous dataplane programs.
+This causes a high upfront cost on per-island resources (CLS/CTM)---\approachshort{} leaves resources for other off-path dataplane applications, but is fairest from 3 cores onwards.
+
+%Thread local storage in CLS for compute/register spilling scales with the number of MEs required (requiring \SIlist{13.77;15.49}{\percent} per-ME for \Indfw{}/\Coopfw) due to precaching and the space needed to hold tile lists.
+%Due to the initial policy cost this falls below the fair share at 3 MEs, but always allows for co-existence with other asynchronous dataplane programs.
 
@@
 \fakepara{Deployability}
+Setup of an \approachshort{} uses two packet types: \emph{setup} which contains learning (hyper-)parameters and most aspects of a tile coded policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
 We found that setup and tiling packets take a mean \SIlist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing for an agent to be swapped between online and offline operation (or repurposed for another task/policy) painlessly.
+Online/offline swaps are useful when an agent should cease learning (i.e., when performance has converged to an acceptable value), or if a detected change in the environment means that more training is needed.
 Tiling packet processing was found to scale linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
@@
 Classical reinforcement learning has seen recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
-\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and local, per-flow state to monitor how flows respond to bandwidth changes and packet drop---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} with more allowance for congestion-unaware protocols.
+\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and local, per-flow state to monitor how flows respond to bandwidth changes and packet drop---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} (observing how flow behaviour reacts to rate limits increasing or decreasing) with more allowance for congestion-unaware protocols.
 Actions then move flows up or down in punishment levels according to a finite state machine.
@@
 \fakepara{Integrating \approachshort}
-Before each Dapper monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{OUT Ring} for any generated actions.
+Before each monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{OUT Ring} for any generated actions.
 As we note in \cref{sec:limitations}, these actions would be placed into a small local hash table and simultaneously exported to the attached controller to be inserted as P4 rules in batches.
@@
 An increased number of (attacking) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
-As we shall examine in \Cref{sec:evaluation}, \approachshort{} far exceeds the action throughput of a host based deployment could handle.
+As shown in \cref{sec:evaluation}, \approachshort{} far exceeds the action throughput of a host based deployment.
 The control plane is then free to use wildcards or specific matches to narrow down (or expand) the set of flows to be considered and controlled, though explicit scheduling techniques are still key in such adversarial environments.
@@
 The two compute models discussed above need not be homogeneously deployed.
-For instance, in a networked deployment a subset of \approachshort{} nodes could be \SI{32}{\bit} \coopfw{} agents, training online, while most other nodes ran \SI{8}{\bit} \indfw{} designs to meet throughput guarantees.
+For instance, in a networked deployment, a subset of \approachshort{} nodes could be \SI{32}{\bit} \coopfw{} agents, training online, while most other nodes ran \SI{8}{\bit} \indfw{} designs to meet throughput guarantees.
 The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
@@
 \section{Conclusion}
-
-?? Future work: allow scaling over multiple islands.
-?? Future work: NetFPGA it.
-
-?? Future directions for community: want to see more general-purpose async compute on-chip! Larger form-factors! More asynchrony!
-
-?? Future directions; transfer learning between classical methods into faster offline-only?
-
-?? Are there other ``old ways of doing things'' worth looking at?
+We have presented \emph{\approachshort{}}, bringing \emph{online reinforcement learning} to the dataplane.
+\approachshort{} has shown how classical RL techniques make online learning possible by simplifying update logic and enable parallel processing.
+These algorithms (and in-NIC deployment) enabled a \SIrange{15}{21}{$\times$} reduction in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput compared to host offloading.
+The deployment environment and asynchronous design were shown to eliminate PCIe and impose minimal impact on carried dataplane traffic.
+\approachshort{}'s designs scale with additional compute resources at deployment to improve upon both decision latency and throughput.
+Our throughput-optimal design, \Indfw{}, improves upon these metrics with \emph{just one worker}.
+
+In future, we aim to examine the relative performance of individual applications driven by \approachshort---both classical and deep RL-based---and how a NetFPGA implementation can offer further latency and throughput improvements.
+A promising avenue here would be to investigate constant transfer learning between a handful of online \approachshort{} agents and other high-throughput offline function approximators such as BNNs.
+%It is our hope that high port-density programmable devices will, in time, expose similar asynchronous, on-chip offloading capability to current SmartNIC devices.
+%We also are interested in `classical' approaches for other tasks which may have been overlooked in favour of more performant (yet complex) solutions, yet enable in-network deployment.
 

### 2021-06-07 paper/paper.tex
It's certainly progress.

@@
 % !BIB program = biber
-\documentclass[sigconf,natbib=false]{acmart}
+\documentclass[
+sigconf,natbib=false
+%,anonymous=true
+]{acmart}
+
+\usepackage[false]{anonymous-acm}
 
@@
 \usepackage{etoolbox}
-\usepackage[binary-units, per-mode=symbol]{siunitx}
+\usepackage[per-mode=symbol]{siunitx}
 \robustify\bfseries
@@
 %\robustify\uline
-\sisetup{detect-all, range-phrase=--, range-units=single, detect-weight=true,detect-inline-weight=math, table-format=1.3}
+\sisetup{detect-all, range-phrase=--, range-units=single, detect-weight=true, table-format=1.3}
 \DeclareSIUnit{\packet}{p}
@@
 
+%\vspace{-1em}
 \section{Introduction}
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/arch-with-p4}
+	\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/arch-with-p4}
 	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
@@
 \end{figure}
-Automatic optimisation, control, and defence of networks are at last becoming commonplace.
-Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing the automation of traffic/multimedia optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
-In RL methods, every change and its measured effects improve future decisions made by an agent.
-
-Parallel to this, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and programmable network hardware~\parencite{DBLP:journals/micro/ZilbermanACM14, netronome-smartnic, xilinx-alveo, barefoot-intel} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
-This has manifested in novel uses of programmable dataplane hardware to accelerate distributed tasks, fine-grained traffic measurement~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}, or per-packet processing.
-The latter is particularly attractive due to the costs of offloading to commodity host machines; input data must cross the PCIe bus several times, throughput requirements often impose the use of batching, and dedicated accelerators must again be contacted over PCIe~\parencite{DBLP:journals/corr/abs-2009-02353}.
-Additionally, in multi-tenant environments, IOMMU contention for both NICs and inference accelerators may impact tail latencies~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
-In response, we have seen non-neural ML techniques~\parencite{DBLP:conf/hotnets/XiongZ19}, neural networks~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome}, and hardware design proposals for CGRA-based map-reduce blocks~\parencite{DBLP:journals/corr/abs-2002-08987} \emph{directly in the dataplane}.
-These works have shown the value in in-network ML: high-throughput, low latency response to network changes, particularly when co-hosted with state extraction.
-
-However, the task of \emph{training} these models online \emph{and} in-network presents an exciting (and challenging) lacuna in the field.
-Online training is particularly important if an agent needs to react to unforeseen (failure) states, such as in DDoS/intrusion prevention.
-For context, \emph{Deep Neural Network} (DNN) training relies upon the backpropagation algorithm to compute gradients, and for sizeable amounts of data to be held in mini-batches for stable training.
-Moreover, DNN-backed RL must often be trained by vast amounts of offline simulation---in some cases compute-years worth~\parencite{DBLP:journals/corr/abs-1912-06680}---which can be counter-productive if the target problem evolves over time or is hard to model.
-These space and compute requirements, at present, mandate that training must occur in commodity hosts or dedicated hardware.
-Moreover, high state-action latencies \emph{caused} by offloading and expensive policy approximations add significant noise to the learning process~\parencite{DBLP:journals/firai/TravnikMSP18} and impact application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
-
-%?? Make a point of explaining why in-NIC... in general!
-%?? Offloading costs
-%?? High-throughput, low response latency to network changes.
-%?? Mention PCIe offloading costs, DNN costs, potential IOMMU contention in virtualisation-based envs over NIC accesses and GPU accesses~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
-
-%?? Using classical RL? cost now is offloading latency
-%?? Using DRL? Other costs.
-%?? Deep RL training infamous for taking  
-%
-%Offloading to a host machine has several drawbacks; the additional latency
-%
-%?? In-network ML has focussed on inference, NOT online training...
-%?? But they have innovations in data formats etc.
-%Moreover, necessary hardware, such as floating-point units, is unavailable in almost all cases.
-%
-%?? P4 caused an explosion of suitable measurement techniques in plane...
-%?? Classes of device exist today
-%?? So... ripe for putting online learning agents here!
-%
-%?? So we do this...
-%?? Classical RL techniques using the advances and observations of other datapath ML.
-%?? path-adjacent, on-chip, in the dataplane (\cref{fig:netro-arch})
-%?? Exploits legacy of many p4 devices to provide general purpose compute.
-%
-%?? Can I fit in the state-slippage shit here?
-
-The question we investigate is: can online RL be brought to the dataplane by returning to computationally simpler methods, to act on locally extracted state?
+Automatic network optimisation, control, and defence are at last becoming commonplace.
+Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing automatic traffic/media optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17,DBLP:conf/conext/GiladSGRS20}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
+In RL methods, every change and its effects improve future decisions made by an agent.
+
+%Parallel to this, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and programmable dataplane hardware~\parencite{DBLP:journals/micro/ZilbermanACM14, netronome-smartnic, xilinx-alveo, barefoot-intel} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
+%This has manifested in novel uses of programmable dataplane hardware to accelerate distributed tasks, fine-grained traffic measurement~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}, or per-packet processing.
+%The latter is particularly attractive due to the costs of offloading to commodity host machines; input data must cross the PCIe bus several times, throughput requirements often impose the use of batching, and dedicated accelerators must again be contacted over PCIe~\parencite{DBLP:journals/corr/abs-2009-02353}.
+%Additionally, in multi-tenant environments, IOMMU contention for both NICs and inference accelerators may impact tail latencies~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
+%In response, we have seen non-neural ML techniques~\parencite{DBLP:conf/hotnets/XiongZ19}, neural networks~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome}, and hardware design proposals for CGRA-based map-reduce blocks~\parencite{DBLP:journals/corr/abs-2002-08987} \emph{directly in the dataplane}.
+%These works have shown the value in in-network ML: high-throughput, low latency response to network changes, particularly when co-hosted with state extraction.
+
+%---
+
+Parallel to this, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} (PDP) hardware~\parencite{DBLP:journals/micro/ZilbermanACM14, netronome-smartnic, xilinx-alveo, barefoot-intel} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
+The promise of PDP hardware is that we can move the entire monitoring and analysis stack \emph{into the dataplane itself, and have it evolve to incorporate new approaches}.
+The P4 ecosystem already presents novel, openly-available, fine-grained traffic measurement techniques to the network such that they can be installed and controlled with ease~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
+In addition, P4's control plane makes it easy to select which flows or packets are monitored in a live network and potentially allow control over traffic at the decision site.
+As a result, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome,DBLP:journals/corr/abs-2002-08987} to take advantage of flow or per-packet state that cannot be efficiently processed or extracted at any other location in the network.
+These works have shown the value in in-network ML: high-throughput, low latency response to network changes.
+While they can exploit on-device state to provide reactive insight, the missing piece of the puzzle is the ability to learn and update these ML analyses online without deferring to another machine in the network.
+Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed by the community}.
+
+It's important to make this feasible, as offloading learning to commodity machines adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is often required due to the complexity of modern ML.
+For context, \emph{Deep Neural Network} (DNN) training relies upon the backpropagation algorithm to compute gradients, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs sizeable amounts of data to be held in mini-batches for stable training.
+These lead to high compute and storage costs, inducing capital expenditure for dedicated accelerators.
+Moreover, high latencies \emph{caused} by offloading and expensive inference harm the learning process~\parencite{DBLP:journals/firai/TravnikMSP18} and impact runtime application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
+If we can bring online learning \emph{to the dataplane}, then we can take advantage of rich, local state while minimising these latencies (and their impact on the learned policy).
+This would also make it easy to train and prototype agent designs which can learn \emph{as the environment evolves}, or when there is too little data to model and simulate a problem.
+
+%Moreover, DNN-backed RL must often be trained by vast amounts of offline simulation~\parencite{DBLP:journals/corr/abs-1912-06680}---in some cases compute-years worth~\parencite{DBLP:journals/corr/abs-1912-06680}---which can be counter-productive if the target problem evolves over time or is hard to model.
+
+%?? rich, local state
+
+%?? Why important to make feasible?
+%?? Why horrid that it hasn't been looked at?
+%This induces capital expenditure on 
+%?? We focus on models \& methods to enable this
+
+To enable \emph{online in-NIC learning}, we return to \emph{classical} reinforcement learning methods and models.
+In particular, we focus on tile-coding with one-step temporal-difference learning algorithms such as Sarsa.
+%These choices have important benefits for in-NIC execution.
+These functions do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays, and have simple update and inference logic.
+Tile-coding in particular admits many optimisations, being an embarrassingly parallel problem.
+Using fixed-point arithmetic, we solve the lack of floating-point support in PDP hardware \emph{and} enable new optimisations.
+%Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair, safeguarding the limited memory of the target devices.
+We also design our solution to operate as closely as possible to the P4 pipeline to use and learn from per-packet state, but outside of the main packet path to prevent packet stalls.
+
+%However, the task of \emph{training} these models online \emph{and} in-network presents an exciting (and challenging) lacuna in the field.
+%Online training is particularly important if an agent needs to react to unforeseen (failure) states, such as in DDoS/intrusion prevention.
+%For context, \emph{Deep Neural Network} (DNN) training relies upon the backpropagation algorithm to compute gradients, and for sizeable amounts of data to be held in mini-batches for stable training.
+%Moreover, DNN-backed RL must often be trained by vast amounts of offline simulation---in some cases compute-years worth~\parencite{DBLP:journals/corr/abs-1912-06680}---which can be counter-productive if the target problem evolves over time or is hard to model.
+%These space and compute requirements, at present, mandate that training must occur in commodity hosts or dedicated hardware.
+%Moreover, high state-action latencies \emph{caused} by offloading and expensive policy approximations add significant noise to the learning process~\parencite{DBLP:journals/firai/TravnikMSP18} and impact application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
+
+%?? Why programmable vs fixed dataplane?
+%?? Entire monitoring and analysis stack in one device
+%?? Why p4? Exploit existing works to get local state (and future works), Ease of runtime reconfig, control over traffic at the decision site.
+
+%?? The chance to exploit
+
+%?? Emphasise related work claims here!
+%?? We're doing novel stuff because ...
+
+%?? quick, reactive insight
+
+%?? without the capex on expensive, powerful hosts
+
+The question we investigate is: can online RL be brought to the dataplane by returning to these computationally simpler methods, to act on locally extracted state?
 Can it be made more efficient \emph{by dataplane hardware}?
-Through this work, we can comfortably answer ``yes'' on both counts.
-In particular, we exploit how SmartNIC devices often expose general purpose compute to provide path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
+Through this work, \approachshort{}, we can comfortably answer ``yes'' on both counts.
+In particular, we exploit how SmartNIC devices often expose general-purpose compute to provide path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
 As many of these devices have engineering and development histories which predate P4, general compute beyond P4's limits~\parencite{p4-psa} is surprisingly common.
 By executing on spare compute units, we prevent packet stalling and offer quick runtime reconfigurability.
-We look backwards to classical RL techniques (tile-coded policies, single-step RL algorithms), alongside the adaptations introduced by existing in-network ML (such as quantisation).
-Such adaptations unlock further improvements in parallel inference and update processing.
-In summary, this paper contributes:
+%We look backwards to classical RL techniques (tile-coded policies, single-step RL algorithms), alongside the adaptations introduced by existing in-network ML (such as quantisation).
+%Such adaptations unlock further improvements in parallel inference and update processing.
+This paper contributes:
 \begin{itemize}
@@
 	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm for tile-coded policies (\cref{alg:parsa}),
-	\item In-depth evaluation of how our approach affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
+	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
 	\item A potential integration of \approachshort{} with state-of-the-art PDP applications to perform fully in-NIC, automated DDoS mitigation (\cref{sec:potential-integrations}).
@@
 %?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
-As ethernet line rate moves beyond \SIlist{40;100}{\giga\bit\per\second}, packet processing deadlines grow tighter in tandem; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second}, giving a \SI{3.99}{\micro\second} deadline on Netronome SmartNICs (having \num{312} P4 pipelines).
+As ethernet moves beyond \SIlist{40;100}{\giga\bit\per\second}, packet processing deadlines grow tighter in tandem; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second}, giving a \SI{3.99}{\micro\second} deadline on Netronome SmartNICs (having \num{312} P4 pipelines).
 In-pipeline execution thus increases the risk of drops or stalled packet transmissions.
@@
 Many compatible devices have legacies long predating these developments. 
-For instance, many-core SOC-based Netronome SmartNIC~\parencite{netronome-smartnic}, and NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19} NICs allow virtually arbitrary programs to be specified and executed.
-This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
+For instance, many-core SOC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
+%This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
 
@@
 In networking, this allows for learning to respond to on-device state or to handle rapidly evolving problems.
-Generally these are agnostic to the policy approximation used, and classical single-step RL algorithms are simple---most of the compute work is occupied by the policy itself.
+RL algorithms are generally agnostic to the policy approximation used, and classical single-step methods are simple---most of the compute work is occupied by the policy itself.
 Consider the single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E} which requires only additions and multiplication to update and learn a policy online.
@@
 %Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
-Consider \cref{fig:state-slip}: the traditional RL formulation assumes that an agent receives a new state at fixed time intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
-These delays (and variance thereof) add noise to the state-action mapping being learned, which has a potent reduction to learning rate and final accuracy, even for simple grid world tasks according to \textcite{DBLP:journals/firai/TravnikMSP18}.
+Consider \cref{fig:state-slip}: the traditional RL formulation assumes that an agent receives new state at fixed time intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
+This adds noise to the state-action mapping being learned, which has harms learning rate and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
 Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
 This principle has influenced the design of real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, showing that asynchrony is necessary for high-speed applications.
-This often requires that state measurements are combined or coalesced~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
-`Stopping the world' in the algorithmic sense causes significant performance degradation, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control, or any time-sensitive control problems.
+This often requires that state measurements are fused~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
+`Stopping the world' causes significant performance degradation, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control, or any time-sensitive control problems.
 
@@
 	\begin{subfigure}{0.45\linewidth}
-		\resizebox{\linewidth}{!}{
+		\centering
+		\resizebox{0.9\linewidth}{!}{
 			\centering
@@
 	\begin{subfigure}{0.45\linewidth}
-		\resizebox{0.75\linewidth}{!}{
+		\centering
+		\resizebox{0.7\linewidth}{!}{
 			\centering
@@
 	\end{subfigure}
-	\caption{Asynchronous RL delays and state slippage (policy updates omitted).\label{fig:state-slip}}
+	\caption{Asynchronous RL delays and state slippage.\label{fig:state-slip}}
 \end{figure}
@@
 
-The solution we propose is to make use of the recent wave of programmable network devices to \emph{bring reinforcement learning to the dataplane}---referring again to \cref{fig:state-slip}, we would place place state measurement ($t_1$), low-cost decision-making processes ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
-In networks, actions are most likely to be installed on backbone switches, bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts.
-Ideally, these functions which comprise an RL agent would all be collocated on the same chip or device, but this is easier said than done. 
+This cost is what drives us to \emph{bring online reinforcement learning to the dataplane}---referring again to \cref{fig:state-slip}, we would place P4-based state measurement ($t_1$), simple policy approximation ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
+In networks, actions are most likely to be installed on backbone switches, bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts---all of which make in-NIC training a perfect fit.
+As such, our goal is to collocate \emph{all} functions which comprise an RL agent on the same chip or device.
 Both programmable devices and the network environment make this more difficult, as we'll examine in the sequel.
 
-?? Return to dimitris note: how is this important to final solution?
+%?? Return to dimitris note: how is this important to final solution?
 
@@
 $$\text{LMEM (ME)} < \text{CLS (Island)} < \text{CTM} < \text{IMEM (Chip)} < \text{EMEM}$$
-
-\subsection{Discussion}
-%?? relate update/execute operations here? i.e., policy execute needs only ALU add (and whatever is needed for tile code), train needs add+mul+shr. or in RL section? Should that come first?
-
-To achieve \emph{online in-NIC learning}, we return our focus to older \emph{classical} reinforcement learning methods.
-In particular, we suggest tile-coding, with one-step temporal-difference learning algorithms such as Sarsa or Q-learning.
-These choices have important benefits for in-NIC execution.
-Simpler function approximators do not require batches of inputs to learn in a more stable way, negating the additional memory required to store experience replays.
-Tile-coding in particular admits many optimisations, being an embarrassingly parallel problem.
-Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair, safeguarding the limited memory of the target devices.
-
-?? How to fit this through?
 
@@
 
-%We open-source our firmware design and control programs for the benefit of the community.\footnote{\url{https://github.com/FelixMcFelix/pdp-rl-paper}---currently private...}
-We open-source our firmware design and control programs for the benefit of the community.\footnote{Link redacted for double-blind submission}.
+We open-source our firmware design and control programs for the benefit of the community\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission}}.
 Moving beyond the Netronome platform, we describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.
@@
 
-The main interaction model is that platform-specific IPC (message rings) is used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
-These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
-Both input and output can occur on any other core of the device, i.e., as part of P4 \texttt{extern} plugins or a dedicated flow state measurement subsystem, while the P4 control plane itself provides granular control over which flows are monitored or affected.
-
-Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match-action tables and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
-Bit depth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
-?? Move to later?
+%The main interaction model is that platform-specific IPC (message rings) is used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
+%These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
+%Both input and output can occur on any other core of the device, i.e., as part of P4 \texttt{extern} plugins or a dedicated flow state measurement subsystem, while the P4 control plane itself provides granular control over which flows are monitored or affected.
+%
+%Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match-action tables and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
+%Bit depth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
 
@@
 If updates are enabled, \approachshort{} then checks an internal hashmap for a previous state-action pair matching the current instruction source, and if found then a policy update is performed.
-Updates are computed using \emph{single-step semi-gradient Sarsa} (\cref{eqn:sg-sarsa}), though modification to support other single-step methods would be trivial.
+Updates are computed using \emph{single-step semi-gradient Sarsa}~\cite[pp. \numrange{217}{221}]{RL2E}, though modification to support other single-step methods would be trivial.
 The new state- or tiles-action pair is then written into storage.
@@
 The key insight of this mechanism is that on-chip reward/state signals enjoy first-class support in much the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-NIC/chip asynchronous processes or the controller.
+As such, \approachshort{} can receive from P4 \texttt{extern}s or another dedicated off-path flow state measurement application.
 
@@
 
+%The main interaction model is that platform-specific IPC (message rings) is used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
+%These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
+%Both input and output can occur on any other core of the device, i.e., as part of P4 \texttt{extern} plugins or a dedicated flow state measurement subsystem, while the P4 control plane itself provides granular control over which flows are monitored or affected.
+
 \subsection{Intra-Agent Communication}\label{sec:intra-agent-communication}
-Even with embarrassingly parallel problems, optimising for latency requires meticulous care in how work is passed out and aggregated.
+Even with parallel problems such as \emph{ParSa}, optimising for latency requires meticulous care in how work is passed out and aggregated.
 This is truer still when moving from the moderately fine-grained, efficient control of classical RL methods ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
-Ordinarily, the serialisation/marshalling of requests, responses, and safe shared data access can incur significant extra overheads.
+Ordinarily, the marshalling of requests, responses, and safe shared data access can incur significant extra overheads.
 On-chip execution and the nature of action preference computation allow us to use lockless atomic aggregation, removing the overheads of explicit messaging/packetisation.
@@
 
+%?? reduce below and make take-homes more generic if possible
+
 Our implementation exploits the locality of threads, cores, and their parent islands in the NFP architecture.
 Policy compute/update tasks and configuration updates are passed between cores using these direct \emph{next neighbour} registers, signalling all child threads in response.
+Such on-chip local signals cost us just $\sim$\SI{20}{\nano\second} latency per relayed message.
 Each task performs atomic adds to a shared preference list, and atomically increments an acknowledgement counter to be periodically checked by the master thread, implementing the wait-free \emph{ParSa} algorithm we introduce (\cref{alg:parsa}).
-
-In earlier designs we had experimented with bounded buffers as in \cref{sec:agent-environment-communication} for this, modified to be located solely in CLS memory, dedicating the master thread to result aggregation.
-We found that this created a performance bottleneck at this final stage, causing significant head-of-line blocking for each of the workers.
-Similarly, our next neighbour work notification scheme ($\sim$\SI{20}{\nano\second} latency per relayed message) was examined against \emph{reflector} and \emph{work queue} IPC mechanisms (\SIlist{58;126}{\nano\second} per messaged core).
-These slower messages can be used in theory to skip ahead into longer ME chains.
+This is essential for aggregation compared to the use of bounded message buffers, which caused significant head-of-line blocking.
+
+%On-chip local messages cost us just $\sim$\SI{20}{\nano\second} latency per relayed message
+%
+%In earlier designs we had experimented with bounded buffers as in \cref{sec:agent-environment-communication} for this, modified to be located solely in CLS memory, dedicating the master thread to result aggregation.
+%We found that this created a performance bottleneck at this final stage, causing significant head-of-line blocking for each of the workers.
+%Similarly, our next neighbour work notification scheme ($\sim$\SI{20}{\nano\second} latency per relayed message) was examined against \emph{reflector} and \emph{work queue} IPC mechanisms (\SIlist{58;126}{\nano\second} per messaged core).
+%These slower messages can be used in theory to skip ahead into longer ME chains.
+
+%Our implementation exploits the locality of threads, cores, and their parent islands in the NFP architecture.
+%Policy compute/update tasks and configuration updates are passed between cores using these direct \emph{next neighbour} registers, signalling all child threads in response.
+%Each task performs atomic adds to a shared preference list, and atomically increments an acknowledgement counter to be periodically checked by the master thread, implementing the wait-free \emph{ParSa} algorithm we introduce (\cref{alg:parsa}).
+
+%In earlier designs we had experimented with bounded buffers as in \cref{sec:agent-environment-communication} for this, modified to be located solely in CLS memory, dedicating the master thread to result aggregation.
+%We found that this created a performance bottleneck at this final stage, causing significant head-of-line blocking for each of the workers.
+%Similarly, our next neighbour work notification scheme ($\sim$\SI{20}{\nano\second} latency per relayed message) was examined against \emph{reflector} and \emph{work queue} IPC mechanisms (\SIlist{58;126}{\nano\second} per messaged core).
+%These slower messages can be used in theory to skip ahead into longer ME chains.
 
@@
 	\tcc{Given message passing mechanisms \emph{scatter} and \emph{recv}, quantised arithmetic functions $Q_\mathit{mul}$ and TileCode, and omitting schedule/config/precache updates.}
+	\tcc{\emph{cfg}.$\alpha$, \emph{cfg}.$\gamma$ are hyperparameters affecting the significance of each update and the degree of forward-planning, respectively.}
 
@@
 	\Const \emph{cfg, policy} = /* ... */\;
+
 	\Let \emph{values}: [AtomicI32; cfg.n\_actions] = \{0\}\;
@@
 \subsection{Reconfigurability}
-\approachshort{} allows for policy dimensions, size, action count, tiling strategies, and algorithmic learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) to be chosen and changed at runtime using at most two control packets.
+\approachshort{} allows for policy design and algorithmic learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) to be chosen and changed at runtime using at most two control packets.
+For instance, the former would be useful at the end of learning, moving from online to offline or when trying to use data from the same vantage point to train a policy for another problem afterwards.
+The latter is useful when an online agent must become more (or less) adaptive to new data (i.e., after a changepoint in traffic is detected).
 This extends to policy data, which may be imported from an offline pre-trained model via many such packets, and exported via the NFP binary support package on the host machine.
 Some aspects must be chosen at compile time; bit depth, work allocation strategy, and maximum policy/tiling/state sizes---these govern core operation or allocated memory regions.
+Choosing a bit depth of \qtylist[list-pair-separator = { or }]{16;8}{\bit} halves or quarters policy memory costs, allowing more complex problems to be modelled using more dimensions or fine-grained policies.
 
@@
 
+%Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match-action tables and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
+%Bit depth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
+
 \subsection{Work Allocation}\label{sec:work-allocation}
-Due to the lack of dynamic memory allocation and to simplify action-value lookups, policies cannot be effectively stored sparsely.
+Due to the lack of dynamic memory allocation on programmable dataplane hardware, and to simplify action-value lookups, policies cannot be effectively stored sparsely.
 As a consequence, tiling space requirements scale exponentially with dimension count, and so higher-dimension tilings must be placed in memory regions large enough to allocate space for them.
-As a result, tilings are split across memory regions, giving different access and compute costs to different tasks---ranging over CLS, CTM and IMEM in increasing latency.
+As a result, tilings are split across memory regions, giving different access and compute costs to different tasks---split by CLS/CTM/IMEM on the NFP, or e.g., BRAM/SRAM/DDR3 on NetFPGA.
 
@@
 Each work item is a separate \emph{tiling} over a number of dimensions, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
-The cost of any individual work item (according to its dimension count and memory location) was empirically measured offline, and we weight the total cost per ME to account for the number of minion threads available.
+The cost of any individual work item (according to its dimension count and memory location) was empirically measured offline, and we weigh the total cost per ME to account for the number of minion threads available.
 This weighting specifically accounts for the presence of the controller thread on the first ME.
@@
 
-\subsection{Variable Quantisation Bit Depth}
-At compile-time, \approachshort{} can be configured to use \SIlist[list-final-separator = { or }]{32;16;8}{\bit} values in input states and its internal policies.
-Naturally, smaller bit depths reduce the storage required for policies, tiling data, and stored state-action pairs---allowing more complex problems to be modelled using more dimensions or fine-grained policies.
-Through the lens of computational efficiency, reduced bit depth should allows action preference lists to be read using fewer I/O operations (as more values may fit into a single machine word).
-
-We had investigated bit-stuffing several such values into a single word for our atomic writeback mechanism (as the platform offers both \SIlist{32;64
-}{\bit} atomic addition).
-This is analogous to SIMD---through clever use of padding bits---but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
-We investigate further performance implications of changing bit depth in the sequel.
+%\subsection{Variable Quantisation Bit Depth}
+%At compile-time, \approachshort{} can be configured to use \SIlist[list-final-separator = { or }]{32;16;8}{\bit} values in input states and its internal policies.
+%Naturally, smaller bit depths reduce the storage required for policies, tiling data, and stored state-action pairs---allowing more complex problems to be modelled using more dimensions or fine-grained policies.
+%Through the lens of computational efficiency, reduced bit depth should allows action preference lists to be read using fewer I/O operations (as more values may fit into a single machine word).
+%
+%We had investigated bit-stuffing several such values into a single word for our atomic writeback mechanism (as the platform offers both \SIlist{32;64
+%}{\bit} atomic addition).
+%This is analogous to SIMD---through clever use of padding bits---but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
+%We investigate further performance implications of changing bit depth in the sequel.
 
@@
 We instead suggest on NFP chips that \texttt{extern}s or datapath stages which apply actions to packet processing should maintain a small local store of state-action pairs, and periodically send these back to the controller for batch rule installation.
-This ensures that the vast majority of installed rules enjoy accelerated performance, while preventing rule installation delay on the newest decisions.
+This ensures that the vast majority of installed rules exhibit accelerated performance, while preventing rule installation delay on the newest decisions.
 Platforms such as Intel Tofino greatly simplify this, with Tofino Native Architecture intrinsics such as \emph{Action Profiles/Selects} allowing a P4 action to be chosen based on a register value (i.e., an RL action).
@@
 \subsection{Targeting other device classes}
-While this design caters primarily to Netronome's SmartNIC devices, many aspects of this design have clear analogues in NICs of similar form-factor, such as their close NetFPGA SUME cousins.
+While this design caters primarily to Netronome's SmartNIC devices, many aspects of this design have clear analogues in NICs of similar form-factor, such as the NetFPGA SUME platform.
 For instance, the availability of other cores can be handled by designing additional off-path functional units, much like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
@@
 \begin{description}
-	\item[\emph{MidServer}] Intel Xeon Bronze 3204 CPUs ($6\times$\SI{1.9}{\giga\hertz}),
-	\item[\emph{HighServer}] Intel Xeon Silver 4208 CPU ($8\times$\SI{2.1}{\giga\hertz}),
-	\item[\emph{Collector}] Intel Core i7-6700K ($4\times$\SI{4.2}{\giga\hertz}).
+	\item[\emph{MidServer}] Intel Xeon Bronze 3204 CPUs (\qtyproduct[product-units=single]{6 x 1.9}{\giga\hertz}),
+	\item[\emph{HighServer}] Intel Xeon Silver 4208 CPU (\qtyproduct[product-units=single]{8 x 2.1}{\giga\hertz}),
+	\item[\emph{Collector}] Intel Core i7-6700K (\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}).
 \end{description}
-Evaluation of \approachshort{} was performed on server blades (\emph{Mid/HighServer}), each hosting a single Netronome Agilio LX 1x40GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
+\approachshort{} was evaluated on server blades (\emph{Mid/HighServer}), each hosting a single Netronome Agilio LX \numproduct{1 x 40}GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
 These servers ran Ubuntu 18.04.5 LTS (4.15.0-140-generic).
@@
 
-Policy sizes are set to those matching a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dimension state vectors, a bias tile and 16 full tiling sets (7$\times$1-dim, 8$\times$2-dim, 1$\times$4-dim), 8 tilings per dimension set, 6 tiles per dimension, choosing between 10 actions.
+Policy sizes are set to those matching a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dimension state vectors, a bias tile and 16 full tiling sets (\numproduct{7x1}-dim, \numproduct{8x2}-dim, \numproduct{1x4}-dim), 8 tilings per dimension set, 6 tiles per dimension, choosing between 10 actions.
 For context, such input state would be various aspects of per-flow state (e.g., IATs, up/down rate) which are combined with other state such as the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim)
@@
 \fakepara{Raw inference and learning performance}
-We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs.
-We compare these against floating point (numpy-based) implementations running on a commodity host machine.
+We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs against floating point (numpy-based) implementations running on a commodity host machine.
 This allows us to demonstrate the performance differences between the \indfw{} and \coopfw{} configurations, particularly in how \indfw's required policy locks impact throughput.
@@
 We compare online learning performance (input states produce an output, and then update the policy with appropriate locking) with offline (input states \emph{only} produce an output) in these cases.
-Online performance marks the number of decisions that can be made per second (and associated latency) when still training a policy.
+Online performance marks the number of decisions that can be made per second (and associated latency) when training a policy.
 Offline performance is crucial for pushing a trained, known-good policy to agents in the network with an expected higher raw decision throughput.
@@
 %?? Also our simple, ahead-of-time, work scheduler.
-We verify that our heuristic, runtime work scheduler (which we term \emph{Balanced}) makes meaningful use of the explicit memory hierarchy and cost of each work item.
-We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} baseline allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ tilings given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
+We verify that our heuristic, runtime work scheduler (termed \emph{Balanced}) makes meaningful use of the explicit memory hierarchy and cost of each work item.
+We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ tilings given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
 We use maximum-size, \SI{32}{\bit} policies as described above.
@@
 We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, placing an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source via the default firmware.
-We perform throughput/loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements (sampling at \SI{2000}{\hertz} for \num{10}$\times$\SI{10}{\second}).
-These choices allow maximum throughput in the former case (relying solely on NIC counters for loss counting).
-In the latter case we aim to minimise host resource contention so as to observe exact latency measurements, have a high enough sample count to detect subtle (aggregate) latency effects, and eliminate \emph{host} receive drops.
+We perform throughput/loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements (sampling at \SI{2000}{\hertz} for \qtyproduct[product-units=single]{10 x 10}{\second}).
+This provides maximum throughput in the former case (relying solely on NIC counters for loss counting).
+In the latter case this minimises host resource contention to observe exact latency measurements, have a high enough sample count to detect subtle (aggregate) latency effects, and eliminate \emph{host} receive drops.
 DPDK was setup using four \SI{1}{\gibi\byte} hugepages.
@@
 We report resource use for \SI{32}{\bit} \Indfw{} and \Coopfw{} agents with maximum policy sizes as above, with hash-tables sized to \num{4096} state-action pairs and \num{16} separate reward values.
-
-?? How do I safely say: these are arbitrary?
+This captures the relative cardinality of network RL traces to rewards, i.e., many input flows will map to one or few reward values (i.e., DDoS attack size estimation per egress-AS, queue occupancy in the case of AQM per output port).
 
@@
 As a result, these bit depths give a \SIrange{2}{4}{$\times$} reduction in memory needed to store a policy (enabling more fine-grained or more complex policies), in exchange for slightly worse latency and throughput.
+We investigated bit-stuffing several such values into a single word for our atomic writeback mechanism (as the platform offers both \SIlist{32;64
+}{\bit} atomic addition).
+This is analogous to SIMD---through clever use of padding bits---but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
 
@@
 All (bar one) of these affected \nth{99} percentile latencies by under \SI{0.38}{\percent} (at most \SI{78}{\nano\second}).
-This slight general degradation can be explained by increased pressure on the NFP's \emph{Command Push-Pull} (CPP) bus, which is responsible for handling cross-island accesses to memory (particularly IMEM/EMEM) and other resources.
+This slight degradation can be explained by increased pressure on the NFP's \emph{Command Push-Pull} (CPP) bus, which is responsible for handling cross-island accesses to memory (particularly IMEM/EMEM) and other resources.
 \approachshort{} places load on the CPP bus through use of its \inring{}/\outring{} rings for external communication and through last-tier policy accesses.
-This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---the Netronome P4 toolchain segments packets, storing metadata (e.g., MAC prepend data) and the first bytes of a packet in a $\SI{256}{\byte}$ CTM block and parking the bodies of processed packets in EMEM.
+This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---Netronome's P4 toolchain segments packets, storing metadata (e.g., MAC prepend data) and the first bytes of a packet in a $\SI{256}{\byte}$ CTM block and parking their payloads in EMEM.
 \SI{256}{\byte} packets overshoot this due to metadata, causing small I/O accesses at a high rate for packets sized around this cutoff point.
@@
 We observe negligible costs on shared bulk EMEM storage ($\sim$\SI{4}{\mebi\byte}), including storage space for a hash-table of \num{4096} state-action pairs and \num{16} reward values.
-The most significant costs arise due to policy data (\SI{405}{\kibi\byte} capacity shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
+The most significant costs arise due to policy data (\SI{405}{\kibi\byte} shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
 This causes a high upfront cost on per-island resources (CLS/CTM)---\approachshort{} leaves resources for other off-path dataplane applications, but is fairest from 3 cores onwards.
@@
 \fakepara{Deployability}
-Setup of an \approachshort{} uses two packet types: \emph{setup} which contains learning (hyper-)parameters and most aspects of a tile coded policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
-We found that setup and tiling packets take a mean \SIlist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing for an agent to be swapped between online and offline operation (or repurposed for another task/policy) painlessly.
+Setup of \approachshort{} uses two packet types: \emph{setup} which contains learning (hyper-)parameters and most aspects of a tile coded policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
+We found that setup and tiling packets take a mean \qtylist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing for an agent to be swapped between online and offline operation (or repurposed for another task/policy) painlessly.
 Online/offline swaps are useful when an agent should cease learning (i.e., when performance has converged to an acceptable value), or if a detected change in the environment means that more training is needed.
-Tiling packet processing was found to scale linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
+Tiling packet processing scales linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
 Online-offline swaps for \Coopfw{} exhibit similar cost, however the need for explicit ahead-of-time heuristic scheduling means that policy/tiling \emph{structure} changes (including the first complete setup) take \SI{422.63}{\micro\second} for the full-size policy described above.
@@
 Additionally, we discuss how network administrators may combine offline (\Indfw) and online (\Coopfw) agents within a network to achieve online learning while maximising action throughput in the majority of the network.
+
+?? How do we move this intuition to the start? ELI5 as dimitris says, to get the full end-to-end pipeline across to the reader.
 
@@
 A recent line of research in the community has been to investigate \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate, low latency packet classification.
-\emph{BaNaNa SPLIT} shows the initial steps of this trend investigating their use as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} in an effort to reduce latency imposed by batching and bulk transfer to the GPU.
+\emph{BaNaNa SPLIT} shows the initial steps of this trend as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} in an effort to reduce latency imposed by batching and bulk transfer to the GPU.
 However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}, with the hypothesis put forward that networking hardware would be a better fit for some parts of the inference pipeline.
@@
 \fakepara{In-network ML \emph{acceleration}}
-Optimisation of distributed neural network training is an area where in-network compute has been key in the case of general NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18} and RL-specific procedures~\parencite{DBLP:conf/isca/LiLYCSH19} using NetFPGAs to introduce floating point adders to process dedicated packet classes.
+Optimisation of distributed neural network training is an area where in-network compute has been key for general NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18} and RL-specific procedures~\parencite{DBLP:conf/isca/LiLYCSH19} using NetFPGAs to introduce floating point adders to process dedicated packet classes.
 Distributed training is a partition-aggregate workload, where gradient vectors are sent back to a controller to update the `true' model---in-NIC processing allows these to be aggregated \emph{in-network}, overcoming incast behaviour and host bottlenecks.
@@
 Their approach, \emph{NeuroCuts}, is able to produce time/space efficient classifiers for use in constrained environments (e.g., network hardware) beyond existing heuristic solutions.
-
 Deep RL techniques have been used for QUIC congestion control optimisation~\parencite{DBLP:journals/corr/abs-1910-04054}.
@@
 	They would additionally like to thank Stefanos Sagkriotis, Mircea Iordache-\c{S}ic\u{a}, and Haruna Adoga for their comments and feedback.
-	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grant~\grantnum{gs-epsrc}{EP/N509668/1}].
+	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grants~\grantnum{gs-epsrc}{EP/N509668/1}, \grantnum{gs-epsrc}{EP/N033957/1}].
 \end{acks}

### 2021-06-08 paper/paper.tex
Still certainly progress...

@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/arch-with-p4}
+	\includegraphics[keepaspectratio, width=0.75\linewidth]{figures/arch-with-p4}
 	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
@@
 
-Parallel to this, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} (PDP) hardware~\parencite{DBLP:journals/micro/ZilbermanACM14, netronome-smartnic, xilinx-alveo, barefoot-intel} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
+%?? Do we want concrete examples of things `pushed down the stack'?
+%
+%?? Consider a DDoS prevention system...
+%?? in the face of new data, Able to start learning again in microseconds
+
+In parallel, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} (PDP) hardware~\parencite{DBLP:journals/micro/ZilbermanACM14, netronome-smartnic, xilinx-alveo, barefoot-intel} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
 The promise of PDP hardware is that we can move the entire monitoring and analysis stack \emph{into the dataplane itself, and have it evolve to incorporate new approaches}.
-The P4 ecosystem already presents novel, openly-available, fine-grained traffic measurement techniques to the network such that they can be installed and controlled with ease~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
+Consider a DDoS mitigation system which could react to dataplane-specific, per-packet state in microseconds, and \emph{start learning again just as quickly} in the face of new data or behaviours.
+The P4 ecosystem already presents novel, openly-available, fine-grained traffic measurement techniques that can be installed and controlled with ease~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
 In addition, P4's control plane makes it easy to select which flows or packets are monitored in a live network and potentially allow control over traffic at the decision site.
 As a result, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome,DBLP:journals/corr/abs-2002-08987} to take advantage of flow or per-packet state that cannot be efficiently processed or extracted at any other location in the network.
-These works have shown the value in in-network ML: high-throughput, low latency response to network changes.
-While they can exploit on-device state to provide reactive insight, the missing piece of the puzzle is the ability to learn and update these ML analyses online without deferring to another machine in the network.
+These works have shown the value of in-network ML: high-throughput, low latency response to network changes.
+While they can exploit on-device state to provide reactive insight, the missing piece of the puzzle is learning and updating these ML analyses online without deferring to another machine in the network.
 Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed by the community}.
 
-It's important to make this feasible, as offloading learning to commodity machines adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is often required due to the complexity of modern ML.
-For context, \emph{Deep Neural Network} (DNN) training relies upon the backpropagation algorithm to compute gradients, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs sizeable amounts of data to be held in mini-batches for stable training.
+It is important to make this feasible, as offloading learning to commodity machines adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
+For context, \emph{Deep Neural Network} (DNN) training relies on the backpropagation algorithm to compute gradients, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many mini-batches of data for stable training.
 These lead to high compute and storage costs, inducing capital expenditure for dedicated accelerators.
@@
 
-To enable \emph{online in-NIC learning}, we return to \emph{classical} reinforcement learning methods and models.
+To enable \emph{online in-NIC learning}, we return to \emph{classical} RL methods and models.
 In particular, we focus on tile-coding with one-step temporal-difference learning algorithms such as Sarsa.
@@
 Using fixed-point arithmetic, we solve the lack of floating-point support in PDP hardware \emph{and} enable new optimisations.
+Moreover, the P4 dataplane can offer runtime control over which flows/packets are monitored.
 %Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair, safeguarding the limited memory of the target devices.
@@
 \begin{itemize}
-	\item An analysis of why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility (\cref{sec:motivation}),
-	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands (\cref{sec:design}),
-	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm for tile-coded policies (\cref{alg:parsa}),
-	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes, and improves upon explicit offloading to produce a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
-	\item A potential integration of \approachshort{} with state-of-the-art PDP applications to perform fully in-NIC, automated DDoS mitigation (\cref{sec:potential-integrations}).
+	\item An analysis of why in-NIC RL is needed and best-placed to interact with the network, made feasible by classical RL methods and quantisation (\cref{sec:motivation}),
+	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands of network traffic analysis (\cref{sec:design}),
+	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm to accelerate tile-coded policy inference and updates (\cref{alg:parsa}),
+	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
+	\item A demonstration of how \approachshort{} would integrate with state-of-the-art PDP applications to perform fully in-NIC, fast, automated DDoS mitigation (\cref{sec:potential-integrations}).
 \end{itemize}
@@
 Complex \emph{Deep Neural Networks} (DNNs), particularly when used as the basis for an RL agent's policy, require vast amounts (e.g., in the worst case compute-years~\parencite{DBLP:journals/corr/abs-1912-06680}) of experience to converge on an accurate parameter set, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
-Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the storage and computational costs of gradient computation via backpropagation remain too significant on embedded hardware.
+Although there has been a vested interest in efficiently \emph{executing} complex function approximators like DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the storage and computational costs of backpropagation remain too significant on embedded hardware.
 
@@
 
-We then return again to what classical RL methods can offer us; in particular, the simple linear function approximation given by \emph{tile coding} \cite[pp.\ \numrange{217}{221}]{RL2E}.
+We return to what classical RL methods can offer us; in particular, the simple linear function approximation given by \emph{tile coding} \cite[pp.\ \numrange{217}{221}]{RL2E}.
 The idea is simple: a policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
@@
 A state vector produces a single `hit' in each tiling, all of which then correspond to a list of action preference lists.
-Inference (choosing an action) is then simply summing over all such lists to obtian a final action preference list, selecting the maximum ($a$).
-Updating at the next timestep uses this same list of lists, adjusting the value of the last selected action using an adjustment value $\delta_t$ computed through Sarsa.
-Crucially, this internal representation ensures that there are no data dependencies between any tile hit calculations for an input state vector.
+Inference (choosing an action) is then simply summing over all such lists to obtian a final action preference list, selecting the maximum.
+Updating at the next timestep uses this same list of lists, adjusting the value of the last selected action using a value $\delta_t$ computed through Sarsa.
+Crucially, this internal representation ensures that there are no data dependencies between any tile calculations for an input.
 This has important benefits for parallelising the key parts of an RL algorithm; both action selection and policy updates are map-reduce problems.
@@
 Updating the policy (with a new $\delta_t$ value) has no aggregate step, and as tile indices map to disjoint memory regions no locks are required.
-As we show in \cref{sec:design}, it is possible to make the inference aggregate step wait-free.
+%As we show in \cref{sec:design}, it is possible to make the inference aggregate step wait-free.
 
@@
 \subsection{Numerical representation for embedded ML}
-A key feature universally lacking from programmable network hardware is floating-point arithmetic support.
-Luckily, the task of bringing machine learning models to resource-constrained environments without these capabilities is well-studied.
+A key feature universally lacking from PDP hardware is floating-point arithmetic support.
+Luckily, the task of bringing ML models to resource-constrained environments without these capabilities is well-studied.
 Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
@@
 To make RL workloads feasible under such constraints, we propose the use of quantised, fixed-point representations such as $Qm.n$ (i.e., $m$\si{bit} signed integers with an $n$\si{bit} fractional part), which allow us to evaluate and update policies using only integer arithmetic.
-This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation ($t_2$, \cref{fig:state-slip}).
-We note that our later parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is made possible only by the use of quantisation to integers, as atomic operations on floating point numbers typically do not exist outside of graphics- or tensor-processing hardware.
-
-When combined with the above P4-driven techniques for in-network flow/device measurement, we at last have the mechanisms to collocate the key processes of an RL agent.
-Moreover, the base P4 dataplane can be used to simplify parsing logic and offer runtime control over which flows/packets are monitored, alongside other packet actions.
-This again fits our goals of integrating RL techniques directly within bump-in-the-wire installations and at end-hosts.
+This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation.
+We note that our later parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is possible only due to the use of $Qm.n$, as atomic operations on floating point numbers typically do not exist outside of graphics-processing hardware.
+
+%When combined with the above P4-driven techniques for in-network flow/device measurement, we at last have the mechanisms to collocate the key processes of an RL agent.
+%Moreover, the base P4 dataplane can be used to simplify parsing logic and offer runtime control over which flows/packets are monitored, alongside other packet actions.
+%This again fits our goals of integrating RL techniques directly within bump-in-the-wire installations and at end-hosts.
 
@@
 
-\subsection{Netronome Platform Fundamentals}
-As we implement this work on Netronome SmartNICs (thus the NFP chip architecture), it is necessary to explain its basics.
-NFP chips are \emph{system-on-a-chip} (SoC) devices, and achieve scalable packet processing through sheer parallelism.
-Aside from an ARM management chip and application specific functional units, most of the chip is composed of \emph{microengines} (MEs), grouped into \emph{islands} of 4 or 12 MEs.
-All 12-ME islands are used by a default P4 pipeline.
-Each ME has \numrange{4}{8} \emph{contexts} (hardware threads) which share a code store and large register file, offering zero-cost context switches triggered by I/O.
-Contexts and MEs may send one another numbered signals, and MEs have a small \emph{next-neighbour} register file for passing values in one direction to the next ME on the same island.
-MEs run a proprietary instruction set, compiled to via a \emph{(Micro-)C} compiler.
-Beyond registers, the platform implements an explicit memory hierarchy scaling in size, location, and access cost as below:
-$$\text{LMEM (ME)} < \text{CLS (Island)} < \text{CTM} < \text{IMEM (Chip)} < \text{EMEM}$$
+%\subsection{Netronome Platform Fundamentals}
+%As we implement this work on Netronome SmartNICs (thus the NFP chip architecture), it is necessary to explain its basics.
+%NFP chips are \emph{system-on-a-chip} (SoC) devices, and achieve scalable packet processing through sheer parallelism.
+%Aside from an ARM management chip and application specific functional units, most of the chip is composed of \emph{microengines} (MEs), grouped into \emph{islands} of 4 or 12 MEs.
+%All 12-ME islands are used by a default P4 pipeline.
+%Each ME has \numrange{4}{8} \emph{contexts} (hardware threads) which share a code store and large register file, offering zero-cost context switches triggered by I/O.
+%Contexts and MEs may send one another numbered signals, and MEs have a small \emph{next-neighbour} register file for passing values in one direction to the next ME on the same island.
+%MEs run a proprietary instruction set, compiled to via a \emph{(Micro-)C} compiler.
+%Beyond registers, the platform implements an explicit memory hierarchy scaling in size, location, and access cost as below:
+%$$\text{LMEM (ME)} < \text{CLS (Island)} < \text{CTM} < \text{IMEM (Chip)} < \text{EMEM}$$
 
@@
 		\centering
-		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/ind}
+		\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/ind}
 		\caption{\Indfw{} (offline throughput-optimal).
@@
 		\centering
-		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/coop}
+		\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/coop}
 		\caption{\Coopfw{} (online-optimal).
@@
 \Cref{fig:netro-arch,fig:single-and-parallel} outline our design and implementation on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
-?? Explain \cref{fig:netro-arch},
-?? Wish-based justification for P4.
-?? \Cref{fig:single-and-parallel} is explained in \cref{sec:action-and-update-computation}
-
-We open-source our firmware design and control programs for the benefit of the community\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission}}.
+We explain relevant NFP architectural details in \cref{sec:netronome-platform-fundamentals}.
+\approachshort{} communicates with the packet pipeline of a P4 dataplane via \texttt{extern} plugins using \inring{} (state, configuration) and \outring{} (action) messages (\cref{fig:netro-arch}).
+Internally, \approachshort{} either has all its cores act independently (\cref{fig:single-and-parallel:single}) or cooperate to solve each task (\cref{fig:single-and-parallel:parallel})---with different latency-throughput benefits (\cref{sec:action-and-update-computation}).
+We open-source our firmware design and control programs for the benefit of the community.\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission.}}
 Moving beyond the Netronome platform, we describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.
@@
 To protect traffic throughput and allow for effective deployment in as many environments as possible, \approachshort{} places RL execution on-chip, \emph{but off the main packet path}, communicating and running parallel to the main P4 dataplane.
-As shown in \cref{fig:netro-arch}, this asynchrony allows coexistence with P4 programs, and imposes minimal/no impact on carried traffic for both bump-in-the-wire deployments and at end-points.
+As shown in \cref{fig:netro-arch}, this asynchrony allows coexistence with P4 programs, and imposes minimal impact on carried traffic for both bump-in-the-wire deployments and at end-points.
 For instance, in the default deployment of a P4 packet processing pipeline on Netronome NFP chips several of these cores go unused (as does spare area on an FPGA design), making this paradigm possible.
@@
 \approachshort{} applies the insights of \textcite{DBLP:journals/firai/TravnikMSP18} to minimise action latency; an action is computed, sent out into the environment, and only then is the underlying policy updated.
-Using one of the below strategies chosen at compile time, a state vector is tile coded, converted into action probabilities, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
+Using one of the below strategies chosen at compile time, a state vector is tile coded, converted into action probabilities, and an action is chosen.
 This is then written out to the environment as in \cref{sec:agent-environment-communication}.
-If updates are enabled, \approachshort{} then checks an internal hashmap for a previous state-action pair matching the current instruction source, and if found then a policy update is performed.
+If online learning is enabled, \approachshort{} then checks an internal hashmap for a previous state-action pair matching the current instruction source, and if found then the policy is updated.
 Updates are computed using \emph{single-step semi-gradient Sarsa}~\cite[pp. \numrange{217}{221}]{RL2E}, though modification to support other single-step methods would be trivial.
@@
 	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel,alg:parsa})] Threads cooperate on processing state vectors, minimising latency.
-Minions have a fixed list of work items, while a controller thread sends compute/update commands before awaiting worker completion.
+\emph{Minions} have a fixed list of work items, while a \emph{controller} thread sends compute/update commands before awaiting worker completion.
 Work items are disjoint, requiring no policy locks.
@@
 \end{description}
-Each offers a different point of optimisation; if updates are disabled, then the \emph{\indfw{}} model can maximise throughput, while the \emph{\coopfw{}} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
+Each offers a different point of optimisation; if updates are disabled, then the \indfw{} model can maximise throughput, while the \coopfw{} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
+These correspond to only executing a trained policy and actively (re-)training a policy, respectively.
 %ParSa is described here at a high level, as an algorithm suited for \emph{any multiprocessor environment}.
@@
 %In general it is also possible for the \emph{Ctl} task to act as an additional \emph{Minion} in its parallel sections; we were limited here by code store requirements on the NFP.
-
 Latency and throughput, as in many networked systems, have different effects upon RL agents according to their design and target problem.
-Higher throughput for an RL agent is a necessity for per-flow or per-packet applications, which can require high decision-per-second rates even after combining state measurements received from the environment, such as flow control in DDoS prevention.
-Equally, lower latency affords an RL agent finer-grained control and learning of a problem, being able to react sooner to new information (e.g., device state in a routing optimisation problem, or queue depth when trying to enforce packet pacing).
-
-We found that, paradoxically, it was more efficient in the \Coopfw{} case \emph{to do more work} by having each worker recompute its tile subset from a stored state.
+Higher RL throughput is a necessity for per-flow or per-packet applications, which can require high decision-per-second rates even after combining state measurements received from the environment, such as flow control in DDoS prevention.
+Equally, lower latency affords an agent finer-grained control and learning of a problem, being able to react sooner to new information (e.g., device state in a routing optimisation problem, or queue depth when trying to enforce packet pacing).
+
+Paradoxically, we found that in the general \emph{ParSa} algorithm it was more efficient \emph{to do more work} by having each worker recompute its tile subset from a stored state.
 It transpired that cacheing this data placed a larger \texttt{memcpy} in the serial section, whose size did not scale at all with bit depth.
 Additionally, we do not use bitshifts in place of division operations in our implementation, due to the strict limits that power-of-two tile widths place upon policy design.
-
-?? Explain that above is a general property, probably still holds particularly at high cpu speed
 
@@
 Through these channels the system \emph{pushes} state vectors, reward measures, and control/setup packets as its inputs, and \emph{pulls} a stream of state-action pairs as its outputs.
-This allows agent decisions to be made asynchronously---preventing packet stalling---and allowing for many parallel decision-making agents to be used if desired.
-The key insight of this mechanism is that on-chip reward/state signals enjoy first-class support in much the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-NIC/chip asynchronous processes or the controller.
+This allows decisions to be made asynchronously---preventing packet stalling---and allowing many parallel decision-making agents to be used if desired.
+The key insight of this mechanism is that on-chip reward/state signals enjoy first-class support in the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-NIC/chip asynchronous processes or the controller.
 As such, \approachshort{} can receive from P4 \texttt{extern}s or another dedicated off-path flow state measurement application.
@@
 Our implementation uses platform-specific IPC (\emph{EMEM ring buffers}) with hardware signalling mechanisms on work arrival to achieve this.
-Due to lack of memory allocation in these environments, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state/configuration/policy data handover.
-Packet headers are extracted and parsed using the tooling and data types autogenerated by the P4 pipeline.
+As PDP hardware lacks dynamic memory allocation, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state/configuration/policy data handover.
+Packet headers are extracted and parsed using the tooling autogenerated by the P4 pipeline.
 We found that this costs a median \SIrange{126}{140}{\nano\second} communication time (local--remote), comparable to message channels in the Rust and Go languages on commodity hardware.
@@
 Even with parallel problems such as \emph{ParSa}, optimising for latency requires meticulous care in how work is passed out and aggregated.
-This is truer still when moving from the moderately fine-grained, efficient control of classical RL methods ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
+This is truer still when moving from the moderately fine-grained control of classical RL methods ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
 Ordinarily, the marshalling of requests, responses, and safe shared data access can incur significant extra overheads.
 On-chip execution and the nature of action preference computation allow us to use lockless atomic aggregation, removing the overheads of explicit messaging/packetisation.
-Moreover, physically adjacent functional units may have special-purpose shared registers or share a small fast cache to accelerate communication further.
+Moreover, physically adjacent functional units/cores often have special-purpose shared registers or share a small fast cache to accelerate communication.
 
@@
 
-Our implementation exploits the locality of threads, cores, and their parent islands in the NFP architecture.
-Policy compute/update tasks and configuration updates are passed between cores using these direct \emph{next neighbour} registers, signalling all child threads in response.
-Such on-chip local signals cost us just $\sim$\SI{20}{\nano\second} latency per relayed message.
-Each task performs atomic adds to a shared preference list, and atomically increments an acknowledgement counter to be periodically checked by the master thread, implementing the wait-free \emph{ParSa} algorithm we introduce (\cref{alg:parsa}).
+Our implementation exploits the locality of cores in the NFP architecture, which can be used on other platforms like NetFPGA.
+Policy compute/update/configure tasks are passed between cores using direct \emph{next neighbour} registers, signalling all child threads in response.
+Such on-chip signals cost just $\sim$\SI{20}{\nano\second} per relayed message.
+Each core performs atomic adds to a shared preference list and an acknowledgement counter checked by the master thread, implementing the wait-free \emph{ParSa} algorithm we introduce (\cref{alg:parsa}).
 This is essential for aggregation compared to the use of bounded message buffers, which caused significant head-of-line blocking.
@@
 \subsection{Reconfigurability}
-\approachshort{} allows for policy design and algorithmic learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) to be chosen and changed at runtime using at most two control packets.
-For instance, the former would be useful at the end of learning, moving from online to offline or when trying to use data from the same vantage point to train a policy for another problem afterwards.
-The latter is useful when an online agent must become more (or less) adaptive to new data (i.e., after a changepoint in traffic is detected).
-This extends to policy data, which may be imported from an offline pre-trained model via many such packets, and exported via the NFP binary support package on the host machine.
-Some aspects must be chosen at compile time; bit depth, work allocation strategy, and maximum policy/tiling/state sizes---these govern core operation or allocated memory regions.
-Choosing a bit depth of \qtylist[list-pair-separator = { or }]{16;8}{\bit} halves or quarters policy memory costs, allowing more complex problems to be modelled using more dimensions or fine-grained policies.
+\approachshort{} allows policy design and algorithmic learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) to be changed at runtime using at most two control packets.
+For instance, design changes are useful at the end of learning (moving from online to offline), or when trying to train a policy for another problem afterwards from the same vantage point.
+Parameter changes are useful when an online agent must become more (or less) adaptive to new data (i.e., after a changepoint in traffic is detected).
+This extends to policy data, which may be imported from an offline pre-trained model via such packets and exported via PCIe to the host machine.
+Some aspects must be chosen at compile time; bit depth, work allocation strategy, and maximum policy/tiling/state sizes---these govern core operation or allocated memory.
+Choosing a bit depth of \qtylist[list-pair-separator = { or }]{16;8}{\bit} halves/quarters policy memory costs, allowing more complex problems to be modelled using more dimensions or fine-grained tiles.
 
@@
 \subsection{Work Allocation}\label{sec:work-allocation}
-Due to the lack of dynamic memory allocation on programmable dataplane hardware, and to simplify action-value lookups, policies cannot be effectively stored sparsely.
-As a consequence, tiling space requirements scale exponentially with dimension count, and so higher-dimension tilings must be placed in memory regions large enough to allocate space for them.
-As a result, tilings are split across memory regions, giving different access and compute costs to different tasks---split by CLS/CTM/IMEM on the NFP, or e.g., BRAM/SRAM/DDR3 on NetFPGA.
-
-We use a simple first-fit work placement algorithm run on the NFP, placing the largest work item into the least loaded context of the least loaded ME.
+Due to the lack of dynamic memory allocation on PDP hardware, and to simplify action-value lookups, policies cannot be stored sparsely.
+As a consequence, tiling space requirements scale exponentially with dimension count, and so higher-dimension tilings must be placed in larger (and slower) memory regions.
+As a result, in many architectures policy parameters are split across such regions, giving different access and compute costs to different tasks.
+
+We use a simple first-fit work placement algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core (as the NFP exposes hardware threads).
 Each work item is a separate \emph{tiling} over a number of dimensions, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
-The cost of any individual work item (according to its dimension count and memory location) was empirically measured offline, and we weigh the total cost per ME to account for the number of minion threads available.
-This weighting specifically accounts for the presence of the controller thread on the first ME.
-This work allocation and any cached per-work-item values are recomputed when any configuration is installed or changed.
+The cost of any individual work item (according to its dimension count and memory location) was empirically measured offline, and we weigh the total cost per core to account for the number of minion threads available.
+This weighting specifically accounts for the presence of the controller thread on the first core.
+Work allocation and any cached per-item data are recomputed when any configuration is installed or changed.
 Naturally, for $n$ tilings/tasks and $m$ threads this procedure is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per task, storing $m/8$ and $\le8$ costs respectively.
@@
 To achieve line-rate match-action table lookup, platforms like NFP use accelerated datastructures (e.g., DCFL~\parencite{DBLP:conf/infocom/TaylorT05}) which must be computed by the controller over the \emph{entire rule set}.
-Even through externs, directly adding new rules is neither feasible nor safe.
-
-We instead suggest on NFP chips that \texttt{extern}s or datapath stages which apply actions to packet processing should maintain a small local store of state-action pairs, and periodically send these back to the controller for batch rule installation.
-This ensures that the vast majority of installed rules exhibit accelerated performance, while preventing rule installation delay on the newest decisions.
+Even through \texttt{extern}s, directly adding new rules is neither feasible nor safe.
+We instead suggest on NFP chips that \texttt{extern}s or datapath stages which apply actions to packets should maintain a small store of state-action pairs, and periodically send these back to the controller for batch installation.
+This ensures that the majority of installed rules exhibit accelerated performance, while preventing installation delay on the newest decisions.
 Platforms such as Intel Tofino greatly simplify this, with Tofino Native Architecture intrinsics such as \emph{Action Profiles/Selects} allowing a P4 action to be chosen based on a register value (i.e., an RL action).
@@
 \subsection{Targeting other device classes}
-While this design caters primarily to Netronome's SmartNIC devices, many aspects of this design have clear analogues in NICs of similar form-factor, such as the NetFPGA SUME platform.
-For instance, the availability of other cores can be handled by designing additional off-path functional units, much like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-Rather than general-purpose cross-core communication, an RL functional unit would be hard-wired to and from the custom actions that interface with it, using the P4$\rightarrow$NetFPGA toolchain to offer the base framework and granular control over packet and flow selection.
-Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
-As our later results show, working on a bit depth below the native word size forces the compiler or programmer to emit excess code to extract and re-align preference values.
+While specifics of this design cater to Netronome devices, many aspects have clear analogues in NICs of similar form-factor, such as the NetFPGA SUME platform.
+For instance, other cores can be introduced by designing additional off-path functional units,  like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+Rather than general-purpose IPC, an RL functional unit would be hard-wired between internal workers and to the custom actions that interface with \approachshort{}, using the P4$\rightarrow$NetFPGA toolchain to offer the base framework and granular control over packet and flow selection.
+%Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
+As our later results show, using a bit depth below the native word size forces the compiler or programmer to emit excess code to extract and re-align preference values.
 This reduces overall efficiency even though considerably fewer reads are made.
-We expect that more bespoke hardware/FPGA co-design would allow this boiler-plate to be removed and enable the SIMD-like optimisations we discuss above without creative and costly bit-stuffing.
-
-Bringing \approachshort{} to high-port density devices such as Intel Tofino-powered switches is more difficult.
+We expect that more bespoke hardware/FPGA co-design would allow this boiler-plate to be removed and enable SIMD-like optimisations we discuss without creative and costly bit-stuffing.
+
+Bringing \approachshort{} to high-port density devices such as Tofino-powered switches is more difficult.
 As discussed in \cref{sec:motivation}, Tofino ASICs map closely to the P4 PSA, meaning that there are no spare asynchronous general-purpose compute units to place this work on.
@@
 \section{Evaluation}\label{sec:evaluation}
-We investigate the performance of \approachshort{} compared to classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{$\times$} speedup in median--\nth{99} state-action latency and \SI{9.9}{$\times$} greater online learning throughput.
+We investigate the performance of \approachshort{} compared to classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{$\times$} speedup in median--\num{99.99}\nthscript{th} state-action latency and \SI{9.9}{$\times$} greater online learning throughput.
 Crucially, we find that in-NIC execution offers tight tail latency bounds compared to host-based approaches.
@@
 
+\subsection{Netronome Platform Fundamentals}\label{sec:netronome-platform-fundamentals}
+As we implement this work on Netronome NFP SmartNICs, it is necessary to explain its basics.
+These \emph{system-on-a-chip} (SoC) devices achieve scalable packet processing through sheer parallelism.
+Most of the chip is composed of \emph{microengines} (MEs), grouped into \emph{islands} of 4 or 12 MEs.
+All 12-ME islands are used by a default P4 pipeline.
+Each ME has \numrange{4}{8} \emph{contexts} (hardware threads) which share a code store.
+%Contexts and MEs may send one another numbered signals, and MEs have a small \emph{next-neighbour} register file for passing values in one direction to the next ME on the same island.
+%MEs run a proprietary instruction set, compiled to via a \emph{(Micro-)C} compiler.
+Beyond registers, the platform implements an explicit memory hierarchy scaling in size, location, and access cost:
+$\text{LMEM (ME)} < \text{CLS (Island)} < \text{CTM} < \text{IMEM (Chip)} < \text{EMEM}$.
+
 \subsection{Experimental Setup}\label{sec:experimental-setup}
-Testing machines had the following specifications, all having \SI{32}{\gibi\byte} RAM:
+Testing machines had the following hardware, all with \SI{32}{\gibi\byte} RAM:
 \begin{description}
@@
 
-We run \approachshort{} on an island with \num{4} MEs of the NFP-6480 for a total of \num{32} hardware threads.
-This is the maximum number of MEs on a single island which is not in use by an existing P4 pipeline.
-Where feasible, we test \SIlist{32;16;8}{\bit} quantised arithmetic firmware versions.
-All \approachshort{} timing measurements were repeated over \num{10000} input state packets (preceded by \num{1000} warmup packets), storing an item processing time which accessed via the control machine, from which throughput was derived.
+We run \approachshort{} on an island with \num{4} MEs of the NFP-6480, totalling \num{32} hardware threads.
+This is the maximum number of MEs on a single island which is not in use by a P4 pipeline.
+Where feasible, we test \SIlist{32;16;8}{\bit} quantised arithmetic.
+All \approachshort{} timing measurements were repeated over \num{10000} input state packets (preceded by \num{1000} warmup packets), retrieving each item processing time over PCIe via the controller machine, from which throughput was derived.
 Host throughput and latency measures were observed over \num{10} trials of \SI{10}{\second} (with \SI{5}{\second} warmup/cooldown times).
@@
 Policy sizes are set to those matching a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dimension state vectors, a bias tile and 16 full tiling sets (\numproduct{7x1}-dim, \numproduct{8x2}-dim, \numproduct{1x4}-dim), 8 tilings per dimension set, 6 tiles per dimension, choosing between 10 actions.
-For context, such input state would be various aspects of per-flow state (e.g., IATs, up/down rate) which are combined with other state such as the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim)
-For \Coopfw{} for instance, this creates \num{129} work items across \num{31} workers.
+For context, such input state would contain various aspects of per-flow state (e.g., IATs, up/down rate) which are combined with other state such as the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim).
+In \Coopfw{} for instance, this creates \num{129} work items across \num{31} workers.
 Although the model application's more successful \emph{SPF} agent design uses 3 actions, we choose a larger action set to investigate the performance of more complex agents.
@@
 \fakepara{Raw inference and learning performance}
-We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs against floating point (numpy-based) implementations running on a commodity host machine.
+We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs against floating point (numpy-based) implementations on a commodity host machine.
 This allows us to demonstrate the performance differences between the \indfw{} and \coopfw{} configurations, particularly in how \indfw's required policy locks impact throughput.
-We also report on the derived throughput measures accounting for the locking behaviour of each.
-We compare online learning performance (input states produce an output, and then update the policy with appropriate locking) with offline (input states \emph{only} produce an output) in these cases.
+%We also report on the derived throughput measures accounting for the locking behaviour of each.
+We compare online learning performance (input states produce an output, and then update the policy) with offline (input states \emph{only} produce an output) in these cases.
 Online performance marks the number of decisions that can be made per second (and associated latency) when training a policy.
 Offline performance is crucial for pushing a trained, known-good policy to agents in the network with an expected higher raw decision throughput.
-State-action latency is a shared property of both cases, with the main difference being the impact on throughput of the update step.
+State-action latency is a shared property of both cases, with the main impact on throughput arising from the update step.
 
 Building on this, we vary the amount of allocated worker threads to show how \approachshort{} can scale to fit available compute resources on a device.
-This is an important aspect for (automated) administration and allocation of compute resources in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of time work scheduling which we examine later.
+This is an important aspect for (automated) administration and allocation of compute resources in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of-time work scheduling which we examine later.
 This also demonstrates the number of cores needed to achieve a given latency/throughput bound on a policy of representative complexity.
@@
 This provides maximum throughput in the former case (relying solely on NIC counters for loss counting).
-In the latter case this minimises host resource contention to observe exact latency measurements, have a high enough sample count to detect subtle (aggregate) latency effects, and eliminate \emph{host} receive drops.
-DPDK was setup using four \SI{1}{\gibi\byte} hugepages.
+In the latter case, this minimises host resource contention to observe exact latency measurements, have a high enough sample count to detect subtle (aggregate) latency effects, and eliminate \emph{host} receive drops.
+DPDK was setup using \qtyproduct[product-units=single]{4 x 1}{\gibi\byte} hugepages.
 Sent traffic was comprised of fixed-size \SIrange{64}{1518}{\byte} packets~\cite{rfc2544}.
@@
 Using the maximum policy size defined above, we investigate how the memory requirements imposed by \approachshort{} vary with the number of dedicated MEs, over and above a base P4 forwarding plane.
-We report resource use for \SI{32}{\bit} \Indfw{} and \Coopfw{} agents with maximum policy sizes as above, with hash-tables sized to \num{4096} state-action pairs and \num{16} separate reward values.
+We report resource use for \SI{32}{\bit} \Indfw{} and \Coopfw{} agents, with hash-tables sized to \num{4096} state-action pairs and \num{16} separate reward values.
 This captures the relative cardinality of network RL traces to rewards, i.e., many input flows will map to one or few reward values (i.e., DDoS attack size estimation per egress-AS, queue occupancy in the case of AQM per output port).
@@
 \Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based implementation of an RL policy\footnote{For brevity, we omit numpy-based integer results---summarised against a floating point implementation, median state-action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}.
-We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using one island (4 MEs) of the NFP-6480.
+We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using 4 MEs of the NFP-6480.
 This corresponds to \SIlist{15.0;21.0}{$\times$} speedups over a \emph{Collector} host.
@@
 \Cref{tab:tputs} compares \approachshort{}'s throughput against host-based execution.
-We choose the worker count on host machines as the number of parallel agents that maximises throughput.
+We choose the worker count on host machines that maximises throughput.
 This equalled the amount of physical cores on each device, while moving beyond this (even below the number of hyper-threads) would hamper tail latencies by an order of magnitude.
@@
 \begin{figure}
-	\resizebox{1.0\linewidth}{!}{
+	\resizebox{0.9\linewidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-core}
@@
 \begin{figure}
-	\resizebox{1.0\linewidth}{!}{
+	\resizebox{0.9\linewidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
 	}
-	\caption{\Coopfw{}'s action latencies improve as additional cores are made available.
+	\caption{\Coopfw{}'s action latencies improve with additional cores.
 This requires \num{3} workers (4 total contexts) to improve upon the state-action latency of single-threaded in-NIC RL.\label{fig:vary-core-latency}}
@@
 
-\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects complete update costs and state-action latency respectively, scaling from a bias tile up to the full DDoS policy size\footnote{Input vectors here all have 20 elements regardless of the policy design.}.
-\Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning performance.
+\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects update costs and state-action latency respectively, scaling from a bias tile up to the full DDoS policy size\footnote{Input vectors here all have 20 elements regardless of the policy design.}.
+\Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning.
 %We note that this is a trivial case, in that the use of \emph{only} a bias tile negates the need for any input state (similar to a multi-arm bandit problem).
-We note that this is a trivial case, as using \emph{only} a bias tile provides/alters a single preference list regardless of input state.
+We note that this is a trivial case, as using \emph{only} a bias tile returns a single preference list regardless of input state.
 
 \begin{figure}
-	\resizebox{1.0\linewidth}{!}{
+	\resizebox{0.9\linewidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-work}
 	}
-	\caption{\Coopfw{} can fully process individual items faster than \Indfw{}---thus has higher online performance---on almost all policy sizes.
+	\caption{\Coopfw{} fully processes updates faster than \Indfw{}---thus has higher online performance---on almost all policy sizes.
 Lower bit depths are more effective on simpler policies.\label{fig:vary-work}}
@@
 
-We found that a key aspect of in-NIC execution is that it allows far tighter bounds on tail latency to be offered compared to host offloading.
+We found that a key aspect of in-NIC execution is that it allows far tighter bounds on tail latency compared to host offloading.
 Examining the state-action latencies in \cref{tab:lats}, we see that \num{99.99}\nthscript{th} percentile latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, respectively.
 Similar results were observed for other bit depths.
-By contrast, host-based tail latencies are at least \SI{40.53}{\percent} greater in our test setup when parallel workers are at or below the physical core count.
-We show the cumulative distributions of these in detail in \cref{fig:lat-cumul}, noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF or traffic processing task---can impact tail latencies further (\emph{Float(Over)}).
+By contrast, host-based tail latencies are at least \SI{40.53}{\percent} greater even when parallel workers are at or below the physical core count.
+We show the cumulative distributions of these in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF/traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
 
 \begin{figure}
-	\resizebox{1.0\linewidth}{!}{
+	\resizebox{0.9\linewidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
@@
 \begin{figure}
-	\resizebox{1.0\linewidth}{!}{
+	\resizebox{0.9\linewidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
@@
 \begin{figure}
-	\resizebox{1.0\linewidth}{!}{
+	\resizebox{0.9\linewidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-core-tput-on-per}
@@
 
-?? How do we move this intuition to the start? ELI5 as dimitris says, to get the full end-to-end pipeline across to the reader.
+%?? How do we move this intuition to the start? ELI5 as dimitris says, to get the full end-to-end pipeline across to the reader.
 
 \subsection{In-Network DDoS Defence}\label{sec:integ-1}
-Classical reinforcement learning has seen recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
-\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and local, per-flow state to monitor how flows respond to bandwidth changes and packet drop---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} (observing how flow behaviour reacts to rate limits increasing or decreasing) with more allowance for congestion-unaware protocols.
+Classical RL has seen recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
+\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and local, per-flow state to monitor how flows respond to bandwidth changes and packet drop---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} (observing how flow behaviour reacts to a change in rate limits) with more allowance for congestion-unaware protocols.
 Actions then move flows up or down in punishment levels according to a finite state machine.
@@
 
-The original work suggests theoretical ``ground truth'' reward functions, whose correct implementation and designs were left as an open challenge. 
-We posit that INDDoS~\parencite{tnms-ddos-victim-ident}, which uses Count-min Sketches to estimate DDoS victim cardinality, could be used as an effective reward function source---i.e., using the number of detected victims as a loss value.
+The original work uses theoretical ``ground truth'' rewards, whose correct implementation and designs were left as an open challenge. 
+We posit that INDDoS~\parencite{tnms-ddos-victim-ident}, which uses Count-min Sketches to estimate DDoS victim cardinality, could be an effective reward function source---i.e., using the number of detected victims as a loss value.
 
 \fakepara{Integrating \approachshort}
-Before each monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{OUT Ring} for any generated actions.
-As we note in \cref{sec:limitations}, these actions would be placed into a small local hash table and simultaneously exported to the attached controller to be inserted as P4 rules in batches.
+Before each monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{\outring{} Ring} for any generated actions.
+As we note in \cref{sec:limitations}, these actions would be placed into a small hash table and simultaneously exported to the attached controller to be inserted as P4 rules in batches.
 After the main monitoring step, packet ingress timestamps would be used to emulate the \emph{Timed Random Sequential} (TRS) scheduler used by the anti-DDoS agents for rate-controlled work, where state vectors would be selected and passed into \approachshort.
@@
 
-A reduction in state-action latency (i.e., with \Coopfw) is, as always, useful for minimising the noise inherent in learning.
+Reducing state-action latency (i.e., with \Coopfw) is useful for minimising the noise inherent in learning.
 However, an agent is limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
@@
 
-We note that the TRS scheduler is designed to handle cases where the number of flows created by an attacker, combining state vectors and information over time while the asynchronous agent is itself busy.
-An increased number of (attacking) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
-As shown in \cref{sec:evaluation}, \approachshort{} far exceeds the action throughput of a host based deployment.
-The control plane is then free to use wildcards or specific matches to narrow down (or expand) the set of flows to be considered and controlled, though explicit scheduling techniques are still key in such adversarial environments.
+We note that the TRS scheduler is designed to handle cases where the number of flows created by an attacker is large, combining state vectors and information over time while the asynchronous agent is itself busy.
+An increased number of (attack) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
+As shown in \cref{sec:evaluation}, \approachshort{} far exceeds the throughput of host offloading.
+The control plane is then free to use wildcards or specific matches to narrow down or expand the set of flows to be considered and controlled, though explicit scheduling techniques are still key in such adversarial environments.
 
@@
 The two compute models discussed above need not be homogeneously deployed.
-For instance, in a networked deployment, a subset of \approachshort{} nodes could be \SI{32}{\bit} \coopfw{} agents, training online, while most other nodes ran \SI{8}{\bit} \indfw{} designs to meet throughput guarantees.
+For instance, in a networked deployment, a subset of \approachshort{} nodes could be \SI{32}{\bit} \coopfw{} agents, training online, while most other nodes run \SI{8}{\bit} \indfw{} designs to meet throughput guarantees.
 The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
@@
 Such designs would more generally enable asynchronous and novel compute in SmartNICs and programmable switches, for instance offering consistent and easy to use communication between workers versus hard-coded ME relationships.
-
 Event-driven versions of the P4 pipeline have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}.

### 2021-06-08 paper/paper.tex
Now that's what I call progress.

@@
 \email{Dimitrios.Pezaros@gla.ac.uk}
+
+\setlength{\abovecaptionskip}{5pt}
+\setlength{\belowcaptionskip}{0pt}
+
+\setlength{\textfloatsep}{6.0pt plus 2.0pt minus 4.0pt}
+\setlength{\floatsep}{6.0pt plus 1.0pt minus 2.0pt}
+
+\setlength{\dbltextfloatsep}{6.0pt plus 2.0pt minus 4.0pt}
+\setlength{\dblfloatsep}{6.0pt plus 1.0pt minus 2.0pt}
 
@@
 	\centering
-	\begin{subfigure}{0.45\linewidth}
+	\begin{subfigure}{0.47\linewidth}
 		\centering
-		\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/ind}
+		\includegraphics[keepaspectratio, width=0.78\linewidth]{figures/ind}
 		\caption{\Indfw{} (offline throughput-optimal).
-\emph{Workers} independently pull commands from (and push state-action pairs to) the environment, locking access to the policy when updates must occur.\label{fig:single-and-parallel:single}}
+\emph{Workers} independently pull commands from (and push actions to) the environment, locking policy access for updates.\label{fig:single-and-parallel:single}}
 	\end{subfigure}
 	\hspace{0.04\linewidth}
-	\begin{subfigure}{0.45\linewidth}
+	\begin{subfigure}{0.47\linewidth}
 		\centering
-		\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/coop}
+		\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/coop}
 		\caption{\Coopfw{} (online-optimal).
@@
 We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using 4 MEs of the NFP-6480.
-This corresponds to \SIlist{15.0;21.0}{$\times$} speedups over a \emph{Collector} host.
+This corresponds to \SIlist{15;21}{$\times$} speedups over a \emph{Collector} host.
 Importantly, \Indfw{} achieves lower median state-action latencies (\SI{2.79}{$\times$}) \emph{and} update times (\SI{2.63}{$\times$}) than a dedicated \emph{Collector} host \emph{while requiring only a single core or dedicated functional unit}.
+
+\newlength{\resultplotwidth}
+\setlength{\resultplotwidth}{0.85\linewidth}
 
@@
 \Cref{tab:tputs} compares \approachshort{}'s throughput against host-based execution.
-We choose the worker count on host machines that maximises throughput.
-This equalled the amount of physical cores on each device, while moving beyond this (even below the number of hyper-threads) would hamper tail latencies by an order of magnitude.
+We chose the worker count on host machines that maximised their throughput.
+This equalled the amount of physical cores on each device---moving beyond this (even below the number of hyper-threads) would hamper tail latencies by an order of magnitude.
 To make the comparison fair in the context of many-core CPU environments, we include per-core throughput.
@@
 While the compute times and latency of \Coopfw{} always outperform the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency.
-Some artefacts of our environment and design choices are visible, such as the addition of new physical cores (MEs) being more significant than contexts, and the presence of some schedule bottlenecks.
+Some artefacts of our environment and design choices are visible, such as the addition of new physical cores being more significant than contexts, and the presence of some schedule bottlenecks.
 Most importantly however, \approachshort{}-\Coopfw{}'s resource demand is tunable at compile time to meet the online training rate and/or action latency required by a task/environment.
@@
 \begin{figure}
-	\resizebox{0.9\linewidth}{!}{
+	\resizebox{\resultplotwidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-core}
@@
 \begin{figure}
-	\resizebox{0.9\linewidth}{!}{
+	\resizebox{\resultplotwidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
@@
 
-\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects update costs and state-action latency respectively, scaling from a bias tile up to the full DDoS policy size\footnote{Input vectors here all have 20 elements regardless of the policy design.}.
+\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects update cost and state-action latency respectively, scaling from a bias tile up to the full DDoS policy size\footnote{Input vectors here all have 20 elements regardless of the policy design.}.
 \Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning.
@@
 \begin{figure}
-	\resizebox{0.9\linewidth}{!}{
+	\resizebox{\resultplotwidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-work}
@@
 Similar results were observed for other bit depths.
-By contrast, host-based tail latencies are at least \SI{40.53}{\percent} greater even when parallel workers are at or below the physical core count.
+By contrast, host-based tail latencies are at least \SI{40.53}{\percent} greater even when the parallel worker count is at or below the physical core count.
 We show the cumulative distributions of these in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF/traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
@@
 \begin{figure}
-	\resizebox{0.9\linewidth}{!}{
+	\resizebox{\resultplotwidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
@@
 		}
-		\caption{Tail latencies suffer in host execution---particularly when oversubscribed.}
+		\caption{Tail latencies suffer in hosts---particularly when oversubscribed.}
 	\end{subfigure}
@@
 A noteworthy trend is that the \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit} designs, with the exception of smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
-This held even though our implementation is optimised to read and write policy tile data in batches (achieving fewer I/O operations as required).
-This is due to the native register width on the NFP being \SI{32}{\bit}, and so the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
-This matches \SI{32}{\bit} becoming dominant in more complex workloads: higher dimension tilings require more arithmetic operations to compute the hit tile index.
-Most of the I/O comes after this step, causing ALU use to dominate.
-This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) \Coopfw{} agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the algorithm in the online case.
-As a result, these bit depths give a \SIrange{2}{4}{$\times$} reduction in memory needed to store a policy (enabling more fine-grained or more complex policies), in exchange for slightly worse latency and throughput.
-We investigated bit-stuffing several such values into a single word for our atomic writeback mechanism (as the platform offers both \SIlist{32;64
+This occurs even though our implementation is optimised to read and write policy tile data in batches (achieving fewer I/O operations as required).
+We see this because the native register width on the NFP is \SI{32}{\bit}, and so the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
+This matches \SI{32}{\bit} becoming dominant in complex workloads: higher dimension tilings require more arithmetic operations to compute.
+Most of the I/O comes \emph{after} this step, causing ALU use to dominate.
+This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the online algorithm.
+Smaller bit depths still give a \SIrange{2}{4}{$\times$} saving in memory for policy storage (enabling more fine-grained or more complex policies), in exchange for slightly worse latency and throughput.
+To overcome this, we investigated bit-stuffing several values into a single word for our atomic writeback mechanism (as the platform offers both \SIlist{32;64
 }{\bit} atomic addition).
-This is analogous to SIMD---through clever use of padding bits---but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
+This is analogous to SIMD through clever use of padding bits, but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
 
 \fakepara{Work allocation}
-\Cref{fig:work-alloc-32} shows that our heuristic first-fit allocator (\emph{Balanced}) is key in achieving consistent sub-\SIlist{35;65}{\micro\second} latencies/processing times, respectively.
-The trend is repeated for all bit depths, noting the latency differences described above.
-The constant difference between \emph{Action} and \emph{Update Prep} was observed to scale linearly with bit depth, matching storage and lookup work in the serial portion between parallel jobs (plots omitted).
-The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its ID, as batching work in contiguous chunks leads to some workers receiving exclusively high-dimensionality tile sets.
-The minor gap in performance lower bounds between the \emph{Random} and \emph{Balanced} allocators suggests that further optimisations can be made in this way.
-We expect that closing or exceeding this gap may require more complex modelling of NFP's hardware thread interactions, which lies far beyond the scope of in-NIC scheduling.
+\Cref{fig:work-alloc-32} shows that our heuristic allocator (\emph{Balanced}) is key in achieving consistent sub-\SIlist{35;65}{\micro\second} latencies/update times, respectively.
+The trend is repeated for all bit depths.
+The constant difference between \emph{Action} and \emph{Update Prep} was observed to scale linearly with bit depth, matching storage and lookup work in the serial portion of \emph{ParSa} (plots omitted).
+The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its index, as batching work in contiguous chunks leads to some workers receiving exclusively high-dimensionality tilings.
+The minor gap in lower bound performance between the \emph{Random} and \emph{Balanced} allocators suggests that further optimisations can be made.
+We expect that closing or exceeding this gap may require more complex modelling of hardware thread interactions, which lies far beyond the scope of in-NIC scheduling.
 Some additional complexity may be tolerated, subject to code store limits---scheduling runs exactly once per configuration change, so does not impact per-action code.
@@
 \begin{figure}
-	\resizebox{0.9\linewidth}{!}{
+	\resizebox{\resultplotwidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
@@
 \begin{figure}
-	\resizebox{0.9\linewidth}{!}{
+	\resizebox{\resultplotwidth}{!}{
 		\includegraphics{../plots/build/rl-perf-tester/vary-core-tput-on-per}
@@
 	\emph{This gives an end-to-end inference latency of \SI{517.9}{\micro\second}}.
-	\item[vNF Offload.] \Textcite{DBLP:journals/cm/CzivaP17} show that more lightweight vNF frameworks like GNF~\parencite{DBLP:journals/cm/CzivaP17} and ClickOS~\parencite{DBLP:conf/nsdi/MartinsAROHBH14} add \SIrange{45}{55}{\micro\second} \emph{additional} latency above PCIe costs.
+	\item[vNF Offload.] \Textcite{DBLP:journals/cm/CzivaP17} show that more lightweight vNF frameworks like GNF~\parencite{DBLP:journals/cm/CzivaP17} and ClickOS~\parencite{DBLP:conf/nsdi/MartinsAROHBH14} add \SIrange{45}{55}{\micro\second} \emph{additional} RTT latency above PCIe costs.
 	\emph{This gives an end-to-end inference latency of \SI{572.9}{\micro\second}}.
@@
 \fakepara{Co-existence with the dataplane}
-We were able to saturate the \SI{40}{\giga\bit\per\second} link for packet sizes at or above \SI{256}{\byte}.
+Our setup saturated the \SI{40}{\giga\bit\per\second} link for packet sizes $\ge$\SI{256}{\byte}.
 %For frame sizes of \SIlist{64;128;256}{\byte} input traffic rates were \SIlist{17.4;32.9;37.0}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2;18.1}{\mega\packet\per\second}).
-For frame sizes of \SIlist{64;128}{\byte} input traffic rates were \SIlist{17.4;32.9}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2}{\mega\packet\per\second}).
-Passing this traffic over the loaded NFP device running \approachshort, no packet losses were incurred at any rate of RL actions.
+For frame sizes of \SIlist{64;128}{\byte}, input traffic rates were \SIlist{17.4;32.9}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2}{\mega\packet\per\second}).
+Passing this traffic over the NFP device running \approachshort, no packet losses were incurred at any rate of RL actions.
 
 We show the effect of RL workloads on the round-trip latencies of cross traffic through \cref{fig:dataplane-coop}.
-As observed latency measures do not obey a normal distribution (particularly \SIlist{256;1518}{\byte}, which are bimodal), we employ a one-tailed \emph{Mann-Whitney U test} to mark statistically significant population increases in latency ($p < 0.05$) with a ``+''.
-In general, we observe that statistically significant latency increases concentrate around smaller packet sizes.
+As observed latencies do not obey a normal distribution (particularly \SIlist{256;1518}{\byte}, which are bimodal), we employ a one-tailed \emph{Mann-Whitney U test} to mark statistically significant population increases in latency ($p < 0.05$) with a ``+''.
+In general, statistically significant latency increases concentrate around smaller packet sizes.
 All (bar one) of these affected \nth{99} percentile latencies by under \SI{0.38}{\percent} (at most \SI{78}{\nano\second}).
 This slight degradation can be explained by increased pressure on the NFP's \emph{Command Push-Pull} (CPP) bus, which is responsible for handling cross-island accesses to memory (particularly IMEM/EMEM) and other resources.
-\approachshort{} places load on the CPP bus through use of its \inring{}/\outring{} rings for external communication and through last-tier policy accesses.
-This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---Netronome's P4 toolchain segments packets, storing metadata (e.g., MAC prepend data) and the first bytes of a packet in a $\SI{256}{\byte}$ CTM block and parking their payloads in EMEM.
+\approachshort{} places load on the CPP bus through its \inring{}/\outring{} EMEM rings and last-tier policy accesses.
+This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---the NFP P4 toolchain segments packets, storing metadata (e.g., MAC prepend) and the first bytes of a packet in a \SI{256}{\byte} CTM block and parking their payloads in EMEM.
 \SI{256}{\byte} packets overshoot this due to metadata, causing small I/O accesses at a high rate for packets sized around this cutoff point.
@@
 		}
-		\caption{Distribution of RTTs for \SI{128}{\byte} packets for baseline and \num{3000} RL actions/s.\label{fig:dataplane-example}}
+		\caption{Distribution of RTTs for \SI{128}{\byte} packets for \numlist{0;3000} RL actions/s.\label{fig:dataplane-example}}
 	\end{subfigure}
@@
 \fakepara{Resource requirements}
-\Cref{tab:resources} demonstrates how \approachshort{} consumes shared and local memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
+\Cref{tab:resources} demonstrates how \approachshort{} consumes shared memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
 As one program is installed per ME, these results represent the minimum and maximum resource use we can command on a single island (i.e., without replacing P4 workers).
-We observe negligible costs on shared bulk EMEM storage ($\sim$\SI{4}{\mebi\byte}), including storage space for a hash-table of \num{4096} state-action pairs and \num{16} reward values.
+We observe negligible costs on shared bulk EMEM storage ($\sim$\SI{4}{\mebi\byte}), incurred due to hashtables for state-action pairs and rewards.
 The most significant costs arise due to policy data (\SI{405}{\kibi\byte} shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
@@
 \caption{NFP memory use due to \approachshort{} using \numlist{1;4} MEs (\SI{32}{\bit}).
-CLS and CTM regions are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM regions are shared between all NFP programs on a NIC.\label{tab:resources}}
+CLS and CTM are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM are shared between all NFP programs on a NIC.\label{tab:resources}}
 \resizebox{\linewidth}{!}{
@@
 \fakepara{Deployability}
-Setup of \approachshort{} uses two packet types: \emph{setup} which contains learning (hyper-)parameters and most aspects of a tile coded policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
+Setup of \approachshort{} uses two packet types: \emph{setup}, which contains learning (hyper-)parameters and most aspects of a tile coded policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
 We found that setup and tiling packets take a mean \qtylist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing for an agent to be swapped between online and offline operation (or repurposed for another task/policy) painlessly.
@@
 Tiling packet processing scales linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
-Online-offline swaps for \Coopfw{} exhibit similar cost, however the need for explicit ahead-of-time heuristic scheduling means that policy/tiling \emph{structure} changes (including the first complete setup) take \SI{422.63}{\micro\second} for the full-size policy described above.
-The time taken for \Coopfw{} to schedule its tasks was found to scale with the number of workers ($m$) and work items ($n$) as described above ($\mathcal{O}{\left(n\log{m}\right)}$).
+Online-offline swaps for \Coopfw{} exhibit similar cost, however the need for explicit scheduling means that policy/tiling \emph{structure} changes (including the first complete setup) take \SI{422.63}{\micro\second} for the full-size policy described above.
+The time taken for \Coopfw{} to schedule its tasks was found to scale with the number of workers ($m$) and work items ($n$) as described earlier (\cref{sec:work-allocation}, $\mathcal{O}{\left(n\log{m}\right)}$).
 Ignoring the trivial solutions, reducing the worker count to \num{1} costs a mean \SI{238.22}{\micro\second}, while placing a single task incurs \SI{53.79}{\micro\second}.
@@
 \fakepara{Why in-NIC?}
-The above approach relies upon co-hosting traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
+The above approach relies on co-hosting traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
 However, packet mirroring, offloading RL computation to a host (potentially over layers of virtualisation), and CPU performance are all sources of additional, consistent state-action latency.
@@
 The two compute models discussed above need not be homogeneously deployed.
-For instance, in a networked deployment, a subset of \approachshort{} nodes could be \SI{32}{\bit} \coopfw{} agents, training online, while most other nodes run \SI{8}{\bit} \indfw{} designs to meet throughput guarantees.
+For instance, in a networked deployment, a subset of \approachshort{} nodes could be \coopfw{} agents, training online, while most other nodes run \indfw{} designs to meet throughput guarantees.
 The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
@@
 \fakepara{In-network ML}
-Taurus~\parencite{DBLP:journals/corr/abs-2002-08987} proposes that efficient line-rate, low-latency per-packet inference can be achieved by placing a run-time configurable grid topology of map-reduce processing units into the packet processing pipeline (implementing CNNs, LSTMs, and some classical approaches such as SVMs).
-Their CGRA-based implementation is able to be repurposed at runtime and achieves sub-\si{\micro\second} extra latency, though online model training is not considered.
-
-\emph{IIsy}~\parencite{DBLP:conf/hotnets/XiongZ19} shows how \emph{classical ML inference} techniques (SVMs, Na\"{i}ve Bayes, Decision Trees, K-Means) can be converted into match-action table structures which are compatible with \emph{any} P4 deployment.
-They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA devices (at line rate in most cases), but can only investigate the use of pre-trained models.
-
-A recent line of research in the community has been to investigate \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate, low latency packet classification.
-\emph{BaNaNa SPLIT} shows the initial steps of this trend as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} in an effort to reduce latency imposed by batching and bulk transfer to the GPU.
-However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}, with the hypothesis put forward that networking hardware would be a better fit for some parts of the inference pipeline.
-This line of work has moved onto full, in-network packet tagging and classification by pre-trained BNNs with \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
+Taurus~\parencite{DBLP:journals/corr/abs-2002-08987} proposes that efficient line-rate, low-latency per-packet inference can be achieved by placing a run-time configurable grid of map-reduce units into the packet processing pipeline (implementing e.g., CNNs, LSTMs, and SVMs).
+Their CGRA-based implementation achieves sub-\si{\micro\second} extra latency, though online model training is not considered.
+
+\emph{IIsy}~\parencite{DBLP:conf/hotnets/XiongZ19} shows how \emph{classical ML inference} techniques (SVMs, Na\"{i}ve Bayes, Decision Trees, K-Means) can be converted into match-action tables compatible with \emph{any} P4 deployment.
+They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA devices (at line rate in most cases), but can only investigate pre-trained models.
+
+A recent line of research in the community has been to investigate \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate packet classification.
+\emph{BaNaNa SPLIT} shows the initial steps of this trend as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} to reduce latency imposed by batching and bulk transfer to the GPU.
+However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}; the authors hypothesise that networking hardware would be a better fit for some parts of the inference pipeline.
+This has moved onto full, in-network packet tagging and classification by pre-trained BNNs with \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
 N3IC achieves packet inference in \SI{45}{\micro\second} on the NFP, and \SI{0.3}{\micro\second} on NetFPGA for \SI{256}{\bit} inputs.
 Comparatively, \approachshort{}-\Coopfw{} can process an identically-sized input (8-dim) in a median \SI{13.83}{\micro\second}.
-Our work handles larger inputs (\SI{640}{\bit}) at lower latencies of \SI{34}{\micro\second}, and offers online learning.
+Our work handles larger inputs (\SI{640}{\bit}) at lower latencies (\SI{34}{\micro\second}), and offers online learning.
 We expect that a NetFPGA implementation of \approachshort{} would enjoy a similar factor of speedup.
@@
 
-\textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantised weights and inputs on NFP hardware, using \emph{in-path} compute rather than our asynchronous model.
-Inference latency on small networks (3 layers, \num{30} layers) can be as high as \SI{500}{\micro\second} on line rate traffic, stressing the value of the asynchronous compute and quantisation we adopt.
+\textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantised weights and inputs on the NFP, using \emph{in-path} compute rather than our asynchronous model.
+Inference latency on small networks (3 layers, \num{30} neurons) can be as high as \SI{500}{\micro\second} on line rate traffic, stressing the value of path-adjacent compute.
 
@@
 Event-driven versions of the P4 pipeline have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}.
-Timer events and device state changes would empower in-network RL use-cases, signalling timesteps for RL agents or a new, effective, fine-grained sources of input state.
+Timer events and device state changes would empower in-network RL use-cases, signalling timesteps for RL agents or new, effective, fine-grained sources of input state.
 
@@
 We have presented \emph{\approachshort{}}, bringing \emph{online reinforcement learning} to the dataplane.
-\approachshort{} has shown how classical RL techniques make online learning possible by simplifying update logic and enable parallel processing.
-These algorithms (and in-NIC deployment) enabled a \SIrange{15}{21}{$\times$} reduction in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput compared to host offloading.
-The deployment environment and asynchronous design were shown to eliminate PCIe and impose minimal impact on carried dataplane traffic.
-\approachshort{}'s designs scale with additional compute resources at deployment to improve upon both decision latency and throughput.
+\approachshort{} has shown how classical RL techniques make online learning possible by simplifying update logic and enabling parallel processing.
+In-NIC use of these algorithms enabled a \SIrange{15}{21}{$\times$} reduction in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput compared to host offloading.
+The deployment environment and asynchronous design were shown to eliminate PCIe delays and impose minimal impact on carried dataplane traffic.
+We also showed how \approachshort{} scales with additional compute resources at deployment to improve on decision latency and throughput.
 Our throughput-optimal design, \Indfw{}, improves upon these metrics with \emph{just one worker}.

### 2021-06-09 paper/paper.tex
Adversarial LaTeX

@@
 
-\setlength{\textfloatsep}{6.0pt plus 2.0pt minus 4.0pt}
-\setlength{\floatsep}{6.0pt plus 1.0pt minus 2.0pt}
-
-\setlength{\dbltextfloatsep}{6.0pt plus 2.0pt minus 4.0pt}
-\setlength{\dblfloatsep}{6.0pt plus 1.0pt minus 2.0pt}
+\setlength{\textfloatsep}{4.0pt plus 2.0pt minus 4.0pt}
+\setlength{\floatsep}{4.0pt plus 1.0pt minus 2.0pt}
+
+\setlength{\dbltextfloatsep}{4.0pt plus 2.0pt minus 4.0pt}
+\setlength{\dblfloatsep}{4.0pt plus 1.0pt minus 2.0pt}
 
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.75\linewidth]{figures/arch-with-p4}
+	\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/arch-with-p4}
 	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
@@
 While the P4 language~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to great interest in network programmability, it requires similar behaviour between device classes.
-This is encoded as the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}---a sensible lower bound on the device capabilities required to implement a P4 dataplane.
+This is encoded by the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}---a sensible lower bound on the device capabilities required to implement a P4 dataplane.
 We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
@@
 Currently, low port density devices like the above are most likely to have specific general-purpose compute, as they are designed to suit high-performance offloading and middlebox development.
-Even Intel Tofino~\parencite{barefoot-intel} ASICs, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
+Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
 Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of deep learning accelerators (which, in turn, lack packet switching functionality).
@@
 Consider \cref{fig:state-slip}: the traditional RL formulation assumes that an agent receives new state at fixed time intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
-This adds noise to the state-action mapping being learned, which has harms learning rate and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
-Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
-This principle has influenced the design of real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, showing that asynchrony is necessary for high-speed applications.
+This adds noise to the state-action mapping being learned, which harms learning rate and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
+%Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
+Reducing these costs requires detailed agent-environment co-design.
+This principle has influenced the design of real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, where asynchrony is necessary for high-speed applications.
 This often requires that state measurements are fused~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
-`Stopping the world' causes significant performance degradation, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control, or any time-sensitive control problems.
+`Stopping the world' causes significant performance loss, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control, or any time-sensitive control problems.
 
@@
 %Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the computational cost of gradient computation via backpropagation remains too significant on embedded hardware.
-Complex \emph{Deep Neural Networks} (DNNs), particularly when used as the basis for an RL agent's policy, require vast amounts (e.g., in the worst case compute-years~\parencite{DBLP:journals/corr/abs-1912-06680}) of experience to converge on an accurate parameter set, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
+Complex \emph{Deep Neural Networks} (DNNs), particularly when used as an RL agent's policy, require vast amounts (possibly compute-years~\parencite{DBLP:journals/corr/abs-1912-06680}) of experience to converge on an accurate parameter set, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
 Although there has been a vested interest in efficiently \emph{executing} complex function approximators like DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the storage and computational costs of backpropagation remain too significant on embedded hardware.
@@
 
-For instance, a policy with $m$ sets of tiles (each having its own set of input dimensions), each containing $n$ overlapping tilings then creates $m \times n$ separate tasks.\footnote{Retrieving each individual action's value can be considered a work item (giving $m \times n \times a$ tasks when there are $a$ actions).
-However, this requires the tile coding operation to repeated $a$ times per tile, which is likely extremely wasteful outside of extreme degrees of parallelism.}
+For instance, a policy with $m$ sets of tiles (each having its own set of input dimensions), each containing $n$ overlapping tilings then creates $m \times n$ separate tasks.\footnote{Retrieving each individual action's value can be considered a work item (giving $m \times n \times a$ tasks for $a$ actions).
+However, this requires tile coding to be repeated $a$ times per tile, which is likely extremely wasteful outside of heavy parallelism.}
 As an example, a real-world policy size which we examine later (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets containing \num{8} tilings creates \num{129} work items.
-
 Computing a tile hit requires one division per dimension, which is an expensive addition to the ALU operations we require from a platform.
 However, if the width of tiles are fixed at powers-of-two, then these may be replaced with right bitshifts.
-As we shall see shortly, this aligns with our choice of numerical representation.
+%As we shall see shortly, this aligns with our choice of numerical representation.
 
@@
 For instance, other cores can be introduced by designing additional off-path functional units,  like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-Rather than general-purpose IPC, an RL functional unit would be hard-wired between internal workers and to the custom actions that interface with \approachshort{}, using the P4$\rightarrow$NetFPGA toolchain to offer the base framework and granular control over packet and flow selection.
+Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the custom actions that interface with it, using the P4$\rightarrow$NetFPGA toolchain to offer granular control over packet and flow selection.
 %Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
-As our later results show, using a bit depth below the native word size forces the compiler or programmer to emit excess code to extract and re-align preference values.
-This reduces overall efficiency even though considerably fewer reads are made.
-We expect that more bespoke hardware/FPGA co-design would allow this boiler-plate to be removed and enable SIMD-like optimisations we discuss without creative and costly bit-stuffing.
+As our later results show, using a bit depth below the native word size forces the compiler to emit excess code to extract and re-align preference values, reducing overall efficiency.
+%This reduces overall efficiency even though considerably fewer reads are made.
+We expect that a more bespoke hardware/FPGA design would allow this to match the desired bit depth and enable SIMD-like optimisations we discuss without creative and costly bit-stuffing.
 
@@
 
-We run \approachshort{} on an island with \num{4} MEs of the NFP-6480, totalling \num{32} hardware threads.
-This is the maximum number of MEs on a single island which is not in use by a P4 pipeline.
+We run \approachshort{} on a\num{4}-ME island of the NFP-6480, totalling \num{32} contexts.
+This is the largest cluster of cores which is not in use by a P4 pipeline.
 Where feasible, we test \SIlist{32;16;8}{\bit} quantised arithmetic.
-All \approachshort{} timing measurements were repeated over \num{10000} input state packets (preceded by \num{1000} warmup packets), retrieving each item processing time over PCIe via the controller machine, from which throughput was derived.
+All \approachshort{} timing measurements were repeated over \num{10000} state packets (preceded by \num{1000} warmup packets), retrieving item processing times over PCIe via the controller machine, from which throughput was derived.
 Host throughput and latency measures were observed over \num{10} trials of \SI{10}{\second} (with \SI{5}{\second} warmup/cooldown times).
@@
 \fakepara{Raw inference and learning performance}
-We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs against floating point (numpy-based) implementations on a commodity host machine.
+We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs against floating point (numpy-based) implementations of Sarsa on a commodity host machine.
 This allows us to demonstrate the performance differences between the \indfw{} and \coopfw{} configurations, particularly in how \indfw's required policy locks impact throughput.
@@
 \fakepara{Raw inference and learning performance}
-\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based implementation of an RL policy\footnote{For brevity, we omit numpy-based integer results---summarised against a floating point implementation, median state-action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}.
+\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy\footnote{For brevity, we omit numpy-based integer results---summarised against a floating point implementation, median state-action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}.
 We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using 4 MEs of the NFP-6480.
@@
 \newlength{\resultplotwidth}
-\setlength{\resultplotwidth}{0.85\linewidth}
+\setlength{\resultplotwidth}{0.9\linewidth}
 
@@
 \Cref{tab:tputs} compares \approachshort{}'s throughput against host-based execution.
-We chose the worker count on host machines that maximised their throughput.
+We set the worker count on host machines which maximised their throughput.
 This equalled the amount of physical cores on each device---moving beyond this (even below the number of hyper-threads) would hamper tail latencies by an order of magnitude.
@@
 	\caption{
-		\Coopfw{}'s online learning performance improves as additional cores are made available, on max size tasks (\num{129} work items).
+		\Coopfw{}'s online learning performance improves with additional cores, on max size tasks (\num{129} work items).
 This requires \num{8} workers to offer greater online throughput than single-threaded in-NIC RL.
@@
 	}
-	\caption{\Coopfw{}'s action latencies improve with additional cores.
+	\caption{\Coopfw{}'s action latencies similarly improve with additional cores.
 This requires \num{3} workers (4 total contexts) to improve upon the state-action latency of single-threaded in-NIC RL.\label{fig:vary-core-latency}}
@@
 \fakepara{Resource requirements}
-\Cref{tab:resources} demonstrates how \approachshort{} consumes shared memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
-As one program is installed per ME, these results represent the minimum and maximum resource use we can command on a single island (i.e., without replacing P4 workers).
-We observe negligible costs on shared bulk EMEM storage ($\sim$\SI{4}{\mebi\byte}), incurred due to hashtables for state-action pairs and rewards.
+\Cref{tab:resources} shows how \approachshort{} consumes shared memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
+As one program is installed per ME, these results represent the minimum and maximum resource use on a single island (i.e., without replacing P4 workers).
+We observe negligible costs on shared EMEM ($\sim$\SI{4}{\mebi\byte}), incurred due to hashtables for past state and rewards.
 The most significant costs arise due to policy data (\SI{405}{\kibi\byte} shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
-This causes a high upfront cost on per-island resources (CLS/CTM)---\approachshort{} leaves resources for other off-path dataplane applications, but is fairest from 3 cores onwards.
+This is a high upfront cost on per-island resources (CLS/CTM)---\approachshort{} leaves resources for other off-path dataplane applications, but is fairest from 3 cores onwards.
 
@@
 As we note in \cref{sec:limitations}, these actions would be placed into a small hash table and simultaneously exported to the attached controller to be inserted as P4 rules in batches.
-After the main monitoring step, packet ingress timestamps would be used to emulate the \emph{Timed Random Sequential} (TRS) scheduler used by the anti-DDoS agents for rate-controlled work, where state vectors would be selected and passed into \approachshort.
+Afterwards, packet ingress timestamps would be used to emulate the \emph{Timed Random Sequential} (TRS) scheduler used by the anti-DDoS agents for rate-controlled work, where state vectors would be selected and passed into \approachshort.
 By design, active flows which are not \emph{judged} after a configurable time are discarded to prune the work set and allow new flows to be seen by the agent.

### 2021-06-09 paper/paper.tex
Two More Paras

@@
 
-\setlength{\textfloatsep}{4.0pt plus 2.0pt minus 4.0pt}
-\setlength{\floatsep}{4.0pt plus 1.0pt minus 2.0pt}
-
-\setlength{\dbltextfloatsep}{4.0pt plus 2.0pt minus 4.0pt}
-\setlength{\dblfloatsep}{4.0pt plus 1.0pt minus 2.0pt}
+\setlength{\textfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
+\setlength{\floatsep}{3.0pt plus 0.0pt minus 2.0pt}
+
+\setlength{\dbltextfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
+\setlength{\dblfloatsep}{3.0pt plus 0.0pt minus 2.0pt}
 
@@
 We discuss recent trends in programmable switch hardware.
-By combining this with insights from the ML/RL communities (past and present), we discuss why in-NIC RL is best-placed to interact with the network, and how classical RL methods and quantisation ensure computational feasibility.
+By combining this with insights from the ML/RL communities (past and present), we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL methods and quantisation make this computationally feasible.
 
@@
 		\end{tikzpicture}}
-		\caption{Reality: costs of measurement and action lead to \emph{state drift}---over a time delay $t_1+t_2+t_3$, inaction transforms state $S$ into $S'$.}
+		\caption{Reality: measurement and action costs lead to \emph{state drift}---over a delay $t_1+t_2+t_3$, inaction transforms state $S$ into $S'$.}
 	\end{subfigure}
@@
 As an example, a real-world policy size which we examine later (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets containing \num{8} tilings creates \num{129} work items.
-Computing a tile hit requires one division per dimension, which is an expensive addition to the ALU operations we require from a platform.
-However, if the width of tiles are fixed at powers-of-two, then these may be replaced with right bitshifts.
+Computing a tile hit requires one division per dimension, which is an expensive ALU operation to require.
+However, if tile widths are fixed at powers-of-two, then these may be replaced with right bitshifts.
 %As we shall see shortly, this aligns with our choice of numerical representation.
@@
 Internally, \approachshort{} either has all its cores act independently (\cref{fig:single-and-parallel:single}) or cooperate to solve each task (\cref{fig:single-and-parallel:parallel})---with different latency-throughput benefits (\cref{sec:action-and-update-computation}).
-We open-source our firmware design and control programs for the benefit of the community.\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission.}}
-Moving beyond the Netronome platform, we describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.
+We open-source our firmware and control programs for the benefit of the community.\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission.}}
+Moving beyond the Netronome platform, we describe how our architecture may be adapted and improved on by bespoke hardware or FPGA-based deployment.
 
@@
 
-To protect traffic throughput and allow for effective deployment in as many environments as possible, \approachshort{} places RL execution on-chip, \emph{but off the main packet path}, communicating and running parallel to the main P4 dataplane.
+To protect traffic throughput and allow effective deployment in as many environments as possible, \approachshort{} places RL execution on-chip, \emph{but off the main packet path}, communicating and running parallel to the main P4 dataplane.
 As shown in \cref{fig:netro-arch}, this asynchrony allows coexistence with P4 programs, and imposes minimal impact on carried traffic for both bump-in-the-wire deployments and at end-points.
-For instance, in the default deployment of a P4 packet processing pipeline on Netronome NFP chips several of these cores go unused (as does spare area on an FPGA design), making this paradigm possible.
+For instance, in the default deployment of a P4 packet processing pipeline on Netronome NFP chips several cores go unused (as does spare area on an FPGA design), making this paradigm possible.
 
@@
 
-Two main firmware models govern how the computation-intensive parts of these tasks (tile-coding, action preference list construction, policy updates) are carried out:
+Two firmware models govern how the compute-heavy parts of these tasks (action selection, policy updates) are carried out:
 \begin{description}
@@
 \subsection{Agent-Environment Communication}\label{sec:agent-environment-communication}
-\approachshort{} uses \emph{multiple-producer/multiple-consumer} (MPMC) messaging channels to communicate with other computational elements; be they P4 programs on the packet path, or additional on-chip analysis and control modules.
-Through these channels the system \emph{pushes} state vectors, reward measures, and control/setup packets as its inputs, and \emph{pulls} a stream of state-action pairs as its outputs.
-This allows decisions to be made asynchronously---preventing packet stalling---and allowing many parallel decision-making agents to be used if desired.
+\approachshort{} uses \emph{multiple-producer/multiple-consumer} (MPMC) messaging channels to communicate with other elements; be they P4 programs on the packet path, or other on-chip analysis and control modules.
+Through these channels a system \emph{pushes} state vectors, reward measures, and setup packets as inputs, and \emph{pulls} a stream of state-action pairs as outputs.
+This allows decisions to be made asynchronously---preventing packet stalling---and allowing many RL agents to be used if desired.
 The key insight of this mechanism is that on-chip reward/state signals enjoy first-class support in the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-NIC/chip asynchronous processes or the controller.
@@
 
-Our implementation uses platform-specific IPC (\emph{EMEM ring buffers}) with hardware signalling mechanisms on work arrival to achieve this.
+Our implementation uses platform-specific IPC (\emph{EMEM ring buffers}) with hardware signalling on work arrival to achieve this.
 As PDP hardware lacks dynamic memory allocation, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state/configuration/policy data handover.
@@
 Even with parallel problems such as \emph{ParSa}, optimising for latency requires meticulous care in how work is passed out and aggregated.
-This is truer still when moving from the moderately fine-grained control of classical RL methods ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
-Ordinarily, the marshalling of requests, responses, and safe shared data access can incur significant extra overheads.
+This is truer still when moving from the moderately-grained control of classical RL ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
+Ordinarily, the marshalling of requests, responses, and shared data access can incur significant overheads.
 On-chip execution and the nature of action preference computation allow us to use lockless atomic aggregation, removing the overheads of explicit messaging/packetisation.
-Moreover, physically adjacent functional units/cores often have special-purpose shared registers or share a small fast cache to accelerate communication.
+Moreover, adjacent functional units/cores often have special-purpose shared registers or share a small fast cache to accelerate communication.
 
@@
 
-Our implementation exploits the locality of cores in the NFP architecture, which can be used on other platforms like NetFPGA.
+Our implementation exploits the locality of cores in the NFP, which can be used on other platforms like NetFPGA.
 Policy compute/update/configure tasks are passed between cores using direct \emph{next neighbour} registers, signalling all child threads in response.
 Such on-chip signals cost just $\sim$\SI{20}{\nano\second} per relayed message.
-Each core performs atomic adds to a shared preference list and an acknowledgement counter checked by the master thread, implementing the wait-free \emph{ParSa} algorithm we introduce (\cref{alg:parsa}).
+Each core performs atomic adds to a shared preference list and an acknowledgement counter checked by the master thread, implementing our wait-free \emph{ParSa} algorithm (\cref{alg:parsa}).
 This is essential for aggregation compared to the use of bounded message buffers, which caused significant head-of-line blocking.
@@
 \subsection{Reconfigurability}
-\approachshort{} allows policy design and algorithmic learning parameters ($\alpha, \gamma, \epsilon$, related falloffs) to be changed at runtime using at most two control packets.
-For instance, design changes are useful at the end of learning (moving from online to offline), or when trying to train a policy for another problem afterwards from the same vantage point.
-Parameter changes are useful when an online agent must become more (or less) adaptive to new data (i.e., after a changepoint in traffic is detected).
-This extends to policy data, which may be imported from an offline pre-trained model via such packets and exported via PCIe to the host machine.
-Some aspects must be chosen at compile time; bit depth, work allocation strategy, and maximum policy/tiling/state sizes---these govern core operation or allocated memory.
+\approachshort{} allows policy design and algorithmic learning parameters to be changed at runtime using at most two control packets.
+For instance, design changes are useful at the end of learning (moving from online to offline), or when trying to train a new policy for another problem from the same vantage point.
+Parameter changes are useful when an online agent must become more (or less) adaptive to new data (i.e., after detecting a changepoint in traffic).
+This extends to policy data, which may be imported from a pre-trained model via such packets and exported via PCIe to the host machine.
+Some aspects must be chosen at compile time; bit depth, \Coopfw/\Indfw, and maximum policy/tiling/state sizes---these govern core operation or allocated memory.
 Choosing a bit depth of \qtylist[list-pair-separator = { or }]{16;8}{\bit} halves/quarters policy memory costs, allowing more complex problems to be modelled using more dimensions or fine-grained tiles.
@@
 In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved (pool 2~\parencite{rfc2474}) DSCP value as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-While we primarily use this to automate parser generation, this also allows for configuration packets to be received from only trusted hosts (over the dataplane if needed) via P4 rules.
-Our control packet generation library (and evaluation frameworks which build upon it) are written in Rust.
+While this mainly automates parser generation, it also allows for configuration to be received from only trusted hosts (over the dataplane if needed) via P4 rules.
+Our control packet generation library and evaluation frameworks which build upon it are written in Rust.
 
@@
 \subsection{Work Allocation}\label{sec:work-allocation}
-Due to the lack of dynamic memory allocation on PDP hardware, and to simplify action-value lookups, policies cannot be stored sparsely.
+Due to the lack of dynamic memory allocation on PDP hardware, and to simplify value lookups, policies cannot be stored sparsely.
 As a consequence, tiling space requirements scale exponentially with dimension count, and so higher-dimension tilings must be placed in larger (and slower) memory regions.
@@
 
-We use a simple first-fit work placement algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core (as the NFP exposes hardware threads).
-Each work item is a separate \emph{tiling} over a number of dimensions, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
-The cost of any individual work item (according to its dimension count and memory location) was empirically measured offline, and we weigh the total cost per core to account for the number of minion threads available.
-This weighting specifically accounts for the presence of the controller thread on the first core.
-Work allocation and any cached per-item data are recomputed when any configuration is installed or changed.
-Naturally, for $n$ tilings/tasks and $m$ threads this procedure is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per task, storing $m/8$ and $\le8$ costs respectively.
+We use a simple first-fit work placement algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core (as the NFP uses hardware threads).
+Each work item is a separate \emph{tiling} over a list of dimensions, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
+The cost of any work item (from its dimension count and memory location) was empirically measured offline, and we weigh the total cost per core based on the number of minion threads available.
+This weighting specifically accounts for the controller thread on the first core.
+Work allocation and cached per-item data are recomputed when policy configuration is installed or changed.
+Naturally, for $n$ tilings and $m$ threads this procedure is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per tiling, storing $m/8$ and $\le8$ costs respectively.
 
@@
 
-As is expected of parallel algorithms, efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
+%As is expected of parallel algorithms, efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
+Efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
 Even in wait-free algorithms, this requires a minimum number of workers to improve upon the latency bounds of a serial approach.
@@
 \subsection{Targeting other device classes}
-While specifics of this design cater to Netronome devices, many aspects have clear analogues in NICs of similar form-factor, such as the NetFPGA SUME platform.
-For instance, other cores can be introduced by designing additional off-path functional units,  like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the custom actions that interface with it, using the P4$\rightarrow$NetFPGA toolchain to offer granular control over packet and flow selection.
+While this design caters to Netronome devices, many aspects have clear analogues in other SmartNICs, such as the NetFPGA SUME.
+For instance, additional off-path functional units can serve as workers, like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the P4$\rightarrow$NetFPGA~\parencite{DBLP:conf/fpga/IbanezBMZ19} actions that interface with it.
 %Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
-As our later results show, using a bit depth below the native word size forces the compiler to emit excess code to extract and re-align preference values, reducing overall efficiency.
+As we show later, using a bit depth below the native word size reduces overall efficiency.
 %This reduces overall efficiency even though considerably fewer reads are made.
-We expect that a more bespoke hardware/FPGA design would allow this to match the desired bit depth and enable SIMD-like optimisations we discuss without creative and costly bit-stuffing.
-
-Bringing \approachshort{} to high-port density devices such as Tofino-powered switches is more difficult.
-As discussed in \cref{sec:motivation}, Tofino ASICs map closely to the P4 PSA, meaning that there are no spare asynchronous general-purpose compute units to place this work on.
-Taking inspiration from real-time programming techniques, a potential solution is to divide RL actions across several received packets (i.e., iteratively computing a portion of the action preference list each time) until any further work would delay outbound transmission.
+We expect an FPGA design would allow this to match the desired bit depth and enable the SIMD-like optimisations we discuss without creative and costly bit-stuffing.
+
+Bringing \approachshort{} to high-port density devices such as Tofino switches is more difficult as they map closely to the P4-PSA, leaving no spare general-purpose compute units.
+Taking inspiration from real-time programming, a potential solution is to divide RL processing across several packets (i.e., computing a portion of the action preference list each time) until any further work would delay outbound transmission.
 This, however, would introduce new issues surrounding concurrent accesses, work splitting, and altered timescales for learning: we leave their treatment and examination to future work.
+
+%While specifics of this design cater to Netronome devices, many aspects have clear analogues in NICs of similar form-factor, such as the NetFPGA SUME platform.
+%For instance, other cores can be introduced by designing additional off-path functional units,  like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+%Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the custom actions that interface with it, using the P4$\rightarrow$NetFPGA toolchain to offer granular control over packet and flow selection.
+%%Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
+%As our later results show, using a bit depth below the native word size forces the compiler to emit excess code to extract and re-align preference values, reducing overall efficiency.
+%%This reduces overall efficiency even though considerably fewer reads are made.
+%We expect that a more bespoke hardware/FPGA design would allow this to match the desired bit depth and enable SIMD-like optimisations we discuss without creative and costly bit-stuffing.
+%
+%Bringing \approachshort{} to high-port density devices such as Tofino-powered switches is more difficult.
+%As discussed in \cref{sec:motivation}, Tofino ASICs map closely to the P4 PSA, meaning that there are no spare asynchronous general-purpose compute units to place this work on.
+%Taking inspiration from real-time programming techniques, a potential solution is to divide RL actions across several received packets (i.e., iteratively computing a portion of the action preference list each time) until any further work would delay outbound transmission.
+%This, however, would introduce new issues surrounding concurrent accesses, work splitting, and altered timescales for learning: we leave their treatment and examination to future work.
 
@@
 We investigate the performance of \approachshort{} compared to classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{$\times$} speedup in median--\num{99.99}\nthscript{th} state-action latency and \SI{9.9}{$\times$} greater online learning throughput.
-Crucially, we find that in-NIC execution offers tight tail latency bounds compared to host-based approaches.
-We report on how \approachshort{} scales to fit available device resources as cores/threads are added, noting that both \Indfw{} and \Coopfw{} designs outperform commodity hosts using just one core in latency and online throughput.
+Crucially, in-NIC execution offers tight tail latency bounds compared to host-based approaches.
+We report on how \approachshort{} scales as additional device resources are added, noting that both in-NIC designs outperform commodity hosts using just one core in latency and online throughput.
 Furthermore, \Indfw{} provides higher per-core offline throughput than host-based approaches, even though our measured hosts exhibit higher clock speeds.
@@
 
-We run \approachshort{} on a\num{4}-ME island of the NFP-6480, totalling \num{32} contexts.
+We run \approachshort{} on a \num{4}-ME island of the NFP-6480, totalling \num{32} contexts.
 This is the largest cluster of cores which is not in use by a P4 pipeline.
@@
 
-Policy sizes are set to those matching a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dimension state vectors, a bias tile and 16 full tiling sets (\numproduct{7x1}-dim, \numproduct{8x2}-dim, \numproduct{1x4}-dim), 8 tilings per dimension set, 6 tiles per dimension, choosing between 10 actions.
-For context, such input state would contain various aspects of per-flow state (e.g., IATs, up/down rate) which are combined with other state such as the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim).
-In \Coopfw{} for instance, this creates \num{129} work items across \num{31} workers.
+Policy sizes are set to those of a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dim state vectors, a bias tile and 16 full tiling sets (\numproduct{7x1}-dim, \numproduct{8x2}-dim, \numproduct{1x4}-dim), 8 tilings per set, 6 tiles per dimension, and 10 actions.
+For context, such input state would contain various aspects of per-flow state (e.g., IATs, rates) which are combined with other state such as the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim).
+In \Coopfw{}, this creates \num{129} work items across \num{31} workers.
 Although the model application's more successful \emph{SPF} agent design uses 3 actions, we choose a larger action set to investigate the performance of more complex agents.
@@
 We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs against floating point (numpy-based) implementations of Sarsa on a commodity host machine.
-This allows us to demonstrate the performance differences between the \indfw{} and \coopfw{} configurations, particularly in how \indfw's required policy locks impact throughput.
+This allows us to demonstrate the performance differences between the \indfw{} and \coopfw{} configurations, particularly in how \indfw's (and hosts') required policy locks impact throughput.
 %We also report on the derived throughput measures accounting for the locking behaviour of each.
@@
 
-Building on this, we vary the amount of allocated worker threads to show how \approachshort{} can scale to fit available compute resources on a device.
-This is an important aspect for (automated) administration and allocation of compute resources in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of-time work scheduling which we examine later.
+Building on this, we vary the amount of worker threads to show how \approachshort{} scales to fit available compute resources on a device.
+This is an important aspect for (automated) allocation of compute resources in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of-time work scheduling which we examine later.
 This also demonstrates the number of cores needed to achieve a given latency/throughput bound on a policy of representative complexity.
@@
 \newlength{\resultplotwidth}
-\setlength{\resultplotwidth}{0.9\linewidth}
+\setlength{\resultplotwidth}{0.88\linewidth}
 
@@
 
-\begin{figure}
+\begin{figure}[t]
 	\resizebox{\resultplotwidth}{!}{
@@
 	}
-	\caption{\Coopfw{}'s action latencies similarly improve with additional cores.
+	\caption{\Coopfw{}'s action latencies similarly improve with more cores.
 This requires \num{3} workers (4 total contexts) to improve upon the state-action latency of single-threaded in-NIC RL.\label{fig:vary-core-latency}}
@@
 We note that this is a trivial case, as using \emph{only} a bias tile returns a single preference list regardless of input state.
-
-\begin{figure}
-	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-work}
-	}
-	\caption{\Coopfw{} fully processes updates faster than \Indfw{}---thus has higher online performance---on almost all policy sizes.
-Lower bit depths are more effective on simpler policies.\label{fig:vary-work}}
-\end{figure}
 
@@
 We show the cumulative distributions of these in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF/traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
+
+\begin{figure}[t]
+	\resizebox{\resultplotwidth}{!}{
+		\includegraphics{../plots/build/rl-perf-tester/vary-work}
+	}
+	\caption{\Coopfw{} fully processes updates faster than \Indfw{}---thus has higher online performance---on almost all policy sizes.
+Lower bit depths are more effective on simpler policies.\label{fig:vary-work}}
+\end{figure}
 
@@
 
-A noteworthy trend is that the \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit} designs, with the exception of smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
-This occurs even though our implementation is optimised to read and write policy tile data in batches (achieving fewer I/O operations as required).
+A noteworthy trend is that \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
+This occurs even though our implementation is optimised to read and write policy data in batches (achieving fewer I/O operations).
 We see this because the native register width on the NFP is \SI{32}{\bit}, and so the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
-This matches \SI{32}{\bit} becoming dominant in complex workloads: higher dimension tilings require more arithmetic operations to compute.
+This matches \SI{32}{\bit} becoming dominant in complex workloads: higher dimension tilings require more arithmetic operations.
 Most of the I/O comes \emph{after} this step, causing ALU use to dominate.
 This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the online algorithm.
-Smaller bit depths still give a \SIrange{2}{4}{$\times$} saving in memory for policy storage (enabling more fine-grained or more complex policies), in exchange for slightly worse latency and throughput.
-To overcome this, we investigated bit-stuffing several values into a single word for our atomic writeback mechanism (as the platform offers both \SIlist{32;64
+Smaller bit depths still give a \SIrange{2}{4}{$\times$} saving in memory for policy storage (enabling more fine-grained or complex policies), in exchange for slightly worse latency and throughput.
+To overcome this, we investigated bit-stuffing several values into a single word during atomic writeback (as the platform offers both \SIlist{32;64
 }{\bit} atomic addition).
@@
 The trend is repeated for all bit depths.
-The constant difference between \emph{Action} and \emph{Update Prep} was observed to scale linearly with bit depth, matching storage and lookup work in the serial portion of \emph{ParSa} (plots omitted).
-The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its index, as batching work in contiguous chunks leads to some workers receiving exclusively high-dimensionality tilings.
+The constant difference between \emph{Action} and \emph{Update Prep} scales linearly with bit depth, matching storage and lookup work in the serial portion of \emph{ParSa} (plots omitted).
+The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its index, as batching work in contiguous chunks gives some workers only high-dimensionality tilings.
 The minor gap in lower bound performance between the \emph{Random} and \emph{Balanced} allocators suggests that further optimisations can be made.
@@
 
-An interesting aspect of \Coopfw{} and the \emph{ParSa} algorithm in general is that adding cores has both diminishing returns and key thresholds to pass.
+An interesting aspect of \Coopfw{} and \emph{ParSa} is that adding cores has both diminishing returns and key thresholds to pass.
 Consider \cref{fig:tput-per-core}, where the throughput per worker decreases over time but occasionally increases sharply.
 Later downward ticks (\numrange{25}{29} workers) correspond to plateaus in throughput.
-This is a problem stemming from the granularity of work items (i.e., tilings in \emph{ParSa}), where the heuristic scheduler cannot find a better solution to a bottleneck until extra cores are allocated.
+This is a problem stemming from the granularity of work items (i.e., tilings in \emph{ParSa}), where our scheduler cannot find a better solution to a bottleneck until extra cores are allocated.
 We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
-While there is still a \SI{4.2}{$\times$} factor of task oversubscription at maximum worker count, it is clear that latencies are bound below by the length of the longest task.
+Though we have a \SI{4.2}{$\times$} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
 
@@
 \fakepara{PDP design for asynchronous compute}
-\Textcite{DBLP:conf/hotnets/StephensAS18} have proposed \emph{PANIC}, which places a routing fabric between distinct packet/data processing elements \emph{in a SmartNIC}.
-Such designs would more generally enable asynchronous and novel compute in SmartNICs and programmable switches, for instance offering consistent and easy to use communication between workers versus hard-coded ME relationships.
-Event-driven versions of the P4 pipeline have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}.
+\emph{PANIC}~\parencite{DBLP:conf/hotnets/StephensAS18} places a routing fabric between distinct packet/data processing elements \emph{in a SmartNIC}.
+Such designs would enable general, asynchronous, and novel compute in SmartNICs and switches, for instance offering consistent and easy to use communication between workers versus hard-coded ME relationships.
+Event-driven versions of P4 have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}.
 Timer events and device state changes would empower in-network RL use-cases, signalling timesteps for RL agents or new, effective, fine-grained sources of input state.

### 2021-06-09 paper/paper.tex
12 pages, 9pt
Depending on TPC response, argh

@@
 %,anonymous=true
+,9pt
 ]{acmart}
 
-\usepackage[false]{anonymous-acm}
+\usepackage[%
+false
+]{anonymous-acm}
 
@@
 
-Bringing \approachshort{} to high-port density devices such as Tofino switches is more difficult as they map closely to the P4-PSA, leaving no spare general-purpose compute units.
-Taking inspiration from real-time programming, a potential solution is to divide RL processing across several packets (i.e., computing a portion of the action preference list each time) until any further work would delay outbound transmission.
-This, however, would introduce new issues surrounding concurrent accesses, work splitting, and altered timescales for learning: we leave their treatment and examination to future work.
+Bringing \approachshort{} to PDP switches such as the Tofino is difficult as they closely match the P4-PSA, leaving no spare general-purpose compute units.
+Taking inspiration from real-time programming, a potential solution is to divide RL processing across several packets (i.e., computing a portion of the preference list each time) until further work would delay outbound transmission.
+This would introduce new issues in concurrent access, work splitting, and altered timescales for learning: we leave their treatment to future work.
 
@@
 \fakepara{Raw inference and learning performance}
-\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy\footnote{For brevity, we omit numpy-based integer results---summarised against a floating point implementation, median state-action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}.
+\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy.\footnote{For brevity, we omit numpy-based integer results---against a floating point implementation, median action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}
 We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using 4 MEs of the NFP-6480.
@@
 By limiting the available workers in software, we show how \Coopfw{}'s policy update time (thus  online throughput---\cref{fig:vary-core}) and state-action latency (\cref{fig:vary-core-latency}) scale with available cores.
-While the compute times and latency of \Coopfw{} always outperform the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency.
+While \Coopfw{} always outperforms the host-based floating point implementations, we observe that there are two distinct crossover points to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency.
 Some artefacts of our environment and design choices are visible, such as the addition of new physical cores being more significant than contexts, and the presence of some schedule bottlenecks.
-Most importantly however, \approachshort{}-\Coopfw{}'s resource demand is tunable at compile time to meet the online training rate and/or action latency required by a task/environment.
+Most importantly however, \Coopfw{}'s resource demand is tunable at compile time to meet the online training rate and/or action latency required by a task/environment.
 
@@
 
-\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects update cost and state-action latency respectively, scaling from a bias tile up to the full DDoS policy size\footnote{Input vectors here all have 20 elements regardless of the policy design.}.
+\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects update cost and state-action latency respectively, scaling from a bias tile up to the full DDoS policy size.\footnote{Input vectors here all have 20 elements regardless of the policy design.}
 \Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning.
@@
 			\includegraphics{../plots/build/rl-perf-tester/latency-cumul}
+			\vspace{-10pt}
 		}
@@
 This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---the NFP P4 toolchain segments packets, storing metadata (e.g., MAC prepend) and the first bytes of a packet in a \SI{256}{\byte} CTM block and parking their payloads in EMEM.
-\SI{256}{\byte} packets overshoot this due to metadata, causing small I/O accesses at a high rate for packets sized around this cutoff point.
-Regardless, this effect is small when observed.
+\SI{256}{\byte} packets overshoot this due to metadata, causing small I/O accesses at a high rate for packets sized around this cutoff.
+%Regardless, this effect is small when observed.
 
@@
 \fakepara{Deployability}
-Setup of \approachshort{} uses two packet types: \emph{setup}, which contains learning (hyper-)parameters and most aspects of a tile coded policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
-We found that setup and tiling packets take a mean \qtylist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing for an agent to be swapped between online and offline operation (or repurposed for another task/policy) painlessly.
-Online/offline swaps are useful when an agent should cease learning (i.e., when performance has converged to an acceptable value), or if a detected change in the environment means that more training is needed.
+Setup of \approachshort{} uses two packet types: \emph{setup}, which contains learning (hyper-)parameters and most aspects of a policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
+We found that setup and tiling packets take a mean \qtylist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing an agent to be swapped between online and offline operation (or repurposed for another task) painlessly.
+Online/offline swaps are useful when an agent should cease learning (i.e., when performance has converged), or if a change in the environment suggests that more training is needed.
 Tiling packet processing scales linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
@@
 \section{Potential Integrations}\label{sec:potential-integrations}
-To show the general applicability and use of \approachshort{}, we propose an ideal integration which would benefit from in-NIC reinforcement learning; online DDoS attack mitigation.
+To show the applicability and use of \approachshort{}, we propose an ideal integration which would benefit from in-NIC RL; online DDoS attack mitigation.
 We support this using other state-of-the-art P4/PDP developments, and discuss how best to balance the concerns of online training with throughput in widespread deployment.
-Additionally, we discuss how network administrators may combine offline (\Indfw) and online (\Coopfw) agents within a network to achieve online learning while maximising action throughput in the majority of the network.
+Additionally, we discuss how network administrators may combine offline (\Indfw) and online (\Coopfw) agents to achieve online learning while maximising action throughput in the majority of the network.
 
@@
 \fakepara{Why in-NIC?}
-The above approach relies on co-hosting traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
-However, packet mirroring, offloading RL computation to a host (potentially over layers of virtualisation), and CPU performance are all sources of additional, consistent state-action latency.
+This approach relies on co-hosting traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
+However, packet mirroring and offloading RL computation to a host (potentially over layers of virtualisation) are all sources of additional, consistent state-action latency.
 Both traffic statistics collection \emph{and} data-driven learning must be executed on such hosts/network functions.
@@
 
-We note that the TRS scheduler is designed to handle cases where the number of flows created by an attacker is large, combining state vectors and information over time while the asynchronous agent is itself busy.
-An increased number of (attack) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
+We note that the TRS scheduler is designed to handle cases large numbers of flows created by an attacker, combining state vectors over time while the asynchronous agent is itself busy.
+A number of (attack) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
 As shown in \cref{sec:evaluation}, \approachshort{} far exceeds the throughput of host offloading.
-The control plane is then free to use wildcards or specific matches to narrow down or expand the set of flows to be considered and controlled, though explicit scheduling techniques are still key in such adversarial environments.
+The control plane can then use wildcards or specific matches to narrow down or expand the set of flows to be controlled, though explicit (TRS) scheduling is still key in such adversarial environments.
 
@@
 The two compute models discussed above need not be homogeneously deployed.
-For instance, in a networked deployment, a subset of \approachshort{} nodes could be \coopfw{} agents, training online, while most other nodes run \indfw{} designs to meet throughput guarantees.
-The control plane would then be responsible for combining, downsampling, and distributing these improved policies between remaining agents.
-This can be taken further still, using policy deltas or execution traces to enable out-of-path transfer learning for more complex function approximators such as neural networks.
+In a networked deployment, a subset of \approachshort{} nodes could be \coopfw{} agents, training online, while most other nodes run \indfw{} designs to meet throughput guarantees.
+The control plane would then combine, downsample, and distribute these improved policies between offline agents.
+This can be taken further still, using policy deltas or execution traces to enable out-of-path transfer learning for more complex models such as neural networks.
 
@@
 \fakepara{In-network ML}
-Taurus~\parencite{DBLP:journals/corr/abs-2002-08987} proposes that efficient line-rate, low-latency per-packet inference can be achieved by placing a run-time configurable grid of map-reduce units into the packet processing pipeline (implementing e.g., CNNs, LSTMs, and SVMs).
-Their CGRA-based implementation achieves sub-\si{\micro\second} extra latency, though online model training is not considered.
-
-\emph{IIsy}~\parencite{DBLP:conf/hotnets/XiongZ19} shows how \emph{classical ML inference} techniques (SVMs, Na\"{i}ve Bayes, Decision Trees, K-Means) can be converted into match-action tables compatible with \emph{any} P4 deployment.
-They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA devices (at line rate in most cases), but can only investigate pre-trained models.
+Taurus~\parencite{DBLP:journals/corr/abs-2002-08987} proposes that efficient line-rate inference can be achieved using a configurable grid of map-reduce units in the packet pipeline (implementing e.g., LSTMs and SVMs).
+On CGRA hardware, they achieve sub-\si{\micro\second} extra latency.
+\emph{IIsy}~\parencite{DBLP:conf/hotnets/XiongZ19} shows how \emph{classical ML inference} (SVMs, Na\"{i}ve Bayes, etc.) can be converted into match-action tables compatible with \emph{any} P4 deployment.
+They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA (at line rate in most cases).
+Neither considers online model training.
 
 A recent line of research in the community has been to investigate \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate packet classification.
-\emph{BaNaNa SPLIT} shows the initial steps of this trend as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} to reduce latency imposed by batching and bulk transfer to the GPU.
-However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}; the authors hypothesise that networking hardware would be a better fit for some parts of the inference pipeline.
-This has moved onto full, in-network packet tagging and classification by pre-trained BNNs with \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
+\emph{BaNaNa SPLIT} shows this as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} to reduce latency imposed by GPU batching and transfer, but fully-connected layers are but can be accelerated further by NICs.
+%However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}; the authors hypothesise that networking hardware would be a better fit for some parts of the inference pipeline.
+Full, in-network packet tagging and classification by pre-trained BNNs is shown by \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
 N3IC achieves packet inference in \SI{45}{\micro\second} on the NFP, and \SI{0.3}{\micro\second} on NetFPGA for \SI{256}{\bit} inputs.
@@
 
-\textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantised weights and inputs on the NFP, using \emph{in-path} compute rather than our asynchronous model.
+\textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantisation on the NFP, using \emph{in-path} compute rather than our asynchronous model.
 Inference latency on small networks (3 layers, \num{30} neurons) can be as high as \SI{500}{\micro\second} on line rate traffic, stressing the value of path-adjacent compute.
 
-We note that none of the other approaches listed here (or that we have seen) tackle the issue of \emph{online learning and control} in-network---we believe \approachshort{} has broken new ground in this regard.
+We stress that none of the other approaches listed here (or that we have seen) tackle the issue of \emph{online learning and control} in-network---we believe \approachshort{} has broken new ground in this regard.
 
@@
 \fakepara{RL for network control}
-\Textcite{DBLP:conf/sigcomm/LiangZJS19} use deep RL techniques on a host machine to train an agent which can build efficient decision tree packet classifiers.
-Their approach, \emph{NeuroCuts}, is able to produce time/space efficient classifiers for use in constrained environments (e.g., network hardware) beyond existing heuristic solutions.
+\Textcite{DBLP:conf/sigcomm/LiangZJS19} use deep RL techniques to train an agent which can build efficient decision tree packet classifiers.
+Their approach, \emph{NeuroCuts}, produces time/space efficient classifiers for use in constrained environments (e.g., network hardware) beyond existing heuristic solutions.
 Deep RL techniques have been used for QUIC congestion control optimisation~\parencite{DBLP:journals/corr/abs-1910-04054}.
@@
 
-In future, we aim to examine the relative performance of individual applications driven by \approachshort---both classical and deep RL-based---and how a NetFPGA implementation can offer further latency and throughput improvements.
-A promising avenue here would be to investigate constant transfer learning between a handful of online \approachshort{} agents and other high-throughput offline function approximators such as BNNs.
+In future, we aim to examine the performance of individual applications driven by \approachshort---both classical and deep RL-based---and how a NetFPGA implementation can offer further latency and throughput improvements.
+A promising avenue here would be to investigate constant transfer learning between online \approachshort{} agents and high-throughput offline function approximators such as BNNs.
 %It is our hope that high port-density programmable devices will, in time, expose similar asynchronous, on-chip offloading capability to current SmartNIC devices.

### 2021-06-09 paper/paper.tex
12 pager, 9pt

@@
 
-It is important to make this feasible, as offloading learning to commodity machines adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
+It is important to make this feasible as offloading learning to commodity machines adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
 For context, \emph{Deep Neural Network} (DNN) training relies on the backpropagation algorithm to compute gradients, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many mini-batches of data for stable training.
@@
 	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
-	\item A demonstration of how \approachshort{} would integrate with state-of-the-art PDP applications to perform fully in-NIC, fast, automated DDoS mitigation (\cref{sec:potential-integrations}).
+	\item A description of how \approachshort{} would integrate with state-of-the-art PDP applications to perform fully in-NIC, fast, automated DDoS mitigation (\cref{sec:potential-integrations}).
 \end{itemize}
@@
 %?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
-As ethernet moves beyond \SIlist{40;100}{\giga\bit\per\second}, packet processing deadlines grow tighter in tandem; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second}, giving a \SI{3.99}{\micro\second} deadline on Netronome SmartNICs (having \num{312} P4 pipelines).
+As ethernet moves beyond \SIlist{40;100}{\giga\bit\per\second}, packet processing deadlines grow tighter in tandem; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second}, giving a \SI{3.99}{\micro\second} deadline on Netronome SmartNICs (which have \num{312} P4 pipelines).
 In-pipeline execution thus increases the risk of drops or stalled packet transmissions.
@@
 
-Currently, low port density devices like the above are most likely to have specific general-purpose compute, as they are designed to suit high-performance offloading and middlebox development.
+Currently, low port-density devices like the above are most likely to have specific general-purpose compute, as they are designed to suit high-performance offloading and middlebox development.
 Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
@@
 RL algorithms are generally agnostic to the policy approximation used, and classical single-step methods are simple---most of the compute work is occupied by the policy itself.
-Consider the single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E} which requires only additions and multiplication to update and learn a policy online.
+The single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to update and learn a policy online.
 
@@
 A state vector produces a single `hit' in each tiling, all of which then correspond to a list of action preference lists.
-Inference (choosing an action) is then simply summing over all such lists to obtian a final action preference list, selecting the maximum.
+Inference (choosing an action) is then simply summing over all such lists to obtain a final action preference list, selecting the maximum.
 Updating at the next timestep uses this same list of lists, adjusting the value of the last selected action using a value $\delta_t$ computed through Sarsa.
@@
 However, this requires tile coding to be repeated $a$ times per tile, which is likely extremely wasteful outside of heavy parallelism.}
-As an example, a real-world policy size which we examine later (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets containing \num{8} tilings creates \num{129} work items.
+As an example, a real-world policy size which we examine later (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets of \num{8} tilings creates \num{129} work items.
 Computing a tile hit requires one division per dimension, which is an expensive ALU operation to require.
@@
 Much of this work goes further still towards integer~\parencite{tensorrt-8bit} or binarised~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16} representations, sacrificing dynamic range for simpler arithmetic operations.
-The works mentioned above use similar techniques to make inference tractable on network hardware, i.e., to bring neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
+The works mentioned above use similar techniques to make inference tractable on network hardware, i.e., to bring pre-trained neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
 
@@
 At a high level, \approachshort{} is designed to use the auxiliary compute exposed by general SmartNIC devices to offer low-latency online learning, scaling according to available on-chip resources at build time.
-As the allocation of cores/chip area is set ahead of time by a framework or system administrator, \approachshort{} (\Coopfw) agents enumerate themselves at runtime initialisation.
+As the allocation of cores/chip area is set ahead of time by a framework or system administrator, \approachshort{}(-\Coopfw) agents enumerate themselves at runtime, during initialisation.
 
@@
 \Cref{fig:netro-arch,fig:single-and-parallel} outline our design and implementation on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
-We explain relevant NFP architectural details in \cref{sec:netronome-platform-fundamentals}.
+We explain relevant NFP architectural details later in \cref{sec:netronome-platform-fundamentals}.
 \approachshort{} communicates with the packet pipeline of a P4 dataplane via \texttt{extern} plugins using \inring{} (state, configuration) and \outring{} (action) messages (\cref{fig:netro-arch}).
@@
 	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel,alg:parsa})] Threads cooperate on processing state vectors, minimising latency.
-\emph{Minions} have a fixed list of work items, while a \emph{controller} thread sends compute/update commands before awaiting worker completion.
+\emph{Minion} threads have a fixed list of work items, while a \emph{controller} thread sends compute/update commands before awaiting worker completion.
 Work items are disjoint, requiring no policy locks.
@@
 \end{description}
-Each offers a different point of optimisation; if updates are disabled, then the \indfw{} model can maximise throughput, while the \coopfw{} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
+Each offers a different point of optimisation; if updates are disabled, then the \indfw{} model can maximise throughput, while the \coopfw{} model is designed to minimise decision latency and needs no locks to update the policy (increasing \emph{online learning} throughput).
 These correspond to only executing a trained policy and actively (re-)training a policy, respectively.
@@
 The key insight of this mechanism is that on-chip reward/state signals enjoy first-class support in the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-NIC/chip asynchronous processes or the controller.
-As such, \approachshort{} can receive from P4 \texttt{extern}s or another dedicated off-path flow state measurement application.
+As such, \approachshort{} can receive input from P4 \texttt{extern}s or other, dedicated off-path flow state measurement applications.
 
@@
 Even with parallel problems such as \emph{ParSa}, optimising for latency requires meticulous care in how work is passed out and aggregated.
-This is truer still when moving from the moderately-grained control of classical RL ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
+This is truer still when moving from the moderately fine-grained control of classical RL ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
 Ordinarily, the marshalling of requests, responses, and shared data access can incur significant overheads.
@@
 
-Our implementation exploits the locality of cores in the NFP, which can be used on other platforms like NetFPGA.
+Our implementation exploits the locality of cores in the NFP, which can be factored into designs on other platforms like NetFPGA.
 Policy compute/update/configure tasks are passed between cores using direct \emph{next neighbour} registers, signalling all child threads in response.
@@
 
-	\Let \emph{values}: [AtomicI32; cfg.n\_actions] = \{0\}\;
+	\Let \emph{values}: [AtomicI32; \emph{cfg.n\_actions}] = \{0\}\;
 	\Let \emph{acks}: AtomicI32 = 0\;
@@
 			\ForAll{state\_pkt \In \inring{}}{
-				Ctl(state\_pkt)\;
+				Ctl(\emph{state\_pkt})\;
 			}
@@
 		\Else{
-			\While{true}{Minion(schedule[id - 1], recv())\;}
+			\While{true}{Minion(\emph{schedule}[$\mathit{id}- 1$], recv())\;}
 		}
@@
 		\If{cfg.online}{
-			\Let \emph{((l\_state, l\_act, l\_val), found\_s)} = \emph{cfg.lookup\_state\_from\_key(state)}\;
-			\Let \emph{(reward, found\_r)} = \emph{cfg.lookup\_reward\_from\_key(state)}\;
+			\Let \emph{((l\_state, l\_act, l\_val), found\_s)} = \emph{cfg}.lookup\_state\_from\_key(\emph{state})\;
+			\Let \emph{(reward, found\_r)} = \emph{cfg}.lookup\_reward\_from\_key(\emph{state})\;
 			\If{found\_s \&\& found\_r}{
@@
 				$\delta_t$ = $Q_\mathit{mul}$(\emph{cfg}.$\alpha$, $\delta_t$)\;
-				\emph{acks} = 0, scatter(Par::Upd($\delta_t$, \emph{action}, \emph{state}))\;
+				\emph{acks} = 0, scatter(Par::Upd($\delta_t$, \emph{l\_act}, \emph{l\_state}))\;
 				\Await acks == \emph{cfg.n\_minions}\;
@@
 				\For{i \In [0..cfg.n\_actions)}{
-					\emph{values}[\emph{i}].atomic\_add(\emph{policy}[\emph{hit}][\emph{a}])\;
+					\emph{values}[\emph{i}].atomic\_add(\emph{policy}[\emph{hit}][\emph{i}])\;
 				}
@@
 		}
-		acks.atomic\_add(1)\;
+		\emph{acks}.atomic\_add(1)\;
 	}
@@
 Even through \texttt{extern}s, directly adding new rules is neither feasible nor safe.
-We instead suggest on NFP chips that \texttt{extern}s or datapath stages which apply actions to packets should maintain a small store of state-action pairs, and periodically send these back to the controller for batch installation.
+We instead suggest on NFP chips that \texttt{extern}s or datapath stages which apply RL actions to packets should maintain a small store of state-action pairs, and periodically send these back to the controller for batch installation.
 This ensures that the majority of installed rules exhibit accelerated performance, while preventing installation delay on the newest decisions.
@@
 Even in wait-free algorithms, this requires a minimum number of workers to improve upon the latency bounds of a serial approach.
-We investigate the exact worker count requirements imposed to break even or improve upon single-threaded execution in the sequel.
+We investigate the exact worker count requirements imposed to break even or improve upon single-threaded execution in \cref{sec:results}.
 
@@
 \subsection{Netronome Platform Fundamentals}\label{sec:netronome-platform-fundamentals}
-As we implement this work on Netronome NFP SmartNICs, it is necessary to explain its basics.
+As we implement this work on Netronome NFP SmartNICs, it is necessary to explain their basics.
 These \emph{system-on-a-chip} (SoC) devices achieve scalable packet processing through sheer parallelism.
@@
 \begin{description}
-	\item[\emph{MidServer}] Intel Xeon Bronze 3204 CPUs (\qtyproduct[product-units=single]{6 x 1.9}{\giga\hertz}),
-	\item[\emph{HighServer}] Intel Xeon Silver 4208 CPU (\qtyproduct[product-units=single]{8 x 2.1}{\giga\hertz}),
+	\item[\emph{MidServer}] Intel Xeon Bronze 3204 (\qtyproduct[product-units=single]{6 x 1.9}{\giga\hertz}),
+	\item[\emph{HighServer}] Intel Xeon Silver 4208 (\qtyproduct[product-units=single]{8 x 2.1}{\giga\hertz}),
 	\item[\emph{Collector}] Intel Core i7-6700K (\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}).
@@
 In \Coopfw{}, this creates \num{129} work items across \num{31} workers.
-Although the model application's more successful \emph{SPF} agent design uses 3 actions, we choose a larger action set to investigate the performance of more complex agents.
+Although their more successful \emph{Guarded} agent design uses 3 actions, we choose a larger action set to investigate the performance of more complex agents.
 
@@
 By limiting the available workers in software, we show how \Coopfw{}'s policy update time (thus  online throughput---\cref{fig:vary-core}) and state-action latency (\cref{fig:vary-core-latency}) scale with available cores.
-While \Coopfw{} always outperforms the host-based floating point implementations, we observe that there are two distinct crossover points to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency.
+While \Coopfw{} always outperforms the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency.
 Some artefacts of our environment and design choices are visible, such as the addition of new physical cores being more significant than contexts, and the presence of some schedule bottlenecks.
@@
 \fakepara{Co-existence with the dataplane}
-Our setup saturated the \SI{40}{\giga\bit\per\second} link for packet sizes $\ge$\SI{256}{\byte}.
+Our setup met \SI{40}{\giga\bit\per\second} for packet sizes $\ge$\SI{256}{\byte}.
 %For frame sizes of \SIlist{64;128;256}{\byte} input traffic rates were \SIlist{17.4;32.9;37.0}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2;18.1}{\mega\packet\per\second}).
@@
 
-We note that the TRS scheduler is designed to handle cases large numbers of flows created by an attacker, combining state vectors over time while the asynchronous agent is itself busy.
+We note that the TRS scheduler is designed to handle large numbers of attack flows, combining seen state vectors over time while the asynchronous agent is itself busy.
 A number of (attack) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
 As shown in \cref{sec:evaluation}, \approachshort{} far exceeds the throughput of host offloading.
-The control plane can then use wildcards or specific matches to narrow down or expand the set of flows to be controlled, though explicit (TRS) scheduling is still key in such adversarial environments.
+The control plane can then use wildcards or specific matches to narrow down or expand the set of flows to be controlled dynamically, though explicit (TRS) scheduling is still key in such adversarial environments.
 
@@
 \textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantisation on the NFP, using \emph{in-path} compute rather than our asynchronous model.
-Inference latency on small networks (3 layers, \num{30} neurons) can be as high as \SI{500}{\micro\second} on line rate traffic, stressing the value of path-adjacent compute.
+Inference latency on small networks (3 layers, \num{30} neurons) can be as high as \SI{500}{\micro\second} on line rate traffic, emphasising the value of path-adjacent compute.
 

### 2021-06-10 paper/paper-10pt.tex
Raw fear, 10pt

--- 2021-06-09 paper/paper.tex
@@
 sigconf,natbib=false
-%,anonymous=true
-,9pt
+,anonymous=true
+,10pt
 ]{acmart}
 
+\usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists]{savetrees}
+
 \usepackage[%
-false
+%false
 ]{anonymous-acm}
@@
 % Copyright
-%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
+\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
 %\setcopyright{none}
@@
 %\settopmatter{printacmref=false, printccs=false, printfolios=true}
-\settopmatter{printfolios=true}
+%\settopmatter{printfolios=true}
+\settopmatter{printacmref=false, printccs=true, printfolios=true}
 
@@
 %% {} with no args suppresses printing of the price
-\acmPrice{15.00}
+%\acmPrice{15.00}
+\acmPrice{}
 
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/arch-with-p4}
+	\includegraphics[keepaspectratio, width=0.65\linewidth]{figures/arch-with-p4}
 	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
@@
 It is important to make this feasible as offloading learning to commodity machines adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
-For context, \emph{Deep Neural Network} (DNN) training relies on the backpropagation algorithm to compute gradients, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many mini-batches of data for stable training.
+For context, \emph{Deep Neural Network} (DNN) training relies on backpropagation to compute gradients, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many mini-batches of data for stable training.
 These lead to high compute and storage costs, inducing capital expenditure for dedicated accelerators.
@@
 
-To enable \emph{online in-NIC learning}, we return to \emph{classical} RL methods and models.
-In particular, we focus on tile-coding with one-step temporal-difference learning algorithms such as Sarsa.
+To enable \emph{online in-NIC learning}, we return to \emph{classical} RL methods.
+In particular, we focus on tile-coding with one-step temporal-difference algorithms such as Sarsa.
 %These choices have important benefits for in-NIC execution.
@@
 
-\section{Preliminaries: RL for in-network computation}\label{sec:motivation}
-We discuss recent trends in programmable switch hardware.
-By combining this with insights from the ML/RL communities (past and present), we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL methods and quantisation make this computationally feasible.
-
-Crucially, we train our focus on devices with a low port density, such as SmartNICs and the NetFPGA.
-The designs of these devices make it reasonable to move policy processing \emph{outside of the packet pipeline}; SmartNICs expose many additional programmable cores, and NetFPGAs allow for the synthesis of independent functional units.
-The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
+\section{Preliminaries: towards in-NIC RL}\label{sec:motivation}
+%\section{Preliminaries: RL for in-network computation}\label{sec:motivation}
+By combining recent trends in PDP hardware with insights from the ML/RL communities (past and present), we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL methods and quantisation make this computationally feasible.
+Crucially, we focus on low port-density devices such as SmartNICs and the NetFPGA.
+The designs of these devices make it reasonable to move processing \emph{outside of the packet pipeline}; SmartNICs expose additional programmable cores, and NetFPGAs allow for the synthesis of independent functional units.
+%The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
+The main benefit is that core control logic can be moved onto the NIC \emph{without impacting packet processing rates}.
 %?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
-As ethernet moves beyond \SIlist{40;100}{\giga\bit\per\second}, packet processing deadlines grow tighter in tandem; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second}, giving a \SI{3.99}{\micro\second} deadline on Netronome SmartNICs (which have \num{312} P4 pipelines).
-In-pipeline execution thus increases the risk of drops or stalled packet transmissions.
+As Ethernet speed increases, packet deadlines grow tighter; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second} at \SI{40}{\giga\bit\per\second}, giving a \SI{3.99}{\micro\second} deadline on Netronome (which has \num{312} P4 pipelines).
+In-pipeline execution thus increases the risk of packet drops or stalls.
 
@@
 
-While the P4 language~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to great interest in network programmability, it requires similar behaviour between device classes.
-This is encoded by the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}---a sensible lower bound on the device capabilities required to implement a P4 dataplane.
-We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
-Many compatible devices have legacies long predating these developments. 
+While P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to great interest in network programmability, it requires similar behaviour between device classes.
+This is encoded by the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}---a lower bound on the device capabilities required to implement a P4 dataplane.
+%We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
+However, many compatible devices have legacies long predating these developments. 
 For instance, many-core SOC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
 %This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
-
-Currently, low port-density devices like the above are most likely to have specific general-purpose compute, as they are designed to suit high-performance offloading and middlebox development.
-Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
-Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of deep learning accelerators (which, in turn, lack packet switching functionality).
-Moreover, the limited memory/block RAM and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
+Currently, low port-density devices like these are most likely to have specific general-purpose compute, as they are designed to for high-performance offloading and middlebox development.
+%Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
+Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of deep learning accelerators (which lack packet switching functionality).
+Moreover, the limited memory and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
 
@@
 %Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
-Consider \cref{fig:state-slip}: the traditional RL formulation assumes that an agent receives new state at fixed time intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
-This adds noise to the state-action mapping being learned, which harms learning rate and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
+Consider \cref{fig:state-slip}: the RL formulation assumes that an agent receives new state at fixed intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
+This adds noise to the state-action mapping, which harms learning and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
 %Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
 Reducing these costs requires detailed agent-environment co-design.
-This principle has influenced the design of real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, where asynchrony is necessary for high-speed applications.
+This principle has influenced real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, where asynchrony is necessary for high-speed applications.
 This often requires that state measurements are fused~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
-`Stopping the world' causes significant performance loss, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control, or any time-sensitive control problems.
+`Stopping the world' causes significant performance loss, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control.
 
@@
 		\centering
-		\resizebox{0.9\linewidth}{!}{
+		\resizebox{0.8\linewidth}{!}{
 			\centering
@@
 		\centering
-		\resizebox{0.7\linewidth}{!}{
+		\resizebox{0.6\linewidth}{!}{
 			\centering
@@
 		\end{tikzpicture}}
-		\caption{Reality: measurement and action costs lead to \emph{state drift}---over a delay $t_1+t_2+t_3$, inaction transforms state $S$ into $S'$.}
+		\caption{Reality: policy costs lead to \emph{state drift}---over a delay $t_1+t_2+t_3$, state $S$ becomes $S'$.}
 	\end{subfigure}
@@
 We return to what classical RL methods can offer us; in particular, the simple linear function approximation given by \emph{tile coding} \cite[pp.\ \numrange{217}{221}]{RL2E}.
-The idea is simple: a policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
-This covers any input-output mapping, such as converting packet header data into a preference list for queue priorities.
+A policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
+This allows any input-output mapping, such as converting packet header data into a preference list for queue priorities.
 A state vector produces a single `hit' in each tiling, all of which then correspond to a list of action preference lists.
-Inference (choosing an action) is then simply summing over all such lists to obtain a final action preference list, selecting the maximum.
-Updating at the next timestep uses this same list of lists, adjusting the value of the last selected action using a value $\delta_t$ computed through Sarsa.
-Crucially, this internal representation ensures that there are no data dependencies between any tile calculations for an input.
-This has important benefits for parallelising the key parts of an RL algorithm; both action selection and policy updates are map-reduce problems.
+Inference (choosing an action) is simply summing over all such lists to obtain a final action preference list, selecting the maximum.
+Updating uses this same list of lists after the next decision is made, adjusting the value of the last selected action using a value $\delta_t$ computed through Sarsa.
+Crucially, this internal representation has no data dependencies between any tile calculations for an input.
+As a result, both action selection and policy updates are map-reduce problems.
 To select an action, we can produce a task for each tiling---retrieving the list of action values contributed by the tile activated by the input state.
@@
 
-For instance, a policy with $m$ sets of tiles (each having its own set of input dimensions), each containing $n$ overlapping tilings then creates $m \times n$ separate tasks.\footnote{Retrieving each individual action's value can be considered a work item (giving $m \times n \times a$ tasks for $a$ actions).
-However, this requires tile coding to be repeated $a$ times per tile, which is likely extremely wasteful outside of heavy parallelism.}
+For instance, a policy with $m$ sets of tiles (each having its own set of input dimensions), each containing $n$ overlapping tilings then creates $m \times n$ separate tasks.\footnote{Work items could be retrieving each individual action's value (giving $m \times n \times a$ tasks for $a$ actions).
+This requires tile coding to be repeated $a$ times per tile, which is likely extremely wasteful outside of heavy parallelism.}
 As an example, a real-world policy size which we examine later (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets of \num{8} tilings creates \num{129} work items.
-Computing a tile hit requires one division per dimension, which is an expensive ALU operation to require.
+Computing a tile hit uses one division per dimension, which is an expensive ALU operation to require.
 However, if tile widths are fixed at powers-of-two, then these may be replaced with right bitshifts.
@@
 
-\subsection{Numerical representation for embedded ML}
-A key feature universally lacking from PDP hardware is floating-point arithmetic support.
-Luckily, the task of bringing ML models to resource-constrained environments without these capabilities is well-studied.
+\subsection{Representations for embedded ML}
+As PDP hardware universally lacks floating-point arithmetic support, the task of bringing ML models to resource-constrained environments without these capabilities is well-studied.
 Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
-Lower bit depths reduce memory footprints, and improve throughput in designs such as \emph{bfloat16}~\parencite{bfloat16-blog} in Google TPUs~\parencite{DBLP:journals/sigops/XieDMKVZT18}, \emph{hfp8}~\parencite{DBLP:conf/nips/SunCCWVSCZG19}, and other floating point formats~\parencite{DBLP:journals/corr/abs-2007-01530}.
-In many cases, accuracy losses for doing so are negligible.
-Much of this work goes further still towards integer~\parencite{tensorrt-8bit} or binarised~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16} representations, sacrificing dynamic range for simpler arithmetic operations.
-The works mentioned above use similar techniques to make inference tractable on network hardware, i.e., to bring pre-trained neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
+Lower bit depths reduce memory footprints, and improve throughput in e.g., \emph{bfloat16}~\parencite{bfloat16-blog,DBLP:journals/sigops/XieDMKVZT18}, \emph{hfp8}~\parencite{DBLP:conf/nips/SunCCWVSCZG19}, and other floating point formats~\parencite{DBLP:journals/corr/abs-2007-01530}.
+In many cases, accuracy losses are negligible.
+Such work goes further still towards integers~\parencite{tensorrt-8bit} or binarisation~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16}, sacrificing dynamic range for cheaper arithmetic.
+Similar techniques are used to make inference tractable on network hardware, i.e., to bring pre-trained neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
 
@@
 \section{Design and Implementation}\label{sec:design}
-Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{} (\approach)}.
-At a high level, \approachshort{} is designed to use the auxiliary compute exposed by general SmartNIC devices to offer low-latency online learning, scaling according to available on-chip resources at build time.
+%Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{} (\approach)}.
+We present our design for an in-NIC, task-independent, online RL system---\emph{\approachshort{} (\approach)}.
+At a high level, \approachshort{} uses auxiliary compute exposed by SmartNIC devices to offer low-latency online learning, scaling with on-chip resources at build time.
 As the allocation of cores/chip area is set ahead of time by a framework or system administrator, \approachshort{}(-\Coopfw) agents enumerate themselves at runtime, during initialisation.
@@
 
-\Cref{fig:netro-arch,fig:single-and-parallel} outline our design and implementation on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
+\Cref{fig:netro-arch,fig:single-and-parallel} outline our design on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
 We explain relevant NFP architectural details later in \cref{sec:netronome-platform-fundamentals}.
@@
 We open-source our firmware and control programs for the benefit of the community.\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission.}}
-Moving beyond the Netronome platform, we describe how our architecture may be adapted and improved on by bespoke hardware or FPGA-based deployment.
+%Moving beyond the Netronome platform, we describe how our architecture may be adapted and improved on by bespoke hardware or FPGA-based deployment.
 
 \subsection{System Model}
-\approachshort{} is a general, task-independent framework for in-network, online training and execution of \emph{any reinforcement learning agent design} using classical methods.
+\approachshort{} is a general, task-independent framework for in-network, online training and execution of \emph{any RL agent design} using classical methods.
 \approachshort{} is agnostic to the meaning of state vectors it receives as inputs and the actions it produces, which are employed by other functional units or the dataplane.
@@
 \approachshort{} runs on one or more cores of a SmartNIC to convert fixed-point state measurements from the environment into a stream of actions using a stored policy.
-As an example, this might be to map flow state and performance measurements into a queueing priority for future packets from that flow, or to compute and apply a rate limit to preserve quality of service.
-These dedicated cores are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
+As an example, this might be to map flow state and performance measurements into a queue priority for future packets, or to compute and apply a rate limit to preserve quality of service.
+These dedicated cores are responsible for processing requests, computing actions, and updating the underlying policy in real time.
 Combined with reward measurements, this policy can then be updated or trained from scratch entirely on the NIC, acting as a fully online RL agent.
@@
 Updates are computed using \emph{single-step semi-gradient Sarsa}~\cite[pp. \numrange{217}{221}]{RL2E}, though modification to support other single-step methods would be trivial.
-The new state- or tiles-action pair is then written into storage.
+The new state-action pair is then written into storage.
 \approachshort{} can be configured to automatically select values of the input state vector as keys for state and reward storage.
-
-Two firmware models govern how the compute-heavy parts of these tasks (action selection, policy updates) are carried out:
+Two firmware models govern how these tasks (action selection, policy updates) are carried out:
 \begin{description}
-	\item[\Indfw{} (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each performs its work sequentially.
+	\item[\Indfw{} (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each works sequentially.
 Computing an action list requires a \emph{read lock} on the policy.
-If an update occurs, the core requests a \emph{write lock} before updating, greatly limiting online throughput.
+If an update occurs, the core requests a \emph{write lock} before updating, limiting online throughput.
 \emph{Tile lists} are stored for update computation.
@@
 %In general it is also possible for the \emph{Ctl} task to act as an additional \emph{Minion} in its parallel sections; we were limited here by code store requirements on the NFP.
-Latency and throughput, as in many networked systems, have different effects upon RL agents according to their design and target problem.
+Latency and throughput have different effects on RL agents according to their design and target problem.
 Higher RL throughput is a necessity for per-flow or per-packet applications, which can require high decision-per-second rates even after combining state measurements received from the environment, such as flow control in DDoS prevention.
-Equally, lower latency affords an agent finer-grained control and learning of a problem, being able to react sooner to new information (e.g., device state in a routing optimisation problem, or queue depth when trying to enforce packet pacing).
+Equally, lower latency affords an agent finer-grained control and learning of a problem, being able to react sooner to new information (e.g., device state in a routing optimisation problem).
 
@@
 \subsection{Agent-Environment Communication}\label{sec:agent-environment-communication}
-\approachshort{} uses \emph{multiple-producer/multiple-consumer} (MPMC) messaging channels to communicate with other elements; be they P4 programs on the packet path, or other on-chip analysis and control modules.
-Through these channels a system \emph{pushes} state vectors, reward measures, and setup packets as inputs, and \emph{pulls} a stream of state-action pairs as outputs.
-This allows decisions to be made asynchronously---preventing packet stalling---and allowing many RL agents to be used if desired.
+\approachshort{} uses \emph{multiple-producer/multiple-consumer} (MPMC) channels to communicate with other elements; be they P4 programs on the packet path, or other on-chip analysis and control modules.
+Through these, a system \emph{pushes} state vectors, reward measures, and setup packets as inputs, and \emph{pulls} a stream of state-action pairs as outputs.
+This allows decisions to be made asynchronously, preventing packet stalling.
 The key insight of this mechanism is that on-chip reward/state signals enjoy first-class support in the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-NIC/chip asynchronous processes or the controller.
@@
 \subsection{Intra-Agent Communication}\label{sec:intra-agent-communication}
-Even with parallel problems such as \emph{ParSa}, optimising for latency requires meticulous care in how work is passed out and aggregated.
+Optimising for latency requires meticulous care in how work is passed out and aggregated.
 This is truer still when moving from the moderately fine-grained control of classical RL ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
-Ordinarily, the marshalling of requests, responses, and shared data access can incur significant overheads.
-On-chip execution and the nature of action preference computation allow us to use lockless atomic aggregation, removing the overheads of explicit messaging/packetisation.
-Moreover, adjacent functional units/cores often have special-purpose shared registers or share a small fast cache to accelerate communication.
+Marshalling of requests and shared data access incur significant overheads, but on-chip execution and the ParSa algorithm allow us to sidestep these via lockless atomic aggregation.
+Moreover, adjacent cores often have special-purpose shared registers or share a small fast cache to accelerate communication.
 
@@
 
-Our implementation exploits the locality of cores in the NFP, which can be factored into designs on other platforms like NetFPGA.
+Our implementation exploits the locality of cores in the NFP, which can be factored into designs on other platforms like NetFPGA as additional off-path functional units, as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
 Policy compute/update/configure tasks are passed between cores using direct \emph{next neighbour} registers, signalling all child threads in response.
@@
 This is essential for aggregation compared to the use of bounded message buffers, which caused significant head-of-line blocking.
+%For instance, additional off-path functional units can serve as workers, like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+%Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the P4$\rightarrow$NetFPGA~\parencite{DBLP:conf/fpga/IbanezBMZ19} actions that interface with it.
 
@@
 \begin{algorithm}
+	\small
 	\caption{ParSa---\emph{Par}allel \emph{Sa}rsa\label{alg:parsa}}
@@
 \subsection{Reconfigurability}
-\approachshort{} allows policy design and algorithmic learning parameters to be changed at runtime using at most two control packets.
+\approachshort{} allows policy design and learning parameters to be changed at runtime using at most two control packets.
 For instance, design changes are useful at the end of learning (moving from online to offline), or when trying to train a new policy for another problem from the same vantage point.
 Parameter changes are useful when an online agent must become more (or less) adaptive to new data (i.e., after detecting a changepoint in traffic).
-This extends to policy data, which may be imported from a pre-trained model via such packets and exported via PCIe to the host machine.
+This extends to policy data, which may be imported from a pre-trained model and exported via PCIe to the host machine.
 Some aspects must be chosen at compile time; bit depth, \Coopfw/\Indfw, and maximum policy/tiling/state sizes---these govern core operation or allocated memory.
@@
 
-In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved (pool 2~\parencite{rfc2474}) DSCP value as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved DSCP value~\parencite{rfc2474} as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
 While this mainly automates parser generation, it also allows for configuration to be received from only trusted hosts (over the dataplane if needed) via P4 rules.
@@
 Due to the lack of dynamic memory allocation on PDP hardware, and to simplify value lookups, policies cannot be stored sparsely.
-As a consequence, tiling space requirements scale exponentially with dimension count, and so higher-dimension tilings must be placed in larger (and slower) memory regions.
+Tiling space requirements then scale exponentially with dimension count, so higher-dimension tilings must be placed in larger, slower memory regions.
 As a result, in many architectures policy parameters are split across such regions, giving different access and compute costs to different tasks.
@@
 We use a simple first-fit work placement algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core (as the NFP uses hardware threads).
-Each work item is a separate \emph{tiling} over a list of dimensions, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
+Each work item is a separate \emph{tiling} over a list of dimensions.%, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
 The cost of any work item (from its dimension count and memory location) was empirically measured offline, and we weigh the total cost per core based on the number of minion threads available.
 This weighting specifically accounts for the controller thread on the first core.
-Work allocation and cached per-item data are recomputed when policy configuration is installed or changed.
+Work allocations are recomputed when policy configuration is installed or changed.
 Naturally, for $n$ tilings and $m$ threads this procedure is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per tiling, storing $m/8$ and $\le8$ costs respectively.
@@
 \subsection{Limitations}\label{sec:limitations}
-Unfortunately, direct rule installation into P4 tables from the SmartNIC itself is not, in general, possible.
-To achieve line-rate match-action table lookup, platforms like NFP use accelerated datastructures (e.g., DCFL~\parencite{DBLP:conf/infocom/TaylorT05}) which must be computed by the controller over the \emph{entire rule set}.
+Direct rule installation into P4 tables from the SmartNIC is not generally possible.
+To achieve line-rate performance, platforms like NFP use accelerated datastructures (e.g., DCFL~\parencite{DBLP:conf/infocom/TaylorT05}) which must be computed over the \emph{entire rule set}.
 Even through \texttt{extern}s, directly adding new rules is neither feasible nor safe.
 We instead suggest on NFP chips that \texttt{extern}s or datapath stages which apply RL actions to packets should maintain a small store of state-action pairs, and periodically send these back to the controller for batch installation.
-This ensures that the majority of installed rules exhibit accelerated performance, while preventing installation delay on the newest decisions.
-Platforms such as Intel Tofino greatly simplify this, with Tofino Native Architecture intrinsics such as \emph{Action Profiles/Selects} allowing a P4 action to be chosen based on a register value (i.e., an RL action).
+%This ensures that the majority of installed rules exhibit accelerated performance, while preventing installation delay on the newest decisions.
+%Platforms such as Intel Tofino greatly simplify this, with Tofino Native Architecture intrinsics such as \emph{Action Profiles/Selects} allowing a P4 action to be chosen based on a register value (i.e., an RL action).
 
 %As is expected of parallel algorithms, efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
-Efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
-Even in wait-free algorithms, this requires a minimum number of workers to improve upon the latency bounds of a serial approach.
-We investigate the exact worker count requirements imposed to break even or improve upon single-threaded execution in \cref{sec:results}.
-
-\subsection{Targeting other device classes}
-While this design caters to Netronome devices, many aspects have clear analogues in other SmartNICs, such as the NetFPGA SUME.
-For instance, additional off-path functional units can serve as workers, like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the P4$\rightarrow$NetFPGA~\parencite{DBLP:conf/fpga/IbanezBMZ19} actions that interface with it.
-%Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
-As we show later, using a bit depth below the native word size reduces overall efficiency.
-%This reduces overall efficiency even though considerably fewer reads are made.
-We expect an FPGA design would allow this to match the desired bit depth and enable the SIMD-like optimisations we discuss without creative and costly bit-stuffing.
-
-Bringing \approachshort{} to PDP switches such as the Tofino is difficult as they closely match the P4-PSA, leaving no spare general-purpose compute units.
-Taking inspiration from real-time programming, a potential solution is to divide RL processing across several packets (i.e., computing a portion of the preference list each time) until further work would delay outbound transmission.
-This would introduce new issues in concurrent access, work splitting, and altered timescales for learning: we leave their treatment to future work.
+%Efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
+Parallelisation adds per-task overheads, which require a minimum number of workers to improve on a serial approach.
+We investigate the exact worker count requirements imposed to overcome single-threaded execution in \cref{sec:results}.
+
+%\subsection{Targeting other device classes}
+%While this design caters to Netronome devices, many aspects have clear analogues in other SmartNICs, such as the NetFPGA SUME.
+%For instance, additional off-path functional units can serve as workers, like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
+%Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the P4$\rightarrow$NetFPGA~\parencite{DBLP:conf/fpga/IbanezBMZ19} actions that interface with it.
+%%Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
+%As we show later, using a bit depth below the native word size reduces overall efficiency.
+%%This reduces overall efficiency even though considerably fewer reads are made.
+%We expect an FPGA design would allow this to match the desired bit depth and enable the SIMD-like optimisations we discuss without creative and costly bit-stuffing.
+%
+%Bringing \approachshort{} to PDP switches such as the Tofino is difficult as they closely match the P4-PSA, leaving no spare general-purpose compute units.
+%Taking inspiration from real-time programming, a potential solution is to divide RL processing across several packets (i.e., computing a portion of the preference list each time) until further work would delay outbound transmission.
+%This would introduce new issues in concurrent access, work splitting, and altered timescales for learning: we leave their treatment to future work.
 
@@
 \subsection{Netronome Platform Fundamentals}\label{sec:netronome-platform-fundamentals}
-As we implement this work on Netronome NFP SmartNICs, it is necessary to explain their basics.
-These \emph{system-on-a-chip} (SoC) devices achieve scalable packet processing through sheer parallelism.
-Most of the chip is composed of \emph{microengines} (MEs), grouped into \emph{islands} of 4 or 12 MEs.
-All 12-ME islands are used by a default P4 pipeline.
+Netronome NFPs achieve scalable packet processing through parallelism.
+The chip is composed of \emph{microengines} (MEs), grouped into \emph{islands} of 4 or 12 MEs.
+%All 12-ME islands are used by a default P4 pipeline.
 Each ME has \numrange{4}{8} \emph{contexts} (hardware threads) which share a code store.
@@
 %MEs run a proprietary instruction set, compiled to via a \emph{(Micro-)C} compiler.
-Beyond registers, the platform implements an explicit memory hierarchy scaling in size, location, and access cost:
+Beyond registers, the platform implements an explicit memory hierarchy scaling in size and access cost:
 $\text{LMEM (ME)} < \text{CLS (Island)} < \text{CTM} < \text{IMEM (Chip)} < \text{EMEM}$.
@@
 \subsection{Experimental Setup}\label{sec:experimental-setup}
-Testing machines had the following hardware, all with \SI{32}{\gibi\byte} RAM:
+Testing machines were as follows, with \SI{32}{\gibi\byte} RAM:
 \begin{description}
@@
 Control programs were built using rustc 1.52.1.
-
-We run \approachshort{} on a \num{4}-ME island of the NFP-6480, totalling \num{32} contexts.
+We run \approachshort{} on a \num{4}-ME island of the NFP-6480 (\num{32} contexts) using \SIlist{32;16;8}{\bit} quantised arithmetic.
 This is the largest cluster of cores which is not in use by a P4 pipeline.
-Where feasible, we test \SIlist{32;16;8}{\bit} quantised arithmetic.
 All \approachshort{} timing measurements were repeated over \num{10000} state packets (preceded by \num{1000} warmup packets), retrieving item processing times over PCIe via the controller machine, from which throughput was derived.
@@
 Policy sizes are set to those of a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dim state vectors, a bias tile and 16 full tiling sets (\numproduct{7x1}-dim, \numproduct{8x2}-dim, \numproduct{1x4}-dim), 8 tilings per set, 6 tiles per dimension, and 10 actions.
-For context, such input state would contain various aspects of per-flow state (e.g., IATs, rates) which are combined with other state such as the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim).
+For context, such input would contain per-flow state (e.g., IATs, rates) which is combined with the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim).
 In \Coopfw{}, this creates \num{129} work items across \num{31} workers.
-Although their more successful \emph{Guarded} agent design uses 3 actions, we choose a larger action set to investigate the performance of more complex agents.
+We choose a larger action than this work to investigate the performance of more complex agents.
 
@@
 \fakepara{Raw inference and learning performance}
-We compare how long it takes for \approachshort{} to compute actions and update its policy per state vector received, and report on the observed throughput of our designs against floating point (numpy-based) implementations of Sarsa on a commodity host machine.
+We compare how long it takes for \approachshort{} to compute actions and policy updates, and report on the throughput of our designs against floating point (numpy-based) implementations of Sarsa on commodity hosts.
 This allows us to demonstrate the performance differences between the \indfw{} and \coopfw{} configurations, particularly in how \indfw's (and hosts') required policy locks impact throughput.
 %We also report on the derived throughput measures accounting for the locking behaviour of each.
-We compare online learning performance (input states produce an output, and then update the policy) with offline (input states \emph{only} produce an output) in these cases.
+We compare online learning performance with offline in these cases.
 Online performance marks the number of decisions that can be made per second (and associated latency) when training a policy.
@@
 
-Building on this, we vary the amount of worker threads to show how \approachshort{} scales to fit available compute resources on a device.
+We then vary the amount of worker threads to show how \approachshort{} scales to fit available compute resources on a device.
 This is an important aspect for (automated) allocation of compute resources in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of-time work scheduling which we examine later.
@@
 We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ tilings given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
-We use maximum-size, \SI{32}{\bit} policies as described above.
+%We use maximum-size, \SI{32}{\bit} policies as described above.
 
 \fakepara{End-to-end RL latency}
-We compare the key RL decision-making latencies we discuss in \cref{fig:state-slip} across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to a \emph{virtual Network Function} (vNF).
+We compare the key RL latencies we discuss in \cref{fig:state-slip} across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to a \emph{virtual Network Function} (vNF).
 
 \fakepara{Co-existence with the dataplane}
-While varying the rate of full RL updates performed by \approachshort{}-\Coopfw{} (\SI{32}{\bit}) from \numrange{0}{16000} actions/s, we measure packet losses and sample latencies of cross traffic forwarded over a P4 pipeline hosted on our SmartNICs.
+While varying the rate of RL updates performed by \approachshort{}-\Coopfw{} (\SI{32}{\bit}) from \numrange{0}{16000} actions/s, we measure packet loss and sample latencies of cross traffic carried over a co-hosted P4 pipeline.
 This allows us to quantify whether on-chip (out-of-path) execution impacts ordinary dataplane behaviour through indirect means: e.g., EMEM cache evictions or hidden resource contention.
@@
 \fakepara{Raw inference and learning performance}
-\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy.\footnote{For brevity, we omit numpy-based integer results---against a floating point implementation, median action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}
-We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using 4 MEs of the NFP-6480.
-This corresponds to \SIlist{15;21}{$\times$} speedups over a \emph{Collector} host.
+\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy.\footnote{For brevity, we omit integer numpy results---against a float implementation, median action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}
+We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using 4 MEs of the NFP-6480 (\SIlist{15;21}{$\times$} speedups over a \emph{Collector} host).
 Importantly, \Indfw{} achieves lower median state-action latencies (\SI{2.79}{$\times$}) \emph{and} update times (\SI{2.63}{$\times$}) than a dedicated \emph{Collector} host \emph{while requiring only a single core or dedicated functional unit}.
+Crucially, in-NIC execution gives far tighter bounds on tail latency compared to host offloading.
+\num{99.99}\nthscript{th} percentile state-action latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, while host-based tail latencies are at least \SI{40.53}{\percent} greater even when the parallel worker count is at or below the physical core count.
+We show the cumulative distributions of these in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF/traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
+
+%We found that a key aspect of in-NIC execution is that it allows far tighter bounds on tail latency compared to host offloading.
+%Examining the state-action latencies in \cref{tab:lats}, we see that \num{99.99}\nthscript{th} percentile latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, respectively.
+%Similar results were observed for other bit depths.
+%By contrast, host-based tail latencies are at least \SI{40.53}{\percent} greater even when the parallel worker count is at or below the physical core count.
+%We show the cumulative distributions of these in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF/traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
 
 \newlength{\resultplotwidth}
-\setlength{\resultplotwidth}{0.88\linewidth}
+\setlength{\resultplotwidth}{0.8\linewidth}
 
@@
 We set the worker count on host machines which maximised their throughput.
-This equalled the amount of physical cores on each device---moving beyond this (even below the number of hyper-threads) would hamper tail latencies by an order of magnitude.
+This equalled the amount of physical cores on each device---moving beyond this would hamper tail latencies by an order of magnitude.
 To make the comparison fair in the context of many-core CPU environments, we include per-core throughput.
@@
 
-By limiting the available workers in software, we show how \Coopfw{}'s policy update time (thus  online throughput---\cref{fig:vary-core}) and state-action latency (\cref{fig:vary-core-latency}) scale with available cores.
-While \Coopfw{} always outperforms the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency.
+By limiting the available workers in software, we show how \Coopfw{}'s policy update time (thus  online throughput---\cref{fig:vary-core}) scales with available cores.
+While \Coopfw{} always outperforms the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency (plot omitted).
 Some artefacts of our environment and design choices are visible, such as the addition of new physical cores being more significant than contexts, and the presence of some schedule bottlenecks.
@@
 	\caption{
-		\Coopfw{}'s online learning performance improves with additional cores, on max size tasks (\num{129} work items).
+		\Coopfw{}'s online learning performance improves with additional cores, on max size tasks.
 This requires \num{8} workers to offer greater online throughput than single-threaded in-NIC RL.
@@
 
-\begin{figure}
-	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
-	}
-	\caption{\Coopfw{}'s action latencies similarly improve with more cores.
+%\begin{figure}
+%	\resizebox{\resultplotwidth}{!}{
+%		\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
+%	}
+%	\caption{\Coopfw{}'s action latencies similarly improve with more cores.
 This requires \num{3} workers (4 total contexts) to improve upon the state-action latency of single-threaded in-NIC RL.\label{fig:vary-core-latency}}
-\end{figure}
+%\end{figure}
 
@@
 
-We found that a key aspect of in-NIC execution is that it allows far tighter bounds on tail latency compared to host offloading.
-Examining the state-action latencies in \cref{tab:lats}, we see that \num{99.99}\nthscript{th} percentile latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, respectively.
-Similar results were observed for other bit depths.
-By contrast, host-based tail latencies are at least \SI{40.53}{\percent} greater even when the parallel worker count is at or below the physical core count.
-We show the cumulative distributions of these in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF/traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
-
 \begin{figure}[t]
@@
 	}
-	\caption{\Coopfw{} fully processes updates faster than \Indfw{}---thus has higher online performance---on almost all policy sizes.
-Lower bit depths are more effective on simpler policies.\label{fig:vary-work}}
+	\caption{\Coopfw{} processes updates faster than \Indfw{}---thus has higher online performance---on almost all policy sizes. Lower bit depths are effective on simple policies.\label{fig:vary-work}}
 \end{figure}
@@
 	}
-	\caption{State-action latency scales with additional work in a similar manner to overall processing time; though \SI{32}{\bit} firmwares become more effective sooner (at \num{3} work items).\label{fig:vary-work-latency}}
+	\caption{State-action latency scales with additional work similarly to \cref{fig:vary-work}; though \SI{32}{\bit} firmwares become more effective sooner (at \num{3} work items).\label{fig:vary-work-latency}}
 \end{figure}
@@
 
-A noteworthy trend is that \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
-This occurs even though our implementation is optimised to read and write policy data in batches (achieving fewer I/O operations).
-We see this because the native register width on the NFP is \SI{32}{\bit}, and so the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
+%A noteworthy trend is that \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
+\SIlist{8;16}{\bit} versions of \approachshort{} underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency})---even though our implementation is optimised to read and write policy data in batches.
+Because the native register width on the NFP is \SI{32}{\bit}, the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
 This matches \SI{32}{\bit} becoming dominant in complex workloads: higher dimension tilings require more arithmetic operations.
@@
 This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the online algorithm.
-Smaller bit depths still give a \SIrange{2}{4}{$\times$} saving in memory for policy storage (enabling more fine-grained or complex policies), in exchange for slightly worse latency and throughput.
+%Smaller bit depths still give a \SIrange{2}{4}{$\times$} saving in memory for policy storage (enabling more fine-grained or complex policies), in exchange for slightly worse latency and throughput.
 To overcome this, we investigated bit-stuffing several values into a single word during atomic writeback (as the platform offers both \SIlist{32;64
@@
 \fakepara{Work allocation}
-\Cref{fig:work-alloc-32} shows that our heuristic allocator (\emph{Balanced}) is key in achieving consistent sub-\SIlist{35;65}{\micro\second} latencies/update times, respectively.
+\Cref{fig:work-alloc-32} shows that our heuristic allocator (\emph{Balanced}) is key in achieving consistently low latencies/update times.
 The trend is repeated for all bit depths.
 The constant difference between \emph{Action} and \emph{Update Prep} scales linearly with bit depth, matching storage and lookup work in the serial portion of \emph{ParSa} (plots omitted).
-The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its index, as batching work in contiguous chunks gives some workers only high-dimensionality tilings.
+%The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its index, as batching work in contiguous chunks gives some workers only high-dimensionality tilings.
 The minor gap in lower bound performance between the \emph{Random} and \emph{Balanced} allocators suggests that further optimisations can be made.
 We expect that closing or exceeding this gap may require more complex modelling of hardware thread interactions, which lies far beyond the scope of in-NIC scheduling.
-Some additional complexity may be tolerated, subject to code store limits---scheduling runs exactly once per configuration change, so does not impact per-action code.
+Additionally, throughput occasionally plateaus due to bottlenecks in the output schedule until extra cores are allocated due to the granularity of the above tilings.
+We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
+%Though we have a \SI{4.2}{$\times$} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
 
@@
 
-\begin{figure}
-	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-core-tput-on-per}
-	}
-	\caption{Throughput per added worker in a \Coopfw{} agent.\label{fig:tput-per-core}}
-\end{figure}
-
-An interesting aspect of \Coopfw{} and \emph{ParSa} is that adding cores has both diminishing returns and key thresholds to pass.
-Consider \cref{fig:tput-per-core}, where the throughput per worker decreases over time but occasionally increases sharply.
-Later downward ticks (\numrange{25}{29} workers) correspond to plateaus in throughput.
-This is a problem stemming from the granularity of work items (i.e., tilings in \emph{ParSa}), where our scheduler cannot find a better solution to a bottleneck until extra cores are allocated.
-We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
-Though we have a \SI{4.2}{$\times$} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
+%\begin{figure}
+%	\resizebox{\resultplotwidth}{!}{
+%		\includegraphics{../plots/build/rl-perf-tester/vary-core-tput-on-per}
+%	}
+%	\caption{Throughput per worker in a \Coopfw{} agent.\label{fig:tput-per-core}}
+%\end{figure}
+
+%An interesting aspect of \Coopfw{} and \emph{ParSa} is that adding cores has both diminishing returns and key thresholds to pass.
+%Consider \cref{fig:tput-per-core}, where the throughput per worker decreases over time but occasionally increases sharply.
+%Later downward ticks (\numrange{25}{29} workers) correspond to plateaus in throughput.
+%This is a problem stemming from the granularity of work items (i.e., tilings in \emph{ParSa}), where our scheduler cannot find a better solution to a bottleneck until extra cores are allocated.
+%We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
+%Though we have a \SI{4.2}{$\times$} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
 
@@
 \begin{description}
-	\item[In-NIC.] As described in \cref{sec:agent-environment-communication}, EMEM rings have a median one way delay cross-island of \SI{140}{\nano\second}, giving a \emph{median \SI{34.63}{\micro\second} end-to-end inference latency}.
-	\item[Dedicated Collector.] Offloads hosted in this manner will employ DPDK to maximise performance, giving one-way PCIe delays of \SIrange{0.9}{2.3}{\micro\second} for network packets.
+	\item[\emph{In-NIC.}] As described in \cref{sec:agent-environment-communication}, EMEM rings have a median one way delay cross-island of \SI{140}{\nano\second}, giving a \emph{median \SI{34.63}{\micro\second} end-to-end inference latency}.
+	\item[\emph{Dedicated Collector.}] Employing DPDK to maximise performance, such hosts will add one-way PCIe delays of \SIrange{0.9}{2.3}{\micro\second} for network packets~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
 	A UDP packet carrying \num{20} elements of state in \approachshort{} is \SI{128}{\byte}, so costs \SI{1}{\micro\second} to forward, and the reply state-action pair invokes a slightly higher cost.
 	\emph{This gives an end-to-end inference latency of \SI{517.9}{\micro\second}}.
-	\item[vNF Offload.] \Textcite{DBLP:journals/cm/CzivaP17} show that more lightweight vNF frameworks like GNF~\parencite{DBLP:journals/cm/CzivaP17} and ClickOS~\parencite{DBLP:conf/nsdi/MartinsAROHBH14} add \SIrange{45}{55}{\micro\second} \emph{additional} RTT latency above PCIe costs.
+	\item[\emph{vNF Offload.}] \Textcite{DBLP:journals/cm/CzivaP17} show that more lightweight vNF frameworks like GNF~\parencite{DBLP:journals/cm/CzivaP17} and ClickOS~\parencite{DBLP:conf/nsdi/MartinsAROHBH14} add \SIrange{45}{55}{\micro\second} \emph{additional} RTT latency above PCIe costs.
 	\emph{This gives an end-to-end inference latency of \SI{572.9}{\micro\second}}.
 \end{description}
-Using these estimates, in-NIC classical RL inference offers \SIlist{14.96;16.54}{$\times$} speedups in latency over collector and vNF deployments respectively (assuming no steering cost in either case).
-We also contrast these against deep RL policies on network tasks, which can take \SI{3}{\milli\second} to compute~\parencite{DBLP:journals/corr/abs-1910-04054}---2 orders of magnitude above \approachshort{} with identically sized (20-dim) input vectors.
+Thus, in-NIC classical RL offers \SIlist{14.96;16.54}{$\times$} speedups in latency over collector and vNF deployments respectively.
+We contrast these against deep RL policies on network tasks, which can take \SI{3}{\milli\second} to compute~\parencite{DBLP:journals/corr/abs-1910-04054}---2 orders of magnitude above \approachshort{} with identically sized (20-dim) input vectors.
 %In the name of fairness, we assume that rule installation uses same mechanism we recommend, but show how bad it can get? I.e. with rule installation costs etc.
@@
 Passing this traffic over the NFP device running \approachshort, no packet losses were incurred at any rate of RL actions.
-
 We show the effect of RL workloads on the round-trip latencies of cross traffic through \cref{fig:dataplane-coop}.
-As observed latencies do not obey a normal distribution (particularly \SIlist{256;1518}{\byte}, which are bimodal), we employ a one-tailed \emph{Mann-Whitney U test} to mark statistically significant population increases in latency ($p < 0.05$) with a ``+''.
+As observed latencies do not obey a normal distribution, we employ a one-tailed \emph{Mann-Whitney U test} to mark statistically significant population increases in latency ($p < 0.05$) with a ``+''.
 In general, statistically significant latency increases concentrate around smaller packet sizes.
@@
 
-The anomalous result is \SI{128}{\byte} packets at \num{3000} RL action/update computations per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase, shown in \cref{fig:dataplane-example}.
-This is observed through a shift of some packet latencies from the mode towards the tail, but no other changes in the distribution.
-In the above context, we believe that the inbound request rate is weakly synchronised with inbound packets, causing a higher level of burstiness around accesses to the CPP bus.
-We expect that dedicated hardware/FPGA designs can avoid this problem by having dedicated \inring{}/\outring{} access mechanisms for an \approachshort{} agent.
+The anomalous result is \SI{128}{\byte} packets at \num{3000} RL updates per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase, shown in \cref{fig:dataplane-example}.
+%This is observed through a shift of some packet latencies from the mode towards the tail, but no other changes in the distribution.
+We believe that the inbound request rate is weakly synchronised with traffic, causing a higher level of burstiness around accesses to the CPP bus.
+We expect that FPGA designs can avoid this problem by having dedicated \inring{}/\outring{} access mechanisms for an \approachshort{} agent.
 
@@
 	\end{subfigure}
-	\caption{Effects on tail latency of cross-traffic caused by different loads of off-path RL compute.
+	\caption{Effects on tail latency of cross-traffic caused by off-path RL compute.
 Statistically significant increases in population latency are concentrated on smaller packet sizes, and are typically sub-\SI{78}{\nano\second}.\label{fig:dataplane-coop}}
@@
 \fakepara{Resource requirements}
-\Cref{tab:resources} shows how \approachshort{} consumes shared memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
+\Cref{tab:resources} shows how \approachshort{} consumes memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
 As one program is installed per ME, these results represent the minimum and maximum resource use on a single island (i.e., without replacing P4 workers).
-We observe negligible costs on shared EMEM ($\sim$\SI{4}{\mebi\byte}), incurred due to hashtables for past state and rewards.
-The most significant costs arise due to policy data (\SI{405}{\kibi\byte} shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation and remain constant regardless of compute unit usage.
+We observe negligible costs on shared EMEM ($\sim$\SI{4}{\mebi\byte}) from state/reward hashtables.
+The most significant costs arise due to policy data (\SI{405}{\kibi\byte} shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation.
 This is a high upfront cost on per-island resources (CLS/CTM)---\approachshort{} leaves resources for other off-path dataplane applications, but is fairest from 3 cores onwards.
@@
 \caption{NFP memory use due to \approachshort{} using \numlist{1;4} MEs (\SI{32}{\bit}).
-CLS and CTM are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM are shared between all NFP programs on a NIC.\label{tab:resources}}
+CLS and CTM are shared between all programs on the same island, while EMEM and IMEM are shared between all NFP programs on a NIC.\label{tab:resources}}
 \resizebox{\linewidth}{!}{
@@
 Setup of \approachshort{} uses two packet types: \emph{setup}, which contains learning (hyper-)parameters and most aspects of a policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
-We found that setup and tiling packets take a mean \qtylist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing an agent to be swapped between online and offline operation (or repurposed for another task) painlessly.
-Online/offline swaps are useful when an agent should cease learning (i.e., when performance has converged), or if a change in the environment suggests that more training is needed.
-Tiling packet processing scales linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
+We found that setup and tiling packets take a mean \qtylist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing an agent to be swapped between online and offline operation (or repurposed) painlessly.
+These swaps are useful when an agent should cease learning (i.e., when performance converges), or if a change in the environment suggests that more training is needed.
+%Tiling packet processing scales linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
 Online-offline swaps for \Coopfw{} exhibit similar cost, however the need for explicit scheduling means that policy/tiling \emph{structure} changes (including the first complete setup) take \SI{422.63}{\micro\second} for the full-size policy described above.
@@
 Policy \emph{data} changes require no additional work in any case, resolving purely to \texttt{memcpy}s.
-
 We observed that firmware installation (i.e., changing from \Indfw{} to \Coopfw{}, bit depth, or increasing maximum policy sizes) took a mean time of \SI{38.83}{\second}.
-In the event that appropriate firmwares are not pre-compiled, we found that compiling and linking both \approachshort{} and the P4 toolchain took around \SI{35}{\second}, while changing only \approachshort{} parameters required around \SI{25}{\second}.
+%In the event that appropriate firmwares are not pre-compiled, we found that compiling and linking both \approachshort{} and the P4 toolchain took around \SI{35}{\second}, while changing only \approachshort{} parameters required around \SI{25}{\second}.
+Compiling and linking both \approachshort{} and the P4 toolchain took around \SI{35}{\second}, while changing only \approachshort{} parameters required around \SI{25}{\second}.
 These results show that \approachshort{} can be easily adapted and altered by network administrators once in place, and illustrates an advantage of SoC-based SmartNICs.
@@
 \section{Potential Integrations}\label{sec:potential-integrations}
-To show the applicability and use of \approachshort{}, we propose an ideal integration which would benefit from in-NIC RL; online DDoS attack mitigation.
-We support this using other state-of-the-art P4/PDP developments, and discuss how best to balance the concerns of online training with throughput in widespread deployment.
-Additionally, we discuss how network administrators may combine offline (\Indfw) and online (\Coopfw) agents to achieve online learning while maximising action throughput in the majority of the network.
+To show the applicability and use of \approachshort{}, we propose an integration which would benefit from in-NIC RL; online DDoS mitigation.
+We support this using state-of-the-art P4 developments, and discuss how to balance online training with throughput in widespread deployment.
+%Additionally, we discuss how network administrators may combine offline (\Indfw) and online (\Coopfw) agents to achieve online learning while maximising action throughput in the majority of the network.
 
@@
 Classical RL has seen recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
-\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and local, per-flow state to monitor how flows respond to bandwidth changes and packet drop---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} (observing how flow behaviour reacts to a change in rate limits) with more allowance for congestion-unaware protocols.
-Actions then move flows up or down in punishment levels according to a finite state machine.
+\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and per-flow state to monitor how flows respond to bandwidth changes and packet drop.
+%---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} (observing how flow behaviour reacts to a change in rate limits) with more allowance for congestion-unaware protocols.
+Actions move flows up or down in punishment levels.
 
 \fakepara{Why in-NIC?}
-This approach relies on co-hosting traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
-However, packet mirroring and offloading RL computation to a host (potentially over layers of virtualisation) are all sources of additional, consistent state-action latency.
-Both traffic statistics collection \emph{and} data-driven learning must be executed on such hosts/network functions.
-Unless running these stages in a dedicated pipeline (adding further processing latency), resource contention between these processes will further impact tail latencies.
+This approach co-hosts traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
+However, packet mirroring and offloading RL computation to a host/vNF are all sources of additional, consistent state-action latency.
+Both traffic measurement \emph{and} data-driven learning must be executed in this way.
+Unless running these stages in a dedicated pipeline (adding further processing latency), resource contention between processes will further impact tail latencies.
 Naturally, this requires high-performance hardware to be stacked at network egress points, potentially beyond reasonable space, power, or ventilation constraints.
@@
 The required global state (load measures from network paths) must still come from elsewhere in the network; this is now the element at highest risk of becoming stale, but the least likely to vary significantly in response to individual actions.
-
 The original work uses theoretical ``ground truth'' rewards, whose correct implementation and designs were left as an open challenge. 
@@
 Before each monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{\outring{} Ring} for any generated actions.
-As we note in \cref{sec:limitations}, these actions would be placed into a small hash table and simultaneously exported to the attached controller to be inserted as P4 rules in batches.
+As we note in \cref{sec:limitations}, these actions would be placed into a local hash table \emph{and} exported to the attached controller to be inserted as P4 rules in batches.
 Afterwards, packet ingress timestamps would be used to emulate the \emph{Timed Random Sequential} (TRS) scheduler used by the anti-DDoS agents for rate-controlled work, where state vectors would be selected and passed into \approachshort.
@@
 
-Reducing state-action latency (i.e., with \Coopfw) is useful for minimising the noise inherent in learning.
-However, an agent is limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
+%Reducing state-action latency (i.e., with \Coopfw) is useful for minimising the noise inherent in learning.
+Agents here are limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
 Accordingly, this use-case benefits most from an increase in \emph{throughput} using \Indfw{}.
-In this context, higher throughput means that network flows are more likely to be judged in every timestep, even when flow cardinality is high---making it more likely that changes in flow behaviour will be observed, acted on, and learned from.
-
-We note that the TRS scheduler is designed to handle large numbers of attack flows, combining seen state vectors over time while the asynchronous agent is itself busy.
-A number of (attack) flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
+Higher throughput means that network flows are more likely to be judged in every timestep, even when flow cardinality is high---making it more likely that changes in flow behaviour will be observed, acted on, and learned from.
+We note that the TRS scheduler is designed to handle large numbers of attack flows, combining new state vectors while the asynchronous agent is itself busy.
+A number of flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
 As shown in \cref{sec:evaluation}, \approachshort{} far exceeds the throughput of host offloading.
@@
 \subsection{Network Deployment Considerations}
-The two compute models discussed above need not be homogeneously deployed.
 In a networked deployment, a subset of \approachshort{} nodes could be \coopfw{} agents, training online, while most other nodes run \indfw{} designs to meet throughput guarantees.
 The control plane would then combine, downsample, and distribute these improved policies between offline agents.
-This can be taken further still, using policy deltas or execution traces to enable out-of-path transfer learning for more complex models such as neural networks.
+This can be taken further, using policy deltas to enable out-of-path transfer learning for more complex models such as neural networks.
 
@@
 \emph{IIsy}~\parencite{DBLP:conf/hotnets/XiongZ19} shows how \emph{classical ML inference} (SVMs, Na\"{i}ve Bayes, etc.) can be converted into match-action tables compatible with \emph{any} P4 deployment.
-They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA (at line rate in most cases).
-Neither considers online model training.
+They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA.
+%Neither considers online model training.
 
 A recent line of research in the community has been to investigate \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate packet classification.
-\emph{BaNaNa SPLIT} shows this as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} to reduce latency imposed by GPU batching and transfer, but fully-connected layers are but can be accelerated further by NICs.
+\emph{BaNaNa SPLIT} shows this as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is carried out on the \emph{CPU} to reduce latency imposed by GPU batching and transfer, but fully-connected layers can be accelerated further by NICs.
 %However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}; the authors hypothesise that networking hardware would be a better fit for some parts of the inference pipeline.
-Full, in-network packet tagging and classification by pre-trained BNNs is shown by \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
+In-network packet tagging and classification by pre-trained BNNs is shown by \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
 N3IC achieves packet inference in \SI{45}{\micro\second} on the NFP, and \SI{0.3}{\micro\second} on NetFPGA for \SI{256}{\bit} inputs.
@@
 We expect that a NetFPGA implementation of \approachshort{} would enjoy a similar factor of speedup.
-No works investigate the runtime training of BNN function approximators in-NIC.
+%No works investigate the runtime training of BNN function approximators in-NIC.
 
@@
 
-We stress that none of the other approaches listed here (or that we have seen) tackle the issue of \emph{online learning and control} in-network---we believe \approachshort{} has broken new ground in this regard.
+We stress that none of these approaches (or that we have seen) tackle the issue of \emph{online learning and control} in-network---we believe \approachshort{} has broken new ground in this regard.
 

### 2021-06-10 paper/paper-10pt.tex
Hmm.

@@
 %This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
-Currently, low port-density devices like these are most likely to have specific general-purpose compute, as they are designed to for high-performance offloading and middlebox development.
+Currently, low port-density devices like these are most likely to have specific general-purpose compute, as they are designed for high-performance offloading and middlebox development.
 %Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
@@
 We use a simple first-fit work placement algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core (as the NFP uses hardware threads).
-Each work item is a separate \emph{tiling} over a list of dimensions.%, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
+Each work item is a separate \emph{tiling} over a list of dimensions.
+%, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
 The cost of any work item (from its dimension count and memory location) was empirically measured offline, and we weigh the total cost per core based on the number of minion threads available.
@@
 To make the comparison fair in the context of many-core CPU environments, we include per-core throughput.
-\Indfw{} achieves \SI{2.82}{$\times$} higher offline throughput than commodity \emph{Collector} hardware in spite of the NFP-6480 having a considerably slower clock speed (\SI{0.29}{$\times$}).
+\Indfw{} achieves \SI{2.82}{$\times$} higher offline throughput per core than commodity \emph{Collector} hardware in spite of the NFP-6480 having a considerably slower clock speed (\SI{0.29}{$\times$}).
 When compounded with the abundance of such weaker chips, in-NIC RL is able to deliver much higher throughput.

### 2021-06-11 paper/paper-10pt.tex
Dimitris feedback on 10pt, some attempts at host ParSa

@@
 \usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists]{savetrees}
+
+%\usepackage{fnpos}
+%\usepackage{dblfnote}
+%\usepackage{ftnright}
 
@@
 \begin{abstract}
-Automatic, data-driven networking is becoming more capable and widely researched, mainly driven by the efficacy of \emph{Deep Reinforcement Learning} (DRL) algorithms.
+Automatic, data-driven networking is becoming more capable and widely researched, partly driven by the efficacy of \emph{Deep Reinforcement Learning} (DRL) algorithms.
 Yet the complexity of both DRL inference and learning force these tasks to be offloaded \emph{away from the dataplane} to hosts, harming latency-sensitive applications.
@@
 In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
-\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, improving on offloading, and are quickly reconfigurable at runtime compared to reinstalling device firmware/designs.
+\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, improving on offloading, and are quickly reconfigurable at runtime compared to reinstalling device firmware.
 
@@
 Automatic network optimisation, control, and defence are at last becoming commonplace.
-Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing automatic traffic/media optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17,DBLP:conf/conext/GiladSGRS20}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
+Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing automatic traffic optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17,DBLP:conf/conext/GiladSGRS20}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
 In RL methods, every change and its effects improve future decisions made by an agent.
@@
 
-To enable \emph{online in-NIC learning}, we return to \emph{classical} RL methods.
-In particular, we focus on tile-coding with one-step temporal-difference algorithms such as Sarsa.
+In this paper, we enable \emph{online in-NIC learning} by returning to \emph{classical} RL methods.
+In particular, we focus on tile coding with one-step temporal-difference algorithms such as Sarsa~\parencite{RL2E}.
 %These choices have important benefits for in-NIC execution.
-These functions do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays, and have simple update and inference logic.
-Tile-coding in particular admits many optimisations, being an embarrassingly parallel problem.
+These functions have lower model capacity, they do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays, and have simple update and inference logic.
+In addition, they have shown promise in other network use cases~\parencite{DBLP:journals/eaai/MalialisK15,DBLP:journals/tnsm/SimpsonRP20}.
+Tile coding in particular admits many optimisations, being an embarrassingly parallel problem.
 Using fixed-point arithmetic, we solve the lack of floating-point support in PDP hardware \emph{and} enable new optimisations.
-Moreover, the P4 dataplane can offer runtime control over which flows/packets are monitored.
+Moreover, the P4 dataplane can offer runtime control over which flows are monitored.
 %Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair, safeguarding the limited memory of the target devices.
@@
 
-The question we investigate is: can online RL be brought to the dataplane by returning to these computationally simpler methods, to act on locally extracted state?
-Can it be made more efficient \emph{by dataplane hardware}?
-Through this work, \approachshort{}, we can comfortably answer ``yes'' on both counts.
+%The question we investigate is: can online RL be brought to the dataplane by returning to these computationally simpler methods, to act on locally extracted state?
+%Can it be made more efficient \emph{by dataplane hardware}?
+%Through this work, \approachshort{}, we can comfortably answer ``yes'' on both counts.
+%Through \approachshort{}, w
+We show that online RL can be brought to the dataplane by such methods, where it can act on locally extracted state and is made \emph{more efficient by dataplane hardware}.
 In particular, we exploit how SmartNIC devices often expose general-purpose compute to provide path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
@@
 	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands of network traffic analysis (\cref{sec:design}),
-	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm to accelerate tile-coded policy inference and updates (\cref{alg:parsa}),
+	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm to accelerate tile coded policy inference and updates (\cref{alg:parsa}),
 	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
@@
 %\section{Preliminaries: RL for in-network computation}\label{sec:motivation}
-By combining recent trends in PDP hardware with insights from the ML/RL communities (past and present), we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL methods and quantisation make this computationally feasible.
-Crucially, we focus on low port-density devices such as SmartNICs and the NetFPGA.
-The designs of these devices make it reasonable to move processing \emph{outside of the packet pipeline}; SmartNICs expose additional programmable cores, and NetFPGAs allow for the synthesis of independent functional units.
+By combining recent trends in PDP hardware with insights from the ML community (past and present), we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL and quantisation make this feasible.
+Crucially, we focus on low port-density devices.
+We envision that this processing is suited to the network edge or key vantage points, rather than core switches.
+These devices' designs make it reasonable to move processing \emph{outside of the packet pipeline}; SmartNICs expose additional cores, and NetFPGAs allow for the synthesis of independent functional units.
 %The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
-The main benefit is that core control logic can be moved onto the NIC \emph{without impacting packet processing rates}.
+The main benefit is that control logic can be moved onto the NIC \emph{without impacting packet processing}.
 %?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
-As Ethernet speed increases, packet deadlines grow tighter; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second} at \SI{40}{\giga\bit\per\second}, giving a \SI{3.99}{\micro\second} deadline on Netronome (which has \num{312} P4 pipelines).
+As Ethernet speed increases, packet deadlines grow tighter; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second} at \SI{40}{\giga\bit\per\second}, giving for example a \SI{3.99}{\micro\second} deadline on Netronome (which has \num{312} P4 pipelines).
 In-pipeline execution thus increases the risk of packet drops or stalls.
@@
 RL algorithms are generally agnostic to the policy approximation used, and classical single-step methods are simple---most of the compute work is occupied by the policy itself.
-The single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to update and learn a policy online.
+The single-step, semi-gradient Sarsa algorithm~\parencite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to update and learn a policy online.
 
@@
 This cost is what drives us to \emph{bring online reinforcement learning to the dataplane}---referring again to \cref{fig:state-slip}, we would place P4-based state measurement ($t_1$), simple policy approximation ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
-In networks, actions are most likely to be installed on backbone switches, bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts---all of which make in-NIC training a perfect fit.
+In networks, actions and analysis are most likely to be installed on bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts---all of which make in-NIC training a perfect fit.
 As such, our goal is to collocate \emph{all} functions which comprise an RL agent on the same chip or device.
@@
 
-We return to what classical RL methods can offer us; in particular, the simple linear function approximation given by \emph{tile coding} \cite[pp.\ \numrange{217}{221}]{RL2E}.
+We return to what classical RL methods can offer us; in particular, the simple linear function approximation given by \emph{tile coding}~\parencite[pp.\ \numrange{217}{221}]{RL2E}.
 A policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
@@
 Inference (choosing an action) is simply summing over all such lists to obtain a final action preference list, selecting the maximum.
-Updating uses this same list of lists after the next decision is made, adjusting the value of the last selected action using a value $\delta_t$ computed through Sarsa.
+Updating uses this same list of lists after the next decision is made, adjusting the value of the last selected action using a \emph{temporal-difference value} ($\delta_t$) computed by Sarsa.
 Crucially, this internal representation has no data dependencies between any tile calculations for an input.
@@
 
-For instance, a policy with $m$ sets of tiles (each having its own set of input dimensions), each containing $n$ overlapping tilings then creates $m \times n$ separate tasks.\footnote{Work items could be retrieving each individual action's value (giving $m \times n \times a$ tasks for $a$ actions).
+For instance, a policy with $m$ tiling sets (each having a set of input dimensions), with $n$ overlapping tilings then creates $m \times n$ separate tasks.\footnote{Work items could be retrieving each individual action's value (giving $m \times n \times a$ tasks for $a$ actions).
 This requires tile coding to be repeated $a$ times per tile, which is likely extremely wasteful outside of heavy parallelism.}
-As an example, a real-world policy size which we examine later (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets of \num{8} tilings creates \num{129} work items.
+As an example, a real-world policy size which we examine (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets of \num{8} tilings creates \num{129} work items.
 Computing a tile hit uses one division per dimension, which is an expensive ALU operation to require.
@@
 %We omit the simple logic for $\epsilon$-greedy action selection (which we implement), and note that modification to other single-step algorithms such as \emph{Q-learning} would be trivial.
-%We detail our communication primitives in \cref{sec:intra-agent-communication}, and our work scheduling strategy in \cref{sec:work-allocation}, eliding the details of tile-coding as they are well-understood.
+%We detail our communication primitives in \cref{sec:intra-agent-communication}, and our work scheduling strategy in \cref{sec:work-allocation}, eliding the details of tile coding as they are well-understood.
 %In general it is also possible for the \emph{Ctl} task to act as an additional \emph{Minion} in its parallel sections; we were limited here by code store requirements on the NFP.
@@
 \subsection{Reconfigurability}
-\approachshort{} allows policy design and learning parameters to be changed at runtime using at most two control packets.
-For instance, design changes are useful at the end of learning (moving from online to offline), or when trying to train a new policy for another problem from the same vantage point.
-Parameter changes are useful when an online agent must become more (or less) adaptive to new data (i.e., after detecting a changepoint in traffic).
+\approachshort{} allows policy design and learning parameters to be changed at runtime using two control packets.
+Design changes are used at the end of learning (moving from online to offline), or when trying to train a new policy for another task from the same vantage point.
+Parameter changes allow an online agent to become more (or less) adaptive to new data (i.e., after detecting a changepoint in traffic).
 This extends to policy data, which may be imported from a pre-trained model and exported via PCIe to the host machine.
@@
 \Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy.\footnote{For brevity, we omit integer numpy results---against a float implementation, median action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}
-We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using 4 MEs of the NFP-6480 (\SIlist{15;21}{$\times$} speedups over a \emph{Collector} host).
+We show that when using 4 MEs of the NFP-6480, \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse (\SIlist{15;21}{$\times$} speedups over a \emph{Collector} host).
 Importantly, \Indfw{} achieves lower median state-action latencies (\SI{2.79}{$\times$}) \emph{and} update times (\SI{2.63}{$\times$}) than a dedicated \emph{Collector} host \emph{while requiring only a single core or dedicated functional unit}.
@@
 
-The anomalous result is \SI{128}{\byte} packets at \num{3000} RL updates per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase, shown in \cref{fig:dataplane-example}.
+An anomaly is \SI{128}{\byte} packets at \num{3000} RL updates per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase, shown in \cref{fig:dataplane-example}.
 %This is observed through a shift of some packet latencies from the mode towards the tail, but no other changes in the distribution.
-We believe that the inbound request rate is weakly synchronised with traffic, causing a higher level of burstiness around accesses to the CPP bus.
+We believe that the inbound request rate is weakly synchronised with traffic, causing higher burstiness of accesses to the CPP bus.
 We expect that FPGA designs can avoid this problem by having dedicated \inring{}/\outring{} access mechanisms for an \approachshort{} agent.

### 2021-06-11 code/rl-perf-tester/src/lib.rs

@@ -1,9 +1,10 @@
 mod config;
 mod constants;
 mod experiment;
+mod parsa;
 mod stress;
 
-pub use self::{config::*, constants::*, experiment::*, stress::*};
+pub use self::{config::*, constants::*, experiment::*, parsa::*, stress::*};
 
 use control::{GlobalConfig, SendStateConfig, Setup, SetupConfig, Tile, TilingsConfig};
 use core::fmt::Debug;

### 2021-06-11 code/rl-perf-tester/src/parsa/mod.rs

@@ -1,0 +1,216 @@
+mod q;
+mod tilecode;
+mod writeback;
+
+pub use self::{q::*, tilecode::*, writeback::*};
+
+use bus::{Bus, BusReader};
+use control::{Setup, Tile};
+use num_traits::{Num, NumCast, PrimInt};
+use std::{
+	cell::UnsafeCell,
+	collections::HashMap,
+	ops::*,
+	sync::{atomic::Ordering, Arc},
+};
+
+type Policy<T> = Arc<Vec<UnsafeCell<Vec<T>>>>;
+
+#[derive(Clone)]
+enum ParsaMessage<T> {
+	Action(Vec<T>),
+	Update(T, usize, Vec<T>),
+	Exit,
+}
+
+fn parsa_control(
+	wb: Writeback,
+	state: &Vec<i32>,
+	setup: &Setup<i32>,
+	bus: &mut Bus<ParsaMessage<i32>>,
+	work_ct: usize,
+	past_states: &mut HashMap<i32, (i32, usize, Vec<i32>)>,
+	past_rewards: &mut HashMap<i32, i32>,
+) -> usize {
+	wb.clear_actions();
+	bus.broadcast(ParsaMessage::Action(state.clone()));
+	wb.gather(work_ct);
+
+	let (val, act) = wb.select_max();
+
+	if setup.do_updates {
+		match (past_states.get(&state[0]),past_rewards.get(&1)) {
+			(Some((l_val, l_act, l_state)), Some(reward)) => {
+				let dt = reward + setup.gamma.q_mul(val, setup.quantiser_shift as i32) - l_val;
+				let dt = dt.q_mul(setup.alpha, setup.quantiser_shift as i32);
+
+				bus.broadcast(ParsaMessage::Update(dt, *l_act, l_state.clone()));
+				wb.gather(work_ct);
+			},
+			_ => {},
+		}
+
+		past_states.insert(state[0], (val, act, state.clone()));
+	}
+
+	act
+}
+
+fn tile_code(
+	state: &Vec<i32>,
+	task: &Task,
+	shift_amts: &Vec<i32>,
+	setup: &Setup<i32>,
+	tile_set_data: &Vec<UsefulTileSet>,
+	adjusted_maxes: &Vec<i32>,
+	widths: &Vec<i32>,
+) -> usize {
+	if task.is_bias {
+		return 0;
+	}
+
+	let mut width_product: usize = 1;
+	let mut local_tile = tile_set_data[task.tiling_set_idx].start_tile;
[…]

### 2021-06-11 code/rl-perf-tester/src/parsa/q.rs

@@ -1,0 +1,79 @@
+use std::ops::*;
+
+pub trait Quantable {
+	fn q_mul(self, other: Self, depth: Self) -> Self;
+	fn q_div(self, other: Self, depth: Self) -> Self;
+}
+
+// pub struct Quant<T: Quantable>{
+// 	pub val: T,
+// 	pub depth:
+// };
+
+// impl<T: Quantable> Add for Quant<T> {
+// 	type Output = Self;
+
+// 	fn add(self, other: Self) -> Self {
+// 		Quant(self.0 + other.0)
+// 	}
+// }
+
+impl Quantable for i32 {
+	fn q_mul(self, other: Self, depth: Self) -> Self {
+		let mut intermediate = self as i64;
+		intermediate *= other as i64;
+		intermediate += 1 << (depth - 1);
+
+		(intermediate >> depth) as i32
+	}
+
+	fn q_div(self, other: Self, depth: Self) -> Self {
+		let intermediate = (self as i64) << depth;
+		(intermediate / other as i64) as i32
+	}
+}
+
+impl Quantable for u32 {
+	fn q_mul(self, other: Self, depth: Self) -> Self {
+		let mut intermediate = self as u64;
+		intermediate *= other as u64;
+		intermediate += 1 << (depth - 1);
+
+		(intermediate >> depth) as u32
+	}
+
+	fn q_div(self, other: Self, depth: Self) -> Self {
+		let intermediate = (self as u64) << depth;
+		(intermediate / other as u64) as u32
+	}
+}
+
+impl Quantable for i16 {
+	fn q_mul(self, other: Self, depth: Self) -> Self {
+		let mut intermediate = self as i32;
+		intermediate *= other as i32;
+		intermediate += 1 << (depth - 1);
+
+		(intermediate >> depth) as i16
+	}
+
+	fn q_div(self, other: Self, depth: Self) -> Self {
+		let intermediate = (self as i32) << depth;
+		(intermediate / other as i32) as i16
+	}
+}
+
+impl Quantable for i8 {
+	fn q_mul(self, other: Self, depth: Self) -> Self {
+		let mut intermediate = self as i16;
+		intermediate *= other as i16;
+		intermediate += 1 << (depth - 1);
+
+		(intermediate >> depth) as i8
+	}
+
+	fn q_div(self, other: Self, depth: Self) -> Self {
+		let intermediate = (self as i16) << depth;
+		(intermediate / other as i16) as i8
+	}
+}

### 2021-06-11 code/rl-perf-tester/src/parsa/tilecode.rs

@@ -1,0 +1,74 @@
+use super::*;
+
+// pub fn tile_code_with_cfg_single<T: Quantable>(
+// 	__addr40 __declspec(emem) tile_t *state,
+// 	__declspec(cls) struct rl_config *cfg,
+// 	uint8_t bias_tile_exists,
+// 	uint8_t work_idx,
+// 	uint16_t tiling_set_idx,
+// 	uint16_t tiling_idx
+// ) {
+// 	uint16_t dim_idx;
+// 	uint32_t local_tile;
+// 	uint32_t width_product = cfg->num_actions;
+
+// 	tile_t emergency_vals[4] = {0};
+
+// 	// check if there's a bias tile.
+// 	// => yes? set 0 has size 1, others have size cfg->tilings_per_set
+// 	// => no? all have size cfg->tilings_per_set
+// 	if (bias_tile_exists && work_idx == 0) {
+// 		return 0;
+// 	}
+
+// 	// can also get by remul'ing
+// 	local_tile = cfg->tiling_sets[tiling_set_idx].start_tile;
+
+// 	local_tile += tiling_idx * cfg->tiling_sets[tiling_set_idx].tiling_size;
+
+// 	// These goes over each dimension in the top-level loop
+// 	for (dim_idx = 0; dim_idx < cfg->tiling_sets[tiling_set_idx].num_dims; ++dim_idx) {
+// 		uint16_t active_dim = cfg->tiling_sets[tiling_set_idx].dims[dim_idx];
+// 		uint8_t reduce_tile;
+// 		tile_t val = state[active_dim];
+
+// 		// This is what varies per tiling in a set.
+// 		tile_t shift = tiling_idx * cfg->shift_amt[active_dim];
+
+// 		tile_t max = cfg->adjusted_maxes[active_dim] - shift;
+// 		tile_t min = cfg->mins[active_dim] - shift;
+
+// 		// to get tile in dim... rescale dimension, divide by window.
+// 		val = (val < max) ? val : max;
+// 		reduce_tile = val == max;
+
+// 		// WE NEED TO CAST VAL AS A UTILE *BEFORE* USE
+// 		// OTHERWISE DIVISION WILL BE WRONG DUE
+// 		// TO POTENTIAL OVERFLOW
+// 		val = (val > min) ? val - min : 0;
+
+// 		// emergency_vals[dim_idx] = val;
+
+// 		local_tile +=
[…]

### 2021-06-11 code/rl-perf-tester/src/parsa/writeback.rs

@@ -1,0 +1,69 @@
+use flume::{Receiver, Sender};
+use std::sync::{
+	atomic::{AtomicI32, AtomicU32, Ordering},
+	Arc,
+};
+
+pub struct WbSite {
+	pub vals: Vec<AtomicI32>,
+	// pub acks: AtomicU32,
+	pub acker: Sender<()>,
+	pub rxer: Receiver<()>,
+}
+
+impl WbSite {
+	pub fn ack(&self) {
+		let _ = self.acker.send(());
+	}
+
+	pub fn gather(&self, target: usize) {
+		let mut count = 0;
+		while target > count {
+			if self.rxer.recv().is_ok() {
+				count += 1;
+			} else {
+				break;
+			}
+		}
+	}
+
+	pub fn clear_actions(&self) {
+		for val in &self.vals {
+			val.store(0, Ordering::Relaxed);
+		}
+	}
+
+	pub fn select_max(&self) -> (i32, usize) {
+		let mut max = None;
+
+		for (i, val) in self.vals.iter().enumerate() {
+			let val_read = val.load(Ordering::Relaxed);
+
+			match max {
+				None => {
+					max = Some((val_read, i));
+				},
+				Some((l_val, l_i)) if val_read > l_val => {
+					max = Some((val_read, i));
+				},
+				_ => {}
+			}
+		}
+
+		max.unwrap()
+	}
+}
+
+pub type Writeback = Arc<WbSite>;
+
+pub fn new_writeback(num_actions: usize) -> Writeback {
+	let (acker, rxer) = flume::unbounded();
+	let mut vals = Vec::new();
+	vals.resize_with(num_actions, || AtomicI32::new(0));
+
+	Arc::new(WbSite {
+		vals,
+		acker,
+		rxer,
+	})
+}

### 2021-06-11 paper/paper-10pt.tex
Before ImageUp

@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.65\linewidth]{figures/arch-with-p4}
+	\includegraphics[keepaspectratio, width=0.6875\linewidth]{figures/arch-with-p4}
 	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
@@
 %These choices have important benefits for in-NIC execution.
-These functions have lower model capacity, they do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays, and have simple update and inference logic.
+While these functions have lower model capacity, they do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays, and have simple update and inference logic.
 In addition, they have shown promise in other network use cases~\parencite{DBLP:journals/eaai/MalialisK15,DBLP:journals/tnsm/SimpsonRP20}.
@@
 However, many compatible devices have legacies long predating these developments. 
-For instance, many-core SOC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
+For instance, many-core SoC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
 %This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
@@
 %Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
-Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of deep learning accelerators (which lack packet switching functionality).
+Regardless, floating-point operations key to ML workloads are very rarely supported outside of deep learning accelerators (which lack packet switching functionality).
 Moreover, the limited memory and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
@@
 In networking, this allows for learning to respond to on-device state or to handle rapidly evolving problems.
-RL algorithms are generally agnostic to the policy approximation used, and classical single-step methods are simple---most of the compute work is occupied by the policy itself.
-The single-step, semi-gradient Sarsa algorithm~\parencite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to update and learn a policy online.
+RL algorithms are generally agnostic to the policy approximation used, and classical single-step methods are simple---most of the compute work is occupied by the policy.
+E.g., single-step, semi-gradient Sarsa~\parencite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to update and learn a policy online.
 
@@
 \subsection{Representations for embedded ML}
-As PDP hardware universally lacks floating-point arithmetic support, the task of bringing ML models to resource-constrained environments without these capabilities is well-studied.
+PDP hardware universally lacks floating-point arithmetic support, yet the task of bringing ML models to resource-constrained environments is well-studied.
 Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
@@
 \emph{Tile lists} are stored for update computation.
-	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel,alg:parsa})] Threads cooperate on processing state vectors, minimising latency.
+	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel,alg:parsa})] Threads cooperate to process state vectors, minimising latency.
 \emph{Minion} threads have a fixed list of work items, while a \emph{controller} thread sends compute/update commands before awaiting worker completion.
@@
 This allows decisions to be made asynchronously, preventing packet stalling.
-The key insight of this mechanism is that on-chip reward/state signals enjoy first-class support in the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-NIC/chip asynchronous processes or the controller.
+The key insight of this mechanism is that on-chip reward or state signals enjoy first-class support in the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-chip processes or the controller.
 As such, \approachshort{} can receive input from P4 \texttt{extern}s or other, dedicated off-path flow state measurement applications.
@@
 Our implementation uses platform-specific IPC (\emph{EMEM ring buffers}) with hardware signalling on work arrival to achieve this.
-As PDP hardware lacks dynamic memory allocation, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state/configuration/policy data handover.
+As PDP hardware lacks dynamic memory allocation, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state or policy data handover.
 Packet headers are extracted and parsed using the tooling autogenerated by the P4 pipeline.
@@
 
-Our implementation exploits the locality of cores in the NFP, which can be factored into designs on other platforms like NetFPGA as additional off-path functional units, as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-Policy compute/update/configure tasks are passed between cores using direct \emph{next neighbour} registers, signalling all child threads in response.
+Our implementation exploits the locality of cores in the NFP. 
+Policy compute or update tasks are passed between cores using direct \emph{next neighbour} registers, signalling all child threads in response.
 Such on-chip signals cost just $\sim$\SI{20}{\nano\second} per relayed message.
+This can be factored into the design of additional async off-path functional units (similar to \textcite{DBLP:conf/isca/LiLYCSH19}) on platforms like NetFPGA.
 Each core performs atomic adds to a shared preference list and an acknowledgement counter checked by the master thread, implementing our wait-free \emph{ParSa} algorithm (\cref{alg:parsa}).
@@
 \subsection{Reconfigurability}
-\approachshort{} allows policy design and learning parameters to be changed at runtime using two control packets.
+\approachshort{} allows policy design and learning parameters to be changed at runtime using at most two control packets.
 Design changes are used at the end of learning (moving from online to offline), or when trying to train a new policy for another task from the same vantage point.
@@
 This extends to policy data, which may be imported from a pre-trained model and exported via PCIe to the host machine.
-Some aspects must be chosen at compile time; bit depth, \Coopfw/\Indfw, and maximum policy/tiling/state sizes---these govern core operation or allocated memory.
+Some aspects must be chosen at compile time; bit depth, \Coopfw/\Indfw, and maximum policy sizes---these govern core operation or allocated memory.
 Choosing a bit depth of \qtylist[list-pair-separator = { or }]{16;8}{\bit} halves/quarters policy memory costs, allowing more complex problems to be modelled using more dimensions or fine-grained tiles.
@@
 In \Coopfw{}, this creates \num{129} work items across \num{31} workers.
-We choose a larger action than this work to investigate the performance of more complex agents.
+We choose a larger action count than this work to investigate the performance of more complex agents.
 
@@
 This is an important aspect for (automated) allocation of compute resources in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of-time work scheduling which we examine later.
-This also demonstrates the number of cores needed to achieve a given latency/throughput bound on a policy of representative complexity.
+This also demonstrates the number of cores needed to achieve a given latency or throughput bound on a policy of representative complexity.
 Moreover, to demonstrate how these costs vary as policy complexity increases, we vary the number of total dimensions included in the tiling.
@@
 We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, placing an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source via the default firmware.
-We perform throughput/loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements (sampling at \SI{2000}{\hertz} for \qtyproduct[product-units=single]{10 x 10}{\second}).
+We perform throughput and loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements (sampling at \SI{2000}{\hertz} for \qtyproduct[product-units=single]{10 x 10}{\second}).
 This provides maximum throughput in the former case (relying solely on NIC counters for loss counting).
@@
 Using the maximum policy size defined above, we investigate how the memory requirements imposed by \approachshort{} vary with the number of dedicated MEs, over and above a base P4 forwarding plane.
-We report resource use for \SI{32}{\bit} \Indfw{} and \Coopfw{} agents, with hash-tables sized to \num{4096} state-action pairs and \num{16} separate reward values.
+We report resource use for \SI{32}{\bit} \Indfw{} and \Coopfw{} agents, with hash tables sized to \num{4096} state-action pairs and \num{16} separate reward values.
 This captures the relative cardinality of network RL traces to rewards, i.e., many input flows will map to one or few reward values (i.e., DDoS attack size estimation per egress-AS, queue occupancy in the case of AQM per output port).
@@
 Some artefacts of our environment and design choices are visible, such as the addition of new physical cores being more significant than contexts, and the presence of some schedule bottlenecks.
-Most importantly however, \Coopfw{}'s resource demand is tunable at compile time to meet the online training rate and/or action latency required by a task/environment.
+Most importantly however, \Coopfw{}'s resource demand is tunable at compile time to meet the online training rate or action latency required by a task.
 
@@
 Most of the I/O comes \emph{after} this step, causing ALU use to dominate.
-This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the online algorithm.
+This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) agents, where hash table accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the online algorithm.
 %Smaller bit depths still give a \SIrange{2}{4}{$\times$} saving in memory for policy storage (enabling more fine-grained or complex policies), in exchange for slightly worse latency and throughput.
@@
 As one program is installed per ME, these results represent the minimum and maximum resource use on a single island (i.e., without replacing P4 workers).
-We observe negligible costs on shared EMEM ($\sim$\SI{4}{\mebi\byte}) from state/reward hashtables.
+We observe negligible costs on shared EMEM ($\sim$\SI{4}{\mebi\byte}) from state and reward hash tables.
 The most significant costs arise due to policy data (\SI{405}{\kibi\byte} shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation.
@@
 Agents here are limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
-Accordingly, this use-case benefits most from an increase in \emph{throughput} using \Indfw{}.
+Accordingly, this use case benefits most from an increase in \emph{throughput} using \Indfw{}.
 Higher throughput means that network flows are more likely to be judged in every timestep, even when flow cardinality is high---making it more likely that changes in flow behaviour will be observed, acted on, and learned from.
@@
 Event-driven versions of P4 have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}.
-Timer events and device state changes would empower in-network RL use-cases, signalling timesteps for RL agents or new, effective, fine-grained sources of input state.
+Timer events and device state changes would empower in-network RL use cases, signalling timesteps for RL agents or new, effective, fine-grained sources of input state.
 

### 2021-09-23 code/control/src/lib.rs
Almost ready to verify...

@@ -114,17 +114,16 @@ pub fn send_state<T: Tile>(cfg: &mut SendStateConfig, state: Vec<T>) {
 	// write_and_send_state
 }
 
-pub fn generate_policy<T: Tile + Serialize>(cfg: &FakePolicyGeneratorConfig<T>) {
+pub fn policy_size<T: Tile>(setup: &Setup<T>, tiling: &TilingSet) -> usize {
 	let mut tiles: usize = 0_usize;
+	let a = setup.n_actions as usize;
+	let s = setup.tilings_per_set as usize;
 
-	let a = cfg.setup.n_actions as usize;
-	let s = cfg.setup.tilings_per_set as usize;
-
-	let n = cfg.setup.tiles_per_dim as usize;
+	let n = setup.tiles_per_dim as usize;
 
 	let mut bias = false;
 
-	for tiling in cfg.tiling.tilings.iter() {
+	for tiling in tiling.tilings.iter() {
 		if tiling.location.is_none() {
 			bias = true;
 		} else {
@@ -133,9 +132,11 @@ pub fn generate_policy<T: Tile + Serialize>(cfg: &FakePolicyGeneratorConfig<T>) 
 		}
 	}
 
-	let cap = (a * s * tiles) + if bias { a } else { 0 };
+	(a * s * tiles) + if bias { a } else { 0 }
+}
 
-	// let mut policy_data = vec![0; cap];
+pub fn generate_policy<T: Tile + Serialize>(cfg: &FakePolicyGeneratorConfig<T>) {
+	let cap = policy_size(&cfg.setup, &cfg.tiling);
 
 	let policy_data: Vec<T> = (0..cap)
 		.map(|i| T::from_int(((i % 20) as i32) - 10))

### 2021-09-23 code/rl-perf-tester/src/config.rs

@@ -32,3 +32,13 @@ pub struct StressConfig<'a> {
 	pub do_hack: bool,
 	pub do_trial_with_hack: Option<u32>,
 }
+
+pub struct VerifyConfig<'a> {
+	pub transport_cfg: TransportConfig,
+	pub sample_seed: u64,
+	pub policy_seed: Option<u64>,
+	pub rtecli_path: &'a str,
+	pub rtsym_path: &'a str,
+	pub n_samples: u64,
+	pub first_sample: u64,
+}

### 2021-09-23 code/rl-perf-tester/src/lib.rs

@@ -3,8 +3,9 @@ mod constants;
 mod experiment;
 mod parsa;
 mod stress;
+mod verify;
 
-pub use self::{config::*, constants::*, experiment::*, parsa::*, stress::*};
+pub use self::{config::*, constants::*, experiment::*, parsa::*, stress::*, verify::*};
 
 use control::{GlobalConfig, SendStateConfig, Setup, SetupConfig, Tile, TilingsConfig};
 use core::fmt::Debug;
@@ -366,6 +367,7 @@ fn get_cycle_ct(config: &Config, source: CycleSource) -> u64 {
 	cycle_ct * 16
 }
 
+#[derive(Debug)]
 enum BlockDataSource {
 	T1Mem,
 	T2Mem,
@@ -387,8 +389,8 @@ impl BlockDataSource {
 		use BlockDataSource::*;
 
 		match self {
-			T1Mem => "_t1_tiles",
-			T2Mem => "_t2_tiles",
+			T1Mem => "i5._t1_tiles",
+			T2Mem => "i5._t2_tiles",
 			T3Mem => "_t3_tiles",
 			AcVals => "_acvals",
 			AcChoice => "_chosenac",
@@ -396,15 +398,12 @@ impl BlockDataSource {
 	}
 }
 
-fn get_nfp_mem(config: &Config, source: BlockDataSource) -> Vec<u8> {
+fn get_nfp_mem(config: &VerifyConfig, source: &BlockDataSource) -> Vec<u8> {
 	// measure and parse output of cycle-estimate.
 	let mem_text = Command::new(config.rtsym_path)
 		.args(&["-v", source.get_register_name()])
 		.output()
 		.expect("Failed to read NFP memory.");
-
-	// each space-delimited group is a 32-bit number.
-	let mut cycle_ct = 0u64;
 
 	let mut out = vec![];
 	let lines = std::str::from_utf8(&mem_text.stdout[..])
@@ -413,6 +412,7 @@ fn get_nfp_mem(config: &Config, source: BlockDataSource) -> Vec<u8> {
 
 	for line in lines {
 		for hex_word in line.split_whitespace().skip(1) {
+			// each space-delimited group is a 32-bit number.
 			let without_prefix = hex_word.trim_start_matches("0x");
 			let x = u32::from_str_radix(without_prefix, 16).expect("Not a hex-formatted word!");
 			out.extend_from_slice(&x.to_be_bytes())

### 2021-09-23 code/rl-perf-tester/src/main.rs

@@ -7,6 +7,7 @@ use rl_perf_tester::{
 	HostConfig,
 	InstallTimes,
 	StressConfig,
+	VerifyConfig,
 	OUTPUT_DIR,
 	PKTGEN_PATH,
 	RTE_PATH,
@@ -283,6 +284,81 @@ fn main() -> Result<(), Box<dyn Error>> {
 				.about("Stress test the main firmware (hosted on another device) using pktgen-dpdk.")
 		)
 		.subcommand(SubCommand::with_name("parsa").about("Test host performance of the Parsa algorithm."))
+		.subcommand(
+			SubCommand::with_name("verify")
+				.arg(
+					Arg::with_name("src-ip")
+						.short("s")
+						.long("src-ip")
+						.value_name("IPV4")
+						.help("Source IP for control packets. Defaults to selected interface.")
+						.takes_value(true)
+						.default_value("192.168.0.4"),
+				)
+				.arg(
+					Arg::with_name("dst-ip")
+						.short("d")
+						.long("dst-ip")
+						.value_name("IPV4")
+						.help("Destination IP for control packets. Defaults to selected interface.")
+						.takes_value(true)
+						.default_value("192.168.1.141"),
+				)
+				.arg(
+					Arg::with_name("src-port")
+						.short("p")
+						.long("src-port")
+						.value_name("U16")
+						.help("Source UDP port for control packets.")
+						.takes_value(true)
+						.default_value("16767"),
+				)
+				.arg(
+					Arg::with_name("dst-port")
+						.short("P")
+						.long("dst-port")
+						.value_name("U16")
+						.help("Destination port for control packets.")
+						.takes_value(true)
+						.default_value("16768"),
+				)
+				.arg(
+					Arg::with_name("n-samples")
+						.short("n")
+						.long("n-samples")
+						.value_name("NUMBER")
+						.help("Number of random samples to test.")
+						.takes_value(true)
+						.default_value("10000"),
+				)
+				.arg(
+					Arg::with_name("first-sample")
+						.short("f")
+						.long("first-sample")
+						.value_name("NUMBER")
+						.help("First random sample to test.")
+						.takes_value(true)
+						.default_value("0"),
+				)
+				.arg(
+					Arg::with_name("seed")
+						.short("x")
+						.long("seed")
+						.value_name("U64_HEX_NUMBER")
+						.help("Random seed for state generation.")
[…]

### 2021-09-23 code/rl-perf-tester/src/parsa/mod.rs

@@ -4,7 +4,7 @@ mod writeback;
 
 pub use self::{q::*, tilecode::*, writeback::*};
 
-use crate::{ProtoSetup, TimeBreakdown, OUTPUT_DIR};
+use crate::{BlockDataSource, ProtoSetup, TimeBreakdown, OUTPUT_DIR};
 use bus::{Bus, BusReader};
 use control::{Setup, Tile, TilingSet};
 use num_traits::{Num, NumCast, PrimInt};
@@ -247,7 +247,7 @@ pub fn parsa_experiment() {
 
 	// do these, and online/offline.
 	// let worker_cts = [num_cpus::get_physical(), 129];
-	let worker_cts = [num_cpus::get_physical()];
+	let worker_cts = [2]; //num_cpus::get_physical() / 1];
 
 	let host = gethostname::gethostname().into_string().unwrap();
 
@@ -594,3 +594,133 @@ struct PrePrep {
 }
 
 unsafe impl Send for PrePrep {}
+
+pub struct Parsa {
+	wb: Writeback,
+	bus: Bus<ParsaMessage<i32>>,
+	past_states: HashMap<i32, (i32, usize, Vec<i32>)>,
+	past_rewards: HashMap<i32, i32>,
+	policy: Policy<i32>,
+	setup: Setup<i32>,
+	task_ct: usize,
+	work_ct: usize,
+}
+
+impl Parsa {
+	pub fn new(setup: Setup<i32>, tiling: &TilingSet, work_ct: usize) -> Self {
+		let wb = new_writeback(setup.n_actions.into());
+		let mut bus = Bus::new(100);
+
+		let (mut tasks, task_ct, preprep) = preprep_state(&setup, &tiling, work_ct);
+
+		// create worker threads
+		for task_set in tasks.drain(..) {
+			let my_wb = wb.clone();
+			let my_bus = bus.add_rx();
+			let my_setup = setup.clone();
+			let my_preprep = preprep.clone();
+
+			std::thread::spawn(|| {
+				parsa_minion_loop(
+					my_wb,
+					my_bus,
+					task_set,
+					my_preprep.policy,
+					my_preprep.shift_amts,
+					my_setup,
+					my_preprep.tile_set_data,
+					my_preprep.adjusted_maxes,
+					my_preprep.widths,
+				);
+			});
+		}
+
+		Self {
+			wb,
+			bus,
+			past_states: HashMap::new(),
+			past_rewards: HashMap::new(),
+			policy: preprep.policy.clone(),
[…]

### 2021-09-23 code/rl-perf-tester/src/verify.rs

@@ -1,0 +1,96 @@
+use super::*;
+
+use control::{PolicyConfig, SendRewardConfig};
+use rand::Rng;
+
+pub fn verify_experiment(config: &VerifyConfig, if_name: &str) {
+	let mut global_base = GlobalConfig::new(Some(if_name));
+	let global = &mut global_base;
+
+	let mut proto = ProtoSetup::default();
+	proto.epsilon = Some(0.0);
+
+	let mut setup: Setup<i32> = proto.instantiate(true);
+	prime_setup_with_timings(&mut setup, &TimeBreakdown::UpdateAll);
+
+	let tiling = crate::generate_tiling(&setup, 28);
+	setup.limit_workers = None;
+
+	let transport = config.transport_cfg;
+
+	let mut state_rng = ChaChaRng::seed_from_u64(config.sample_seed);
+
+	let policy = if let Some(policy_seed) = config.policy_seed {
+		unimplemented!()
+	} else {
+		control::Policy::Quantised {
+			data: control::PolicyFormat::Dense(vec![0; control::policy_size(&setup, &tiling)]),
+		}
+	};
+
+	let shift = setup.quantiser_shift;
+
+	let mut setup_cfg = SetupConfig {
+		global,
+		transport,
+		setup: setup.clone(),
+	};
+
+	control::setup::<i32>(&mut setup_cfg);
+
+	control::tilings(&mut TilingsConfig {
+		global,
+		tiling: tiling.clone(),
+		transport,
+	});
+
+	// TODO: send policy packets
+	control::insert_policy::<i32>(&mut PolicyConfig {
+		global,
+		transport,
+
+		policy: Some(policy),
+		tiling: tiling.clone(),
+		setup: setup.clone(),
+	});
+
+	let mut host_impl = Parsa::new(setup.clone(), &tiling, 1);
+
+	for i in 0..config.n_samples {
+		// take state vector
+		let mut v = crate::generate_state(&setup, &mut state_rng);
+		v[0] = 15;
+
+		let reward_measure = state_rng.gen_range(-1.0..=1.0);
+
+		if i < config.first_sample {
+			continue;
+		}
+
+		// reward value handling
+		let q_reward = i32::from_float_with_quantiser(reward_measure, shift);
+		host_impl.reward(1, q_reward);
+		control::send_reward::<i32>(
+			&mut SendRewardConfig { global, transport },
+			1,
+			reward_measure,
+			shift.into(),
+		);
+
[…]

### 2021-10-08 paper/poster-abstract.tex
Poster abstract
Just want someone to see this.

--- 2021-06-10 paper/paper-10pt.tex
@@
 sigconf,natbib=false
-,anonymous=true
-,10pt
+%,anonymous=true
 ]{acmart}
 
-\usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists]{savetrees}
+%\usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists]{savetrees}
+
+%\usepackage{fnpos}
+%\usepackage{dblfnote}
+%\usepackage{ftnright}
 
@@
 
-\usepackage[maxnames=2,maxbibnames=99,mincrossrefs=99,sortcites,style=numeric-comp
+\usepackage[maxnames=1,maxbibnames=1,mincrossrefs=99,sortcites,style=numeric-comp
 %,backend=biber
@@
 % Copyright
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
+%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
 %\setcopyright{none}
@@
 %\settopmatter{printacmref=false, printccs=false, printfolios=true}
-%\settopmatter{printfolios=true}
-\settopmatter{printacmref=false, printccs=true, printfolios=true}
+\settopmatter{printfolios=true, printccs=true}
+%\settopmatter{printacmref=false, printccs=true, printfolios=true}
 
@@
 %Conference
-\acmConference[SOSR '21]{Symposium of SDx Research 2021}{September \numrange{20}{21}, 2021}{Online}
+\acmConference[CoNEXT '21]{the \nth{17} International COnference on emerging Networking EXperiments and Technologies}{December \numrange{7}{10}, 2021}{Online}
 \acmYear{2021}
@@
 \begin{document}
-	\title{Revisiting~the~Classics: Online~RL~in~the~Programmable~Dataplane}
+	\title{Poster: Online RL in the Programmable Dataplane with \approachshort{}}
 
@@
 \begin{abstract}
-Automatic, data-driven networking is becoming more capable and widely researched, mainly driven by the efficacy of \emph{Deep Reinforcement Learning} (DRL) algorithms.
-Yet the complexity of both DRL inference and learning force these tasks to be offloaded \emph{away from the dataplane} to hosts, harming latency-sensitive applications.
-Online learning of such policies cannot occur in the dataplane, despite being useful techniques when problems evolve or are hard to model.
+%Automatic, data-driven networking is becoming more capable and widely researched, partly driven by the efficacy of \emph{Deep Reinforcement Learning} (DRL) algorithms.
+%Yet the complexity of both DRL inference and learning force these tasks to be offloaded \emph{away from the dataplane} to hosts, harming latency-sensitive applications.
+%Online learning of such policies cannot occur in the dataplane, despite being useful techniques when problems evolve or are hard to model.
 
-We present \emph{\approachshort{}}---\approach{}---the first work to bring \emph{online reinforcement learning} to the dataplane.
-\approachshort{} makes online learning possible in constrained SmartNIC hardware by returning to classical RL techniques which simplify update logic and enable parallel processing.
-Compared to offloading, we achieve a \SI{21}{$\times$} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{$\times$} improvement in online throughput for real-world policy designs.
-In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
-\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, improving on offloading, and are quickly reconfigurable at runtime compared to reinstalling device firmware/designs.
+Reinforcement learning (RL) is a key tool in data-driven networking for learning to control systems online.
+While recent research has shown how to offload machine learning tasks to the dataplane (reducing processing latency), online learning remains an open challenge unless the model is moved back to a host CPU, harming latency-sensitive applications.
+Our poster introduces \emph{\approachshort{}}---\approach{}---the first work to bring \emph{online reinforcement learning} to the dataplane.
+\approachshort{} makes online learning possible in SmartNIC/NPU hardware by returning to classical RL techniques---avoiding neural networks.
+This simplifies update logic, enabling online learning, and benefits well from the parallelism common to SmartNICs.
+We show that our implementation on Netronome SmartNIC hardware offers concrete latency improvements over host execution.
+
+%Compared to offloading, we achieve a \SI{21}{$\times$} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{$\times$} improvement in online throughput for real-world policy designs.
+%In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
+%\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, improving on offloading, and are quickly reconfigurable at runtime compared to reinstalling device firmware.
 
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.65\linewidth]{figures/arch-with-p4}
-	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
+		\includegraphics[keepaspectratio, width=0.6875\linewidth]{figures/arch-with-p4}
+		\caption{\approachshort{} brings low-latency, online reinforcement learning to SoC- and NPU-based SmartNICs.
+Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
+\end{figure}
+\begin{figure}
+	\centering
+%	\includegraphics[keepaspectratio, width=0.6875\linewidth]{figures/arch-with-p4}
+%	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
 SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible alongside P4 dataplanes.
 Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
+	\centering
+	\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/coop}
+	\caption{Parallel policy execution.
+A single \emph{controller} delegates RL computation and updates to many \emph{minion} threads, using a shared atomic writeback.\label{fig:single-and-parallel:parallel}}
 \end{figure}
-Automatic network optimisation, control, and defence are at last becoming commonplace.
-Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing automatic traffic/media optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17,DBLP:conf/conext/GiladSGRS20}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
-In RL methods, every change and its effects improve future decisions made by an agent.
+Automatic network optimisation, control, and defence are at last becoming commonplace through data-driven techniques such as RL methods, where every change and its effects improve future decisions made by an agent.
+In tandem, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} (PDP) hardware have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
+This ecosystem exposes not only runtime reconfigurable packet processing, but per-packet and per-flow traffic measurement state that can greatly aid network operation~\parencite{DBLP:conf/sosr/GhasemiBR17}.
 
@@
 %?? in the face of new data, Able to start learning again in microseconds
+\fakepara{Machine learning in the dataplane}
+To process this fine-grained state both at line rate \emph{and} at low latency, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:journals/corr/abs-2009-02353}.
+All works in this category, while powerful, operate by converting a \emph{pre-trained} model into a suitable representation, such as a binary neural network or string of match-action tables.
+%These works have shown the value of in-network ML: high-throughput, low latency response to network changes.
+These can react to on-device state, yet the missing piece of the puzzle is learning and updating these ML analyses \emph{online}, without deferring to another machine in the network.
+%Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed by the community}.
+This lacuna has yet to be addressed by the community.
 
-In parallel, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} (PDP) hardware~\parencite{DBLP:journals/micro/ZilbermanACM14, netronome-smartnic, xilinx-alveo, barefoot-intel} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
-The promise of PDP hardware is that we can move the entire monitoring and analysis stack \emph{into the dataplane itself, and have it evolve to incorporate new approaches}.
-Consider a DDoS mitigation system which could react to dataplane-specific, per-packet state in microseconds, and \emph{start learning again just as quickly} in the face of new data or behaviours.
-The P4 ecosystem already presents novel, openly-available, fine-grained traffic measurement techniques that can be installed and controlled with ease~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
-In addition, P4's control plane makes it easy to select which flows or packets are monitored in a live network and potentially allow control over traffic at the decision site.
-As a result, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome,DBLP:journals/corr/abs-2002-08987} to take advantage of flow or per-packet state that cannot be efficiently processed or extracted at any other location in the network.
-These works have shown the value of in-network ML: high-throughput, low latency response to network changes.
-While they can exploit on-device state to provide reactive insight, the missing piece of the puzzle is learning and updating these ML analyses online without deferring to another machine in the network.
-Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed by the community}.
+\fakepara{Problems of online training}
+Programmable dataplane hardware, being designed solely for efficient packet processing, lacks floating-point arithmetic support even in the case of more general purpose NPU-type SmartNICs.
+Additionally, deep neural network training relies on backpropagation, can require memorizing many a sizeable replay buffer for RL, and needs many mini-batches of data for stable training.
+The natural solution is to pass training data to a host machine or dedicated accelerator, yet this adds delays in crossing the PCIe bus, kernel-user handover, and buffering---adding latency while learning.
+The above (NIC-suitable) inference schemes still, even if their data formats solve the issue of floating point use.
+Online learning thus requires careful choices in both sample data formats, and the learned function approximation scheme.
 
-It is important to make this feasible as offloading learning to commodity machines adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
-For context, \emph{Deep Neural Network} (DNN) training relies on backpropagation to compute gradients, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many mini-batches of data for stable training.
-These lead to high compute and storage costs, inducing capital expenditure for dedicated accelerators.
-Moreover, high latencies \emph{caused} by offloading and expensive inference harm the learning process~\parencite{DBLP:journals/firai/TravnikMSP18} and impact runtime application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
-If we can bring online learning \emph{to the dataplane}, then we can take advantage of rich, local state while minimising these latencies (and their impact on the learned policy).
-This would also make it easy to train and prototype agent designs which can learn \emph{as the environment evolves}, or when there is too little data to model and simulate a problem.
+\section{\approachshort{}'s design}
+\approachshort{} makes RL workloads feasible under on SmartNICs by using quantised fixed-point representations of action values, tile-coded policies, and one-step temporal-difference RL algorithms such as Sarsa~\parencite{RL2E}.
+These allow us to evaluate and update policies using \emph{only integer arithmetic}.
+We observe that tile coded inference is a map-reduce problem (\cref{fig:par-tilecode}), and over integers admits a novel \emph{wait-free Sarsa} RL implementation (exploiting the parallelism built in to SmartNICs).
+In addition to their runtime simplicity, while these functions have lower model capacity they do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays.
 
-%Moreover, DNN-backed RL must often be trained by vast amounts of offline simulation~\parencite{DBLP:journals/corr/abs-1912-06680}---in some cases compute-years worth~\parencite{DBLP:journals/corr/abs-1912-06680}---which can be counter-productive if the target problem evolves over time or is hard to model.
+\fakepara{Interacting with \approachshort{}}
+To prevent packet stalls but maintain access to fine-grained state, \approachshort{} places RL execution on-NIC, \emph{but off the main packet path}, parallel to the main dataplane (\cref{fig:netro-arch}).
+The P4 (or device-specific) dataplane pushes state vectors and reward measurements to \approachshort{}'s \inring{} ring, and pulls output actions when it is able to, imposing minimal impact on carried traffic for both bump-in-the-wire deployments and at end-points.
+Once an action is enqueued on its \outring{} ring, \approachshort{} updates its policy.
 
-%?? rich, local state
+\fakepara{Inference/Updates}
+State is mirrored to a set of packet processing threads, which compute hit tiles from a locally held work set.
+During inference, these threads atomically add values to a shared action preference list, while policy updates are guaranteed by tile-coding to proceed without data races (\cref{fig:single-and-parallel:parallel}).
+In the case of our Netronome implementation, these run on lightweight cores known as \emph{microengines} (MEs)---which could be substituted for suitable FUs on CGRA/FPGA hardware.
+Additionally, we have a single-threaded implementation, \indfw, which forgoes this cooperative model to maximise offline throughput (i.e., by having many parallel agents handle state).
 
-%?? Why important to make feasible?
-%?? Why horrid that it hasn't been looked at?
-%This induces capital expenditure on 
-%?? We focus on models \& methods to enable this
+\begin{figure}
+	\centering
+	\resizebox{0.6\linewidth}{!}{
+		\begin{tikzpicture}
+			\node at (0,0) {
+				\begin{tikzpicture}
+					\draw[step=0.5cm,color=uofgcobalt,opacity=0.7,shift={(0,0)},label=above:{Tiling 0}] (-0.5,-0.5) grid (1,1);
+					\fill[uofgcobalt,opacity=0.5] (0.5,-0.5) rectangle (1,0);
+					\node[color=uofgcobalt] (t1g) at (0,1.2) {\footnotesize Tiling 1};
 
-To enable \emph{online in-NIC learning}, we return to \emph{classical} RL methods.
-In particular, we focus on tile-coding with one-step temporal-difference algorithms such as Sarsa.
-%These choices have important benefits for in-NIC execution.
-These functions do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays, and have simple update and inference logic.
-Tile-coding in particular admits many optimisations, being an embarrassingly parallel problem.
-Using fixed-point arithmetic, we solve the lack of floating-point support in PDP hardware \emph{and} enable new optimisations.
-Moreover, the P4 dataplane can offer runtime control over which flows/packets are monitored.
-%Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair, safeguarding the limited memory of the target devices.
-We also design our solution to operate as closely as possible to the P4 pipeline to use and learn from per-packet state, but outside of the main packet path to prevent packet stalls.
+					\draw[step=0.5cm,color=uofgpumpkin,opacity=0.9,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
+					\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
+					\node[color=uofgpumpkin!50!uofgrust] (t2g) at (0.25,-0.95) {\footnotesize Tiling 2};
 
-%However, the task of \emph{training} these models online \emph{and} in-network presents an exciting (and challenging) lacuna in the field.
-%Online training is particularly important if an agent needs to react to unforeseen (failure) states, such as in DDoS/intrusion prevention.
-%For context, \emph{Deep Neural Network} (DNN) training relies upon the backpropagation algorithm to compute gradients, and for sizeable amounts of data to be held in mini-batches for stable training.
-%Moreover, DNN-backed RL must often be trained by vast amounts of offline simulation---in some cases compute-years worth~\parencite{DBLP:journals/corr/abs-1912-06680}---which can be counter-productive if the target problem evolves over time or is hard to model.
-%These space and compute requirements, at present, mandate that training must occur in commodity hosts or dedicated hardware.
-%Moreover, high state-action latencies \emph{caused} by offloading and expensive policy approximations add significant noise to the learning process~\parencite{DBLP:journals/firai/TravnikMSP18} and impact application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
+					\node[circle, black, draw,
+					fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
+					%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];
 
-%?? Why programmable vs fixed dataplane?
-%?? Entire monitoring and analysis stack in one device
-%?? Why p4? Exploit existing works to get local state (and future works), Ease of runtime reconfig, control over traffic at the decision site.
+					\draw[->] (-0.25,-0.5)--(-0.25,0.85);
+					\draw[->] (-0.25,-0.5)--(1.1,-0.5);
 
-%?? The chance to exploit
+					\node at (1,-0.7) {\footnotesize 1};
+					\node at (-0.4,0.75) {\footnotesize 1};
+					\node at (-0.35,-0.6) {\footnotesize 0};
+				\end{tikzpicture}
+			};
 
-%?? Emphasise related work claims here!
-%?? We're doing novel stuff because ...
+			\node (policy-head) at (2.5,1.2) {Policy};
+			\draw[color=uofgcobalt,opacity=0.7] (2,0) rectangle ++(2,1) node[pos=.5] (t1p) {Tiling 1};
+			\fill[uofgcobalt,opacity=0.25] (3.33,0) rectangle ++(0.67,0.33);
 
-%?? quick, reactive insight
+			\draw[color=uofgpumpkin,opacity=0.9] (2,-1) rectangle ++(2,1) node[pos=.5] (t2p) {Tiling 2};
+			\fill[uofgpumpkin,opacity=0.25] (2.67,-0.67) rectangle ++(0.67,0.33);
 
-%?? without the capex on expensive, powerful hosts
+			\draw (2,-2) rectangle ++(2,1) node[pos=.5] (tdot) {$\cdots$};
 
-The question we investigate is: can online RL be brought to the dataplane by returning to these computationally simpler methods, to act on locally extracted state?
-Can it be made more efficient \emph{by dataplane hardware}?
-Through this work, \approachshort{}, we can comfortably answer ``yes'' on both counts.
-In particular, we exploit how SmartNIC devices often expose general-purpose compute to provide path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
-As many of these devices have engineering and development histories which predate P4, general compute beyond P4's limits~\parencite{p4-psa} is surprisingly common.
-By executing on spare compute units, we prevent packet stalling and offer quick runtime reconfigurability.
-%We look backwards to classical RL techniques (tile-coded policies, single-step RL algorithms), alongside the adaptations introduced by existing in-network ML (such as quantisation).
-%Such adaptations unlock further improvements in parallel inference and update processing.
-This paper contributes:
-\begin{itemize}
-	\item An analysis of why in-NIC RL is needed and best-placed to interact with the network, made feasible by classical RL methods and quantisation (\cref{sec:motivation}),
-	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands of network traffic analysis (\cref{sec:design}),
-	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm to accelerate tile-coded policy inference and updates (\cref{alg:parsa}),
-	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
-	\item A description of how \approachshort{} would integrate with state-of-the-art PDP applications to perform fully in-NIC, fast, automated DDoS mitigation (\cref{sec:potential-integrations}).
-\end{itemize}
+			\draw [->,color=uofgcobalt, bend left] (t1g) to (t1p.west);
+			\draw [->,color=uofgpumpkin, bend right] (t2g) to (t2p.west);
+
+			\node (act-list) at (1,-2.5) {$\bm{a}=\left[ \cdots \right]$};
+
+			\draw [->, bend right] (t1p.west) to (act-list);
+			\draw [->, bend right] (t2p.west) to (act-list);
+			\draw [->, bend right] (tdot.west) to (act-list);
+
+		\end{tikzpicture}
+	}
+	\caption{Tile-coding: actions preferences are aggregated from \emph{disjoint} tile queries---a map-reduce problem.
+To update, gradients are simply the tiles activated during the forward pass with no aggregation.\label{fig:par-tilecode}}
+\end{figure}
+
+\section{Evaluation}
+We compare our implementation of \approachshort{} against a tile-coded RL agent written in numpy on a powerful host (i7-6700K@\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}), using policy dimensions used by existing DDoS prevention work\parencite{DBLP:journals/tnsm/SimpsonRP20}, across various quantisation bit depths.
+\Cref{fig:lat-cumul} shows that our main design (\coopfw) achieves order-of-magnitude improvements over our single-threaded variant (\indfw) and a host implementation, as well as considerably tighter tail behaviour on the SmartNIC.
+Note that lower bit depths have higher execution costs due to the native \qty{32}{\bit} width of Netronome registers.
+On update times, we see \qty{63.12}{\micro\second} at 99.99\nthscript{th} percentile for a \qty{32}{\bit} policy of this size.
+We relate a \qty{15}{$\times$} latency (\coopfw) and \qty{2.82}{$\times$} offline throughput-per-core (\indfw) in spite of the Netronome's considerably slower clock speed (\qty{0.29}{$\times$}), showing the value of exploiting SmartNIC parallelism.
+
+\begin{figure}
+	\centering
+	\begin{subfigure}{0.45\linewidth}
+		\resizebox{1.0\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/latency-cumul}
+			\vspace{-10pt}
+		}
+		\caption{\approachshort's \Coopfw{} design achieves consistent, tight latency bounds.}
+	\end{subfigure}
+	\hspace{0.05\linewidth}
+	\begin{subfigure}{0.45\linewidth}
+		\resizebox{1.0\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/latency-cumul-broad}
+		}
+		\caption{Tail latencies suffer in hosts---particularly when oversubscribed.}
+	\end{subfigure}
+	\caption{Cumulative state-action latency plots for \approachshort{} and host-based execution at different quantisation settings.\label{fig:lat-cumul}}
+\end{figure}
 
@@
 
-\section{Preliminaries: towards in-NIC RL}\label{sec:motivation}
-%\section{Preliminaries: RL for in-network computation}\label{sec:motivation}
-By combining recent trends in PDP hardware with insights from the ML/RL communities (past and present), we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL methods and quantisation make this computationally feasible.
-Crucially, we focus on low port-density devices such as SmartNICs and the NetFPGA.
-The designs of these devices make it reasonable to move processing \emph{outside of the packet pipeline}; SmartNICs expose additional programmable cores, and NetFPGAs allow for the synthesis of independent functional units.
-%The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
-The main benefit is that core control logic can be moved onto the NIC \emph{without impacting packet processing rates}.
-%?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
-As Ethernet speed increases, packet deadlines grow tighter; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second} at \SI{40}{\giga\bit\per\second}, giving a \SI{3.99}{\micro\second} deadline on Netronome (which has \num{312} P4 pipelines).
-In-pipeline execution thus increases the risk of packet drops or stalls.
-
-%However, this requires that there be a reasonably consistent interface and behaviour between device classes.
-%This need for consistency underpins the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}, which defines a conceptual model of match-action tables divided into ingress and egress pipelines.
-%The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.
-
-%?? TODO: trim down. Too ``thesis-y'', rather than ``conference-y''.
-%?? Save this elsewhere 'til then?
-
-\subsection{Programmable hardware capabilities}
-%The introduction of the P4 language~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to explosive growth in the research community surrounding in-network computation and offloading.
-%Providing a single language which is compatible with network devices of many form factors has been instrumental in the development of novel fine-grained traffic measurement approaches~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
-%However, this requires that there be a reasonably consistent interface and behaviour between device classes.
-%This need for consistency underpins the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}, which defines a conceptual model of match-action tables divided into ingress and egress pipelines.
-%The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.
-
-%We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
-%Many compatible devices have legacies long predating these developments. 
-%For instance, many-core SOC-based Netronome SmartNIC~\parencite{netronome-smartnic}, and NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14} NICs allow virtually arbitrary programs to be specified and executed.
-%The first of these compiles from P4$\rightarrow$Micro-C$\rightarrow$bytecode (with Micro-C externs), while the latter can combine the P4$\rightarrow$NetFPGA toolchain~\parencite{DBLP:conf/fpga/IbanezBMZ19} with arbitrary circuit externs---both exposing further device capabilities beyond the specification.
-%This pattern extends to other SmartNICs, such as NVIDIA BlueField~\parencite{nvidia-bluefield} and Xilinx Alveo~\parencite{xilinx-alveo} devices.
-
-While P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to great interest in network programmability, it requires similar behaviour between device classes.
-This is encoded by the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}---a lower bound on the device capabilities required to implement a P4 dataplane.
-%We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
-However, many compatible devices have legacies long predating these developments. 
-For instance, many-core SOC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
-%This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
-Currently, low port-density devices like these are most likely to have specific general-purpose compute, as they are designed for high-performance offloading and middlebox development.
-%Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
-Regardless, floating-point operations key to ML/RL workloads are very rarely supported outside of deep learning accelerators (which lack packet switching functionality).
-Moreover, the limited memory and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
-
-%\begin{insight}
-%	Current low-port density programmable network devices often have spare resources and extra capabilities beyond the PSA specification which can aid asynchronous, local compute.
-%\end{insight}
-
-\subsection{Reinforcement Learning}
-%\emph{Reinforcement learning} (RL) algorithms are methods of training an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task \cite{RL2E}.
-%Like most machine learning methods, an RL algorithm uses gradient information to update the parameters used to approximate a function.
-%In RL contexts, this is a state$\rightarrow$action function known as a \emph{policy}.
-%When training, agents are given \emph{reward measurements} and a learned policy acts to maximise the \emph{expected discounted reward} received.
-%When a pre-trained policy is deployed, this signal is not required.
-%
-%It's useful to consider how these algorithms differ from other ML use cases, such as classifiers.
-%The main differences lie in how this gradient information is used, and combined with \emph{reward measurements} received from the environment.
-%Rather than adapting the learned parameters along the gradient using an error value from a target value, because the optimal actions aren't known they adjust values using a \emph{Markov Decision Process} (MDP)---capturing state trajectories to adjust value based on past and future decisions.
-
-\emph{Reinforcement learning} (RL) algorithms train an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task~\parencite{RL2E}.
-Like most ML methods, RL algorithms use gradient information to update the parameters used to approximate a function with \emph{reward measurements} from the environment.
-RL's \emph{Markov Decision Process} (MDP) structure allows online learning of a state-action map in a model-free way, and can step through local minima when needed compared to ML.
-In networking, this allows for learning to respond to on-device state or to handle rapidly evolving problems.
-RL algorithms are generally agnostic to the policy approximation used, and classical single-step methods are simple---most of the compute work is occupied by the policy itself.
-The single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to update and learn a policy online.
-
-%Consider the single-step, semi-gradient Sarsa algorithm~\cite[pp. \numrange{217}{221}]{RL2E}:
-%\begin{subequations}
-%	\begin{gather}
-%		\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
-%		\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
-%	\end{gather}%
-%	\label{eqn:sg-sarsa}%
-%	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, $\acval{S}{A}{\wvec{}}$ denotes the \emph{value} of an action $A$ taken in state $S$ according to the policy $\wvec{}$, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
-$\gamma,\alpha\in[0,1]$ are the discount factor and learning rate (governing the degree of forward planning and policy stability, respectively).
-%\end{subequations}
-%In essence, at each timestep the policy parameters ($\wvec{}$) are increased along the gradient ($\nabla{\acvalblank}$) using a fixed learning rate ($\alpha$) and a computed adjustment ($\delta_t$).
-%This adjustment is equal to the difference between the chosen action $A$'s value in state $S$ and the reward received ($R_{t+1} - \acval{S_t}{A_t}{\wvec{}}$), \emph{plus} some part of the \emph{next} action's value ($\gamma\acval{S_{t+1}}{A_{t+1}}{\wvec{}}$).
-%To give some context on the design space here, other algorithms may employ separate state value approximations, use the entirety of an agent's trace, or be tailored to characteristics of the policy approximator (e.g., how neural networks benefit from batching).
-
-%\begin{insight}
-%	Classical, single-step RL algorithms are computationally simple and are agnostic to the policy approximation used.
-Furthermore, they require only additions and multiplication to update and learn a policy online.
-%\end{insight}
-
-\fakepara{RL in asynchronous environments}
-%There remains some degree of divergence between the theory and implementation of RL agents.
-%Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
-%The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
-%Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
-Consider \cref{fig:state-slip}: the RL formulation assumes that an agent receives new state at fixed intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
-This adds noise to the state-action mapping, which harms learning and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
-%Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
-Reducing these costs requires detailed agent-environment co-design.
-This principle has influenced real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, where asynchrony is necessary for high-speed applications.
-This often requires that state measurements are fused~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
-`Stopping the world' causes significant performance loss, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control.
-
-\begin{figure}
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\resizebox{0.8\linewidth}{!}{
-			\centering
-			\begin{tikzpicture}
-				\node[circle, draw] (state) {$S$};
-				\node[circle, draw] (state') at ($(state) + (0, -2)$) {$S'$};
-
-				\node (agent) at ($(2, 0) + (state)$) {Agent};
-				\node[below of=agent] (action) {Action $A$};
-
-				\node[right of=state'] {$+ R$};
-
-				\draw[->] (state) -- (state') node[midway, right] {$A$};
-
-				\draw[dotted, ->, bend left = 30] (state) -- (agent);
-				\draw[->] (agent) -- (action);
-				\draw[dotted, ->] (action) -- (state);
-		\end{tikzpicture}}
-		\caption{Theory: state measurement, action computation, and learning are zero-cost.}
-	\end{subfigure}
-	\hspace{0.05\linewidth}
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\resizebox{0.6\linewidth}{!}{
-			\centering
-			\begin{tikzpicture}
-				\node[circle, draw] (state) {$S$};
-				\node[circle, draw] (state') at ($(state) + (0, -1.5)$) {$S'$};
-				\node[circle, draw] (state'') at ($(state') + (0, -1.5)$) {$S''$};
-
-				\node (agent) at ($(2, 0) + (state)$) {Agent};
-				\node[below of=agent] (action) {Action $A$};
-
-				\node[right of=state''] {$+ R$};
-
-				\draw[->] (state) -- (state') node[midway, right] {$\varnothing$};
-				\draw[->] (state') -- (state'') node[midway, right] {$A$};
-
-				\draw[dotted, ->, bend left = 30] (state) -- (agent) node[midway, above] {$t_1$};
-				\draw[->] (agent) -- (action) node[midway, right] {$t_2$};
-				\draw[dotted, ->] (action) -- (state') node[midway, below] {$t_3$};
-		\end{tikzpicture}}
-		\caption{Reality: policy costs lead to \emph{state drift}---over a delay $t_1+t_2+t_3$, state $S$ becomes $S'$.}
-	\end{subfigure}
-	\caption{Asynchronous RL delays and state slippage.\label{fig:state-slip}}
-\end{figure}
-
-%?? Find some cites citing the relevance of this problem wrt. self-driving cars, robotics, etc.
-
-This cost is what drives us to \emph{bring online reinforcement learning to the dataplane}---referring again to \cref{fig:state-slip}, we would place P4-based state measurement ($t_1$), simple policy approximation ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
-In networks, actions are most likely to be installed on backbone switches, bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts---all of which make in-NIC training a perfect fit.
-As such, our goal is to collocate \emph{all} functions which comprise an RL agent on the same chip or device.
-Both programmable devices and the network environment make this more difficult, as we'll examine in the sequel.
-
-%?? Return to dimitris note: how is this important to final solution?
-
-%\begin{insight}
-%	Online control problems benefit from low-latency, local execution and training (i.e., in-NIC/switch).
-%\end{insight}
-
-\subsection{Efficient policy approximation}
-%Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
-%This can then make \emph{online} learning of the task an attractive prospect, using a single available stream of experience.
-%\emph{Deep neural networks} (DNNs), particularly when used as the basis for an RL agent's policy, require vast amounts of experience to converge on an accurate parameter set.
-%In many problem domains, this equates to training from compute-years worth of distributed offline simulations, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
-%Achieving stable learning requires sizeable batches of gradients to be computed, potentially using the entire experience replay from each simulation.
-%Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the computational cost of gradient computation via backpropagation remains too significant on embedded hardware.
-Complex \emph{Deep Neural Networks} (DNNs), particularly when used as an RL agent's policy, require vast amounts (possibly compute-years~\parencite{DBLP:journals/corr/abs-1912-06680}) of experience to converge on an accurate parameter set, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
-Although there has been a vested interest in efficiently \emph{executing} complex function approximators like DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the storage and computational costs of backpropagation remain too significant on embedded hardware.
-
-%?? Probably want some cites on the need for batching in NN methods, even though this is understood. DL book?
-
-We return to what classical RL methods can offer us; in particular, the simple linear function approximation given by \emph{tile coding} \cite[pp.\ \numrange{217}{221}]{RL2E}.
-A policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
-This allows any input-output mapping, such as converting packet header data into a preference list for queue priorities.
-A state vector produces a single `hit' in each tiling, all of which then correspond to a list of action preference lists.
-Inference (choosing an action) is simply summing over all such lists to obtain a final action preference list, selecting the maximum.
-Updating uses this same list of lists after the next decision is made, adjusting the value of the last selected action using a value $\delta_t$ computed through Sarsa.
-Crucially, this internal representation has no data dependencies between any tile calculations for an input.
-As a result, both action selection and policy updates are map-reduce problems.
-To select an action, we can produce a task for each tiling---retrieving the list of action values contributed by the tile activated by the input state.
-An aggregate step sums up all action preference lists, and the action with the highest value is then selected.
-Updating the policy (with a new $\delta_t$ value) has no aggregate step, and as tile indices map to disjoint memory regions no locks are required.
-%As we show in \cref{sec:design}, it is possible to make the inference aggregate step wait-free.
-
-For instance, a policy with $m$ sets of tiles (each having its own set of input dimensions), each containing $n$ overlapping tilings then creates $m \times n$ separate tasks.\footnote{Work items could be retrieving each individual action's value (giving $m \times n \times a$ tasks for $a$ actions).
-This requires tile coding to be repeated $a$ times per tile, which is likely extremely wasteful outside of heavy parallelism.}
-As an example, a real-world policy size which we examine later (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets of \num{8} tilings creates \num{129} work items.
-Computing a tile hit uses one division per dimension, which is an expensive ALU operation to require.
-However, if tile widths are fixed at powers-of-two, then these may be replaced with right bitshifts.
-%As we shall see shortly, this aligns with our choice of numerical representation.
-
-%\begin{insight}
-%	Tile coded policies are embarrassingly parallel in both inference \emph{and} learning.
-Moreover, in some instances the process can be altered to require only bitshifts, additions, and multiplication.
-%\end{insight}
-
-\subsection{Representations for embedded ML}
-As PDP hardware universally lacks floating-point arithmetic support, the task of bringing ML models to resource-constrained environments without these capabilities is well-studied.
-Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
-Lower bit depths reduce memory footprints, and improve throughput in e.g., \emph{bfloat16}~\parencite{bfloat16-blog,DBLP:journals/sigops/XieDMKVZT18}, \emph{hfp8}~\parencite{DBLP:conf/nips/SunCCWVSCZG19}, and other floating point formats~\parencite{DBLP:journals/corr/abs-2007-01530}.
-In many cases, accuracy losses are negligible.
-Such work goes further still towards integers~\parencite{tensorrt-8bit} or binarisation~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16}, sacrificing dynamic range for cheaper arithmetic.
-Similar techniques are used to make inference tractable on network hardware, i.e., to bring pre-trained neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
-
-To make RL workloads feasible under such constraints, we propose the use of quantised, fixed-point representations such as $Qm.n$ (i.e., $m$\si{bit} signed integers with an $n$\si{bit} fractional part), which allow us to evaluate and update policies using only integer arithmetic.
-This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation.
-We note that our later parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is possible only due to the use of $Qm.n$, as atomic operations on floating point numbers typically do not exist outside of graphics-processing hardware.
-
-%When combined with the above P4-driven techniques for in-network flow/device measurement, we at last have the mechanisms to collocate the key processes of an RL agent.
-%Moreover, the base P4 dataplane can be used to simplify parsing logic and offer runtime control over which flows/packets are monitored, alongside other packet actions.
-%This again fits our goals of integrating RL techniques directly within bump-in-the-wire installations and at end-hosts.
-
-%\begin{insight}
-%	Fixed point quantisation allows us to perform numerical calculations using only integer arithmetic.
-%	Existing literature to-date supports the effectiveness and use of these techniques on embedded ML deployments.
-%\end{insight}
-
-%\subsection{Netronome Platform Fundamentals}
-%As we implement this work on Netronome SmartNICs (thus the NFP chip architecture), it is necessary to explain its basics.
-%NFP chips are \emph{system-on-a-chip} (SoC) devices, and achieve scalable packet processing through sheer parallelism.
-%Aside from an ARM management chip and application specific functional units, most of the chip is composed of \emph{microengines} (MEs), grouped into \emph{islands} of 4 or 12 MEs.
-%All 12-ME islands are used by a default P4 pipeline.
-%Each ME has \numrange{4}{8} \emph{contexts} (hardware threads) which share a code store and large register file, offering zero-cost context switches triggered by I/O.
-%Contexts and MEs may send one another numbered signals, and MEs have a small \emph{next-neighbour} register file for passing values in one direction to the next ME on the same island.
-%MEs run a proprietary instruction set, compiled to via a \emph{(Micro-)C} compiler.
-%Beyond registers, the platform implements an explicit memory hierarchy scaling in size, location, and access cost as below:
-%$$\text{LMEM (ME)} < \text{CLS (Island)} < \text{CTM} < \text{IMEM (Chip)} < \text{EMEM}$$
-
-\section{Design and Implementation}\label{sec:design}
-%Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{} (\approach)}.
-We present our design for an in-NIC, task-independent, online RL system---\emph{\approachshort{} (\approach)}.
-At a high level, \approachshort{} uses auxiliary compute exposed by SmartNIC devices to offer low-latency online learning, scaling with on-chip resources at build time.
-As the allocation of cores/chip area is set ahead of time by a framework or system administrator, \approachshort{}(-\Coopfw) agents enumerate themselves at runtime, during initialisation.
-
-\begin{figure*}
-	\centering
-	\begin{subfigure}{0.47\linewidth}
-		\centering
-		\includegraphics[keepaspectratio, width=0.78\linewidth]{figures/ind}
-		\caption{\Indfw{} (offline throughput-optimal).
-\emph{Workers} independently pull commands from (and push actions to) the environment, locking policy access for updates.\label{fig:single-and-parallel:single}}
-	\end{subfigure}
-	\hspace{0.04\linewidth}
-	\begin{subfigure}{0.47\linewidth}
-		\centering
-		\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/coop}
-		\caption{\Coopfw{} (online-optimal).
-A single \emph{controller} delegates RL computation and updates to many \emph{minion} threads, who operate on independent subtasks.\label{fig:single-and-parallel:parallel}}
-	\end{subfigure}
-	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to either latency or throughput needs.\label{fig:single-and-parallel}}
-\end{figure*}
-
-\Cref{fig:netro-arch,fig:single-and-parallel} outline our design on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
-We explain relevant NFP architectural details later in \cref{sec:netronome-platform-fundamentals}.
-\approachshort{} communicates with the packet pipeline of a P4 dataplane via \texttt{extern} plugins using \inring{} (state, configuration) and \outring{} (action) messages (\cref{fig:netro-arch}).
-Internally, \approachshort{} either has all its cores act independently (\cref{fig:single-and-parallel:single}) or cooperate to solve each task (\cref{fig:single-and-parallel:parallel})---with different latency-throughput benefits (\cref{sec:action-and-update-computation}).
-We open-source our firmware and control programs for the benefit of the community.\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission.}}
-%Moving beyond the Netronome platform, we describe how our architecture may be adapted and improved on by bespoke hardware or FPGA-based deployment.
-
-\subsection{System Model}
-\approachshort{} is a general, task-independent framework for in-network, online training and execution of \emph{any RL agent design} using classical methods.
-\approachshort{} is agnostic to the meaning of state vectors it receives as inputs and the actions it produces, which are employed by other functional units or the dataplane.
-However, in-NIC/in-network execution specifically benefits packet-, flow-, and network-level learning, control, and optimisation tasks.
-
-\approachshort{} runs on one or more cores of a SmartNIC to convert fixed-point state measurements from the environment into a stream of actions using a stored policy.
-As an example, this might be to map flow state and performance measurements into a queue priority for future packets, or to compute and apply a rate limit to preserve quality of service.
-These dedicated cores are responsible for processing requests, computing actions, and updating the underlying policy in real time.
-Combined with reward measurements, this policy can then be updated or trained from scratch entirely on the NIC, acting as a fully online RL agent.
-An input state vector \emph{always} induces an action and, if desired, updates the policy using either an included reward or one retrieved from memory according to a (configurable) key placed alongside the state.
-This allows for simultaneous control and learning over independent systems by the same agent (i.e., optimising several flows with their own reward measures, such as DDoS mitigation in an AS where each next-hop AS might have their own `health' metric).
-
-%However, high-speed data networks impose inviolable per-packet deadlines.
-
-To protect traffic throughput and allow effective deployment in as many environments as possible, \approachshort{} places RL execution on-chip, \emph{but off the main packet path}, communicating and running parallel to the main P4 dataplane.
-As shown in \cref{fig:netro-arch}, this asynchrony allows coexistence with P4 programs, and imposes minimal impact on carried traffic for both bump-in-the-wire deployments and at end-points.
-For instance, in the default deployment of a P4 packet processing pipeline on Netronome NFP chips several cores go unused (as does spare area on an FPGA design), making this paradigm possible.
-
-%The main interaction model is that platform-specific IPC (message rings) is used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
-%These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
-%Both input and output can occur on any other core of the device, i.e., as part of P4 \texttt{extern} plugins or a dedicated flow state measurement subsystem, while the P4 control plane itself provides granular control over which flows are monitored or affected.
-%
-%Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match-action tables and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
-%Bit depth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
-
-\subsection{Action and Update Computation}\label{sec:action-and-update-computation}
-\approachshort{} applies the insights of \textcite{DBLP:journals/firai/TravnikMSP18} to minimise action latency; an action is computed, sent out into the environment, and only then is the underlying policy updated.
-Using one of the below strategies chosen at compile time, a state vector is tile coded, converted into action probabilities, and an action is chosen.
-This is then written out to the environment as in \cref{sec:agent-environment-communication}.
-If online learning is enabled, \approachshort{} then checks an internal hashmap for a previous state-action pair matching the current instruction source, and if found then the policy is updated.
-Updates are computed using \emph{single-step semi-gradient Sarsa}~\cite[pp. \numrange{217}{221}]{RL2E}, though modification to support other single-step methods would be trivial.
-The new state-action pair is then written into storage.
-\approachshort{} can be configured to automatically select values of the input state vector as keys for state and reward storage.
-Two firmware models govern how these tasks (action selection, policy updates) are carried out:
-\begin{description}
-	\item[\Indfw{} (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each works sequentially.
-Computing an action list requires a \emph{read lock} on the policy.
-If an update occurs, the core requests a \emph{write lock} before updating, limiting online throughput.
-\emph{Tile lists} are stored for update computation.
-	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel,alg:parsa})] Threads cooperate on processing state vectors, minimising latency.
-\emph{Minion} threads have a fixed list of work items, while a \emph{controller} thread sends compute/update commands before awaiting worker completion.
-Work items are disjoint, requiring no policy locks.
-\emph{State vectors} are stored for update computation.
-\end{description}
-Each offers a different point of optimisation; if updates are disabled, then the \indfw{} model can maximise throughput, while the \coopfw{} model is designed to minimise decision latency and needs no locks to update the policy (increasing \emph{online learning} throughput).
-These correspond to only executing a trained policy and actively (re-)training a policy, respectively.
-%ParSa is described here at a high level, as an algorithm suited for \emph{any multiprocessor environment}.
-%We omit the simple logic for $\epsilon$-greedy action selection (which we implement), and note that modification to other single-step algorithms such as \emph{Q-learning} would be trivial.
-%We detail our communication primitives in \cref{sec:intra-agent-communication}, and our work scheduling strategy in \cref{sec:work-allocation}, eliding the details of tile-coding as they are well-understood.
-%In general it is also possible for the \emph{Ctl} task to act as an additional \emph{Minion} in its parallel sections; we were limited here by code store requirements on the NFP.
-Latency and throughput have different effects on RL agents according to their design and target problem.
-Higher RL throughput is a necessity for per-flow or per-packet applications, which can require high decision-per-second rates even after combining state measurements received from the environment, such as flow control in DDoS prevention.
-Equally, lower latency affords an agent finer-grained control and learning of a problem, being able to react sooner to new information (e.g., device state in a routing optimisation problem).
-
-Paradoxically, we found that in the general \emph{ParSa} algorithm it was more efficient \emph{to do more work} by having each worker recompute its tile subset from a stored state.
-It transpired that cacheing this data placed a larger \texttt{memcpy} in the serial section, whose size did not scale at all with bit depth.
-Additionally, we do not use bitshifts in place of division operations in our implementation, due to the strict limits that power-of-two tile widths place upon policy design.
-
-\subsection{Agent-Environment Communication}\label{sec:agent-environment-communication}
-\approachshort{} uses \emph{multiple-producer/multiple-consumer} (MPMC) channels to communicate with other elements; be they P4 programs on the packet path, or other on-chip analysis and control modules.
-Through these, a system \emph{pushes} state vectors, reward measures, and setup packets as inputs, and \emph{pulls} a stream of state-action pairs as outputs.
-This allows decisions to be made asynchronously, preventing packet stalling.
-The key insight of this mechanism is that on-chip reward/state signals enjoy first-class support in the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-NIC/chip asynchronous processes or the controller.
-As such, \approachshort{} can receive input from P4 \texttt{extern}s or other, dedicated off-path flow state measurement applications.
-
-Our implementation uses platform-specific IPC (\emph{EMEM ring buffers}) with hardware signalling on work arrival to achieve this.
-As PDP hardware lacks dynamic memory allocation, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state/configuration/policy data handover.
-Packet headers are extracted and parsed using the tooling autogenerated by the P4 pipeline.
-We found that this costs a median \SIrange{126}{140}{\nano\second} communication time (local--remote), comparable to message channels in the Rust and Go languages on commodity hardware.
-
-%The main interaction model is that platform-specific IPC (message rings) is used to \emph{push} configuration, state vectors, and reward measurements to the RL system.
-%These same mechanisms are used by other cores on the same device to \emph{pull} output actions from the RL system.
-%Both input and output can occur on any other core of the device, i.e., as part of P4 \texttt{extern} plugins or a dedicated flow state measurement subsystem, while the P4 control plane itself provides granular control over which flows are monitored or affected.
-
-\subsection{Intra-Agent Communication}\label{sec:intra-agent-communication}
-Optimising for latency requires meticulous care in how work is passed out and aggregated.
-This is truer still when moving from the moderately fine-grained control of classical RL ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
-Marshalling of requests and shared data access incur significant overheads, but on-chip execution and the ParSa algorithm allow us to sidestep these via lockless atomic aggregation.
-Moreover, adjacent cores often have special-purpose shared registers or share a small fast cache to accelerate communication.
-
-%?? reduce below and make take-homes more generic if possible
-
-Our implementation exploits the locality of cores in the NFP, which can be factored into designs on other platforms like NetFPGA as additional off-path functional units, as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-Policy compute/update/configure tasks are passed between cores using direct \emph{next neighbour} registers, signalling all child threads in response.
-Such on-chip signals cost just $\sim$\SI{20}{\nano\second} per relayed message.
-Each core performs atomic adds to a shared preference list and an acknowledgement counter checked by the master thread, implementing our wait-free \emph{ParSa} algorithm (\cref{alg:parsa}).
-This is essential for aggregation compared to the use of bounded message buffers, which caused significant head-of-line blocking.
-%For instance, additional off-path functional units can serve as workers, like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-%Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the P4$\rightarrow$NetFPGA~\parencite{DBLP:conf/fpga/IbanezBMZ19} actions that interface with it.
-
-%On-chip local messages cost us just $\sim$\SI{20}{\nano\second} latency per relayed message
-%
-%In earlier designs we had experimented with bounded buffers as in \cref{sec:agent-environment-communication} for this, modified to be located solely in CLS memory, dedicating the master thread to result aggregation.
-%We found that this created a performance bottleneck at this final stage, causing significant head-of-line blocking for each of the workers.
-%Similarly, our next neighbour work notification scheme ($\sim$\SI{20}{\nano\second} latency per relayed message) was examined against \emph{reflector} and \emph{work queue} IPC mechanisms (\SIlist{58;126}{\nano\second} per messaged core).
-%These slower messages can be used in theory to skip ahead into longer ME chains.
-
-%Our implementation exploits the locality of threads, cores, and their parent islands in the NFP architecture.
-%Policy compute/update tasks and configuration updates are passed between cores using these direct \emph{next neighbour} registers, signalling all child threads in response.
-%Each task performs atomic adds to a shared preference list, and atomically increments an acknowledgement counter to be periodically checked by the master thread, implementing the wait-free \emph{ParSa} algorithm we introduce (\cref{alg:parsa}).
-
-%In earlier designs we had experimented with bounded buffers as in \cref{sec:agent-environment-communication} for this, modified to be located solely in CLS memory, dedicating the master thread to result aggregation.
-%We found that this created a performance bottleneck at this final stage, causing significant head-of-line blocking for each of the workers.
-%Similarly, our next neighbour work notification scheme ($\sim$\SI{20}{\nano\second} latency per relayed message) was examined against \emph{reflector} and \emph{work queue} IPC mechanisms (\SIlist{58;126}{\nano\second} per messaged core).
-%These slower messages can be used in theory to skip ahead into longer ME chains.
-
-\begin{algorithm}
-	\small
-	\caption{ParSa---\emph{Par}allel \emph{Sa}rsa\label{alg:parsa}}
-	\SetKw{Let}{let}
-	\SetKw{Enum}{enum}
-	\SetKw{In}{in}
-	\SetKw{Await}{await}
-	\SetKw{Const}{const}
-	\SetKwProg{parsa}{Function \emph{ParSa}}{}{end}
-	\SetKwProg{control}{Function \emph{Ctl}}{}{end}
-	\SetKwProg{minion}{Function \emph{Minion}}{}{end}
-	\SetKwProg{tilecode}{Function \emph{TileCode}}{}{end}
-
-	\tcc{Given message passing mechanisms \emph{scatter} and \emph{recv}, quantised arithmetic functions $Q_\mathit{mul}$ and TileCode, and omitting schedule/config/precache updates.}
-	\tcc{\emph{cfg}.$\alpha$, \emph{cfg}.$\gamma$ are hyperparameters affecting the significance of each update and the degree of forward-planning, respectively.}
-
-	\Enum Par \{ Act(\emph{state}), Upd(\emph{delta, action, state}) \}\;
-	\Const \emph{cfg, policy} = /* ... */\;
-
-	\Let \emph{values}: [AtomicI32; \emph{cfg.n\_actions}] = \{0\}\;
-	\Let \emph{acks}: AtomicI32 = 0\;
-	\parsa{id, schedule}{
-		\uIf{id==0}{
-			\ForAll{state\_pkt \In \inring{}}{
-				Ctl(\emph{state\_pkt})\;
-			}
-		}
-		\Else{
-			\While{true}{Minion(\emph{schedule}[$\mathit{id}- 1$], recv())\;}
-		}
-	}
-
-	\control{state}{
-		\emph{values, acks} = \{0\}, scatter(Par::Act(\emph{state}))\;
-		acquire slot for \outring, copy \emph{state} into slot\;
-		\Await acks == \emph{cfg.n\_minions}\;
-		\Let \emph{action} = argmax(\emph{values})\;
-		write \emph{action} into \outring{} slot, enqueue\;
-		\If{cfg.online}{
-			\Let \emph{((l\_state, l\_act, l\_val), found\_s)} = \emph{cfg}.lookup\_state\_from\_key(\emph{state})\;
-			\Let \emph{(reward, found\_r)} = \emph{cfg}.lookup\_reward\_from\_key(\emph{state})\;
-			\If{found\_s \&\& found\_r}{
-				\Let $\delta_t$ = $\mathit{reward} + Q_\mathit{mul}$(\emph{cfg}.$\gamma$, \emph{values}[\emph{action}]) $-$ \emph{l\_val}\;
-				$\delta_t$ = $Q_\mathit{mul}$(\emph{cfg}.$\alpha$, $\delta_t$)\;
-				\emph{acks} = 0, scatter(Par::Upd($\delta_t$, \emph{l\_act}, \emph{l\_state}))\;
-				\Await acks == \emph{cfg.n\_minions}\;
-			}
-			\emph{cfg}.store\_state(\emph{state}, \emph{action}, \emph{values}[\emph{action}])\;
-		}
-	}
-
-	\minion{tasks, msg}{
-		\Switch{msg}{
-		\uCase{Par::Act(\emph{s})}{
-			\ForAll{task \In tasks}{
-				\Let \emph{hit} = TileCode(\emph{s}, \emph{task})\;
-				\For{i \In [0..cfg.n\_actions)}{
-					\emph{values}[\emph{i}].atomic\_add(\emph{policy}[\emph{hit}][\emph{i}])\;
-				}
-			}
-		}
-		\uCase{Par::Upd($\delta$, \emph{a}, \emph{s})}{
-			\ForAll{task \In tasks}{
-				\Let \emph{hit} = TileCode(\emph{s}, \emph{task})\;
-				\emph{policy}[\emph{hit}][\emph{a}] += $\delta$\;
-			}
-		}
-		}
-		\emph{acks}.atomic\_add(1)\;
-	}
-
-%	\tilecode{state, task}{
-%		\Let \emph{hit} = \emph{cfg}.get\_task(task)\;
-%	}
-\end{algorithm}
-
-\subsection{Reconfigurability}
-\approachshort{} allows policy design and learning parameters to be changed at runtime using at most two control packets.
-For instance, design changes are useful at the end of learning (moving from online to offline), or when trying to train a new policy for another problem from the same vantage point.
-Parameter changes are useful when an online agent must become more (or less) adaptive to new data (i.e., after detecting a changepoint in traffic).
-This extends to policy data, which may be imported from a pre-trained model and exported via PCIe to the host machine.
-Some aspects must be chosen at compile time; bit depth, \Coopfw/\Indfw, and maximum policy/tiling/state sizes---these govern core operation or allocated memory.
-Choosing a bit depth of \qtylist[list-pair-separator = { or }]{16;8}{\bit} halves/quarters policy memory costs, allowing more complex problems to be modelled using more dimensions or fine-grained tiles.
-
-In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved DSCP value~\parencite{rfc2474} as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-While this mainly automates parser generation, it also allows for configuration to be received from only trusted hosts (over the dataplane if needed) via P4 rules.
-Our control packet generation library and evaluation frameworks which build upon it are written in Rust.
-
-%Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match-action tables and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
-%Bit depth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.
-
-\subsection{Work Allocation}\label{sec:work-allocation}
-Due to the lack of dynamic memory allocation on PDP hardware, and to simplify value lookups, policies cannot be stored sparsely.
-Tiling space requirements then scale exponentially with dimension count, so higher-dimension tilings must be placed in larger, slower memory regions.
-As a result, in many architectures policy parameters are split across such regions, giving different access and compute costs to different tasks.
-
-We use a simple first-fit work placement algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core (as the NFP uses hardware threads).
-Each work item is a separate \emph{tiling} over a list of dimensions.
-%, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
-The cost of any work item (from its dimension count and memory location) was empirically measured offline, and we weigh the total cost per core based on the number of minion threads available.
-This weighting specifically accounts for the controller thread on the first core.
-Work allocations are recomputed when policy configuration is installed or changed.
-Naturally, for $n$ tilings and $m$ threads this procedure is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per tiling, storing $m/8$ and $\le8$ costs respectively.
-
-%\subsection{Variable Quantisation Bit Depth}
-%At compile-time, \approachshort{} can be configured to use \SIlist[list-final-separator = { or }]{32;16;8}{\bit} values in input states and its internal policies.
-%Naturally, smaller bit depths reduce the storage required for policies, tiling data, and stored state-action pairs---allowing more complex problems to be modelled using more dimensions or fine-grained policies.
-%Through the lens of computational efficiency, reduced bit depth should allows action preference lists to be read using fewer I/O operations (as more values may fit into a single machine word).
-%
-%We had investigated bit-stuffing several such values into a single word for our atomic writeback mechanism (as the platform offers both \SIlist{32;64
-%}{\bit} atomic addition).
-%This is analogous to SIMD---through clever use of padding bits---but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
-%We investigate further performance implications of changing bit depth in the sequel.
-
-%Potential: low-latency train for some specific points, and high throughput for others? Think on the exact note I took...
-%
-%``can have accel'd offline, train online subset w/ my approach''
-%
-%THis might mean "do 32-bit online", then downsample to 8-bit policy for high-throughput mode.
-%
-%Can changes in trained policy be used to transfer to a more complex function approximator elsewhere?
-
-\subsection{Limitations}\label{sec:limitations}
-Direct rule installation into P4 tables from the SmartNIC is not generally possible.
-To achieve line-rate performance, platforms like NFP use accelerated datastructures (e.g., DCFL~\parencite{DBLP:conf/infocom/TaylorT05}) which must be computed over the \emph{entire rule set}.
-Even through \texttt{extern}s, directly adding new rules is neither feasible nor safe.
-We instead suggest on NFP chips that \texttt{extern}s or datapath stages which apply RL actions to packets should maintain a small store of state-action pairs, and periodically send these back to the controller for batch installation.
-%This ensures that the majority of installed rules exhibit accelerated performance, while preventing installation delay on the newest decisions.
-%Platforms such as Intel Tofino greatly simplify this, with Tofino Native Architecture intrinsics such as \emph{Action Profiles/Selects} allowing a P4 action to be chosen based on a register value (i.e., an RL action).
-
-%As is expected of parallel algorithms, efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
-%Efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
-Parallelisation adds per-task overheads, which require a minimum number of workers to improve on a serial approach.
-We investigate the exact worker count requirements imposed to overcome single-threaded execution in \cref{sec:results}.
-
-%\subsection{Targeting other device classes}
-%While this design caters to Netronome devices, many aspects have clear analogues in other SmartNICs, such as the NetFPGA SUME.
-%For instance, additional off-path functional units can serve as workers, like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-%Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the P4$\rightarrow$NetFPGA~\parencite{DBLP:conf/fpga/IbanezBMZ19} actions that interface with it.
-%%Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
-%As we show later, using a bit depth below the native word size reduces overall efficiency.
-%%This reduces overall efficiency even though considerably fewer reads are made.
-%We expect an FPGA design would allow this to match the desired bit depth and enable the SIMD-like optimisations we discuss without creative and costly bit-stuffing.
-%
-%Bringing \approachshort{} to PDP switches such as the Tofino is difficult as they closely match the P4-PSA, leaving no spare general-purpose compute units.
-%Taking inspiration from real-time programming, a potential solution is to divide RL processing across several packets (i.e., computing a portion of the preference list each time) until further work would delay outbound transmission.
-%This would introduce new issues in concurrent access, work splitting, and altered timescales for learning: we leave their treatment to future work.
-
-%While specifics of this design cater to Netronome devices, many aspects have clear analogues in NICs of similar form-factor, such as the NetFPGA SUME platform.
-%For instance, other cores can be introduced by designing additional off-path functional units,  like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-%Rather than general-purpose IPC, \approachshort{} would be hard-wired between internal workers and to the custom actions that interface with it, using the P4$\rightarrow$NetFPGA toolchain to offer granular control over packet and flow selection.
-%%Further optimisations arise from this model: a NetFPGA design would be able to replicate and optimise further on this concept of dedicated, low-latency communication between functional units.
-%As our later results show, using a bit depth below the native word size forces the compiler to emit excess code to extract and re-align preference values, reducing overall efficiency.
-%%This reduces overall efficiency even though considerably fewer reads are made.
-%We expect that a more bespoke hardware/FPGA design would allow this to match the desired bit depth and enable SIMD-like optimisations we discuss without creative and costly bit-stuffing.
-%
-%Bringing \approachshort{} to high-port density devices such as Tofino-powered switches is more difficult.
-%As discussed in \cref{sec:motivation}, Tofino ASICs map closely to the P4 PSA, meaning that there are no spare asynchronous general-purpose compute units to place this work on.
-%Taking inspiration from real-time programming techniques, a potential solution is to divide RL actions across several received packets (i.e., iteratively computing a portion of the action preference list each time) until any further work would delay outbound transmission.
-%This, however, would introduce new issues surrounding concurrent accesses, work splitting, and altered timescales for learning: we leave their treatment and examination to future work.
-
-\section{Evaluation}\label{sec:evaluation}
-We investigate the performance of \approachshort{} compared to classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{$\times$} speedup in median--\num{99.99}\nthscript{th} state-action latency and \SI{9.9}{$\times$} greater online learning throughput.
-Crucially, in-NIC execution offers tight tail latency bounds compared to host-based approaches.
-We report on how \approachshort{} scales as additional device resources are added, noting that both in-NIC designs outperform commodity hosts using just one core in latency and online throughput.
-Furthermore, \Indfw{} provides higher per-core offline throughput than host-based approaches, even though our measured hosts exhibit higher clock speeds.
-Finally, we show that \approachshort{} has minimal impact on dataplane cross-traffic carried by its parent device.
-
-\subsection{Netronome Platform Fundamentals}\label{sec:netronome-platform-fundamentals}
-Netronome NFPs achieve scalable packet processing through parallelism.
-The chip is composed of \emph{microengines} (MEs), grouped into \emph{islands} of 4 or 12 MEs.
-%All 12-ME islands are used by a default P4 pipeline.
-Each ME has \numrange{4}{8} \emph{contexts} (hardware threads) which share a code store.
-%Contexts and MEs may send one another numbered signals, and MEs have a small \emph{next-neighbour} register file for passing values in one direction to the next ME on the same island.
-%MEs run a proprietary instruction set, compiled to via a \emph{(Micro-)C} compiler.
-Beyond registers, the platform implements an explicit memory hierarchy scaling in size and access cost:
-$\text{LMEM (ME)} < \text{CLS (Island)} < \text{CTM} < \text{IMEM (Chip)} < \text{EMEM}$.
-
-\subsection{Experimental Setup}\label{sec:experimental-setup}
-Testing machines were as follows, with \SI{32}{\gibi\byte} RAM:
-\begin{description}
-	\item[\emph{MidServer}] Intel Xeon Bronze 3204 (\qtyproduct[product-units=single]{6 x 1.9}{\giga\hertz}),
-	\item[\emph{HighServer}] Intel Xeon Silver 4208 (\qtyproduct[product-units=single]{8 x 2.1}{\giga\hertz}),
-	\item[\emph{Collector}] Intel Core i7-6700K (\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}).
-\end{description}
-\approachshort{} was evaluated on server blades (\emph{Mid/HighServer}), each hosting a single Netronome Agilio LX \numproduct{1 x 40}GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
-These servers ran Ubuntu 18.04.5 LTS (4.15.0-140-generic).
-We additionally use a more powerful consumer-grade machine (\emph{Collector}) for estimating host performance when offloaded to a network function, running Ubuntu 18.04.4 LTS (4.15.0-96-generic).
-Control programs were built using rustc 1.52.1.
-We run \approachshort{} on a \num{4}-ME island of the NFP-6480 (\num{32} contexts) using \SIlist{32;16;8}{\bit} quantised arithmetic.
-This is the largest cluster of cores which is not in use by a P4 pipeline.
-All \approachshort{} timing measurements were repeated over \num{10000} state packets (preceded by \num{1000} warmup packets), retrieving item processing times over PCIe via the controller machine, from which throughput was derived.
-Host throughput and latency measures were observed over \num{10} trials of \SI{10}{\second} (with \SI{5}{\second} warmup/cooldown times).
-We differ this from the NFP as hosts need to run numpy-based agents in parallel via separate processes; this also allows us to investigate the effects of oversubscription on host latencies.
-
-Policy sizes are set to those of a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dim state vectors, a bias tile and 16 full tiling sets (\numproduct{7x1}-dim, \numproduct{8x2}-dim, \numproduct{1x4}-dim), 8 tilings per set, 6 tiles per dimension, and 10 actions.
-For context, such input would contain per-flow state (e.g., IATs, rates) which is combined with the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim).
-In \Coopfw{}, this creates \num{129} work items across \num{31} workers.
-We choose a larger action than this work to investigate the performance of more complex agents.
-
-\subsection{Experiments}
-
-\fakepara{Raw inference and learning performance}
-We compare how long it takes for \approachshort{} to compute actions and policy updates, and report on the throughput of our designs against floating point (numpy-based) implementations of Sarsa on commodity hosts.
-This allows us to demonstrate the performance differences between the \indfw{} and \coopfw{} configurations, particularly in how \indfw's (and hosts') required policy locks impact throughput.
-%We also report on the derived throughput measures accounting for the locking behaviour of each.
-We compare online learning performance with offline in these cases.
-Online performance marks the number of decisions that can be made per second (and associated latency) when training a policy.
-Offline performance is crucial for pushing a trained, known-good policy to agents in the network with an expected higher raw decision throughput.
-State-action latency is a shared property of both cases, with the main impact on throughput arising from the update step.
-
-We then vary the amount of worker threads to show how \approachshort{} scales to fit available compute resources on a device.
-This is an important aspect for (automated) allocation of compute resources in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of-time work scheduling which we examine later.
-This also demonstrates the number of cores needed to achieve a given latency/throughput bound on a policy of representative complexity.
-Moreover, to demonstrate how these costs vary as policy complexity increases, we vary the number of total dimensions included in the tiling.
-
-\fakepara{Work allocation}
-%?? Also our simple, ahead-of-time, work scheduler.
-We verify that our heuristic, runtime work scheduler (termed \emph{Balanced}) makes meaningful use of the explicit memory hierarchy and cost of each work item.
-We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ tilings given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
-%We use maximum-size, \SI{32}{\bit} policies as described above.
-
-\fakepara{End-to-end RL latency}
-We compare the key RL latencies we discuss in \cref{fig:state-slip} across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to a \emph{virtual Network Function} (vNF).
-
-\fakepara{Co-existence with the dataplane}
-While varying the rate of RL updates performed by \approachshort{}-\Coopfw{} (\SI{32}{\bit}) from \numrange{0}{16000} actions/s, we measure packet loss and sample latencies of cross traffic carried over a co-hosted P4 pipeline.
-This allows us to quantify whether on-chip (out-of-path) execution impacts ordinary dataplane behaviour through indirect means: e.g., EMEM cache evictions or hidden resource contention.
-%In particular, we stress test both decisions and policy updates of our \emph{Parallel} design at full throughput.
-
-We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, placing an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source via the default firmware.
-We perform throughput/loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements (sampling at \SI{2000}{\hertz} for \qtyproduct[product-units=single]{10 x 10}{\second}).
-This provides maximum throughput in the former case (relying solely on NIC counters for loss counting).
-In the latter case, this minimises host resource contention to observe exact latency measurements, have a high enough sample count to detect subtle (aggregate) latency effects, and eliminate \emph{host} receive drops.
-DPDK was setup using \qtyproduct[product-units=single]{4 x 1}{\gibi\byte} hugepages.
-Sent traffic was comprised of fixed-size \SIrange{64}{1518}{\byte} packets~\cite{rfc2544}.
-CPU clock scaling was disabled on \emph{HighServer}.
-
-\fakepara{Resource requirements}
-Using the maximum policy size defined above, we investigate how the memory requirements imposed by \approachshort{} vary with the number of dedicated MEs, over and above a base P4 forwarding plane.
-We report resource use for \SI{32}{\bit} \Indfw{} and \Coopfw{} agents, with hash-tables sized to \num{4096} state-action pairs and \num{16} separate reward values.
-This captures the relative cardinality of network RL traces to rewards, i.e., many input flows will map to one or few reward values (i.e., DDoS attack size estimation per egress-AS, queue occupancy in the case of AQM per output port).
-
-%\fakepara{The impact of bit depth}
-%We further investigate how bit depth affects the accuracy of policy by varying the number of bits used to represent the mantissa of action values and input state.
-
-\fakepara{Deployability}
-By timing agent setup and compile times, we describe the runtime costs needed for an administrator to repurpose an installed agent in a live network.
-
-\subsection{Results and Discussion}\label{sec:results}
-\fakepara{Raw inference and learning performance}
-\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy.\footnote{For brevity, we omit integer numpy results---against a float implementation, median action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}
-We show that \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse using 4 MEs of the NFP-6480 (\SIlist{15;21}{$\times$} speedups over a \emph{Collector} host).
-Importantly, \Indfw{} achieves lower median state-action latencies (\SI{2.79}{$\times$}) \emph{and} update times (\SI{2.63}{$\times$}) than a dedicated \emph{Collector} host \emph{while requiring only a single core or dedicated functional unit}.
-Crucially, in-NIC execution gives far tighter bounds on tail latency compared to host offloading.
-\num{99.99}\nthscript{th} percentile state-action latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, while host-based tail latencies are at least \SI{40.53}{\percent} greater even when the parallel worker count is at or below the physical core count.
-We show the cumulative distributions of these in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF/traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
-
-%We found that a key aspect of in-NIC execution is that it allows far tighter bounds on tail latency compared to host offloading.
-%Examining the state-action latencies in \cref{tab:lats}, we see that \num{99.99}\nthscript{th} percentile latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, respectively.
-%Similar results were observed for other bit depths.
-%By contrast, host-based tail latencies are at least \SI{40.53}{\percent} greater even when the parallel worker count is at or below the physical core count.
-%We show the cumulative distributions of these in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF/traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
-
-\newlength{\resultplotwidth}
-\setlength{\resultplotwidth}{0.8\linewidth}
-
-\begin{table}
-	\caption{Latencies and computation times for \approachshort{} versus commodity hardware hosts.
-On-device execution is crucial in not only lowering latencies, but in reducing tail latencies.
-Lower is better, with the best marked \emph{in bold}.\label{tab:lats}}
-	\resizebox{\linewidth}{!}{
-		\expandableinput ../tables/build/latency-only32
-	}
-\end{table}
-
-\begin{table}
-	\caption{Action and update throughputs for \approachshort{} versus commodity hardware hosts.
-Most designs cannot scale online performance with additional cores.
-Higher is better, with the best marked \emph{in bold}.\label{tab:tputs}}
-	\resizebox{\linewidth}{!}{
-		\expandableinput ../tables/build/tput-only32
-	}
-\end{table}
-
-\Cref{tab:tputs} compares \approachshort{}'s throughput against host-based execution.
-We set the worker count on host machines which maximised their throughput.
-This equalled the amount of physical cores on each device---moving beyond this would hamper tail latencies by an order of magnitude.
-To make the comparison fair in the context of many-core CPU environments, we include per-core throughput.
-\Indfw{} achieves \SI{2.82}{$\times$} higher offline throughput per core than commodity \emph{Collector} hardware in spite of the NFP-6480 having a considerably slower clock speed (\SI{0.29}{$\times$}).
-When compounded with the abundance of such weaker chips, in-NIC RL is able to deliver much higher throughput.
-%As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} that of a dedicated collector machine, due to the locking requirements mandating coordinated division of work.
-As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} that of a dedicated collector machine, as the write lock around policy updates creates a bottleneck.
-
-By limiting the available workers in software, we show how \Coopfw{}'s policy update time (thus  online throughput---\cref{fig:vary-core}) scales with available cores.
-While \Coopfw{} always outperforms the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency (plot omitted).
-Some artefacts of our environment and design choices are visible, such as the addition of new physical cores being more significant than contexts, and the presence of some schedule bottlenecks.
-Most importantly however, \Coopfw{}'s resource demand is tunable at compile time to meet the online training rate and/or action latency required by a task/environment.
-
-\begin{figure}[t]
-	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-core}
-	}
-	\caption{
-		\Coopfw{}'s online learning performance improves with additional cores, on max size tasks.
-This requires \num{8} workers to offer greater online throughput than single-threaded in-NIC RL.
-Sharper performance increases occur when a new physical core is added (\numrange{7}{8}) or the scheduler works around a bottleneck (\numrange{13}{14}).\label{fig:vary-core}}
-\end{figure}
-
-%\begin{figure}
-%	\resizebox{\resultplotwidth}{!}{
-%		\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
-%	}
-%	\caption{\Coopfw{}'s action latencies similarly improve with more cores.
-This requires \num{3} workers (4 total contexts) to improve upon the state-action latency of single-threaded in-NIC RL.\label{fig:vary-core-latency}}
-%\end{figure}
-
-\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects update cost and state-action latency respectively, scaling from a bias tile up to the full DDoS policy size.\footnote{Input vectors here all have 20 elements regardless of the policy design.}
-\Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning.
-%We note that this is a trivial case, in that the use of \emph{only} a bias tile negates the need for any input state (similar to a multi-arm bandit problem).
-We note that this is a trivial case, as using \emph{only} a bias tile returns a single preference list regardless of input state.
-
-\begin{figure}[t]
-	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-work}
-	}
-	\caption{\Coopfw{} processes updates faster than \Indfw{}---thus has higher online performance---on almost all policy sizes. Lower bit depths are effective on simple policies.\label{fig:vary-work}}
-\end{figure}
-
-\begin{figure}
-	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
-	}
-	\caption{State-action latency scales with additional work similarly to \cref{fig:vary-work}; though \SI{32}{\bit} firmwares become more effective sooner (at \num{3} work items).\label{fig:vary-work-latency}}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\begin{subfigure}{0.45\linewidth}
-		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/rl-perf-tester/latency-cumul}
-			\vspace{-10pt}
-		}
-		\caption{\approachshort's \Coopfw{} design achieves consistent, tight latency bounds.}
-	\end{subfigure}
-	\hspace{0.05\linewidth}
-	\begin{subfigure}{0.45\linewidth}
-		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/rl-perf-tester/latency-cumul-broad}
-		}
-		\caption{Tail latencies suffer in hosts---particularly when oversubscribed.}
-	\end{subfigure}
-	\caption{Cumulative state-action latency plots for \approachshort{} and host-based execution.\label{fig:lat-cumul}}
-\end{figure}
-
-%A noteworthy trend is that \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
-\SIlist{8;16}{\bit} versions of \approachshort{} underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency})---even though our implementation is optimised to read and write policy data in batches.
-Because the native register width on the NFP is \SI{32}{\bit}, the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
-This matches \SI{32}{\bit} becoming dominant in complex workloads: higher dimension tilings require more arithmetic operations.
-Most of the I/O comes \emph{after} this step, causing ALU use to dominate.
-This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) agents, where hashtable accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the online algorithm.
-%Smaller bit depths still give a \SIrange{2}{4}{$\times$} saving in memory for policy storage (enabling more fine-grained or complex policies), in exchange for slightly worse latency and throughput.
-To overcome this, we investigated bit-stuffing several values into a single word during atomic writeback (as the platform offers both \SIlist{32;64
-}{\bit} atomic addition).
-This is analogous to SIMD through clever use of padding bits, but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
-
-\fakepara{Work allocation}
-\Cref{fig:work-alloc-32} shows that our heuristic allocator (\emph{Balanced}) is key in achieving consistently low latencies/update times.
-The trend is repeated for all bit depths.
-The constant difference between \emph{Action} and \emph{Update Prep} scales linearly with bit depth, matching storage and lookup work in the serial portion of \emph{ParSa} (plots omitted).
-%The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its index, as batching work in contiguous chunks gives some workers only high-dimensionality tilings.
-The minor gap in lower bound performance between the \emph{Random} and \emph{Balanced} allocators suggests that further optimisations can be made.
-We expect that closing or exceeding this gap may require more complex modelling of hardware thread interactions, which lies far beyond the scope of in-NIC scheduling.
-Additionally, throughput occasionally plateaus due to bottlenecks in the output schedule until extra cores are allocated due to the granularity of the above tilings.
-We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
-%Though we have a \SI{4.2}{$\times$} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
-
-\begin{figure}
-	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
-	}
-	\caption{Action/update compute times in a \SI{32}{\bit} \Coopfw{} agent under different work schedulers.\label{fig:work-alloc-32}}
-\end{figure}
-
-%\begin{figure}
-%	\resizebox{\resultplotwidth}{!}{
-%		\includegraphics{../plots/build/rl-perf-tester/vary-core-tput-on-per}
-%	}
-%	\caption{Throughput per worker in a \Coopfw{} agent.\label{fig:tput-per-core}}
-%\end{figure}
-
-%An interesting aspect of \Coopfw{} and \emph{ParSa} is that adding cores has both diminishing returns and key thresholds to pass.
-%Consider \cref{fig:tput-per-core}, where the throughput per worker decreases over time but occasionally increases sharply.
-%Later downward ticks (\numrange{25}{29} workers) correspond to plateaus in throughput.
-%This is a problem stemming from the granularity of work items (i.e., tilings in \emph{ParSa}), where our scheduler cannot find a better solution to a bottleneck until extra cores are allocated.
-%We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
-%Though we have a \SI{4.2}{$\times$} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
-
-\fakepara{End-to-end RL latency}
-%?? PCIe RTT is 10us (neugebauer, BNN NFP paper), NFP RTT is 18us (own measurements). EMEM Ring one-way is 120ns.
-%?? Cziva~\parencite[pp.113]{DBLP:phd/ethos/Cziva18} discusses vNF times.
-%?? reference the inference times brought up in Taurus?
-Referring to \cref{fig:state-slip}, we take $t_2$ from \cref{tab:lats} for host and in-NIC processing times, and substitute $t_1+t_3$ for the state packet round-trip time to the decision site:
-\begin{description}
-	\item[\emph{In-NIC.}] As described in \cref{sec:agent-environment-communication}, EMEM rings have a median one way delay cross-island of \SI{140}{\nano\second}, giving a \emph{median \SI{34.63}{\micro\second} end-to-end inference latency}.
-	\item[\emph{Dedicated Collector.}] Employing DPDK to maximise performance, such hosts will add one-way PCIe delays of \SIrange{0.9}{2.3}{\micro\second} for network packets~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
-	A UDP packet carrying \num{20} elements of state in \approachshort{} is \SI{128}{\byte}, so costs \SI{1}{\micro\second} to forward, and the reply state-action pair invokes a slightly higher cost.
-	\emph{This gives an end-to-end inference latency of \SI{517.9}{\micro\second}}.
-	\item[\emph{vNF Offload.}] \Textcite{DBLP:journals/cm/CzivaP17} show that more lightweight vNF frameworks like GNF~\parencite{DBLP:journals/cm/CzivaP17} and ClickOS~\parencite{DBLP:conf/nsdi/MartinsAROHBH14} add \SIrange{45}{55}{\micro\second} \emph{additional} RTT latency above PCIe costs.
-	\emph{This gives an end-to-end inference latency of \SI{572.9}{\micro\second}}.
-\end{description}
-Thus, in-NIC classical RL offers \SIlist{14.96;16.54}{$\times$} speedups in latency over collector and vNF deployments respectively.
-We contrast these against deep RL policies on network tasks, which can take \SI{3}{\milli\second} to compute~\parencite{DBLP:journals/corr/abs-1910-04054}---2 orders of magnitude above \approachshort{} with identically sized (20-dim) input vectors.
-%In the name of fairness, we assume that rule installation uses same mechanism we recommend, but show how bad it can get? I.e. with rule installation costs etc.
-
-\fakepara{Co-existence with the dataplane}
-Our setup met \SI{40}{\giga\bit\per\second} for packet sizes $\ge$\SI{256}{\byte}.
-%For frame sizes of \SIlist{64;128;256}{\byte} input traffic rates were \SIlist{17.4;32.9;37.0}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2;18.1}{\mega\packet\per\second}).
-For frame sizes of \SIlist{64;128}{\byte}, input traffic rates were \SIlist{17.4;32.9}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2}{\mega\packet\per\second}).
-Passing this traffic over the NFP device running \approachshort, no packet losses were incurred at any rate of RL actions.
-We show the effect of RL workloads on the round-trip latencies of cross traffic through \cref{fig:dataplane-coop}.
-As observed latencies do not obey a normal distribution, we employ a one-tailed \emph{Mann-Whitney U test} to mark statistically significant population increases in latency ($p < 0.05$) with a ``+''.
-In general, statistically significant latency increases concentrate around smaller packet sizes.
-All (bar one) of these affected \nth{99} percentile latencies by under \SI{0.38}{\percent} (at most \SI{78}{\nano\second}).
-This slight degradation can be explained by increased pressure on the NFP's \emph{Command Push-Pull} (CPP) bus, which is responsible for handling cross-island accesses to memory (particularly IMEM/EMEM) and other resources.
-\approachshort{} places load on the CPP bus through its \inring{}/\outring{} EMEM rings and last-tier policy accesses.
-This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---the NFP P4 toolchain segments packets, storing metadata (e.g., MAC prepend) and the first bytes of a packet in a \SI{256}{\byte} CTM block and parking their payloads in EMEM.
-\SI{256}{\byte} packets overshoot this due to metadata, causing small I/O accesses at a high rate for packets sized around this cutoff.
-%Regardless, this effect is small when observed.
-
-The anomalous result is \SI{128}{\byte} packets at \num{3000} RL updates per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase, shown in \cref{fig:dataplane-example}.
-%This is observed through a shift of some packet latencies from the mode towards the tail, but no other changes in the distribution.
-We believe that the inbound request rate is weakly synchronised with traffic, causing a higher level of burstiness around accesses to the CPP bus.
-We expect that FPGA designs can avoid this problem by having dedicated \inring{}/\outring{} access mechanisms for an \approachshort{} agent.
-
-\begin{figure}
-	\centering
-	\begin{subfigure}{0.45\linewidth}
-		\resizebox{1.0\linewidth}{!}{
-%			\includegraphics{../plots/build/stress/heat-latency-median}
-			\includegraphics{../plots/build/stress/heat-latency-two-9-rel}
-%			\includegraphics{../plots/build/stress/heat-latency-two-9-perc}
-		}
-		\caption{Deviations in \nth{99} percentile cross-traffic RTTs.\label{fig:dataplane-heat}}
-	\end{subfigure}
-	\hspace{0.05\linewidth}
-		\begin{subfigure}{0.45\linewidth}
-		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/stress/histo-128B-0-3-trim}
-		}
-		\caption{Distribution of RTTs for \SI{128}{\byte} packets for \numlist{0;3000} RL actions/s.\label{fig:dataplane-example}}
-	\end{subfigure}
-	\caption{Effects on tail latency of cross-traffic caused by off-path RL compute.
-Statistically significant increases in population latency are concentrated on smaller packet sizes, and are typically sub-\SI{78}{\nano\second}.\label{fig:dataplane-coop}}
-\end{figure}
-
-\fakepara{Resource requirements}
-\Cref{tab:resources} shows how \approachshort{} consumes memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
-As one program is installed per ME, these results represent the minimum and maximum resource use on a single island (i.e., without replacing P4 workers).
-We observe negligible costs on shared EMEM ($\sim$\SI{4}{\mebi\byte}) from state/reward hashtables.
-The most significant costs arise due to policy data (\SI{405}{\kibi\byte} shared IMEM, \SI{90}{\kibi\byte} local CTM, \SI{15}{\kibi\byte} local CLS), which can be halved or quartered using \SIlist{16;8}{\bit} quantisation.
-This is a high upfront cost on per-island resources (CLS/CTM)---\approachshort{} leaves resources for other off-path dataplane applications, but is fairest from 3 cores onwards.
-
-%Thread local storage in CLS for compute/register spilling scales with the number of MEs required (requiring \SIlist{13.77;15.49}{\percent} per-ME for \Indfw{}/\Coopfw) due to precaching and the space needed to hold tile lists.
-%Due to the initial policy cost this falls below the fair share at 3 MEs, but always allows for co-existence with other asynchronous dataplane programs.
-
-\begin{table}
-\caption{NFP memory use due to \approachshort{} using \numlist{1;4} MEs (\SI{32}{\bit}).
-CLS and CTM are shared between all programs on the same island, while EMEM and IMEM are shared between all NFP programs on a NIC.\label{tab:resources}}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{
-		@{}c
-		S[table-format=4.2] S[table-format=2.2]
-		S[table-format=4.2] S[table-format=2.2]
-		S[table-format=4.2] S[table-format=2.2]
-		S[table-format=2.2] S[table-format=2.2]
-		S[table-format=3.2] S[table-format=2.2]
-		@{}
-	}
-	\toprule Firmware & \multicolumn{2}{c}{EMEM} & \multicolumn{2}{c}{EMEM Cache} & \multicolumn{2}{c}{IMEM} & \multicolumn{2}{c}{i5.CLS} & \multicolumn{2}{c}{i5.CTM}\\
-	& \multicolumn{1}{c}{\si{\mebi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} \\
-	\midrule Base P4 & 6776.67 & 88.24 & 268.52 & 2.91 & 858.28 & 10.48 & 0.00 & 0.00 & 0.00 & 0.00 \\
-	\Indfw(1) & 6780.21 & 88.28 & 2541.08 & 27.57 & 1263.28 & 15.42 & 24.75 & 38.67 & 94.25 & 36.82 \\
-	\Indfw(4) & 6780.22 & 88.28 & 2545.33 & 27.62 & 1263.28 & 15.42 & 51.18 & 79.97 & 107.00 & 41.80 \\
-	\Coopfw(1) & 6779.12 & 88.27 & 1773.59 & 19.24 & 1263.28 & 15.42 & 22.41 & 35.01 & 90.00 & 35.16 \\
-	\Coopfw(4) & 6779.12 & 88.27 & 1769.84 & 19.20 & 1263.28 & 15.42 & 52.16 & 81.49 & 90.00 & 35.16 \\
-	\bottomrule
-\end{tabular}
-}
-\end{table}
-
-%\fakepara{The impact of bit depth}
-%?? \Cref{fig:quant-acc}
-%?? \SI{5}{\bit} mantissa suffices for $\ge$ \SI{90}{\percent} relative accuracy, so \SI{8}{\bit} is fine (1S + 2E + 5M), \SI{16}{\bit} is better.
-%
-%\begin{figure}
-%	\includegraphics{../plots/build/marl-quant/accuracy-binary}
-%	\caption{Normalised accuracy of a converted, pre-trained floating-point tile-coded policy after conversion to $32Qn$ fixed-point.}\label{fig:quant-acc}
-%\end{figure}
-
-\fakepara{Deployability}
-Setup of \approachshort{} uses two packet types: \emph{setup}, which contains learning (hyper-)parameters and most aspects of a policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
-We found that setup and tiling packets take a mean \qtylist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing an agent to be swapped between online and offline operation (or repurposed) painlessly.
-These swaps are useful when an agent should cease learning (i.e., when performance converges), or if a change in the environment suggests that more training is needed.
-%Tiling packet processing scales linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
-Online-offline swaps for \Coopfw{} exhibit similar cost, however the need for explicit scheduling means that policy/tiling \emph{structure} changes (including the first complete setup) take \SI{422.63}{\micro\second} for the full-size policy described above.
-The time taken for \Coopfw{} to schedule its tasks was found to scale with the number of workers ($m$) and work items ($n$) as described earlier (\cref{sec:work-allocation}, $\mathcal{O}{\left(n\log{m}\right)}$).
-Ignoring the trivial solutions, reducing the worker count to \num{1} costs a mean \SI{238.22}{\micro\second}, while placing a single task incurs \SI{53.79}{\micro\second}.
-Policy \emph{data} changes require no additional work in any case, resolving purely to \texttt{memcpy}s.
-We observed that firmware installation (i.e., changing from \Indfw{} to \Coopfw{}, bit depth, or increasing maximum policy sizes) took a mean time of \SI{38.83}{\second}.
-%In the event that appropriate firmwares are not pre-compiled, we found that compiling and linking both \approachshort{} and the P4 toolchain took around \SI{35}{\second}, while changing only \approachshort{} parameters required around \SI{25}{\second}.
-Compiling and linking both \approachshort{} and the P4 toolchain took around \SI{35}{\second}, while changing only \approachshort{} parameters required around \SI{25}{\second}.
-These results show that \approachshort{} can be easily adapted and altered by network administrators once in place, and illustrates an advantage of SoC-based SmartNICs.
-
-\section{Potential Integrations}\label{sec:potential-integrations}
-To show the applicability and use of \approachshort{}, we propose an integration which would benefit from in-NIC RL; online DDoS mitigation.
-We support this using state-of-the-art P4 developments, and discuss how to balance online training with throughput in widespread deployment.
-%Additionally, we discuss how network administrators may combine offline (\Indfw) and online (\Coopfw) agents to achieve online learning while maximising action throughput in the majority of the network.
-
-%?? How do we move this intuition to the start? ELI5 as dimitris says, to get the full end-to-end pipeline across to the reader.
-
-\subsection{In-Network DDoS Defence}\label{sec:integ-1}
-Classical RL has seen recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
-\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and per-flow state to monitor how flows respond to bandwidth changes and packet drop.
-%---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} (observing how flow behaviour reacts to a change in rate limits) with more allowance for congestion-unaware protocols.
-Actions move flows up or down in punishment levels.
-
-\fakepara{Why in-NIC?}
-This approach co-hosts traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
-However, packet mirroring and offloading RL computation to a host/vNF are all sources of additional, consistent state-action latency.
-Both traffic measurement \emph{and} data-driven learning must be executed in this way.
-Unless running these stages in a dedicated pipeline (adding further processing latency), resource contention between processes will further impact tail latencies.
-Naturally, this requires high-performance hardware to be stacked at network egress points, potentially beyond reasonable space, power, or ventilation constraints.
-The solution to implement and improve upon this work using \approachshort{} is to place its RL agents on SmartNICs at AS edge nodes---a bump-in-the-wire deployment.
-
-%?? Any way to move some of this sooner? I.e., this is ``why move DDoS prevention here'', Stefanos suggested backporting some of this to justify ``why \emph{RL} in-NIC''?
-
-\fakepara{Inputs}
-To collate the required inputs and state, we then examine the recent innovations of the community.
-Low-latency, pure-P4 solutions to extract and record per-flow TCP state directly in the dataplane such as Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} and Sonata~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18} are well-studied.
-In fact, the statistics offered by Dapper are a super-set of the local input state values employed by \emph{Guarded} agents, offering an opportunity to further improve their efficacy. 
-We propose placing such monitors in the P4 dataplane, existing on-chip alongside the \approachshort{} agent.
-The required global state (load measures from network paths) must still come from elsewhere in the network; this is now the element at highest risk of becoming stale, but the least likely to vary significantly in response to individual actions.
-The original work uses theoretical ``ground truth'' rewards, whose correct implementation and designs were left as an open challenge. 
-We posit that INDDoS~\parencite{tnms-ddos-victim-ident}, which uses Count-min Sketches to estimate DDoS victim cardinality, could be an effective reward function source---i.e., using the number of detected victims as a loss value.
-
-\fakepara{Integrating \approachshort}
-Before each monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{\outring{} Ring} for any generated actions.
-As we note in \cref{sec:limitations}, these actions would be placed into a local hash table \emph{and} exported to the attached controller to be inserted as P4 rules in batches.
-Afterwards, packet ingress timestamps would be used to emulate the \emph{Timed Random Sequential} (TRS) scheduler used by the anti-DDoS agents for rate-controlled work, where state vectors would be selected and passed into \approachshort.
-By design, active flows which are not \emph{judged} after a configurable time are discarded to prune the work set and allow new flows to be seen by the agent.
-The tight bounds on execution time known \emph{a priori} make it easy to calculate the maximum number of decisions which can be made per deadline.
-Reward values would then be separately inserted by a modified INDDoS table.
-
-%?? Can I have a P4 code snippet here?
-
-%Reducing state-action latency (i.e., with \Coopfw) is useful for minimising the noise inherent in learning.
-Agents here are limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
-Accordingly, this use-case benefits most from an increase in \emph{throughput} using \Indfw{}.
-Higher throughput means that network flows are more likely to be judged in every timestep, even when flow cardinality is high---making it more likely that changes in flow behaviour will be observed, acted on, and learned from.
-We note that the TRS scheduler is designed to handle large numbers of attack flows, combining new state vectors while the asynchronous agent is itself busy.
-A number of flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
-As shown in \cref{sec:evaluation}, \approachshort{} far exceeds the throughput of host offloading.
-The control plane can then use wildcards or specific matches to narrow down or expand the set of flows to be controlled dynamically, though explicit (TRS) scheduling is still key in such adversarial environments.
-
-%\subsection{Something host-specific?}\label{sec:integ-2}
-%Accelerate something at an end-host... Flow control? Something datacentre-y? A novel CCA? What?
-
-\subsection{Network Deployment Considerations}
-In a networked deployment, a subset of \approachshort{} nodes could be \coopfw{} agents, training online, while most other nodes run \indfw{} designs to meet throughput guarantees.
-The control plane would then combine, downsample, and distribute these improved policies between offline agents.
-This can be taken further, using policy deltas to enable out-of-path transfer learning for more complex models such as neural networks.
-
-\section{Related Work}
-\fakepara{In-network ML}
-Taurus~\parencite{DBLP:journals/corr/abs-2002-08987} proposes that efficient line-rate inference can be achieved using a configurable grid of map-reduce units in the packet pipeline (implementing e.g., LSTMs and SVMs).
-On CGRA hardware, they achieve sub-\si{\micro\second} extra latency.
-\emph{IIsy}~\parencite{DBLP:conf/hotnets/XiongZ19} shows how \emph{classical ML inference} (SVMs, Na\"{i}ve Bayes, etc.) can be converted into match-action tables compatible with \emph{any} P4 deployment.
-They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA.
-%Neither considers online model training.
-
-A recent line of research in the community has been to investigate \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate packet classification.
-\emph{BaNaNa SPLIT} shows this as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is carried out on the \emph{CPU} to reduce latency imposed by GPU batching and transfer, but fully-connected layers can be accelerated further by NICs.
-%However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}; the authors hypothesise that networking hardware would be a better fit for some parts of the inference pipeline.
-In-network packet tagging and classification by pre-trained BNNs is shown by \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
-N3IC achieves packet inference in \SI{45}{\micro\second} on the NFP, and \SI{0.3}{\micro\second} on NetFPGA for \SI{256}{\bit} inputs.
-Comparatively, \approachshort{}-\Coopfw{} can process an identically-sized input (8-dim) in a median \SI{13.83}{\micro\second}.
-Our work handles larger inputs (\SI{640}{\bit}) at lower latencies (\SI{34}{\micro\second}), and offers online learning.
-We expect that a NetFPGA implementation of \approachshort{} would enjoy a similar factor of speedup.
-%No works investigate the runtime training of BNN function approximators in-NIC.
-
-\textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantisation on the NFP, using \emph{in-path} compute rather than our asynchronous model.
-Inference latency on small networks (3 layers, \num{30} neurons) can be as high as \SI{500}{\micro\second} on line rate traffic, emphasising the value of path-adjacent compute.
-
-We stress that none of these approaches (or that we have seen) tackle the issue of \emph{online learning and control} in-network---we believe \approachshort{} has broken new ground in this regard.
-
-\fakepara{In-network ML \emph{acceleration}}
-Optimisation of distributed neural network training is an area where in-network compute has been key for general NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18} and RL-specific procedures~\parencite{DBLP:conf/isca/LiLYCSH19} using NetFPGAs to introduce floating point adders to process dedicated packet classes.
-Distributed training is a partition-aggregate workload, where gradient vectors are sent back to a controller to update the `true' model---in-NIC processing allows these to be aggregated \emph{in-network}, overcoming incast behaviour and host bottlenecks.
-
-\fakepara{RL for network control}
-\Textcite{DBLP:conf/sigcomm/LiangZJS19} use deep RL techniques to train an agent which can build efficient decision tree packet classifiers.
-Their approach, \emph{NeuroCuts}, produces time/space efficient classifiers for use in constrained environments (e.g., network hardware) beyond existing heuristic solutions.
-Deep RL techniques have been used for QUIC congestion control optimisation~\parencite{DBLP:journals/corr/abs-1910-04054}.
-A key facet of this work is the need for asynchronous RL in networks, where pauses for DNN-based inference can significantly harm throughput.
-
-\fakepara{PDP design for asynchronous compute}
-\emph{PANIC}~\parencite{DBLP:conf/hotnets/StephensAS18} places a routing fabric between distinct packet/data processing elements \emph{in a SmartNIC}.
-Such designs would enable general, asynchronous, and novel compute in SmartNICs and switches, for instance offering consistent and easy to use communication between workers versus hard-coded ME relationships.
-Event-driven versions of P4 have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}.
-Timer events and device state changes would empower in-network RL use-cases, signalling timesteps for RL agents or new, effective, fine-grained sources of input state.
-
-\section{Conclusion}
-We have presented \emph{\approachshort{}}, bringing \emph{online reinforcement learning} to the dataplane.
-\approachshort{} has shown how classical RL techniques make online learning possible by simplifying update logic and enabling parallel processing.
-In-NIC use of these algorithms enabled a \SIrange{15}{21}{$\times$} reduction in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput compared to host offloading.
-The deployment environment and asynchronous design were shown to eliminate PCIe delays and impose minimal impact on carried dataplane traffic.
-We also showed how \approachshort{} scales with additional compute resources at deployment to improve on decision latency and throughput.
-Our throughput-optimal design, \Indfw{}, improves upon these metrics with \emph{just one worker}.
-
-In future, we aim to examine the performance of individual applications driven by \approachshort---both classical and deep RL-based---and how a NetFPGA implementation can offer further latency and throughput improvements.
-A promising avenue here would be to investigate constant transfer learning between online \approachshort{} agents and high-throughput offline function approximators such as BNNs.
-%It is our hope that high port-density programmable devices will, in time, expose similar asynchronous, on-chip offloading capability to current SmartNIC devices.
-%We also are interested in `classical' approaches for other tasks which may have been overlooked in favour of more performant (yet complex) solutions, yet enable in-network deployment.
+\section{Next Steps}
+We intend to measure the effects of our data format and function approximation choices on overall accuracy and convergence times for synthetic data, comparing different bit depths against tile-coded floating-point numbers and more expressive functions such as neural networks.
+Moreover, we intend to build \approachshort{} into ultra low latency control scenarios, such as short-flow prioritisation or active queue management.
 
 \begin{acks}
-	The authors would like to thank Rhys Simpson for his comments, and for discussions on the soundness of SIMD-like optimisations.
-	They would additionally like to thank Stefanos Sagkriotis, Mircea Iordache-\c{S}ic\u{a}, and Haruna Adoga for their comments and feedback.
+	%	The authors would like to thank Rhys Simpson for his comments, and for discussions on the soundness of SIMD-like optimisations.
+	%	They would additionally like to thank Stefanos Sagkriotis, Mircea Iordache-\c{S}ic\u{a}, and Haruna Adoga for their comments and feedback.
+	%	We thank the anonymous reviewers of SOSR'21 and XXX'22 for their comments.
 	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grants~\grantnum{gs-epsrc}{EP/N509668/1}, \grantnum{gs-epsrc}{EP/N033957/1}].
@@
 %\bibliography{reference}
+
 \printbibliography
 
-%\begin{appendices}
-%	\section{Old ``What could we do''?}
-%	How do we evaluate this without looking into classification performance?
-%	\begin{itemize}
-%		\item Analytically?
-%		\begin{enumerate}
-%			\item Prove that tile splitting scheme over memory hierarchy can improve performance.
-%			\begin{itemize}
-%				\item Hit rate of different tiles according to tile dimensionality?
-%				\item Access time for each tier of memory hierarchy.
-%			\end{itemize}
-%		\end{enumerate}
-%		\item Experimentally?
-%		\begin{enumerate}
-%			\item Create ``VNF'' testbed: mirror traffic from switch to a host for telemetry processing/flow state.
-Send flow state over to another node who computes RL actions.
-RL actions installed on switch.
-%			\begin{itemize}
-%				\item Why not co-host RL computation with telemetry? Could be separate functions, could want to ensure that neither impacts the performance of the other.
-%			\end{itemize}
-%			\item Create ``PDP'' testbed: Load firmware, pass in traffic. Traffic processing/action compute/table mod occurs in PDP.
-%			\item What traffic? Healthy mix of attack (UDP), Discord legit Opus VoIP traffic, HTTP bulk-transfer traffic.
-%			\item Measure $t_{1 \cdots 3}$ according to \cref{fig:state-slip} in both VNF and PDP contexts.
-%			\begin{description}
-%				\item[$t_1$]
-%				\begin{description}
-%					\item[PDP] Packet ingress $\rightarrow$ INT Island $\rightarrow$ State arrival in Control ME.
-%					\item[VNF] Packet mirror time $\rightarrow$ telemetry VNF $\rightarrow$ state arrival in agent vnf.
-%				\end{description}
-%				
-%				\item[$t_2$]
-%				\begin{description}
-%					\item[PDP] Time to compute action over parallel Policy MEs (RL Island, SmartNIC).
-%					\item[VNF] Time to compute action in host program (on VNF) from received state.
-%				\end{description}
-%				
-%				\item[$t_3$]
-%				\begin{description}
-%					\item[PDP] Modify local tables across ME/Island boundaries (same device), verify rule present (or get ACK).
-%					\item[VNF] Generate OpenFlow control messages, send OF message to switch, verify rule present (or get ACK).
-%				\end{description}
-%			\end{description}
-%			\item HYPOTHESIS: PDP will present a reduction in these three times for per-flow state.
-%			\item QUESTION: Does same apply to learning?
-%			\item NOTE: since we're not actually assessing the \emph{quality} of output decisions, we can cut corners on what data is sent out and where.
-I.e., in-band network telemetry island can be skipped, if we move over representative volumes of state to/from the RL island with approximately correct delay as recorded by other studies.
-%		\end{enumerate}
-%	\end{itemize}
-%	I can think of one way to evaluate this if we employ classification performance:
-%	\begin{enumerate}
-%		\item Run the existing work~\parencite{DBLP:journals/tnsm/SimpsonRP20}, acquire a policy for one node after full info sharing.
-%		\item Load this policy onto the RL Island (\cref{fig:netro-arch}).
-%		\item Either feed in raw state packets, or feed in traffic if we can get INT working.
-%		\item For each level of quantisation needed, examine average performance.
-%		\item Don't compare against old numbers for final perf: use a ``VNF'' setup similar to above---this will probably suffer a bit compared to old results, where everything was on the same machine.
-Make it relative/normalised.
-%	\end{enumerate}
-%
-%	\section{Intermediate steps?}
-%	System?
-%	\begin{itemize}
-%		\item Setup other core as event source.
-%		\begin{itemize}
-%			\item \emph{\num{2} week.}
-%			\item Idea: proxy for idea of INT core mentioned elsewhere, combined with the main periodic event sending that the last RL work relied upon.
-So, say "you can put that state management and extraction on another core, here's what the comms cost is" or similar.
-%		\end{itemize}
-%		\item Fixes to hash table use, signalling.
-%		\begin{itemize}
-%			\item \emph{\num{1} week.}
-%			\item Blocks progress on optimisation, correct multi-stream RL.
-%		\end{itemize}
-%		\item Optimise/parallelise action compute/learning.
-%		\begin{itemize}
-%			\item \emph{\num{2} week.}
-%			\item Value? Proves this approach can scale according to resource availability/demand on the deployment device, at compile-time anyhow.
-%		\end{itemize}
-%	\end{itemize}
-%	
-%	Experiments?
-%	\begin{itemize}
-%		\item Measure sum of times $t_{1...3}$ described elsewhere.
-%		\begin{itemize}
-%			\item \emph{\num{1} week.}
-%			\item Currently have individual, need ``end-to-end''.
-%		\end{itemize}
-%		\item Throughput impact testing.
-%		\begin{itemize}
-%			\item \emph{\num{1} week.}
-%			\item Requires current impl, creation of basic testbed and packet forwarding setup.
-%			\item Use-case independent.
-%		\end{itemize}
-%		\item Measure execution, training costs for varying policy sizes.
-%		\begin{itemize}
-%			\item \emph{\num{1.5} weeks.}
-%			\item Policies need not be meaningful, and can be garbage data. Use-case independent.
-%			\item Partially blocked on multicore accelerated/optimised execution.
-%		\end{itemize}
-%		\item Measure rule installation cost on different platforms.
-%		\begin{itemize}
-%			\item \emph{\num{1.5} weeks.}
-%			\item Blocked on creation of basic testbed and packet forwarding setup.
-%		\end{itemize}
-%		\item Investigate stationarity.
-%		\begin{itemize}
-%			\item \emph{\numrange{2}{3} weeks.}
-%			\item Requires full use-case implementation.
-%			\item Requires careful thought about sensible changes to normality.
-%			\item Might be bolstered by comparison against an offline-trained system?
-%		\end{itemize}
-%	\end{itemize}
-%	
-%	\section{Everything we could possibly evaluate}
-%	
-%	Evaluation will need to be divided into 2 main categories:
-%	\begin{enumerate}
-%		\item Simulation/emulation based on mininet.
-%		\begin{itemize}
-%			\item These can only assess algorithmic properties, as performance measurements in mininet (i.e., a full mult-agent system) are not representative.
-%		\end{itemize}
-%		\item Testbed setup (2/3 configurations):
-%		\begin{itemize}
-%			\item These will assess key system performance properties of a \emph{single agent}, in control of a single switch/NIC as a bump in the wire.
-Bump-in-the-wire device will carry traffic between two hosts.
-%			\item \emph{Time permitting, may need to send state vectors in packets rather than do feature/state extraction on-chip}.
-%			\item \emph{Necessary.} RL runs on bump-in-the-wire SmartNIC.
-%			\item \emph{Either/both.} P4-based off-path RL.
-Custom packet actions (e.g., random drop) will be implemented in P4 with externs.
-Packets/digests mirrored to state extraction machine.
-Agent may be co-hosted with state-extraction.
-Rules inserted via P4-device specific RTE.
-%			\item \emph{Either/both.} OvS-based off-path RL.
-Custom packet actions are already implemented in OvS as OpenFlow extensions.
-Packets/digests mirrored to state extraction machine.
-Agent may be co-hosted with state-extraction.
-%		\end{itemize}
-%	\end{enumerate}
-%	
-%	\subsection{What is this different from, or better than?}
-%	List below mainly for the benefit of helping enumerate everything I can compare against, and what their drawbacks are.
-%	
-%	\begin{itemize}
-%		\item (Deep) neural network-backed RL.
-%		\begin{itemize}
-%			\item NNs infeasible to update online.
-Ordinarily this requires larger volumes of data than can be gathered in real time, but becomes substantially more difficult after change of representation (quantisation, binarisation).
-An online system (such as this) should be able to adapt to changes in underlying traffic, optimisation criteria, etc., dynamically.
-%			\item Complexity of function approximation means that policy updates are more computationally complex to compute.
-For instance, NNs require use of the backpropagation algorithm, rather than the simpler classical schemes which need only the tile set already identified in action selection.
-%			\item Only heavily quantised (i.e., binary neural networks) suitable for in-switch execution.
-%		\end{itemize}
-%		
-%		\item MARL and other ``agents as network functions'' designs.
-%		\begin{itemize}
-%			\item Action, state collection, updates etc.
-happen on nodes outside of the main packet path in this paradigm.
-Thus, state measurements, action computation, and so on, should induce the state drift described above.
-Also makes these approaches ``slower to react'' in theory.
-%		\end{itemize}
-%	\end{itemize}
-%	
-%	At present, there are no other RL works in PDPs, let alone with online policy updates.
-%	Other existing RL use cases suitable for adaptation to the ``bump-in-the-wire'' model are proving tricky to find.
-%	
-%	\subsection{Emulation metrics}
-%	Key (MVP) metrics will be \textbf{bolded}, and expanded with a subsubsection.
-%	
-%	\textbf{Quantisation accuracy}, quantised training accuracy, \textbf{stationarity}.
-%	
-%	\subsubsection{Quantisation accuracy}
-%	Examination of whether lower resolution quantisation impacts RL agent decision accuracy substantially.
-%	
-%	This has been done using MARL in mininet, see \cref{fig:quant-acc}.
-%	
-%	\subsubsection{Interaction with stationarity}
-%	Learning time of the system as compared to known changes in normality.
-%	
-%	I.e., in the DDoS case: how long does an agent take to respond to the onset of an attack?
-%	A change in attack strategy?
-%	The conclusion of an attack?
-%	There are no citations on \emph{how} DDoS attacks evolve, just that they do, and that this can occur on various timescales.
-%	
-%	\subsection{Testbed metrics}
-%	Key (MVP) metrics will be \textbf{bolded}, and expanded with a subsubsection.
-%	
-%	\textbf{Traffic throughput (bytes, packets)}, \textbf{packet processing delay}, \textbf{action execution time}, \textbf{learning time}, from-scratch learning accuracy, \textbf{rule installation delay/cost}, host and NIC memory costs, policy installation time, ``accuracy'' in the DDoS case.
-%	
-%	\subsubsection{Traffic throughput}
-%	Main research question: does moving RL logic onto the NIC reduce maximum throughput of the NIC, even if calculation is on a different functional unit?
-%	This is relevant in both \si{\byte\per\second} and packets per second.
-%	
-%	This is application independent, and would assess whether the RL units (acting independently on the netronome) affect throughput compared with a basic P4 program at different work intensities.
-%	Ideally, this would show no effect on forwarding performance.
-%	
-%	\subsubsection{Packet processing delay}
-%	Main research question: does the presence of off-path (but on-chip) RL execution meaningfully increase packet processing delay?
-How does this compare to a P4 program with an equivalent number of tables?
-What about to OvS?
-%	
-%	\subsubsection{Action execution \& learning time}
-%	Main research question: does moving RL logic from the host to the NIC substantially reduce delays from state measurement, to state receipt, to action computation and installation?
-%	These constitute $t_{1...3}$ discussed above.
-%	Primarily, these need to be looked at using varying policy sizes and distribution across memory regions.
-%	
-%	We ideally want these both with and without any acceleration from extra cores on the netronome: the point of this being that we can demonstrate that the system (or a similar NetFPGA system) can scale upwards according to available system resources or demands.
-%	
-%	\subsubsection{Rule installation delay/cost.}
-%	Main research question: how long does it take to install rules in the above testbed setups? What is the downtime of so doing?
-%	
-%	This is application-dependent: see earlier discussions on rule batching and local extern/hash-table use.
-%\end{appendices}
-
 \end{document}

### 2021-10-12 paper/poster-abstract.tex
Poster-er, with Reward lookup/storage fixes.
Not fixed on Ind yet, that one's gonna bite me in the proverbial, I assume...

@@
 sigconf,natbib=false
-%,anonymous=true
+,anonymous=true
 ]{acmart}
@@
 
-%\DeclareFieldFormat[inproceedings]{doi}{}
-%\DeclareFieldFormat[article]{doi}{}
-%\DeclareFieldFormat*{url}{}
+\DeclareFieldFormat[inproceedings]{doi}{}
+\DeclareFieldFormat[article]{doi}{}
+\DeclareFieldFormat*{url}{}
 %\DeclareFieldFormat[online]{url}{\mkbibacro{URL}\addcolon\space\url{#1}}
@@
 %\DeclareFieldFormat[article]{doi}{}
-%\DeclareFieldFormat[inproceedings]{editor}{}
-%\DeclareFieldFormat[proceedings]{editor}{}
-%\DeclareFieldFormat[article]{editor}{}
+\DeclareFieldFormat[inproceedings]{editor}{}
+\DeclareFieldFormat[proceedings]{editor}{}
+\DeclareFieldFormat[article]{editor}{}
+\DeclareFieldFormat[inproceedings]{isbn}{}
+\DeclareFieldFormat[proceedings]{isbn}{}
+\DeclareFieldFormat[article]{isbn}{}
 
@@
 	\centering
-		\includegraphics[keepaspectratio, width=0.6875\linewidth]{figures/arch-with-p4}
+		\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/arch-with-p4}
 		\caption{\approachshort{} brings low-latency, online reinforcement learning to SoC- and NPU-based SmartNICs.
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/coop}
+	\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/coop}
 	\caption{Parallel policy execution.
@@
 \end{figure}
-Automatic network optimisation, control, and defence are at last becoming commonplace through data-driven techniques such as RL methods, where every change and its effects improve future decisions made by an agent.
-In tandem, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} (PDP) hardware have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
+Automatic network optimisation, control, and defence are becoming commonplace through data-driven techniques such as RL methods, where every change and its effects improve future decisions made by an agent.
+In tandem, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} hardware have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
 This ecosystem exposes not only runtime reconfigurable packet processing, but per-packet and per-flow traffic measurement state that can greatly aid network operation~\parencite{DBLP:conf/sosr/GhasemiBR17}.
@@
 \fakepara{Machine learning in the dataplane}
-To process this fine-grained state both at line rate \emph{and} at low latency, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:journals/corr/abs-2009-02353}.
-All works in this category, while powerful, operate by converting a \emph{pre-trained} model into a suitable representation, such as a binary neural network or string of match-action tables.
+To process this fine-grained state both at line rate \emph{and} at low latency, there has been keen interest in executing ML models in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:journals/corr/abs-2009-02353}.
+Works in this category, while powerful, operate by converting a \emph{pre-trained} model into a suitable representation, such as a binary neural network or string of match-action tables.
 %These works have shown the value of in-network ML: high-throughput, low latency response to network changes.
@@
 \fakepara{Problems of online training}
-Programmable dataplane hardware, being designed solely for efficient packet processing, lacks floating-point arithmetic support even in the case of more general purpose NPU-type SmartNICs.
+Programmable dataplane hardware, being designed solely for efficient packet processing, lacks floating-point arithmetic support, even in the case of more general purpose NPU-type SmartNICs.
 Additionally, deep neural network training relies on backpropagation, can require memorizing many a sizeable replay buffer for RL, and needs many mini-batches of data for stable training.
 The natural solution is to pass training data to a host machine or dedicated accelerator, yet this adds delays in crossing the PCIe bus, kernel-user handover, and buffering---adding latency while learning.
-The above (NIC-suitable) inference schemes still, even if their data formats solve the issue of floating point use.
-Online learning thus requires careful choices in both sample data formats, and the learned function approximation scheme.
+The above (NIC-suitable) inference schemes still must be trained offline, even though their data formats solve the issue of floating point use.
+Online learning thus requires careful choices in sample data formats and the learned function approximation scheme.
 
 \section{\approachshort{}'s design}
-\approachshort{} makes RL workloads feasible under on SmartNICs by using quantised fixed-point representations of action values, tile-coded policies, and one-step temporal-difference RL algorithms such as Sarsa~\parencite{RL2E}.
-These allow us to evaluate and update policies using \emph{only integer arithmetic}.
-We observe that tile coded inference is a map-reduce problem (\cref{fig:par-tilecode}), and over integers admits a novel \emph{wait-free Sarsa} RL implementation (exploiting the parallelism built in to SmartNICs).
-In addition to their runtime simplicity, while these functions have lower model capacity they do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays.
+\approachshort{} makes RL workloads feasible on SmartNICs by using quantised fixed-point representations of action values, tile-coded policies, and one-step temporal-difference RL algorithms such as Sarsa~\parencite{RL2E}.
+These allow us to evaluate and update policies using \emph{only integer arithmetic}, and are computationally simple.
+We observe that tile coded inference is a map-reduce problem (\cref{fig:par-tilecode}), and thus over integers admits a novel \emph{wait-free Sarsa} RL algorithm (exploiting the parallelism built in to SmartNICs).
+While these functions have lower theoretical model capacity, they do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays or minibatches.
 
@@
 \fakepara{Inference/Updates}
-State is mirrored to a set of packet processing threads, which compute hit tiles from a locally held work set.
-During inference, these threads atomically add values to a shared action preference list, while policy updates are guaranteed by tile-coding to proceed without data races (\cref{fig:single-and-parallel:parallel}).
+State is sent to a set of packet processing threads, which compute hit tiles from a locally held work set.
+During inference, these threads atomically add values to a shared action preference list, while tile-coding guarantees that policy updates proceed without data races (\cref{fig:single-and-parallel:parallel}).
 In the case of our Netronome implementation, these run on lightweight cores known as \emph{microengines} (MEs)---which could be substituted for suitable FUs on CGRA/FPGA hardware.
-Additionally, we have a single-threaded implementation, \indfw, which forgoes this cooperative model to maximise offline throughput (i.e., by having many parallel agents handle state).
+Additionally, we have a single-threaded implementation, \indfw, which forgoes this cooperative model to maximise offline throughput (by having many parallel single-threaded agents).
 
@@
 \section{Evaluation}
-We compare our implementation of \approachshort{} against a tile-coded RL agent written in numpy on a powerful host (i7-6700K@\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}), using policy dimensions used by existing DDoS prevention work\parencite{DBLP:journals/tnsm/SimpsonRP20}, across various quantisation bit depths.
+We compare our implementation of \approachshort{} against a tile-coded RL agent written in numpy on a powerful host (i7-6700K@\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}), using the policy dimensions of existing DDoS prevention work~\parencite{DBLP:journals/tnsm/SimpsonRP20}, across various quantisation bit depths.
 \Cref{fig:lat-cumul} shows that our main design (\coopfw) achieves order-of-magnitude improvements over our single-threaded variant (\indfw) and a host implementation, as well as considerably tighter tail behaviour on the SmartNIC.

### 2021-10-12 code/nfp/rl-test/packet/packet.c

@@ -586,9 +586,9 @@ void state_packet(
 void reward_packet(__declspec(cls) struct rl_config *cfg, union pad_tile value, uint32_t reward_insert_loc) {
 	// Switch on self->reward_key
 	// if shared, place into key 0 I guess?
-	uint32_t loc = 0;
+	int32_t loc = 0;
 	uint64_t long_key = reward_insert_loc;
-	int32_t added = 0;
+	int32_t added = 1;
 
 	switch (cfg->reward_key.kind) {
 		case KEY_SRC_SHARED:

### 2021-10-12 code/nfp/rl-test/rl-prims.c

@@ -258,30 +258,38 @@ __intrinsic void update_action_preferences(__addr40 _declspec(emem) uint32_t *ti
 }
 
 // Selects the state-action / reward key from a given state vector.
-_intrinsic tile_t select_key(struct key_source key_src, tile_t *state_vec) {
+_intrinsic struct key_select_result select_key(struct key_source key_src, tile_t *state_vec) {
+	struct key_select_result out = {0};
+
 	switch(key_src.kind) {
 		case KEY_SRC_SHARED:
+			out.skip_tcam = 1;
 			break;
 		case KEY_SRC_FIELD:
-			return state_vec[key_src.body.field_id];
+			out.loc = state_vec[key_src.body.field_id];
 		case KEY_SRC_VALUE:
-			return key_src.body.value;
+			// out.skip_tcam = 1;
+			out.loc = key_src.body.value;
+			break;
 	}
 
-	// Assumes shared in usual case.
-	return 0;
+	return out;
 }
 
-_intrinsic tile_t fat_select_key(struct key_source key_src, __addr40 __declspec(emem) tile_t *state_vec) {
+_intrinsic struct key_select_result fat_select_key(struct key_source key_src, __addr40 __declspec(emem) tile_t *state_vec) {
+	struct key_select_result out = {0};
+
 	switch(key_src.kind) {
 		case KEY_SRC_SHARED:
+			out.skip_tcam = 1;
 			break;
 		case KEY_SRC_FIELD:
-			return state_vec[key_src.body.field_id];
+			out.loc = state_vec[key_src.body.field_id];
 		case KEY_SRC_VALUE:
-			return key_src.body.value;
+			// out.skip_tcam = 1;
+			out.loc = key_src.body.value;
+			break;
 	}
 
-	// Assumes shared in usual case.
-	return 0;
+	return out;
 }

### 2021-10-12 code/nfp/rl-test/rl.c

@@ -261,21 +261,30 @@ void state_packet_delegate(
 	if (cfg->do_updates) {
 		// This dummies the access cost.
 		// For some reason, it bugs out if struct size is lte 32bit?
-		uint64_t reward_key = fat_select_key(cfg->reward_key, (__declspec(emem) __addr40 tile_t *)pkt->packet_payload);
-		uint64_t state_key = fat_select_key(cfg->state_key, (__declspec(emem) __addr40 tile_t *)pkt->packet_payload);
+		struct key_select_result reward_key =
+			fat_select_key(cfg->reward_key, (__declspec(emem) __addr40 tile_t *)pkt->packet_payload);
+		struct key_select_result state_key =
+			fat_select_key(cfg->state_key, (__declspec(emem) __addr40 tile_t *)pkt->packet_payload);
 
 		int32_t state_added = 0;
 
-		int32_t reward_found = CAMHT_LOOKUP(reward_map, &reward_key);
+		int32_t reward_found = reward_key.loc;
+		int32_t state_found = state_key.loc;
 
-		// This is dummied until I figure out the mechanism better.
-		int32_t state_found = CAMHT_LOOKUP_IDX_ADD(state_action_map, &state_key, &state_added);
 		uint32_t changed_key = 0;
 
 		uint8_t updating_on_this_cycle;
 
-		if (state_action_map_key_tbl[state_found] != state_key) {
-			state_action_map_key_tbl[state_found] = state_key;
+		if (!reward_key.skip_tcam) {
+			reward_found = CAMHT_LOOKUP(reward_map, &reward_key.loc);
+		}
+
+		if (!state_key.skip_tcam) {
+			state_found = CAMHT_LOOKUP_IDX_ADD(state_action_map, &state_key.loc, &state_added);
+		}
+
+		if (state_action_map_key_tbl[state_found] != state_key.loc) {
+			state_action_map_key_tbl[state_found] = state_key.loc;
 			changed_key = 1;
 		}
 

### 2021-10-12 code/nfp/rl-test/rl.h

@@ -221,4 +221,9 @@ struct state_action_pair {
 
 #endif /* WORKER_BARGAIN_BUCKET_SIMD */
 
+struct key_select_result {
+	uint64_t loc;
+	uint8_t skip_tcam;
+};
+
 #endif /* !_RL_H_ */

### 2021-10-12 code/nfp/rl-test/worker_config_marker.h

@@ -1,5 +1,7 @@
 #ifndef _CONFIG_MARKER_H
 #define _CONFIG_MARKER_H
-#define WORK_ALLOC_STRAT ALLOC_FILL_HEAVY
+// #define ENABLE_STRESS_TEST
 #define _RL_CONFIG_SOLO_REMOTE_WORK
+#define WORK_ALLOC_STRAT ALLOC_FILL_HEAVY
+#define VERIFY_OUTPUTS
 #endif /* !_CONFIG_MARKER_H_ */

### 2021-10-13 paper/noms-short-paper.tex
One page shorter, made ugly as per IEEE reqs.
3 pages to go.

--- 2021-06-11 paper/paper-10pt.tex
@@
 \documentclass[
-sigconf,natbib=false
-,anonymous=true
+%sigconf,natbib=false
+%,anonymous=true
+conference
 ,10pt
-]{acmart}
-
-\usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists]{savetrees}
+]{IEEEtran}
+
+%\usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists]{savetrees}
+%\usepackage[subtle, indent, lists]{savetrees}
 
@@
 %\usepackage{ftnright}
-
-\usepackage[%
-%false
-]{anonymous-acm}
 
@@
 
-\usepackage[title]{appendix}
+%\usepackage[title]{appendix}
 
@@
 \usepackage[font=footnotesize]{subcaption}
-\usepackage{awesomebox}
 
@@
 \usepackage{url}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage[nameinlink]{cleveref}
@@
 
+\newcommand{\subparagraph}{}
+\usepackage{titlesec}
+\titlespacing*{\section}{0pt}{1.5ex}{0.7ex}
+\titlespacing*{\subsection}{0pt}{1.5ex}{0.7ex}
+\titlespacing*{\paragraph}{0pt}{1.5ex}{0.7ex}
+\titlespacing*{\subsubsection}{0pt}{1.5ex}{0.7ex}
+
 % End Imports
@@
 % Copyright
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
-%\setcopyright{none}
-\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-%\settopmatter{printacmref=false, printccs=false, printfolios=true}
-%\settopmatter{printfolios=true}
-\settopmatter{printacmref=false, printccs=true, printfolios=true}
-
-% DOI
-\acmDOI{12345}
-
-% ISBN
-\acmISBN{67890}
-
-%Conference
-\acmConference[SOSR '21]{Symposium of SDx Research 2021}{September \numrange{20}{21}, 2021}{Online}
-\acmYear{2021}
-%\copyrightyear{}
-
-%% {} with no args suppresses printing of the price
-%\acmPrice{15.00}
-\acmPrice{}
 
@@
 
-\newcounter{insightc}
-\newenvironment{insight}
-	{
-		\begin{tipblock}\refstepcounter{insightc}\textbf{Insight \theinsightc:}\em
-	}
-	{
-		\end{tipblock}
-	}
+%\newcounter{insightc}
+%\newenvironment{insight}
+%	{
+%		\begin{tipblock}\refstepcounter{insightc}\textbf{Insight \theinsightc:}\em
+%	}
+%	{
+%		\end{tipblock}
+%	}
+
+\usepackage[hang,flushmargin]{footmisc}
 
@@
 %	\author{Paper \# XXX, XXX pages}
-	 \author{Kyle A. Simpson}
-	 \orcid{0000-0001-8068-9909}
-	 \affiliation{%
-	   \institution{University of Glasgow}
-	   \city{Glasgow} 
-	   \country{Scotland}
-	 }
-	 \email{k.simpson.1@research.gla.ac.uk}
-	 \author{Dimitrios P. Pezaros}
-\orcid{0000-0003-0939-378X}
-	 \affiliation{%
-	\institution{University of Glasgow}
-	\city{Glasgow} 
-	\country{Scotland}
-}
-\email{Dimitrios.Pezaros@gla.ac.uk}
-
-\setlength{\abovecaptionskip}{5pt}
-\setlength{\belowcaptionskip}{0pt}
-
-\setlength{\textfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
-\setlength{\floatsep}{3.0pt plus 0.0pt minus 2.0pt}
-
-\setlength{\dbltextfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
-\setlength{\dblfloatsep}{3.0pt plus 0.0pt minus 2.0pt}
+	 \author{Kyle A. Simpson, Dimitrios P. Pezaros,\\University of Glasgow, Glasgow, Scotland}
+
+%\setlength{\abovecaptionskip}{5pt}
+%\setlength{\belowcaptionskip}{0pt}
+%
+%\setlength{\textfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
+%\setlength{\floatsep}{3.0pt plus 0.0pt minus 2.0pt}
+%
+%\setlength{\dbltextfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
+%\setlength{\dblfloatsep}{3.0pt plus 0.0pt minus 2.0pt}
 
@@
 %\renewcommand{\shortauthors}{Simpson \emph{et al}.}
+
+\maketitle
 
@@
 \approachshort{} makes online learning possible in constrained SmartNIC hardware by returning to classical RL techniques which simplify update logic and enable parallel processing.
-Compared to offloading, we achieve a \SI{21}{$\times$} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{$\times$} improvement in online throughput for real-world policy designs.
+Compared to offloading, we achieve a \SI{21}{\texttimes} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{\texttimes} improvement in online throughput for real-world policy designs.
 In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
@@
 
-\begin{CCSXML}
-	<ccs2012>
-	<concept>
-	<concept_id>10003033.10003068.10003069</concept_id>
-	<concept_desc>Networks~Data path algorithms</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10003033.10003099.10003102</concept_id>
-	<concept_desc>Networks~Programmable networks</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10003752.10010070.10010071.10010261</concept_id>
-	<concept_desc>Theory of computation~Reinforcement learning</concept_desc>
-	<concept_significance>300</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10010147.10010257.10010258.10010261</concept_id>
-	<concept_desc>Computing methodologies~Reinforcement learning</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Networks~Data path algorithms}
-\ccsdesc[500]{Networks~Programmable networks}
-\ccsdesc[500]{Computing methodologies~Reinforcement learning}
-
 %\keywords{ACM proceedings, \LaTeX, text tagging}
 
-\maketitle
-
-\setlength{\aweboxleftmargin}{0.12\linewidth}
-\setlength{\aweboxcontentwidth}{1.97\linewidth}
+%\setlength{\aweboxleftmargin}{0.12\linewidth}
+%\setlength{\aweboxcontentwidth}{1.97\linewidth}
 
@@
 	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm to accelerate tile coded policy inference and updates (\cref{alg:parsa}),
-	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{$\times$} latency reduction compared to commodity hardware (\SI{21}{$\times$} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
+	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{\texttimes} latency reduction compared to commodity hardware (\SI{21}{\texttimes} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
 	\item A description of how \approachshort{} would integrate with state-of-the-art PDP applications to perform fully in-NIC, fast, automated DDoS mitigation (\cref{sec:potential-integrations}).
@@
 
-\begin{figure}
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\resizebox{0.8\linewidth}{!}{
-			\centering
-			\begin{tikzpicture}
-				\node[circle, draw] (state) {$S$};
-				\node[circle, draw] (state') at ($(state) + (0, -2)$) {$S'$};
-
-				\node (agent) at ($(2, 0) + (state)$) {Agent};
-				\node[below of=agent] (action) {Action $A$};
-
-				\node[right of=state'] {$+ R$};
-
-				\draw[->] (state) -- (state') node[midway, right] {$A$};
-
-				\draw[dotted, ->, bend left = 30] (state) -- (agent);
-				\draw[->] (agent) -- (action);
-				\draw[dotted, ->] (action) -- (state);
-		\end{tikzpicture}}
-		\caption{Theory: state measurement, action computation, and learning are zero-cost.}
-	\end{subfigure}
-	\hspace{0.05\linewidth}
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\resizebox{0.6\linewidth}{!}{
-			\centering
-			\begin{tikzpicture}
-				\node[circle, draw] (state) {$S$};
-				\node[circle, draw] (state') at ($(state) + (0, -1.5)$) {$S'$};
-				\node[circle, draw] (state'') at ($(state') + (0, -1.5)$) {$S''$};
-
-				\node (agent) at ($(2, 0) + (state)$) {Agent};
-				\node[below of=agent] (action) {Action $A$};
-
-				\node[right of=state''] {$+ R$};
-
-				\draw[->] (state) -- (state') node[midway, right] {$\varnothing$};
-				\draw[->] (state') -- (state'') node[midway, right] {$A$};
-
-				\draw[dotted, ->, bend left = 30] (state) -- (agent) node[midway, above] {$t_1$};
-				\draw[->] (agent) -- (action) node[midway, right] {$t_2$};
-				\draw[dotted, ->] (action) -- (state') node[midway, below] {$t_3$};
-		\end{tikzpicture}}
-		\caption{Reality: policy costs lead to \emph{state drift}---over a delay $t_1+t_2+t_3$, state $S$ becomes $S'$.}
-	\end{subfigure}
-	\caption{Asynchronous RL delays and state slippage.\label{fig:state-slip}}
-\end{figure}
+%\begin{figure}
+%	\begin{subfigure}{0.45\linewidth}
+%		\centering
+%		\resizebox{0.8\linewidth}{!}{
+%			\centering
+%			\begin{tikzpicture}
+%				\node[circle, draw] (state) {$S$};
+%				\node[circle, draw] (state') at ($(state) + (0, -2)$) {$S'$};
+%				
+%				\node (agent) at ($(2, 0) + (state)$) {Agent};
+%				\node[below of=agent] (action) {Action $A$};
+%				
+%				\node[right of=state'] {$+ R$};
+%				
+%				\draw[->] (state) -- (state') node[midway, right] {$A$};
+%				
+%				\draw[dotted, ->, bend left = 30] (state) -- (agent);
+%				\draw[->] (agent) -- (action);
+%				\draw[dotted, ->] (action) -- (state);
+%		\end{tikzpicture}}
+%		\caption{Theory: state measurement, action computation, and learning are zero-cost.}
+%	\end{subfigure}
+%	\hspace{0.05\linewidth}
+%	\begin{subfigure}{0.45\linewidth}
+%		\centering
+%		\resizebox{0.6\linewidth}{!}{
+%			\centering
+%			\begin{tikzpicture}
+%				\node[circle, draw] (state) {$S$};
+%				\node[circle, draw] (state') at ($(state) + (0, -1.5)$) {$S'$};
+%				\node[circle, draw] (state'') at ($(state') + (0, -1.5)$) {$S''$};
+%				
+%				\node (agent) at ($(2, 0) + (state)$) {Agent};
+%				\node[below of=agent] (action) {Action $A$};
+%				
+%				\node[right of=state''] {$+ R$};
+%				
+%				\draw[->] (state) -- (state') node[midway, right] {$\phi$}; %$\varnothing$};
+%				\draw[->] (state') -- (state'') node[midway, right] {$A$};
+%				
+%				\draw[dotted, ->, bend left = 30] (state) -- (agent) node[midway, above] {$t_1$};
+%				\draw[->] (agent) -- (action) node[midway, right] {$t_2$};
+%				\draw[dotted, ->] (action) -- (state') node[midway, below] {$t_3$};
+%		\end{tikzpicture}}
+%		\caption{Reality: policy costs lead to \emph{state drift}---over a delay $t_1+t_2+t_3$, state $S$ becomes $S'$.}
+%	\end{subfigure}
+%	\caption{Asynchronous RL delays and state slippage.\label{fig:state-slip}}
+%\end{figure}
 
@@
 
-\begin{figure*}
-	\centering
-	\begin{subfigure}{0.47\linewidth}
-		\centering
-		\includegraphics[keepaspectratio, width=0.78\linewidth]{figures/ind}
-		\caption{\Indfw{} (offline throughput-optimal).
+%\begin{figure*}
+%	\centering
+%	\begin{subfigure}{0.47\linewidth}
+%		\centering
+%		\includegraphics[keepaspectratio, width=0.78\linewidth]{figures/ind}
+%		\caption{\Indfw{} (offline throughput-optimal).
 \emph{Workers} independently pull commands from (and push actions to) the environment, locking policy access for updates.\label{fig:single-and-parallel:single}}
-	\end{subfigure}
-	\hspace{0.04\linewidth}
-	\begin{subfigure}{0.47\linewidth}
+%	\end{subfigure}
+%	\hspace{0.04\linewidth}
+%	\begin{subfigure}{0.47\linewidth}
+%		\centering
+%		\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/coop}
+%		\caption{\Coopfw{} (online-optimal).
+A single \emph{controller} delegates RL computation and updates to many \emph{minion} threads, who operate on independent subtasks.\label{fig:single-and-parallel:parallel}}
+%	\end{subfigure}
+%	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to either latency or throughput needs.\label{fig:single-and-parallel}}
+%\end{figure*}
+
+\begin{figure}
 	\centering
 	\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/coop}
-		\caption{\Coopfw{} (online-optimal).
+	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to either latency or throughput needs---\Coopfw{} shows this optimally online.
 A single \emph{controller} delegates RL computation and updates to many \emph{minion} threads, who operate on independent subtasks.\label{fig:single-and-parallel:parallel}}
-	\end{subfigure}
-	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to either latency or throughput needs.\label{fig:single-and-parallel}}
-\end{figure*}
+\end{figure}
 
@@
 Internally, \approachshort{} either has all its cores act independently (\cref{fig:single-and-parallel:single}) or cooperate to solve each task (\cref{fig:single-and-parallel:parallel})---with different latency-throughput benefits (\cref{sec:action-and-update-computation}).
-We open-source our firmware and control programs for the benefit of the community.\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission.}}
+%We open-source our firmware and control programs for the benefit of the community.\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission.}}
+We open-source our firmware and control programs for the benefit of the community.\footnote{Link redacted during submission.}
 %Moving beyond the Netronome platform, we describe how our architecture may be adapted and improved on by bespoke hardware or FPGA-based deployment.
@@
 Two firmware models govern how these tasks (action selection, policy updates) are carried out:
-\begin{description}
+\begin{LaTeXdescription}
 	\item[\Indfw{} (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each works sequentially.
@@
 \emph{State vectors} are stored for update computation.
-\end{description}
+\end{LaTeXdescription}
 Each offers a different point of optimisation; if updates are disabled, then the \indfw{} model can maximise throughput, while the \coopfw{} model is designed to minimise decision latency and needs no locks to update the policy (increasing \emph{online learning} throughput).
@@
 \section{Evaluation}\label{sec:evaluation}
-We investigate the performance of \approachshort{} compared to classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{$\times$} speedup in median--\num{99.99}\nthscript{th} state-action latency and \SI{9.9}{$\times$} greater online learning throughput.
+We investigate the performance of \approachshort{} compared to classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{\texttimes} speedup in median--\num{99.99}\nthscript{th} state-action latency and \SI{9.9}{\texttimes} greater online learning throughput.
 Crucially, in-NIC execution offers tight tail latency bounds compared to host-based approaches.
@@
 Testing machines were as follows, with \SI{32}{\gibi\byte} RAM:
-\begin{description}
+\begin{LaTeXdescription}
 	\item[\emph{MidServer}] Intel Xeon Bronze 3204 (\qtyproduct[product-units=single]{6 x 1.9}{\giga\hertz}),
@@
 	\item[\emph{Collector}] Intel Core i7-6700K (\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}).
-\end{description}
+\end{LaTeXdescription}
 \approachshort{} was evaluated on server blades (\emph{Mid/HighServer}), each hosting a single Netronome Agilio LX \numproduct{1 x 40}GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
@@
 \Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy.\footnote{For brevity, we omit integer numpy results---against a float implementation, median action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}
-We show that when using 4 MEs of the NFP-6480, \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse (\SIlist{15;21}{$\times$} speedups over a \emph{Collector} host).
-Importantly, \Indfw{} achieves lower median state-action latencies (\SI{2.79}{$\times$}) \emph{and} update times (\SI{2.63}{$\times$}) than a dedicated \emph{Collector} host \emph{while requiring only a single core or dedicated functional unit}.
+We show that when using 4 MEs of the NFP-6480, \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse (\SIlist{15;21}{\texttimes} speedups over a \emph{Collector} host).
+Importantly, \Indfw{} achieves lower median state-action latencies (\SI{2.79}{\texttimes}) \emph{and} update times (\SI{2.63}{\texttimes}) than a dedicated \emph{Collector} host \emph{while requiring only a single core or dedicated functional unit}.
 Crucially, in-NIC execution gives far tighter bounds on tail latency compared to host offloading.
@@
 \newlength{\resultplotwidth}
-\setlength{\resultplotwidth}{0.8\linewidth}
+\setlength{\resultplotwidth}{0.849\linewidth}
+%\setlength{\resultplotwidth}{0.9\linewidth}
+%\setlength{\resultplotwidth}{0.8\linewidth}
 
@@
 To make the comparison fair in the context of many-core CPU environments, we include per-core throughput.
-\Indfw{} achieves \SI{2.82}{$\times$} higher offline throughput per core than commodity \emph{Collector} hardware in spite of the NFP-6480 having a considerably slower clock speed (\SI{0.29}{$\times$}).
+\Indfw{} achieves \SI{2.82}{\texttimes} higher offline throughput per core than commodity \emph{Collector} hardware in spite of the NFP-6480 having a considerably slower clock speed (\SI{0.29}{\texttimes}).
 When compounded with the abundance of such weaker chips, in-NIC RL is able to deliver much higher throughput.
-%As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} that of a dedicated collector machine, due to the locking requirements mandating coordinated division of work.
-As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{$\times$} that of a dedicated collector machine, as the write lock around policy updates creates a bottleneck.
+%As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{\texttimes} that of a dedicated collector machine, due to the locking requirements mandating coordinated division of work.
+As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{\texttimes} that of a dedicated collector machine, as the write lock around policy updates creates a bottleneck.
 
@@
 \begin{figure}[t]
+	\centering
 	\resizebox{\resultplotwidth}{!}{
@@
 \begin{figure}
+	\centering
 	\resizebox{\resultplotwidth}{!}{
@@
 	\begin{subfigure}{0.45\linewidth}
+		\centering
 		\resizebox{1.0\linewidth}{!}{
@@
 This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) agents, where hash table accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the online algorithm.
-%Smaller bit depths still give a \SIrange{2}{4}{$\times$} saving in memory for policy storage (enabling more fine-grained or complex policies), in exchange for slightly worse latency and throughput.
+%Smaller bit depths still give a \SIrange{2}{4}{\texttimes} saving in memory for policy storage (enabling more fine-grained or complex policies), in exchange for slightly worse latency and throughput.
 To overcome this, we investigated bit-stuffing several values into a single word during atomic writeback (as the platform offers both \SIlist{32;64
@@
 We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
-%Though we have a \SI{4.2}{$\times$} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
+%Though we have a \SI{4.2}{\texttimes} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
 
 \begin{figure}
+	\centering
 	\resizebox{\resultplotwidth}{!}{
@@
 %We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
-%Though we have a \SI{4.2}{$\times$} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
+%Though we have a \SI{4.2}{\texttimes} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
 
@@
 Referring to \cref{fig:state-slip}, we take $t_2$ from \cref{tab:lats} for host and in-NIC processing times, and substitute $t_1+t_3$ for the state packet round-trip time to the decision site:
-\begin{description}
+\begin{LaTeXdescription}
 	\item[\emph{In-NIC.}] As described in \cref{sec:agent-environment-communication}, EMEM rings have a median one way delay cross-island of \SI{140}{\nano\second}, giving a \emph{median \SI{34.63}{\micro\second} end-to-end inference latency}.
@@
 	\emph{This gives an end-to-end inference latency of \SI{572.9}{\micro\second}}.
-\end{description}
-Thus, in-NIC classical RL offers \SIlist{14.96;16.54}{$\times$} speedups in latency over collector and vNF deployments respectively.
+\end{LaTeXdescription}
+Thus, in-NIC classical RL offers \SIlist{14.96;16.54}{\texttimes} speedups in latency over collector and vNF deployments respectively.
 We contrast these against deep RL policies on network tasks, which can take \SI{3}{\milli\second} to compute~\parencite{DBLP:journals/corr/abs-1910-04054}---2 orders of magnitude above \approachshort{} with identically sized (20-dim) input vectors.
@@
 \approachshort{} has shown how classical RL techniques make online learning possible by simplifying update logic and enabling parallel processing.
-In-NIC use of these algorithms enabled a \SIrange{15}{21}{$\times$} reduction in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput compared to host offloading.
+In-NIC use of these algorithms enabled a \SIrange{15}{21}{\texttimes} reduction in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput compared to host offloading.
 The deployment environment and asynchronous design were shown to eliminate PCIe delays and impose minimal impact on carried dataplane traffic.
@@
 
-\begin{acks}
+\fakepara{Acknowledgements}
 The authors would like to thank Rhys Simpson for his comments, and for discussions on the soundness of SIMD-like optimisations.
 They would additionally like to thank Stefanos Sagkriotis, Mircea Iordache-\c{S}ic\u{a}, and Haruna Adoga for their comments and feedback.
-	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grants~\grantnum{gs-epsrc}{EP/N509668/1}, \grantnum{gs-epsrc}{EP/N033957/1}].
-\end{acks}
+We thank the anonymous reviewers of SOSR'21 and NOMS'22 for their comments.
+This work was supported in part by the Engineering and Physical Sciences Research Council [grants EP/N509668/1, EP/N033957/1].
 
@@
 %\bibliography{reference}
+
+\clearpage
+
 \printbibliography

### 2021-10-14 paper/poster-abstract.tex
Texttimes in poster.

@@
 On update times, we see \qty{63.12}{\micro\second} at 99.99\nthscript{th} percentile for a \qty{32}{\bit} policy of this size.
-We relate a \qty{15}{$\times$} latency (\coopfw) and \qty{2.82}{$\times$} offline throughput-per-core (\indfw) in spite of the Netronome's considerably slower clock speed (\qty{0.29}{$\times$}), showing the value of exploiting SmartNIC parallelism.
+We relate a \qty{15}{\texttimes} latency (\coopfw) and \qty{2.82}{\texttimes} offline throughput-per-core (\indfw) in spite of the Netronome's considerably slower clock speed (\qty{0.29}{\texttimes}), showing the value of exploiting SmartNIC parallelism.
 

### 2021-10-14 paper/noms-short-paper.tex
1.3 pages to go...

@@
 %,anonymous=true
-conference
+conference,
+comsoc,
 ,10pt
@@
 
-%\usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists]{savetrees}
+\usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists,tracking=normal]{savetrees}
 %\usepackage[subtle, indent, lists]{savetrees}
@@
 \newcommand\bmmax{2}
+\usepackage[cmintegrals]{newtxmath}
 \usepackage{bm}
@@
 
+%\usepackage{xparse}
+%\ExplSyntaxOn
+%\NewDocumentCommand{\crefnames}{ m m m }
+%{
+%	\clist_map_inline:nn { #1 } { \crefname{##1}{#2}{#3} }
+%}
+%\ExplSyntaxOff
+%
+%\crefnames{section,subsection}{\S}{\S\S}
+\crefformat{section}{\S{}#2#1#3}
+\crefformat{subsection}{\S{}#2#1#3}
+
 \newcommand{\fakepara}[1]{\noindent\textbf{#1:}}
@@
 \usepackage[hang,flushmargin]{footmisc}
+\usepackage{orcidlink}
+
+\newcommand{\mytitle}{Revisiting~the~Classics: Online~RL~in~the~Programmable~Dataplane}
+
+\hypersetup{
+	colorlinks,
+	citecolor=black,
+	filecolor=black,
+	linkcolor=black,
+	urlcolor=black,
+	pdftitle={\mytitle{}},
+	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros}
+}
+\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}}}
+
+\title{\mytitle}
+\author{Kyle~A.~Simpson~\orcidlink{0000-0001-8068-9909}, Dimitrios~P.~Pezaros~\orcidlink{0000-0003-0939-378X}\\\emph{University of Glasgow}, Glasgow, Scotland\\\email{k.simpson.1@research.gla.ac.uk}, \email{Dimitrios.Pezaros@Glasgow.ac.uk}}
 
 \begin{document}
-	\title{Revisiting~the~Classics: Online~RL~in~the~Programmable~Dataplane}
 
@@
 %	\author{Paper \# XXX, XXX pages}
-	 \author{Kyle A. Simpson, Dimitrios P. Pezaros,\\University of Glasgow, Glasgow, Scotland}
-
-%\setlength{\abovecaptionskip}{5pt}
-%\setlength{\belowcaptionskip}{0pt}
-%
-%\setlength{\textfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
-%\setlength{\floatsep}{3.0pt plus 0.0pt minus 2.0pt}
-%
-%\setlength{\dbltextfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
-%\setlength{\dblfloatsep}{3.0pt plus 0.0pt minus 2.0pt}
+
+\setlength{\abovecaptionskip}{0pt}
+\setlength{\belowcaptionskip}{0pt}
+
+\setlength{\textfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
+\setlength{\floatsep}{3.0pt plus 0.0pt minus 2.0pt}
+
+\setlength{\dbltextfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
+\setlength{\dblfloatsep}{3.0pt plus 0.0pt minus 2.0pt}
 
@@
 We present \emph{\approachshort{}}---\approach{}---the first work to bring \emph{online reinforcement learning} to the dataplane.
-\approachshort{} makes online learning possible in constrained SmartNIC hardware by returning to classical RL techniques which simplify update logic and enable parallel processing.
+\approachshort{} makes online learning possible in constrained SmartNIC hardware by returning to classical RL techniques---avoiding neural networks.
+Our design allows weak yet highly parallel SmartNIC NPUs to be competitive against commodity x86 hosts, despite having fewer features and slower cores.
 Compared to offloading, we achieve a \SI{21}{\texttimes} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{\texttimes} improvement in online throughput for real-world policy designs.
 In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
-\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, improving on offloading, and are quickly reconfigurable at runtime compared to reinstalling device firmware.
-
+\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, and are quickly reconfigurable at runtime compared to reinstalling device firmware.
 \end{abstract}
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.6875\linewidth]{figures/arch-with-p4}
+	\includegraphics[keepaspectratio, width=0.65\linewidth]{figures/arch-with-p4}
 	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
@@
 Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing automatic traffic optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17,DBLP:conf/conext/GiladSGRS20}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
-In RL methods, every change and its effects improve future decisions made by an agent.
+In RL methods, every change and its effects improve future decisions.
 
@@
 The promise of PDP hardware is that we can move the entire monitoring and analysis stack \emph{into the dataplane itself, and have it evolve to incorporate new approaches}.
-Consider a DDoS mitigation system which could react to dataplane-specific, per-packet state in microseconds, and \emph{start learning again just as quickly} in the face of new data or behaviours.
+%Consider a DDoS mitigation system which could react to dataplane-specific, per-packet state in microseconds, and \emph{start learning again just as quickly} in the face of new data or behaviours.
 The P4 ecosystem already presents novel, openly-available, fine-grained traffic measurement techniques that can be installed and controlled with ease~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
@@
 
-It is important to make this feasible as offloading learning to commodity machines adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
-For context, \emph{Deep Neural Network} (DNN) training relies on backpropagation to compute gradients, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many mini-batches of data for stable training.
-These lead to high compute and storage costs, inducing capital expenditure for dedicated accelerators.
-Moreover, high latencies \emph{caused} by offloading and expensive inference harm the learning process~\parencite{DBLP:journals/firai/TravnikMSP18} and impact runtime application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
+It is important to make this feasible; offloading to commodity hosts adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
+For context, deep neural network training relies on backpropagation, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many minibatches for stability.
+These lead to high compute and storage costs, inducing capital expenditure for dedicated accelerators to overcome the batching needed to operate at line-rate---while BrainWave~\parencite{DBLP:conf/isca/FowersOPMLLAHAG18} can reduce batching (and thus tail latencies) by \qty{32}{\texttimes}~\parencite{Duarte2019}, inference still takes $\mathcal{O}{\left(\text{\si{\milli\second}}\right)}$.
+Moreover, novel DMA techniques such as \emph{GPUDirect}~\parencite{gpudirect} halve but do not eliminate such PCIe transfers.
+The high latencies \emph{caused} by offloading and inference harm learning~\parencite{DBLP:journals/firai/TravnikMSP18} and runtime application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
 If we can bring online learning \emph{to the dataplane}, then we can take advantage of rich, local state while minimising these latencies (and their impact on the learned policy).
@@
 We show that online RL can be brought to the dataplane by such methods, where it can act on locally extracted state and is made \emph{more efficient by dataplane hardware}.
-In particular, we exploit how SmartNIC devices often expose general-purpose compute to provide path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
+In particular, we exploit how SmartNIC devices often expose general-purpose compute~\parencite{DBLP:conf/sigcomm/LiuCSKPG19} to provide path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
 As many of these devices have engineering and development histories which predate P4, general compute beyond P4's limits~\parencite{p4-psa} is surprisingly common.
@@
 	\item An analysis of why in-NIC RL is needed and best-placed to interact with the network, made feasible by classical RL methods and quantisation (\cref{sec:motivation}),
-	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands of network traffic analysis (\cref{sec:design}),
-	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm to accelerate tile coded policy inference and updates (\cref{alg:parsa}),
+	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands of network traffic control (\cref{sec:design}),
+	\item \emph{ParSa}, a wait-free, parallel RL algorithm to accelerate tile coded policy inference and updates (\cref{alg:parsa}),
 	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{\texttimes} latency reduction compared to commodity hardware (\SI{21}{\texttimes} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
@@
 Crucially, we focus on low port-density devices.
-We envision that this processing is suited to the network edge or key vantage points, rather than core switches.
+%We envision that this processing is suited to the network edge or key vantage points, rather than core switches.
 These devices' designs make it reasonable to move processing \emph{outside of the packet pipeline}; SmartNICs expose additional cores, and NetFPGAs allow for the synthesis of independent functional units.
 %The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
-The main benefit is that control logic can be moved onto the NIC \emph{without impacting packet processing}.
+The main benefit is that RL logic can be moved onto the NIC \emph{without impacting packet processing}.
 %?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
-As Ethernet speed increases, packet deadlines grow tighter; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second} at \SI{40}{\giga\bit\per\second}, giving for example a \SI{3.99}{\micro\second} deadline on Netronome (which has \num{312} P4 pipelines).
-In-pipeline execution thus increases the risk of packet drops or stalls.
+%As Ethernet speed increases, packet deadlines grow tighter; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second} at \SI{40}{\giga\bit\per\second}, giving for example a \SI{3.99}{\micro\second} deadline on Netronome (which has \num{312} P4 pipelines).
+%In-pipeline execution thus increases the risk of packet drops or stalls.
 
@@
 %We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
-However, many compatible devices have legacies long predating these developments. 
-For instance, many-core SoC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
+However, many compatible devices long predate these developments. 
+Many-core SoC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
 %This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
-Currently, low port-density devices like these are most likely to have specific general-purpose compute, as they are designed for high-performance offloading and middlebox development.
+Currently, low port-density devices like these are most likely to have specific general-purpose compute and high degrees of parallelism~\parencite{DBLP:conf/sigcomm/LiuCSKPG19}, as they are designed for high-performance offloading and middlebox development.
 %Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
-Regardless, floating-point operations key to ML workloads are very rarely supported outside of deep learning accelerators (which lack packet switching functionality).
-Moreover, the limited memory and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
+%Regardless, floating-point operations key to ML workloads are very rarely supported outside of deep learning accelerators (which lack packet switching functionality).
+%Moreover, the limited memory and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.
 
@@
 
-\emph{Reinforcement learning} (RL) algorithms train an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task~\parencite{RL2E}.
-Like most ML methods, RL algorithms use gradient information to update the parameters used to approximate a function with \emph{reward measurements} from the environment.
-RL's \emph{Markov Decision Process} (MDP) structure allows online learning of a state-action map in a model-free way, and can step through local minima when needed compared to ML.
-In networking, this allows for learning to respond to on-device state or to handle rapidly evolving problems.
-RL algorithms are generally agnostic to the policy approximation used, and classical single-step methods are simple---most of the compute work is occupied by the policy.
-E.g., single-step, semi-gradient Sarsa~\parencite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to update and learn a policy online.
+\emph{Reinforcement learning} (RL) trains an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task~\parencite{RL2E}.
+Like most ML methods, RL algorithms use gradients to update the parameters used to approximate a function with \emph{reward measurements} from the environment.
+RL's Markov Decision Process structure allows online learning of a state-action map in a model-free way, and can step through local minima when needed compared to ML.
+In networking, this allows for learning from on-device state or handling rapidly evolving problems.
+RL algorithms are agnostic to the policy approximation used so long as they are differentiable, and are computationally simple.
+E.g., Sarsa~\parencite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to learn a policy online.
 
@@
 
-\fakepara{RL in asynchronous environments}
-%There remains some degree of divergence between the theory and implementation of RL agents.
-%Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
-%The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
-%Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
-Consider \cref{fig:state-slip}: the RL formulation assumes that an agent receives new state at fixed intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
-This adds noise to the state-action mapping, which harms learning and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
-%Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
-Reducing these costs requires detailed agent-environment co-design.
-This principle has influenced real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, where asynchrony is necessary for high-speed applications.
-This often requires that state measurements are fused~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
-`Stopping the world' causes significant performance loss, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control.
+%\fakepara{RL in asynchronous environments}
+%%There remains some degree of divergence between the theory and implementation of RL agents.
+%%Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
+%%The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
+%%Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
+%Consider \cref{fig:state-slip}: the RL formulation assumes that an agent receives new state at fixed intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
+%This adds noise to the state-action mapping, which harms learning and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
+%%Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
+%Reducing these costs requires detailed agent-environment co-design.
+%This principle has influenced real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, where asynchrony is necessary for high-speed applications.
+%This often requires that state measurements are fused~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
+%`Stopping the world' causes significant performance loss, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control.
 
@@
 
-This cost is what drives us to \emph{bring online reinforcement learning to the dataplane}---referring again to \cref{fig:state-slip}, we would place P4-based state measurement ($t_1$), simple policy approximation ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
-In networks, actions and analysis are most likely to be installed on bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts---all of which make in-NIC training a perfect fit.
-As such, our goal is to collocate \emph{all} functions which comprise an RL agent on the same chip or device.
-Both programmable devices and the network environment make this more difficult, as we'll examine in the sequel.
+%This cost is what drives us to \emph{bring online reinforcement learning to the dataplane}---referring again to \cref{fig:state-slip}, we would place P4-based state measurement ($t_1$), simple policy approximation ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
+%In networks, actions and analysis are most likely to be installed on bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts---all of which make in-NIC training a perfect fit.
+%As such, our goal is to collocate \emph{all} functions which comprise an RL agent on the same chip or device.
+%Both programmable devices and the network environment make this more difficult, as we'll examine in the sequel.
 
@@
 
-\subsection{Efficient policy approximation}
+\subsection{Tile-coded policy approximation}
 %Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
@@
 %Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the computational cost of gradient computation via backpropagation remains too significant on embedded hardware.
-Complex \emph{Deep Neural Networks} (DNNs), particularly when used as an RL agent's policy, require vast amounts (possibly compute-years~\parencite{DBLP:journals/corr/abs-1912-06680}) of experience to converge on an accurate parameter set, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
-Although there has been a vested interest in efficiently \emph{executing} complex function approximators like DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the storage and computational costs of backpropagation remain too significant on embedded hardware.
+%Complex \emph{Deep Neural Networks} (DNNs), particularly when used as an RL agent's policy, require vast amounts (possibly compute-years~\parencite{DBLP:journals/corr/abs-1912-06680}) of experience to converge on an accurate parameter set, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
+%Although there has been a vested interest in efficiently \emph{executing} complex function approximators like DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the storage and computational costs of backpropagation remain too significant on embedded hardware.
 
@@
 
-We return to what classical RL methods can offer us; in particular, the simple linear function approximation given by \emph{tile coding}~\parencite[pp.\ \numrange{217}{221}]{RL2E}.
+Returning to classical RL, we examine the linear function approximation of \emph{tile coding}~\parencite[pp.\ \numrange{217}{221}]{RL2E}.
 A policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
-This allows any input-output mapping, such as converting packet header data into a preference list for queue priorities.
-A state vector produces a single `hit' in each tiling, all of which then correspond to a list of action preference lists.
-Inference (choosing an action) is simply summing over all such lists to obtain a final action preference list, selecting the maximum.
-Updating uses this same list of lists after the next decision is made, adjusting the value of the last selected action using a \emph{temporal-difference value} ($\delta_t$) computed by Sarsa.
+%This allows any input-output mapping, such as converting packet header data into a preference list for queue priorities.
+A state vector produces a single `hit' in each tiling, all of which then correspond to a list of action preference lists---where computing a hit requires division by a known tile width.
+Inference is simply summing over all such lists to obtain a final preference list, selecting the highest-value action.
+Learning uses this same list after the next decision is made, adjusting the value of the previous action using a \emph{temporal-difference value} ($\delta_t$) computed by Sarsa.
 Crucially, this internal representation has no data dependencies between any tile calculations for an input.
-As a result, both action selection and policy updates are map-reduce problems.
-To select an action, we can produce a task for each tiling---retrieving the list of action values contributed by the tile activated by the input state.
-An aggregate step sums up all action preference lists, and the action with the highest value is then selected.
-Updating the policy (with a new $\delta_t$ value) has no aggregate step, and as tile indices map to disjoint memory regions no locks are required.
+%As a result, both action selection and policy updates are map-reduce problems.
+%To select an action, we can produce a task for each tiling---retrieving the list of action values contributed by the tile activated by the input state.
+%An aggregate step sums up all action preference lists, and the action with the highest value is then selected.
+%Updating the policy (with a new $\delta_t$ value) has no aggregate step, and as tile indices map to disjoint memory regions no locks are required.
 %As we show in \cref{sec:design}, it is possible to make the inference aggregate step wait-free.
 
-For instance, a policy with $m$ tiling sets (each having a set of input dimensions), with $n$ overlapping tilings then creates $m \times n$ separate tasks.\footnote{Work items could be retrieving each individual action's value (giving $m \times n \times a$ tasks for $a$ actions).
-This requires tile coding to be repeated $a$ times per tile, which is likely extremely wasteful outside of heavy parallelism.}
-As an example, a real-world policy size which we examine (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets of \num{8} tilings creates \num{129} work items.
-Computing a tile hit uses one division per dimension, which is an expensive ALU operation to require.
-However, if tile widths are fixed at powers-of-two, then these may be replaced with right bitshifts.
+%For instance, a policy with $m$ tiling sets (each having a set of input dimensions), with $n$ overlapping tilings then creates $m \times n$ separate tasks.
+%As an example, a real-world policy size which we examine (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets of \num{8} tilings creates \num{129} work items.
+%Computing a tile hit uses one division per dimension, which is an expensive ALU operation to require.
+%However, if tile widths are fixed at powers-of-two, then these may be replaced with right bitshifts.
 %As we shall see shortly, this aligns with our choice of numerical representation.
@@
 
-\subsection{Representations for embedded ML}
-PDP hardware universally lacks floating-point arithmetic support, yet the task of bringing ML models to resource-constrained environments is well-studied.
-Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
-Lower bit depths reduce memory footprints, and improve throughput in e.g., \emph{bfloat16}~\parencite{bfloat16-blog,DBLP:journals/sigops/XieDMKVZT18}, \emph{hfp8}~\parencite{DBLP:conf/nips/SunCCWVSCZG19}, and other floating point formats~\parencite{DBLP:journals/corr/abs-2007-01530}.
-In many cases, accuracy losses are negligible.
-Such work goes further still towards integers~\parencite{tensorrt-8bit} or binarisation~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16}, sacrificing dynamic range for cheaper arithmetic.
-Similar techniques are used to make inference tractable on network hardware, i.e., to bring pre-trained neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
-
-To make RL workloads feasible under such constraints, we propose the use of quantised, fixed-point representations such as $Qm.n$ (i.e., $m$\si{bit} signed integers with an $n$\si{bit} fractional part), which allow us to evaluate and update policies using only integer arithmetic.
-This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation.
-We note that our later parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is possible only due to the use of $Qm.n$, as atomic operations on floating point numbers typically do not exist outside of graphics-processing hardware.
+%\subsection{Representations for embedded ML}
+%PDP hardware universally lacks floating-point arithmetic support, yet the task of bringing ML models to resource-constrained environments is well-studied.
+%Quantisation and alternative data formats have been suggested to make ML inference feasible on resource- and power-limited platforms, work around hardware constraints, or compute faster and more efficiently.
+%Lower bit depths reduce memory footprints, and improve throughput in \emph{bfloat16}~\parencite{bfloat16-blog,DBLP:journals/sigops/XieDMKVZT18}, \emph{hfp8}~\parencite{DBLP:conf/nips/SunCCWVSCZG19}, and other floating point formats~\parencite{DBLP:journals/corr/abs-2007-01530}.
+%In many cases, accuracy losses are negligible.
+%Such work goes further still towards integers~\parencite{tensorrt-8bit} or binarisation~\parencite{DBLP:journals/corr/MiyashitaLM16,DBLP:conf/eccv/RastegariORF16,DBLP:journals/corr/KimS16,DBLP:conf/nips/HubaraCSEB16}, sacrificing dynamic range for cheaper arithmetic.
+%Similar techniques are used to make inference tractable on network hardware, i.e., to bring pre-trained neural networks to the dataplane~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
+%
+%To make RL workloads feasible under such constraints, we propose the use of quantised, fixed-point representations such as $Qm.n$ (i.e., $m$\si{bit} signed integers with an $n$\si{bit} fractional part), which allow us to evaluate and update policies using only integer arithmetic.
+%This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation.
+%We note that our later parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is possible only due to the use of $Qm.n$, as atomic operations on floating point numbers typically do not exist outside of graphics-processing hardware.
 
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/coop}
-	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to either latency or throughput needs---\Coopfw{} shows this optimally online.
+	\includegraphics[keepaspectratio, width=0.7\linewidth]{figures/coop}
+	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to either latency or throughput needs---\Coopfw{} is the optimal online design.
 A single \emph{controller} delegates RL computation and updates to many \emph{minion} threads, who operate on independent subtasks.\label{fig:single-and-parallel:parallel}}
@@
 
-\Cref{fig:netro-arch,fig:single-and-parallel} outline our design on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
+\Cref{fig:netro-arch} outlines our design on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
 We explain relevant NFP architectural details later in \cref{sec:netronome-platform-fundamentals}.
 \approachshort{} communicates with the packet pipeline of a P4 dataplane via \texttt{extern} plugins using \inring{} (state, configuration) and \outring{} (action) messages (\cref{fig:netro-arch}).
-Internally, \approachshort{} either has all its cores act independently (\cref{fig:single-and-parallel:single}) or cooperate to solve each task (\cref{fig:single-and-parallel:parallel})---with different latency-throughput benefits (\cref{sec:action-and-update-computation}).
+Internally, \approachshort{} either has all its cores act independently, or cooperate to solve each task (\cref{fig:single-and-parallel:parallel})---with different latency-throughput benefits.
 %We open-source our firmware and control programs for the benefit of the community.\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission.}}
@@
 
+\subsection{Challenges, solutions, and insights}\label{sec:challenges-solutions-and-insights}
+\fakepara{Network and execution latency}
+State measurement, policy computation, and action installation take time to carry out, including serialisation to MAT-friendly rule formats~\parencite{DBLP:conf/infocom/TaylorT05}.
+In the midst of these tasks, system state evolves over time---adding noise to the state-action mapping being learned, harming learning and accuracy~\parencite{DBLP:journals/firai/TravnikMSP18}.
+To solve this, we \emph{co-locate all RL agent functions on PDP hardware}.
+
+\fakepara{Lack of floating-point}
+PDP hardware, being designed solely for efficient packet processing, lacks floating-point arithmetic support even in more general purpose NPU SmartNICs.
+Luckily, the embedded ML literature offers many low-precision floating-point formats~\parencite{bfloat16-blog,DBLP:journals/sigops/XieDMKVZT18,DBLP:conf/nips/SunCCWVSCZG19,DBLP:journals/corr/abs-2007-01530} and fixed-point representations without dynamic range~\parencite{DBLP:conf/usenix/ZhangLX00020,DBLP:conf/nsdi/SapioC0NKKKMPR21}.
+The latter set requires only integer arithmetic, and allows us to express and update policy parameters online.
+Thus, we use \emph{quantised fixed-point ($Qm.n$) representations of action values}.
+
+\fakepara{Costs of \emph{online} function approximation}
+Existing ML works on PDP operate by converting a \emph{pre-trained} model into a suitable representation, such as a binary neural network or chain of MATs.
+However, NIC-suitable inference schemes must still be trained offline, even though their data formats eschew floating point.
+Additionally, DNN training relies on backpropagation, can require memorizing a sizeable replay buffer for RL, and needs many mini-batches of data for stable training.
+To solve this, we employ techniques from more classic RL literature: \emph{tile-coded policies}, which have trivial gradients, and \emph{one-step temporal-difference RL algorithms such as Sarsa~\parencite{RL2E}}.
+
+%\fakepara{Heavy compute in the packet path can cause stalls/losses}
+%To solve this, we employ techniques from more classical RL literature: \emph{sarsa}, and \emph{tile-coding}.
+
+\fakepara{Insights}
+We observe that tile coded inference is a map-reduce problem, where each tile hit accesses separate data which are then aggregated together (and updates are can be performed locklessly with no aggregation), enabling us to exploit SmartNICs' parallelism.
+For instance, a policy with $m$ tiling sets (each having a set of input dimensions), with $n$ overlapping tilings then creates $m \times n$ separate tasks.
+What is key is that fixed-point numbers also enable the use of atomic arithmetic, and thus our scheme admits a novel \emph{wait-free Sarsa} RL algorithm.
+Moreover, individual integer operations are cheaper than floating-point, and we may tweak policy memory cost at compile time by choosing the desired integer size.
+%?? Excellent fit for the structure of smartnic hardware.
+%?? In networks, actions and analysis are most likely to be installed on bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts---all of which make in-NIC training a perfect fit.
+%?? This is not only essential in performing this work, but also serves as a mechanism for reducing the processing and memory costs of function approximation. (i.e., can choose depth and int > float)
+
 \subsection{System Model}
@@
 \approachshort{} is agnostic to the meaning of state vectors it receives as inputs and the actions it produces, which are employed by other functional units or the dataplane.
-However, in-NIC/in-network execution specifically benefits packet-, flow-, and network-level learning, control, and optimisation tasks.
+However, in-NIC/in-network execution specifically benefits packet-, flow-, and network-level control tasks.
 
@@
 \begin{LaTeXdescription}
-	\item[\Indfw{} (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each works sequentially.
-Computing an action list requires a \emph{read lock} on the policy.
-If an update occurs, the core requests a \emph{write lock} before updating, limiting online throughput.
-\emph{Tile lists} are stored for update computation.
 	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel,alg:parsa})] Threads cooperate to process state vectors, minimising latency.
@@
 \emph{State vectors} are stored for update computation.
+	\item[\Indfw{} (fig.
+omitted)] Separate threads listen for new states, and each works sequentially.
+Computing an action list requires a \emph{read lock} on the policy.
+If an update occurs, the core requests a \emph{write lock} before updating, limiting online throughput.
+\emph{Tile lists} are stored for update computation.
 \end{LaTeXdescription}
@@
 It transpired that cacheing this data placed a larger \texttt{memcpy} in the serial section, whose size did not scale at all with bit depth.
-Additionally, we do not use bitshifts in place of division operations in our implementation, due to the strict limits that power-of-two tile widths place upon policy design.
+%Additionally, we do not use bitshifts in place of division operations in our implementation, due to the strict limits that power-of-two tile widths place upon policy design.
 
@@
 Packet headers are extracted and parsed using the tooling autogenerated by the P4 pipeline.
-We found that this costs a median \SIrange{126}{140}{\nano\second} communication time (local--remote), comparable to message channels in the Rust and Go languages on commodity hardware.
+We found that this costs a median \SIrange{126}{140}{\nano\second} communication time (local--remote), comparable to message channels in Rust and Go on commodity hardware.
 
@@
 \begin{algorithm}
-	\small
+%	\small
+	\footnotesize
 	\caption{ParSa---\emph{Par}allel \emph{Sa}rsa\label{alg:parsa}}
@@
 	\SetKw{Const}{const}
-	\SetKwProg{parsa}{Function \emph{ParSa}}{}{end}
-	\SetKwProg{control}{Function \emph{Ctl}}{}{end}
-	\SetKwProg{minion}{Function \emph{Minion}}{}{end}
-	\SetKwProg{tilecode}{Function \emph{TileCode}}{}{end}
+	\SetKwProg{parsa}{fn \emph{ParSa}}{}{end}
+	\SetKwProg{control}{fn \emph{Ctl}}{}{end}
+	\SetKwProg{minion}{fn \emph{Minion}}{}{end}
+	\SetKwProg{tilecode}{fn \emph{TileCode}}{}{end}
 
@@
 \subsection{Reconfigurability}
-\approachshort{} allows policy design and learning parameters to be changed at runtime using at most two control packets.
+\approachshort{} allows policy design and parameters to be changed at runtime using at most two control packets.
 Design changes are used at the end of learning (moving from online to offline), or when trying to train a new policy for another task from the same vantage point.
@@
 This extends to policy data, which may be imported from a pre-trained model and exported via PCIe to the host machine.
-Some aspects must be chosen at compile time; bit depth, \Coopfw/\Indfw, and maximum policy sizes---these govern core operation or allocated memory.
-Choosing a bit depth of \qtylist[list-pair-separator = { or }]{16;8}{\bit} halves/quarters policy memory costs, allowing more complex problems to be modelled using more dimensions or fine-grained tiles.
+Some aspects must be chosen at compile time; bit depth, \Coopfw/\Indfw, and maximum policy sizes.
+Choosing a bit depth of \qtylist[list-pair-separator = { or }]{16;8}{\bit} halves/quarters policy memory costs, allowing complex problems to be modelled using more dimensions or fine-grained tiles.
 
 In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved DSCP value~\parencite{rfc2474} as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-While this mainly automates parser generation, it also allows for configuration to be received from only trusted hosts (over the dataplane if needed) via P4 rules.
-Our control packet generation library and evaluation frameworks which build upon it are written in Rust.
+While this automates parser generation, it also allows for configuration to be received from only trusted hosts (over the dataplane if needed) via P4 rules.
+Our control packet generation library and evaluation frameworks are written in Rust.
 
@@
 
-We use a simple first-fit work placement algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core (as the NFP uses hardware threads).
+We use a simple first-fit work placement algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core.
 Each work item is a separate \emph{tiling} over a list of dimensions.
@@
 Naturally, for $n$ tilings and $m$ threads this procedure is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per tiling, storing $m/8$ and $\le8$ costs respectively.
+Although we omit relevant plots for space, we observe that this offers \qtylist{1.33; 1.11}{\texttimes} speedup over na\"{i}ve and stride-modulo schedules.
 
@@
 We additionally use a more powerful consumer-grade machine (\emph{Collector}) for estimating host performance when offloaded to a network function, running Ubuntu 18.04.4 LTS (4.15.0-96-generic).
+Host execution occurs on the CPU using a numpy-based Sarsa implementation.
 Control programs were built using rustc 1.52.1.
@@
 Policy sizes are set to those of a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dim state vectors, a bias tile and 16 full tiling sets (\numproduct{7x1}-dim, \numproduct{8x2}-dim, \numproduct{1x4}-dim), 8 tilings per set, 6 tiles per dimension, and 10 actions.
-For context, such input would contain per-flow state (e.g., IATs, rates) which is combined with the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim).
-In \Coopfw{}, this creates \num{129} work items across \num{31} workers.
+For context, such input would contain per-flow state (e.g., IATs, rates) combined with the last action taken (2-dim tilings) and loads along the ingress-egress path (4-dim).
+In \Coopfw{}, this creates \num{129} tasks across \num{31} workers.
 We choose a larger action count than this work to investigate the performance of more complex agents.
@@
 
-\fakepara{Work allocation}
-%?? Also our simple, ahead-of-time, work scheduler.
-We verify that our heuristic, runtime work scheduler (termed \emph{Balanced}) makes meaningful use of the explicit memory hierarchy and cost of each work item.
-We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ tilings given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
+%\fakepara{Work allocation}
+%%?? Also our simple, ahead-of-time, work scheduler.
+%We verify that our heuristic, runtime work scheduler (termed \emph{Balanced}) makes meaningful use of the explicit memory hierarchy and cost of each work item.
+%We compare it versus several baselines: a \emph{Na\"{i}ve} chunked scheduler, a \emph{Random} allocation, and a simple \emph{Modular} allocator (where a worker $j$ out of $w$ takes $k$ tilings given by $\mathit{work}_j=\left\{\left(j + i \times w\right) \bmod n_{\mathit{items}} \mid i \in [0,k) \right\}$).
 %We use maximum-size, \SI{32}{\bit} policies as described above.
@@
 \fakepara{End-to-end RL latency}
-We compare the key RL latencies we discuss in \cref{fig:state-slip} across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to a \emph{virtual Network Function} (vNF).
+We compare the key RL latencies we discuss in \cref{sec:challenges-solutions-and-insights} across 3 scenarios: completely in-NIC (\approachshort{}), offloading RL decisions to a SmartNIC's controller machine, and offloading to a \emph{virtual Network Function} (vNF).
 
@@
 
-We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, placing an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source via the default firmware.
+We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, placing an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source.
 We perform throughput and loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements (sampling at \SI{2000}{\hertz} for \qtyproduct[product-units=single]{10 x 10}{\second}).
@@
 Crucially, in-NIC execution gives far tighter bounds on tail latency compared to host offloading.
-\num{99.99}\nthscript{th} percentile state-action latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, while host-based tail latencies are at least \SI{40.53}{\percent} greater even when the parallel worker count is at or below the physical core count.
-We show the cumulative distributions of these in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates, scans, or another vNF/traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
+\num{99.99}\nthscript{th} percentile state-action latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, while host tail latencies are at least \SI{40.53}{\percent} greater---even when not oversubscribed.
+We show their cumulative distributions in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates or another traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
 
@@
 \newlength{\resultplotwidth}
-\setlength{\resultplotwidth}{0.849\linewidth}
+%\setlength{\resultplotwidth}{0.849\linewidth}
 %\setlength{\resultplotwidth}{0.9\linewidth}
-%\setlength{\resultplotwidth}{0.8\linewidth}
+\setlength{\resultplotwidth}{0.8\linewidth}
 
@@
 \Cref{tab:tputs} compares \approachshort{}'s throughput against host-based execution.
-We set the worker count on host machines which maximised their throughput.
-This equalled the amount of physical cores on each device---moving beyond this would hamper tail latencies by an order of magnitude.
+We set the worker count on host machines equal to the number of physical cores on each device---moving beyond this would hamper tail latencies by an order of magnitude.
 To make the comparison fair in the context of many-core CPU environments, we include per-core throughput.
@@
 By limiting the available workers in software, we show how \Coopfw{}'s policy update time (thus  online throughput---\cref{fig:vary-core}) scales with available cores.
-While \Coopfw{} always outperforms the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome our own \Indfw{}; \num{8} workers for online throughput, and \num{3} workers for state-action latency (plot omitted).
-Some artefacts of our environment and design choices are visible, such as the addition of new physical cores being more significant than contexts, and the presence of some schedule bottlenecks.
-Most importantly however, \Coopfw{}'s resource demand is tunable at compile time to meet the online training rate or action latency required by a task.
+While \Coopfw{} always outperforms the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome \Indfw{}; \num{8} workers for online throughput, and \num{3} for state-action latency (plot omitted).
+Some artefacts of our environment and design choices are visible; the addition of new physical cores is more significant than contexts, and some schedule bottlenecks are visible.
+Most importantly, \Coopfw{}'s resource demand is tunable at compile time to meet the online training rate or action latency required by a task.
 
@@
 	\caption{
-		\Coopfw{}'s online learning performance improves with additional cores, on max size tasks.
+		\Coopfw{}'s online learning performance improves with additional cores.
 This requires \num{8} workers to offer greater online throughput than single-threaded in-NIC RL.
@@
 
-\Cref{fig:vary-work,fig:vary-work-latency} show how policy complexity affects update cost and state-action latency respectively, scaling from a bias tile up to the full DDoS policy size.\footnote{Input vectors here all have 20 elements regardless of the policy design.}
+\Cref{fig:vary-work} shows how policy complexity affects update cost, scaling from a bias tile up to the full DDoS policy size.\footnote{Input vectors here all have 20 elements regardless of the policy design.}
 \Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning.
@@
 	}
-	\caption{\Coopfw{} processes updates faster than \Indfw{}---thus has higher online performance---on almost all policy sizes. Lower bit depths are effective on simple policies.\label{fig:vary-work}}
+	\caption{\Coopfw{} processes updates faster than \Indfw{}---thus has higher online performance---on almost all policy sizes.
+Lower bit depths are effective on simple policies.
+State-action latency scales similarly.\label{fig:vary-work}}
 \end{figure}
 
-\begin{figure}
-	\centering
-	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
-	}
-	\caption{State-action latency scales with additional work similarly to \cref{fig:vary-work}; though \SI{32}{\bit} firmwares become more effective sooner (at \num{3} work items).\label{fig:vary-work-latency}}
-\end{figure}
+%\begin{figure}
+%	\centering
+%	\resizebox{\resultplotwidth}{!}{
+%		\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
+%	}
+%	\caption{State-action latency scales with additional work similarly to \cref{fig:vary-work}; though \SI{32}{\bit} firmwares become more effective sooner (at \num{3} work items).\label{fig:vary-work-latency}}
+%\end{figure}
 
@@
 	\end{subfigure}
-	\caption{Cumulative state-action latency plots for \approachshort{} and host-based execution.\label{fig:lat-cumul}}
+	\caption{Cumulative state-action latency plots for \approachshort{} versus hosts.\label{fig:lat-cumul}}
 \end{figure}
@@
 %A noteworthy trend is that \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
-\SIlist{8;16}{\bit} versions of \approachshort{} underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency})---even though our implementation is optimised to read and write policy data in batches.
+\SIlist{8;16}{\bit} versions of \approachshort{} underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portion of \cref{fig:vary-work})---even though our implementation is optimised to read and write policy data in batches.
 Because the native register width on the NFP is \SI{32}{\bit}, the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
@@
 Most of the I/O comes \emph{after} this step, causing ALU use to dominate.
-This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (\cref{fig:vary-work-latency}, 3 dims) agents, where hash table accesses and a \texttt{memcpy} of the state vector fall into the serial portion of the online algorithm.
+This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (plot omitted, 3 dims) agents, as state management falls into the serial portion of the online algorithm.
 %Smaller bit depths still give a \SIrange{2}{4}{\texttimes} saving in memory for policy storage (enabling more fine-grained or complex policies), in exchange for slightly worse latency and throughput.
@@
 
-\fakepara{Work allocation}
-\Cref{fig:work-alloc-32} shows that our heuristic allocator (\emph{Balanced}) is key in achieving consistently low latencies/update times.
-The trend is repeated for all bit depths.
-The constant difference between \emph{Action} and \emph{Update Prep} scales linearly with bit depth, matching storage and lookup work in the serial portion of \emph{ParSa} (plots omitted).
-%The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its index, as batching work in contiguous chunks gives some workers only high-dimensionality tilings.
-The minor gap in lower bound performance between the \emph{Random} and \emph{Balanced} allocators suggests that further optimisations can be made.
-We expect that closing or exceeding this gap may require more complex modelling of hardware thread interactions, which lies far beyond the scope of in-NIC scheduling.
-Additionally, throughput occasionally plateaus due to bottlenecks in the output schedule until extra cores are allocated due to the granularity of the above tilings.
-We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
-%Though we have a \SI{4.2}{\texttimes} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
-
-\begin{figure}
-	\centering
-	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
-	}
-	\caption{Action/update compute times in a \SI{32}{\bit} \Coopfw{} agent under different work schedulers.\label{fig:work-alloc-32}}
-\end{figure}
+%\fakepara{Work allocation}
+%\Cref{fig:work-alloc-32} shows that our heuristic allocator (\emph{Balanced}) is key in achieving consistently low latencies/update times.
+%The trend is repeated for all bit depths.
+%The constant difference between \emph{Action} and \emph{Update Prep} scales linearly with bit depth, matching storage and lookup work in the serial portion of \emph{ParSa} (plots omitted).
+%%The severe underperformance of the \emph{Na\"{i}ve} allocator confirms that work item complexity is correlated with its index, as batching work in contiguous chunks gives some workers only high-dimensionality tilings.
+%The minor gap in lower bound performance between the \emph{Random} and \emph{Balanced} allocators suggests that further optimisations can be made.
+%We expect that closing or exceeding this gap may require more complex modelling of hardware thread interactions, which lies far beyond the scope of in-NIC scheduling.
+%Additionally, throughput occasionally plateaus due to bottlenecks in the output schedule until extra cores are allocated due to the granularity of the above tilings.
+%We measured individual work items in state-action computation to take a mean \SIlist{5.2; 6.2; 9.7; 11.0}{\micro\second} for bias, CLS, CTM and IMEM tilings respectively.
+%%Though we have a \SI{4.2}{\texttimes} factor of task oversubscription, it is clear that latencies are eventually bound below by the length of the longest task.
+%
+%\begin{figure}
+%	\centering
+%	\resizebox{\resultplotwidth}{!}{
+%		\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
+%	}
+%	\caption{Action/update compute times in a \SI{32}{\bit} \Coopfw{} agent under different work schedulers.\label{fig:work-alloc-32}}
+%\end{figure}
 
@@
 %?? reference the inference times brought up in Taurus?
-Referring to \cref{fig:state-slip}, we take $t_2$ from \cref{tab:lats} for host and in-NIC processing times, and substitute $t_1+t_3$ for the state packet round-trip time to the decision site:
+We take inference times from \cref{tab:lats} for host and in-NIC processing, adding on the packet round-trip time to the decision site as state-action latency:
 \begin{LaTeXdescription}
 	\item[\emph{In-NIC.}] As described in \cref{sec:agent-environment-communication}, EMEM rings have a median one way delay cross-island of \SI{140}{\nano\second}, giving a \emph{median \SI{34.63}{\micro\second} end-to-end inference latency}.
-	\item[\emph{Dedicated Collector.}] Employing DPDK to maximise performance, such hosts will add one-way PCIe delays of \SIrange{0.9}{2.3}{\micro\second} for network packets~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
+	\item[\emph{Dedicated Collector.}] Employing DPDK to maximise performance, such hosts add one-way PCIe packet delays of \SIrange{0.9}{2.3}{\micro\second}~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
 	A UDP packet carrying \num{20} elements of state in \approachshort{} is \SI{128}{\byte}, so costs \SI{1}{\micro\second} to forward, and the reply state-action pair invokes a slightly higher cost.
@@
 
-An anomaly is \SI{128}{\byte} packets at \num{3000} RL updates per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase, shown in \cref{fig:dataplane-example}.
+An anomaly is \SI{128}{\byte} packets at \num{3000} RL updates per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase in \cref{fig:dataplane-example}.
 %This is observed through a shift of some packet latencies from the mode towards the tail, but no other changes in the distribution.
-We believe that the inbound request rate is weakly synchronised with traffic, causing higher burstiness of accesses to the CPP bus.
+We believe that the inbound request rate is weakly synchronised with traffic, causing burstiness of accesses to the CPP bus.
 We expect that FPGA designs can avoid this problem by having dedicated \inring{}/\outring{} access mechanisms for an \approachshort{} agent.
@@
 		}
-		\caption{Distribution of RTTs for \SI{128}{\byte} packets for \numlist{0;3000} RL actions/s.\label{fig:dataplane-example}}
+		\caption{RTT distribution for \SI{128}{\byte} packets at \numlist{0;3000} RL actions/s.\label{fig:dataplane-example}}
 	\end{subfigure}
@@
 \caption{NFP memory use due to \approachshort{} using \numlist{1;4} MEs (\SI{32}{\bit}).
-CLS and CTM are shared between all programs on the same island, while EMEM and IMEM are shared between all NFP programs on a NIC.\label{tab:resources}}
+CLS and CTM are shared between all programs on the same island, EMEM and IMEM are shared between all programs on a NIC.\label{tab:resources}}
 \resizebox{\linewidth}{!}{
@@
 \fakepara{Deployability}
-Setup of \approachshort{} uses two packet types: \emph{setup}, which contains learning (hyper-)parameters and most aspects of a policy, and \emph{tiling}, which provides a list of state indices for tiling sets.
-We found that setup and tiling packets take a mean \qtylist{27.03;16.69}{\micro\second} to be processed on \Indfw{}, allowing an agent to be swapped between online and offline operation (or repurposed) painlessly.
+Setup of \approachshort{} uses two packet types: \emph{setup}, which contains policy shape and learning parameters, and \emph{tiling}, which provides a list of state indices for tiling sets.
+These packets took a mean \qtylist{27.03;16.69}{\micro\second} respectively to be processed on \Indfw{}, allowing an agent to be swapped between online and offline operation (or repurposed) painlessly.
 These swaps are useful when an agent should cease learning (i.e., when performance converges), or if a change in the environment suggests that more training is needed.
@@
 Online-offline swaps for \Coopfw{} exhibit similar cost, however the need for explicit scheduling means that policy/tiling \emph{structure} changes (including the first complete setup) take \SI{422.63}{\micro\second} for the full-size policy described above.
-The time taken for \Coopfw{} to schedule its tasks was found to scale with the number of workers ($m$) and work items ($n$) as described earlier (\cref{sec:work-allocation}, $\mathcal{O}{\left(n\log{m}\right)}$).
+The time taken for \Coopfw{} to schedule its tasks was found to scale with the number of workers ($m$) and work items ($n$) as described earlier ($\mathcal{O}{\left(n\log{m}\right)}$).
 Ignoring the trivial solutions, reducing the worker count to \num{1} costs a mean \SI{238.22}{\micro\second}, while placing a single task incurs \SI{53.79}{\micro\second}.
@@
 %Reducing state-action latency (i.e., with \Coopfw) is useful for minimising the noise inherent in learning.
-Agents here are limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
+Agents are limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
 Accordingly, this use case benefits most from an increase in \emph{throughput} using \Indfw{}.
-Higher throughput means that network flows are more likely to be judged in every timestep, even when flow cardinality is high---making it more likely that changes in flow behaviour will be observed, acted on, and learned from.
+Higher throughput means that network flows are more likely to be judged in every timestep, even at high flow cardinality---making it more likely that changes in flow behaviour will be acted on and learned from.
 We note that the TRS scheduler is designed to handle large numbers of attack flows, combining new state vectors while the asynchronous agent is itself busy.
@@
 %No works investigate the runtime training of BNN function approximators in-NIC.
-
 \textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantisation on the NFP, using \emph{in-path} compute rather than our asynchronous model.
@@
 
-We stress that none of these approaches (or that we have seen) tackle the issue of \emph{online learning and control} in-network---we believe \approachshort{} has broken new ground in this regard.
+We stress that none of these approaches (or that we have seen) tackle \emph{online learning and control} in-network---we believe \approachshort{} has broken new ground in this regard.
 
@@
 We have presented \emph{\approachshort{}}, bringing \emph{online reinforcement learning} to the dataplane.
-\approachshort{} has shown how classical RL techniques make online learning possible by simplifying update logic and enabling parallel processing.
-In-NIC use of these algorithms enabled a \SIrange{15}{21}{\texttimes} reduction in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput compared to host offloading.
-The deployment environment and asynchronous design were shown to eliminate PCIe delays and impose minimal impact on carried dataplane traffic.
-We also showed how \approachshort{} scales with additional compute resources at deployment to improve on decision latency and throughput.
-Our throughput-optimal design, \Indfw{}, improves upon these metrics with \emph{just one worker}.
-
+%\approachshort{} has shown how classical RL techniques make online learning possible by simplifying update logic and enabling parallel processing.
+In-NIC use of these classical algorithms makes this possible, and enables significant reductions in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput, with minimal impact on dataplane cross-traffic.
+%The deployment environment and asynchronous design were shown to eliminate PCIe delays and impose minimal impact on carried dataplane traffic.
+%We also showed how \approachshort{} scales with additional compute resources at deployment to improve on decision latency and throughput.
+%Our throughput-optimal design, \Indfw{}, improves upon these metrics with \emph{just one worker}.
 In future, we aim to examine the performance of individual applications driven by \approachshort---both classical and deep RL-based---and how a NetFPGA implementation can offer further latency and throughput improvements.

### 2021-10-14 code/rl-perf-tester/src/parsa/mod.rs

@@ -5,14 +5,7 @@ mod writeback;
 pub use self::{q::*, tilecode::*, writeback::*};
 
 use crate::{
-	fw_consts::{
-		T1_FIRST_TILE,
-		T1_TILE_COUNT,
-		T2_FIRST_TILE,
-		T2_TILE_COUNT,
-		T3_FIRST_TILE,
-		T3_TILE_COUNT,
-	},
+	fw_consts::T1_FIRST_TILE,
 	BlockDataSource,
 	ProtoSetup,
 	TimeBreakdown,

### 2021-10-18 paper/noms-short-paper.tex
Terrible, terrible things.

@@
 \begin{abstract}
-Automatic, data-driven networking is becoming more capable and widely researched, partly driven by the efficacy of \emph{Deep Reinforcement Learning} (DRL) algorithms.
-Yet the complexity of both DRL inference and learning force these tasks to be offloaded \emph{away from the dataplane} to hosts, harming latency-sensitive applications.
+Data-driven networking is becoming more capable and widely researched, partly driven by the efficacy of \emph{Deep Reinforcement Learning} (DRL) algorithms.
+Yet the complexity of both DRL inference and learning force these tasks to be pushed \emph{away from the dataplane} to hosts, harming latency-sensitive applications.
 Online learning of such policies cannot occur in the dataplane, despite being useful techniques when problems evolve or are hard to model.
@@
 Our design allows weak yet highly parallel SmartNIC NPUs to be competitive against commodity x86 hosts, despite having fewer features and slower cores.
-Compared to offloading, we achieve a \SI{21}{\texttimes} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{\texttimes} improvement in online throughput for real-world policy designs.
+Compared to hosts, we achieve a \SI{21}{\texttimes} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{\texttimes} improvement in online throughput for real-world policy designs.
 In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
-\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, and are quickly reconfigurable at runtime compared to reinstalling device firmware.
+\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, and is quickly reconfigurable at runtime compared to reinstalling device firmware.
 \end{abstract}
@@
 SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible alongside P4 dataplanes.
-Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
+Classical RL policy methods are the key to making this feasible.\label{fig:netro-arch}}
 \end{figure}
@@
 %Consider a DDoS mitigation system which could react to dataplane-specific, per-packet state in microseconds, and \emph{start learning again just as quickly} in the face of new data or behaviours.
-The P4 ecosystem already presents novel, openly-available, fine-grained traffic measurement techniques that can be installed and controlled with ease~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
-In addition, P4's control plane makes it easy to select which flows or packets are monitored in a live network and potentially allow control over traffic at the decision site.
-As a result, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome,DBLP:journals/corr/abs-2002-08987} to take advantage of flow or per-packet state that cannot be efficiently processed or extracted at any other location in the network.
+The P4 ecosystem already presents novel, openly-available, and fine-grained traffic measurement techniques~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}, and its control plane makes it easy to select which flows or packets are monitored in a live network.
+As a result, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome,DBLP:journals/corr/abs-2002-08987} to take advantage of flow or per-packet state that cannot be efficiently processed or extracted at any other location.
 These works have shown the value of in-network ML: high-throughput, low latency response to network changes.
 While they can exploit on-device state to provide reactive insight, the missing piece of the puzzle is learning and updating these ML analyses online without deferring to another machine in the network.
-Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed by the community}.
+Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed}.
 
 It is important to make this feasible; offloading to commodity hosts adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
-For context, deep neural network training relies on backpropagation, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many minibatches for stability.
-These lead to high compute and storage costs, inducing capital expenditure for dedicated accelerators to overcome the batching needed to operate at line-rate---while BrainWave~\parencite{DBLP:conf/isca/FowersOPMLLAHAG18} can reduce batching (and thus tail latencies) by \qty{32}{\texttimes}~\parencite{Duarte2019}, inference still takes $\mathcal{O}{\left(\text{\si{\milli\second}}\right)}$.
+For context, deep neural network training relies on backpropagation, can take vast amounts of offline simulation~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many minibatches for stability.
+These induce high costs for compute, storage, or dedicated accelerators to overcome the batching needed to operate at line-rate---while BrainWave~\parencite{DBLP:conf/isca/FowersOPMLLAHAG18} can reduce batching (and thus tail latencies) by \qty{32}{\texttimes}, inference still takes $\mathcal{O}{\left(\text{\si{\milli\second}}\right)}$~\parencite{Duarte2019}.
 Moreover, novel DMA techniques such as \emph{GPUDirect}~\parencite{gpudirect} halve but do not eliminate such PCIe transfers.
-The high latencies \emph{caused} by offloading and inference harm learning~\parencite{DBLP:journals/firai/TravnikMSP18} and runtime application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
-If we can bring online learning \emph{to the dataplane}, then we can take advantage of rich, local state while minimising these latencies (and their impact on the learned policy).
+The high latencies \emph{caused} by steering and inference harm learning~\parencite{DBLP:journals/firai/TravnikMSP18} and runtime application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
+If we can bring online learning \emph{to the dataplane}, then we can take advantage of rich, local state while minimising this impact on the learned policy.
 This would also make it easy to train and prototype agent designs which can learn \emph{as the environment evolves}, or when there is too little data to model and simulate a problem.
@@
 In this paper, we enable \emph{online in-NIC learning} by returning to \emph{classical} RL methods.
-In particular, we focus on tile coding with one-step temporal-difference algorithms such as Sarsa~\parencite{RL2E}.
+In particular, we focus on tile coding with algorithms such as Sarsa~\parencite{RL2E}.
 %These choices have important benefits for in-NIC execution.
@@
 In addition, they have shown promise in other network use cases~\parencite{DBLP:journals/eaai/MalialisK15,DBLP:journals/tnsm/SimpsonRP20}.
-Tile coding in particular admits many optimisations, being an embarrassingly parallel problem.
+%Tile coding in particular admits many optimisations, being an embarrassingly parallel problem.
 Using fixed-point arithmetic, we solve the lack of floating-point support in PDP hardware \emph{and} enable new optimisations.
@@
 %Finally, the choice of single-step algorithms (as opposed to $n$-step or Monte Carlo methods) bounds the amount of per-trace state required for online learning to just the last state-action pair, safeguarding the limited memory of the target devices.
-We also design our solution to operate as closely as possible to the P4 pipeline to use and learn from per-packet state, but outside of the main packet path to prevent packet stalls.
+We also design our solution to operate close to the P4 pipeline to access per-packet state, but outside of the main packet path to prevent packet stalls.
 
@@
 \begin{itemize}
-	\item An analysis of why in-NIC RL is needed and best-placed to interact with the network, made feasible by classical RL methods and quantisation (\cref{sec:motivation}),
-	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands of network traffic control (\cref{sec:design}),
+%	\item An analysis of why in-NIC RL is needed and best-placed to interact with the network, made feasible by classical RL methods and quantisation (\cref{sec:motivation}),
+	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands of network control (\cref{sec:design}),
 	\item \emph{ParSa}, a wait-free, parallel RL algorithm to accelerate tile coded policy inference and updates (\cref{alg:parsa}),
-	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{\texttimes} latency reduction compared to commodity hardware (\SI{21}{\texttimes} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
-	\item A description of how \approachshort{} would integrate with state-of-the-art PDP applications to perform fully in-NIC, fast, automated DDoS mitigation (\cref{sec:potential-integrations}).
+	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes, and improves on host deployment with a \SI{15}{\texttimes} latency reduction compared to commodity hardware (\SI{21}{\texttimes} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
+	\item A description of how \approachshort{} would integrate with state-of-the-art PDP applications to perform fully in-NIC, automated DDoS mitigation (\cref{sec:potential-integrations}).
 \end{itemize}
@@
 %\section{Preliminaries: RL for in-network computation}\label{sec:motivation}
-By combining recent trends in PDP hardware with insights from the ML community (past and present), we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL and quantisation make this feasible.
-Crucially, we focus on low port-density devices.
+%By combining recent trends in PDP hardware with insights from the ML community, we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL and quantisation make this feasible.
+%Crucially, we focus on low port-density devices.
+%These devices' designs make it reasonable to move processing \emph{outside of the packet pipeline}; SmartNICs expose additional cores, and NetFPGAs allow for the synthesis of independent functional units.
+%The main benefit is that RL logic can be moved onto the NIC \emph{without impacting packet processing}.
+%The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
+%The main benefit is that RL logic can be moved onto the NIC \emph{without impacting packet processing}.
 %We envision that this processing is suited to the network edge or key vantage points, rather than core switches.
-These devices' designs make it reasonable to move processing \emph{outside of the packet pipeline}; SmartNICs expose additional cores, and NetFPGAs allow for the synthesis of independent functional units.
-%The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
-The main benefit is that RL logic can be moved onto the NIC \emph{without impacting packet processing}.
 %?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
@@
 %In-pipeline execution thus increases the risk of packet drops or stalls.
-
 %However, this requires that there be a reasonably consistent interface and behaviour between device classes.
@@
 %The PSA presents a sensible lower bound on the device capabilities required to implement a P4 dataplane, but the reality is more complex and interesting.
-
 %?? TODO: trim down. Too ``thesis-y'', rather than ``conference-y''.
 %?? Save this elsewhere 'til then?
-
 \subsection{Programmable hardware capabilities}
@@
 
-While P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to great interest in network programmability, it requires similar behaviour between device classes.
-This is encoded by the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}---a lower bound on the device capabilities required to implement a P4 dataplane.
+While P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to great interest in network programmability, it requires similar behaviour between device classes---this is encoded by the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}.
+%---a lower bound on the device capabilities required to implement a P4 dataplane.
 %We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
 However, many compatible devices long predate these developments. 
-Many-core SoC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
+Many-core SoC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow arbitrary programs to be specified and executed.
 %This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
-Currently, low port-density devices like these are most likely to have specific general-purpose compute and high degrees of parallelism~\parencite{DBLP:conf/sigcomm/LiuCSKPG19}, as they are designed for high-performance offloading and middlebox development.
+Currently, low port-density devices like these are most likely to have general-purpose compute and high degrees of parallelism~\parencite{DBLP:conf/sigcomm/LiuCSKPG19}, as they are designed for high-performance offloading and middlebox development.
 %Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
@@
 
-\emph{Reinforcement learning} (RL) trains an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task~\parencite{RL2E}.
-Like most ML methods, RL algorithms use gradients to update the parameters used to approximate a function with \emph{reward measurements} from the environment.
-RL's Markov Decision Process structure allows online learning of a state-action map in a model-free way, and can step through local minima when needed compared to ML.
+\emph{Reinforcement learning} (RL) trains an \emph{agent} to choose an optimal sequence of \emph{actions} from any \emph{state} in pursuit of a given task~\parencite{RL2E}.
+Like most ML methods RL uses gradients to update the parameters used to approximate a function, aided by \emph{reward measurements} from the environment.
+RL's MDP structure allows online learning of a state-action map in a model-free way, and can step through local minima when needed compared to ML.
 In networking, this allows for learning from on-device state or handling rapidly evolving problems.
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.7\linewidth]{figures/coop}
-	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to either latency or throughput needs---\Coopfw{} is the optimal online design.
-A single \emph{controller} delegates RL computation and updates to many \emph{minion} threads, who operate on independent subtasks.\label{fig:single-and-parallel:parallel}}
+	\includegraphics[keepaspectratio, width=0.65\linewidth]{figures/coop}
+	\caption{\approachshort{} scales to fit device capacity according to either latency or throughput needs---\Coopfw{} is the optimal online design.
+A single \emph{controller} delegates RL inference and updates to many \emph{minions}, who complete independent subtasks.\label{fig:single-and-parallel:parallel}}
 \end{figure}
 
-\Cref{fig:netro-arch} outlines our design on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec \emph{can and should be used} to drive asynchronous environmental control.
+\Cref{fig:netro-arch} outlines our design on Netronome SmartNIC hardware in pursuit of this goal: unused device resources beyond the P4-PSA spec are used to drive asynchronous environmental control.
 We explain relevant NFP architectural details later in \cref{sec:netronome-platform-fundamentals}.
@@
 %We open-source our firmware and control programs for the benefit of the community.\footnote{\textanon{\url{https://github.com/FelixMcFelix/pdp-rl-paper}}{Link redacted for double-blind submission.}}
-We open-source our firmware and control programs for the benefit of the community.\footnote{Link redacted during submission.}
+We open-source our firmware and control programs for the benefit of the community~\parencite{pdp-rl-github}.
 %Moving beyond the Netronome platform, we describe how our architecture may be adapted and improved on by bespoke hardware or FPGA-based deployment.
@@
 \fakepara{Network and execution latency}
-State measurement, policy computation, and action installation take time to carry out, including serialisation to MAT-friendly rule formats~\parencite{DBLP:conf/infocom/TaylorT05}.
+State measurement, policy inference, and action installation take time, including network latency and serialisation to MAT-friendly formats~\parencite{DBLP:conf/infocom/TaylorT05}.
 In the midst of these tasks, system state evolves over time---adding noise to the state-action mapping being learned, harming learning and accuracy~\parencite{DBLP:journals/firai/TravnikMSP18}.
@@
 PDP hardware, being designed solely for efficient packet processing, lacks floating-point arithmetic support even in more general purpose NPU SmartNICs.
-Luckily, the embedded ML literature offers many low-precision floating-point formats~\parencite{bfloat16-blog,DBLP:journals/sigops/XieDMKVZT18,DBLP:conf/nips/SunCCWVSCZG19,DBLP:journals/corr/abs-2007-01530} and fixed-point representations without dynamic range~\parencite{DBLP:conf/usenix/ZhangLX00020,DBLP:conf/nsdi/SapioC0NKKKMPR21}.
-The latter set requires only integer arithmetic, and allows us to express and update policy parameters online.
+Luckily, the embedded ML literature offers many low-precision floating-point formats~\parencite{bfloat16-blog,DBLP:journals/sigops/XieDMKVZT18,DBLP:conf/nips/SunCCWVSCZG19,DBLP:journals/corr/abs-2007-01530} and fixed-point representations~\parencite{DBLP:conf/usenix/ZhangLX00020,DBLP:conf/nsdi/SapioC0NKKKMPR21}.
+The latter set requires only integer arithmetic, which allows us to express and update policy parameters in-NIC.
 Thus, we use \emph{quantised fixed-point ($Qm.n$) representations of action values}.
@@
 \fakepara{Costs of \emph{online} function approximation}
-Existing ML works on PDP operate by converting a \emph{pre-trained} model into a suitable representation, such as a binary neural network or chain of MATs.
-However, NIC-suitable inference schemes must still be trained offline, even though their data formats eschew floating point.
+NIC-suitable inference schemes must still be trained offline, even though their data formats eschew floating-point---existing PDP ML converts a \emph{pre-trained} model into a suitable representation, such as a binary neural network or chain of MATs.
 Additionally, DNN training relies on backpropagation, can require memorizing a sizeable replay buffer for RL, and needs many mini-batches of data for stable training.
@@
 \fakepara{Insights}
-We observe that tile coded inference is a map-reduce problem, where each tile hit accesses separate data which are then aggregated together (and updates are can be performed locklessly with no aggregation), enabling us to exploit SmartNICs' parallelism.
+We observe that tile coded inference is a map-reduce problem, where each tile hit accesses separate data which are then aggregated together, enabling us to exploit SmartNICs' parallelism.
+Crucially, updates can be performed locklessly without aggregation.
 For instance, a policy with $m$ tiling sets (each having a set of input dimensions), with $n$ overlapping tilings then creates $m \times n$ separate tasks.
@@
 \subsection{System Model}
-\approachshort{} is a general, task-independent framework for in-network, online training and execution of \emph{any RL agent design} using classical methods.
+\approachshort{} is a task-independent framework for in-network, online training and execution of \emph{any RL agent design} using classical methods.
 \approachshort{} is agnostic to the meaning of state vectors it receives as inputs and the actions it produces, which are employed by other functional units or the dataplane.
@@
 \approachshort{} runs on one or more cores of a SmartNIC to convert fixed-point state measurements from the environment into a stream of actions using a stored policy.
-As an example, this might be to map flow state and performance measurements into a queue priority for future packets, or to compute and apply a rate limit to preserve quality of service.
-These dedicated cores are responsible for processing requests, computing actions, and updating the underlying policy in real time.
-Combined with reward measurements, this policy can then be updated or trained from scratch entirely on the NIC, acting as a fully online RL agent.
-An input state vector \emph{always} induces an action and, if desired, updates the policy using either an included reward or one retrieved from memory according to a (configurable) key placed alongside the state.
-This allows for simultaneous control and learning over independent systems by the same agent (i.e., optimising several flows with their own reward measures, such as DDoS mitigation in an AS where each next-hop AS might have their own `health' metric).
-
+As an example, this might map flow performance measurements into queue priorities, or to compute and apply a rate limit to preserve quality of service.
+These dedicated cores process requests, compute actions, and update the policy in real time using reward measures.
+This policy can be trained from scratch entirely on the NIC, acting as a fully online RL agent.
+An input state vector \emph{always} induces an action, and may update the policy using either an included reward, or one retrieved from memory based on the input state.
+This allows for simultaneous control and learning of independent systems by the same agent (i.e., optimising several flows with their own reward measures, such as DDoS mitigation in an AS where each next-hop AS might have their own `health' metric).
 %However, high-speed data networks impose inviolable per-packet deadlines.
-
 To protect traffic throughput and allow effective deployment in as many environments as possible, \approachshort{} places RL execution on-chip, \emph{but off the main packet path}, communicating and running parallel to the main P4 dataplane.
 As shown in \cref{fig:netro-arch}, this asynchrony allows coexistence with P4 programs, and imposes minimal impact on carried traffic for both bump-in-the-wire deployments and at end-points.
-For instance, in the default deployment of a P4 packet processing pipeline on Netronome NFP chips several cores go unused (as does spare area on an FPGA design), making this paradigm possible.
+For instance, the default P4 pipeline on Netronome has several cores go unused (similarly to spare area on an FPGA design), making this paradigm possible.
 
@@
 \subsection{Action and Update Computation}\label{sec:action-and-update-computation}
-\approachshort{} applies the insights of \textcite{DBLP:journals/firai/TravnikMSP18} to minimise action latency; an action is computed, sent out into the environment, and only then is the underlying policy updated.
-Using one of the below strategies chosen at compile time, a state vector is tile coded, converted into action probabilities, and an action is chosen.
+\approachshort{} applies the insights of \textcite{DBLP:journals/firai/TravnikMSP18} to minimise action latency; an action is computed, sent out into the environment, and only then is the policy updated.
+Using one of the below strategies, a state vector is tile coded, converted into action probabilities, and an action is chosen.
 This is then written out to the environment as in \cref{sec:agent-environment-communication}.
-If online learning is enabled, \approachshort{} then checks an internal hashmap for a previous state-action pair matching the current instruction source, and if found then the policy is updated.
+If online learning is enabled, \approachshort{} then checks an internal hashmap for a prior state-action pair matching the current trace, and if this is found then the policy is updated.
 Updates are computed using \emph{single-step semi-gradient Sarsa}~\cite[pp. \numrange{217}{221}]{RL2E}, though modification to support other single-step methods would be trivial.
 The new state-action pair is then written into storage.
-\approachshort{} can be configured to automatically select values of the input state vector as keys for state and reward storage.
-Two firmware models govern how these tasks (action selection, policy updates) are carried out:
+\approachshort{} can be configured to select values of the input state vector as keys for state and reward storage.
+Two firmware models govern how these tasks are carried out:
 \begin{LaTeXdescription}
@@
 \end{LaTeXdescription}
-Each offers a different point of optimisation; if updates are disabled, then the \indfw{} model can maximise throughput, while the \coopfw{} model is designed to minimise decision latency and needs no locks to update the policy (increasing \emph{online learning} throughput).
-These correspond to only executing a trained policy and actively (re-)training a policy, respectively.
+Each offers a different point of optimisation; \indfw{} maximises throughput if updates are disabled, while \coopfw{} is designed to minimise decision latency and needs no locks to update the policy (increasing \emph{online learning} throughput).
+These correspond to raw inference and learning performance, respectively.
 %ParSa is described here at a high level, as an algorithm suited for \emph{any multiprocessor environment}.
@@
 %In general it is also possible for the \emph{Ctl} task to act as an additional \emph{Minion} in its parallel sections; we were limited here by code store requirements on the NFP.
-Latency and throughput have different effects on RL agents according to their design and target problem.
-Higher RL throughput is a necessity for per-flow or per-packet applications, which can require high decision-per-second rates even after combining state measurements received from the environment, such as flow control in DDoS prevention.
-Equally, lower latency affords an agent finer-grained control and learning of a problem, being able to react sooner to new information (e.g., device state in a routing optimisation problem).
-
-Paradoxically, we found that in the general \emph{ParSa} algorithm it was more efficient \emph{to do more work} by having each worker recompute its tile subset from a stored state.
-It transpired that cacheing this data placed a larger \texttt{memcpy} in the serial section, whose size did not scale at all with bit depth.
+Latency and throughput have different effects on RL agents according to their target problem.
+Higher RL throughput is a necessity for per-flow/packet applications, which can require high decision rates even after combining state measurements, such as flow control or DDoS prevention.
+Equally, lower latency affords an agent finer-grained control and learning of a problem, able to react sooner to new information (e.g., in a routing optimisation problem).
+
+%Paradoxically, we found that in the general \emph{ParSa} algorithm it was more efficient \emph{to do more work} by having each worker recompute its tile subset from a stored state.
+%It transpired that cacheing this data placed a larger \texttt{memcpy} in the serial section, whose size did not scale at all with bit depth.
 %Additionally, we do not use bitshifts in place of division operations in our implementation, due to the strict limits that power-of-two tile widths place upon policy design.
@@
 \subsection{Agent-Environment Communication}\label{sec:agent-environment-communication}
-\approachshort{} uses \emph{multiple-producer/multiple-consumer} (MPMC) channels to communicate with other elements; be they P4 programs on the packet path, or other on-chip analysis and control modules.
+\approachshort{} uses \emph{multiple-producer/multiple-consumer} (MPMC) channels to communicate with other elements on the NIC; be they P4 programs on the packet path, or other on-chip analysis modules.
 Through these, a system \emph{pushes} state vectors, reward measures, and setup packets as inputs, and \emph{pulls} a stream of state-action pairs as outputs.
 This allows decisions to be made asynchronously, preventing packet stalling.
-The key insight of this mechanism is that on-chip reward or state signals enjoy first-class support in the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-chip processes or the controller.
-As such, \approachshort{} can receive input from P4 \texttt{extern}s or other, dedicated off-path flow state measurement applications.
-
-Our implementation uses platform-specific IPC (\emph{EMEM ring buffers}) with hardware signalling on work arrival to achieve this.
-As PDP hardware lacks dynamic memory allocation, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state or policy data handover.
-Packet headers are extracted and parsed using the tooling autogenerated by the P4 pipeline.
-We found that this costs a median \SIrange{126}{140}{\nano\second} communication time (local--remote), comparable to message channels in Rust and Go on commodity hardware.
+%The key insight of this mechanism is that on-chip reward or state signals enjoy first-class support in the same manner as packets from the P4 dataplane, allowing agents to act on environmental signals from other on-chip processes or the controller.
+As such, \approachshort{} can receive input from P4 \texttt{extern}s or other, dedicated off-path flow measurement applications in the same manner.
+We use platform-specific IPC (EMEM ring buffers) to achieve this, using a shared freelist of buffers for packet payloads.
+%As PDP hardware lacks dynamic memory allocation, P4 pipeline threads request buffers for packet payloads using a shared freelist to enable state or policy data handover.
+%Packet headers are extracted and parsed using the tooling autogenerated by the P4 pipeline.
+Each message takes a median \SIrange{126}{140}{\nano\second} communication time (local--remote), comparable to message channels in Rust and Go on commodity hardware.
 
@@
 This is truer still when moving from the moderately fine-grained control of classical RL ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
-Marshalling of requests and shared data access incur significant overheads, but on-chip execution and the ParSa algorithm allow us to sidestep these via lockless atomic aggregation.
+Marshalling and data mutexes incur significant overheads, but on-chip execution and the ParSa algorithm allow us to sidestep these via lockless atomic aggregation.
 Moreover, adjacent cores often have special-purpose shared registers or share a small fast cache to accelerate communication.
-
 %?? reduce below and make take-homes more generic if possible
-
 Our implementation exploits the locality of cores in the NFP. 
-Policy compute or update tasks are passed between cores using direct \emph{next neighbour} registers, signalling all child threads in response.
+Policy tasks are passed between cores using direct \emph{next neighbour} registers, signalling all child threads in response.
 Such on-chip signals cost just $\sim$\SI{20}{\nano\second} per relayed message.
 This can be factored into the design of additional async off-path functional units (similar to \textcite{DBLP:conf/isca/LiLYCSH19}) on platforms like NetFPGA.
-Each core performs atomic adds to a shared preference list and an acknowledgement counter checked by the master thread, implementing our wait-free \emph{ParSa} algorithm (\cref{alg:parsa}).
-This is essential for aggregation compared to the use of bounded message buffers, which caused significant head-of-line blocking.
+To aggregate, each core performs atomic adds to a shared preference list and an acknowledgement counter checked by the master thread, implementing our wait-free \emph{ParSa} algorithm (\cref{alg:parsa}).
+%This is essential for aggregation compared to the use of bounded message buffers, which caused significant head-of-line blocking.
 %For instance, additional off-path functional units can serve as workers, like the floating-point adders used by \textcite{DBLP:conf/isca/LiLYCSH19}.
@@
 Choosing a bit depth of \qtylist[list-pair-separator = { or }]{16;8}{\bit} halves/quarters policy memory costs, allowing complex problems to be modelled using more dimensions or fine-grained tiles.
-
-In our implementation, configuration packets are carried over UDP and signalled to the P4 parser using a reserved DSCP value~\parencite{rfc2474} as used by \textcite{DBLP:conf/isca/LiLYCSH19}.
-While this automates parser generation, it also allows for configuration to be received from only trusted hosts (over the dataplane if needed) via P4 rules.
-Our control packet generation library and evaluation frameworks are written in Rust.
+In our implementation, configuration packets are carried over UDP and signalled to the dataplane using a reserved DSCP value~\parencite{rfc2474,DBLP:conf/isca/LiLYCSH19}.
+While this simplifies parser generation, it also allows for configuration to be received from only trusted hosts (over the dataplane if needed) via P4 rules.
+Our control library and evaluation frameworks are written in Rust.
 
@@
 Tiling space requirements then scale exponentially with dimension count, so higher-dimension tilings must be placed in larger, slower memory regions.
-As a result, in many architectures policy parameters are split across such regions, giving different access and compute costs to different tasks.
-
-We use a simple first-fit work placement algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core.
+As a result policy parameters are split across such regions, giving different access and compute costs to different tasks.
+We use a simple first-fit scheduling algorithm run in \approachshort{}, placing the largest work item into the least loaded thread of the least loaded core.
 Each work item is a separate \emph{tiling} over a list of dimensions.
@@
 Direct rule installation into P4 tables from the SmartNIC is not generally possible.
-To achieve line-rate performance, platforms like NFP use accelerated datastructures (e.g., DCFL~\parencite{DBLP:conf/infocom/TaylorT05}) which must be computed over the \emph{entire rule set}.
+To achieve line-rate performance, platforms like NFP use accelerated datastructures (e.g., DCFL~\parencite{DBLP:conf/infocom/TaylorT05}) computed over the \emph{entire rule set}.
 Even through \texttt{extern}s, directly adding new rules is neither feasible nor safe.
-We instead suggest on NFP chips that \texttt{extern}s or datapath stages which apply RL actions to packets should maintain a small store of state-action pairs, and periodically send these back to the controller for batch installation.
+We instead suggest that \texttt{extern}s or datapath stages which apply RL actions to packets should maintain a small store of state-action pairs, and periodically send these back to the controller for batch installation.
 %This ensures that the majority of installed rules exhibit accelerated performance, while preventing installation delay on the newest decisions.
 %Platforms such as Intel Tofino greatly simplify this, with Tofino Native Architecture intrinsics such as \emph{Action Profiles/Selects} allowing a P4 action to be chosen based on a register value (i.e., an RL action).
-
 %As is expected of parallel algorithms, efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
 %Efficient division of work, communication, and aggregation add per-task overheads in both the serial and parallel portions.
-Parallelisation adds per-task overheads, which require a minimum number of workers to improve on a serial approach.
-We investigate the exact worker count requirements imposed to overcome single-threaded execution in \cref{sec:results}.
+Parallelisation also adds per-task overheads which require a minimum number of workers to improve on a serial approach---we measure this crossover point in \cref{sec:results}
+%We investigate the exact worker count requirements imposed to overcome single-threaded execution in \cref{sec:results}.
 
@@
 \section{Evaluation}\label{sec:evaluation}
-We investigate the performance of \approachshort{} compared to classical RL techniques executed on commodity host machines, with \Coopfw{} offering a \SIrange{15}{21}{\texttimes} speedup in median--\num{99.99}\nthscript{th} state-action latency and \SI{9.9}{\texttimes} greater online learning throughput.
+We investigate the performance of \approachshort{} compared to classical RL techniques executed on commodity hosts, with \Coopfw{} offering a \SIrange{15}{21}{\texttimes} speedup in median--\num{99.99}\nthscript{th} state-action latency and \SI{9.9}{\texttimes} greater online learning throughput.
 Crucially, in-NIC execution offers tight tail latency bounds compared to host-based approaches.
-We report on how \approachshort{} scales as additional device resources are added, noting that both in-NIC designs outperform commodity hosts using just one core in latency and online throughput.
+We report on how \approachshort{} scales as additional device resources are added, noting that both in-NIC designs outperform commodity hosts \emph{using just one core} in latency and online throughput.
 Furthermore, \Indfw{} provides higher per-core offline throughput than host-based approaches, even though our measured hosts exhibit higher clock speeds.
@@
 \subsection{Netronome Platform Fundamentals}\label{sec:netronome-platform-fundamentals}
-Netronome NFPs achieve scalable packet processing through parallelism.
+NFPs achieve scalable packet processing through parallelism.
 The chip is composed of \emph{microengines} (MEs), grouped into \emph{islands} of 4 or 12 MEs.
 %All 12-ME islands are used by a default P4 pipeline.
-Each ME has \numrange{4}{8} \emph{contexts} (hardware threads) which share a code store.
+Each ME has \numrange{4}{8} \emph{contexts} (threads) which share a code store.
 %Contexts and MEs may send one another numbered signals, and MEs have a small \emph{next-neighbour} register file for passing values in one direction to the next ME on the same island.
 %MEs run a proprietary instruction set, compiled to via a \emph{(Micro-)C} compiler.
-Beyond registers, the platform implements an explicit memory hierarchy scaling in size and access cost:
+Beyond registers, the platform has an explicit memory hierarchy in size and access cost:
 $\text{LMEM (ME)} < \text{CLS (Island)} < \text{CTM} < \text{IMEM (Chip)} < \text{EMEM}$.
@@
 \end{LaTeXdescription}
-\approachshort{} was evaluated on server blades (\emph{Mid/HighServer}), each hosting a single Netronome Agilio LX \numproduct{1 x 40}GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
+\approachshort{} was evaluated on server blades (\emph{Mid/HighServer}), each hosting a single Netronome Agilio LX \numproduct{40}GbE (NFP-6480, \SI{1.2}{\giga\hertz}).
 These servers ran Ubuntu 18.04.5 LTS (4.15.0-140-generic).
@@
 Control programs were built using rustc 1.52.1.
-We run \approachshort{} on a \num{4}-ME island of the NFP-6480 (\num{32} contexts) using \SIlist{32;16;8}{\bit} quantised arithmetic.
+We run \approachshort{} on a \num{4}-ME island of the NFP-6480 (\num{32} contexts) using \SIlist{32;16;8}{\bit} arithmetic.
 This is the largest cluster of cores which is not in use by a P4 pipeline.
-All \approachshort{} timing measurements were repeated over \num{10000} state packets (preceded by \num{1000} warmup packets), retrieving item processing times over PCIe via the controller machine, from which throughput was derived.
+All \approachshort{} timing measurements were repeated over \num{10000} state packets (preceded by \num{1000} warmup packets), retrieving item processing times over PCIe from which throughput was derived.
 Host throughput and latency measures were observed over \num{10} trials of \SI{10}{\second} (with \SI{5}{\second} warmup/cooldown times).
-We differ this from the NFP as hosts need to run numpy-based agents in parallel via separate processes; this also allows us to investigate the effects of oversubscription on host latencies.
-
+We differ this from the NFP as hosts need to run numpy-based agents in parallel as separate processes; this also allows us to investigate the effects of oversubscription.
 Policy sizes are set to those of a real-world DDoS control application~\parencite{DBLP:journals/tnsm/SimpsonRP20}: 20-dim state vectors, a bias tile and 16 full tiling sets (\numproduct{7x1}-dim, \numproduct{8x2}-dim, \numproduct{1x4}-dim), 8 tilings per set, 6 tiles per dimension, and 10 actions.
@@
 In \Coopfw{}, this creates \num{129} tasks across \num{31} workers.
-We choose a larger action count than this work to investigate the performance of more complex agents.
+%We choose a larger action count than this work to investigate the performance of more complex agents.
+We choose a larger action count to investigate the performance of more complex agents.
 
@@
 
-\fakepara{Raw inference and learning performance}
-We compare how long it takes for \approachshort{} to compute actions and policy updates, and report on the throughput of our designs against floating point (numpy-based) implementations of Sarsa on commodity hosts.
-This allows us to demonstrate the performance differences between the \indfw{} and \coopfw{} configurations, particularly in how \indfw's (and hosts') required policy locks impact throughput.
+\fakepara{Inference and learning}
+We compare how long it takes for \approachshort{} to compute actions and policy updates, and report on its throughput against a floating point (numpy-based) implementation of Sarsa on commodity hosts.
+This lets us demonstrate the performance differences between \indfw{} and \coopfw{}, particularly in how \indfw's (and hosts') required policy locks impact throughput.
 %We also report on the derived throughput measures accounting for the locking behaviour of each.
@@
 Online performance marks the number of decisions that can be made per second (and associated latency) when training a policy.
-Offline performance is crucial for pushing a trained, known-good policy to agents in the network with an expected higher raw decision throughput.
-State-action latency is a shared property of both cases, with the main impact on throughput arising from the update step.
+Offline performance is crucial for pushing a trained, known-good policy to agents with an expected higher raw decision throughput.
+State-action latency is a shared property of both cases; the main impact on throughput arises from the update step.
 
 We then vary the amount of worker threads to show how \approachshort{} scales to fit available compute resources on a device.
-This is an important aspect for (automated) allocation of compute resources in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of-time work scheduling which we examine later.
-This also demonstrates the number of cores needed to achieve a given latency or throughput bound on a policy of representative complexity.
-Moreover, to demonstrate how these costs vary as policy complexity increases, we vary the number of total dimensions included in the tiling.
+This is important for planning in an intelligent dataplane---particularly when cohabiting with other dataplane programs---and has effects on ahead-of-time work scheduling which we examine later.
+This also demonstrates the number of cores needed to achieve a given latency or throughput bound on a real-world policy.
+Moreover, to demonstrate how these costs vary in larger policies, we vary the total number of dimensions in each tiling.
 
@@
 
-\fakepara{Co-existence with the dataplane}
-While varying the rate of RL updates performed by \approachshort{}-\Coopfw{} (\SI{32}{\bit}) from \numrange{0}{16000} actions/s, we measure packet loss and sample latencies of cross traffic carried over a co-hosted P4 pipeline.
-This allows us to quantify whether on-chip (out-of-path) execution impacts ordinary dataplane behaviour through indirect means: e.g., EMEM cache evictions or hidden resource contention.
+\fakepara{Coexistence with the dataplane}
+While varying the rate of RL updates performed by \Coopfw{} (\SI{32}{\bit}) from \numrange{0}{16000} actions/s, we measure packet loss and latencies of cross traffic carried over a co-hosted P4 pipeline.
+This allows us to quantify whether on-chip (out-of-path) execution impacts ordinary dataplane behaviour indirectly: e.g., EMEM cache evictions or hidden resource contention.
 %In particular, we stress test both decisions and policy updates of our \emph{Parallel} design at full throughput.
-
-We perform these tests using Pktgen-DPDK~\cite{pktgen-dpdk}, placing an NFP in \emph{MidServer} as the device under test and connecting \emph{HighServer} over a \SI{40}{\giga\bit\per\second} DAC cable as the traffic source.
-We perform throughput and loss tests using \num{7}/\num{1} transmit/receive queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} transmit/receive queue at \SI{10}{\percent} send rate for \num{200000} measurements (sampling at \SI{2000}{\hertz} for \qtyproduct[product-units=single]{10 x 10}{\second}).
-This provides maximum throughput in the former case (relying solely on NIC counters for loss counting).
-In the latter case, this minimises host resource contention to observe exact latency measurements, have a high enough sample count to detect subtle (aggregate) latency effects, and eliminate \emph{host} receive drops.
+We test an NFP in \emph{MidServer} using Pktgen-DPDK~\cite{pktgen-dpdk}, connecting \emph{HighServer} as the traffic source over a \SI{40}{\giga\bit\per\second} cable.
+We perform loss tests using \num{7}/\num{1} Tx/Rx queues at \SI{100}{\percent} send rate for 10 bursts of \SI{30}{\second}, and perform latency tests using \num{1}/\num{1} Tx/Rx queue at \SI{10}{\percent} send rate for \num{200000} measurements (sampling at \SI{2000}{\hertz} for \qtyproduct[product-units=single]{10 x 10}{\second}).
+This maximises throughput in the former case, relying on NIC counters for loss detection.
+The latter minimises host resource contention to observe accurate latencies, observe enough samples to detect subtle (aggregate) latency effects, and eliminate \emph{host} receive drops.
 DPDK was setup using \qtyproduct[product-units=single]{4 x 1}{\gibi\byte} hugepages.
-Sent traffic was comprised of fixed-size \SIrange{64}{1518}{\byte} packets~\cite{rfc2544}.
+Sent traffic comprised fixed-size \qtyrange{64}{1518}{\byte} packets~\cite{rfc2544}.
 CPU clock scaling was disabled on \emph{HighServer}.
@@
 \fakepara{Resource requirements}
-Using the maximum policy size defined above, we investigate how the memory requirements imposed by \approachshort{} vary with the number of dedicated MEs, over and above a base P4 forwarding plane.
-We report resource use for \SI{32}{\bit} \Indfw{} and \Coopfw{} agents, with hash tables sized to \num{4096} state-action pairs and \num{16} separate reward values.
-This captures the relative cardinality of network RL traces to rewards, i.e., many input flows will map to one or few reward values (i.e., DDoS attack size estimation per egress-AS, queue occupancy in the case of AQM per output port).
+Using the policy size defined above, we investigate how the memory requirements imposed by \approachshort{} vary with the number of dedicated MEs, over and above a base P4 forwarding plane.
+We report resource use for \qty{32}{\bit} \Indfw{} and \Coopfw{} agents, with hash tables sized to \num{4096} state-action pairs and \num{16} separate reward values.
+This captures the relative cardinality of network RL traces to rewards; many input flows will map to one or few reward values (i.e., DDoS attack size estimation per egress-AS, queue occupancy per output port in AQM).
 
@@
 \subsection{Results and Discussion}\label{sec:results}
-\fakepara{Raw inference and learning performance}
-\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy.\footnote{For brevity, we omit integer numpy results---against a float implementation, median action latencies are \SI{14.6}{\percent} worse, with \SI{7.9}{\percent} longer update times.}
-We show that when using 4 MEs of the NFP-6480, \Coopfw{} achieves sub-\SI{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \SI{1}{\micro\second} worse (\SIlist{15;21}{\texttimes} speedups over a \emph{Collector} host).
-Importantly, \Indfw{} achieves lower median state-action latencies (\SI{2.79}{\texttimes}) \emph{and} update times (\SI{2.63}{\texttimes}) than a dedicated \emph{Collector} host \emph{while requiring only a single core or dedicated functional unit}.
+\fakepara{Inference and learning}
+\Cref{tab:lats} shows how \approachshort{} compares in latency with a numpy-based RL policy.\footnote{For brevity, we omit integer numpy results---against a float implementation, median action latencies are \qty{14.6}{\percent} worse, with \qty{7.9}{\percent} longer update times.}
+When using 4 MEs of the NFP-6480, \Coopfw{} achieves sub-\qty{35}{\micro\second} median latency, with \nth{99} and \num{99.99}\nthscript{th} percentile latencies less than \qty{1}{\micro\second} worse (\qtylist{15;21}{\texttimes} speedups over \emph{Collector}).
+Importantly, \Indfw{} achieves lower median latencies (\SI{2.79}{\texttimes}) \emph{and} update times (\SI{2.63}{\texttimes}) than a dedicated \emph{Collector} \emph{using only a single core or functional unit}.
 Crucially, in-NIC execution gives far tighter bounds on tail latency compared to host offloading.
-\num{99.99}\nthscript{th} percentile state-action latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, while host tail latencies are at least \SI{40.53}{\percent} greater---even when not oversubscribed.
+\num{99.99}\nthscript{th} percentile state-action latencies exceed the median by \SIlist{0.58;0.66}{\percent} for \Indfw{} and \Coopfw{}, while \emph{host} tail latencies are at least \SI{40.53}{\percent} greater.
 We show their cumulative distributions in detail (\cref{fig:lat-cumul}), noting how just one additional CPU-intensive task---potentially automated system updates or another traffic processing task---impacts tail latencies further (\emph{Float(Over)}).
@@
 \begin{table}
-	\caption{Latencies and computation times for \approachshort{} versus commodity hardware hosts.
-On-device execution is crucial in not only lowering latencies, but in reducing tail latencies.
+	\caption{Latencies and computation times for \approachshort{} versus commodity hosts.
+On-device execution is crucial in lowering latencies \emph{and} reducing tail latencies.
 Lower is better, with the best marked \emph{in bold}.\label{tab:lats}}
@@
 \begin{table}
-	\caption{Action and update throughputs for \approachshort{} versus commodity hardware hosts.
+	\caption{Action and update throughputs for \approachshort{} versus commodity hosts.
 Most designs cannot scale online performance with additional cores.
@@
 
-\Cref{tab:tputs} compares \approachshort{}'s throughput against host-based execution.
-We set the worker count on host machines equal to the number of physical cores on each device---moving beyond this would hamper tail latencies by an order of magnitude.
+\Cref{tab:tputs} compares \approachshort{}'s throughput against hosts.
+We set the worker count on host machines equal to their number of physical cores---moving beyond this would hamper tail latencies by an order of magnitude.
 To make the comparison fair in the context of many-core CPU environments, we include per-core throughput.
-\Indfw{} achieves \SI{2.82}{\texttimes} higher offline throughput per core than commodity \emph{Collector} hardware in spite of the NFP-6480 having a considerably slower clock speed (\SI{0.29}{\texttimes}).
-When compounded with the abundance of such weaker chips, in-NIC RL is able to deliver much higher throughput.
+\Indfw{} achieves \SI{2.82}{\texttimes} higher offline throughput per core than commodity \emph{Collector}, in spite of the NFP-6480 having a considerably slower clock speed (\SI{0.29}{\texttimes}).
+Due to the \emph{abundance} of such weaker chips, in-NIC RL is able to deliver much higher throughput.
 %As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{\texttimes} that of a dedicated collector machine, due to the locking requirements mandating coordinated division of work.
-As anticipated, the \Coopfw{} strategy is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{\texttimes} that of a dedicated collector machine, as the write lock around policy updates creates a bottleneck.
-
-By limiting the available workers in software, we show how \Coopfw{}'s policy update time (thus  online throughput---\cref{fig:vary-core}) scales with available cores.
-While \Coopfw{} always outperforms the host-based floating point implementations, we observe that there are two distinct crossover points which must be met to overcome \Indfw{}; \num{8} workers for online throughput, and \num{3} for state-action latency (plot omitted).
-Some artefacts of our environment and design choices are visible; the addition of new physical cores is more significant than contexts, and some schedule bottlenecks are visible.
-Most importantly, \Coopfw{}'s resource demand is tunable at compile time to meet the online training rate or action latency required by a task.
+As anticipated, \Coopfw{} is key in achieving serviceable throughput in an online learning agent, \SI{9.9}{\texttimes} that of a dedicated collector machine.
+
+By limiting workers, we show how \Coopfw{}'s policy update time (thus online throughput---\cref{fig:vary-core}) scales with available cores.
+While \Coopfw{} always outperforms hosts, we observe there are two distinct crossover points which must be met to overcome \Indfw{}; \num{8} workers for online throughput, and \num{3} for latency (plot omitted).
+Some artefacts of our environment and design are visible; the addition of new physical cores is more significant than contexts, and some schedule bottlenecks are visible.
+Most importantly, \Coopfw{}'s resource demand is tunable at compile time to meet the training rate or action latency required by a task.
+\Cref{fig:vary-work} shows how policy complexity affects update cost, scaling from a bias tile up to the full DDoS policy size.
+\Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning.
+%We note that this is a trivial case, in that the use of \emph{only} a bias tile negates the need for any input state (similar to a multi-arm bandit problem).
+We note that this is a trivial case, as using \emph{only} a bias tile returns a single preference list regardless of input state.
 
@@
 
-\Cref{fig:vary-work} shows how policy complexity affects update cost, scaling from a bias tile up to the full DDoS policy size.\footnote{Input vectors here all have 20 elements regardless of the policy design.}
-\Coopfw{} always produces an action in less time than \Indfw{}, but requires at least one state-based tile to excel in online learning.
-%We note that this is a trivial case, in that the use of \emph{only} a bias tile negates the need for any input state (similar to a multi-arm bandit problem).
-We note that this is a trivial case, as using \emph{only} a bias tile returns a single preference list regardless of input state.
-
 \begin{figure}[t]
@@
 %A noteworthy trend is that \SIlist{8;16}{\bit} versions of \approachshort{} consistently underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portions of \cref{fig:vary-work,fig:vary-work-latency}).
-\SIlist{8;16}{\bit} versions of \approachshort{} underperform compared to \SI{32}{\bit}, except for smaller workloads (seen in the zoomed portion of \cref{fig:vary-work})---even though our implementation is optimised to read and write policy data in batches.
-Because the native register width on the NFP is \SI{32}{\bit}, the compiler must emit extra instructions around arithmetic operations to correctly load and update values.
-This matches \SI{32}{\bit} becoming dominant in complex workloads: higher dimension tilings require more arithmetic operations.
-Most of the I/O comes \emph{after} this step, causing ALU use to dominate.
-This also explains why \SI{32}{\bit} becomes the best choice at different policy complexities for online (\cref{fig:vary-work}, 10 dims) and offline (plot omitted, 3 dims) agents, as state management falls into the serial portion of the online algorithm.
+\SIlist{8;16}{\bit} agents underperform against \SI{32}{\bit}, except for smaller policies (zoomed portion of \cref{fig:vary-work})---even though \approachshort{} is optimised to access policy data in batches.
+As the native register width on the NFP is \SI{32}{\bit}, the compiler emits extra instructions around ALU operations to correctly load and store values.
+This explains what we see in larger policies, as higher dimension tilings require more arithmetic, and most of the I/O comes \emph{after} computing each hit tile, causing ALU use to dominate.
+This also explains why the crossover point differs for online (\cref{fig:vary-work}, 10 dims) and offline (plot omitted, 3 dims) agents: state management falls into the serial portion of the online algorithm.
 %Smaller bit depths still give a \SIrange{2}{4}{\texttimes} saving in memory for policy storage (enabling more fine-grained or complex policies), in exchange for slightly worse latency and throughput.
-To overcome this, we investigated bit-stuffing several values into a single word during atomic writeback (as the platform offers both \SIlist{32;64
+To overcome this, we investigated bit-stuffing values into a single word during writeback (as the platform offers both \SIlist{32;64
 }{\bit} atomic addition).
-This is analogous to SIMD through clever use of padding bits, but we found that manipulating tiles into the correct format added \SI{10}{\percent} extra overhead.
+This is analogous to SIMD through use of padding bits, but we found that manipulating tiles into the correct format added \SI{10}{\percent} overhead.
 
@@
 %?? reference the inference times brought up in Taurus?
-We take inference times from \cref{tab:lats} for host and in-NIC processing, adding on the packet round-trip time to the decision site as state-action latency:
+To determine state-action latencies, we take inference times from \cref{tab:lats} for host and in-NIC processing, and add the packet RTT to the inference site:
 \begin{LaTeXdescription}
-	\item[\emph{In-NIC.}] As described in \cref{sec:agent-environment-communication}, EMEM rings have a median one way delay cross-island of \SI{140}{\nano\second}, giving a \emph{median \SI{34.63}{\micro\second} end-to-end inference latency}.
-	\item[\emph{Dedicated Collector.}] Employing DPDK to maximise performance, such hosts add one-way PCIe packet delays of \SIrange{0.9}{2.3}{\micro\second}~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
+	\item[\emph{In-NIC.}] As described in \cref{sec:agent-environment-communication}, EMEM rings have a median one-way delay cross-island of \SI{140}{\nano\second}, giving a \emph{median \SI{34.63}{\micro\second} end-to-end inference latency}.
+	\item[\emph{Dedicated Collector.}] Employing DPDK, such hosts add one-way PCIe packet delays of \SIrange{0.9}{2.3}{\micro\second}~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018}.
 	A UDP packet carrying \num{20} elements of state in \approachshort{} is \SI{128}{\byte}, so costs \SI{1}{\micro\second} to forward, and the reply state-action pair invokes a slightly higher cost.
@@
 \end{LaTeXdescription}
-Thus, in-NIC classical RL offers \SIlist{14.96;16.54}{\texttimes} speedups in latency over collector and vNF deployments respectively.
-We contrast these against deep RL policies on network tasks, which can take \SI{3}{\milli\second} to compute~\parencite{DBLP:journals/corr/abs-1910-04054}---2 orders of magnitude above \approachshort{} with identically sized (20-dim) input vectors.
+Thus, in-NIC RL offers \SIlist{14.96;16.54}{\texttimes} lower latency over collector and vNF deployments respectively.
+We contrast these against deep RL for network tasks, which can take \SI{3}{\milli\second}~\parencite{DBLP:journals/corr/abs-1910-04054}---2 orders of magnitude above \approachshort{} with identically sized inputs.
 %In the name of fairness, we assume that rule installation uses same mechanism we recommend, but show how bad it can get? I.e. with rule installation costs etc.
 
-\fakepara{Co-existence with the dataplane}
+\fakepara{Coexistence with the dataplane}
 Our setup met \SI{40}{\giga\bit\per\second} for packet sizes $\ge$\SI{256}{\byte}.
 %For frame sizes of \SIlist{64;128;256}{\byte} input traffic rates were \SIlist{17.4;32.9;37.0}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2;18.1}{\mega\packet\per\second}).
-For frame sizes of \SIlist{64;128}{\byte}, input traffic rates were \SIlist{17.4;32.9}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2}{\mega\packet\per\second}).
-Passing this traffic over the NFP device running \approachshort, no packet losses were incurred at any rate of RL actions.
-We show the effect of RL workloads on the round-trip latencies of cross traffic through \cref{fig:dataplane-coop}.
-As observed latencies do not obey a normal distribution, we employ a one-tailed \emph{Mann-Whitney U test} to mark statistically significant population increases in latency ($p < 0.05$) with a ``+''.
-In general, statistically significant latency increases concentrate around smaller packet sizes.
-All (bar one) of these affected \nth{99} percentile latencies by under \SI{0.38}{\percent} (at most \SI{78}{\nano\second}).
-This slight degradation can be explained by increased pressure on the NFP's \emph{Command Push-Pull} (CPP) bus, which is responsible for handling cross-island accesses to memory (particularly IMEM/EMEM) and other resources.
-\approachshort{} places load on the CPP bus through its \inring{}/\outring{} EMEM rings and last-tier policy accesses.
-This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---the NFP P4 toolchain segments packets, storing metadata (e.g., MAC prepend) and the first bytes of a packet in a \SI{256}{\byte} CTM block and parking their payloads in EMEM.
+For frames of \SIlist{64;128}{\byte}, input traffic rates were \SIlist{17.4;32.9}{\giga\bit\per\second} respectively (\SIlist[per-symbol=p,sticky-per=true]{33.9;32.2}{\mega\packet\per\second}).
+Passing this traffic over the NFP device running \approachshort, no packet losses occurred at any rate of RL actions.
+We show the effect of RL workloads on the RTTs of cross traffic via \cref{fig:dataplane-coop}.
+As observed RTTs were not normally distributed, we employed a one-tailed \emph{Mann-Whitney U test}, marking population increases in latency ($p < 0.05$) with a ``+''.
+Statistically significant increases concentrate around smaller packet sizes; all (bar one) of these affected \nth{99} percentile latencies by under \SI{0.38}{\percent} ($\le$\SI{78}{\nano\second}).
+This slight effect can be explained by increased pressure on the NFP's \emph{Command Push-Pull} (CPP) bus, which handles cross-island accesses to memory and other resources.
+\approachshort{} uses the CPP bus through its \inring{}/\outring{} EMEM rings and last-tier policy accesses.
+This also explains the sensitivity of \SI{256}{\byte} packets to \approachshort{}---the NFP segments packets, storing metadata (e.g., MAC prepend) and the first bytes of a packet in a \SI{256}{\byte} CTM block and parking their payloads in EMEM.
 \SI{256}{\byte} packets overshoot this due to metadata, causing small I/O accesses at a high rate for packets sized around this cutoff.
 %Regardless, this effect is small when observed.
-
-An anomaly is \SI{128}{\byte} packets at \num{3000} RL updates per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase in \cref{fig:dataplane-example}.
+An anomaly is \SI{128}{\byte} packets at \num{3000} RL updates per second, causing a \SI{222}{\nano\second} (\SI{1.18}{\percent}) increase.
 %This is observed through a shift of some packet latencies from the mode towards the tail, but no other changes in the distribution.
-We believe that the inbound request rate is weakly synchronised with traffic, causing burstiness of accesses to the CPP bus.
-We expect that FPGA designs can avoid this problem by having dedicated \inring{}/\outring{} access mechanisms for an \approachshort{} agent.
+We believe the inbound request rate is weakly synchronised with traffic, causing bursty accesses to the CPP bus.
+We expect that FPGA designs can avoid this by having dedicated \inring{}/\outring{} mechanisms for an \approachshort{} agent.
 
@@
 \fakepara{Resource requirements}
-\Cref{tab:resources} shows how \approachshort{} consumes memory as it scales to fit a device's compute resources, compared with a simple P4 forwarding application.
+\Cref{tab:resources} shows how \approachshort{} consumes memory as it scales to additional cores, compared with a simple P4 forwarding application.
 As one program is installed per ME, these results represent the minimum and maximum resource use on a single island (i.e., without replacing P4 workers).
@@
 \fakepara{Deployability}
-Setup of \approachshort{} uses two packet types: \emph{setup}, which contains policy shape and learning parameters, and \emph{tiling}, which provides a list of state indices for tiling sets.
-These packets took a mean \qtylist{27.03;16.69}{\micro\second} respectively to be processed on \Indfw{}, allowing an agent to be swapped between online and offline operation (or repurposed) painlessly.
-These swaps are useful when an agent should cease learning (i.e., when performance converges), or if a change in the environment suggests that more training is needed.
+Setup of \approachshort{} uses two packet types: \emph{setup}, which contains policy shape and learning parameters, and \emph{tiling}, which provides a list of indices for tiling sets.
+Handling these packets took a mean \qtylist{27.03;16.69}{\micro\second} respectively on \Indfw{}, allowing an agent to be swapped from online to offline painlessly, i.e., after convergence.
+%These swaps are useful when an agent should cease learning (i.e., when performance converges), or if a change in the environment suggests that more training is needed.
 %Tiling packet processing scales linearly with the number of \emph{tiling sets}, due to the needed precaching of tile set boundaries.
-Online-offline swaps for \Coopfw{} exhibit similar cost, however the need for explicit scheduling means that policy/tiling \emph{structure} changes (including the first complete setup) take \SI{422.63}{\micro\second} for the full-size policy described above.
-The time taken for \Coopfw{} to schedule its tasks was found to scale with the number of workers ($m$) and work items ($n$) as described earlier ($\mathcal{O}{\left(n\log{m}\right)}$).
-Ignoring the trivial solutions, reducing the worker count to \num{1} costs a mean \SI{238.22}{\micro\second}, while placing a single task incurs \SI{53.79}{\micro\second}.
-Policy \emph{data} changes require no additional work in any case, resolving purely to \texttt{memcpy}s.
-We observed that firmware installation (i.e., changing from \Indfw{} to \Coopfw{}, bit depth, or increasing maximum policy sizes) took a mean time of \SI{38.83}{\second}.
+\Coopfw{} exhibits similar costs, however the scheduler causes policy \emph{structure} changes to take \SI{422.63}{\micro\second} for a max-size policy---we found that this scales as described earlier ($\mathcal{O}{\left(n\log{m}\right)}$).
+%Ignoring the trivial solutions, reducing the worker count to \num{1} costs a mean \SI{238.22}{\micro\second}, while placing a single task incurs \SI{53.79}{\micro\second}.
+Policy \emph{data} changes require no additional work, resolving purely to \texttt{memcpy}s.
+Firmware installation (i.e., changing \Indfw{} to \Coopfw{} or bit depth) took a mean \SI{38.83}{\second}.
 %In the event that appropriate firmwares are not pre-compiled, we found that compiling and linking both \approachshort{} and the P4 toolchain took around \SI{35}{\second}, while changing only \approachshort{} parameters required around \SI{25}{\second}.
-Compiling and linking both \approachshort{} and the P4 toolchain took around \SI{35}{\second}, while changing only \approachshort{} parameters required around \SI{25}{\second}.
-These results show that \approachshort{} can be easily adapted and altered by network administrators once in place, and illustrates an advantage of SoC-based SmartNICs.
+Compiling and linking \approachshort{} and the P4 toolchain took \SI{35}{\second}, while changing only \approachshort{} parameters required \SI{25}{\second}.
+These results show that \approachshort{} can be easily adapted by network administrators once in place, and illustrates an advantage of SoC-based SmartNICs.
 
 \section{Potential Integrations}\label{sec:potential-integrations}
-To show the applicability and use of \approachshort{}, we propose an integration which would benefit from in-NIC RL; online DDoS mitigation.
-We support this using state-of-the-art P4 developments, and discuss how to balance online training with throughput in widespread deployment.
+%To show the applicability of \approachshort{}, we propose an integration which would benefit from in-NIC RL; online DDoS mitigation.
+%We support this using recent P4 developments, and discuss how to balance online training with throughput in deployment.
 %Additionally, we discuss how network administrators may combine offline (\Indfw) and online (\Coopfw) agents to achieve online learning while maximising action throughput in the majority of the network.
@@
 Classical RL has seen recent use in real-time, adaptive DDoS mitigation~\parencite{DBLP:journals/tnsm/SimpsonRP20}.
-\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of global network state and per-flow state to monitor how flows respond to bandwidth changes and packet drop.
+\Citeauthor{DBLP:journals/tnsm/SimpsonRP20}'s \emph{Guarded} agent design uses a mixture of network and per-flow state to monitor how flows respond to bandwidth changes and packet loss.
 %---applying the observations made by SPIFFY~\parencite{DBLP:conf/ndss/KangGS16} (observing how flow behaviour reacts to a change in rate limits) with more allowance for congestion-unaware protocols.
 Actions move flows up or down in punishment levels.
-
-\fakepara{Why in-NIC?}
-This approach co-hosts traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
-However, packet mirroring and offloading RL computation to a host/vNF are all sources of additional, consistent state-action latency.
-Both traffic measurement \emph{and} data-driven learning must be executed in this way.
-Unless running these stages in a dedicated pipeline (adding further processing latency), resource contention between processes will further impact tail latencies.
-Naturally, this requires high-performance hardware to be stacked at network egress points, potentially beyond reasonable space, power, or ventilation constraints.
-The solution to implement and improve upon this work using \approachshort{} is to place its RL agents on SmartNICs at AS edge nodes---a bump-in-the-wire deployment.
+%\fakepara{Why in-NIC?}
+%This approach co-hosts traffic measurement and analysis alongside OpenFlow-compatible switches at the edge nodes of an AS.
+%However, packet mirroring and offloading RL computation to a host/vNF are all sources of additional, consistent state-action latency.
+%Both traffic measurement \emph{and} data-driven learning must be executed in this way.
+%Unless running these stages in a dedicated pipeline (adding further processing latency), resource contention between processes will further impact tail latencies.
+%Naturally, this requires high-performance hardware to be stacked at network egress points, potentially beyond reasonable space, power, or ventilation constraints.
+To implement and improve upon this work using \approachshort{}, we would place its RL agents on SmartNICs at AS edge nodes---a bump-in-the-wire deployment.
 
@@
 \fakepara{Inputs}
-To collate the required inputs and state, we then examine the recent innovations of the community.
+%To collate the required inputs and state, we then examine the recent innovations of the community.
 Low-latency, pure-P4 solutions to extract and record per-flow TCP state directly in the dataplane such as Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} and Sonata~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18} are well-studied.
-In fact, the statistics offered by Dapper are a super-set of the local input state values employed by \emph{Guarded} agents, offering an opportunity to further improve their efficacy. 
-We propose placing such monitors in the P4 dataplane, existing on-chip alongside the \approachshort{} agent.
+%In fact, the statistics offered by Dapper are a super-set of the local input state values employed by \emph{Guarded} agents, offering an opportunity to further improve their efficacy. 
+We propose placing such monitors in the P4 dataplane, existing on-chip alongside \approachshort{}.
 The required global state (load measures from network paths) must still come from elsewhere in the network; this is now the element at highest risk of becoming stale, but the least likely to vary significantly in response to individual actions.
-The original work uses theoretical ``ground truth'' rewards, whose correct implementation and designs were left as an open challenge. 
-We posit that INDDoS~\parencite{tnms-ddos-victim-ident}, which uses Count-min Sketches to estimate DDoS victim cardinality, could be an effective reward function source---i.e., using the number of detected victims as a loss value.
+%The original work uses theoretical ``ground truth'' rewards, whose correct implementation and designs were left as an open challenge. 
+We posit that INDDoS~\parencite{tnms-ddos-victim-ident}, which estimates DDoS victim cardinality, would be an effective reward function source.
 
 \fakepara{Integrating \approachshort}
-Before each monitoring action, we require that the table hosting this hybrid solution polls \approachshort{}'s \emph{\outring{} Ring} for any generated actions.
-As we note in \cref{sec:limitations}, these actions would be placed into a local hash table \emph{and} exported to the attached controller to be inserted as P4 rules in batches.
-Afterwards, packet ingress timestamps would be used to emulate the \emph{Timed Random Sequential} (TRS) scheduler used by the anti-DDoS agents for rate-controlled work, where state vectors would be selected and passed into \approachshort.
-By design, active flows which are not \emph{judged} after a configurable time are discarded to prune the work set and allow new flows to be seen by the agent.
-The tight bounds on execution time known \emph{a priori} make it easy to calculate the maximum number of decisions which can be made per deadline.
+Before flow monitoring, this solution polls \approachshort{}'s \emph{\outring{} Ring}---these actions would be placed into a local hash table \emph{and} exported to the controller to be batch-inserted as P4 rules.
+Packet ingress timestamps would be used to emulate the TRS scheduler used by anti-DDoS agents for rate-controlled work, selecting state vectors for \approachshort.
+%By design, active flows which are not \emph{judged} after a configurable time are discarded to prune the work set and allow new flows to be seen by the agent.
+The tight bounds on \approachshort's execution time make it easy to calculate the maximum number of decisions which can be made per deadline.
 Reward values would then be separately inserted by a modified INDDoS table.
@@
 %?? Can I have a P4 code snippet here?
-
 %Reducing state-action latency (i.e., with \Coopfw) is useful for minimising the noise inherent in learning.
-Agents are limited by the fact that it can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
+It can take at least one RTT for meaningful changes to occur in a flow's behaviour ($\mathcal{O}{\left(\si{\milli\second}\right)}$ in a transit AS/ISP).
 Accordingly, this use case benefits most from an increase in \emph{throughput} using \Indfw{}.
-Higher throughput means that network flows are more likely to be judged in every timestep, even at high flow cardinality---making it more likely that changes in flow behaviour will be acted on and learned from.
-We note that the TRS scheduler is designed to handle large numbers of attack flows, combining new state vectors while the asynchronous agent is itself busy.
-A number of flows beyond the maximum throughput simply makes it take longer in expectation for a flow to be reassessed.
-As shown in \cref{sec:evaluation}, \approachshort{} far exceeds the throughput of host offloading.
-The control plane can then use wildcards or specific matches to narrow down or expand the set of flows to be controlled dynamically, though explicit (TRS) scheduling is still key in such adversarial environments.
+Higher throughput means that network flows are more likely to be judged in \emph{every} timestep---changes in flow behaviour are more likely to be acted on and learned from.
+%We note that the TRS scheduler is designed to handle large numbers of attack flows, combining new state vectors while the asynchronous agent is itself busy.
+Flows exceeding maximum throughput simply cause it take longer \emph{in expectation} for a flow to be reassessed.
+As shown in \cref{sec:evaluation}, \approachshort{} far exceeds the throughput of host offloading, making in-NIC execution ideal.
+The control plane can then dynamically narrow down or expand the set of flows to be monitored.
 
@@
 \subsection{Network Deployment Considerations}
-In a networked deployment, a subset of \approachshort{} nodes could be \coopfw{} agents, training online, while most other nodes run \indfw{} designs to meet throughput guarantees.
-The control plane would then combine, downsample, and distribute these improved policies between offline agents.
-This can be taken further, using policy deltas to enable out-of-path transfer learning for more complex models such as neural networks.
+In a network, a subset of \approachshort{} nodes could be \coopfw{} agents, training online, while most other nodes run \indfw{} to meet throughput guarantees.
+The control plane would then combine and distribute these improved policies between offline agents.
+This can be taken further, using policy deltas to enable transfer learning for more complex models such as neural networks.
 
@@
 \fakepara{In-network ML}
-Taurus~\parencite{DBLP:journals/corr/abs-2002-08987} proposes that efficient line-rate inference can be achieved using a configurable grid of map-reduce units in the packet pipeline (implementing e.g., LSTMs and SVMs).
+Taurus~\parencite{DBLP:journals/corr/abs-2002-08987} proposes efficient line-rate inference using a configurable grid of map-reduce units in the packet pipeline (implementing e.g., LSTMs and SVMs).
 On CGRA hardware, they achieve sub-\si{\micro\second} extra latency.
 \emph{IIsy}~\parencite{DBLP:conf/hotnets/XiongZ19} shows how \emph{classical ML inference} (SVMs, Na\"{i}ve Bayes, etc.) can be converted into match-action tables compatible with \emph{any} P4 deployment.
-They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA.
+They achieve mean \SI{2.62}{\micro\second} extra latency on NetFPGA.
 %Neither considers online model training.
 
-A recent line of research in the community has been to investigate \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate packet classification.
-\emph{BaNaNa SPLIT} shows this as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is carried out on the \emph{CPU} to reduce latency imposed by GPU batching and transfer, but fully-connected layers can be accelerated further by NICs.
+A recent line of research is the use of \emph{Binarised Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate packet classification.
+\emph{BaNaNa SPLIT}~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731} shows this as an offload mechanism for fully-connected layers.
 %However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}; the authors hypothesise that networking hardware would be a better fit for some parts of the inference pipeline.
-In-network packet tagging and classification by pre-trained BNNs is shown by \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
-N3IC achieves packet inference in \SI{45}{\micro\second} on the NFP, and \SI{0.3}{\micro\second} on NetFPGA for \SI{256}{\bit} inputs.
-Comparatively, \approachshort{}-\Coopfw{} can process an identically-sized input (8-dim) in a median \SI{13.83}{\micro\second}.
+In-network packet tagging and classification by pre-trained BNNs is shown by \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}, achieving packet inference in \SI{45}{\micro\second} on the NFP, and \SI{0.3}{\micro\second} on NetFPGA for \SI{256}{\bit} inputs.
+Comparatively, \approachshort{}-\Coopfw{} can process an identically-sized input in a median \SI{13.83}{\micro\second}.
 Our work handles larger inputs (\SI{640}{\bit}) at lower latencies (\SI{34}{\micro\second}), and offers online learning.
@@
 \textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantisation on the NFP, using \emph{in-path} compute rather than our asynchronous model.
-Inference latency on small networks (3 layers, \num{30} neurons) can be as high as \SI{500}{\micro\second} on line rate traffic, emphasising the value of path-adjacent compute.
+Inference latency on small networks can be as high as \SI{500}{\micro\second} on line rate traffic, emphasising the value of path-adjacent compute.
 
@@
 \fakepara{In-network ML \emph{acceleration}}
-Optimisation of distributed neural network training is an area where in-network compute has been key for general NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18} and RL-specific procedures~\parencite{DBLP:conf/isca/LiLYCSH19} using NetFPGAs to introduce floating point adders to process dedicated packet classes.
-Distributed training is a partition-aggregate workload, where gradient vectors are sent back to a controller to update the `true' model---in-NIC processing allows these to be aggregated \emph{in-network}, overcoming incast behaviour and host bottlenecks.
+Optimisation of distributed neural network training is an area where in-network compute has been key for general NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18} and RL-specific procedures~\parencite{DBLP:conf/isca/LiLYCSH19}, using NetFPGAs to implement floating point adders.
+In-NIC processing allows gradient packets to be aggregated \emph{in-network}, overcoming incast behaviour and host bottlenecks.
 
 \fakepara{RL for network control}
-\Textcite{DBLP:conf/sigcomm/LiangZJS19} use deep RL techniques to train an agent which can build efficient decision tree packet classifiers.
-Their approach, \emph{NeuroCuts}, produces time/space efficient classifiers for use in constrained environments (e.g., network hardware) beyond existing heuristic solutions.
+\emph{NeuroCuts}~\parencite{DBLP:conf/sigcomm/LiangZJS19} uses deep RL to train an agent which can build efficient decision tree packet classifiers for use in constrained environments (e.g., network hardware).
 Deep RL techniques have been used for QUIC congestion control optimisation~\parencite{DBLP:journals/corr/abs-1910-04054}.
@@
 \emph{PANIC}~\parencite{DBLP:conf/hotnets/StephensAS18} places a routing fabric between distinct packet/data processing elements \emph{in a SmartNIC}.
-Such designs would enable general, asynchronous, and novel compute in SmartNICs and switches, for instance offering consistent and easy to use communication between workers versus hard-coded ME relationships.
+Such designs would enable novel asynchronous compute in SmartNICs and switches, for instance offering consistent and easy to use communication between workers versus hard-coded ME relationships.
 Event-driven versions of P4 have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}.
@@
 \section{Conclusion}
-We have presented \emph{\approachshort{}}, bringing \emph{online reinforcement learning} to the dataplane.
+We have presented \emph{\approachshort{}}, bringing \emph{online RL} to the dataplane.
 %\approachshort{} has shown how classical RL techniques make online learning possible by simplifying update logic and enabling parallel processing.
-In-NIC use of these classical algorithms makes this possible, and enables significant reductions in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvement in online learning throughput, with minimal impact on dataplane cross-traffic.
+In-NIC use of classical RL algorithms makes this possible, and enables significant reductions in median--\num{99.99}\nthscript{th} inference times and order of magnitude improvements in online learning throughput, with minimal impact on dataplane cross-traffic.
 %The deployment environment and asynchronous design were shown to eliminate PCIe delays and impose minimal impact on carried dataplane traffic.
@@
 \fakepara{Acknowledgements}
-The authors would like to thank Rhys Simpson for his comments, and for discussions on the soundness of SIMD-like optimisations.
+The authors would like to thank Rhys Simpson for his comments and discussions on the soundness of SIMD-like optimisations.
 They would additionally like to thank Stefanos Sagkriotis, Mircea Iordache-\c{S}ic\u{a}, and Haruna Adoga for their comments and feedback.

### 2021-10-18 paper/poster-abstract.tex
Terrible, terrible things.

@@
 sigconf,natbib=false
-,anonymous=true
+%,anonymous=true
 ]{acmart}
@@
 %\setcopyright{none}
-\setcopyright{acmcopyright}
+%\setcopyright{acmcopyright}
 %\setcopyright{acmlicensed}
@@
 %\settopmatter{printacmref=false, printccs=false, printfolios=true}
-\settopmatter{printfolios=true, printccs=true}
+%\settopmatter{printfolios=true, printccs=true}
 %\settopmatter{printacmref=false, printccs=true, printfolios=true}
 
-% DOI
-\acmDOI{12345}
-
-% ISBN
-\acmISBN{67890}
-
-%Conference
-\acmConference[CoNEXT '21]{the \nth{17} International COnference on emerging Networking EXperiments and Technologies}{December \numrange{7}{10}, 2021}{Online}
 \acmYear{2021}
-%\copyrightyear{}
-
-%% {} with no args suppresses printing of the price
-%\acmPrice{15.00}
-\acmPrice{}
+\copyrightyear{2021}
+\setcopyright{acmcopyright}
+\acmConference{CoNEXT '21}{December 7--10, 2021}{Virtual Event, Germany}
+\acmPrice{TBA} % Use the data from your email from ACM rights management
+\acmDOI{TBA} % Use the data from your email from ACM rights management
+\acmISBN{TBA} % Use the data from your email from ACM rights management
 

### 2021-10-19 paper/poster-abstract.tex
Camera-ready poster.

@@
 
-\acmYear{2021}
-\copyrightyear{2021}
-\setcopyright{acmcopyright}
-\acmConference{CoNEXT '21}{December 7--10, 2021}{Virtual Event, Germany}
-\acmPrice{TBA} % Use the data from your email from ACM rights management
-\acmDOI{TBA} % Use the data from your email from ACM rights management
-\acmISBN{TBA} % Use the data from your email from ACM rights management
+\acmYear{2021}\copyrightyear{2021}
+\acmConference[CoNEXT '21]{The 17th International Conference on emerging Networking EXperiments and Technologies}{December 7--10, 2021}{Virtual Event, Germany}
+\acmBooktitle{The 17th International Conference on emerging Networking EXperiments and Technologies (CoNEXT '21), December 7--10, 2021, Virtual Event, Germany}
+\acmPrice{15.00}
+\acmDOI{10.1145/3485983.3493345}
+\acmISBN{978-1-4503-9098-9/21/12}
 
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/coop}
+	\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/coop}
 	\caption{Parallel policy execution.

### 2021-10-25 paper/poster-abstract.tex
Camera-readier poster.

@@
 \acmYear{2021}\copyrightyear{2021}
+\setcopyright{rightsretained}
 \acmConference[CoNEXT '21]{The 17th International Conference on emerging Networking EXperiments and Technologies}{December 7--10, 2021}{Virtual Event, Germany}
@@
 Programmable dataplane hardware, being designed solely for efficient packet processing, lacks floating-point arithmetic support, even in the case of more general purpose NPU-type SmartNICs.
-Additionally, deep neural network training relies on backpropagation, can require memorizing many a sizeable replay buffer for RL, and needs many mini-batches of data for stable training.
+Additionally, deep neural network training relies on backpropagation, can require memorizing many sizeable replay buffers for RL, and needs many mini-batches of data for stable training.
 The natural solution is to pass training data to a host machine or dedicated accelerator, yet this adds delays in crossing the PCIe bus, kernel-user handover, and buffering---adding latency while learning.
@@
 We compare our implementation of \approachshort{} against a tile-coded RL agent written in numpy on a powerful host (i7-6700K@\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}), using the policy dimensions of existing DDoS prevention work~\parencite{DBLP:journals/tnsm/SimpsonRP20}, across various quantisation bit depths.
-\Cref{fig:lat-cumul} shows that our main design (\coopfw) achieves order-of-magnitude improvements over our single-threaded variant (\indfw) and a host implementation, as well as considerably tighter tail behaviour on the SmartNIC.
+\Cref{fig:lat-cumul} shows that our main design (\coopfw) achieves order-of-magnitude latency improvements over our single-threaded variant (\indfw) and a host implementation, as well as considerably tighter tail behaviour on the SmartNIC.
 Note that lower bit depths have higher execution costs due to the native \qty{32}{\bit} width of Netronome registers.
 On update times, we see \qty{63.12}{\micro\second} at 99.99\nthscript{th} percentile for a \qty{32}{\bit} policy of this size.
-We relate a \qty{15}{\texttimes} latency (\coopfw) and \qty{2.82}{\texttimes} offline throughput-per-core (\indfw) in spite of the Netronome's considerably slower clock speed (\qty{0.29}{\texttimes}), showing the value of exploiting SmartNIC parallelism.
+We relate a \qty{15}{\texttimes} latency (\coopfw) and \qty{2.82}{\texttimes} offline throughput-per-core (\indfw) improvement in spite of the Netronome's slower clock speed (\qty{0.29}{\texttimes}), showing the value of exploiting SmartNIC parallelism.
 
@@
 		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/rl-perf-tester/latency-cumul}
+			\includegraphics{plots/latency-cumul}
 			\vspace{-10pt}
@@
 		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/rl-perf-tester/latency-cumul-broad}
+			\includegraphics{plots/latency-cumul-broad}
 		}

### 2021-12-01 poster/render-table.tex
CoNEXT Poster rendered/done.

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    % ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

   \usepackage{graphicx}

\usepackage{booktabs}
\newcommand{\approach}{On Path Learning}
\newcommand{\approachshort}{OPaL}
\newcommand{\Coopfw}{\emph{CoOp}}
\newcommand{\coopfw}{\Coopfw}
\newcommand{\Indfw}{\emph{Ind}}
\newcommand{\indfw}{\Indfw}
\newcommand{\inring}{\textsc{In}}
\newcommand{\outring}{\textsc{Out}}
\newcommand{\seidr}{Sei\dh{}r}
\newcommand{\seidrsafe}{Seidr}

\usepackage[per-mode=symbol]{siunitx}
\sisetup{detect-all, range-phrase=--, range-units=single, detect-weight=true, table-format=1.3}
\DeclareSIUnit{\packet}{p}

\usepackage[sfdefault]{AlegreyaSans}

\begin{document}

\resizebox{\linewidth}{!}{
\begin{tabular}{@{}ccS[table-format=2]S[table-format=3.3]S[table-format=2.3]S[table-format=1.3]S[table-format=1.3]@{}}
\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{1}{c}{Workers} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{2}{c}{Throughput/core (k actions/s)}\\
\cmidrule(lr){4-5}\cmidrule(lr){6-7} & & & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online} & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online}\\ \midrule
Float & Collector & 4 & 7.673 \pm 0.049 & 1.627 \pm 0.031 & 1.918 \pm 0.012 & \multicolumn{1}{c}{---}\\
 & MidServer & 6 & 5.584 \pm 0.030 & 0.791 \pm 0.012 & 0.931 \pm 0.005 & \multicolumn{1}{c}{---}\\
Int32 & \approachshort{}-\Indfw{} & 32 & \bfseries 172.875 \pm 0.229 & 4.333 \pm 0.005 & \bfseries 5.402 \pm 0.007 & \multicolumn{1}{c}{---}\\
 & \approachshort{}-\Coopfw{} & 32 & 29.166 \pm 0.173 & \bfseries 16.141 \pm 0.073 & 0.911 \pm 0.005 & \bfseries 0.504 \pm 0.002\\
\bottomrule
\end{tabular}
}

\end{document}

### 2022-01-06 paper/noms-short-paper.tex
Camera readied!

@@
 
+%\usepackage[T1]{fontenc}
+
 \title{\mytitle}
-\author{Kyle~A.~Simpson~\orcidlink{0000-0001-8068-9909}, Dimitrios~P.~Pezaros~\orcidlink{0000-0003-0939-378X}\\\emph{University of Glasgow}, Glasgow, Scotland\\\email{k.simpson.1@research.gla.ac.uk}, \email{Dimitrios.Pezaros@Glasgow.ac.uk}}
-
+\author{Kyle~A.~Simpson~{\orcidlink{0000-0001-8068-9909}~\scriptsize[0000-0001-8068-9909]}, Dimitrios~P.~Pezaros~{\orcidlink{0000-0003-0939-378X}~\scriptsize[0000-0003-0939-378X]}\\\emph{University of Glasgow}, Glasgow, Scotland\\\email{k.simpson.1@research.gla.ac.uk}, \email{Dimitrios.Pezaros@Glasgow.ac.uk}}
+
+%\IEEEpubid{\makebox[\columnwidth]{978-1-6654-0601-7/22\/\$31.00 ~\copyright~2022 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
 \begin{document}
@@
 %\renewcommand{\shortauthors}{Simpson \emph{et al}.}
+
+%\IEEEoverridecommandlockouts \IEEEpubid{\makebox[\columnwidth]{978-1-6654-0601-7/22\/\$31.00 ~\copyright~2022 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
 
@@
 \section{Introduction}
-\begin{figure}
-	\centering
-	\includegraphics[keepaspectratio, width=0.65\linewidth]{figures/arch-with-p4}
-	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
-SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible alongside P4 dataplanes.
-Classical RL policy methods are the key to making this feasible.\label{fig:netro-arch}}
-\end{figure}
 Automatic network optimisation, control, and defence are at last becoming commonplace.
@@
 I intend to present how reinforcement learning execution and network telemetry will differ in the above metrics when compared to a vNF-based deployment (i.e., software on external commodity servers).
+
+%\vspace{-0.3em}
 
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[keepaspectratio, width=0.65\linewidth]{figures/arch-with-p4}
+	\vspace{-0.5em}
+	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane.
+SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible alongside P4 dataplanes.
+Classical RL policy methods are the key to making this feasible.\label{fig:netro-arch}}
+\end{figure}
+
 \section{Design and Implementation}\label{sec:design}
@@
 	\centering
-	\includegraphics[keepaspectratio, width=0.65\linewidth]{figures/coop}
+	\includegraphics[keepaspectratio, width=0.69\linewidth]{figures/coop-stretch}
+	\vspace{-0.5em}
 	\caption{\approachshort{} scales to fit device capacity according to either latency or throughput needs---\Coopfw{} is the optimal online design.
@@
 	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-core}
-	}
+		\includegraphics{../plots/build/rl-perf-tester/vary-core-bigtext}
+	}
+\vspace{-0.5em}
 	\caption{
@@
 	\resizebox{\resultplotwidth}{!}{
-		\includegraphics{../plots/build/rl-perf-tester/vary-work}
-	}
+		\includegraphics{../plots/build/rl-perf-tester/vary-work-bigtext}
+	}
+\vspace{-0.5em}
 	\caption{\Coopfw{} processes updates faster than \Indfw{}---thus has higher online performance---on almost all policy sizes.
@@
 	\centering
-	\begin{subfigure}{0.45\linewidth}
+	\begin{subfigure}{0.525\linewidth}
 		\resizebox{1.0\linewidth}{!}{
 %			\includegraphics{../plots/build/stress/heat-latency-median}
-			\includegraphics{../plots/build/stress/heat-latency-two-9-rel}
+			\includegraphics{../plots/build/stress/heat-latency-two-9-rel-bigtext}
 %			\includegraphics{../plots/build/stress/heat-latency-two-9-perc}
 		}
-		\caption{Deviations in \nth{99} percentile cross-traffic RTTs.\label{fig:dataplane-heat}}
+		\vspace{-1.75em}
+		\caption{\nth{99} \%tile cross-traffic RTT changes.\label{fig:dataplane-heat}}
 	\end{subfigure}
-	\hspace{0.05\linewidth}
-		\begin{subfigure}{0.45\linewidth}
+%	\hspace{0.05\linewidth}
+		\begin{subfigure}{0.46\linewidth}
 		\resizebox{1.0\linewidth}{!}{
-			\includegraphics{../plots/build/stress/histo-128B-0-3-trim}
-		}
-		\caption{RTT distribution for \SI{128}{\byte} packets at \numlist{0;3000} RL actions/s.\label{fig:dataplane-example}}
+			\includegraphics{../plots/build/stress/histo-128B-0-3-trim-div}
+		}
+		\caption{RTT distribution for \SI{128}{\byte} packets at \numlist{0;3000} actions/s.\label{fig:dataplane-example}}
 	\end{subfigure}

### 2022-11-18 ngni-talk/arista-talk.tex
Forgot to upload some talk (sub-)variants

--- 2021-06-04 ngni-talk/talk.tex
@@
 \robustify\bfseries
-\sisetup{detect-all, range-phrase=--, range-units=single}
+\sisetup{detect-all, range-phrase=--, range-units=single, per-mode=symbol}
 
@@
 
-\title{Revisiting the Classics: Online RL in the Programmable Dataplane}
+\title{Revisiting the Classics: Online RL in the Programmable Dataplane \emph{(IEEE NOMS '22)}}
 \author{\vspace{-1em}\textbf{Kyle A. Simpson}, Dimitrios P. Pezaros\\
@@
 \institute{University of Glasgow}
-\date{\nth{4} June, 2021}
+\date{\nth{27} October, 2022}
 
@@
 				\item How? Train model off-NIC, convert to \alert{binary neural network}\footnotemark, or \alert{decision tree}\footnotemark.
+				\item Bespoke architectures like \emph{Taurus}\footnotemark.
 				\item Limits? No online training, cost of backprop algo (expensive!), vast data needs.
@@
 \footcitetext{DBLP:conf/hotnets/XiongZ19}
+\setcounter{footnote}{3}
+\footcitetext{DBLP:journals/corr/abs-2002-08987}
 \end{frame}
@@
 				\item PCIe access times $\mathcal{O}(\si{\micro\second})$\footnotemark
-				\item Crossing VMs/vNFS has $\sim$\SI{10}{$\times$} higher cost\footnotemark
+				\item Crossing VMs/vNFS has $\sim$\qty{10}{\times} higher cost\footnotemark
 				\item MATs recompiled $\mathcal{O}(\si{\second})$. Many rules $\implies$ batching.
@@
 	\end{columns}
-	\setcounter{footnote}{3}
+	\setcounter{footnote}{4}
 	\footcitetext{DBLP:conf/sigcomm/NeugebauerAZAL018}
-	\setcounter{footnote}{4}
+	\setcounter{footnote}{5}
 	\footcitetext{DBLP:journals/cm/CzivaP17}
@@
 \section{Background and Design}
+
+\begin{frame}{A primer on the Netronome NFP}
+	\begin{columns}
+		\begin{column}{0.6\linewidth}
+			\begin{itemize}
+				\item PCIe NIC @ \qtyrange{40}{100}{\giga\bit\per\second}.
+				\item \alert{NPU-type device}---many weak cores.
+				\item One program per core, cooperative scheduling between HW threads.
+				\item \textbf{Programming Model:}
+				\begin{itemize}
+					\item (P4) -> $\mu$C -> Proprietary ASM.
+					\item Explicit locality of compute \& memory.
+				\end{itemize}
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.4\linewidth}
+			\centering
+			\resizebox{0.8\linewidth}{!}{
+				\includegraphics{images/smartnic.png}
+			}
+		\end{column}
+	\end{columns}
+\end{frame}
 
@@
 
-	\textbf{One NFP island $\implies$ \alert{\SI{15}{$\times$} median S$\rightarrow$A speedup}.}\pause
+	\textbf{One NFP island $\implies$ \alert{\qty{15}{\times} median S$\rightarrow$A speedup}.}\pause
 
 	\textbf{Far tighter tail than host offload!}
+\end{frame}
+
+\begin{frame}{Scheduler performance}
+	\begin{figure}
+		\resizebox{0.8\linewidth}{!}{
+			\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
+		}
+		\caption{Action/update compute times in a \SI{32}{\bit} \Coopfw{} agent under different work schedulers.\label{fig:work-alloc-32}}
+	\end{figure}
 \end{frame}
@@
 
+\begin{frame}{Feasibility on other architectures?}
+	\begin{itemize}[<+->]
+		\item \textbf{Tofino etc.?} \alert{Unlikely}.
+		\item \textbf{Taurus?\footcite{DBLP:journals/corr/abs-2002-08987}} Exec model \emph{almost} right, but \alert{read-only}.
+		\item \textbf{FPGA?} Probably an effective match to this model + algorithm.
+		\item Research gap: hardware architectures built to collate and apply batches of samples, gradient updates?
+	\end{itemize}
+\end{frame}
+
 \begin{frame}[standout]
@@
 
-\begin{frame}{Scheduler performance}
-	\begin{figure}
-		\resizebox{0.8\linewidth}{!}{
-			\includegraphics{../plots/build/rl-perf-tester/work-strat-32bit}
-		}
-		\caption{Action/update compute times in a \SI{32}{\bit} \Coopfw{} agent under different work schedulers.\label{fig:work-alloc-32}}
-	\end{figure}
-\end{frame}
-
 \begin{frame}{Per-worker throughput}
